### 2017-01-23 chapters/ch2-bottom-up.tex
First commit

\section{Requirements}
\begin{itemize}
\item Permissionless
\item Byzantine fault tolerant
\item No PoW
\item Works under churn
\item Underlying data structure is TrustChain
\item Detects forks or double-spends
\item No step in the protocol blocks transactions
\item Application independent
\end{itemize}

\section{Assumptions}
\begin{itemize}
\item The total number of valiators is $N$. The number of faulty validators is
  no more than $f$, where $n > 3f$ (this value depends on BFT consensus
  algorithm that we use).
\item Validators have the complete history of the previously agreed set of
  transactions.
% \item Single connected component for TrustChain.
% \item No sybil attack % 
\item Weak synchrony, where messages are guaranteed to be delivered after
  $\Delta$ (also depends on BFT consensus algorithm). 
% TODO this depends on the BFT protocol and the broadcast protocol
% \item Rough (inaccurate to some degree) global clock
\end{itemize}

\section{Protocol}
The goal of the bottom-up consensus protocol is to validate a set of new
transactions against the previously validated transactions, and then disseminate
these transactions. For a set of transactions to be valid, they need to adhere
to the TrustChain construction. Out solution addresses the issue of Proof of Work
(wasteful) and classical BFT algorithms (does not scale).

To describe the protocol, we make two assumptions, later we show to these
assumptions are removed.
\begin{enumerate}
\item We assume there exist a set of validator nodes for every consensus run and
  for every round in a consensus run that are selected in a fair way and is
  known to all nodes. We address this assumption in \cref{sec:consensus-group}
  and \cref{sec:validator-group}.
\item We assume there exist a set of valid transactions from the previous
  consensus run. Every node knows the Merkle hash of this set of transactions.
  Assumption addressed in \cref{sec:bootstrapping}.
\end{enumerate}
\Cref{fig:protocol-init} shows the state of the system. Before explaining the
four phases of the protocol, we describe the concept of checkpointing.

\begin{figure}[htb]
	\centering
	\includegraphics[width=0.7\textwidth]{figures/protocol-init}
	\caption{The initial state of the bottom-up consensus protocol. The thin
      lines are hash pointers to transactions and $ck_{x,n}$ represents a
      checkpoint block, where $x$ is the node ID and $n$ is the consensus run.}
    \label{fig:protocol-init}
\end{figure}

\subsection{Checkpointing}
A checkpoint block is a self-transaction where its input and output only involve
the node itself. Node $x$ creates the checkpoint block using
$$
ck_{x,n} = H(pk_x || H(txs_{n-1})),
$$
where $H$ is a cryptographically secure hash function, $pk_{x}$ is public key of
$x$ and $txs_{n-1}$ is the validated set of transaction in consensus run
$n-1$.

Transactions between two checkpoints form a chain which may or may not be
validated. For example, the chain between $ck_{1,n-2}$ and
$ck_{1,n-1}$ in \cref{fig:protocol-init} is validated, any chain after
$ck_{1,n-1}$ is not validated. Only one checkpoint block is allowed for
every agreed set of transactions $txs_{n}$.

The goal of every honest node is to validate their unvalidated chain. They do
this by sending it to the validators. The validators then performs the actual
validation with other validators and reach consensus. We describe these in
detail next.

\subsection{Setup phase}

To prepare a chain on node $x$ for validation in consensus run $n$, $x$ first
communicates to the validators of consensus run $n-1$ for the valid set of
blocks (recall that we assume every node knows the validators), and then
generates a new checkpoint block (\cref{fig:protocol-setup}). This completes
the setup phase.

\begin{figure}[htb]
	\centering
	\includegraphics[width=0.7\textwidth]{figures/protocol-setup}
	\caption{In the setup phase, nodes obtains the new set of validated
      transaction from the validators and generate checkpoint blocks (top of the
      figure).}
    \label{fig:protocol-setup}
\end{figure}

\subsection{Group consensus phase}
Every node (validator or not) belong to a \emph{consensus group} (again we
assume this is known). Validators that belong to the same consensus group is
called a \emph{validator group}.

Nodes send their unvalidated chain to their respective validator groups. The
validator collect those chains for a fixed amount of time and then rejects any
new chains. Next, the validator groups perform a BFT consensus algorithm (e.g.
PBFT~\cite{castro1999practical}) and reach consensus on a set of valid
transactions. At the end of this phase, the validators in every validator group
should have a set of transactions that has reached group consensus.

If the BFT consensus algorithm is sufficient when there is only one consensus
group, then there is no need to perform the group merge phase. But most
algorithms do not scale well~\cite{vukolic2015quest}.

\subsection{Group merge phase}
Given a set of valid transaction for every validator group, we need to merge
them to achieve global consensus. We do this using a bottom-up approach, similar
to divide and conquer algorithms such as merge sort. The algorithm works in
rounds, but it does not assume a synchronous environment.

Every validator group merges with another validator group (assuming the total
number of groups is even). We assume the pairing is known just like how every
node knows the identities of the validators, but we remove the assumption later.
Validators from both groups share their validated transactions and run a BFT
algorithm to agree on the final result, which is simply the union of the two
previously validated transactions. By this point the group is merged with a new
set of validators.

The group merge phase proceeds to the next round by repeating itself until there
is only a single validator group. Groups may not finish the same round at the
same time, thus groups that finish early need to wait until its pair is ready
(by exchanging messages).
% If the pair never finishes, it implies that the majority of the nodes in the
% pair have been infiltrated by attackers. In this case, the host node

\subsection{Dissemination phase}
The validator in the final round publishes the set of validated transactions
(possibly using BitTorrent) and the Merkle hash of the validated transactions
(possibly in a hierarchical way). They also create a self-transaction in their
own chain that contains the Merkle hash.

\subsection{Consensus group formation}\label{sec:consensus-group}
Until now we always assumed the consensus groups and validator groups are known,
in this and the subsequent sections we describe a novel technique to perform
group formation.

Group formation is a hard problem in distributed systems because it is
essentially a consensus problem---every node need to know the group membership
information of every other node. Even worse, running a BFT algorithm may not be
possible in this case because we have dynamic group sizes.

Our novel idea is to piggyback on the latest consensus result. Every node knows
the latest consensus result $txs_{n-1}$ (if they don't they can ask the
validators) or the Merkle hash of it $H(txs_{n-1})$, then they perform some
deterministic computation to find the luck value
$$l_{x,n-1} = H(H(txs_{n-1}) || pk_{x})$$
for every node that participated in $txs_{n-1}$. The consensus group for $x$ in
consensus run $n$ on round $k$ (initially at 0) is then
$$
g_{x,n,k} = l_{x, n-1} \mod 2^{m-k},
$$
where $m$ can be for example 8, it depends on the estimated number of users
and the target consensus group size.

Computing the cryptographic hash for every node in the set may become expensive,
and poor clients may fail to do it in time. We can work around this by having
the validators send the membership information to all members of their consensus
group.

\subsection{Validator group formation}\label{sec:validator-group}
Validators all belong to a consensus group, but only lucky nodes are allowed to
be validators. To find the luckiness, we use the same luck value $l_{x,n-1}$,
and check for the inequality
$$
l_{x, n-1} < D,
$$
where $D$ is the difficulty value and a system parameter (undecided). Nodes that
satisfies the inequality are considered lucky.

Recall that validator groups merge and create a new set of valid transactions.
So the validator groups need to answer the following two questions.
\begin{enumerate}
\item Which other validator group to pair up (this information need to be
  symmetrical)?
  \item Who are the new validators in the merged validator group?
\end{enumerate}
The answer for (1) is encoded in $g_{x,n,k}$, namely $g_{x,n,k} = g_{x,n,k-1}
\mod 2^{m-k}$. The information for (2) is the first $c$ validators ordered by
their luck, where $c$ is a system parameter (undecided).

In essence, every node runs a deterministic algorithm on the latest consensus
result to determine the consensus group, validator group, etc. for the next
consensus run. With this, we circumvent the permissionless problem with
classical BFT algorithms.

\subsection{Bootstrapping}\label{sec:bootstrapping}
The protocol need to be bootstrapped, in this section we provide a possible
method.

From the genesis block up until some consensus run $n$, there would be no
validator group formation, all validation are performed by servers run by the
developer. This property can be implemented in the client software (similar to
how Bitcoin performs protocol upgrades). From consensus run $n+1$ and onwards,
the validator groups begin to form and the system begins to run the full
bottom-up consensus protocol.

\section{Choosing the right BFT consensus algorithm}
PBFT~\cite{castro1999practical} assumes weak synchrony. It uses a primary node
to initiate the algorithm which may not fit nicely with our model. But it is
somewhat successful in Tendermint and Hyperledger (citation and verification
needed).

HoneyBadgerBFT~\cite{miller2016honey} and Cachin's secure broadcast
protocol~\cite{cachin2001secure} may suite our needs more, but further
investigation is needed. The advantage of these two protocols over PBFT is that
they work in the asynchronous environment.

The Swirlds~\cite{baird2016swirlds} algorithm is a departure from the classical
BFT algorithms. Nodes use ``gossip about gossip'' to determine whether a
majority of the population has heard about a particular transaction. So
agreement is achieved not after one execution of the algorithm but after the
nodes have eventually exchanged enough messages---we call this ``eventual
consensus''. The downside of Swirlds is that they assume static membership, thus
it is not permissionless. Swirlds may be applicable to our design where we can
use our group formation technique to select the validators and run Swirlds on
top (assuming Swirlds scales well). This eleminates the need for the recursive
(bottom-up) phase.

% TODO Swirlds need good scalability and convergence properties to be useful for
% us. With our bootstrapping, perhaps we can infer the member count in Swirlds
% from its history!

\section{Limitations, Discussion and Questions}
\begin{itemize}
\item What's the mechanisms for publishing double-spends?
\item Doesn't prevent spam, so DoS is possible. But we can work around this by
  ignoring spammy nodes.
\item If there is stake in the system, we can use it to elect validator, this
  would prevent the sybil attack. But I argue sybil attack is application
  dependent.
\item No incentive in this system, but again it's application specific.
\item If a node is offline for some time, it will not be in any consensus group.
  To join a consensus group, it must make a transaction to a node that is in a
  consensus group.
\item Not all the validators are online (chrun), but we can model it using the
  Sleep consensus model~\cite{bentov2016sleepy}.
\item This model does not require a single connected component. 
\item I personally don't see a lot of benefit with the idea of spontaneously
  growing the set of valid transaction by traversing the DAG, in the end
  everything must be validated anyway, and I don't see any major advantages of
  that technique. It also must assume single connected component, where as this
  approach does not.
\end{itemize}

\subsection{Merging consensus groups does not guarantee BFT with a probability
  of 1}
Our assumption is that $n > 3f$, this can be alternatively written as $n = 3f +
1$, for the set of all validators. Suppose the validators are put into $c$
groups and each group has $\frac{n}{c}$ validators. The bottom-up consensus
protocol requires every validator group to reach consensus, thus if a group has
$\frac{n}{3c}$ or more Byzantine nodes, then the group cannot reach consensus.
We claim if Byzantine nodes can freely choose validator groups, then they can
prevent consensus for $c-1$ groups.

We can write
$$
f = \frac{n-1}{3} = x \frac{n}{3c},
$$
where $x$ is the number validator groups that the Byzantine group can
infiltrate. Suppose $x = c$, then
$$f = \frac{n-1}{3} < c \frac{n}{3c} = \frac{n}{3},$$
this implies that the number Byzantine nodes is insufficient to infiltrate all
$c$ groups. Now suppose $x = c - 1$, then we claim
$$f > (c-1) \frac{n}{3c}$$
for $n > c$. This is easy to see
\begin{align*}
  f = \frac{n-1}{3} &> (c-1)\frac{n}{3c} \\
  n - 1 &> n\frac{c-1}{c} \\
  \frac{n-1}{n} &> \frac{c-1}{c} \\
  n &> c.
\end{align*}
This result implies that there is a sufficient number of Byzantine nodes to
infiltrate $c - 1$ groups.

While this disheartening, there may be a few ways to improve on it or look at it
in a different light.
\begin{itemize}
\item The analysis assumes that the Byzantine nodes can choose which group to
  join, but in our protocol they cannot. Thus the aforementioned result is the
  worst case scenario.
\item Perhaps we do not aim BFT with a probability of 1, but try to achieve BFT
  with a sufficiently high probability. Randomisation of group membership helps
  here.
\item Does our approach break down when a consensus run is infiltrated by
  Byzantine nodes? Can we recover afterwards? My intuition says yes, because
  honest nodes will always detect malicious transactions due to the TrustChain
  structure. Thus they will not accept the ``validated'' transactions from the
  Byzantine nodes, and they can publish a proof of malicious transaction in
  future consensus runs.
\item There is always one consensus group that is not infiltrated, can it do
  something about the situation?
\item The bottom-up approach is aimed to make BFT consensus algorithms scale. In
  some applications it may be sufficient even without the bottom-up approach
  (HoneyBadgerBFT can handle thousands of transactions on 64
  nodes~\cite{miller2016honey}).
\item The comparison (theoretical and practical) between bottom-up and
  non-bottom-up would make an interesting thesis topic.
\end{itemize}

\subsection{Tail bounding the number of Byzantine nodes in a consensus group}
We want to answer the following question. \emph{Suppose there is an urn that has
  $N$ balls, $M = (2N+1)/3$ of them are white and $N - M$ of them are black.
  $N/c$ balls are drawn without replacement. What is the probability that the
  number of white balls drawn is lower than the expected number of white balls?}
This question maps directly to our problem of Byzantine nodes, where the urn is
the set of all validators and they form $c$ groups each of size $N/c$. The white
balls represent honest nodes and the black balls represent Byzantine nodes.

The problem is modelled by the hypergeometric distribution\footnote{This is a
  variation of the geometric distribution where the balls are picked without
  replacement.} and we can solve it using tail inequalities.

Using the notation and definition in~\cite{skala2013hypergeometric}, the tail
inequality (not necessarily tight) is given by
\[
\Pr[ i \le E[i] - tn ] \le e^{-2t^2n},
\]
where $i$ is the random variable (the number of white balls), $E[i] = n
\frac{M}{N}$ is the expected value and $n = N/c$ is the number of balls drawn.
For our problem, we are interested in the probability that the number of white
balls is lower than the expected value. Thus we can write $\Pr[i \le E[i] - 1]$,
and then
\begin{align*}
  tn &= 1 \\
  tN/c &= 1 \\
  t &= c/N.
\end{align*}
Finally we get
\begin{align*}
  \Pr[ i \le E[i] - 1] &\le e^{-2 (c/N)^2 (N/c)} \\
  \Pr[ i \le E[i] - 1] &\le e^{-2 (c/N)}
\end{align*}

To minimise the probability, we must maximise $c/N$. But $c/N < 1$, so parameter
tuning cannot help us to reach a vanishingly small probability. This may be the
nail in the coffin for the current algorithm.

\subsection{How to set the initial validator group size?}
This can be modelled as the occupancy problem.

%  LocalWords:  BFT Merkle checkpointing permissionless setup

### 2017-01-23 chapters/ch2-bottom-up.tex
Clarify setup phase, settings for Emacs

@@
 To prepare a chain on node $x$ for validation in consensus run $n$, $x$ first
-communicates to the validators of consensus run $n-1$ for the valid set of
-blocks (recall that we assume every node knows the validators), and then
-generates a new checkpoint block (\cref{fig:protocol-setup}). This completes
-the setup phase.
+communicates to the validators of consensus run $n-1$ for the Merkle hash of the
+valid set of blocks (recall that we assume every node knows the validators), and
+then generates a new checkpoint block (\cref{fig:protocol-setup}). This
+completes the setup phase.
 
@@
 %  LocalWords:  BFT Merkle checkpointing permissionless setup
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "../thesis"
+%%% End:

### 2017-01-23 chapters/ch2-bottom-up.tex
Is the traditional Byzantine Agreement too strict?

@@
 
-The Swirlds~\cite{baird2016swirlds} algorithm is a departure from the classical
-BFT algorithms. Nodes use ``gossip about gossip'' to determine whether a
-majority of the population has heard about a particular transaction. So
-agreement is achieved not after one execution of the algorithm but after the
-nodes have eventually exchanged enough messages---we call this ``eventual
-consensus''. The downside of Swirlds is that they assume static membership, thus
-it is not permissionless. Swirlds may be applicable to our design where we can
-use our group formation technique to select the validators and run Swirlds on
-top (assuming Swirlds scales well). This eleminates the need for the recursive
-(bottom-up) phase.
-
-% TODO Swirlds need good scalability and convergence properties to be useful for
-% us. With our bootstrapping, perhaps we can infer the member count in Swirlds
-% from its history!
-
 \section{Limitations, Discussion and Questions}
@@
 
+\section{Is the traditional Byzantine Agreement too strict?}
+If nodes are in agreement, they share the same state machine. But is this too
+strict for our use case?
+
+Suppose there is set of transactions $A \rightarrow B \rightarrow C \rightarrow
+D$ that depend on each other. Then half of the nodes agree on a subset of those
+transactions, i.e. $\{A, B, C\}$. Another half agree on all four transactions.
+This is not Byzantine agreement in the traditional sense, because there does not
+exist a majority that share the same state machine. But indirectly they all
+agree on the set intersection. Can we build a model that support this type of
+consensus?
+
 %  LocalWords:  BFT Merkle checkpointing permissionless setup

### 2017-01-23 chapters/ch2-bottom-up.tex
Some ideas and questions. Thanks @Zhijieren

@@
   approach does not.
+\item What if no or very few transactions happen in a consensus run? How do we
+  select validators in this case? Perhaps select validators according to a
+  different metric. Closeness metric between the hash and some value perhaps.
+\item Given a transaction, how does a node efficiently validate its G
+  authenticity?
+\item Force nodes to only create checkpoint when their previous checkpointed
+  chain has been validated?
 \end{itemize}
@@
 
-\section{Is the traditional Byzantine Agreement too strict?}
-If nodes are in agreement, they share the same state machine. But is this too
-strict for our use case?
-
-Suppose there is set of transactions $A \rightarrow B \rightarrow C \rightarrow
-D$ that depend on each other. Then half of the nodes agree on a subset of those
-transactions, i.e. $\{A, B, C\}$. Another half agree on all four transactions.
-This is not Byzantine agreement in the traditional sense, because there does not
-exist a majority that share the same state machine. But indirectly they all
-agree on the set intersection. Can we build a model that support this type of
-consensus?
-
 %  LocalWords:  BFT Merkle checkpointing permissionless setup

### 2017-02-02 chapters/ch2-checkpoint.tex
Checkpoint consensus, some definitions

\section{Preliminaries}

\subsection{Requirements}
\begin{itemize}
\item Permissionless
\item Byzantine fault tolerant
\item No PoW
\item Works under churn
\item Underlying data structure is TrustChain
\item Detects forks or double-spends
\item No step in the protocol blocks transactions
\item Application independent
\end{itemize}

\subsection{Assumptions}
\begin{itemize}
\item We elect $N$ consensus promoters in every round, we assume the number of
  faulty promoters is $f$ and $N = 3f + 1$.
\item Promoters have the complete history of the previously agreed set of
  transactions.
% \item Single connected component for TrustChain.
% \item No sybil attack % 
\end{itemize}

\subsection{Notations}
\begin{itemize}

\item $y = H(x)$ is a cryptographically secure hash function (random oracle),
  the domain $x$ is infinite and the range is $y \in \{0, 2^{256}-1\}$.

\item Every node in the system has an identifier $i$ and a blockchain
  $$B_i = \{b_{i,j} : j \in \{1 \dots h\} \},$$
  where $h$ is the height of the chain. Note that $h = |B_i|$

\item Each block $b_{i,j}$ has a type $t \in \{ \tau, \gamma \}$, denoted by
  $b^t_{i,j}$. Blocks without the superscript can be of any type.

\item $T(b_{i, j}) = \{ \tau, \gamma \}$ is the type function, where its domain
  is a block and outputs the corresponding type of the block.

\item A block of type $\tau$ is a \emph{transaction block}. It is a six-tuple,
  i.e. $$b^\tau_{i,j} = (H(b_{i,j-1}), h_s, h_r, s_s, s_r, m).$$ $h_s$ and $h_r$
  denote the height (the sequence number for when the transaction is made) of
  the sender and receiver respectively. $s_s$ and $s_r$ denote the signature of
  the sender and the receiver respectively.

\item A block of type $\gamma$ is a \emph{checkpoint block}. It is a
  three-tuple, i.e.
  $$b^\gamma_{i,j} = (H(b_{i,j}), H(\mathcal{C}_r), p)$$
  where $\mathcal{C}_r$ is the consensus result in round $r$ and $p \in {0,1}$
  which indicates whether $i$ wish to become a promoter in the following
  consensus round.

\item Given the input $b^\tau_{i,j}$, we define the
  \emph{get-neighbouring-checkpoints function}
  $$C(b^\tau_{i,j}) = (b^\gamma_{i,a}, b^\gamma_{i,b})$$
  where $a = \argmin_{k, k < j, T(b_{i,k}) = \gamma}(j - k)$ and $b =
  \argmin_{k, k > j, T(b_{i,k}) = \gamma}(k - j)$.

\item Given two $\gamma$ transactions, we define \emph{get-piece function}
  $$P(b^\gamma_{i,a}, b^\gamma_{i,b}) = \{b_{i,j} : b_{i,j} \in B_i, a \le j \le
  b\}.$$

\item Given two blockchains and a message, we define the \emph{do-transaction
    function}
  $$X_\tau(B_s, B_r, m) = (B'_s, B'_r)$$
  where $B'_s = \{(H(b_{s,h_s}), h_s + 1, h_r + 1, s_s, s_r, m)\} \cup B_s$,
  $B'_r = \{(H(b_{r, h_r}), h_s + 1, h_r + 1, s_s, s_r, m)\} \cup B_r$, $h_s =
  |B_s|$ and $h_r = |B_r|$.

\item Given a blockchain, we define the \emph{do-checkpoint function}
  $$X_\gamma(B_i, r, p) = \{ (H(b_{i,h}), H(\mathcal{C}_r), p) \}\cup B_i$$ where
  $h = |B_i|$, and $\mathcal{C}_r$ is the latest consensus result.

\item Note that $X_\tau$ and $X_\gamma$ perform a state transition.

\item The result of a consensus in round $r$ is a set of two-tuple of
  checkpoints. Namely, $\mathcal{C}_r = \{ (b^\gamma_{i,a}, b^\gamma_{i,b}) : a
  < b, \text{agreed by the promoters}\}$.
\end{itemize}

\section{Checkpoint consensus}

\subsection{Promoter registration}
Node $i$ can register as a promoter when the latest consensus result is
announced (suppose after the completion of round $r-1$), then it generates a new
block using $b = T_\gamma(B_i, r-1, 1)$. The current promoters (in round $r$)
may decide to include $b$ in the new consensus result. If $b$ is in it, then $i$
becomes one of the promoter of round $r+1$.

We can fix the number of promotors to $N$ by sorting the promotors by their
``luck value''' and taking the first $N$.

\subsection{Setup phase}
We begin in the state where $\mathcal{C}_{r-1}$ has just been agreed but has not
been disseminated yet.
\begin{lemma}
  Nodes in the network can always verify the validity of the consensus result.
\end{lemma}
\begin{proof}
\end{proof}

\begin{lemma}
  The new set of promoter for the next consensus round is consistent with
  respect to all the nodes in the network.
\end{lemma}

\begin{lemma}
  Promoters waiting for a some time $\Delta$ to collect transactions does not
  violate the asynchronous assumption.
\end{lemma}

\begin{corollary}
  The setup phase satisfies the validity, correctness and termination
  properties.
\end{corollary}

\subsection{Consensus phase}

\section{Fraud detection}
Here we provide two techniques for fraud detection. The first guarantees fraud
detection but is not practical. The second is a randomised solution that detects
fraud with a high probability.

\subsection{Breadth first search}

\subsection{Random sampling}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../thesis"
%%% End:

### 2017-02-03 chapters/ch2-checkpoint.tex
Validation proof WIP

@@
 \begin{itemize}
+\item Asynchronous network
+\item Private and authenticated channel
 \item We elect $N$ consensus promoters in every round, we assume the number of
@@
 \item Promoters have the complete history of the previously agreed set of
-  transactions.
+  transactions (TX).
 % \item Single connected component for TrustChain.
@@
 
-\item A block of type $\tau$ is a \emph{transaction block}. It is a six-tuple,
-  i.e. $$b^\tau_{i,j} = (H(b_{i,j-1}), h_s, h_r, s_s, s_r, m).$$ $h_s$ and $h_r$
-  denote the height (the sequence number for when the transaction is made) of
-  the sender and receiver respectively. $s_s$ and $s_r$ denote the signature of
-  the sender and the receiver respectively.
+\item A block of type $\tau$ is a \emph{transaction block} or \emph{TX block}.
+  It is a six-tuple, i.e. $$b^\tau_{i,j} = (H(b_{i,j-1}), h_s, h_r, s_s, s_r,
+  m).$$ $h_s$ and $h_r$ denote the height (the sequence number for when the TX
+  is made) of the sender and receiver respectively. $s_s$ and $s_r$ denote the
+  signature of the sender and the receiver respectively. $i = \{s, r\}$ and $j =
+  \{h_s, h_r\}$ and we call the block a sender block or a receiver block
+  respectively.
 
-\item A block of type $\gamma$ is a \emph{checkpoint block}. It is a
-  three-tuple, i.e.
+\item Given two TX blocks $b^\tau_{i,j}$ and $b^\tau_{i',j'}$, but $i \neq i'$.
+  If $h_s = h'_s$, $h_r = h'_r$, $m = m'$ and the signatures are valid, then we
+  call them a \emph{pair}.
+
+\item If there exist two TX blocks $b^\tau_{i,j}$ and $b^\tau_{i',j'}$, where
+  $s_s$ and $s_s'$ is created by the same public key and $h_s = h_s'$ but $i
+  \neq i'$, then we call this a \emph{fork}.
+
+\item A TX block has two \emph{inputs} blocks, to retrieve the input we define
+  the \emph{get-tx-input function} $I^\tau$. Let the pair of $b^\tau_{i,j}$ be
+  $b^\tau_{i',j'}$, then
+  $$I^\tau(b^\tau_{i,j}) = (b_{i, j-1}, b_{i', j'-1}).$$
+
+\item A block of type $\gamma$ is a \emph{checkpoint block} or \emph{CP block}.
+  It is a three-tuple, i.e.
   $$b^\gamma_{i,j} = (H(b_{i,j}), H(\mathcal{C}_r), p)$$
@@
 
+\item A CP block has one input, we define the \emph{get-cp-input
+    function} $$I^\gamma(b^\gamma_{i,j}) = b_{i,j-1}.$$
+
 \item Given the input $b^\tau_{i,j}$, we define the
-  \emph{get-neighbouring-checkpoints function}
+  \emph{get-nearest-cp function}
   $$C(b^\tau_{i,j}) = (b^\gamma_{i,a}, b^\gamma_{i,b})$$
@@
 
-\item Given two $\gamma$ transactions, we define \emph{get-piece function}
+\item Given two $\gamma$ blocks, we define \emph{get-piece function}
   $$P(b^\gamma_{i,a}, b^\gamma_{i,b}) = \{b_{i,j} : b_{i,j} \in B_i, a \le j \le
@@
 
-\item Given two blockchains and a message, we define the \emph{do-transaction
-    function}
+\item Given two blockchains and a message, we define the \emph{do-tx-function}
   $$X_\tau(B_s, B_r, m) = (B'_s, B'_r)$$
@@
 
-\item Given a blockchain, we define the \emph{do-checkpoint function}
+\item Given a blockchain, we define the \emph{do-cp function}
   $$X_\gamma(B_i, r, p) = \{ (H(b_{i,h}), H(\mathcal{C}_r), p) \}\cup B_i$$ where
@@
 
-\item The result of a consensus in round $r$ is a set of two-tuple of
-  checkpoints. Namely, $\mathcal{C}_r = \{ (b^\gamma_{i,a}, b^\gamma_{i,b}) : a
-  < b, \text{agreed by the promoters}\}$.
+\item The result of a consensus in round $r$ is a set of two-tuple of CP.
+  Namely, $\mathcal{C}_r = \{ (b^\gamma_{i,a}, b^\gamma_{i,b}) : a < b,
+  \text{agreed by the promoters}\}$.
 \end{itemize}
@@
 We begin in the state where $\mathcal{C}_{r-1}$ has just been agreed but has not
-been disseminated yet.
+been disseminated yet. The promoters
+
 \begin{lemma}
-  Nodes in the network can always verify the validity of the consensus result.
+  If a node sees a valid $\mathcal{C}_r$ and another node sees a valid
+  $\mathcal{C}'_r$, then $\mathcal{C}_r = \mathcal{C}'_r$.
 \end{lemma}
@@
 \subsection{Consensus phase}
+We need an atomic broadcast algorithm for the consensus phase. We use a similar
+but simplified construction as~\cite{miller2016honey}, where atomic broadcast is
+constructed from the reliable broadcast\footnote{Reliable broadcast solves the
+  Byzantine generals problem.}~\cite{bracha1984asynchronous} and asynchronous
+common subset (ACS). The ACS protocol requires a binary Byzantine agreement
+protocol, and for that it needs a trusted dealer to distributed the secret
+shares. Promoters can check whether the secret shares are valid, but they cannot
+prevent the dealer from disclosing the secrets.
+
+There techniques that uses no dealers. First is to use
+PBFT~\cite{castro1999practical}, but we must change our asynchronous assumption
+into the weak synchrony assumption. Second is to use an inefficient binary
+Byzantine agreement protocol where its message complexity is $O(N^3)$ rather
+than $O(N^2)$ and becomes a bottleneck.
+
+Suppose we use a dealer, what is the effect to the algorithm if the dealer is
+malicious?
 
@@
 
-\subsection{Breadth first search}
+\subsection{Depth first search}
+\begin{algorithmic}
+\State{$bs = \emptyset$}
 
+\Procedure{reverse-dfs}{$G, b$}
+\State {$bs \gets {b} \cup bs$}
+\If{$t = \text{genesis}$}
+    \State \textbf{return} $1$
+\EndIf
+\ForAll{$d \in I(b)$}
+\If{\Call{valid}{$d$} = 0}
+   \State {\textbf{return} 0}
+\EndIf
+\EndFor
+
+\State {\textbf{return} 1}
+\EndProcedure
+\end{algorithmic}
+
+\begin{algorithmic}
+\Procedure{valid}{$b, bs$}
+
+\State $(b^\gamma_a, b^\gamma_b) = C(b)$
+\State $p = P(b^\gamma_a, b^\gamma_b)$
+\If{$(b^\gamma_a, b^\gamma_b)$ is in a consensus result}
+  \If{$p$ is a valid chain}
+    \If{$b$ does not have a \emph{fork} in $bs$}
+      \State {\textbf{return} $1$}
+    \EndIf
+  \EndIf
+\EndIf
+
+\State \textbf{return} $0$
+\EndProcedure
+\end{algorithmic}
+
+\begin{lemma}
+  The \verb!reverse-dfs! algorithm traverses all nodes that may have an effect
+  on block $b$.
+\end{lemma}
+\begin{proof}
+  Recall that the TrustChain is a DAG, which always has a topological ordering.
+  That is, for every edge $(b', b)$, $b'$ comes before $b$ in the ordering, and
+  nodes before $b$ are its ancestors. If we reverse all the edges, then $b'$
+  comes after $b$ in the ordering, and nodes before $b$ are its descendents.
+
+  \verb!reverse-dfs! is essentially running DFS with all the edges reversed. If
+  the DFS outputs the visited nodes in reverse postordering, then the output is
+  in topological order and contains all of $b$'s original ancestors. Due to the
+  hash pointers, only the ancestors can have an effect on $b$.
+\end{proof}
+
+\begin{lemma}
+  \verb!verify!...
+\end{lemma}
 \subsection{Random sampling}

### 2017-02-10 chapters/ch2-checkpoint.tex
Improve definition and write consensus protocol

--- 2017-02-02 chapters/ch2-checkpoint.tex
@@
 
-\subsection{Notations}
+\subsection{Notation and definition}
 \begin{itemize}
 
-\item $y = H(x)$ is a cryptographically secure hash function (random oracle),
-  the domain $x$ is infinite and the range is $y \in \{0, 2^{256}-1\}$.
+\item $y = \texttt{h}(x)$ is a cryptographically secure hash function (random
+  oracle), the domain $x$ is infinite and the range is $y \in \{0, 2^{256}-1\}$.
 
-\item Every node in the system has an identifier $i$ and a blockchain
-  $$B_i = \{b_{i,j} : j \in \{1 \dots h\} \},$$
-  where $h$ is the height of the chain. Note that $h = |B_i|$
+\item We model our system in the permissionless setting, where each
+  participating party has a unique identity $i$, and a chain $B_i$.
 
-\item Each block $b_{i,j}$ has a type $t \in \{ \tau, \gamma \}$, denoted by
-  $b^t_{i,j}$. Blocks without the superscript can be of any type.
+\item A chain is a collection of blocks
+  $B_i = \{b_{i,j} : j \in \{1 \dots h\} \},$
+  the blocks are linked together using hash pointers, similar to bitcoin. All
+  blocks contain a reference to the previous block, the very first block with no
+  references is the genesis block.
 
-\item $T(b_{i, j}) = \{ \tau, \gamma \}$ is the type function, where its domain
-  is a block and outputs the corresponding type of the block.
+  The sequence number of the block begins at 0 on the genesis block and is
+  incremented for every new block. The height of the chain is $h = |B_i|$.
 
-\item A block of type $\tau$ is a \emph{transaction block}. It is a six-tuple,
-  i.e. $$b^\tau_{i,j} = (H(b_{i,j-1}), h_s, h_r, s_s, s_r, m).$$ $h_s$ and $h_r$
-  denote the height (the sequence number for when the transaction is made) of
-  the sender and receiver respectively. $s_s$ and $s_r$ denote the signature of
-  the sender and the receiver respectively.
+\item There are two sets of blocks $T_i$ and $C_i$, where $T_i \cup
+  C_i = B_i$ and $T_i \cap C_i = \emptyset$. This can be seen as the
+  block type, where $t_{i,j}$ and $c_{i,j}$ to represent a
+  \emph{transaction block (TX block)} and \emph{checkpoint block (CP block)}
+  respectively.
 
-\item A block of type $\gamma$ is a \emph{checkpoint block}. It is a
-  three-tuple, i.e.
-  $$b^\gamma_{i,j} = (H(b_{i,j}), H(\mathcal{C}_r), p)$$
+\item $\texttt{typeof}: B_i \rightarrow \{ \tau, \gamma \}$ returns the
+  corresponding type of the block.
+
+\item A block of type $\tau$ is a six-tuple, i.e. $$t_{i,j} =
+  (\texttt{h}(b_{i,j-1}), h_s, h_r, s_s, s_r, m).$$ $h_s$ and $h_r$ denote the
+  height (the sequence number for when the TX is made) of the sender and
+  receiver respectively. $s_s$ and $s_r$ are the signatures of the sender and
+  the receiver respectively. $i = \{s, r\}$ and $j = \{h_s, h_r\}$ depending on
+  whether $i$ is the sender or the receiver.
+
+\item Given two TX blocks $t_{i,j}$ and $t_{i',j'}$, but $i \neq i'$. If $h_s =
+  h'_s$, $h_r = h'_r$, $m = m'$ and the signatures are valid, then we call them
+  a \emph{pair}.
+
+\item If there exist two TX blocks $t_{i,j}$ and $t_{i',j'}$, where $s_s$ and
+  $s_s'$ is created by the same public key, $h_s = h_s'$, but $i \neq i'$, then
+  we call this a \emph{fork}.
+
+\item A block of type $\gamma$ is a three-tuple, i.e.
+  $$c_{i,j} = (\texttt{h}(b_{i,j}), H(\mathcal{C}_r), p)$$
   where $\mathcal{C}_r$ is the consensus result in round $r$ and $p \in {0,1}$
@@
 
-\item Given the input $b^\tau_{i,j}$, we define the
-  \emph{get-neighbouring-checkpoints function}
-  $$C(b^\tau_{i,j}) = (b^\gamma_{i,a}, b^\gamma_{i,b})$$
-  where $a = \argmin_{k, k < j, T(b_{i,k}) = \gamma}(j - k)$ and $b =
-  \argmin_{k, k > j, T(b_{i,k}) = \gamma}(k - j)$.
+\item We define
+  $$\texttt{newtx}: B_i \times B_j \times M \rightarrow B_i \times B_j$$
+  as a function that creates new TX blocks. Its functionality is to
+  extend the given chains using the following rule. If the input is $(B_s, B_r,
+  m)$ then the output is $(B'_s, B'_r)$ where $B'_s = \{(\texttt{h}(b_{s,h_s}), h_s + 1,
+  h_r + 1, s_s, s_r, m)\} \cup B_s$, $B'_r = \{(\texttt{h}(b_{r, h_r}), h_s + 1, h_r + 1,
+  s_s, s_r, m)\} \cup B_r$, $h_s = |B_s|$ and $h_r = |B_r|$.
 
-\item Given two $\gamma$ transactions, we define \emph{get-piece function}
-  $$P(b^\gamma_{i,a}, b^\gamma_{i,b}) = \{b_{i,j} : b_{i,j} \in B_i, a \le j \le
-  b\}.$$
+\item We define
+  $$\texttt{newcp}: B_i \times \mathbb{R}_{\ge 1} \times \{0, 1\} \rightarrow B_i$$
+  as a function that creates new CP blocks. Concretely, given $(B_i, r, p)$, it
+  results in $\{ (\texttt{h}(b_{i,h}), H(\mathcal{C}_r), p) \}\cup B_i$ where $h =
+  |B_i|$, and $\mathcal{C}_r$ is the latest consensus result at round $r$.
 
-\item Given two blockchains and a message, we define the \emph{do-transaction
-    function}
-  $$X_\tau(B_s, B_r, m) = (B'_s, B'_r)$$
-  where $B'_s = \{(H(b_{s,h_s}), h_s + 1, h_r + 1, s_s, s_r, m)\} \cup B_s$,
-  $B'_r = \{(H(b_{r, h_r}), h_s + 1, h_r + 1, s_s, s_r, m)\} \cup B_r$, $h_s =
-  |B_s|$ and $h_r = |B_r|$.
+\item Note that in the actual system, \texttt{newtx} and \texttt{newcp} perform
+  a state transition.
 
-\item Given a blockchain, we define the \emph{do-checkpoint function}
-  $$X_\gamma(B_i, r, p) = \{ (H(b_{i,h}), H(\mathcal{C}_r), p) \}\cup B_i$$ where
-  $h = |B_i|$, and $\mathcal{C}_r$ is the latest consensus result.
+\item The result of a consensus in round $r$ is a set of two-tuple of CP.
+  Namely, $\mathcal{C}_r = \{ (c_{i,j}, c_{i,k}) : j < k, \text{agreed by the
+    promoters}\}$.
+\end{itemize}
 
-\item Note that $X_\tau$ and $X_\gamma$ perform a state transition.
+\subsection{Properties}
+\begin{itemize}
+\item A TX block has two \emph{ancestor} blocks. Let a pair be $t_{i,j}$ and
+  $t_{i',j'}$, then $(b_{i, j-1}, b_{i', j'-1})$ is the input block of $t_{i,j}$.
 
-\item The result of a consensus in round $r$ is a set of two-tuple of
-  checkpoints. Namely, $\mathcal{C}_r = \{ (b^\gamma_{i,a}, b^\gamma_{i,b}) : a
-  < b, \text{agreed by the promoters}\}$.
+\item A CP block has one \emph{ancestor} block, that is simply the block with
+  the previous sequence number, i.e. the ancestor of $c_{i,j}$ is $b_{i,j-1}$.
+
+\item Every TX block $t_{i,j}$ is enclosed by two CP blocks $(c_{i,a},
+  c_{i,b})$, where $a = \argmin_{k, k < j, \texttt{typeof}(b_{i,k}) = \gamma}(j
+  - k)$ and $b = \argmin_{k, k > j, \texttt{typeof}(b_{i,k}) = \gamma}(k - j)$.
+
+\item The blocks between two $\gamma$ blocks $(c_{i,a}, c_{i,b})$ is $\{b_{i,j}
+  : b_{i,j} \in B_i, a \le j \le b\}$.
+
 \end{itemize}
@@
 
+\subsection{Luck value}
+First we define the luck value $l_{i,j} = \texttt{h}(k_i || c_{i,j})$, where $k$ is the
+public key of $i$. A lower luck value equates to higher luck.
+
 \subsection{Promoter registration}
@@
 announced (suppose after the completion of round $r-1$), then it generates a new
-block using $b = T_\gamma(B_i, r-1, 1)$. The current promoters (in round $r$)
-may decide to include $b$ in the new consensus result. If $b$ is in it, then $i$
-becomes one of the promoter of round $r+1$.
+block $b_{i,j} = \texttt{newcp}(B_i, r-1, 1)$. The current promoters (in round
+$r$) may decide to include $b_{i,j}$ in the new consensus result. If $b$ is in
+it, then $i$ becomes one of the promoter of round $r+1$.
 
-We can fix the number of promotors to $N$ by sorting the promotors by their
+We can fix the number of promoters to $N$ by sorting the promoters by their
 ``luck value''' and taking the first $N$.
@@
 \subsection{Setup phase}
+The setup phase should satisfy the BFT conditions regarding the promoter
+selection, that is:
+\begin{enumerate}
+\item \emph{Agreement}: If any correct node outputs a promoter $p$, then every
+  correct node outputs $p$.
+\item \emph{Total Order}: If one correct node outputs the sequence of promoters
+  $\{p_1, p_i, \dots, p_n\}$ and another has output $\{p'_0, p'_1, \dots,
+  p'_{n'} \}$, then $p_i = p'_i$ for $i \le \min(n, n')$.
+\item \emph{Liveness}: All $N - f$ correct nodes terminate eventually.
+\end{enumerate}
+
 We begin in the state where $\mathcal{C}_{r-1}$ has just been agreed but has not
-been disseminated yet.
+been disseminated yet. The exact technique to disseminate $\mathcal{C}_{r-1}$ is
+irrelevant, broadcasting or gossiping are both sufficient. In fact,
+dissemination is not necessary, nodes interested in the result can simply query
+the promoters that created $\mathcal{C}_{r-1}$.
+
 \begin{lemma}
-  Nodes in the network can always verify the validity of the consensus result.
+  If a node sees a valid $\mathcal{C}_{r-1}$ and another node sees a valid
+  $\mathcal{C}'_{r-1}$, then $\mathcal{C}_{r-1} = \mathcal{C}'_{r-1}$.
 \end{lemma}
 \begin{proof}
+  (sketch) $\mathcal{C}_{r-1}$ is signed by at least $N-f$ promoters from round
+  $r-1$.
 \end{proof}
+
+The potential promoters now need to first discover whether
+they are the first $N$ lucky promoters.
 
@@
 \end{lemma}
+\begin{proof}
+  (sketch) All nodes use the same deterministic function to compute the luck
+  value.
+\end{proof}
+
+Nodes should wait for some time to collect the CP blocks, so they wait for some
+time $\Delta$ before moving on to the next phase.
 
@@
 \end{lemma}
+\begin{proof}
+  (sketch) This can be seen as a long delay in the asynchronous system.
+\end{proof}
 
 \begin{corollary}
-  The setup phase satisfies the validity, correctness and termination
-  properties.
+  The setup phase satisfies the validity, agreement and liveness properties.
 \end{corollary}
+\begin{proof}
+  (sketch) Validity and agreement is satisfied because the promoters are
+  computed using a deterministic algorithm using the latest consensus result.
+  Liveness is satisfied because $\mathcal{C}_{r-1}$ is eventually propagated to
+  all node by gossiping.
+\end{proof}
 
 \subsection{Consensus phase}
+The consensus phase should satisfy the BFT conditions regarding the CP blocks,
+that is:
+\begin{enumerate}
+\item \emph{Agreement}: If any correct node outputs a CP block $c$, then every
+  correct node outputs $c$.
+\item \emph{Total Order}: If one correct node outputs the sequence of CP blocks
+  $\{c_1, c_i, \dots, c_n\}$ and another has output $\{c'_0, c'_1, \dots,
+  c'_{n'} \}$, then $c_i = c'_i$ for $i \le \min(n, n')$.
+\item \emph{Liveness}: All $N - f$ correct nodes terminate eventually.
+\end{enumerate}
+
+We need an atomic broadcast algorithm for the consensus phase. We use a similar
+but simplified construction as~\cite{miller2016honey}, where atomic broadcast is
+constructed from the reliable broadcast\footnote{Reliable broadcast solves the
+  Byzantine generals problem.}~\cite{bracha1984asynchronous} and asynchronous
+common subset (ACS). The ACS protocol requires a binary Byzantine agreement
+protocol, and for that it needs a trusted dealer to distributed the secret
+shares. Promoters can check whether the secret shares are valid, but they cannot
+prevent the dealer from disclosing the secrets.
+
+There techniques that uses no dealers. First is to use
+PBFT~\cite{castro1999practical}, but we must change our asynchronous assumption
+into the weak synchrony assumption. Most likely this is not possible because of
+the ``When to start?'' problem. It's difficult to give a bounded delay for
+propagating the consensus result.
+
+Second is to use an inefficient binary Byzantine agreement protocol where its
+message complexity is $O(N^3)$ rather than $O(N^2)$ and becomes a bottleneck. Or
+a suboptimal one, e.g. $n/5$ instead of $n/3$.
+
+Suppose we use a dealer, what is the effect to the algorithm if the dealer is
+malicious?
 
 \section{Fraud detection}
-Here we provide two techniques for fraud detection. The first guarantees fraud
-detection but is not practical. The second is a randomised solution that detects
-fraud with a high probability.
 
-\subsection{Breadth first search}
+\subsection{Threat model}
+\begin{itemize}
+\item What is the influence of a fork for future transactions?
+\item What if both parties (sender and receiver) are malicious?
+\end{itemize}
 

### 2017-02-10 chapters/ch2-checkpoint.tex
Existence testing

@@
   a state transition.
+
+\item We define
+  $$\texttt{getr}: C_i \rightarrow \mathbb{R}_{\ge 1}$$
+  as a function that gets the consensus round number used to create the given CP
+  block.
+
+\item The CP blocks that follows a pair of TX blocks must be created using the
+  same $\mathcal{C}_r$, otherwise the transaction is invalid. Concretely, given
+  a pair $t_{i,j}$ and $t_{i',j'}$, then there exist $c_{i, k}$ and $c_{i', k'}$
+  where $j < k$, $j' < k'$, $\{ \texttt{typeof}(b_{i, x}) : x \in \{j, \dots k -
+  1\}\}$ are all $\tau$, $ \{ \texttt{typeof}(b_{i', x}) : x \in \{ j', \dots k'
+  - 1\} \}$ are all $\tau$ and $\texttt{getr}(c_{i, k}) = \texttt{getr}(c_{i',
+    k'})$. This constraint is not the result of the consensus protocol, but the
+  validation protocol.
 
@@
 \section{Fraud detection}
+Fraud detection is application specific. In the context of electronic cash, a
+fraudulent transaction will have an effect in subsequent transactions. In the
+context of P2P file sharing, the transactions are reports of the number of bytes
+uploaded and/or downloaded, thus they are independent of each other.
+
+We focus in the latter case, where trasactions are independent.
 
@@
 
+\subsection{Existence testing}
+Suppose node $i$ created a TX $t_{i,j}$ and its pair is $t_{i',j'}$. Our
+protocol states that there exist two CP block $c_{i,k}$ and $c_{i',k'}$ such
+that $k>j$ and $k' > j'$, but all blocks between $(k, j)$ and $(k', j')$ are of
+type $\tau$. To verify whether $t_{i', j'}$ exists, $i$ performs the following
+tasks:
+\begin{enumerate}
+\item Waits for $c_{i', k'}$ to be in consensus.
+\item Finds $c_{i', k''}$ such that $k'' < j' < k'$ and $k''$ is also in
+  consensus (in some previous consensus round).
+\item Asks $i'$ for $\{b_{i', x} : x \in \{ k'', \dots k' \}\}$, and verified
+  that all the blocks follow the rules of TrustChain and contains $t_{i', j'}$.
+\end{enumerate}
+
 \subsection{Random sampling}

### 2017-02-13 chapters/ch2-checkpoint.tex
Notes on random sampling

@@
 \subsection{Luck value}
-First we define the luck value $l_{i,j} = \texttt{h}(k_i || c_{i,j})$, where $k$ is the
-public key of $i$. A lower luck value equates to higher luck.
+First we define the luck value $l_{i,j} = \texttt{h}(k_i || c_{i,j})$, where $k$
+is the public key of $i$. A lower luck value equates to higher luck.
 
@@
 ``luck value''' and taking the first $N$.
+
+\subsection{Promoter invitation}
+The output of promoter registration is a random set of $N$ promoters. If 1/3 of
+the population is malicious, then we cannot guarantee that the chosen promoters
+satisfy the $< n/3$ requirement.
+
+Promoter invitation is an attempt to involve human in the protocol. A naive
+method method is to use $N$ tickets, and then distribute them to trusted nodes.
+Nodes with the ticket can forward it to others. We have to rely on the humans to
+always forward the tickets to other honest humans. Finally, the nodes that hold
+a ticket are promoters. The result is that we have a permissioned system.
 
@@
 
+Node $i$ may wish to perform existence testing for all its unverified TX blocks.
+
 \subsection{Random sampling}
+Existence testing detects fraud when a single party of the TX is malicious. To
+detect fraud when both parties are malicious, every node performs existence
+testing on all TX blocks between CP blocks as if they are some other node. The
+assignment is determined using CP block digest where every node is assigned to
+the nearest node that has a higher digest, cycle back if there is no higher
+digest.
+
+If the number of malicious parties is low, then it's more probable that a fork
+is detected. If where there are a large number of malicious parties, we repeat
+the random sampling process.
 

### 2017-02-16 chapters/ch2-checkpoint.tex
Write some validation proofs

--- 2017-02-10 chapters/ch2-checkpoint.tex
@@
 
-\subsection{Notation and definition}
+\subsection{Notation, definition and properties}
 \begin{itemize}
@@
 
-\item Given two TX blocks $t_{i,j}$ and $t_{i',j'}$, but $i \neq i'$. If $h_s =
-  h'_s$, $h_r = h'_r$, $m = m'$ and the signatures are valid, then we call them
-  a \emph{pair}.
+\item Given two TX blocks $t_{i,j}$ and $t_{i',j'}$, if $i \ne i'$, $i = s$, $i'
+  = r$, $h_s = h'_s$, $h_r = h'_r$, $m = m'$ and the signatures are valid, then
+  we call them a \emph{pair}. Note that given one TX block, the pair can be
+  determined directly.
 
@@
 
-\item The result of a consensus in round $r$ is a set of two-tuple of CP.
-  Namely, $\mathcal{C}_r = \{ (c_{i,j}, c_{i,k}) : j < k, \text{agreed by the
-    promoters}\}$.
-\end{itemize}
+\item We define
+  $$\texttt{round}: C_i \rightarrow \mathbb{R}_{\ge 1}$$
+  as a function that gets the consensus round number used to create the given CP
+  block.
 
-\subsection{Properties}
-\begin{itemize}
+\item The CP blocks that follows a pair of TX blocks must be created using the
+  same $\mathcal{C}_r$, otherwise the transaction is invalid. Concretely, given
+  a pair $t_{i,j}$ and $t_{i',j'}$, then there exist $c_{i, k}$ and $c_{i', k'}$
+  where $j < k$, $j' < k'$, $\{ \texttt{typeof}(b_{i, x}) : x \in \{j, \dots k -
+  1\}\}$ are all $\tau$, $ \{ \texttt{typeof}(b_{i', x}) : x \in \{ j', \dots k'
+  - 1\} \}$ are all $\tau$ and $\texttt{round}(c_{i, k}) = \texttt{round}(c_{i',
+    k'})$. This constraint is not the result of the consensus protocol, but the
+  validation protocol.
+
+\item The result of a consensus in round $r$ is a set of CP blocks. Namely,
+  $$\mathcal{C}_r = \{ c_{i,j} : \text{agreed by the promoters}\}.$$
+
+\item A TX block can be valid, invalid or unknown. All TX blocks begin as
+  unknown, they can be validated using our validation protocol.
+
 \item A TX block has two \emph{ancestor} blocks. Let a pair be $t_{i,j}$ and
@@
 \item Every TX block $t_{i,j}$ is enclosed by two CP blocks $(c_{i,a},
-  c_{i,b})$, where $a = \argmin_{k, k < j, \texttt{typeof}(b_{i,k}) = \gamma}(j
-  - k)$ and $b = \argmin_{k, k > j, \texttt{typeof}(b_{i,k}) = \gamma}(k - j)$.
+  c_{i,b})$, where
+  $$a = \argmin_{k, k < j, \texttt{typeof}(b_{i,k}) = \gamma}(j - k),$$
+  $$b = \argmin_{k, k > j, \texttt{typeof}(b_{i,k}) = \gamma}(k - j).$$
+  We call this the \emph{enclosure} of $t_{i,j}$.
 
-\item The blocks between two $\gamma$ blocks $(c_{i,a}, c_{i,b})$ is $\{b_{i,j}
-  : b_{i,j} \in B_i, a \le j \le b\}$.
+\item The \emph{piece} of $t_{i,j}$ defined by the enclosure $(c_{i,a},
+  c_{i,b})$ is $\{b_{i,j} : a \le j \le b\}$.
+  We define
+  $$ \texttt{pieces}: B_i \rightarrow P(B_i) $$
+  as the function that returns the pieces ($P$ denotes power set).
+
+\item Every TX block $t_{i,j}$ is enclosed by two \emph{agreed} CP blocks
+  $(c_{i,a}, c_{i,b})$ (CP blocks that are in some consensus result), where
+  $$a = \argmin_{k, k < j, \texttt{typeof}(b_{i,k}) = \gamma}(j - k),$$
+  $$b = \argmin_{k, k > j, \texttt{typeof}(b_{i,k}) = \gamma}(k - j).$$
+  We call this the \emph{agreed enclosure} of $t_{i,j}$.
+
+\item The \emph{agreed piece} of $t_{i,j}$ defined by the agreed enclosure
+  $(c_{i,a}, c_{i,b})$ is $\{b_{i,j} : a \le j \le b\}$. Note that \emph{piece}
+  $\subseteq$ \emph{agreed piece} for some TX block.
+  We define
+  $$ \texttt{a-pieces}: B_i \rightarrow P(B_i) $$
+  as the function that returns the agreed pieces.
 
@@
 \subsection{Luck value}
-First we define the luck value $l_{i,j} = \texttt{h}(k_i || c_{i,j})$, where $k$ is the
-public key of $i$. A lower luck value equates to higher luck.
+First we define the luck value $l_{i,j} = \texttt{h}(k_i || c_{i,j})$, where $k$
+is the public key of $i$. A lower luck value equates to higher luck. We assume
+an application agnostic system and do not attempt to defend against the sybil
+attack.
+
+An alternative is to use \emph{proof of work}. This defends the sybil attack.
+But an incentive is needed for the nodes that expend their CPU resources.
 
@@
 
-\begin{lemma}
+\begin{lemma}\label{lem:con_consistent}
   If a node sees a valid $\mathcal{C}_{r-1}$ and another node sees a valid
@@
 
+\begin{lemma}\label{lem:con_alive}
+  Eventually all node sees a valid $\mathcal{C}_{r-1}.$  
+\end{lemma}
+\begin{proof}
+  (sketch) Liveness is satisfied because $\mathcal{C}_{r-1}$ is eventually propagated to
+  all node by gossiping.
+\end{proof}
+
 The potential promoters now need to first discover whether
@@
 
-\begin{lemma}
+\begin{lemma}\label{lem:promoter_consistent}
   The new set of promoter for the next consensus round is consistent with
@@
 Nodes should wait for some time to collect the CP blocks, so they wait for some
-time $\Delta$ before moving on to the next phase.
-
-\begin{lemma}
-  Promoters waiting for a some time $\Delta$ to collect transactions does not
-  violate the asynchronous assumption.
-\end{lemma}
-\begin{proof}
-  (sketch) This can be seen as a long delay in the asynchronous system.
-\end{proof}
+time $\Delta$ before moving on to the next phase. Note that promoters waiting
+for a some time $\Delta$ to collect transactions does not violate the
+asynchronous assumption because this behaviour can be seen as a long delay in
+the asynchronous system.
 
 \begin{corollary}
-  The setup phase satisfies the validity, agreement and liveness properties.
+  Setup phase satisfies agreement and total order because the protocol is run
+  deterministicly on the same input (\cref{lem:con_consistent},
+  \cref{lem:promoter_consistent}). It also satisfies liveness due to
+  \cref{lem:con_alive}.
 \end{corollary}
-\begin{proof}
-  (sketch) Validity and agreement is satisfied because the promoters are
-  computed using a deterministic algorithm using the latest consensus result.
-  Liveness is satisfied because $\mathcal{C}_{r-1}$ is eventually propagated to
-  all node by gossiping.
-\end{proof}
 
@@
 
-There techniques that uses no dealers. First is to use
+There are techniques that uses no dealers. First is to use
 PBFT~\cite{castro1999practical}, but we must change our asynchronous assumption
@@
 
+\begin{lemma}
+  Consensus phase satisfied agreement, total order and liveness.
+\end{lemma}
+\begin{proof}
+  Defer proof? Refer to the papers.
+\end{proof}
+
+\begin{theorem}
+  Checkpoint consensus satisfied agreement, total order and liveness.
+\end{theorem}
+\begin{proof}
+  (sketch) Both phases are asynchronous so we do not need to make assumptions on
+  when begin phase begins. Both phases also satisfy the agreement, total order
+  and liveness.
+\end{proof}
+
+\section{Validation}\label{sec:validation}
+
+Since we reach consensus on checkpoints rather than all the transactions, we
+need to detect fraud, such as forks. Further, we need to ensure the system is
+secure in a sense that valid transactions cannot be forged into invalid
+transactions once it has reached consensus, and vice versa.
+
+First, we define the requirements of a valid transaction $t_{i,j}$ as follows,
+much of it is derived from the TrustChain model.
+
+\begin{enumerate}
+\item There exist a pair $t_{i', j'}$ that satisfies the pair definition.
+\item $t_{i,j}$ and $t_{i', j'}$ is created using \texttt{newtx}, i.e. valid
+  signatures and hash pointers, etc..
+\item There exist an agreed piece that contains $t_{i, j}$ and another agreed
+  piece that contains $t_{i', j'}$. All the blocks in the agreed pieces have
+  valid hash pointers, the blocks in the pieces do not need to be valid.
+\item The CP blocks $c_{i, k}, c_{i', k'}$ that immediately follow $t_{i, j}$
+  and $t_{i', j'}$ are created using \texttt{newcp} using the same consensus
+  result $\mathcal{C}_r$ as input and are in the agreed pieces.
+\end{enumerate}
+
+Blocks that do not satisfy the definition above are not necessarily invalid. It
+may be the case that the validity cannot be determined due to incomplete
+information. For such cases we say its validity is unknown.
+Now we define an invalid transaction.
+
+\begin{enumerate}
+\item TODO
+\end{enumerate}
+
+Now we define the properties that are desired by the validation protocol.
+\begin{enumerate}
+\item \emph{Correctness}: The validation protocol outputs the correct result
+  according to the aforementioned requirements.
+\item \emph{Agreement}: If any correct node decides on the validity of a 
+  transaction, then all other correct nodes are able to reach the same
+  conclusion.
+\item \emph{Liveness}: Any valid transactions can be validated eventually.
+\item \emph{Unforgeability}: If some transaction is determined to be valid, then
+  it cannot be changed to an invalid transaction, the opposite also applies.
+\end{enumerate}
+
+Note that the input of the validation protocol is a TX block, so these
+properties hold with respect to TX blocks. In practice, if a node has a set of
+TX blocks that are in the unknown state, then it must run the validation
+protocol to determine whether they are valid. Further, these conditions do not
+imply all invalid transactions (forks) can be found globally, only the validity
+of the TX blocks that the honest nodes are interested in can be determined. The
+advantage of this scheme is that it saves a lot of computational and bandwidth
+costs because nodes only run the validation protocol on the TX blocks of their
+own interest.
+
+\subsection{Validation protocol}
+Assume node $u$ is aware of all the past consensus results $\mathcal{C}_r$.
+Suppose $u$ wish to validate $t_{i,j}$. It performs the following.
+
+\begin{enumerate}
+\item Determine the pair $t_{i', j'}$.
+\item Find the agreed enclosure for $t_{i,j}$ and $t_{i', j'}$ from
+  $\mathcal{C}_r$, otherwise return ``unknown''.
+%\item Find the smallest $c_{i', b}$ such that $b > j'$ is in consensus.
+\item Query $i$ and $i'$ for the agreed pieces and ensure hash pointers are
+  correct. Otherwise return ``unknown''.
+\item Check that $t_{i,j}$ and $t_{i', j'}$ are in the agreed pieces and are
+  created correctly using \texttt{newtx}. Otherwise return ``invalid''.
+\item Check the checkpoints $c_{i, k}$ and $c_{i', k'}$ that immediately follow
+  $t_{i,j}$ and $t_{i', j'}$ are in the agreed pieces and are created in the
+  same round, i.e. $\texttt{round}(c_{i, k}) = \texttt{round}(c_{i', k'})$.
+  Otherwise return ``invalid''.
+\item Return ``valid''.
+\end{enumerate}
+
+Most likely $u = i$ or $u = i'$, because they are incentivised to check the
+validity of their TX blocks that are of unknown validity.
+
+\subsection{Analysis}
+In this section we analyse the validation protocol with respect to the four
+properties in \cref{sec:validation}.
+
+\begin{lemma}
+  The validation protocol is correct.
+\end{lemma}
+\begin{proof}
+  Correctness follows directly from the protocol specification, namely it
+  directly implements the validation requirements.
+\end{proof}
+
+\begin{lemma}
+  The validation protocol satisfies the agreement property.
+\end{lemma}
+\begin{proof}
+  We proof by contradiction. Suppose two nodes $i$ and $j$ where $i \ne j$ runs
+  the protocol on the same TX block $t_{k, l}$, $i$ outputs ``valid'' and $j$
+  outputs ``invalid''. $i$ and $j$ would have both picked the same enclosure
+  $c_{k, a}$ and $c_{k, b}$, as these are determined from the consensus result.
+  Then $k$ must produce two agreed pieces that start with $c_{k,a}$ and end with
+  $c_{k, b}$ with valid hash pointers, but one satisfies the validation protocol
+  and the other one doesn't (e.g. missing $t_{k, l}$). Producing these two
+  pieces is equivalent to producing a collision. Due to the properties of
+  cryptographic hash functions this is not possible, thus a contradiction.
+\end{proof}
+
+\begin{lemma}
+  The validation protocol does not satisfy the liveness property.
+\end{lemma}
+\begin{proof}
+  (sketch) Nodes may be offline.
+\end{proof}
+
+There are many ways around the liveness. As long as the TX is validated once,
+then the agreed pieces can be gossiped.
+
+\begin{lemma}
+  A valid TX block cannot be made into an invalid TX block and vice versa.
+\end{lemma}
+\begin{proof}
+  We proof by contradiction. Suppose $t_{k, l}$ is determined to be valid using
+  the enclosure $c_{k, a}$ and $c_{k, b}$ and the agreed piece $p$. Then $t_{k,
+    l}$ is forged into an invalid transaction. To forge it into an invalid
+  transaction the attacker must either tamper with the data within the block or
+  remove $t_{k, l}$ from $p$. In other words create $p'$ where $p \ne p'$,
+  enclosed by $c_{k, a}$ and $c_{k, b}$, and has valid hash pointers. Producing
+  $p'$ is equivalent to finding a collision, thus a contradiction.
+
+  TODO proof ``vice versa''.
+\end{proof}
+
+\subsection{Time and message complexity}
+For a single transaction, the message complexity is linear with respect to the
+size of the agreed pieces for both part of the TX pair. Time complexity is
+constant since we only perfrom two queries (to $i$ and $i'$).
+
 \section{Fraud detection}
+Existence testing detects fraud when a single party of the TX is malicious. To
+detect fraud when both parties are malicious, every node performs existence
+testing on all TX blocks between CP blocks as if they are some other node. The
+assignment is determined using CP block digest where every node is assigned to
+the nearest node that has a higher digest, cycle back if there is no higher
+digest.
 
-\subsection{Threat model}
-\begin{itemize}
-\item What is the influence of a fork for future transactions?
-\item What if both parties (sender and receiver) are malicious?
-\end{itemize}
-
-\subsection{Random sampling}
+If the number of malicious parties is low, then it's more probable that a fork
+is detected. If where there are a large number of malicious parties, we repeat
+the random sampling process.
 

### 2017-03-08 chapters/ch2-checkpoint.tex
Write signature dissemination

@@
 
-\item A block of type $\gamma$ is a three-tuple, i.e.
-  $$c_{i,j} = (\texttt{h}(b_{i,j}), H(\mathcal{C}_r), p)$$
+\item A block of type $\gamma$ is a five-tuple, i.e.
+  $$c_{i,j} = (\texttt{h}(b_{i,j}), H(\mathcal{C}_r), r, p, s)$$
   where $\mathcal{C}_r$ is the consensus result in round $r$ and $p \in {0,1}$
   which indicates whether $i$ wish to become a promoter in the following
-  consensus round.
+  consensus round, finally $s$ is a signature of the block.
 
@@
 \item The result of a consensus in round $r$ is a set of CP blocks. Namely,
-  $$\mathcal{C}_r = \{ c_{i,j} : \text{agreed by the promoters}\}.$$
+  $$\mathcal{C}_r = (r, \{ c_{i,j} : \text{agreed by the promoters}\}).$$
 
@@
 
+The consensus result $\mathcal{C}_{r-1}$ does not contain all the signatures of
+the promoters. Thus, the signatures must also be disseminated by the promoters.
+This can be done in the same way as consensus result dissemination. Nodes must
+collect $f + 1$ signatures in order to trust $\mathcal{C}_{r-1}$ because there
+is at least one honest node in $f + 1$ nodes. If there is a sufficient number of
+signatures on the consensus result, we say it is \emph{valid}.
+
 \begin{lemma}\label{lem:con_consistent}
@@
 \begin{proof}
-  (sketch) $\mathcal{C}_{r-1}$ is signed by at least $N-f$ promoters from round
-  $r-1$.
+  $\mathcal{C}_{r-1}$ and $\mathcal{C}'_{r-1}$ are signed by at two groups of $f
+  + 1$ promoters (does not have to overlap) from round $r-1$. In both groups, at
+  least one node is honest. Thus the consensus result is what is actually
+  agreed.
 \end{proof}

### 2017-03-15 chapters/ch2-checkpoint.tex
Presentation background section

@@
 \item A block of type $\gamma$ is a five-tuple, i.e.
-  $$c_{i,j} = (\texttt{h}(b_{i,j}), H(\mathcal{C}_r), r, p, s)$$
+  $$c_{i,j} = (\texttt{h}(b_{i,j-1}), \texttt{h}(\mathcal{C}_r), r, p, s)$$
   where $\mathcal{C}_r$ is the consensus result in round $r$ and $p \in {0,1}$

### 2017-03-15 presentation.tex
Presentation background section

% Copyright 2004 by Till Tantau <tantau@users.sourceforge.net>.
%
% In principle, this file can be redistributed and/or modified under
% the terms of the GNU Public License, version 2.
%
% However, this file is supposed to be a template to be modified
% for your own needs. For this reason, if you use this file as a
% template and not specifically distribute it as part of a another
% package/program, I grant the extra permission to freely copy and
% modify this file as you see fit and even to delete this copyright
% notice.

\documentclass{beamer}

% There are many different themes available for Beamer. A comprehensive
% list with examples is given here:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
% You can uncomment the themes below if you would like to use a different
% one:
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{boxes}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
\usetheme{default}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
%\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

\usepackage{graphicx}

\title{Presentation Title}

% A subtitle is optional and this may be deleted
\subtitle{Optional Subtitle}

\author{K.~Cong}
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[Delft University of Technology] % (optional, but mostly needed)
{
  Faculty of Electrical Engineering, Mathematics and Computer Science\\
  Delft University of Technology}
% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.

% \date{Conference Name, 2013}
\date{2017}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

% \subject{Theoretical Computer Science}
% This is only inserted into the PDF information catalog. Can be left
% out.

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

% \pgfdeclareimage[height=0.5cm]{university-logo}{university-logo-filename}
% \logo{\pgfuseimage{university-logo}}

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
\AtBeginSubsection[]
{
  \begin{frame}<beamer>{Outline}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}

% Let's get started
\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
\end{frame}

% Section and subsections will appear in the presentation overview
% and table of contents.
\section{Background}

\subsection{TrustChain}

\begin{frame}{TrustChain}{}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{figures/trustchain-good}
  \centering
  \caption{\emph{TX block} is a six-tuple: $t_{i,j} = (\texttt{h}(b_{i,j-1}),
    h_s, h_r, s_s, s_r, m)$, one transaction results in two TX blocks---\emph{a
      pair}.}
  \end{figure}

\end{frame}

\begin{frame}{TrustChain}{}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{figures/trustchain-bad}
  \centering
  \caption{Fork is two correctly signed TX blocks that has the same $h_s$ but
    involve different receivers. Only one TX block may be in consensus.}
  \end{figure}

\end{frame}

\begin{frame}{TrustChain}
  \begin{itemize}
    \item Everyone has their own chain
    \item Transactions are on arbitrary data $m$
    \item Transactions make the chains intertwined
    \item No consensus (my thesis)
  \end{itemize}
\end{frame}

\section{My Thesis}
\subsection{TrustChain with Checkpoints}
\begin{frame}{TrustChain with Checkpoints}

  \begin{figure}[h]
  \includegraphics[width=0.7\textwidth]{figures/trustchain-bad-cp}
  \centering
  \caption{CP block is a five-tuple: $c_{i,j} = (\texttt{h}(b_{i,j-1}),
    \texttt{h}(\mathcal{C}_r), r, p, s)$, $\mathcal{C}_r$ is the consensus
    result at round $r$, $p =$ promoter indicator, $s =$ signature.}
  \end{figure}

\end{frame}

\subsection{Consensus Protocol Overview}
\begin{frame}{Consensus Protocol Overview}
  \begin{enumerate}
    \item $N$ lucky nodes are selected at random to act as promoters.
    \item Every node sends CP blocks to promoters.
    \item Promoters run a BFT (Byzantine Fault Tolerant) consensus algorithm to
      agree on a set of CP blocks---$\mathcal{C}_r$.
    \item Disseminate $\mathcal{C}_r$.
    \item Any node can verify that their transaction is in consensus.
    \item Repeat for next round.
  \end{enumerate}
\end{frame}

\subsection{Promoter Registration}
\begin{frame}{Promoter Registration}

\end{frame}

\begin{frame}{Blocks}
\begin{block}{Block Title}
You can also highlight sections of your presentation in a block, with it's own title
\end{block}
\begin{theorem}
There are separate environments for theorems, examples, definitions and proofs.
\end{theorem}
\begin{example}
Here is an example of an example block.
\end{example}
\end{frame}

% Placing a * after \section means it will not show in the
% outline or table of contents.
\section*{Summary}

\begin{frame}{Summary}
  \begin{itemize}
  \item
    The \alert{first main message} of your talk in one or two lines.
  \item
    The \alert{second main message} of your talk in one or two lines.
  \item
    Perhaps a \alert{third message}, but not more than that.
  \end{itemize}

  \begin{itemize}
  \item
    Outlook
    \begin{itemize}
    \item
      Something you haven't solved.
    \item
      Something else you haven't solved.
    \end{itemize}
  \end{itemize}
\end{frame}

% All of the following is optional and typically not needed. 
\appendix
\section<presentation>*{\appendixname}
\subsection<presentation>*{For Further Reading}

\begin{frame}[allowframebreaks]
  \frametitle<presentation>{For Further Reading}

  \begin{thebibliography}{10}

  \beamertemplatebookbibitems
  % Start with overview books.

  \bibitem{Author1990}
    A.~Author.
    \newblock {\em Handbook of Everything}.
    \newblock Some Press, 1990.

  \beamertemplatearticlebibitems
  % Followed by interesting articles. Keep the list short.

  \bibitem{Someone2000}
    S.~Someone.
    \newblock On this and that.
    \newblock {\em Journal of This and That}, 2(1):50--100,
    2000.
  \end{thebibliography}
\end{frame}

\end{document}

### 2017-03-15 presentation.tex
Presentation of current work

@@
     \item Transactions make the chains intertwined
+    \item Transactions are irrefutable due to hash pointers
     \item No consensus (my thesis)
@@
 
-\subsection{Consensus Protocol Overview}
-\begin{frame}{Consensus Protocol Overview}
+\subsection{Protocol Overview}
+\begin{frame}{Protocol Overview}
   \begin{enumerate}
-    \item $N$ lucky nodes are selected at random to act as promoters.
-    \item Every node sends CP blocks to promoters.
+    \item $n$ lucky nodes are selected at random to act as promoters.
     \item Promoters run a BFT (Byzantine Fault Tolerant) consensus algorithm to
-      agree on a set of CP blocks---$\mathcal{C}_r$.
-    \item Disseminate $\mathcal{C}_r$.
-    \item Any node can verify that their transaction is in consensus.
+      agree on a set of CP blocks.
+    \item Disseminate the consensus result (the CP blocks).
     \item Repeat for next round.
+    \item Any interested node can validate that their transaction.
   \end{enumerate}
@@
 
+  \begin{figure}[h]
+  \includegraphics[width=0.7\textwidth]{figures/trustchain-1}
+  \centering
+  \caption{We start in the state where $\mathcal{C}_{r - 1}$ has just been agreed but
+    not yet propagated.}
+  \end{figure}
+
 \end{frame}
 
-\begin{frame}{Blocks}
-\begin{block}{Block Title}
-You can also highlight sections of your presentation in a block, with it's own title
-\end{block}
-\begin{theorem}
-There are separate environments for theorems, examples, definitions and proofs.
-\end{theorem}
-\begin{example}
-Here is an example of an example block.
-\end{example}
+\begin{frame}{Promoter Registration}
+
+  \begin{figure}[h]
+  \includegraphics[width=0.7\textwidth]{figures/trustchain-2}
+  \centering
+  \caption{Nodes receive consensus result and set promoters indicator $p$, then
+    send the new CP blocks to promoters of round $r$.}
+  \end{figure}
+
 \end{frame}
 
-% Placing a * after \section means it will not show in the
-% outline or table of contents.
-\section*{Summary}
+\begin{frame}{Promoter Registration}
 
-\begin{frame}{Summary}
+  \begin{figure}[h]
+  \includegraphics[width=0.7\textwidth]{figures/trustchain-3}
+  \centering
+  \caption{Transactions carry on as usual in round $r$. Note that our CP blocks
+    (round $r - 1$) has not reached consensus yet.}
+  \end{figure}
+
+\end{frame}
+
+\begin{frame}{Promoter Registration}
+
+  \begin{figure}[h]
+  \includegraphics[width=0.7\textwidth]{figures/trustchain-4}
+  \centering
+  \caption{CP blocks at round $r-1$ should be in $\mathcal{C}_r$. If we're
+    lucky: $\texttt{h}(k_i || c_{i,j}) < T$, then we're responsible for
+    consensus of round $r + 1$.}
+  \end{figure}
+
+\end{frame}
+
+\subsection{Consensus}
+\begin{frame}{Consensus}
+  \begin{enumerate}
+    \item Nodes send CP blocks to the promoters.
+    \item The promoters' identities are encoded in the consensus result.
+    \item Promoters run the some asynchronous BFT consensus algorithm to agree on
+      a set of CP blocks---$\mathcal{C}_r$.
+    \item $n = 3t + 1$ is the optimal for BFT consensus.
+    \item $\mathcal{C}_r$ and the signatures are disseminated.
+    \item Nodes create new CP blocks when they receive $t + 1$ good signatures and
+      $\mathcal{C}_r$.
+  \end{enumerate}
+
+  \begin{theorem}
+    Assuming the promoters satisfy $n = 3t + 1$. The promoter registration and
+    the consensus protocol satisfied agreement, total order and liveness.
+  \end{theorem}
+\end{frame}
+
+\subsection{Validation}
+
+\begin{frame}{Validation}
+Assume node $u$ is aware of all the past consensus results $\mathcal{C}_r$.
+Suppose $u$ wish to validate $t_{i,j}$. It performs the following.
+
+\begin{enumerate}
+\item Determine the pair $t_{i', j'}$.
+\item Find the agreed enclosure for $t_{i,j}$ and $t_{i', j'}$ from
+  $\mathcal{C}_r$, otherwise return ``unknown''.
+%\item Find the smallest $c_{i', b}$ such that $b > j'$ is in consensus.
+\item Query $i$ and $i'$ for the agreed pieces and ensure hash pointers are
+  correct. Otherwise return ``unknown''.
+\item Check that $t_{i,j}$ and $t_{i', j'}$ are in the agreed pieces and are
+  created correctly using \texttt{newtx}. Otherwise return ``invalid''.
+\item Check the checkpoints $c_{i, k}$ and $c_{i', k'}$ that immediately follow
+  $t_{i,j}$ and $t_{i', j'}$ are in the agreed pieces and are created in the
+  same round, i.e. $\texttt{round}(c_{i, k}) = \texttt{round}(c_{i', k'})$.
+  Otherwise return ``invalid''.
+\item Return ``valid''.
+\end{enumerate}
+\end{frame}
+
+\begin{frame}{Validation}
+  In essence, given a TX, ask the receiver to proof that it has a set of
+  transactions that produces some CP block, the CP block should be in the
+  consensus result and the pair of the TX should be in that set.
+
+  \begin{theorem}
+    If at least one party of every transaction is honest, then forking and other
+    forms of tampering is guaranteed to be detected if the malicious party is
+    alive.
+  \end{theorem}
+\end{frame}
+
+\subsection{Implementation}
+\begin{frame}{Implementation}
   \begin{itemize}
-  \item
-    The \alert{first main message} of your talk in one or two lines.
-  \item
-    The \alert{second main message} of your talk in one or two lines.
-  \item
-    Perhaps a \alert{third message}, but not more than that.
-  \end{itemize}
-
-  \begin{itemize}
-  \item
-    Outlook
-    \begin{itemize}
-    \item
-      Something you haven't solved.
-    \item
-      Something else you haven't solved.
-    \end{itemize}
+    \item Currently on going---using Python and Twisted
+    \item Completed BFT consensus algorithm
+    \item Completed local TrustChain
+    \item Next step is networked TrustChain and validation protocol
   \end{itemize}
 \end{frame}
-
-% All of the following is optional and typically not needed. 
-\appendix
-\section<presentation>*{\appendixname}
-\subsection<presentation>*{For Further Reading}
-
-\begin{frame}[allowframebreaks]
-  \frametitle<presentation>{For Further Reading}
-
-  \begin{thebibliography}{10}
-
-  \beamertemplatebookbibitems
-  % Start with overview books.
-
-  \bibitem{Author1990}
-    A.~Author.
-    \newblock {\em Handbook of Everything}.
-    \newblock Some Press, 1990.
-
-  \beamertemplatearticlebibitems
-  % Followed by interesting articles. Keep the list short.
-
-  \bibitem{Someone2000}
-    S.~Someone.
-    \newblock On this and that.
-    \newblock {\em Journal of This and That}, 2(1):50--100,
-    2000.
-  \end{thebibliography}
-\end{frame}
-
 \end{document}

### 2017-03-16 presentation.tex
Interesting research question on ZK

@@
   \centering
-  \caption{\emph{TX block} is a six-tuple: $t_{i,j} = (\texttt{h}(b_{i,j-1}),
-    h_s, h_r, s_s, s_r, m)$, one transaction results in two TX blocks---\emph{a
-      pair}.}
+  \caption{Every node has a chain. \emph{TX block} is a six-tuple: $t_{i,j} =
+    (\texttt{h}(b_{i,j-1}), h_s, h_r, s_s, s_r, m)$, one transaction results in
+    two TX blocks---\emph{a pair}.}
   \end{figure}
@@
   \centering
-  \caption{Fork is two correctly signed TX blocks that has the same $h_s$ but
-    involve different receivers. Only one TX block may be in consensus.}
+  \caption{Fork is two correctly signed TX blocks that belong to the same sender
+    and has the same hash pointer but involve different receivers. Only one TX
+    block may be in consensus.}
   \end{figure}
@@
     \item Transactions are on arbitrary data $m$
-    \item Transactions make the chains intertwined
+    % \item Transactions make the chains intertwined
     \item Transactions are irrefutable due to hash pointers
@@
       a set of CP blocks---$\mathcal{C}_r$.
-    \item $n = 3t + 1$ is the optimal for BFT consensus.
+    \item $n \ge 3t + 1$ is the optimal for BFT consensus.
     \item $\mathcal{C}_r$ and the signatures are disseminated.
@@
   \begin{theorem}
-    Assuming the promoters satisfy $n = 3t + 1$. The promoter registration and
-    the consensus protocol satisfied agreement, total order and liveness.
+    Assuming the promoters satisfy $n \ge 3t + 1$. The promoter registration and
+    the consensus protocol satisfies agreement, total order and liveness.
   \end{theorem}
@@
   \end{theorem}
+
+  \begin{block}{Question}
+    Given a starting digest and an ending digest, is it possible for the prover
+    to proof to the verifier that it has a set of blocks linked by hash pointers
+    that ``fits'' inside those digests in \emph{zero knowledge}.
+  \end{block}
 \end{frame}
@@
     \item Completed BFT consensus algorithm
-    \item Completed local TrustChain
+    \item Completed local TrustChain with Checkpoints
     \item Next step is networked TrustChain and validation protocol

### 2017-04-12 chapters/ch2-checkpoint.tex
Prepare for tomorrow's presentation

--- 2017-03-08 chapters/ch2-checkpoint.tex
@@
 
-\item A block of type $\gamma$ is a five-tuple, i.e.
-  $$c_{i,j} = (\texttt{h}(b_{i,j}), H(\mathcal{C}_r), r, p, s)$$
+% TODO remove the p indication
+\item A block of type $\gamma$ is a six-tuple, i.e.
+  $$c_{i,j} = (\texttt{h}(b_{i,j-1}), \texttt{h}(\mathcal{C}_r), h, r, p, s)$$
   where $\mathcal{C}_r$ is the consensus result in round $r$ and $p \in {0,1}$
@@
 \begin{lemma}
-  Consensus phase satisfied agreement, total order and liveness.
+  Consensus phase satisfies agreement, total order and liveness.
 \end{lemma}
@@
 \begin{theorem}
-  Checkpoint consensus satisfied agreement, total order and liveness.
+  Checkpoint consensus satisfies agreement, total order and liveness.
 \end{theorem}

### 2017-04-12 presentation.tex
Prepare for tomorrow's presentation

@@
   \centering
-  \caption{CP block is a five-tuple: $c_{i,j} = (\texttt{h}(b_{i,j-1}),
-    \texttt{h}(\mathcal{C}_r), r, p, s)$, $\mathcal{C}_r$ is the consensus
+  \caption{CP block is a six-tuple: $c_{i,j} = (\texttt{h}(b_{i,j-1}),
+    \texttt{h}(\mathcal{C}_r), h, r, p, s)$, $\mathcal{C}_r$ is the consensus
     result at round $r$, $p =$ promoter indicator, $s =$ signature.}
@@
 
-\subsection{Implementation}
-\begin{frame}{Implementation}
+\subsection{Implementation and Experimentation Results}
+\begin{frame}{Implementation (4 weeks ago)}
   \begin{itemize}
@@
 \end{frame}
+
+\begin{frame}{Implementation}
+  \begin{itemize}
+    \item Completed BFT consensus algorithm, with erasure coding
+    \item Completed TrustChain with Checkpoints, with experiments on DAS5
+    \item Validation function is working
+    \item Next step is to build validation into a network protocol and do experiments
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Experiment with 500 nodes---no gossip}
+  \begin{figure}[h]
+  \includegraphics[width=0.95\textwidth]{figures/consensus-500-no-gossip}
+  \centering
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Experiment with 500 nodes---with gossip}
+  \begin{figure}[h]
+  \includegraphics[width=0.95\textwidth]{figures/consensus-500}
+  \centering
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Experiment with 500 nodes---with gossip}
+  \begin{block}{Conclusion}
+    \begin{itemize}
+      \item The number of facilitators affects the consensus latency ($O(N\log{N})$?).
+      \item The transaction rate has little or not effect on the average consensus latency.
+    \end{itemize}
+  \end{block}
+\end{frame}
+
+\begin{frame}{Planned experiments}
+  \begin{itemize}
+    \item Consensus latency (y-axis) vs population size (x-axis)---to
+      investigate how population size impacts consensus latency.
+    \item Global transaction rate vs population size---to investigate
+      the scalability property.
+    \item Global validation rate vs population size---to also investigate the
+      scalability property.
+  \end{itemize}
+\end{frame}
 \end{document}

### 2017-04-24 chapters/ch2-checkpoint.tex
Privacy preserving validation

@@
 
+\section{Privacy}
+Currently the validation protocol requires TX blocks to be sent in full. We can
+do better by modifying how the chain is computed to enhance user's privacy.
+First, recall that hash pointers are calculated on the whole block (TX or CP),
+i.e. $\texttt{h}(b_{s, h_s})$. In the validation stage, $b_{s, h_s}$ must be
+provided to recompuate the digest and check that it matches the block at $h_s - 1$.
+To improve this algorithm, we introduce a new property \emph{digest} for each
+block, which simply is $d_{s, h_s} = \texttt{h}(b_{s, h_s})$, $d$ is computed
+dynamically. Then the chain of hash pointers is formed using $\texttt{h}(d_{s,
+  h_s})$. During the validation phase, there is no need to provide the complete
+block, only the \emph{digest} is sufficient. Of course, the pair of the TX of
+interest must be provided in full.
+
 %%% Local Variables:

### 2017-05-15 thesis/chapters/ch1-intro.tex
Organise into directories

--- 2017-03-15 chapters/ch2-checkpoint.tex
@@
-\section{Preliminaries}
-
-\subsection{Requirements}
-\begin{itemize}
-\item Permissionless
-\item Byzantine fault tolerant
-\item No PoW
-\item Works under churn
-\item Underlying data structure is TrustChain
-\item Detects forks or double-spends
-\item No step in the protocol blocks transactions
-\item Application independent
-\end{itemize}
-
-\subsection{Assumptions}
-\begin{itemize}
-\item Asynchronous network
-\item Private and authenticated channel
-\item We elect $N$ consensus promoters in every round, we assume the number of
-  faulty promoters is $f$ and $N = 3f + 1$.
-\item Promoters have the complete history of the previously agreed set of
-  transactions (TX).
-% \item Single connected component for TrustChain.
-% \item No sybil attack % 
-\end{itemize}
-
-\subsection{Notation, definition and properties}
-\begin{itemize}
-
-\item $y = \texttt{h}(x)$ is a cryptographically secure hash function (random
-  oracle), the domain $x$ is infinite and the range is $y \in \{0, 2^{256}-1\}$.
-
-\item We model our system in the permissionless setting, where each
-  participating party has a unique identity $i$, and a chain $B_i$.
-
-\item A chain is a collection of blocks
-  $B_i = \{b_{i,j} : j \in \{1 \dots h\} \},$
-  the blocks are linked together using hash pointers, similar to bitcoin. All
-  blocks contain a reference to the previous block, the very first block with no
-  references is the genesis block.
-
-  The sequence number of the block begins at 0 on the genesis block and is
-  incremented for every new block. The height of the chain is $h = |B_i|$.
-
-\item There are two sets of blocks $T_i$ and $C_i$, where $T_i \cup
-  C_i = B_i$ and $T_i \cap C_i = \emptyset$. This can be seen as the
-  block type, where $t_{i,j}$ and $c_{i,j}$ to represent a
-  \emph{transaction block (TX block)} and \emph{checkpoint block (CP block)}
-  respectively.
-
-\item $\texttt{typeof}: B_i \rightarrow \{ \tau, \gamma \}$ returns the
-  corresponding type of the block.
-
-\item A block of type $\tau$ is a six-tuple, i.e. $$t_{i,j} =
-  (\texttt{h}(b_{i,j-1}), h_s, h_r, s_s, s_r, m).$$ $h_s$ and $h_r$ denote the
-  height (the sequence number for when the TX is made) of the sender and
-  receiver respectively. $s_s$ and $s_r$ are the signatures of the sender and
-  the receiver respectively. $i = \{s, r\}$ and $j = \{h_s, h_r\}$ depending on
-  whether $i$ is the sender or the receiver.
-
-\item Given two TX blocks $t_{i,j}$ and $t_{i',j'}$, if $i \ne i'$, $i = s$, $i'
-  = r$, $h_s = h'_s$, $h_r = h'_r$, $m = m'$ and the signatures are valid, then
-  we call them a \emph{pair}. Note that given one TX block, the pair can be
-  determined directly.
-
-\item If there exist two TX blocks $t_{i,j}$ and $t_{i',j'}$, where $s_s$ and
-  $s_s'$ is created by the same public key, $h_s = h_s'$, but $i \neq i'$, then
-  we call this a \emph{fork}.
-
-\item A block of type $\gamma$ is a five-tuple, i.e.
-  $$c_{i,j} = (\texttt{h}(b_{i,j-1}), \texttt{h}(\mathcal{C}_r), r, p, s)$$
-  where $\mathcal{C}_r$ is the consensus result in round $r$ and $p \in {0,1}$
-  which indicates whether $i$ wish to become a promoter in the following
-  consensus round, finally $s$ is a signature of the block.
-
-\item We define
-  $$\texttt{newtx}: B_i \times B_j \times M \rightarrow B_i \times B_j$$
-  as a function that creates new TX blocks. Its functionality is to
-  extend the given chains using the following rule. If the input is $(B_s, B_r,
-  m)$ then the output is $(B'_s, B'_r)$ where $B'_s = \{(\texttt{h}(b_{s,h_s}), h_s + 1,
-  h_r + 1, s_s, s_r, m)\} \cup B_s$, $B'_r = \{(\texttt{h}(b_{r, h_r}), h_s + 1, h_r + 1,
-  s_s, s_r, m)\} \cup B_r$, $h_s = |B_s|$ and $h_r = |B_r|$.
-
-\item We define
-  $$\texttt{newcp}: B_i \times \mathbb{R}_{\ge 1} \times \{0, 1\} \rightarrow B_i$$
-  as a function that creates new CP blocks. Concretely, given $(B_i, r, p)$, it
-  results in $\{ (\texttt{h}(b_{i,h}), H(\mathcal{C}_r), p) \}\cup B_i$ where $h =
-  |B_i|$, and $\mathcal{C}_r$ is the latest consensus result at round $r$.
-
-\item Note that in the actual system, \texttt{newtx} and \texttt{newcp} perform
-  a state transition.
-
-\item We define
-  $$\texttt{round}: C_i \rightarrow \mathbb{R}_{\ge 1}$$
-  as a function that gets the consensus round number used to create the given CP
-  block.
-
-\item The CP blocks that follows a pair of TX blocks must be created using the
-  same $\mathcal{C}_r$, otherwise the transaction is invalid. Concretely, given
-  a pair $t_{i,j}$ and $t_{i',j'}$, then there exist $c_{i, k}$ and $c_{i', k'}$
-  where $j < k$, $j' < k'$, $\{ \texttt{typeof}(b_{i, x}) : x \in \{j, \dots k -
-  1\}\}$ are all $\tau$, $ \{ \texttt{typeof}(b_{i', x}) : x \in \{ j', \dots k'
-  - 1\} \}$ are all $\tau$ and $\texttt{round}(c_{i, k}) = \texttt{round}(c_{i',
-    k'})$. This constraint is not the result of the consensus protocol, but the
-  validation protocol.
-
-\item The result of a consensus in round $r$ is a set of CP blocks. Namely,
-  $$\mathcal{C}_r = (r, \{ c_{i,j} : \text{agreed by the promoters}\}).$$
-
-\item A TX block can be valid, invalid or unknown. All TX blocks begin as
-  unknown, they can be validated using our validation protocol.
-
-\item A TX block has two \emph{ancestor} blocks. Let a pair be $t_{i,j}$ and
-  $t_{i',j'}$, then $(b_{i, j-1}, b_{i', j'-1})$ is the input block of $t_{i,j}$.
-
-\item A CP block has one \emph{ancestor} block, that is simply the block with
-  the previous sequence number, i.e. the ancestor of $c_{i,j}$ is $b_{i,j-1}$.
-
-\item Every TX block $t_{i,j}$ is enclosed by two CP blocks $(c_{i,a},
-  c_{i,b})$, where
-  $$a = \argmin_{k, k < j, \texttt{typeof}(b_{i,k}) = \gamma}(j - k),$$
-  $$b = \argmin_{k, k > j, \texttt{typeof}(b_{i,k}) = \gamma}(k - j).$$
-  We call this the \emph{enclosure} of $t_{i,j}$.
-
-\item The \emph{piece} of $t_{i,j}$ defined by the enclosure $(c_{i,a},
-  c_{i,b})$ is $\{b_{i,j} : a \le j \le b\}$.
-  We define
-  $$ \texttt{pieces}: B_i \rightarrow P(B_i) $$
-  as the function that returns the pieces ($P$ denotes power set).
-
-\item Every TX block $t_{i,j}$ is enclosed by two \emph{agreed} CP blocks
-  $(c_{i,a}, c_{i,b})$ (CP blocks that are in some consensus result), where
-  $$a = \argmin_{k, k < j, \texttt{typeof}(b_{i,k}) = \gamma}(j - k),$$
-  $$b = \argmin_{k, k > j, \texttt{typeof}(b_{i,k}) = \gamma}(k - j).$$
-  We call this the \emph{agreed enclosure} of $t_{i,j}$.
-
-\item The \emph{agreed piece} of $t_{i,j}$ defined by the agreed enclosure
-  $(c_{i,a}, c_{i,b})$ is $\{b_{i,j} : a \le j \le b\}$. Note that \emph{piece}
-  $\subseteq$ \emph{agreed piece} for some TX block.
-  We define
-  $$ \texttt{a-pieces}: B_i \rightarrow P(B_i) $$
-  as the function that returns the agreed pieces.
-
-\end{itemize}
-
-\section{Checkpoint consensus}
-
-\subsection{Luck value}
-First we define the luck value $l_{i,j} = \texttt{h}(k_i || c_{i,j})$, where $k$
-is the public key of $i$. A lower luck value equates to higher luck. We assume
-an application agnostic system and do not attempt to defend against the sybil
-attack.
-
-An alternative is to use \emph{proof of work}. This defends the sybil attack.
-But an incentive is needed for the nodes that expend their CPU resources.
-
-\subsection{Promoter registration}
-Node $i$ can register as a promoter when the latest consensus result is
-announced (suppose after the completion of round $r-1$), then it generates a new
-block $b_{i,j} = \texttt{newcp}(B_i, r-1, 1)$. The current promoters (in round
-$r$) may decide to include $b_{i,j}$ in the new consensus result. If $b$ is in
-it, then $i$ becomes one of the promoter of round $r+1$.
-
-We can fix the number of promoters to $N$ by sorting the promoters by their
-``luck value''' and taking the first $N$.
-
-\subsection{Promoter invitation}
-The output of promoter registration is a random set of $N$ promoters. If 1/3 of
-the population is malicious, then we cannot guarantee that the chosen promoters
-satisfy the $< n/3$ requirement.
-
-Promoter invitation is an attempt to involve human in the protocol. A naive
-method method is to use $N$ tickets, and then distribute them to trusted nodes.
-Nodes with the ticket can forward it to others. We have to rely on the humans to
-always forward the tickets to other honest humans. Finally, the nodes that hold
-a ticket are promoters. The result is that we have a permissioned system.
-
-\subsection{Setup phase}
-The setup phase should satisfy the BFT conditions regarding the promoter
-selection, that is:
-\begin{enumerate}
-\item \emph{Agreement}: If any correct node outputs a promoter $p$, then every
-  correct node outputs $p$.
-\item \emph{Total Order}: If one correct node outputs the sequence of promoters
-  $\{p_1, p_i, \dots, p_n\}$ and another has output $\{p'_0, p'_1, \dots,
-  p'_{n'} \}$, then $p_i = p'_i$ for $i \le \min(n, n')$.
-\item \emph{Liveness}: All $N - f$ correct nodes terminate eventually.
-\end{enumerate}
-
-We begin in the state where $\mathcal{C}_{r-1}$ has just been agreed but has not
-been disseminated yet. The exact technique to disseminate $\mathcal{C}_{r-1}$ is
-irrelevant, broadcasting or gossiping are both sufficient. In fact,
-dissemination is not necessary, nodes interested in the result can simply query
-the promoters that created $\mathcal{C}_{r-1}$.
-
-The consensus result $\mathcal{C}_{r-1}$ does not contain all the signatures of
-the promoters. Thus, the signatures must also be disseminated by the promoters.
-This can be done in the same way as consensus result dissemination. Nodes must
-collect $f + 1$ signatures in order to trust $\mathcal{C}_{r-1}$ because there
-is at least one honest node in $f + 1$ nodes. If there is a sufficient number of
-signatures on the consensus result, we say it is \emph{valid}.
-
-\begin{lemma}\label{lem:con_consistent}
-  If a node sees a valid $\mathcal{C}_{r-1}$ and another node sees a valid
-  $\mathcal{C}'_{r-1}$, then $\mathcal{C}_{r-1} = \mathcal{C}'_{r-1}$.
-\end{lemma}
-\begin{proof}
-  $\mathcal{C}_{r-1}$ and $\mathcal{C}'_{r-1}$ are signed by at two groups of $f
-  + 1$ promoters (does not have to overlap) from round $r-1$. In both groups, at
-  least one node is honest. Thus the consensus result is what is actually
-  agreed.
-\end{proof}
-
-\begin{lemma}\label{lem:con_alive}
-  Eventually all node sees a valid $\mathcal{C}_{r-1}.$  
-\end{lemma}
-\begin{proof}
-  (sketch) Liveness is satisfied because $\mathcal{C}_{r-1}$ is eventually propagated to
-  all node by gossiping.
-\end{proof}
-
-The potential promoters now need to first discover whether
-they are the first $N$ lucky promoters.
-
-\begin{lemma}\label{lem:promoter_consistent}
-  The new set of promoter for the next consensus round is consistent with
-  respect to all the nodes in the network.
-\end{lemma}
-\begin{proof}
-  (sketch) All nodes use the same deterministic function to compute the luck
-  value.
-\end{proof}
-
-Nodes should wait for some time to collect the CP blocks, so they wait for some
-time $\Delta$ before moving on to the next phase. Note that promoters waiting
-for a some time $\Delta$ to collect transactions does not violate the
-asynchronous assumption because this behaviour can be seen as a long delay in
-the asynchronous system.
-
-\begin{corollary}
-  Setup phase satisfies agreement and total order because the protocol is run
-  deterministicly on the same input (\cref{lem:con_consistent},
-  \cref{lem:promoter_consistent}). It also satisfies liveness due to
-  \cref{lem:con_alive}.
-\end{corollary}
-
-\subsection{Consensus phase}
-The consensus phase should satisfy the BFT conditions regarding the CP blocks,
-that is:
-\begin{enumerate}
-\item \emph{Agreement}: If any correct node outputs a CP block $c$, then every
-  correct node outputs $c$.
-\item \emph{Total Order}: If one correct node outputs the sequence of CP blocks
-  $\{c_1, c_i, \dots, c_n\}$ and another has output $\{c'_0, c'_1, \dots,
-  c'_{n'} \}$, then $c_i = c'_i$ for $i \le \min(n, n')$.
-\item \emph{Liveness}: All $N - f$ correct nodes terminate eventually.
-\end{enumerate}
-
-We need an atomic broadcast algorithm for the consensus phase. We use a similar
-but simplified construction as~\cite{miller2016honey}, where atomic broadcast is
-constructed from the reliable broadcast\footnote{Reliable broadcast solves the
-  Byzantine generals problem.}~\cite{bracha1984asynchronous} and asynchronous
-common subset (ACS). The ACS protocol requires a binary Byzantine agreement
-protocol, and for that it needs a trusted dealer to distributed the secret
-shares. Promoters can check whether the secret shares are valid, but they cannot
-prevent the dealer from disclosing the secrets.
-
-There are techniques that uses no dealers. First is to use
-PBFT~\cite{castro1999practical}, but we must change our asynchronous assumption
-into the weak synchrony assumption. Most likely this is not possible because of
-the ``When to start?'' problem. It's difficult to give a bounded delay for
-propagating the consensus result.
-
-Second is to use an inefficient binary Byzantine agreement protocol where its
-message complexity is $O(N^3)$ rather than $O(N^2)$ and becomes a bottleneck. Or
-a suboptimal one, e.g. $n/5$ instead of $n/3$.
-
-Suppose we use a dealer, what is the effect to the algorithm if the dealer is
-malicious?
-
-\begin{lemma}
-  Consensus phase satisfied agreement, total order and liveness.
-\end{lemma}
-\begin{proof}
-  Defer proof? Refer to the papers.
-\end{proof}
-
-\begin{theorem}
-  Checkpoint consensus satisfied agreement, total order and liveness.
-\end{theorem}
-\begin{proof}
-  (sketch) Both phases are asynchronous so we do not need to make assumptions on
-  when begin phase begins. Both phases also satisfy the agreement, total order
-  and liveness.
-\end{proof}
-
-\section{Validation}\label{sec:validation}
-
-Since we reach consensus on checkpoints rather than all the transactions, we
-need to detect fraud, such as forks. Further, we need to ensure the system is
-secure in a sense that valid transactions cannot be forged into invalid
-transactions once it has reached consensus, and vice versa.
-
-First, we define the requirements of a valid transaction $t_{i,j}$ as follows,
-much of it is derived from the TrustChain model.
-
-\begin{enumerate}
-\item There exist a pair $t_{i', j'}$ that satisfies the pair definition.
-\item $t_{i,j}$ and $t_{i', j'}$ is created using \texttt{newtx}, i.e. valid
-  signatures and hash pointers, etc..
-\item There exist an agreed piece that contains $t_{i, j}$ and another agreed
-  piece that contains $t_{i', j'}$. All the blocks in the agreed pieces have
-  valid hash pointers, the blocks in the pieces do not need to be valid.
-\item The CP blocks $c_{i, k}, c_{i', k'}$ that immediately follow $t_{i, j}$
-  and $t_{i', j'}$ are created using \texttt{newcp} using the same consensus
-  result $\mathcal{C}_r$ as input and are in the agreed pieces.
-\end{enumerate}
-
-Blocks that do not satisfy the definition above are not necessarily invalid. It
-may be the case that the validity cannot be determined due to incomplete
-information. For such cases we say its validity is unknown.
-Now we define an invalid transaction.
-
-\begin{enumerate}
-\item TODO
-\end{enumerate}
-
-Now we define the properties that are desired by the validation protocol.
-\begin{enumerate}
-\item \emph{Correctness}: The validation protocol outputs the correct result
-  according to the aforementioned requirements.
-\item \emph{Agreement}: If any correct node decides on the validity of a 
-  transaction, then all other correct nodes are able to reach the same
-  conclusion.
-\item \emph{Liveness}: Any valid transactions can be validated eventually.
-\item \emph{Unforgeability}: If some transaction is determined to be valid, then
-  it cannot be changed to an invalid transaction, the opposite also applies.
-\end{enumerate}
-
-Note that the input of the validation protocol is a TX block, so these
-properties hold with respect to TX blocks. In practice, if a node has a set of
-TX blocks that are in the unknown state, then it must run the validation
-protocol to determine whether they are valid. Further, these conditions do not
-imply all invalid transactions (forks) can be found globally, only the validity
-of the TX blocks that the honest nodes are interested in can be determined. The
-advantage of this scheme is that it saves a lot of computational and bandwidth
-costs because nodes only run the validation protocol on the TX blocks of their
-own interest.
-
-\subsection{Validation protocol}
-Assume node $u$ is aware of all the past consensus results $\mathcal{C}_r$.
-Suppose $u$ wish to validate $t_{i,j}$. It performs the following.
-
-\begin{enumerate}
-\item Determine the pair $t_{i', j'}$.
-\item Find the agreed enclosure for $t_{i,j}$ and $t_{i', j'}$ from
-  $\mathcal{C}_r$, otherwise return ``unknown''.
-%\item Find the smallest $c_{i', b}$ such that $b > j'$ is in consensus.
-\item Query $i$ and $i'$ for the agreed pieces and ensure hash pointers are
-  correct. Otherwise return ``unknown''.
-\item Check that $t_{i,j}$ and $t_{i', j'}$ are in the agreed pieces and are
-  created correctly using \texttt{newtx}. Otherwise return ``invalid''.
-\item Check the checkpoints $c_{i, k}$ and $c_{i', k'}$ that immediately follow
-  $t_{i,j}$ and $t_{i', j'}$ are in the agreed pieces and are created in the
-  same round, i.e. $\texttt{round}(c_{i, k}) = \texttt{round}(c_{i', k'})$.
-  Otherwise return ``invalid''.
-\item Return ``valid''.
-\end{enumerate}
-
-Most likely $u = i$ or $u = i'$, because they are incentivised to check the
-validity of their TX blocks that are of unknown validity.
-
-\subsection{Analysis}
-In this section we analyse the validation protocol with respect to the four
-properties in \cref{sec:validation}.
-
-\begin{lemma}
-  The validation protocol is correct.
-\end{lemma}
-\begin{proof}
-  Correctness follows directly from the protocol specification, namely it
-  directly implements the validation requirements.
-\end{proof}
-
-\begin{lemma}
-  The validation protocol satisfies the agreement property.
-\end{lemma}
-\begin{proof}
-  We proof by contradiction. Suppose two nodes $i$ and $j$ where $i \ne j$ runs
-  the protocol on the same TX block $t_{k, l}$, $i$ outputs ``valid'' and $j$
-  outputs ``invalid''. $i$ and $j$ would have both picked the same enclosure
-  $c_{k, a}$ and $c_{k, b}$, as these are determined from the consensus result.
-  Then $k$ must produce two agreed pieces that start with $c_{k,a}$ and end with
-  $c_{k, b}$ with valid hash pointers, but one satisfies the validation protocol
-  and the other one doesn't (e.g. missing $t_{k, l}$). Producing these two
-  pieces is equivalent to producing a collision. Due to the properties of
-  cryptographic hash functions this is not possible, thus a contradiction.
-\end{proof}
-
-\begin{lemma}
-  The validation protocol does not satisfy the liveness property.
-\end{lemma}
-\begin{proof}
-  (sketch) Nodes may be offline.
-\end{proof}
-
-There are many ways around the liveness. As long as the TX is validated once,
-then the agreed pieces can be gossiped.
-
-\begin{lemma}
-  A valid TX block cannot be made into an invalid TX block and vice versa.
-\end{lemma}
-\begin{proof}
-  We proof by contradiction. Suppose $t_{k, l}$ is determined to be valid using
-  the enclosure $c_{k, a}$ and $c_{k, b}$ and the agreed piece $p$. Then $t_{k,
-    l}$ is forged into an invalid transaction. To forge it into an invalid
-  transaction the attacker must either tamper with the data within the block or
-  remove $t_{k, l}$ from $p$. In other words create $p'$ where $p \ne p'$,
-  enclosed by $c_{k, a}$ and $c_{k, b}$, and has valid hash pointers. Producing
-  $p'$ is equivalent to finding a collision, thus a contradiction.
-
-  TODO proof ``vice versa''.
-\end{proof}
-
-\subsection{Time and message complexity}
-For a single transaction, the message complexity is linear with respect to the
-size of the agreed pieces for both part of the TX pair. Time complexity is
-constant since we only perfrom two queries (to $i$ and $i'$).
-
-\section{Fraud detection}
-Existence testing detects fraud when a single party of the TX is malicious. To
-detect fraud when both parties are malicious, every node performs existence
-testing on all TX blocks between CP blocks as if they are some other node. The
-assignment is determined using CP block digest where every node is assigned to
-the nearest node that has a higher digest, cycle back if there is no higher
-digest.
-
-If the number of malicious parties is low, then it's more probable that a fork
-is detected. If where there are a large number of malicious parties, we repeat
-the random sampling process.
+This is a story about a scalable blockchain framework and Byzantine
+generals\dots
 

### 2017-05-15 thesis/chapters/ch3-implementation.tex
Organise into directories

\section{Implementation}
Python. Twisted.

\section{Evaluation}
\begin{itemize}
  \item How fast is the consensus algorithm? Possibly plot graph of time versus
    the number of nodes.
  \item Does the promoter registration phase add a lot of extra overhead?
  \item What's the rate of transaction such that they can be verified ``on
    time'', i.e. without a growing backlog?
  \item Our global validation rate is somewhat equivalent to the transaction
    rate in other systems. Does the validation rate scale with respect to the
    number of nodes? In theory it should. Plot validation rate vs number of
    nodes, we expect it to be almost linear.
\end{itemize}

### 2017-05-28 thesis/chapters/ch1-intro.tex
Write some intro

@@
-This is a story about a scalable blockchain framework and Byzantine
-generals\dots
+TODO some positive intro about blockchain
+
+One of the key issues in many blockchain systems today is that they are not scalable.
+Bitcoin~\cite{bitcoin}, the largest permissionless\footnote{Explain permissionless} blockchain system
+in terms of market capitalisation~\cite{bitcoinmarketcap} has a maximum transaction rate of merely 7 transaction per second (TX/s).
+This is due to the consensus mechanism in Bitcoin, namely proof-of-work (PoW),
+miners can only create new blocks every 10 minutes and every block cannot be larger than 1 megabyte.
+Payment processors in use today such as Visa can handle transaction rates in the order of thousands~\cite{visa}.
+While Bitcoin may be a revolutionary phenomenon, it clearly cannot be ubiquitous in its current state.
+
+An different approach is to not reach global consensus at all.
+For instance in TrustChain~\cite{multichain} and Tangle~\cite{tangle},
+nodes in the network only store their personal ledger.
+Since consensus is left out, nodes can perform transactions as fast as their machine and network allows.
+The downside of this approach is that it cannot prevent fraud (it is possible to detect fraud).
+To examplify, a malicious node Mallory may claim she has 3 units of currency to Alice,
+but in reality Mallory already spent all of it on Bob.
+If there is no global consensus and Bob and Alice never communicate,
+then the 3 units that Alice is about to receive is nonexistent.
+
+The scalability property of TrustChain and Tangle are exceptionally desirable.
+The global consensus mechanism of Bitcoin and many other blockchain systems are also worthwhile for detecting or preventing fraud.
+These two properties may seem mutually exclusive, but in this work, we demonstrate the opposite.
+Specifically, we answer the following research question in the affirmative.
+\emph{Is it possible to design a blockchain fabric that can reach global consensus on the state of the system 
+and also scalable?} We define scalability as a property where if more nodes join the system, then the transaction rate should increases.
+% this is not completely true, there's a cap
+
+Our primary insight came from observing the differences between how transactional systems work in the real world and how they work in blockchain systems like Bitcoin. 
+Take a restaurant owner for example, most of the time the customer is honest and pays the bill.
+There is no need for the customer or the restaurant owner to report the transaction to any central authority 
+because both parties are happy with the transaction.
+On the other hand, if the customer leaves without paying the bill,
+then the restaurant owner would report the incident to some central authority, e.g. the police.
+On the contrary, in blockchain systems, every transaction is effectively sent to the miners,
+which can be seen as a collective authority.
+This consequently lead to limited scalability because every transaction must be validated by the authority even when most of the transaction are legitimate.
+
+Using the aforementioned insight,
+we explore an alternative consensus model for blockchain systems where transactions themself do not reach consensus,
+but nevertheless verifiable at a later stage by any node in the network.
+Informally, out model works as follows.
+Every node stores its own blockchain and every block is one transaction, same as the TrustChain construction.
+We randomly selected nodes in every round, the selected ones are called facilitators.
+They reach consensus not on the individual transactions,
+but on the state of every chain represented by a single digest, we call this state the checkpoint.
+If a checkpoint of some node is in consensus, 
+then that node can proof to any other node that it holds a set of transactions that computes (form a chain) to the checkpoint.
+This immediately show that those transactions are tamper-proof.
+
+We begin the detailed discussion by formulating the model \autoref{sec:model}. 
+Next, we analyse the correctness, security and performance of our design in \autoref{sec:analysis},
+this is where we present our theorems.
+Implementation and experimental results are discussed in \autoref{sec:implementation}.
+Finally, we compare our system with other state-of-the-art blockchain systems and conclude in Sections \ref{sec:related} and \ref{sec:conclusion} respectively.
 

### 2017-05-29 thesis/chapters/ch1-intro.tex
Write trustchain intro

@@
 
-We begin the detailed discussion by formulating the model \autoref{sec:model}. 
-Next, we analyse the correctness, security and performance of our design in \autoref{sec:analysis},
+We begin the detailed discussion by formulating the model \Cref{sec:model}. 
+Next, we analyse the correctness, security and performance of our design in \Cref{sec:analysis},
 this is where we present our theorems.
-Implementation and experimental results are discussed in \autoref{sec:implementation}.
+Implementation and experimental results are discussed in \Cref{sec:implementation}.
 Finally, we compare our system with other state-of-the-art blockchain systems and conclude in Sections \ref{sec:related} and \ref{sec:conclusion} respectively.

### 2017-05-29 thesis/chapters/ch2-model.tex
Write trustchain intro

What does our model capture?
What are the functionalities of our system?
What's necessary in a model?
Which formalism to use? (reason for the choice)

\section{Universally Composable Framework}

\section{Protocol Overview}
The protocol can be described in four parts, the extended TrustChain datastructure, 
the consensus protocol, the transaction protocol and finally the validation protocol.
We first describe how the protocols work individually and then explain how they fit together.

\subsection{Extended TrustChain}
Extended TrustChain naturally builds on top of TrustChain, thus we first describe the standard TrustChain.
Our description has minor differences compared to the description in~\cite{trustchain}.
This is to help with the description of the extended TrustChain.
However, the two descriptions are functionally the same.

\subsubsection*{Standard TrustChain}
In TrustChain, every node has their ``personal'' chain. 
Initially, the chain only contains a genesis block.
When a node wishes to add a new transaction (TX), a new TX block is generated and is appended to the chain.
A TX block must have a valid hash pointer pointing to the previous block
and a reference\footnote{TODO This is the difference?} to its \emph{pair}.
Suppose Alice made a transaction with Bob, then both parties must create a TX block to acknowledge that the transaction took place.
The pair of Alice's TX block is the corresponding TX on Bob's chain and vice versa.
An example of 3 nodes is shown in \Cref{fig:trustchain-bad}.
\begin{figure}
    \includegraphics[width=0.8\textwidth]{trustchain-bad}
    \centering
    \caption{Every block is denoted by $t_{i,j}$, where $i$ is the node ID and $j$ is the sequence number of the block.
    Thus we have three nodes and three corresponding chains in this example.
    The arrows represent hash pointers and the dotted lines represent references.
    The blocks at the ends of one dotted line are pairs of each other.
    The red block after $t_{b, 5}$ indicate a fork.}
    \label{fig:trustchain-bad}
\end{figure}

If every node follows the rules of TrustChain and we only consider hash pointers,
then the chain effectively forms a singly linked list.
However, if a node violates the rules, then a \emph{fork} may happen.
That is, there may be more than one TX block with a hash pointer pointing back to the same block.
In \Cref{fig:trustchain-bad}, node $b$ (in the middle chain) created two TX blocks that both point to $t_{b, 5}$.
If this is a ledger system it can be seen as a double spend, where the currency accumulated up until $t_{b, 5}$ are spent twice.

\subsubsection*{Extended TrustChain}
We are now ready to explain the extended TrustChain, which we abbreviate to ETC.
In ETC, we introduce a new type of block, namely checkpoint (CP) block.
In contract to TX blocks, CP blocks do not store transactions or contain references.
They capture the state of the chain and the state of the whole system.
In particular, the state of the chain is captured with a hash pointer.
The state of the whole system is captured in the content of the CP block,
namely as a digest of the latest \emph{consensus result} which we explain in \Cref{sec:overview-cons}.
A visual representation is shown in \Cref{fig:trustchain-bad-cp}.
\begin{figure}
    \includegraphics[width=0.8\textwidth]{trustchain-bad-cp}
    \centering
    \caption{TODO}
    \label{fig:trustchain-bad-cp}
\end{figure}

\subsection{Consensus Protocol}\label{sec:overview-cons}
Before describing our consensus protocol, we take a brief detour to explain Byzantine consensus,
which is a fundamental building block of our consensus protocol.

\subsubsection*{Byzantine consensus}
Byzantine consensus is also known as \emph{atomic broadcast}.
Roughly speaking, atomic broadcast need to satisfy the following properties.
\begin{enumerate}
\item TODO
\end{enumerate}
We stress that Byzantine consensus is not the same as Byzantine agreement or the Byzantine general's problem.
Byzantine agreement is TODO

The literature on Byzantine consensus and atomic broadcast is rich, some noteable ones include TODO.
Thus in the ETC consensus protocol, we assume there exist an "off-the-shelf" Byzantine consensus algorithm which we can use
(in our implementation we use HoneyBadgerBFT and motivate our choice in TODO).

\subsubsection{ETC consensus protocol} 
We explain the consensus protocol using the example in

\subsection{Transaction Protocol}

\subsection{Validation Protocol}

\subsection{Putting Everything Together}

\section{Extended TrustChain Protocol}

\section{Protocol Extensions}

### 2017-05-29 thesis/formalities/abstract.tex
Use TUD template

%De aankondiging bevat de spreker, titel, plaats, datum en tijd, samenstelling van de afstudeercommissie en een korte samenvatting (maximaal 25 regels).
\thispagestyle{empty}

\noindent \textbf{Author}\\
\begin{tabular}{l}
TODO AUTHOR\\
\\
\end{tabular}\\
\noindent \textbf{Title}\\
\begin{tabular}{l}
TODO TITLE\\
\\
\end{tabular}\\
\noindent \textbf{MSc presentation}\\
\begin{tabular}{l}
% <MM> DD, YYYY (like \today)
TODO GRADUATION DATE\\
\\
\end{tabular}

\vspace{1.1cm}

\noindent \textbf{Graduation Committee}\\
\begin{tabular}{ll}
TODO GRADUATION COMMITTEE & Delft University of Technology \\
% The order of listing the names: Graduation prof, supervisor(s), others ordered by title + alphabetical
%examples:
%prof. dr. ir. H. J. Sips (chair) & Delft University of Technology \\
%ir. dr. D. H. J. Epema           & Delft University of Technology \\
\end{tabular}

\begin{abstract} %de abstract bevat alleen een korte samenvatting van de inhoud van het onderzoek
TODO ABSTRACT
\end{abstract}

\clearpage

### 2017-05-29 thesis/formalities/preface.tex
Use TUD template

\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
TODO MOTIVATION FOR RESEARCH TOPIC

\vspace{1\baselineskip}

\noindent
TODO ACKNOWLEDGEMENTS

\vspace{1\baselineskip}

\noindent
TODO AUTHOR

\vspace{1\baselineskip}

\noindent
Delft, The Netherlands

\noindent
\today

### 2017-05-29 thesis/appendices/consensus-example.tex
First draft of protocol overview

\chapter{Consensus Example}
\label{app:consensus-example}

\begin{figure}[htb]
    \frame{\includegraphics[trim={0 0 0 11cm}, clip, width=1.0\textwidth]{trustchain-1}}
    \centering
    \caption{Initial state}
    \label{fig:trustchain-1}
\end{figure}

\begin{figure}[htb]
    \frame{\includegraphics[trim={0 0 0 8cm}, clip, width=1.0\textwidth]{trustchain-2}}
    \centering
    \caption{Initial state}
    \label{fig:trustchain-2}
\end{figure}

\begin{figure}
    \frame{\includegraphics[trim={0 0 0 2cm}, clip, width=1.0\textwidth]{trustchain-3}}
    \centering
    \caption{Initial state}
    \label{fig:trustchain-3}
\end{figure}

\begin{figure}
    \frame{\includegraphics[width=1.0\textwidth]{trustchain-4}}
    \centering
    \caption{Initial state}
    \label{fig:trustchain-4}
\end{figure}

### 2017-05-29 thesis/chapters/ch2-model.tex
Use TUD template
First draft of protocol overview

@@
+\chapter{System Model}
+\label{ch:model}
+
+TODO
 What does our model capture?
@@
 A TX block must have a valid hash pointer pointing to the previous block
-and a reference\footnote{TODO This is the difference?} to its \emph{pair}.
+and a reference\footnote{This is different from the original TrustChain definition found in~\cite{trustchain}.
+In there, a TX block has two outgoing edges which are hash pointers to the two parties involved in the transaction.
+This work uses one outgoing edge and a reference.} to its \emph{pair}.
 Suppose Alice made a transaction with Bob, then both parties must create a TX block to acknowledge that the transaction took place.
@@
 An example of 3 nodes is shown in \Cref{fig:trustchain-bad}.
+
 \begin{figure}
-    \includegraphics[width=0.8\textwidth]{trustchain-bad}
+    \includegraphics[width=0.9\textwidth]{trustchain-bad}
     \centering
@@
 A visual representation is shown in \Cref{fig:trustchain-bad-cp}.
+
 \begin{figure}
-    \includegraphics[width=0.8\textwidth]{trustchain-bad-cp}
+    \includegraphics[width=0.9\textwidth]{trustchain-bad-cp}
     \centering
-    \caption{TODO}
+    \caption{The circles represent CP blocks,
+    they also have hash pointers (arrow) but do not have references (dotted line).
+    Note that the sequence number counter do not change, it is shared with TX blocks.}
     \label{fig:trustchain-bad-cp}
@@
 \end{enumerate}
+TODO downsides (can't run with too many nodes) (high message complexity)
 We stress that Byzantine consensus is not the same as Byzantine agreement or the Byzantine general's problem.
@@
 
-\subsubsection{ETC consensus protocol} 
-We explain the consensus protocol using the example in
+In our case, every node do not propose some arbitrary data, but a set of CP blocks.
+Thus the result of the consensus is the set union of all the legitimate proposals.
+
+\subsubsection*{ETC consensus protocol} 
+The consensus protocol runs continuously in rounds.
+That is because a blockchain systems always need to reach consensus on new values, or CP blocks in our case.
+This can be seen as running infinitly many rounds of some Byzantine consensus algorithm,
+starting a new execution immediately after the previous one is completed.
+
+As we mentioned earlier, the high message complexity prohibits us from running a Byzantine consensus algorithm on a large network.
+Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks,
+and then use them as the proposal of the Byzantine consensus algorithm.
+The facilitators are elected using a \emph{luck value}, which is computed using $H(\Cons_r || pk_i)$,
+where $\Cons_r$ is the consensus result in round $r$ and $pk_i$ is the public key of $i$.
+Intuitively, the election is guaranteed to be random 
+because the output of a cryptographically secure hash function is unpredictable and $\Cons_r$ cannot be determined in advance.
+
+A visual explaination can be found in \Cref{app:consensus-example},
+it walks through the steps needed for a node to be selected as a facilitator.
 
 \subsection{Transaction Protocol}
+The TX protocol~\footnote{It is also known as True Halves, devised and implemented by Ewout Bongers.
+See \url{https://github.com/Tribler/tribler/pull/2135} for more information.}
+consist of two messages, \texttt{TX\_REQ} and \texttt{TX\_RESP}.
+When a node  wishes to create a transaction,
+it generates a TX block with the corresponding hash pointer, sequence number, counterparty and the transaction message itself.
+It then sends a \texttt{TX\_REQ} message containing the TX block to the counterparty.
+Upon receiving a TX block at the counterparty, it creates its own version of the TX block (the pair of the sender's TX block),
+and sends it back to the sender in a \texttt{TX\_RESP} message.
+At the end of the protocol, both parties should have both versions of the TX block.
 
 \subsection{Validation Protocol}
+The consensus and transaction protocol by themselves do not provide a mechanism to detect forks or other forms of tamperaing.
+This is the goal of the validation protocol.
 
-\subsection{Putting Everything Together}
+It also has two message---\texttt{VD\_REQ} and \texttt{VD\_RESP}.
+When a node wishes to validate a transaction,
+it sends a \texttt{VD\_REQ} with the sequence number $s$ to the party that created the TX block of the transaction.
+Upon receiving a \text{VD\_REQ}, the node finds two \emph{agreed} CP blocks that surrounds the TX block and computes the \emph{fragment}.
+An agreed CP block means that it has reached consensus.
+A fragment of two CP blocks is a section of the chain where the beginning and the end are defined by those two CP blocks.
+With the fragment, the node replies with a \texttt{VD\_RESP} message.
+Upon receiving a \texttt{VD\_RESP}, the node checks that the fragment is created correctly.
+Loosely, that means the agreed CP blocks have actually reached consensus,
+the fragment has the correct hash pointers,
+and TX block containing $s$ exists and is valid.
+
+\subsection{Putting The Protocols Together}
+The final protocol ($\Preal$) is essentially the concurrent composition of the three aforementioned protocols.
+Our subprotocol design gives us the highly desireable non-blocking property.
+In particular, we do not need to ``freeze'' the state of the chain for some communication to complete in order to create a block.
+For instance, a node may start the consensus protocol, and while it is running, the node may still perform transactions.
+By the time the consensus protocol is done, the new CP block is added to whatever the state that the chain is in.
+It is not necessary to keep the chain immutable while the consensus protocol is running.
 

### 2017-05-29 presentation/presentation.tex
Organise into directories
Add experimental results

--- 2017-04-12 presentation.tex
@@
 \usepackage{graphicx}
+\graphicspath{{../figures/}}
 
@@
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{figures/trustchain-good}
+  \includegraphics[width=0.7\textwidth]{trustchain-good}
   \centering
@@
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{figures/trustchain-bad}
+  \includegraphics[width=0.7\textwidth]{trustchain-bad}
   \centering
@@
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{figures/trustchain-bad-cp}
+  \includegraphics[width=0.7\textwidth]{trustchain-bad-cp}
   \centering
@@
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{figures/trustchain-1}
+  \includegraphics[width=0.7\textwidth]{trustchain-1}
   \centering
@@
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{figures/trustchain-2}
+  \includegraphics[width=0.7\textwidth]{trustchain-2}
   \centering
@@
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{figures/trustchain-3}
+  \includegraphics[width=0.7\textwidth]{trustchain-3}
   \centering
@@
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{figures/trustchain-4}
+  \includegraphics[width=0.7\textwidth]{trustchain-4}
   \centering
@@
 
-\begin{frame}{Experiment with 500 nodes---no gossip}
+\begin{frame}{Consensus duration vs population size (static neighbour)}
   \begin{figure}[h]
-  \includegraphics[width=0.95\textwidth]{figures/consensus-500-no-gossip}
+  \includegraphics[width=0.95\textwidth]{result-normal-cons-vs-pop}
   \centering
@@
 
-\begin{frame}{Experiment with 500 nodes---with gossip}
+\begin{frame}{Throughput vs population size (static neighbour)}
   \begin{figure}[h]
-  \includegraphics[width=0.95\textwidth]{figures/consensus-500}
+  \includegraphics[width=0.95\textwidth]{result-normal-throughput-vs-pop}
   \centering
@@
 
-\begin{frame}{Experiment with 500 nodes---with gossip}
-  \begin{block}{Conclusion}
-    \begin{itemize}
-      \item The number of facilitators affects the consensus latency ($O(N\log{N})$?).
-      \item The transaction rate has little or not effect on the average consensus latency.
-    \end{itemize}
-  \end{block}
+\begin{frame}{Consensus duration vs population size (random neighbour)}
+  \begin{figure}[h]
+  \includegraphics[width=0.95\textwidth]{result-rand-cons-vs-pop}
+  \centering
+  \end{figure}
 \end{frame}
 
-\begin{frame}{Planned experiments}
-  \begin{itemize}
-    \item Consensus latency (y-axis) vs population size (x-axis)---to
-      investigate how population size impacts consensus latency.
-    \item Global transaction rate vs population size---to investigate
-      the scalability property.
-    \item Global validation rate vs population size---to also investigate the
-      scalability property.
-  \end{itemize}
+\begin{frame}{Throughput vs population size (random neighbour)}
+  \begin{figure}[h]
+  \includegraphics[width=0.95\textwidth]{result-rand-throughput-vs-pop}
+  \centering
+  \end{figure}
 \end{frame}
+
 \end{document}

### 2017-05-30 thesis/chapters/ch2-model.tex
Write architecture intro and UC intro

@@
-\chapter{System Model}
+\chapter{System Architecture}
 \label{ch:model}
 
-TODO
-What does our model capture?
-What are the functionalities of our system?
-What's necessary in a model?
-Which formalism to use? (reason for the choice)
+The primary goal guiding our design is scalability.
+As mentioned in the Introduction, having a scalable blockchain system while still keeping global consensus
+allows the system to be ubiquitous and realise the full potential of blockchain.
 
-\section{Universally Composable Framework}
+Our secondary goal is to design an application neutral system.
+In particular, it should act as a framework that provides the building blocks of blockchain systems,
+namely transaction and consensus.
+Application developers using the framework should be able to create any blockchain based application they wish.
+Further, we do not impose on a consensus algorithm,
+the application developer is free to choose between proof-of-work, proof-of-stake,
+Byzantine consensus as long as it  satisfies the properties in \Cref{sec:todo}TODO.
+
+Due to the nature of our system, we do not explicitly address the Sybil attack~\cite{douceur2002sybil}.
+Sybil defence mechanism always require some form of reputation score from the application.
+For example, social network based Sybil defence mechanims use graph structure of real-world relationships~\cite{yu2006sybilguard}.
+Online marketplaces such as Amazon use the rating of buyer and sellers.
+Thus with a application neutral framework it is not possible to design a Sybil defence mechanism.
+On the other hand, our system has no restrictions on the Sybil defence technique.
+
+The third and final goal is security.
+Our system should be unaffected in the presence of powerful adversaries.
+Security is often difficult to verify, especially when it is not formalised, therefore we require our design to be provably secure.
+To summarise, our system design is designed with the following goals in mind.
+\begin{itemize}
+    \item Application neutrality,
+    \item scalability and
+    \item security.
+\end{itemize}
+
+We begin the chapter with an informal overview of our design in \Cref{sec:protocol-overview}.
+Next, we introduce the Universally Composable (UC) Framework and its explain its importance in \Cref{sec:uc-intro}.
+The UC framework is used in \Cref{sec:formal-model} to formally model our design.
+Finally, we give some possible extensions of our design in \Cref{sec:protocol-extensions}.
 
 \section{Protocol Overview}
+\label{sec:protocol-overview}
 The protocol can be described in four parts, the extended TrustChain datastructure, 
@@
 
-\section{Extended TrustChain Protocol}
+\section{Universally Composable Framework}
+\label{sec:uc-intro}
+
+In order to analyse the security of a system, a formal notion of security is required.
+For instance, what does it mean if an encryption algorithm is secure?
+One may say it is secure if the adversay cannot learn any information about the plaintext from the ciphertext.
+But what if the adversay has some background information, for instance she may know it is English.
+Do we then still say the encryption algorithm is secure?
+Goldwasser and Micali introduced the notion of semantic security~\cite{goldwasser1982probabilistic}.
+That is, imagine two worlds, a real world and an ideal world.
+In the real world, the adversay is given the ciphertext,
+and in the ideal world the adversay has nothing.
+Then the encryption algorithm is semantically secure if the amount of information that the adversay can learn in the real world is just as much as the ideal world.
+While our description is informal, the notion of security can be formally captured in this way.
+
+Security sensitive distributed systems such as secure multi-party computation and blockchain systems also require a formal notion of security so that they can be analysed.
+Fortunately, the idea from semantic security can also be applied in a distributed setting.
+In an ideal world, we create an ideal functionality $\Fideal$ that performs all the computation on behalf of every node in the network.
+The nodes act as dummies and only relay messages between $\Fideal$ and the environment.
+Thus $\Fideal$ essentially becomes the specification of the distributed protocol.
+If the execution of the real world protocol is indistinguishable from the ideal emulation in the presence of some adversay,
+then the real world protocol is secure as per the ideal specification.
+This is in essence the Universally Composable (UC) framework.
+
+We use the UC framework not only because it suits our needs, it is also the only framework used in modelling blockchain systems to the best of our knowledge
+~\cite{todo}TODO.
+In the remainder of this section we give an overview of the UC framework.
+The framework is formally described in~\cite{canetti2001universally}.
+
+TODO
+
+\section{Formal Specification}
+\label{sec:formal-model}
 
 \section{Protocol Extensions}
+\label{sec:protocol-extensions}

### 2017-06-02 thesis/chapters/ch2-model.tex
First protocol

@@
 
-Our secondary goal is to design an application neutral system.
-In particular, it should act as a framework that provides the building blocks of blockchain systems,
-namely transaction and consensus.
-Application developers using the framework should be able to create any blockchain based application they wish.
+The secondary goal is to design an application neutral system.
+In particular, it should act as a framework that provides the building blocks of blockchain based applications.
+Application developers using the framework should be able to create any application they wish.
 Further, we do not impose on a consensus algorithm,
@@
 Online marketplaces such as Amazon use the rating of buyer and sellers.
-Thus with a application neutral framework it is not possible to design a Sybil defence mechanism.
-On the other hand, our system has no restrictions on the Sybil defence technique.
+Thus it is not possible to design a Sybil defence mechanism with a a application neutral framework.
+On the other hand, our system also has no restrictions on the Sybil defence technique
+and application designers can pick the best mechanism for their application.
 
@@
 and then use them as the proposal of the Byzantine consensus algorithm.
-The facilitators are elected using a \emph{luck value}, which is computed using $H(\Cons_r || pk_i)$,
-where $\Cons_r$ is the consensus result in round $r$ and $pk_i$ is the public key of $i$.
+The facilitators are elected using a \emph{luck value}, which is computed using $H(\C_r || pk_i)$,
+where $\C_r$ is the consensus result in round $r$ and $pk_i$ is the public key of $i$.
 Intuitively, the election is guaranteed to be random 
-because the output of a cryptographically secure hash function is unpredictable and $\Cons_r$ cannot be determined in advance.
+because the output of a cryptographically secure hash function is unpredictable and $\C_r$ cannot be determined in advance.
 
@@
 
-We use the UC framework not only because it suits our needs, it is also the only framework used in modelling blockchain systems to the best of our knowledge
+We use the UC framework not only because it suits our needs, 
+it is also the only framework used in modelling blockchain systems to the best of our knowledge
 ~\cite{todo}TODO.
 In the remainder of this section we give an overview of the UC framework.
-The framework is formally described in~\cite{canetti2001universally}.
+A detailed treatment  can be found in~\cite{canetti2001universally}.
 
-TODO
+\subsection{Model of Computation}
+The model of computation considered in the UC framework is the interactive Turing machine (ITM).
+Specifically, an ITM is an extension of the Turing machine with externally writable tapes which are the following.
+\begin{enumerate}
+\item input tape---TODO
+\item incoming communication tape---TODO
+\item subroutine output tape---TODO
+\end{enumerate}
+ITM can be seen as a specification or an algorithm, a machine running an ITM is an ITM instance (ITI).
+To communicate, an ITI can write on the externally writable tapes of other ITIs.
+The writing ITI then pauses exeuction, the receiving ITI begins execution.
+Consequently, only one ITI is running at any point in time.
+
+\subsection{Simulation-based Security}
+Simulation-based security, also known as ideal/real paradigm is a model for defining security.
+The model consist of a set of , the environment $\E$, the adversay $\A$,
+the protocol ITM $\Preal$ and zero or more ideal functionalities $\mathcal{F}_0, \mathcal{F}_1, \dots$.
+The 
+Other than the machines running the protocol under consideration, the model contains two extra entities,
+the environment $\E$ and the adversary $\A$.
+$\E$ can be seen as users of the protocol, it can only provide input and receive output from $\A$ and the protocol.
+
+Control function TODO
+
+\subsection{Universal Composability}
 
@@
 
+The formal model follows the same structure as \ref{fig:todo}TODO.
+
+\begin{framed}
+\textbf{Protocol}\,$\Preal$
+
+The ETC protocol. On initialisation do the following.
+\begin{itemize}
+\item Generate public and private key pair, $pk_i$ and $sk_i$.
+\item Set personal chain $C := \{genesis\}$.
+\item Set checkpoint buffer $\C' := \varnothing$.
+\item Set the facilitators $F$ to the bootstrap nodes provided by $\E$.
+\end{itemize}
+
+Run the protocol as specified below after initialisation.
+\begin{itemize}
+
+\item Upon (\texttt{tx\_init}, $m$, $p$) from $\E$, 
+    $C := C \cup \{new\_tx(m)\}$,
+    send (\texttt{tx\_req}, latext TX in $C$) to $p$.
+\item Upon (\texttt{tx\_req}, $m$) from $p \in \P$,
+    $C := C \cup \{new\_tx(m)\}$,
+    let $m'$ be the latest TX in $C$ and set $m$ to be the other half of $m'$,
+    send (\texttt{tx\_resp}, $m'$) to $p$.
+    Output the complete TX to $\E$.
+\item Upon (\texttt{tx\_resp}, $m'$) from $p \in \P$,
+    find the corresponding pair of $m'$ and name it $m$,
+    add $m'$ as the other half of $m$.
+    Output the complete TX to $\E$.
+
+\item Upon (\texttt{vd\_init}, $s$) from $\E$,
+    if the sequence number $s$ does not exist in $C$ or the block with $s$ does not have the other half then do nothing.
+    Otherwise send (\texttt{vd\_req}, $s'$) to $p'$ where $s'$ is the sequence number of the other half and $p'$ is the counterparty.
+\item Upon (\texttt{vd\_req}, $s'$) from $p \in \P$,
+    if the sequence number $s'$ does not exist then do nothing.
+    Otherwise send (\texttt{vd\_resp}, $agreed\_fragment(s')$) to $p$.
+\item Upon (\texttt{vd\_resp}, $x$) from $p \in \P$,
+    run $validate(x)$ and output result to $\E$.
+
+\item Upon (\texttt{consensus}, $r$, $D$) from $\F_\text{BFT}$,
+    TODO,
+    send (\texttt{propose})
+
+\item Upon (\texttt{checkpoint}, $c$) from $p \in \P$,
+
+\end{itemize}
+
+\end{framed}
+
+\begin{framed}
+\textbf{Functionality}\,$\F_\text{BFT}$
+
+The ideal Byzantine consensus protocol, parameterised by...
+
+\begin{itemize}
+\item Upon (\texttt{propose}, $r$, $C$),
+\item Upon (\texttt{fetch}),
+\end{itemize}
+\end{framed}
+
+\subsection{Discussion}
+
+\subsubsection{Global clock for synchronisation}
+Where there is no Byzantine corruption, we conjecture that our protocol runs in the asynchronous setting.
+However, security of many Byzantine consensus algorithms, especially the one we adopted---HoneyBadgerBFT,
+fall apart when there is dynamic corruption.
+In order to enforce that the corrupted machines do not change when running an instance of some Byzantine consensus algorithm,
+we introduce synchrony.
+
+We make use of a global clock... TODO
+
+Dynamic corruption only between different clock ticks is enforced by the control function.
+
+\subsubsection{Static corruption versus dynamic corruption}
+A subtlety exists when modelling the type of corruption.
+In dynamic corruption\footnote{In \cite{canetti2001universally}, dynamic corruption is termed Byzantine corruption.},
+$\A$ take full control of a number of machines and also learns all of their states.
+The states include private keys.
+Thus, using the dynamic corruption model from \cite{canetti2001universally} we cannot guarantee security as
+$\A$ can corrupt different machines over time and eventually learn all the private keys..
+
+First is to use static corruption, where the corrupted machines are fixed.
+While this circumvents our problem, it is a much weaker model.
+Alternatively, we modify the aforementioned dynamic corruption model and weaken the adversary's abilities.
+In particular, we introduce that forgetful adversary that only remembers the state of the currently corrupted nodes and forget the state of the recovered nodes.
+
 \section{Protocol Extensions}

### 2017-06-06 thesis/chapters/ch2-model.tex
Writing

--- 2017-05-29 thesis/chapters/ch2-model.tex
@@
-\chapter{System Model}
+\chapter{System Architecture}
 \label{ch:model}
 
-TODO
-What does our model capture?
-What are the functionalities of our system?
-What's necessary in a model?
-Which formalism to use? (reason for the choice)
+The primary goal guiding our design is scalability.
+As mentioned in the Introduction, having a scalable blockchain system while still keeping global consensus
+allows the system to be ubiquitous and realise the full potential of blockchain.
 
-\section{Universally Composable Framework}
+The secondary goal is to design an application neutral system.
+In particular, it should act as a framework that provides the building blocks of blockchain based applications.
+Application developers using the framework should be able to create any application they wish.
+Further, we do not impose on a consensus algorithm,
+the application developer is free to choose between proof-of-work, proof-of-stake,
+Byzantine consensus as long as it  satisfies the properties in \Cref{sec:todo}TODO.
 
-\section{Protocol Overview}
-The protocol can be described in four parts, the extended TrustChain datastructure, 
-the consensus protocol, the transaction protocol and finally the validation protocol.
-We first describe how the protocols work individually and then explain how they fit together.
+Due to the nature of our system, we do not explicitly address the Sybil attack~\cite{douceur2002sybil}.
+Sybil defence mechanism always require some form of reputation score from the application.
+For example, social network based Sybil defence mechanims use graph structure of real-world relationships~\cite{yu2006sybilguard}.
+Online marketplaces such as Amazon use the rating of buyer and sellers.
+Thus it is not possible to design a Sybil defence mechanism with a a application neutral framework.
+On the other hand, our system also has no restrictions on the Sybil defence technique
+and application designers can pick the best mechanism for their application.
+
+The third and final goal is security.
+Our system should be unaffected in the presence of powerful adversaries.
+Security is often difficult to verify, especially when it is not formalised, therefore we require our design to be provably secure.
+To summarise, our system design is designed with the following goals in mind.
+\begin{itemize}
+    \item Application neutrality,
+    \item scalability and
+    \item security.
+\end{itemize}
+
+We begin the chapter with an intuitive overview of the architecture in \Cref{sec:system-overview}.
+Next, we give the formal description, starting with the model and assumptions in \Cref{sec:model-assumptions}.
+Then, the three protocols which make up the complete system,
+namely consensus protocol (\Cref{sec:cons-protocol}), transaction protocol (\Cref{sec:tx-protocol}) and validation protocol (\Cref{sec:vd-protocol}).
+Finally, the possible extensions are described in \Cref{sec:protocol-extensions}.
+
+\section{System Overview}
+\label{sec:system-overview}
+The system consist of one data structure---Extended TrustChain,
+and three protocols---consensus protocol, transaction protocol and validation protocol.
+We first describe each component individually and then explain how they fit together in \Cref{sec:combined-protocol}.
 
@@
 \subsubsection*{Standard TrustChain}
-In TrustChain, every node has their ``personal'' chain. 
+In TrustChain, every node has a ``personal'' chain. 
 Initially, the chain only contains a genesis block.
@@
 This work uses one outgoing edge and a reference.} to its \emph{pair}.
-Suppose Alice made a transaction with Bob, then both parties must create a TX block to acknowledge that the transaction took place.
-The pair of Alice's TX block is the corresponding TX on Bob's chain and vice versa.
-An example of 3 nodes is shown in \Cref{fig:trustchain-bad}.
+As a result, a single transaction generates two TX blocks, one on each party's chain.
+An example of is shown in \Cref{fig:trustchain-bad}.
 
@@
 \subsubsection*{Extended TrustChain}
-We are now ready to explain the extended TrustChain, which we abbreviate to ETC.
-In ETC, we introduce a new type of block, namely checkpoint (CP) block.
+We are now ready to explain the Extended TrustChain, which we abbreviate to ETC.
+In ETC, we introduce a new type of block---checkpoint (CP) block.
 In contract to TX blocks, CP blocks do not store transactions or contain references.
-They capture the state of the chain and the state of the whole system.
+Their purpose is to capture the state of the chain and the state of the whole system.
 In particular, the state of the chain is captured with a hash pointer.
@@
 
-\subsubsection*{ETC consensus protocol} 
+\subsubsection*{Consensus Protocol} 
 The consensus protocol runs continuously in rounds.
@@
 Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks,
-and then use them as the proposal of the Byzantine consensus algorithm.
-The facilitators are elected using a \emph{luck value}, which is computed using $H(\Cons_r || pk_i)$,
-where $\Cons_r$ is the consensus result in round $r$ and $pk_i$ is the public key of $i$.
+and then use the CP blocks as proposals of the Byzantine consensus algorithm.
+The facilitators are elected using a \emph{luck value}, which is computed using $H(\C_r || pk_i)$,
+where $\C_r$ is the consensus result in round $r$ and $pk_i$ is the public key of $i$.
 Intuitively, the election is guaranteed to be random 
-because the output of a cryptographically secure hash function is unpredictable and $\Cons_r$ cannot be determined in advance.
+because the output of a cryptographically secure hash function is unpredictable and $\C_r$ cannot be determined in advance.
 
@@
 \subsection{Transaction Protocol}
-The TX protocol~\footnote{It is also known as True Halves, devised and implemented by Ewout Bongers.
+The TX protocol~\footnote{It is also known as True Halves.
 See \url{https://github.com/Tribler/tribler/pull/2135} for more information.}
-consist of two messages, \texttt{TX\_REQ} and \texttt{TX\_RESP}.
+consist of two messages, \texttt{tx\_req} and \texttt{tx\_resp}.
 When a node  wishes to create a transaction,
 it generates a TX block with the corresponding hash pointer, sequence number, counterparty and the transaction message itself.
-It then sends a \texttt{TX\_REQ} message containing the TX block to the counterparty.
+It then sends a \texttt{tx\_req} message containing the TX block to the counterparty.
 Upon receiving a TX block at the counterparty, it creates its own version of the TX block (the pair of the sender's TX block),
-and sends it back to the sender in a \texttt{TX\_RESP} message.
+and sends it back to the sender in a \texttt{tx\_resp} message.
 At the end of the protocol, both parties should have both versions of the TX block.
@@
 
-It also has two message---\texttt{VD\_REQ} and \texttt{VD\_RESP}.
+It also has two message---\texttt{vd\_req} and \texttt{vd\_resp}.
 When a node wishes to validate a transaction,
-it sends a \texttt{VD\_REQ} with the sequence number $s$ to the party that created the TX block of the transaction.
-Upon receiving a \text{VD\_REQ}, the node finds two \emph{agreed} CP blocks that surrounds the TX block and computes the \emph{fragment}.
+it sends a \texttt{vd\_req} with the sequence number $s$ to the party that created the TX block of the transaction.
+Upon receiving a \text{vd\_req}, the node finds two \emph{agreed} CP blocks that surrounds the TX block and computes the \emph{fragment}.
 An agreed CP block means that it has reached consensus.
 A fragment of two CP blocks is a section of the chain where the beginning and the end are defined by those two CP blocks.
-With the fragment, the node replies with a \texttt{VD\_RESP} message.
-Upon receiving a \texttt{VD\_RESP}, the node checks that the fragment is created correctly.
+With the fragment, the node replies with a \texttt{vd\_resp} message.
+Upon receiving a \texttt{vd\_resp}, the node checks that the fragment is created correctly.
 Loosely, that means the agreed CP blocks have actually reached consensus,
@@
 
-\subsection{Putting The Protocols Together}
-The final protocol ($\Preal$) is essentially the concurrent composition of the three aforementioned protocols.
+\subsection{Combined Protocol}
+\label{sec:combined-protocol}
+The final protocol is essentially the concurrent composition of the three aforementioned protocols,
+all making use the Extended TrustChain data structure.
+
 Our subprotocol design gives us the highly desireable non-blocking property.
@@
 
-\section{Extended TrustChain Protocol}
+\section{Model and Assumptions}
+\label{sec:model-assumptions}
+
+For notational clarify, we use the following convention (adapted from~\cite{miller2016honey}) throughout this work.
+\begin{itemize}
+\item Lower case (e.g. $x$) denotes a scalar object.
+\item Upper case (e.g. $X$) denotes a set or a constant.
+\item Sans serif (e.g. $\textsf{fn}(\cdot)$) denotes a function.
+\item Typewrite (e.g. $\texttt{ack}$) denotes message type.
+\end{itemize}
+
+We assume a distributed network where nodes are fully connected,
+the channels are reliable\footnote{
+    Reliability can be achieved in unreliable networks by resending messages or using some error correction code.
+},
+but messages may be re-ordered and delayed by at most some time $\Delta$.
+Nodes are identified by their public key thus we assume there exist a Public Key Infrastructure (PKI).
+TODO this is weak synchrony?
+In addition, we assume the existance of cryptographically secure hash function $\textsf{H}(\cdot)$,
+that has preimage resistance, second preimage resistance and collision resistance.
+
+In our model we consider $N$ nodes, which is the population size.
+$n$ of them are facilitators, $t$ out of $n$ are malicious and the inequality
+$n \ge 3t + 1$ must hold.
+
+The primary data structure used in our system is Extended TrustChain.
+Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
+The chain consist of blocks $B_u = \{ b_{u, i} : k \in \{ 0, \dots, h - 1 \} \}$,
+where $b_{u, i}$ is the $i$th block of $u$,
+and $h = |B_u|$.
+There are two types of blocks, TX blocks and CP blocks.
+If $T_u$ is the set of TX blocks of $u$ and $C_u$ is the set of CP blocks of $u$,
+then it must be the case that $T_u \cup C_u = B_u$ and $T_u \cap C_u = \varnothing$.
+The notation $b_{u, i}$ is generic over the block type.
+We assume there exist a function $\textsf{typeof}: B_u \rightarrow \{ \tau, \gamma \}$ that returns the type of the block,
+where $\tau$ represents the TX type and $\gamma$ represents the CP type.
+
+\subsection{TX Block}
+The TX block is a six-tuple, i.e $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$.
+We describe each item in turn.
+\begin{enumerate}
+\item $\textsf{H}(b_{u, i - 1})$ is the hash pointer to the previous block.
+\item $seq_u$ is the sequence which should equal $i$.
+\item $txid$ is a cryptographically secure random number representing the transaction identifier.
+\item $pk_v$ is the public key of the counterparty.
+\item $m$ is the transaction message.
+\item $sig_u$ is the signature created using $sk_u$ on the concatination of the binary representation of the five items above.
+\end{enumerate}
+The fact that we have no constraint on the content of $m$ is in alignment with our design goal---application neutrality.
+
+TX blocks come in pairs.
+In particular, for every block $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$
+there exist one and only one \emph{pair} $t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), seq_v, txid, pk_u, m, sig_v \rangle$.
+Note that the $txid$ and $m$ are the same, and the public keys refer to each other.
+Thus, given a TX block, these properties allow us to identify its pair.
+
+% TODO Node $v$ may cheat and create more than one pair for $t_{u, i}. we discuss later
+
+\subsection{CP Block}
+
+The CP block is a five-tuple, 
+i.e. $c_{u, i} = \langle \textsf{H}(b_{u, i-1}), seq_u, \textsf{H}(\C_r), r, sig_u \rangle$,
+where $\C_r$ is the consensus result in round $r$, the other items are the same as the TX block definition.
+Note that unlike in our prior work~\cite{implicitconsensus}, CP blocks and TX blocks do not have independent sequence numbers.
+
+The genesis block in the chain must be a CP block in the form of
+$c_{u, 0} = \langle \textsf{H}(\bot), 0,  \textsf{H}(\bot), 0, sig_u \rangle$
+where $\textsf{H}(\bot)$ can be interpreted as applying the hash function on an empty string.
+The genesis block is unique due to every node due to $sig_u$.
+
+\subsection{Consensus Result}
+Our consensus protocol runs in rounds as discussed in \Cref{sec:system-overview}.
+Every round is identified by a round number $r$, which is incremented on every new round.
+The consensus result is a tuple, i.e. $\C_r = \langle r, C \rangle$,
+where $C$ is a set of CP blocks agreed by the facilitators of round $r$.
+
+\subsection{Chain Properties}
+Here we define a few important properties which results from the interleaving nature of CP and TX blocks.
+
+If there exist a tuple $\langle c_{u,a}, c_{u, b} \rangle$ for a TX block $t_{u, i}$,
+where 
+$$a = \argmin_{k, k < i, \texttt{typeof}(b_{u,k}) = \gamma}(i - k)$$
+$$b = \argmin_{k, k > i, \texttt{typeof}(b_{u,k}) = \gamma}(k - i),$$
+then $\langle c_{u,a}, c_{u, b} \rangle$ is the \emph{enclosure} of $t_{u, i}$.
+Some TX blocks may not have any subsequent CP blocks, then its enclosure is $\bot$.
+
+If the enclosure of some TX block is $\langle c_{u,a}, c_{u, b} \rangle$,
+then its \emph{fragment} is computed as $\{ b_{u, i} : a \le i \le b \}$.
+For convenience, the function $\textsf{fragment}(\cdot)$ represents the fragment of some TX block if it exists, otherwise $\bot$.
+
+\emph{Agreed enclosure} is the same as enclosure with an extra constraint where the CP blocks must be in some consensus result $\C_r$.
+Similarly, \emph{agreed fragment} is computed using agreed enclosure.
+We define its function to be $\textsf{agreed\_fragment}(\cdot)$
+
+The length of the fragment is constrained by $L$,
+namely $\forall t |\textsf{fragment}(t)| \le L$.
+The purpose to prevent spam and encourage nodes to create more CP blocks.
+$L$ should be sufficiently high so that busy nodes are not hindered by it.
+
+\section{Consensus Protocol}
+\label{sec:cons-protocol}
+
+Our consensus protocol runs on top of the model describe before, it makes heavy use of CP blocks.
+We describe the protocol step by step.
+Starting with the bootstrap phase and then moving on to the actual consensus phase.
+
+\subsection{Bootstrap Phase}
+Recall that facilitators are computed from the consensus result,
+but the consensus result is agreed by the facilitators.
+Thus we have a dependency cycle.
+The goal of the bootstrap phase is to give us a starting point for the cycle.
+
+Our bootstrap phase runs as follows.
+First we assume all $N$ nodes start simultaneously.
+The facilitators for round 0 are hard coded in the program,
+in practice this could be the machines controlled by the developer.
+If some node $u$ not a facilitator,
+it sends the message $\langle \texttt{cp\_msg}, c_{u, 0} \rangle$ to all the facilitators.
+If $u$ is a facilitator, it does the same, but also waits for some time $\Delta + 1$ to collect messages of type \texttt{cp\_msg}.
+After $\Delta + 1$ elapses, it begins the atomic broadcast protocol using the collected CP blocks as the proposal.
+$\C_0$ is agreed at the end of the protocol.
+This concludes the bootstrap phase.
+
+The bootstrap phase can be seen as an idealised version of the setup phase combined with the consensus phase,
+which we describe next.
+
+\subsection{Setup phase}
+The setup phase begins immediately after some consensus result is agreed, but not yet disseminate.
+This could be right after the bootstrap phase.
+The goal is to reach agreement on a list of new facilitators between every node.
+Concretely, for every consensus round $r$, the following is required.
+\begin{enumerate}
+\item \emph{Agreement}: If any correct node receives $\C_r$,
+  then every correct node receives $\C_r$.
+\item \emph{Liveness}: All $N - t$ correct nodes receives a value eventually.
+\end{enumerate}
+
+The setup phase works as follows.
+Assume that facilitators of round $r-1$ have just agreed on $\C_{r-1}$.
+They then immediately broadcast two messages to all the nodes---
+first the consensus message $\langle \texttt{cons\_msg}, \C_{r-1} \rangle$,
+and second the signature message $\langle \texttt{cons\_sig}, r, sig \rangle$.
+
+We make two remarks regarding those message.
+First, broadcasting is inefficient in terms of message  complexity.
+Gossiping is often better in practice but it introduces additional complexity due to its probabilistic nature.
+Thus, to keep the analysis clear in \Cref{sec:analysis}, we use broadcasting.
+Second, recall that channels are not authenticated, 
+and there are no signatures in $\C_{r - 1}$.
+If a non-facilitator sees some $\C_{r - 1}$, it cannot immediately trust it because it may have been forged.
+Thus, To guarantee authenticity, every facilitator sends an additional message that is the signature of $\C_r$.
+
+Continuing with the setup phase.
+Upon receiving $\C_{r - 1}$ and at least $t + 1$ valid signatures by $u$, $u$ performs two asks.
+First, it computes the new facilitators using $\textsf{derive\_facilitator}(C, n)$ (\Cref{alg:facilitator})
+and updates its facilitator list to the result.
+Second, it creates a new CP block using $\textsf{new\_cp}(\C_{r - 1}, r - 1)$ (\Cref{alg:new-cp}).
+This concludes the setup phase.
+
+\begin{algorithm}
+\caption{Function $\textsf{derive\_facilitator}(C, n)$ takes a list of CP blocks $C$ and an integer $n$,
+sort evey element in $C$ by its luck value (the $\lambda$-expression), and outputs the smallest $n$ elements.}
+\label{alg:facilitator}
+\begin{algorithmic}
+\State $\textbf{take} (n, \textbf{sort} (\textbf{map}(\lambda x.\textsf{H}(x || pk \text{ of x}), C)))$
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Function $\textsf{new\_cp}(\C_r, r)$ runs in the context of the caller $u$. TODO}
+\label{alg:new-cp}
+\begin{algorithmic}
+\State $h \gets |B_u|$
+\State $c_{u, h} \gets \langle \textsf{H}(b_{u, h-1}), h, \textsf{H}(\C_r), r, sig_u \rangle$
+\State $B_u \gets B_u \cup c_{u, h}$
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Consensus Phase}
+The consensus phase begins after the new facilitators are selected.
+Assuming the facilitators are of round $r$,
+the goal is to agree on some $\C_r$ between the facilitators.
+There are two scenarios in the consensus phase.
+First is when the node is a facilitator, the other is when the node is not.
+If the node is a facilitator, it must satisfy the following conditions.
+\begin{enumerate}
+\item \emph{Agreement}: If any correct node outputs a CP block $c$, then every
+  correct node outputs $c$.
+\item \emph{Total Order}: If one correct node outputs the sequence of CP blocks
+  $\{c_1, c_i, \dots, c_n\}$ and another has output $\{c'_0, c'_1, \dots,
+  c'_{n'} \}$, then $c_i = c'_i$ for $i \le \min(n, n')$.
+\item \emph{Liveness}: All $n - t$ correct nodes terminate eventually.
+\end{enumerate}
+
+\section{Transaction Protocol}
+\label{sec:tx-protocol}
+
+\begin{algorithm}
+
+    \caption{Function $\textsf{tx\_init}(pk_v, m)$ generates a new TX block,
+    appends it to the caller $u$'s chain and sends it to the counterparty.
+    It is executed in the private context of $u$, i.e. it has access to the $sk_u$ and $B_u$.
+    The necessary arguments are the public key of the counterparty $pk_v$ and the transaction message $m$.}
+    \label{alg:tx-init}
+
+    \begin{algorithmic}
+
+    \State $h \gets |B_u|$
+    \State $txid \in_R \{0, 1\}^{256}$ \Comment A uniformly distributed 256 bit random value.
+    \State $b_{u, h} \gets \langle \textsf{H}(b_{u, h - 1}), h, txid, pk_v, m, sig_u \rangle$
+    \State $B_u \gets B_u \cup \{ b_{u, h} \}$
+
+    \State \textbf{Send} $\langle \texttt{tx\_req}, b_{u,h} \rangle$ to $v$
+
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+    \caption{The TX protocol which runs in the context of node $u$.}
+    \label{alg:tx-proto}
+
+    \begin{algorithmic}
+        \Function{Create}{$z$}
+        \State x
+        \EndFunction
+
+    \end{algorithmic}
+
+\end{algorithm}
+
+\section{Validation Protocol}
+\label{sec:vd-protocol}
 
 \section{Protocol Extensions}
+\label{sec:protocol-extensions}
+
+% \section{Universally Composable Framework}
+% \label{sec:uc-intro}
+
+% In order to analyse the security of a system, a formal notion of security is required.
+% For instance, what does it mean if an encryption algorithm is secure?
+% One may say it is secure if the adversay cannot learn any information about the plaintext from the ciphertext.
+% But what if the adversay has some background information, for instance she may know it is English.
+% Do we then still say the encryption algorithm is secure?
+% Goldwasser and Micali introduced the notion of semantic security~\cite{goldwasser1982probabilistic}.
+% That is, imagine two worlds, a real world and an ideal world.
+% In the real world, the adversay is given the ciphertext,
+% and in the ideal world the adversay has nothing.
+% Then the encryption algorithm is semantically secure if the amount of information that the adversay can learn in the real world is just as much as the ideal world.
+% While our description is informal, the notion of security can be formally captured in this way.
+
+% Security sensitive distributed systems such as secure multi-party computation and blockchain systems also require a formal notion of security so that they can be analysed.
+% Fortunately, the idea from semantic security can also be applied in a distributed setting.
+% In an ideal world, we create an ideal functionality $\Fideal$ that performs all the computation on behalf of every node in the network.
+% The nodes act as dummies and only relay messages between $\Fideal$ and the environment.
+% Thus $\Fideal$ essentially becomes the specification of the distributed protocol.
+% If the execution of the real world protocol is indistinguishable from the ideal emulation in the presence of some adversay,
+% then the real world protocol is secure as per the ideal specification.
+% This is in essence the Universally Composable (UC) framework.
+
+% We use the UC framework not only because it suits our needs, 
+% it is also the only framework used in modelling blockchain systems to the best of our knowledge
+% ~\cite{todo}TODO.
+% In the remainder of this section we give an overview of the UC framework.
+% A detailed treatment  can be found in~\cite{canetti2001universally}.
+
+% \subsection{Model of Computation}
+% The model of computation considered in the UC framework is the interactive Turing machine (ITM).
+% Specifically, an ITM is an extension of the Turing machine with externally writable tapes which are the following.
+% \begin{enumerate}
+% \item input tape---TODO
+% \item incoming communication tape---TODO
+% \item subroutine output tape---TODO
+% \end{enumerate}
+% ITM can be seen as a specification or an algorithm, a machine running an ITM is an ITM instance (ITI).
+% To communicate, an ITI can write on the externally writable tapes of other ITIs.
+% The writing ITI then pauses exeuction, the receiving ITI begins execution.
+% Consequently, only one ITI is running at any point in time.
+
+% \subsection{Simulation-based Security}
+% Simulation-based security, also known as ideal/real paradigm is a model for defining security.
+% The model consist of a set of , the environment $\E$, the adversay $\A$,
+% the protocol ITM $\Preal$ and zero or more ideal functionalities $\mathcal{F}_0, \mathcal{F}_1, \dots$.
+% The 
+% Other than the machines running the protocol under consideration, the model contains two extra entities,
+% the environment $\E$ and the adversary $\A$.
+% $\E$ can be seen as users of the protocol, it can only provide input and receive output from $\A$ and the protocol.
+
+% Control function TODO
+
+% \subsection{Universal Composability}
+
+% \section{Formal Specification}
+% \label{sec:formal-model}
+
+% The formal model follows the same structure as \ref{fig:todo}TODO.
+
+% \begin{figure}[h]
+% \begin{framed}
+% \small  % end with \normalsize
+
+% \textbf{Protocol}\,$\Preal$
+
+% The ETC protocol. On initialisation do the following.
+% \begin{itemize}
+% \item Generate public and private key pair, $pk$ and $sk$.
+% \item Set personal chain $C := \{genesis()\}$.
+% \item Set checkpoint buffer $\C' := \varnothing$.
+% \item Set the facilitators $F$ to the bootstrap nodes provided by $\E$.
+% \item Set the latest round $r_{\text{latest}} = 0$.
+% \end{itemize}
+
+% Run the protocol as specified below after initialisation.
+% \begin{itemize}
+
+% \item Upon (\texttt{consensus}, $D$) from $\F^r_\text{BFT}$,
+%     if $r_{\text{latest}} >= r$ then do nothing and pause,
+%     otherwise do $try_add_cp(D, r)$
+%     send (\texttt{propose})
+
+% \item Upon (\texttt{consensus}, $D$, $r$) from $p \in \P$,
+
+% \item Upon (\texttt{checkpoint}, $c$) from $p \in \P$,
+
+% \item Upon (\texttt{tx\_init}, $m$, $p$) from $\E$, 
+%     $C := C \cup \{new\_tx(m)\}$,
+%     send (\texttt{tx\_req}, latext TX in $C$) to $p$.
+
+% \item Upon (\texttt{tx\_req}, $m$) from $p \in \P$,
+%     $C := C \cup \{new\_tx(m)\}$,
+%     let $m'$ be the latest TX in $C$ and set $m$ to be the other half of $m'$,
+%     send (\texttt{tx\_resp}, $m'$) to $p$.
+%     Output the complete TX to $\E$.
+
+% \item Upon (\texttt{tx\_resp}, $m'$) from $p \in \P$,
+%     find the corresponding pair of $m'$ and name it $m$,
+%     add $m'$ as the other half of $m$.
+%     Output the complete TX to $\E$.
+
+% \item Upon (\texttt{vd\_init}, $s$) from $\E$,
+%     if the sequence number $s$ does not exist in $C$ or the block with $s$ does not have the other half then do nothing.
+%     Otherwise send (\texttt{vd\_req}, $s'$) to $p'$ where $s'$ is the sequence number of the other half and $p'$ is the counterparty.
+
+% \item Upon (\texttt{vd\_req}, $s'$) from $p \in \P$,
+%     if the sequence number $s'$ does not exist then do nothing.
+%     Otherwise send (\texttt{vd\_resp}, $agreed\_fragment(s')$) to $p$.
+
+% \item Upon (\texttt{vd\_resp}, $x$) from $p \in \P$,
+%     run $validate(x)$ and output result to $\E$.
+
+% \end{itemize}
+
+% \normalsize
+% \end{framed}
+% \end{figure}
+
+% \begin{framed}
+% \textbf{Functionality}\,$\F^r_\text{BFT}$
+
+% The ideal Byzantine consensus protocol, parameterised by the round number $r$.
+
+% \begin{itemize}
+% \item Upon (\texttt{propose}, $r$, $C$),
+% \item Upon (\texttt{fetch}),
+% \end{itemize}
+% \end{framed}
+
+% \subsection{Discussion}
+
+% \subsubsection{Global clock for synchronisation}
+% Where there is no Byzantine corruption, we conjecture that our protocol runs in the asynchronous setting.
+% However, security of many Byzantine consensus algorithms, especially the one we adopted---HoneyBadgerBFT,
+% fall apart when there is dynamic corruption.
+% In order to enforce that the corrupted machines do not change when running an instance of some Byzantine consensus algorithm,
+% we introduce synchrony.
+
+% We make use of a global clock... TODO
+
+% Dynamic corruption only between different clock ticks is enforced by the control function.
+
+% \subsubsection{Static corruption versus dynamic corruption}
+% A subtlety exists when modelling the type of corruption.
+% In dynamic corruption\footnote{In \cite{canetti2001universally}, dynamic corruption is termed Byzantine corruption.},
+% $\A$ take full control of a number of machines and also learns all of their states.
+% The states include private keys.
+% Thus, using the dynamic corruption model from \cite{canetti2001universally} we cannot guarantee security as
+% $\A$ can corrupt different machines over time and eventually learn all the private keys..
+
+% First is to use static corruption, where the corrupted machines are fixed.
+% While this circumvents our problem, it is a much weaker model.
+% Alternatively, we modify the aforementioned dynamic corruption model and weaken the adversary's abilities.
+% In particular, we introduce that forgetful adversary that only remembers the state of the currently corrupted nodes and forget the state of the recovered nodes.

### 2017-06-06 thesis/chapters/ch2-model.tex
Writing TX protocol

@@
 
-\subsection{Transaction Protocol}
-The TX protocol~\footnote{It is also known as True Halves.
-See \url{https://github.com/Tribler/tribler/pull/2135} for more information.}
-consist of two messages, \texttt{tx\_req} and \texttt{tx\_resp}.
-When a node  wishes to create a transaction,
-it generates a TX block with the corresponding hash pointer, sequence number, counterparty and the transaction message itself.
-It then sends a \texttt{tx\_req} message containing the TX block to the counterparty.
-Upon receiving a TX block at the counterparty, it creates its own version of the TX block (the pair of the sender's TX block),
-and sends it back to the sender in a \texttt{tx\_resp} message.
-At the end of the protocol, both parties should have both versions of the TX block.
+\subsection{Transaction and Validation}
+The TX protocol is a simple request and response protocol.
+The nodes exchange one round of messages and create new TX blocks on their respective chains.
+Thus, as we mentioned before, one transaction should result in two TX blocks.
 
-\subsection{Validation Protocol}
 The consensus and transaction protocol by themselves do not provide a mechanism to detect forks or other forms of tamperaing.
-This is the goal of the validation protocol.
+Thus we need a validation protocol to counteract malicious behaviour.
+When a node wish to validate one of its TX, it asks the counterparty for the \emph{fragment} of the TX.
+A fragment of a TX is a section of the chain beginning and ending with CP blocks that contains the TX.
+Upon the counterparty's response, the node checks that the CP blocks are in consensus, the hash pointers are valid and his TX is actually in the fragment.
+The TX is valid if these conditions are satisfied.
+Intuitively, this works because it is hard (because hash collision is hard)
+ to create a different chain that begins and ends with the same two CP blocks but with a different middle section.
 
-It also has two message---\texttt{vd\_req} and \texttt{vd\_resp}.
-When a node wishes to validate a transaction,
-it sends a \texttt{vd\_req} with the sequence number $s$ to the party that created the TX block of the transaction.
-Upon receiving a \text{vd\_req}, the node finds two \emph{agreed} CP blocks that surrounds the TX block and computes the \emph{fragment}.
-An agreed CP block means that it has reached consensus.
-A fragment of two CP blocks is a section of the chain where the beginning and the end are defined by those two CP blocks.
-With the fragment, the node replies with a \texttt{vd\_resp} message.
-Upon receiving a \texttt{vd\_resp}, the node checks that the fragment is created correctly.
-Loosely, that means the agreed CP blocks have actually reached consensus,
-the fragment has the correct hash pointers,
-and TX block containing $s$ exists and is valid.
+% The TX protocol~\footnote{It is also known as True Halves.
+% See \url{https://github.com/Tribler/tribler/pull/2135} for more information.}
+% consist of two messages, \texttt{tx\_req} and \texttt{tx\_resp}.
+% When a node  wishes to create a transaction,
+% it generates a TX block with the corresponding hash pointer, sequence number, counterparty and the transaction message itself.
+% It then sends a \texttt{tx\_req} message containing the TX block to the counterparty.
+% Upon receiving a TX block at the counterparty, it creates its own version of the TX block (the pair of the sender's TX block),
+% and sends it back to the sender in a \texttt{tx\_resp} message.
+% At the end of the protocol, both parties should have both versions of the TX block.
+
+% \subsection{Validation Protocol}
+% The consensus and transaction protocol by themselves do not provide a mechanism to detect forks or other forms of tamperaing.
+% This is the goal of the validation protocol.
+
+% It also has two message---\texttt{vd\_req} and \texttt{vd\_resp}.
+% When a node wishes to validate a transaction,
+% it sends a \texttt{vd\_req} with the sequence number $s$ to the party that created the TX block of the transaction.
+% Upon receiving a \text{vd\_req}, the node finds two \emph{agreed} CP blocks that surrounds the TX block and computes the \emph{fragment}.
+% An agreed CP block means that it has reached consensus.
+% A fragment of two CP blocks is a section of the chain where the beginning and the end are defined by those two CP blocks.
+% With the fragment, the node replies with a \texttt{vd\_resp} message.
+% Upon receiving a \texttt{vd\_resp}, the node checks that the fragment is created correctly.
+% Loosely, that means the agreed CP blocks have actually reached consensus,
+% the fragment has the correct hash pointers,
+% and TX block containing $s$ exists and is valid.
 
@@
 and $h = |B_u|$.
+We often use $b_{u, h}$ to denote the latest block.
 There are two types of blocks, TX blocks and CP blocks.
@@
 \subsection{Bootstrap Phase}
+\label{sec:bootstrap}
 Recall that facilitators are computed from the consensus result,
@@
 There are two scenarios in the consensus phase.
-First is when the node is a facilitator, the other is when the node is not.
-If the node is a facilitator, it must satisfy the following conditions.
+First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators.
+Second if the node is a facilitator, it waits for some duration $\delta$ and collect messages of type \texttt{cp\_msg}.
+After $\delta$ is elapsed, it begins the atomic broadcast algorithm.
+In particular, the facilitators must satisfy the following conditions.
 \begin{enumerate}
@@
 
+We remark that this procedure is the same as what is described in the bootstrap phase (\Cref{sec:bootstrap}),
+but the precondition and the waiting time is different.
+In particular, the bootstrap stage assumed that every node initiated and sent the \texttt{cp\_msg} simultaneously,
+but here we make no such assumption. Further, the bootstrap waits for duration $\Delta$ in order to collect all the CP blocks,
+here we wait for some $\delta$ which is a system variable.
+We make no assumption on the value of $\delta$ in relation to $\Delta$,
+thus it may be the case that not all \texttt{cp\_msg} are delivered before the the facilitators begins the atomic broadcast protocol.
+
+At the core of the consensus phase is the atomic broadcast protocol.
+While using any atomic broadcast protocol will suffice,
+we use a (minimal) simplification of HoneyBadgerBFT as it is designed for blockchain systems and runs in fully asynchronous networks.
+The transactions in HoneyBadgerBFT are first queued in a buffer, the main consensus algorithm starts only when the buffer reaches an optimal size.
+We do not have an infinite stream of CP blocks, thus buffering is unsuitable.
+The main algorithms are kept the same, namely Asynchronous Common Subset, Reliable Broadcast and Binary Agreement.
+At the end of the consensus sphase, some $\C_r$ should be agreed upon.
+This bring us back to the setup phase and the cycle can be started again.
+
 \section{Transaction Protocol}
@@
 
+The TX protocol, shown in \Cref{alg:tx-proto}, is run by all nodes.
+It is also known as True Halves TODO.
+Node that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
+$txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$.
+Then the initiator sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.
+
 \begin{algorithm}
-
-    \caption{Function $\textsf{tx\_init}(pk_v, m)$ generates a new TX block,
-    appends it to the caller $u$'s chain and sends it to the counterparty.
+    \caption{Function $\textsf{new\_tx}(pk_v, m, txid)$ generates a new TX block and appends it to the caller $u$'s chain.
     It is executed in the private context of $u$, i.e. it has access to the $sk_u$ and $B_u$.
-    The necessary arguments are the public key of the counterparty $pk_v$ and the transaction message $m$.}
-    \label{alg:tx-init}
+    The necessary arguments are the public key of the counterparty $pk_v$, the transaction message $m$ and the transaction identifier $txid$.}
+    \label{alg:new-tx}
 
     \begin{algorithmic}
-
     \State $h \gets |B_u|$
-    \State $txid \in_R \{0, 1\}^{256}$ \Comment A uniformly distributed 256 bit random value.
-    \State $b_{u, h} \gets \langle \textsf{H}(b_{u, h - 1}), h, txid, pk_v, m, sig_u \rangle$
-    \State $B_u \gets B_u \cup \{ b_{u, h} \}$
-
-    \State \textbf{Send} $\langle \texttt{tx\_req}, b_{u,h} \rangle$ to $v$
-
+    \State $t_{u, h} \gets \langle \textsf{H}(b_{u, h - 1}), h, txid, pk_v, m, sig_u \rangle$
+    \State $B_u \gets B_u \cup \{ t_{u, h} \}$
     \end{algorithmic}
@@
     \begin{algorithmic}
-        \Function{Create}{$z$}
-        \State x
-        \EndFunction
+        \Upon $\langle \texttt{tx\_req}, t_{v, j} \rangle$
+        \State $txid, pk_v, m \gets t_{v, j}$ \Comment Unpack the TX
+        \State $\textsf{new\_tx}(pk_u, m, txid)$
+        \State store $t_{v, j}$ as the other half of $t_{u, h}$
+        \State send $\langle \texttt{tx\_reqp}, t_{u, h} \rangle$ to $v$
 
+        \Upon $\langle \texttt{tx\_resp}, t_{v, j} \rangle$
+        \State $txid, pk_v, m \gets t_{v, j}$ \Comment Unpack the TX
+        \State store $t_{v, j}$ as the other half of the TX with identifier $txid$
     \end{algorithmic}
+\end{algorithm}
 
-\end{algorithm}
+A key feature of the TX protocol is that it is non-blocking.
+TODO
 

### 2017-06-07 thesis/chapters/ch2-model.tex
Write validation

@@
 
-% The TX protocol~\footnote{It is also known as True Halves.
-% See \url{https://github.com/Tribler/tribler/pull/2135} for more information.}
-% consist of two messages, \texttt{tx\_req} and \texttt{tx\_resp}.
-% When a node  wishes to create a transaction,
-% it generates a TX block with the corresponding hash pointer, sequence number, counterparty and the transaction message itself.
-% It then sends a \texttt{tx\_req} message containing the TX block to the counterparty.
-% Upon receiving a TX block at the counterparty, it creates its own version of the TX block (the pair of the sender's TX block),
-% and sends it back to the sender in a \texttt{tx\_resp} message.
-% At the end of the protocol, both parties should have both versions of the TX block.
-
-% \subsection{Validation Protocol}
-% The consensus and transaction protocol by themselves do not provide a mechanism to detect forks or other forms of tamperaing.
-% This is the goal of the validation protocol.
-
-% It also has two message---\texttt{vd\_req} and \texttt{vd\_resp}.
-% When a node wishes to validate a transaction,
-% it sends a \texttt{vd\_req} with the sequence number $s$ to the party that created the TX block of the transaction.
-% Upon receiving a \text{vd\_req}, the node finds two \emph{agreed} CP blocks that surrounds the TX block and computes the \emph{fragment}.
-% An agreed CP block means that it has reached consensus.
-% A fragment of two CP blocks is a section of the chain where the beginning and the end are defined by those two CP blocks.
-% With the fragment, the node replies with a \texttt{vd\_resp} message.
-% Upon receiving a \texttt{vd\_resp}, the node checks that the fragment is created correctly.
-% Loosely, that means the agreed CP blocks have actually reached consensus,
-% the fragment has the correct hash pointers,
-% and TX block containing $s$ exists and is valid.
-
 \subsection{Combined Protocol}
@@
 \begin{itemize}
-\item Lower case (e.g. $x$) denotes a scalar object.
+\item Lower case (e.g. $x$) denotes a scalar object or a tuple.
 \item Upper case (e.g. $X$) denotes a set or a constant.
@@
 \begin{algorithmic}
-\State $\textbf{take} (n, \textbf{sort} (\textbf{map}(\lambda x.\textsf{H}(x || pk \text{ of x}), C)))$
+\State $\textbf{take} (n, \textbf{sort} (\textbf{map}(\lambda x.\textsf{H}(x || pk \text{ of } x), C)))$
 \end{algorithmic}
@@
 
+TODO assume the wait time is much longer than network delay.
+
 At the core of the consensus phase is the atomic broadcast protocol.
@@
 The TX protocol, shown in \Cref{alg:tx-proto}, is run by all nodes.
-It is also known as True Halves TODO.
+It is also known as True Halves, first described by Veldhuisen~\cite[Chapter~3.2]{truehalves}.
 Node that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
@@
 \begin{algorithm}
-    \caption{The TX protocol which runs in the context of node $u$.}
+    \caption{The TX protocol which runs in the context of node $u$. TODO}
     \label{alg:tx-proto}
@@
     \begin{algorithmic}
-        \Upon $\langle \texttt{tx\_req}, t_{v, j} \rangle$
-        \State $txid, pk_v, m \gets t_{v, j}$ \Comment Unpack the TX
+        \Upon $\langle \texttt{tx\_req}, t_{v, j} \rangle$ from $v$
+        \State $txid, pk_v, m \gets t_{v, j}$ \Comment unpack the TX
         \State $\textsf{new\_tx}(pk_u, m, txid)$
-        \State store $t_{v, j}$ as the other half of $t_{u, h}$
+        \State store $t_{v, j}$ as the pair of $t_{u, h}$
         \State send $\langle \texttt{tx\_reqp}, t_{u, h} \rangle$ to $v$
 
-        \Upon $\langle \texttt{tx\_resp}, t_{v, j} \rangle$
-        \State $txid, pk_v, m \gets t_{v, j}$ \Comment Unpack the TX
-        \State store $t_{v, j}$ as the other half of the TX with identifier $txid$
+        \Upon $\langle \texttt{tx\_resp}, t_{v, j} \rangle$ from $v$
+        \State $txid, pk_v, m \gets t_{v, j}$ \Comment unpack the TX
+        \State store $t_{v, j}$ as the pair of the TX with identifier $txid$
     \end{algorithmic}
@@
 A key feature of the TX protocol is that it is non-blocking.
-TODO
+At no time in \Cref{alg:new-tx} or \Cref{alg:tx-proto} do we need to hold the chain state and wait for some message to be delivered before committing a new block to the chain.
+This allows for high concurrency where we can call $\textsf{new\_tx}(\cdot)$ multiple times without waiting for the corresponding \texttt{tx\_resp} messages.
 
@@
 \label{sec:vd-protocol}
+
+Up to this point, we do not provide a mechanism to detect forks or other forms of tampering or forging, and we cannot detect malicious parties.
+The validation protocol aims to solve this issue.
+The protocol is also a request-response protocol, just like the transaction protocol.
+But before explaining the protocol itself, we first define what it means to be valid.
+
+\subsection{Validity Definition}
+A \emph{valid} TX must be created by following the TX protocol.
+Concretely, a TX $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), i, txid_u, pk_v, m_u, sig_u \rangle$ should satisfy the following conditions.
+\begin{enumerate}
+    \item \emph{Existance of pair}:
+        There exists $t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), j, txid, pk_u, m, sig_v \rangle$,
+        i.e. the pair, where $txid_u = txid_v$, $m_u = m_v$ and the signatures are valid.
+    \item \emph{Existance of fragment}:
+        There exists $\textsf{agreed\_fragment}(t_{u, i})$---call it $F_{u, i}$ and
+        $\textsf{agreed\_fragment}(t_{v, j})$---call it $F_{v, j}$,
+        where the hash pointers are valid and the enclosing CP blocks have valid signatures.
+    \item \emph{Only one pair}:
+        There is only one transaction with $txid$ in $F_{v, j}$.
+    \item \emph{Small fragment}:
+        The $|F_{u, i}| \le L$ and $|F_{v, j}| \le L$.
+    \item \emph{Same consensus round}:
+        Suppose the last block in $F_{u,i}$ and the last block in $F_{v, j}$ are respectively $c_{u, p}$ and $c_{v, q}$. 
+        Then these they must be created from the same consensus round.
+\end{enumerate}
+We make three remarks regarding the definition.
+Firstly, the existance of fragment is the most important property.
+It ensures that the pair of TX blocks exist between a pair of CP blocks that are in consensus.
+This implicitly proves that the TX blocks are in consensus (we show a proof in \Cref{sec:analysis} TODO).
+Secondly, the same consensus round property is to encourage node to remain active in the network.
+Recall that the facilitators waits for duration $\Delta$, which is much longer than the network delay.
+Thus the \texttt{cp\_msg} should be delivered to the facilitators before the consensus starts.
+Lastly, TX blocks that do not satisfy the requirement above are not invalid.
+In fact, most new TX blocks cannot satisfy the Existance of fragment property because the CP blocks following the new transactions have not reached consensus.
+Moreoever, very new TX blocks may not satisfy the Existance of pair property because the counterparty may have not yet sent \texttt{tx\_resp}.
+Next we define \emph{invalid} transactions.
+\begin{enumerate}
+    \item \emph{Bad pair}:
+        $txid_u = txid_v \wedge m_u \ne m_v$.
+    \item \emph{Invalid counterparty}:
+        $txid_u = txid_v$, but $pk_u$ in $t_{v, j}$ does not match $u$'s public key.
+    \item \emph{Invalid hash pointer}:
+        Suppose $F_{u, i}$ and $F_{v, j}$ are available, but the transactions in the fragments do not form a chain.
+    \item \emph{Large fragment}:
+        $F_{u, i} > L \vee F_{v, j} > L$.
+    \item \emph{Different consensus round}:
+        The last block in $F_{u,i}$ and the last block in $F_{v, j}$ are respectively $c_{u, p}$ and $c_{v, q}$. 
+        But these CP blocks are not created from the same consensus round.
+\end{enumerate}
+The final two properties are the negation of the final two properties in the valid definition.
+The other three properties are for detecting tampering.
+Any transaction that do not fall into the valid or invalid category are known as \emph{unknown} transactions.
+
+\subsection{Validation Protocol}
+With the validity definition, we are ready to construct a protocol for determining the validity of transactions.
+The protocol is a simple response and request protocol (\Cref{alg:vd-proto}).
+If $u$ wishes to validate some TX with ID $txid$ and counterparty $v$, it sends $\langle \texttt{vd\_req}, txid \rangle$ to $v$.
+The desired properties of the validation protocol are as follows.
+\begin{enumerate}
+    \item \emph{Correctness}:
+        The validation protocol outputs the correct result
+        according to the aforementioned validity definition.
+    \item \emph{Agreement}:
+        If any correct node decides on the validity of a transaction,
+        then all other correct nodes are able to reach the same conclusion.
+    \item \emph{Liveness}:
+        Any valid transactions can be validated eventually.
+    \item \emph{Unforgeability}:
+        If some transaction is valid, it cannot be forged into an invalid transaction.
+        If some transaction is invalid, it cannot be forged into a valid transaction.
+\end{enumerate}
+
+\begin{algorithm}
+\caption{Validation protocol}
+\label{alg:vd-proto}
+
+\begin{algorithmic}
+    \Upon $\langle \texttt{vd\_req}, txid \rangle$ from $v$
+        \State $t_{u, i} \gets \text{the transaction identified by } txid$
+        \State $F_{u, i} \gets \textsf{agreed\_fragment}(t_{u, i})$
+        \State send $\langle \texttt{vd\_resp}, txid, F_{u, i} \rangle$ to $v$
+
+    \Upon $\langle \texttt{vd\_resp}, txid, F_{v, j} \rangle$ from $v$
+        \State $t_{u, i} \gets \text{the transaction identified by } txid$
+        \State set the validity of $t_{u, i}$ to $\textsf{get\_validity}(t_{u, i}, F_{v, j})$
+\end{algorithmic}
+\end{algorithm}
+
+The core of the validation protocol lies within the $\textsf{get\_validity}(\cdot)$ function (\Cref{alg:get-validity}).
+It essentially implements all the aforementioned validity requirements.
+If they are all satisfied, the function returns \emph{valid}.
+Otherwise, depending on which stage the function failed, it will return either \emph{unknown} or \emph{invalid}.
+We prove in the following chapter that the validation protocol satisfies the most of the desired properties.
+
+\begin{algorithm}
+\caption{Function $\textsf{get\_validity}(t_{u, i}, F_{v, j})$ runs in the private context of $u$.
+$t_{u, i}$ is the transaction that $u$ wishes to verify, and $F_{v, j}$ is the corresponding fragment received from $v$.}
+\label{alg:get-validity}
+
+\begin{algorithmic}
+
+    \State $F_{u, i} \gets \textsf{agreed\_fragment}(t_{u, i})$
+    \If{$F_{u, i} = \bot$}
+        \State \Return \emph{unknown}
+    \EndIf
+    \State
+
+    \State $c_{v, a} \gets \textsf{first}(F_{v, j})$
+    \State $c_{v, b} \gets \textsf{last}(F_{v, j})$
+    \If{$c_{v, a}$ or $c_{v, b}$ are not in consensus}
+        \State \Return \emph{unknown}
+    \EndIf
+    \State
+
+    \State $txid, pk_v, m \gets t_{u, i}$
+    \If{number of blocks of $txid$ in $F_{v, j} = 0$}
+        \State \Return \emph{unknown}
+    \EndIf
+    \State
+
+    \If{number of blocks of $txid$ in $F_{v, j} > 1$}
+        \State \Return \emph{invalid}
+    \EndIf
+    \State
+
+    % txid are equal now, but bad hash pointer
+    \If{sequence number in $F_{v, j}$ is correct (sequential)}
+        \If{hash pointers in $F_{v, j}$ is wrong}
+            \State \Return \emph{invalid}
+        \EndIf
+    \EndIf
+    \State
+
+    \State $txid', pk'_u, m' \gets t_{v, j}$
+    \If{$m \ne m' \vee pk_u \ne pk_u \vee |F_{v, j}| > L$}
+        \State \Return \emph{invalid}
+    \EndIf
+    \State
+
+    \State $c_{u, b} \gets \textsf{last}(F_{u, i})$
+    \If{$c_{u, b}$ is not created using the same $\C_r$ as $c_{v, b}$}
+        \State \Return \emph{invalid}
+    \EndIf
+    \State
+
+    \State \Return \emph{valid}
+\end{algorithmic}
+\end{algorithm}
+
+We make two remarks.
+First, just like the TX protocol, we do not block at any part of the protocol.
+Second, suppose some $F_{v, j}$ validates $t_{u, i}$, then that does not imply that $t_{u, i}$ only has one pair $t_{v, j}$.
+Our validity requirement only requires that there is only one $t_{v, j}$ in the correct consensus round.
+The counterparty may create any number of fake pairs in a later consensus rounds.
+But these fake pairs only pollutes the chain of $v$ and can never be validated because the round is incorrect.
 

### 2017-06-08 thesis/chapters/ch2-model.tex
Fix TODOs

@@
 Application developers using the framework should be able to create any application they wish.
-Further, we do not impose on a consensus algorithm,
-the application developer is free to choose between proof-of-work, proof-of-stake,
-Byzantine consensus as long as it  satisfies the properties in \Cref{sec:todo}TODO.
+Further, we do not impose on a consensus algorithm, as long as it satisfies the properties of atomic broadcast which we describe in~\Cref{sec:overview-cons}.
 
@@
 \subsection{Consensus Protocol}\label{sec:overview-cons}
-Before describing our consensus protocol, we take a brief detour to explain Byzantine consensus,
-which is a fundamental building block of our consensus protocol.
+Before describing our consensus protocol, we take a brief detour to explain \emph{atomic broadcast} primitive and contrast it with other well known primitives.
+It is a fundamental building block of our consensus protocol.
 
-\subsubsection*{Byzantine consensus}
-Byzantine consensus is also known as \emph{atomic broadcast}.
-Roughly speaking, atomic broadcast need to satisfy the following properties.
+\subsubsection*{Atomic broadcast}
+Atomic broadcast is an especially useful primitive for blockchain systems.
+It allows nodes to propose values (which can be transactions),
+and the goal is to reach a common ordering of the values---roughly speaking,
+the output is an ordered set of the propsal.
+Concretely, atomic broadcast needs to satisfy the following properties (adapted from~\cite{miller2016honey}).
 \begin{enumerate}
-\item TODO
+\item \emph{Agreement}:
+    If any correct node outputs a value $v$, then every correct node outputs $v$.
+\item \emph{Total order}:
+    If one correct node has output the sequence of values $v_0, v_1, \dots, v_j$
+    and another has output $v'_0, v'_1, \dots, v_j'$, then $v_i = v'_i$ for $i \le \min(j, j')$.
+\item \emph{Liveness}:
+    If $v$ is an input to all the correct nodes, then it is eventually output by all the correct nodes.
 \end{enumerate}
-TODO downsides (can't run with too many nodes) (high message complexity)
-We stress that Byzantine consensus is not the same as Byzantine agreement or the Byzantine general's problem.
-Byzantine agreement is TODO
+The liveness property also implies censorship resilience,
+i.e. an adversay cannot block some $v$ if all the correct nodes are committed to it.
 
-The literature on Byzantine consensus and atomic broadcast is rich, some noteable ones include TODO.
-Thus in the ETC consensus protocol, we assume there exist an "off-the-shelf" Byzantine consensus algorithm which we can use
-(in our implementation we use HoneyBadgerBFT and motivate our choice in TODO).
+The atomic broadcast properties can be seen as a stronger version of reliable broadcast.
+In reliable broadcast (e.g. Bracha's reliable broadcast~\cite{bracha1984asynchronous}),
+there is only one sender and the following properties must be satisfied.
+\begin{enumerate}
+    \item \emph{Agreement}:
+        If any two correct nodes deliver $v$ and $v'$, then $v = v'$.
+    \item \emph{Totality}:
+        If any correct node delivers $v$, then all correct nodes deliver $v$.
+    \item \emph{Validity}:
+        If the sender is correct and inputs $v$, then all correct nodes deliver $v$.
+\end{enumerate}
+The main different here is that reliable broadcast does not guarantee order, it only guarantees delivery.
+If there are more than one sender, or even a single sender with multiple messages,
+the order of delivery may be different on different nodes.
 
-In our case, every node do not propose some arbitrary data, but a set of CP blocks.
-Thus the result of the consensus is the set union of all the legitimate proposals.
+Atomic broadcast also is not the same as Byzantine agreement (BA).
+In particular, BA algorithms lets nodes propose a binary value,
+then output a binary value with respect to the following properties.
+\begin{enumerate}
+    \item \emph{Agreement}:
+        If any correct node outputs $v$, then every correct node outputs $v$.
+    \item \emph{Termination}:
+        If all correct nodes make a proposal, then every correct node outputs a value.
+    \item \emph{Validity}:
+        If any correct node outputs $v$, then at least one correct node proposed $v$.
+\end{enumerate}
+The primary difference here is that BA only considers only binary value,
+and the goal is to reach agreement on the binary value according with respect to the validity property.
+Whereas the agreement is on all the proposals from honest nodes in atomic broadcast.
+
+The main drawback with atomic broadcast (and other similar protocols, i.e. reliable broadcast and Byzantine agreement)
+is the high message complexity.
+Typically, such protocols have the complexity of $O(n^2)$, where $n$ is the number of participants.
+This may work with a small number of nodes, but it is infeasible for blockchain systems where thousands of nodes are involved.
+
+% The literature on Byzantine consensus and atomic broadcast is rich, some noteable ones include TODO.
+% Thus in the ETC consensus protocol, we assume there exist an "off-the-shelf" Byzantine consensus algorithm which we can use
+% (in our implementation we use HoneyBadgerBFT and motivate our choice in TODO).
+
+% In our case, every node do not propose some arbitrary data, but a set of CP blocks.
+% Thus the result of the consensus is the set union of all the legitimate proposals.
 
 \subsubsection*{Consensus Protocol} 
-The consensus protocol runs continuously in rounds.
-That is because a blockchain systems always need to reach consensus on new values, or CP blocks in our case.
+The consensus protocol runs continuously in rounds
+because a blockchain systems always need to reach consensus on new values, which are CP blocks in our case.
 This can be seen as running infinitly many rounds of some Byzantine consensus algorithm,
@@
 As we mentioned earlier, the high message complexity prohibits us from running a Byzantine consensus algorithm on a large network.
-Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks,
-and then use the CP blocks as proposals of the Byzantine consensus algorithm.
+Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks and use them as the proposal.
 The facilitators are elected using a \emph{luck value}, which is computed using $H(\C_r || pk_i)$,
@@
 },
-but messages may be re-ordered and delayed by at most some time $\Delta$.
+but messages may be re-ordered and delayed by at most some time $\Delta$,
+this is sometimes known as a $\Delta$-synchronous network.
 Nodes are identified by their public key thus we assume there exist a Public Key Infrastructure (PKI).
-TODO this is weak synchrony?
 In addition, we assume the existance of cryptographically secure hash function $\textsf{H}(\cdot)$,
@@
 $n \ge 3t + 1$ must hold.
+
+We use a restricted version of the adaptive corruption model.
+The first restriction is that corrupted node can only change across rounds.
+That is, if a round has started, the corrupted nodes cannot changed until the next round.
+The second restriction is that the adversary, presumably controlling all the corrupted nodes, is forgetful.
+Namely the adversary may learn the internal state such as the private key of a corrupted node,
+but if the node recovers, then the adversary must forget the private key.
+Otherwise the adversary can eventually learn all the private keys and sabotage the system.
 
@@
 \begin{algorithm}
-\caption{Function $\textsf{new\_cp}(\C_r, r)$ runs in the context of the caller $u$. TODO}
+\caption{Function $\textsf{new\_cp}(\C_r, r)$ runs in the context of the caller $u$.
+It creates a new CP block and appends it to $u$'s chain.}
 \label{alg:new-cp}
@@
 First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators.
-Second if the node is a facilitator, it waits for some duration $\delta$ and collect messages of type \texttt{cp\_msg}.
-After $\delta$ is elapsed, it begins the atomic broadcast algorithm.
+Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D >> \Delta$.
+After $D$ is elapsed, it begins the atomic broadcast algorithm.
 In particular, the facilitators must satisfy the following conditions.
@@
 In particular, the bootstrap stage assumed that every node initiated and sent the \texttt{cp\_msg} simultaneously,
-but here we make no such assumption. Further, the bootstrap waits for duration $\Delta$ in order to collect all the CP blocks,
-here we wait for some $\delta$ which is a system variable.
-We make no assumption on the value of $\delta$ in relation to $\Delta$,
-thus it may be the case that not all \texttt{cp\_msg} are delivered before the the facilitators begins the atomic broadcast protocol.
-
-TODO assume the wait time is much longer than network delay.
+but here we make no such assumption.
+Further, the bootstrap waits for duration $\Delta + 1$ in order to collect all the CP blocks,
+here we wait for some duration $D$---a system parameter,
+it must be much larger than $\Delta + 1$ to ensure that all honest \texttt{cp\_msg} are collected.
 
@@
 \begin{algorithm}
-    \caption{The TX protocol which runs in the context of node $u$. TODO}
+    \caption{The TX protocol which runs in the context of node $u$.}
     \label{alg:tx-proto}
@@
         \State store $t_{v, j}$ as the pair of $t_{u, h}$
-        \State send $\langle \texttt{tx\_reqp}, t_{u, h} \rangle$ to $v$
+        \State send $\langle \texttt{tx\_resp}, t_{u, h} \rangle$ to $v$
 
@@
 \subsection{Validity Definition}
-A \emph{valid} TX must be created by following the TX protocol.
-Concretely, a TX $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), i, txid_u, pk_v, m_u, sig_u \rangle$ should satisfy the following conditions.
-\begin{enumerate}
-    \item \emph{Existance of pair}:
-        There exists $t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), j, txid, pk_u, m, sig_v \rangle$,
-        i.e. the pair, where $txid_u = txid_v$, $m_u = m_v$ and the signatures are valid.
-    \item \emph{Existance of fragment}:
-        There exists $\textsf{agreed\_fragment}(t_{u, i})$---call it $F_{u, i}$ and
-        $\textsf{agreed\_fragment}(t_{v, j})$---call it $F_{v, j}$,
-        where the hash pointers are valid and the enclosing CP blocks have valid signatures.
-    \item \emph{Only one pair}:
-        There is only one transaction with $txid$ in $F_{v, j}$.
-    \item \emph{Small fragment}:
-        The $|F_{u, i}| \le L$ and $|F_{v, j}| \le L$.
-    \item \emph{Same consensus round}:
-        Suppose the last block in $F_{u,i}$ and the last block in $F_{v, j}$ are respectively $c_{u, p}$ and $c_{v, q}$. 
-        Then these they must be created from the same consensus round.
-\end{enumerate}
-We make three remarks regarding the definition.
-Firstly, the existance of fragment is the most important property.
-It ensures that the pair of TX blocks exist between a pair of CP blocks that are in consensus.
-This implicitly proves that the TX blocks are in consensus (we show a proof in \Cref{sec:analysis} TODO).
-Secondly, the same consensus round property is to encourage node to remain active in the network.
-Recall that the facilitators waits for duration $\Delta$, which is much longer than the network delay.
-Thus the \texttt{cp\_msg} should be delivered to the facilitators before the consensus starts.
-Lastly, TX blocks that do not satisfy the requirement above are not invalid.
-In fact, most new TX blocks cannot satisfy the Existance of fragment property because the CP blocks following the new transactions have not reached consensus.
-Moreoever, very new TX blocks may not satisfy the Existance of pair property because the counterparty may have not yet sent \texttt{tx\_resp}.
-Next we define \emph{invalid} transactions.
-\begin{enumerate}
-    \item \emph{Bad pair}:
-        $txid_u = txid_v \wedge m_u \ne m_v$.
-    \item \emph{Invalid counterparty}:
-        $txid_u = txid_v$, but $pk_u$ in $t_{v, j}$ does not match $u$'s public key.
-    \item \emph{Invalid hash pointer}:
-        Suppose $F_{u, i}$ and $F_{v, j}$ are available, but the transactions in the fragments do not form a chain.
-    \item \emph{Large fragment}:
-        $F_{u, i} > L \vee F_{v, j} > L$.
-    \item \emph{Different consensus round}:
-        The last block in $F_{u,i}$ and the last block in $F_{v, j}$ are respectively $c_{u, p}$ and $c_{v, q}$. 
-        But these CP blocks are not created from the same consensus round.
-\end{enumerate}
-The final two properties are the negation of the final two properties in the valid definition.
-The other three properties are for detecting tampering.
-Any transaction that do not fall into the valid or invalid category are known as \emph{unknown} transactions.
+A transaction can be in one of three states in terms of validity---\emph{valid}, \emph{invalid} and \emph{unknown}.
+Given a fragment $F$, the validity of the transaction $t$ is captured by the function $\textsf{get\_validity}(t, F)$ (\Cref{alg:get-validity}).
+The first four conditions (up to \Cref{line:valid-fragment}) essentially check whether the fragment is the one that the verifier needs.
+If it is not, then the verifier cannot make any decision and return \emph{unknown}.
+This is likely to be the case for new transactions because $\textsf{agreed\_fragment}(\cdot)$ would be $\bot$.
+The next two conditions checks for tampering or missing blocks, if any of these misconducts are detected, then the TX is invalid.
+
+Note that the validity is on a transaction (two TX blocks with the same $txid$), rather than on one TX block owned by a single party.
+It is defined this way because the malicious sender may create new TX blocks in their own chain but never send \texttt{tx\_req} messages.
+In that case, it may seem that the counterparty, who is honest, purposefully ommitted TX blocks.
+But in reality, it was the malicious sender who did not follow the protocol.
+Thus, in such cases the whole transaction, identified by its $txid$ is marked as invalid.
+
+\begin{algorithm}
+\caption{Function $\textsf{get\_validity}(t_{u, i}, F_{v, j})$ runs in the private context of $u$.
+$t_{u, i}$ is the transaction that $u$ wishes to verify, and $F_{v, j}$ is the corresponding fragment received from $v$.}
+\label{alg:get-validity}
+
+\begin{algorithmic}[1]
+
+    \State $F_{u, i} \gets \textsf{agreed\_fragment}(t_{u, i})$
+    \If{$F_{u, i} = \bot$}
+        \State \Return \emph{unknown}
+    \EndIf \Comment $u$ has agreed fragment
+    \State
+
+    \State $c_{v, a} \gets \textsf{first}(F_{v, j})$
+    \State $c_{v, b} \gets \textsf{last}(F_{v, j})$
+    \If{$c_{v, a}$ or $c_{v, b}$ are not in consensus}
+        \State \Return \emph{unknown}
+    \EndIf \Comment $v$ has agreed fragment
+    \State
+
+    % txid are equal now, but bad hash pointer
+    \If{sequence number in $F_{v, j}$ is correct (sequential)}
+        \If{hash pointers in $F_{v, j}$ is wrong}
+            \State \Return \emph{unknown}
+        \EndIf
+    \EndIf \Comment correct TrustChain structure
+    \State
+
+    \State $c_{u, b} \gets \textsf{last}(F_{u, i})$
+    \If{$c_{u, b}$ is not created using the same $\C_r$ as $c_{v, b}$}
+        \State \Return \emph{unknown}
+    \EndIf \Comment correct consensus round
+    \State
+    \label{line:valid-fragment}
+    \State $txid, pk_v, m \gets t_{u, i}$
+    \If{number of blocks of $txid$ in $F_{v, j} \ne 1$}
+        \State \Return \emph{invalid}
+    \EndIf \Comment TX exists
+    \State
+
+    \State $txid', pk'_u, m' \gets t_{v, j}$
+    \If{$m \ne m' \vee pk_u \ne pk_u \vee |F_{v, j}| > L$}
+        \State \Return \emph{invalid}
+    \EndIf \Comment no tampering
+    \State
+
+    \State \Return \emph{valid}
+\end{algorithmic}
+\end{algorithm}
 
@@
     \item \emph{Agreement}:
-        If any correct node decides on the validity of a transaction,
-        then all other correct nodes are able to reach the same conclusion.
+        If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
+        then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
     \item \emph{Liveness}:
@@
 
-The core of the validation protocol lies within the $\textsf{get\_validity}(\cdot)$ function (\Cref{alg:get-validity}).
-It essentially implements all the aforementioned validity requirements.
-If they are all satisfied, the function returns \emph{valid}.
-Otherwise, depending on which stage the function failed, it will return either \emph{unknown} or \emph{invalid}.
-We prove in the following chapter that the validation protocol satisfies the most of the desired properties.
-
-\begin{algorithm}
-\caption{Function $\textsf{get\_validity}(t_{u, i}, F_{v, j})$ runs in the private context of $u$.
-$t_{u, i}$ is the transaction that $u$ wishes to verify, and $F_{v, j}$ is the corresponding fragment received from $v$.}
-\label{alg:get-validity}
-
-\begin{algorithmic}
-
-    \State $F_{u, i} \gets \textsf{agreed\_fragment}(t_{u, i})$
-    \If{$F_{u, i} = \bot$}
-        \State \Return \emph{unknown}
-    \EndIf
-    \State
-
-    \State $c_{v, a} \gets \textsf{first}(F_{v, j})$
-    \State $c_{v, b} \gets \textsf{last}(F_{v, j})$
-    \If{$c_{v, a}$ or $c_{v, b}$ are not in consensus}
-        \State \Return \emph{unknown}
-    \EndIf
-    \State
-
-    \State $txid, pk_v, m \gets t_{u, i}$
-    \If{number of blocks of $txid$ in $F_{v, j} = 0$}
-        \State \Return \emph{unknown}
-    \EndIf
-    \State
-
-    \If{number of blocks of $txid$ in $F_{v, j} > 1$}
-        \State \Return \emph{invalid}
-    \EndIf
-    \State
-
-    % txid are equal now, but bad hash pointer
-    \If{sequence number in $F_{v, j}$ is correct (sequential)}
-        \If{hash pointers in $F_{v, j}$ is wrong}
-            \State \Return \emph{invalid}
-        \EndIf
-    \EndIf
-    \State
-
-    \State $txid', pk'_u, m' \gets t_{v, j}$
-    \If{$m \ne m' \vee pk_u \ne pk_u \vee |F_{v, j}| > L$}
-        \State \Return \emph{invalid}
-    \EndIf
-    \State
-
-    \State $c_{u, b} \gets \textsf{last}(F_{u, i})$
-    \If{$c_{u, b}$ is not created using the same $\C_r$ as $c_{v, b}$}
-        \State \Return \emph{invalid}
-    \EndIf
-    \State
-
-    \State \Return \emph{valid}
-\end{algorithmic}
-\end{algorithm}
-
 We make two remarks.
@@
 \label{sec:protocol-extensions}
+
+(Move to future work?)
+
+Up to this point, we have discussed our protocol in the context of the model and assumptions defined in \Cref{sec:model-assumptions}.
+In this section, we remove a few assumptions and discuss how our architecture is adapted.
 

### 2017-06-08 thesis/chapters/ch3-analysis.tex
Start analysis

\chapter{Analysis}
\label{ch:analysis}

Three aspects:
checkpoints (consensus)
    for every round we need fair lottery, consistent facilitators, agreement, total order, liveness
    for the whole thing, we need global liveness
transactions (tamperproofness)
    implicit consensus on transactions
    maintain chain structure (not possible, but we can do it probabilistically)
performance

### 2017-06-09 thesis/chapters/ch2-model.tex
Start analysis
Rework consensus definition

@@
 \subsection{Consensus Protocol}\label{sec:overview-cons}
-Before describing our consensus protocol, we take a brief detour to explain \emph{atomic broadcast} primitive and contrast it with other well known primitives.
+Before describing our consensus protocol, we take a brief detour to explain asynchronous common subset primitive and contrast it with other well known primitives.
 It is a fundamental building block of our consensus protocol.
 
-\subsubsection*{Atomic broadcast}
-Atomic broadcast is an especially useful primitive for blockchain systems.
-It allows nodes to propose values (which can be transactions),
-and the goal is to reach a common ordering of the values---roughly speaking,
-the output is an ordered set of the propsal.
-Concretely, atomic broadcast needs to satisfy the following properties (adapted from~\cite{miller2016honey}).
+\subsubsection*{Asynchronous common subset}
+TODO give history/background of ACS
+Asynchronous common subset (ACS) is an especially useful primitive for blockchain systems.
+It allows any party to propose a value and the result is the set union of all the proposed values by the majority.
+Concretely, ACS needs to satisfy the following properties (adapted from~\cite{miller2016honey}).
 \begin{enumerate}
+    \item \emph{Validity}:
+        If any correct party outputs a set $C$,
+        then $|C| \ge n - t$ and $C$ contains the input of at least $n - 2t$ parties.
     \item \emph{Agreement}:
-    If any correct node outputs a value $v$, then every correct node outputs $v$.
-\item \emph{Total order}:
-    If one correct node has output the sequence of values $v_0, v_1, \dots, v_j$
-    and another has output $v'_0, v'_1, \dots, v_j'$, then $v_i = v'_i$ for $i \le \min(j, j')$.
-\item \emph{Liveness}:
-    If $v$ is an input to all the correct nodes, then it is eventually output by all the correct nodes.
+        If a correct party outputs $C$, then every party outputs $C$.
+    \item \emph{Totality}:
+        If $n - f$ party receive an input, then all correct parties produce an output.
 \end{enumerate}
-The liveness property also implies censorship resilience,
-i.e. an adversay cannot block some $v$ if all the correct nodes are committed to it.
 
-The atomic broadcast properties can be seen as a stronger version of reliable broadcast.
-In reliable broadcast (e.g. Bracha's reliable broadcast~\cite{bracha1984asynchronous}),
-there is only one sender and the following properties must be satisfied.
-\begin{enumerate}
-    \item \emph{Agreement}:
-        If any two correct nodes deliver $v$ and $v'$, then $v = v'$.
-    \item \emph{Totality}:
-        If any correct node delivers $v$, then all correct nodes deliver $v$.
-    \item \emph{Validity}:
-        If the sender is correct and inputs $v$, then all correct nodes deliver $v$.
-\end{enumerate}
-The main different here is that reliable broadcast does not guarantee order, it only guarantees delivery.
-If there are more than one sender, or even a single sender with multiple messages,
-the order of delivery may be different on different nodes.
+ACS has the nice property of censorship resiliance when compared to other consensus algorithms.
+For instance, Hyperledger and Tender mint uses Practical Byzantine Fault Tolerance (PBFT) as their consensus algorithm.
+In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectivly filter transactions.
+In contract, every party in ACS are involved in the proposal phase,
+and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
 
-Atomic broadcast also is not the same as Byzantine agreement (BA).
-In particular, BA algorithms lets nodes propose a binary value,
-then output a binary value with respect to the following properties.
-\begin{enumerate}
-    \item \emph{Agreement}:
-        If any correct node outputs $v$, then every correct node outputs $v$.
-    \item \emph{Termination}:
-        If all correct nodes make a proposal, then every correct node outputs a value.
-    \item \emph{Validity}:
-        If any correct node outputs $v$, then at least one correct node proposed $v$.
-\end{enumerate}
-The primary difference here is that BA only considers only binary value,
-and the goal is to reach agreement on the binary value according with respect to the validity property.
-Whereas the agreement is on all the proposals from honest nodes in atomic broadcast.
-
-The main drawback with atomic broadcast (and other similar protocols, i.e. reliable broadcast and Byzantine agreement)
-is the high message complexity.
-Typically, such protocols have the complexity of $O(n^2)$, where $n$ is the number of participants.
-This may work with a small number of nodes, but it is infeasible for blockchain systems where thousands of nodes are involved.
+The main drawback with ACS and other BFT protocols is the high message complexity.
+Typically, such protocols have the complexity of $O(n^2)$, where $n$ is the number of parties.
+Hence, it may work with a small number of nodes,
+but it is infeasible for blockchain systems where thousands of nodes are involved.
 
@@
 
+TODO agreed fragment must be consecutive
+
 \section{Consensus Protocol}
@@
 
-Our consensus protocol runs on top of the model describe before, it makes heavy use of CP blocks.
-We describe the protocol step by step.
+Our consensus protocol runs on top of the model described above.
+It is directly related to the creation of CP blocks.
+The objectives of the protocol are to allow honest nodes always make progress (by creating new CP blocks),
+the consensus result in every round should be correct and the facilitators election should be unbiased.
+Concretely, we define the necessary properties as follows.
+\begin{definition}
+\label{def:consensus}
+\textbf{\emph{Necessary properties of the consensus protocol}}
+
+For any $r \in \{ x \in \mathbb{Z}_+ : x < \infty \}$, starting at the genesis round (round 0),
+the system will eventually reach $r$\textsuperscript{th} round.
+Further, for every non-genesis round $r$ ($r > 0$), the following properties must be satisfied.
+\begin{enumerate}
+    \item \emph{Fair lottery}:
+        Every node with a CP block in $\C_{r - 1}$,
+        should have an equal probability to be elected as a facilitator in round $r$.
+    \item \emph{Consistent facilitators}:
+        If one honest node decides on a list of facilitators in round $r$,
+        every other node honest decides on the same list.
+    \item \emph{Validity}:
+        If any correct facilitator outputs a set $C$, then $|C| \ge n - t$ and $C$ contains the input of at least $n - 2t$ facilitators.
+        Note that the input is a set of CP blocks, so the output is a set of a set of CP blocks.
+    \item \emph{Agreement}:
+        If a correct facilitator outputs $C$, then every facilitator outputs $C$.
+    \item \emph{Totality}:
+        If $n - f$ facilitators receive an input, then all correct facilitators produce an output.
+\end{enumerate}
+\end{definition}
+Note that the final three properties are on the facilitators rather than on every node,
+they are the properties ACS. 
+The fair lottery and consistent facilitator are properties on every node,
+they are prerequites for ACS.
+
+We proceed to describe the protocol step by step.
 Starting with the bootstrap phase and then moving on to the actual consensus phase.
@@
 Thus we have a dependency cycle.
-The goal of the bootstrap phase is to give us a starting point for the cycle.
+The goal of the bootstrap phase is to give us a starting point in the cycle.
 
@@
 First we assume all $N$ nodes start simultaneously.
-The facilitators for round 0 are hard coded in the program,
-in practice this could be the machines controlled by the developer.
-If some node $u$ not a facilitator,
+The facilitators for round 0 are hard coded in the program.
+In practice this could be the machines controlled by the developer.
+If some node $u$ is not a facilitator,
 it sends the message $\langle \texttt{cp\_msg}, c_{u, 0} \rangle$ to all the facilitators.
@@
 This could be right after the bootstrap phase.
-The goal is to reach agreement on a list of new facilitators between every node.
-Concretely, for every consensus round $r$, the following is required.
-\begin{enumerate}
-\item \emph{Agreement}: If any correct node receives $\C_r$,
-  then every correct node receives $\C_r$.
-\item \emph{Liveness}: All $N - t$ correct nodes receives a value eventually.
-\end{enumerate}
+The goal is to reach agreement on a list of new facilitators between every node
+and should correspond to the fair lottery and consistent facilitators in \Cref{def:consensus}.
 
@@
 and second the signature message $\langle \texttt{cons\_sig}, r, sig \rangle$.
-
-We make two remarks regarding those message.
-First, broadcasting is inefficient in terms of message  complexity.
-Gossiping is often better in practice but it introduces additional complexity due to its probabilistic nature.
-Thus, to keep the analysis clear in \Cref{sec:analysis}, we use broadcasting.
-Second, recall that channels are not authenticated, 
+The reason for sending \texttt{cons\_sig} is the following.
+Recall that channels are not authenticated, 
 and there are no signatures in $\C_{r - 1}$.
@@
 
-Continuing with the setup phase.
-Upon receiving $\C_{r - 1}$ and at least $t + 1$ valid signatures by $u$, $u$ performs two asks.
-First, it computes the new facilitators using $\textsf{derive\_facilitator}(C, n)$ (\Cref{alg:facilitator})
+Continuing, upon receiving $\C_{r - 1}$ and at least $t + 1$ valid signatures by $u$, $u$ performs two asks.
+First, it computes the new facilitators using $\textsf{get\_facilitator}(C, n)$ (\Cref{alg:facilitator})
 and updates its facilitator list to the result.
@@
 \begin{algorithm}
-\caption{Function $\textsf{derive\_facilitator}(C, n)$ takes a list of CP blocks $C$ and an integer $n$,
+\caption{Function $\textsf{get\_facilitator}(C, n)$ takes a list of CP blocks $C$ and an integer $n$,
 sort evey element in $C$ by its luck value (the $\lambda$-expression), and outputs the smallest $n$ elements.}
@@
 First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators.
-Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D >> \Delta$.
-After $D$ is elapsed, it begins the atomic broadcast algorithm.
-In particular, the facilitators must satisfy the following conditions.
-\begin{enumerate}
-\item \emph{Agreement}: If any correct node outputs a CP block $c$, then every
-  correct node outputs $c$.
-\item \emph{Total Order}: If one correct node outputs the sequence of CP blocks
-  $\{c_1, c_i, \dots, c_n\}$ and another has output $\{c'_0, c'_1, \dots,
-  c'_{n'} \}$, then $c_i = c'_i$ for $i \le \min(n, n')$.
-\item \emph{Liveness}: All $n - t$ correct nodes terminate eventually.
-\end{enumerate}
+Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D \gg \Delta$.
+After $D$ is elapsed, it begins the ACS protocol and some $\C_r$ should be agreed upon by the end of it.
+The facilitators must satisfy the validity, agreement and totality properties from \Cref{def:consensus}.
+This bring us back to the setup phase and the cycle can be started again.
+
+At the core of the consensus phase is the ACS protocol.
+While any ACS protocol that satisfies the standard definition will work,
+we use a simplification of HoneyBadgerBFT as our ACS protocol
+because it is the only (to the best of our knowledge) consensus algorithms designed for blockchain systems.
+We do not use the full HoneyBadgerBFT due to the following.
+First, the transactions in HoneyBadgerBFT are first queued in a buffer and the main consensus algorithm starts only when the buffer reaches an optimal size.
+We do not have an infinite stream of CP blocks, thus buffering is unsuitable.
+Second, HoneyBadgerBFT uses threshold encryption to hide the content of the transactions.
+But we do not reach consensus on transactions, only CP blocks, so hiding CP blocks is meaningless for us as it contains no transactional information.
 
@@
 but here we make no such assumption.
-Further, the bootstrap waits for duration $\Delta + 1$ in order to collect all the CP blocks,
+Furthermore, in the bootstrap phase nodes wait for duration $\Delta + 1$ to collect all the CP blocks,
 here we wait for some duration $D$---a system parameter,
-it must be much larger than $\Delta + 1$ to ensure that all honest \texttt{cp\_msg} are collected.
-
-At the core of the consensus phase is the atomic broadcast protocol.
-While using any atomic broadcast protocol will suffice,
-we use a (minimal) simplification of HoneyBadgerBFT as it is designed for blockchain systems and runs in fully asynchronous networks.
-The transactions in HoneyBadgerBFT are first queued in a buffer, the main consensus algorithm starts only when the buffer reaches an optimal size.
-We do not have an infinite stream of CP blocks, thus buffering is unsuitable.
-The main algorithms are kept the same, namely Asynchronous Common Subset, Reliable Broadcast and Binary Agreement.
-At the end of the consensus sphase, some $\C_r$ should be agreed upon.
-This bring us back to the setup phase and the cycle can be started again.
+which is much larger than $\Delta + 1$ to ensure that all honest \texttt{cp\_msg} are collected.
 
@@
 The desired properties of the validation protocol are as follows.
+\begin{definition}
+\label{def:validation}
 \begin{enumerate}
@@
 \end{enumerate}
+\end{definition}
 
@@
 In this section, we remove a few assumptions and discuss how our architecture is adapted.
+
+Gossip?
+
+Churn?
 

### 2017-06-09 thesis/chapters/ch3-analysis.tex
Rework consensus definition

@@
 \label{ch:analysis}
+
+Up to this point we described our system specification in detail.
+Of course, specification along does not establish any truths.
+In this chapter, we prove two aspects of our system.
+First is correctness,
+where we show that the consensus protocol and the validation protocol satisfies their desired properties
+(\Cref{def:consensus} and \Cref{def:validation} respectively).
+The second is performance,
+where we prove the lower bound of our throughput and show that it out performs classical blockchain systems.
+
+\section{Correctness}
+Our objective in this section is to establish correctness,
+in particular our system should have the properties described in \Cref{def:consensus}.
+
+We first give the proof outline.
 
@@
     maintain chain structure (not possible, but we can do it probabilistically)
-performance
+
+\section{Performance}

### 2017-06-12 thesis/chapters/ch1-intro.tex
First draft of protocol overview
(trying to) write some proof

@@
-TODO some positive intro about blockchain
+\chapter{Introduction}
+\label{ch:intro}
 
-One of the key issues in many blockchain systems today is that they are not scalable.
+Blockchains is one of the most prominent but also controversial technologies of our time.
+Never before did we have an incorruptable ledger that can be shared with mutually untrustful parties.
+% as long as the majority have a common interest
+The ledger in fact can contain any type of data.
+Thus enabling applications in myriad of fields,
+such as finance~\cite{todo}, heath care~\cite{todo}, logistics~\cite{todo}, energy~\cite{todo} and so on.
+However, not everyone is convinced on the practicality of blockchain systems.
+Bram Cohen, the inventor of BitTorrent said
+``It's getting a lot more media attention than the actual impact it's having so far.''~\cite{cohen}
+Forbes listed a few reasons on why we should be skeptical about blockchains~\cite{forbes},
+some of them include its accessibility for the layperson and its scalability issues.
+
+Scalability is indeed one of the key challenges we face in blockchain systems today.
 Bitcoin~\cite{bitcoin}, the largest permissionless\footnote{Explain permissionless} blockchain system
@@
 Finally, we compare our system with other state-of-the-art blockchain systems and conclude in Sections \ref{sec:related} and \ref{sec:conclusion} respectively.
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: "../thesis"
-%%% End:

### 2017-06-12 thesis/chapters/ch2-model.tex
(trying to) write some proof

@@
 Concretely, ACS needs to satisfy the following properties (adapted from~\cite{miller2016honey}).
+\begin{definition}
+\label{def:acs}
 \begin{enumerate}
@@
 Nodes are identified by their public key thus we assume there exist a Public Key Infrastructure (PKI).
-In addition, we assume the existance of cryptographically secure hash function $\textsf{H}(\cdot)$,
-that has preimage resistance, second preimage resistance and collision resistance.
+% TODO everyone has one public key?
+Finally, we use the random oracle model, calls to the random oracle are denoted by $\textsf{H}(\cdot)$.
 
@@
 Further, for every non-genesis round $r$ ($r > 0$), the following properties must be satisfied.
+% TODO rework this -> agreement, validity, fairness, termination
 \begin{enumerate}
@@
 \label{sec:bootstrap}
+% TODO rework this -> bootstrap is simply setting the facilitators in a fair way, no consensus
 Recall that facilitators are computed from the consensus result,
@@
 \subsection{Setup phase}
+\label{sec:setup-phase}
+% TODO setup phase begins with known facilitators from F_r
 The setup phase begins immediately after some consensus result is agreed, but not yet disseminate.
@@
 
-Continuing, upon receiving $\C_{r - 1}$ and at least $t + 1$ valid signatures by $u$, $u$ performs two asks.
+Continuing, upon receiving $\C_{r - 1}$ and at least $n - f$ valid signatures by $u$, $u$ performs two asks.
 First, it computes the new facilitators using $\textsf{get\_facilitator}(C, n)$ (\Cref{alg:facilitator})
@@
 \subsection{Consensus Phase}
+\label{sec:consensus-phase}
 The consensus phase begins after the new facilitators are selected.
@@
 which is much larger than $\Delta + 1$ to ensure that all honest \texttt{cp\_msg} are collected.
+
+% TODO consensus phase ends when F_{r + 1} is selected
+% TODO remove duplicate CP blocks
 

### 2017-06-12 thesis/chapters/ch3-analysis.tex
(trying to) write some proof

@@
 \section{Correctness}
-Our objective in this section is to establish correctness,
-in particular our system should have the properties described in \Cref{def:consensus}.
+Our first objective in this section is to establish truths regarding the correctness of our protocol.
+We do this in two parts.
+First we use mathematical induction to show that properties in \Cref{def:consensus} holds for all round.
+Building on top that, if \Cref{def:consensus} is true, then we can show that many properties in \Cref{def:validation} is true.
 
-We first give the proof outline.
+\textbf{\emph{Correctness of consensus}}
 
-Three aspects:
-checkpoints (consensus)
-    for every round we need fair lottery, consistent facilitators, agreement, total order, liveness
-    for the whole thing, we need global liveness
-transactions (tamperproofness)
-    implicit consensus on transactions
-    maintain chain structure (not possible, but we can do it probabilistically)
+$\forall r \in \mathbb{N}$, the following properties must be satisfied.
+\begin{enumerate}
+    \item \emph{Agreement}:
+        If one honest node outputs a list of facilitators $F_r$,
+        every other node honest decides on $F_r$
+    \item \emph{Validity}:
+        If any honest node outputs $F_r$, then $|F_r| > n$ and $|F_r|$ must contain at least $n - t$ honest nodes.
+    \item \emph{Fairness}:
+        Every node should have an equal probability of becoming a facilitator.
+    \item \emph{Termination}:
+        At the end of the round, every node outputs some $F_r$.
+\end{enumerate}
+
+\subsection{Correctness of Consensus}
+
+We begin our analysis by establishing the fact that the $\textsf{get\_facilitator}(\cdot)$ is fair.
+\begin{lemma}
+\label{lemma:fairness}
+\textbf{\emph{Fairness of facilitator election}}
+
+Every node with a CP block in $\C_r$, should have an equal probability to be elected as a facilitator.
+\end{lemma}
+\begin{proof}
+This directly follows from the random oracle model.
+Recall that the luck value is computed using $\textsf{H}(\C_r, || pk_i)$.
+Since $pk_i$ is unique for every node that has a CP block in $\C_r$, the output of $\textsf{H}(\cdot)$ is uniformly random.
+This implies that the ordered sequence by luck value is uniformly random.
+\end{proof}
+
+Using \Cref{lemma:fairness}, we show that our consensus protocol satisfied \Cref{def:consensus}.
+\begin{lemma}
+\textbf{\emph{Correctness of consensus}}
+$\forall r \in \mathbb{N}$, agreement, validity, fairness and termination holds.
+\end{lemma}
+
+\begin{proof}
+We proof by mathematical induction.
+
+In the base case, agreement, validity fairness and termination follows directly from the bootstrap protocol, due to the bootstrap oracle.
+Note that the result is $F_1$, which indicates the facilitators that are agreed in round 1 and are responsible for driving the ACS protocol in round 2.
+
+For the inductive step, we assume that the properties hold for round $r$.
+Then, to start round $r + 1$, the honest nodes begin sending CP blocks to $F_r$. % TODO but they don't send at the same time!
+Since the honest region in $F_r$ waits for $D$ and $D \gg \Delta$,
+the CP blocks of the honest nodes are guaranteed to be received by at least $n - t$ facilitators.
+The agreement property of ACS (from \Cref{def:acs}) ensures that the consensus result $\C_{r + 1}$ is in consensus. 
+The validity property of ACS ensures that $\C_{r + 1}$ contains the input of at least $n - 2t$ parties, but this is a quorum containing at least one honest facilitator,
+thus $\C_{r + 1}$ contains the CP blocks of all facilitators.
+Observe that $F_{r + 1}$ is computed using the deterministic function $\textsf{get\_facilitators}(\cdot)$.
+Thus agreement of $F_{r + 1}$ follows directly from agreement of ACS.
+Due to the assumption that the adversary cannot corrupt more than $t$ nodes,
+validity of $F_{r + 1}$ also follows from validity of ACS.
+The fairness property follows from \Cref{lemma:fairness}.
+Finally, the termination property holds because $D$ eventually elapses and then ACS eventually terminates.
+This completes our proof.
+
+\end{proof}
+
+\subsection{Correctness of Validation}
+
+Remark: chain structure cannot be satisfied, but can be probabilistically checked
 

### 2017-06-12 thesis/chapters/ch2-model.tex
Give structure

--- 2017-06-09 thesis/chapters/ch2-model.tex
@@
 Concretely, ACS needs to satisfy the following properties (adapted from~\cite{miller2016honey}).
+\begin{definition}
+\label{def:acs}
+\textbf{\emph{Properties of ACS}}
+
 \begin{enumerate}
@@
 this is sometimes known as a $\Delta$-synchronous network.
-Nodes are identified by their public key thus we assume there exist a Public Key Infrastructure (PKI).
-In addition, we assume the existance of cryptographically secure hash function $\textsf{H}(\cdot)$,
-that has preimage resistance, second preimage resistance and collision resistance.
+We assume there exist a Public Key Infrastructure (PKI), and nodes are identified by their unique and permanent public key.
+Finally, we use the random oracle model, i.e. calls to the random oracle are denoted by $\textsf{H}(\cdot)$.
 
@@
 
-\emph{Agreed enclosure} is the same as enclosure with an extra constraint where the CP blocks must be in some consensus result $\C_r$.
-Similarly, \emph{agreed fragment} is computed using agreed enclosure.
+\emph{Agreed enclosure} is the same as enclosure with an extra constraint where the CP blocks must be in some consensus result $\C_r$ and also must be the smallest enclosure.
+That is, suppose a chain is in the form
+    $\{c_{i}, c_{i+1}, t_{i+2}, c_{i+3}\}$\footnote{Usually the notation is of the form $c_{u, i}$, but the node identity is not important here so we simplify it to $c_{i}$}
+    and $c_{i}, c_{i+1}, c_{i+3}$ are in $\C_r, \C_{r+1}, \C_{r+3}$ respectively,
+    then the agreed enclosure of $t_{i+2}$ is $\langle c_{i+1}, c_{i+3}\rangle$ and cannot be $\langle c_{i}, c_{i+3}\rangle$.
+Similarly, \emph{agreed fragment} is computed using the agreed enclosure.
 We define its function to be $\textsf{agreed\_fragment}(\cdot)$
@@
 
-TODO agreed fragment must be consecutive
-
 \section{Consensus Protocol}
@@
 It is directly related to the creation of CP blocks.
-The objectives of the protocol are to allow honest nodes always make progress (by creating new CP blocks),
-the consensus result in every round should be correct and the facilitators election should be unbiased.
-Concretely, we define the necessary properties as follows.
+The objectives of the protocol are to
+    allow honest nodes always make progress (by creating new CP blocks),
+    compute correct consensus result in every round
+    and have unbiased election of facilitators.
+Concretely, we define the necessary properties as follows,
+it is collectively known as the correctness properties of the consensus protocol.
 \begin{definition}
 \label{def:consensus}
-\textbf{\emph{Necessary properties of the consensus protocol}}
+\textbf{\emph{Correctness properties of the consensus protocol}}
 
-For any $r \in \{ x \in \mathbb{Z}_+ : x < \infty \}$, starting at the genesis round (round 0),
-the system will eventually reach $r$\textsuperscript{th} round.
-Further, for every non-genesis round $r$ ($r > 0$), the following properties must be satisfied.
+$\forall r \in \mathbb{N}$, the following properties must be satisfied.
 \begin{enumerate}
-    \item \emph{Fair lottery}:
-        Every node with a CP block in $\C_{r - 1}$,
-        should have an equal probability to be elected as a facilitator in round $r$.
-    \item \emph{Consistent facilitators}:
-        If one honest node decides on a list of facilitators in round $r$,
-        every other node honest decides on the same list.
+    \item \emph{Agreement}:
+        If one honest node outputs a list of facilitators $\F_r$,
+        every other node honest decides on $\F_r$
     \item \emph{Validity}:
-        If any correct facilitator outputs a set $C$, then $|C| \ge n - t$ and $C$ contains the input of at least $n - 2t$ facilitators.
-        Note that the input is a set of CP blocks, so the output is a set of a set of CP blocks.
-    \item \emph{Agreement}:
-        If a correct facilitator outputs $C$, then every facilitator outputs $C$.
-    \item \emph{Totality}:
-        If $n - f$ facilitators receive an input, then all correct facilitators produce an output.
+        If any honest node outputs $\F_r$, then $|\F_r| > n$ and $|\F_r|$ must contain at least $n - t$ honest nodes.
+    \item \emph{Fairness}:
+        Every node should have an equal probability of becoming a facilitator.
+    \item \emph{Termination}:
+        At the end of the round, every node outputs some $\F_r$.
 \end{enumerate}
+
 \end{definition}
@@
 
-Our bootstrap phase runs as follows.
-First we assume all $N$ nodes start simultaneously.
-The facilitators for round 0 are hard coded in the program.
-In practice this could be the machines controlled by the developer.
-If some node $u$ is not a facilitator,
-it sends the message $\langle \texttt{cp\_msg}, c_{u, 0} \rangle$ to all the facilitators.
-If $u$ is a facilitator, it does the same, but also waits for some time $\Delta + 1$ to collect messages of type \texttt{cp\_msg}.
-After $\Delta + 1$ elapses, it begins the atomic broadcast protocol using the collected CP blocks as the proposal.
-$\C_0$ is agreed at the end of the protocol.
+Imagine that there is some bootstrap oracle, that initiates the code on every node.
+The code satisfied all the properties in \Cref{def:consensus}.
+Namely every node has the same set of valid facilitators $\F_1$ that are randomly choosen.
 This concludes the bootstrap phase.
 
-The bootstrap phase can be seen as an idealised version of the setup phase combined with the consensus phase,
-which we describe next.
+In practice, the bootstrap oracle is most likely the software developer and some of the desired properties cannot be achieved.
+In particular, it is not possible to have the fairness property because it is unlikely that the developer knows the identity of every node in advance.
 
-\subsection{Setup phase}
-The setup phase begins immediately after some consensus result is agreed, but not yet disseminate.
-This could be right after the bootstrap phase.
-The goal is to reach agreement on a list of new facilitators between every node
-and should correspond to the fair lottery and consistent facilitators in \Cref{def:consensus}.
+\subsection{Consensus Phase}
+\label{sec:consensus-phase}
+The consensus phase begins when $\F_r$ is available to all the nodes.
+Note that $\F_r$ indicates the facilitators that were elected using result of round $r$ and are responsible for driving the ACS protocol in round $r + 1$.
+The goal is to reach agreement on a set of new facilitators $\F_{r + 1}$ that satisfied \Cref{def:consensus}.
 
-The setup phase works as follows.
-Assume that facilitators of round $r-1$ have just agreed on $\C_{r-1}$.
-They then immediately broadcast two messages to all the nodes---
-first the consensus message $\langle \texttt{cons\_msg}, \C_{r-1} \rangle$,
+There are two scenarios in the consensus phase.
+First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators.
+Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D \gg \Delta$.
+Invalid messages are removed, namely blocks with invalid signatures and duplicate blocks signed by the same key.
+After $D$ elapses, it begins the ACS protocol and some $\C_{r + 1}$ should be agreed upon by the end of it.
+
+At the core of the consensus phase is the ACS protocol.
+While any ACS protocol that satisfies the standard definition will work,
+we use a simplification of HoneyBadgerBFT as our ACS protocol
+because it is the only (to the best of our knowledge) consensus algorithms designed for blockchain systems.
+We do not use the full HoneyBadgerBFT due to the following.
+First, the transactions in HoneyBadgerBFT are first queued in a buffer and the main consensus algorithm starts only when the buffer reaches an optimal size.
+We do not have an infinite stream of CP blocks, thus buffering is unsuitable.
+Second, HoneyBadgerBFT uses threshold encryption to hide the content of the transactions.
+But we do not reach consensus on transactions, only CP blocks, so hiding CP blocks is meaningless for us as it contains no transactional information.
+
+When $\F_{r}$ reaches agreement on $\C_{r+1}$,
+they immediately broadcast two messages to all the nodes---
+first the consensus message $\langle \texttt{cons\_msg}, \C_{r+1} \rangle$,
 and second the signature message $\langle \texttt{cons\_sig}, r, sig \rangle$.
@@
 If a non-facilitator sees some $\C_{r - 1}$, it cannot immediately trust it because it may have been forged.
-Thus, To guarantee authenticity, every facilitator sends an additional message that is the signature of $\C_r$.
+Thus, To guarantee authenticity, every facilitator sends an additional message that is the signature of $\C_{r + 1}$.
 
-Continuing, upon receiving $\C_{r - 1}$ and at least $t + 1$ valid signatures by $u$, $u$ performs two asks.
-First, it computes the new facilitators using $\textsf{get\_facilitator}(C, n)$ (\Cref{alg:facilitator})
+Continuing, upon receiving $\C_{r + 1}$ and at least $n - f$ valid signatures by some $u$, $u$ performs two asks.
+First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1}, r + 1)$ (\Cref{alg:new-cp}).
+Second, it computes the new facilitators using $\textsf{get\_facilitator}(C, n)$ (\Cref{alg:facilitator})
 and updates its facilitator list to the result.
-Second, it creates a new CP block using $\textsf{new\_cp}(\C_{r - 1}, r - 1)$ (\Cref{alg:new-cp}).
-This concludes the setup phase.
+This concludes the consensus phase and brings us back to the beginning of the consensus phase.
 
@@
 \begin{algorithmic}
-\State $\textbf{take} (n, \textbf{sort} (\textbf{map}(\lambda x.\textsf{H}(x || pk \text{ of } x), C)))$
+\State $\textsf{take} (n, \textsf{sort\_by} (\lambda x.\textsf{H}(x || pk \text{ of } x), C))$
 \end{algorithmic}
@@
 \end{algorithm}
-
-\subsection{Consensus Phase}
-The consensus phase begins after the new facilitators are selected.
-Assuming the facilitators are of round $r$,
-the goal is to agree on some $\C_r$ between the facilitators.
-There are two scenarios in the consensus phase.
-First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators.
-Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D \gg \Delta$.
-After $D$ is elapsed, it begins the ACS protocol and some $\C_r$ should be agreed upon by the end of it.
-The facilitators must satisfy the validity, agreement and totality properties from \Cref{def:consensus}.
-This bring us back to the setup phase and the cycle can be started again.
-
-At the core of the consensus phase is the ACS protocol.
-While any ACS protocol that satisfies the standard definition will work,
-we use a simplification of HoneyBadgerBFT as our ACS protocol
-because it is the only (to the best of our knowledge) consensus algorithms designed for blockchain systems.
-We do not use the full HoneyBadgerBFT due to the following.
-First, the transactions in HoneyBadgerBFT are first queued in a buffer and the main consensus algorithm starts only when the buffer reaches an optimal size.
-We do not have an infinite stream of CP blocks, thus buffering is unsuitable.
-Second, HoneyBadgerBFT uses threshold encryption to hide the content of the transactions.
-But we do not reach consensus on transactions, only CP blocks, so hiding CP blocks is meaningless for us as it contains no transactional information.
-
-We remark that this procedure is the same as what is described in the bootstrap phase (\Cref{sec:bootstrap}),
-but the precondition and the waiting time is different.
-In particular, the bootstrap stage assumed that every node initiated and sent the \texttt{cp\_msg} simultaneously,
-but here we make no such assumption.
-Furthermore, in the bootstrap phase nodes wait for duration $\Delta + 1$ to collect all the CP blocks,
-here we wait for some duration $D$---a system parameter,
-which is much larger than $\Delta + 1$ to ensure that all honest \texttt{cp\_msg} are collected.
 
@@
 
+TODO make $w$ the verified of some TX between $u$ and $v$.
+TODO remark that $w$ calls the function.
+
 \begin{algorithm}
-\caption{Function $\textsf{get\_validity}(t_{u, i}, F_{v, j})$ runs in the private context of $u$.
-$t_{u, i}$ is the transaction that $u$ wishes to verify, and $F_{v, j}$ is the corresponding fragment received from $v$.}
+\caption{Function $\textsf{get\_validity}(t_{u, i}, F_{v, j})$ validates the transaction $t_{u, i}$.
+$F_{v, j}$ is the corresponding fragment received from $v$.}
+We assume there exist a valid $F_{u, i}$, namely the agreed fragment of $t_{u, i}$. 
+The caller is $w$, it may be $u$ but this is not necessary.
 \label{alg:get-validity}
@@
 
-    \State $F_{u, i} \gets \textsf{agreed\_fragment}(t_{u, i})$
-    \If{$F_{u, i} = \bot$}
-        \State \Return \emph{unknown}
-    \EndIf \Comment $u$ has agreed fragment
-    \State
+    % \State $F_{u, i} \gets \textsf{agreed\_fragment}(t_{u, i})$
+    % \If{$F_{u, i} = \bot$}
+    %     \State \Return \emph{unknown}
+    % \EndIf \Comment $u$ has agreed fragment
+    % \State
 
@@
 \label{def:validation}
+\textbf{\emph{Implicit consensus on transactions}}
+
 \begin{enumerate}
@@
 Churn?
+
+Sybil?
 

### 2017-06-12 thesis/chapters/ch3-analysis.tex
Give structure

@@
 Building on top that, if \Cref{def:consensus} is true, then we can show that many properties in \Cref{def:validation} is true.
-
-\textbf{\emph{Correctness of consensus}}
-
-$\forall r \in \mathbb{N}$, the following properties must be satisfied.
-\begin{enumerate}
-    \item \emph{Agreement}:
-        If one honest node outputs a list of facilitators $F_r$,
-        every other node honest decides on $F_r$
-    \item \emph{Validity}:
-        If any honest node outputs $F_r$, then $|F_r| > n$ and $|F_r|$ must contain at least $n - t$ honest nodes.
-    \item \emph{Fairness}:
-        Every node should have an equal probability of becoming a facilitator.
-    \item \emph{Termination}:
-        At the end of the round, every node outputs some $F_r$.
-\end{enumerate}
 
@@
 In the base case, agreement, validity fairness and termination follows directly from the bootstrap protocol, due to the bootstrap oracle.
-Note that the result is $F_1$, which indicates the facilitators that are agreed in round 1 and are responsible for driving the ACS protocol in round 2.
+Note that the result is $\F_1$, which indicates the facilitators that are agreed in round 1 and are responsible for driving the ACS protocol in round 2.
 
 For the inductive step, we assume that the properties hold for round $r$.
-Then, to start round $r + 1$, the honest nodes begin sending CP blocks to $F_r$. % TODO but they don't send at the same time!
-Since the honest region in $F_r$ waits for $D$ and $D \gg \Delta$,
+Then, to start round $r + 1$, the honest nodes begin sending CP blocks to $\F_r$. % TODO but they don't send at the same time!
+Since the honest region in $\F_r$ waits for $D$ and $D \gg \Delta$,
 the CP blocks of the honest nodes are guaranteed to be received by at least $n - t$ facilitators.
@@
 thus $\C_{r + 1}$ contains the CP blocks of all facilitators.
-Observe that $F_{r + 1}$ is computed using the deterministic function $\textsf{get\_facilitators}(\cdot)$.
-Thus agreement of $F_{r + 1}$ follows directly from agreement of ACS.
+Observe that $\F_{r + 1}$ is computed using the deterministic function $\textsf{get\_facilitators}(\cdot)$.
+Thus agreement of $\F_{r + 1}$ follows directly from agreement of ACS.
 Due to the assumption that the adversary cannot corrupt more than $t$ nodes,
-validity of $F_{r + 1}$ also follows from validity of ACS.
+validity of $\F_{r + 1}$ also follows from validity of ACS.
 The fairness property follows from \Cref{lemma:fairness}.

### 2017-06-12 thesis/chapters/ch2-model.tex
More proofs

@@
 Otherwise the adversary can eventually learn all the private keys and sabotage the system.
+Finally, we assume computational security.
+That is, the adversary can run polynomial-time algorithms but not exponential-time algorithms.
 
@@
 
-Up to this point, we do not provide a mechanism to detect forks or other forms of tampering or forging, and we cannot detect malicious parties.
+Up to this point, we do not provide a mechanism to detect forks or other forms of tampering or forging.
 The validation protocol aims to solve this issue.
 The protocol is also a request-response protocol, just like the transaction protocol.
-But before explaining the protocol itself, we first define what it means to be valid.
+But before explaining the protocol itself, we first define what it means for a transaction to be valid.
 
@@
 A transaction can be in one of three states in terms of validity---\emph{valid}, \emph{invalid} and \emph{unknown}.
-Given a fragment $F$, the validity of the transaction $t$ is captured by the function $\textsf{get\_validity}(t, F)$ (\Cref{alg:get-validity}).
+Given a fragment $F_{v, j}$, the validity of the transaction $t_{u, i}$ is captured by the function $\textsf{get\_validity}(t, F)$ (\Cref{alg:get-validity}).
 The first four conditions (up to \Cref{line:valid-fragment}) essentially check whether the fragment is the one that the verifier needs.
@@
 
-TODO make $w$ the verified of some TX between $u$ and $v$.
-TODO remark that $w$ calls the function.
+Further, the caller of $\textsf{get\_validity}(t_u{u, i}, F_{v, i})$ is not necessarily 
+$u$\footnote{In practice it often is because after completing the TX protocol the parties are incentivised to check that the counterparty ``did the right thing''.}
+Any node $w$ may call $\textsf{get\_validity}(t_u{u, i}, F_{v, i})$ as long as the caller $w$ has an agreed fragment of $t_{u, i}$---$F_{u, i}$.
+$F_{u, i}$ may be readily available if $w = u$ or it may be from some other \texttt{vd\_resp} message, which we describe next in the validation protocol.
 
@@
 
+    \If{$|F_{v, j}| > L$}
+        \State \Return \emph{unknown}
+    \EndIf \Comment fragment not too long
+    \State
+
     % txid are equal now, but bad hash pointer
@@
     \State $txid', pk'_u, m' \gets t_{v, j}$
-    \If{$m \ne m' \vee pk_u \ne pk_u \vee |F_{v, j}| > L$}
+    \If{$m \ne m' \vee pk_u \ne pk'_u \vee |F_{v, j}| > L$}
         \State \Return \emph{invalid}
@@
 \label{def:validation}
-\textbf{\emph{Implicit consensus on transactions}}
+\textbf{\emph{Consensus on transactions}}
 
@@
         then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
+    % \item \emph{Unforgeability}:
+    %     If some transaction is valid, it cannot be forged into an invalid transaction.
+    %     If some transaction is invalid, it cannot be forged into a valid transaction.
     \item \emph{Liveness}:
         Any valid transactions can be validated eventually.
-    \item \emph{Unforgeability}:
-        If some transaction is valid, it cannot be forged into an invalid transaction.
-        If some transaction is invalid, it cannot be forged into a valid transaction.
 \end{enumerate}

### 2017-06-12 thesis/chapters/ch3-analysis.tex
More proofs

@@
 \begin{lemma}
+\label{lemma:consensus}
 \textbf{\emph{Correctness of consensus}}
+
 $\forall r \in \mathbb{N}$, agreement, validity, fairness and termination holds.
@@
 the CP blocks of the honest nodes are guaranteed to be received by at least $n - t$ facilitators.
-The agreement property of ACS (from \Cref{def:acs}) ensures that the consensus result $\C_{r + 1}$ is in consensus. 
+The agreement property of ACS (from \Cref{def:acs}) ensures that the $\C_{r + 1}$ is in consensus. 
+Observe that $\F_{r + 1}$ is computed using the deterministic function $\textsf{get\_facilitators}(\cdot)$.
+Thus agreement of $\F_{r + 1}$ follows directly from the agreement of ACS.
 The validity property of ACS ensures that $\C_{r + 1}$ contains the input of at least $n - 2t$ parties, but this is a quorum containing at least one honest facilitator,
-thus $\C_{r + 1}$ contains the CP blocks of all facilitators.
-Observe that $\F_{r + 1}$ is computed using the deterministic function $\textsf{get\_facilitators}(\cdot)$.
-Thus agreement of $\F_{r + 1}$ follows directly from agreement of ACS.
+thus $\C_{r + 1}$ contains the CP blocks of all nodes.
 Due to the assumption that the adversary cannot corrupt more than $t$ nodes,
 validity of $\F_{r + 1}$ also follows from validity of ACS.
-The fairness property follows from \Cref{lemma:fairness}.
-Finally, the termination property holds because $D$ eventually elapses and then ACS eventually terminates.
-This completes our proof.
-
+Since all honest nodes are in $\C_{r + 1}$, the fairness property follows directly from \Cref{lemma:fairness}.
+Finally, the termination property holds because $D$ eventually elapses and then ACS eventually terminates
+(the totality property of ACS).
+This completes the proof.
 \end{proof}
@@
 \subsection{Correctness of Validation}
+The consensus protocol (on CP blocks and facilitators) is the backbone for consensus on transactions.
+In this section we build on top of \Cref{lemma:consensus} to show that most properties in \Cref{def:validation} can be satisfied.
 
-Remark: chain structure cannot be satisfied, but can be probabilistically checked
+\begin{lemma}
+\textbf{\emph{Correctness of the validation protocol}}
+
+The validation protocol outputs the correct result
+according to the aforementioned validity definition.
+\end{lemma}
+\begin{proof}
+Follows directly from the definition.
+\end{proof}
+
+\begin{lemma}
+\label{lemma:validation-agreement}
+\textbf{\emph{Agreement of the validation protocol}}
+
+If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
+then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
+\end{lemma}
+\begin{proof}
+We proof by contradiction.
+Without loss of generality, for some transaction $t$ with an agreed fragment $F$,
+node $u$ decides \emph{valid} but node $v$ decides \emph{invalid}.
+Then there exist a fragment $F' = \{ \dots, t', c'\}$ which $u$ received that contains a valid pair of $t$---$t'$.
+There also exist a fragment $F'' = \{ \dots, t'', c''\}$ which $v$ received that does not contain or contains an invalid pair---$t''$.
+In both cases, the $\textsf{get\_validity}(\cdot)$ must have reached \Cref{line:valid-fragment},
+meaning that $c' = c''$, otherwise the result would be \emph{unknown}.
+Since $c' (= c'') = \langle \textsf{H}(t'), \dots \rangle$ we must have $\textsf{H}(t') = \textsf{H}(t'')$ and $t' \ne t''$ (because $t''$ is invalid).
+In other words, the sender of $F''$ must be able to create some $t''$ that has the same digest as $t'$.
+But we assumed that the adversary can only perform polynomial-time algorithm, but in order to find $t''$ it needs to query the random oracle exponentially many times.
+Thus we have a contradiction and this completes our proof.
+\end{proof}
+
+A consequence \Cref{lemma:validation-agreement} is unforgeability.
+That is, no polynomial time adversary is able to create two chains $F = \{ \dots, t, c\}$ $F' = \{ \dots, t', c\}$ with correct hash pointers and the same end of chain $c$.
+
+Now we show a negative result, where the liveness property cannot be attained.
+\begin{lemma}
+\textbf{\emph{Impossibility of liveness of the validation protocol}}
+
+There exist a valid transactions that cannot be validated eventually.
+\end{lemma}
+\begin{proof}
+We proof by providing a counterexample.
+Suppose nodes $u$ and $v$ correctly perfomed the TX protocol which resulted a transaction $t$.
+Then when $u$ wants to validate $t$, it does so by sending \texttt{vd\_req} message to $v$.
+$v$ can act maliciously and ignore all \texttt{vd\_req} message, thus $t$ can never be validated.
+\end{proof}
+Although this is a negative result, it does not put the adversary in an advantageous position.
+If the adversary is observed to ignore validation requests, then the honest nodes may prefer not to transact with her in the future.
+Thus, to stay relevant in the system, the adversary need to comply to the protocol.
+
+\subsection{Summary}
+We have shown that
+
+TODO: chain structure cannot be satisfied, but can be probabilistically checked
 

### 2017-06-13 thesis/chapters/ch3-analysis.tex
Write bandwidth

@@
 \label{lemma:fairness}
-\textbf{\emph{Fairness of facilitator election}}
-
+% (Fairness of facilitator election)
 Every node with a CP block in $\C_r$, should have an equal probability to be elected as a facilitator.
@@
 \label{lemma:consensus}
-\textbf{\emph{Correctness of consensus}}
-
+% (Correctness of consensus)
 $\forall r \in \mathbb{N}$, agreement, validity, fairness and termination holds.
@@
 \begin{lemma}
-\textbf{\emph{Correctness of the validation protocol}}
-
+% (Correctness of the validation protocol)
 The validation protocol outputs the correct result
-according to the aforementioned validity definition.
+according to the validity definition.
 \end{lemma}
 \begin{proof}
-Follows directly from the definition.
+The algorithm (\Cref{alg:get-validity}) is the validity definition.
 \end{proof}
 
-\begin{lemma}
-\label{lemma:validation-agreement}
-\textbf{\emph{Agreement of the validation protocol}}
-
+\begin{theorem}
+\label{theorem:validation-agreement}
+% (Agreement of the validation protocol)
 If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
 then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
-\end{lemma}
+\end{theorem}
 \begin{proof}
@@
 There also exist a fragment $F'' = \{ \dots, t'', c''\}$ which $v$ received that does not contain or contains an invalid pair---$t''$.
-In both cases, the $\textsf{get\_validity}(\cdot)$ must have reached \Cref{line:valid-fragment},
-meaning that $c' = c''$, otherwise the result would be \emph{unknown}.
+In both cases, the $\textsf{get\_validity}(\cdot)$ must have reached \Cref{line:valid-fragment}.
+Due to \Cref{lemma:consensus}, we have $c' = c''$, otherwise the result would be \emph{unknown}.
 Since $c' (= c'') = \langle \textsf{H}(t'), \dots \rangle$ we must have $\textsf{H}(t') = \textsf{H}(t'')$ and $t' \ne t''$ (because $t''$ is invalid).
@@
 
-A consequence \Cref{lemma:validation-agreement} is unforgeability.
+\Cref{theorem:validation-agreement} is our first major result.
+It shows that consensus on CP blocks would lead to consensus on TX blocks when the nodes are running the validation protocol.
+One of the main advantages over running a consensus algorithm on all the transactions is that 
+the rate of transaction is no longer dependent on the consensus algorithm---ACS (up to $L$ transactions).
+This enables horizontal scalability where adding new nodes would lead to higher global transaction rate.
+In addition, a convenient consequence \Cref{theorem:validation-agreement} is unforgeability.
 That is, no polynomial time adversary is able to create two chains $F = \{ \dots, t, c\}$ $F' = \{ \dots, t', c\}$ with correct hash pointers and the same end of chain $c$.
+% Another way to think about it is if a transaction can be forged, then it cannot be agreed.
 
+However, not everything is perfect.
 Now we show a negative result, where the liveness property cannot be attained.
 \begin{lemma}
-\textbf{\emph{Impossibility of liveness of the validation protocol}}
-
 There exist a valid transactions that cannot be validated eventually.
@@
 
-\subsection{Summary}
-We have shown that
-
-TODO: chain structure cannot be satisfied, but can be probabilistically checked
+% \subsection{Chain structure}
+% In this section we discuss how the adversary may tamper with the chain structure and its effects.
+% The adversary can create invalid blocks, which are blocks that have an invalid hash pointer.
+% The result is some chain with a broken link.
 
 \section{Performance}
+TODO intro
+
+\subsection{Bandwidth Per Transaction}
+To make and validate a transaction, the bandwidth required is of $O(l)$,
+where $l$ is the length of the agreed fragment.
+This can be seen from the fact that the largest message by far is the \texttt{vd\_resp} message,
+which contains the agreed fragment,
+the other messages (\texttt{tx\_req}, \texttt{tx\_resp} and \texttt{vd\_req}) are constant factors.
+If we assume that every node performs transactions at a constant rate of $c$ per second.
+Then $l = O(N) \cdot c$, where $O(N)$ is the message complexity of ACS (TODO explain this somewhere),
+thus the bandwidth per transaction is $O(N)$.
+TODO how does message complexity translate to time complexity?
+This is natural because consensus duration would be longer if there are more CP blocks, which means that the agreed fragments are longer.
+This behaviour is also verified experimentally in \Cref{ch:implementation}.
+
+\subsection{Global Throughput}
+TODO

### 2017-06-14 thesis/chapters/ch2-model.tex
Write ACS complexity

@@
 \end{itemize}
+Further, we use $a || b$ to dedote concatination of the binary representations of $a$ and $b$.
 
@@
 We assume there exist a Public Key Infrastructure (PKI), and nodes are identified by their unique and permanent public key.
-Finally, we use the random oracle model, i.e. calls to the random oracle are denoted by $\textsf{H}(\cdot)$.
+Finally, we use the random oracle model, i.e. calls to the random oracle are denoted by $\textsf{H}: \{0, 1\}^* \rightarrow \{0, 1\}^\lambda$,
+where $\{0, 1\}^*$ denotes the space of finite binary strings and $\lambda$ is the security parameter~\cite{bellare1993random}.
 

### 2017-06-14 thesis/chapters/ch3-analysis.tex
Write ACS complexity

@@
 
-\subsection{Bandwidth Per Transaction}
+\subsection{Message Complexity of ACS}
+\label{sec:acs-complexity}
+The message complexity of ACS is $O(n^2|v| + \lambda n^3 \log n)$~\cite{miller2016honey},
+where $|v|$ is the size of largest message and $\lambda$ is the security parameter.
+Note that the security parameter is the same as the one for our random oracle described in \Cref{sec:model-assumptions}.
+In particular, it is from the use of $\textsf{H}(\cdot)$ in the reliable broadcast phase in ACS.
+In our system, we wish to understand the scalability properties.
+Thus we consider the complexity as a function of $N$ rather than $n$ or $\lambda$.
+Since $|v|$ is at most all the CP blocks from every node, we have $|v| = N$.
+Therefore the message complexity of ACS in our system is $O(N)$.
+
+\subsection{Bandwidth Requirement for Transactions}
 To make and validate a transaction, the bandwidth required is of $O(l)$,
@@
 the other messages (\texttt{tx\_req}, \texttt{tx\_resp} and \texttt{vd\_req}) are constant factors.
-If we assume that every node performs transactions at a constant rate of $c$ per second.
-Then $l = O(N) \cdot c$, where $O(N)$ is the message complexity of ACS (TODO explain this somewhere),
-thus the bandwidth per transaction is $O(N)$.
-TODO how does message complexity translate to time complexity?
+If we assume that every node performs transactions at a constant rate of $r_{\text{tx}}$ per second.
+Then $l = (D_{\text{acs}} + D) \cdot r_{\text{tx}}$, where $D_{\text{acs}}$ is the duration an instance of ACS.
+But from \Cref{sec:acs-complexity}, we know that $D_{\text{acs}}$ is of $O(N)$, thus the bandwidth per transaction is $O(N)$.
+% TODO how does message complexity translate to time complexity?
+% TODO computational task minimal compared to messages, thus the sending of message is expensive?
 This is natural because consensus duration would be longer if there are more CP blocks, which means that the agreed fragments are longer.

### 2017-06-14 thesis/chapters/ch1-intro.tex
Write global throughput

@@
 
-We begin the detailed discussion by formulating the model \Cref{sec:model}. 
-Next, we analyse the correctness, security and performance of our design in \Cref{sec:analysis},
+We begin the detailed discussion by formulating the model \Cref{ch:model}. 
+Next, we analyse the correctness, security and performance of our design in \Cref{ch:analysis},
 this is where we present our theorems.
-Implementation and experimental results are discussed in \Cref{sec:implementation}.
-Finally, we compare our system with other state-of-the-art blockchain systems and conclude in Sections \ref{sec:related} and \ref{sec:conclusion} respectively.
+Implementation and experimental results are discussed in \Cref{ch:implementation}.
+Finally, we compare our system with other state-of-the-art blockchain systems and conclude in Sections \ref{ch:related} and \ref{ch:conclusion} respectively.

### 2017-06-14 thesis/chapters/ch3-analysis.tex
Write global throughput

@@
-\chapter{Analysis}
+\chapter{Fault Tolerance and Security Analysis}
 \label{ch:analysis}
@@
 
-\section{Correctness}
+\section{Correctness in the Presense of Faults}
 Our first objective in this section is to establish truths regarding the correctness of our protocol.
@@
 Therefore the message complexity of ACS in our system is $O(N)$.
+Since we use a constant $n$, $O(N)$ message complexity also holds for a single facilitator.
+
+% We argue that with a $O(N)$ message complexity we also have $O(N)$ time complexity.
+% It is know that the running time is $O(\log n)$~\cite{miller2016honey}, which translates to $O(1)$ if our $n$ is constant.
+% On the other hand, in many distributed algorithms, computational cost is small when compared to the communication cost.
+% But in our system, we are more interested in the effect of the message size.
 
@@
 \subsection{Global Throughput}
-TODO
+Suppose every node has some fixed bandwidth capacity $C$ (unit of communication per second),
+they make transactions at $r_{\text{tx}}$ per second.
+Then we have the inequality $C \ge r_{\text{tx}} l$, where $l$ is the length of the the agree fragment as before.
+Rearranging, we get $\frac{C}{l} \ge r_{\text{tx}}$.
+With this, we consider two cases, first is  when all the nodes are running at maximum capacity.
+Recall that $l$ is of $O(N)$, so as the the population increases, 
+the transaction rate must decrease in order to maintain the inverse relationship,
+thus $r_\text{tx}$ is of $O(N^{-1})$. % TODO is this formal? use omega?
+Therefore, the global throughput would be $O(N^{-1})N = O(1)$.
+In the second case, nodes are not running at maximum capacity and $r_{\text{tx}}$ is maintained.
+Therefore, if every node runs at $r_{\text{tx}}$,
+the global throughput becomes $O(N)$ until $N$ is too large and we go back to the first case.
+
+The upshot of this analysis is that our system scales nicely at a global throughput of $O(N)$ until the population gets too large.
+Then we maintain a constant global throughput.
+This result falls a bit short of what we envisioned in the introduction.
+However, the population is not the number of nodes that use the system, but the number of nodes that are online during a single round.
+Furthermore, this result is for the worst case where every transaction needs an agreed fragment to be transmitted.
+In practice, nodes are able to cache agreed fragments.
+For instance, if $u$ and $v$ make $x$ transactions in a single round,
+then only one agreed fragment need to be exchanged as it contains all the transactions rather than $x$ agreed fragments.

### 2017-06-15 thesis/chapters/ch2-model.tex
Write implementation and experiment setup

@@
 TODO give history/background of ACS
+TODO more detailed explination with algorithm and new section.
 Asynchronous common subset (ACS) is an especially useful primitive for blockchain systems.

### 2017-06-15 thesis/chapters/ch4-implementation.tex
Start analysis
Write implementation and experiment setup

--- 2017-05-15 thesis/chapters/ch3-implementation.tex
@@
+\chapter{Implementation and Experimental Results}
+\label{ch:implementation}
+
+Thus far, we have only discussed our system from a theoretical perspective.
+Henceforth, we evaluate out system experimentally and compare the results with the theoretical analysis.
+We begin this chapter by a description of the implementation in \Cref{sec:implementation}.
+Then, we move on to describing our experimental setup in \Cref{sec:experimental-setup}.
+Finally, \Cref{sec:evaluation} presents our experimental results and our evaluation.
+Our experiment primarily focuses on the consensus duration and the throughput.
+
 \section{Implementation}
-Python. Twisted.
+\label{sec:implementation}
+
+The prototype implementation is done in the event driven paradigm, using the Python programming language.
+We use Twisted as our networking library.
+The code can be found on GitHub\footnote{\url{https://github.com/kc1212/consensus-thesis-code}}.
+
+The structure of the implementation is primarily made up of three modules---\texttt{acs}, \texttt{trustchain} and \texttt{node}.
+\texttt{acs}, as its name suggests, implements ACS.
+We use liberasurecode\footnote{\url{https://github.com/openstack/liberasurecode}} as the Reed-Solomon error correcting code library, used in RBC.
+An implementation detail is that liberasurecode cannot create more than 32 code blocks\footnote{
+  The 32 code blocks limitation is hardcoded in the source file, see
+  \url{https://github.com/openstack/liberasurecode/blob/0794b31c623e4cede76d66be730719d24debcca9/include/erasurecode/erasurecode.h#L35}
+}.
+The \texttt{acs} module provides a small interface to the caller to start and stop the consensus process and also receive results.
+The \texttt{trustchain} module implements the Extended TrustChain data structure.
+It also provides the essential algorithm necessary to interact with Extended TrustChain such as 
+$\textsf{new\_tx}(\cdot)$, $\textsf{new\_cp}(\cdot)$, $\textsf{agreed\_fragment}(\cdot)$ and so on.
+Since this is a prototype implementation, we only store the data structure in memory and not on disk.
+Finally, the \texttt{node} module ties everything together.
+It implements the the consensus phase, the transaction protocol and the validation protocol.
+
+Every node keeps a persistent TCP connection with every other node.
+This creates a fully connected network for our experiment.
+It is certainly not idea in real world scenarios where nodes may have limited resources.
+But as a prototype, it is sufficient to create a system that has just over a thousand nodes, which is enough for us to experiment on.
+
+\section{Experimental Setup}
+\label{sec:experimental-setup}
+
+There are two asppects of the experimental setup.
+First is the nature of the experiment.
+Second is the physical setup.
+We describe these in turn.
+
+The nature and the goal of the experiment is to run the three protocols---consensus protocol,
+transaction protocol and validation protocol---simultaneously and analyse the throughput and consensus duration.
+There are two types of parameters which we must consider.
+First are the fixed parameters $r_{\text{tx}}$ and $D$.
+These are selected from emperical evidence, we found that $D = 30$ seconds to be more than enough for all CP blocks to reach consensus.
+We also found that using $r_{\text{tx}} = 8$ gave us good throughput without putting too much demand on the bandwidth which is must also be used for ACS.
+These parameters are fixed for all our experiments.
+The second set of parameters can be seen as the domain,
+and we run our experiment for every combination of these parameters.
+Concretely, there are three of them.
+The number of facilitators $n \ in \{ 8, 12, \dots, 32\}$.
+The population size $N \in \{100, 200, \dots, 1200\}$.
+Finally the two modes of transaction.
+The first mode or the ideal mode is that nodes only transact with their immediate neighbour. 
+This minimises the bandwidth required per validated transaction because agreed fragment can be cached.
+The second mode is in the other extreme, where  every transaction is with a random node out of the $N$ nodes in the system,
+thus the agreed fragment is unlikely to be cached.
+Unfortunately, the maximum $n$ is 32 because the limitation in liberasurecode mentioned in \Cref{sec:implementation}.
+$N$ stops at 1200 is due to our physical setup, which we describe next.
+
+The experiment is run on the DAS-5 (The Distributed ASCI Supercomputer 5).
+From now on, we use "machines" to refer to DAS-5 nodes and nodes to refer to a running instance in our system.
+On DAS-5 we use up to 24 machine, for each machine we use 50 nodes.
+This gives us the aforementioned 1200 number.
+With this setup, we cannot run more nodes because the every machine only has 65535 ports available (some of them are reserved).
+But 50 nodes each need 1200 TCP connections which is 60000 TCP connections per machine.
+Thus we set $N$ to be at most 1200\footnote{While it is possible to have many more TCP connections per machine,
+but it requires additional network interface which is something we do not control on the DAS-5.}.
+% TODO contradicts with the fully connected argument before
+
+To coordinate nodes on many different machine,
+we employ a discovery server to inform every node the IP addresses and port numbers of every other node.
+It is only run before the experiment and is not used during the experiment.
+
+\section{Evaluation}
+\label{sec:evaluation}
+
+\subsection{Consensus Duration}
+
+\subsection{Global Throughput}
 

### 2017-06-15 thesis/chapters/ch4-implementation.tex
Better figures

@@
 
+\begin{figure}[h]
+  \centering
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{result-normal-cons-vs-facilitator}
+  \end{subfigure}
+
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{result-rand-cons-vs-facilitator}
+  \end{subfigure}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{result-normal-cons-vs-pop}
+  \end{subfigure}
+
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{result-rand-cons-vs-pop}
+  \end{subfigure}
+\end{figure}
+
 \subsection{Global Throughput}
 
+\begin{figure}[h]
+  \centering
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{result-normal-throughput-vs-pop}
+  \end{subfigure}
+
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{result-rand-throughput-vs-pop}
+  \end{subfigure}
+\end{figure}
 \section{Evaluation}

### 2017-06-17 thesis/chapters/ch5-related-work.tex
Write some related work

\chapter{Related Work}
\label{ch:related}

Having analysed our system both theoretically and experimentally,
we dedicate this chapter on comparing our results with related work.
Blockchain technology has seen a surge in recent years from both the industry and academia.
There are many approaches and innovation.
We classify by their consensus approach and divide them into the following categories: 
\begin{enumerate}
    \item classical blockchain systems,
    \item Byzantine fault tolerance consensus based consensus,
    \item hybrid systems and
    \item blockchain without global consensus.
\end{enumerate}

\section{Classical Blockchain Systems}
This category represent systems with a probabilistic probabilistic consensus algorithm.
That is, transactions never reach consensus with a probability of 1.
The typical examples are proof-of-work based systems such as Bitcoin, Ethereum and other Altcoins.
In Bitcoin, the level of consensus of a block\footnote{Note that a block in Bitcoin contain many transactions whereas our TX block only contain a single transaction.}
is determined how deep it is in the Bitcoin blockchain, also called the number of confirmation.
The probability of a block being orphaned drops exponentially as the depth increases~\cite{bitcoin}.
Nevertheless, the probability of the highest block being orphaned is non-negligible.
The advantage of this type of consensus is that it can be used in a large network and is reasonably secure.
Attackers can not out pace honest users in finding new blocks unless they have a majority of the hash power.
The disadvantage however is that transactions are never in consensus with a probability of 1.
Also, the performance is limited due to the fact that blocks are of fixed size and are generated at fixed intervals.

Our system significantly improves upon Bitcoin and other classical blockchain systems in the performance, scalability and consistency department.
The results described in \Cref{ch:analysis} and \Cref{ch:implementation}, show that we have horizontal scalability,
where more nodes result in more global throughput.
Further, we do not have the aforementioned probabilistic behaviour,
once some consensus result is decided, it cannot be orphaned.
The leadership election is also not ideal in classical blockchain systems.
Mining can be seen as a technique to elect a single leader.
The leader has full control of what goes into the block thus it may selectively censor transactions.
We use ACS, so as long as the CP block is in $n - 2t$ nodes, it is guaranteed to be in consensus.

However, the security aspect falls short of the ``honest majority'' security model that classical blockchains claim to 
have\footnote{Recently it was shown that doing selfish-ming would give the adversary an unfair advantage when she only controls 25\% of the mining power~\cite{eyal2014majority}}.
Our system risks going into erroneous state if the inequality $n \ge 3t + 1$ is not satisfied.
Classical blockchain systems also have an incentive mechanism, thus they do not depend on altruistic nodes and encourages participation.
Our system on the other hand does not have an incentive mechanism because we make no assumptions on the application.

\section{Byzantine Fault Tolerance Consensus Based Systems}

The Byzantine fault tolerance problem is a classical problem is distributed systems literature.
It is first described by Leslie Lamport as the Byzantine general's problem in 1982~\cite{lamport1982byzantine}.

### 2017-06-18 thesis/chapters/ch2-model.tex
Better Byzantine consensus background

--- 2017-06-14 thesis/chapters/ch2-model.tex
@@
 \subsection{Consensus Protocol}\label{sec:overview-cons}
-Before describing our consensus protocol, we take a brief detour to explain asynchronous common subset primitive and contrast it with other well known primitives.
-It is a fundamental building block of our consensus protocol.
 
-\subsubsection*{Asynchronous common subset}
-TODO give history/background of ACS
-Asynchronous common subset (ACS) is an especially useful primitive for blockchain systems.
-It allows any party to propose a value and the result is the set union of all the proposed values by the majority.
-Concretely, ACS needs to satisfy the following properties (adapted from~\cite{miller2016honey}).
-\begin{definition}
-\label{def:acs}
-\textbf{\emph{Properties of ACS}}
+% The consensus protocol uses an existing Byzantine consensus algorithm (explained in detail in \Cref{sec:xyz}) and runs it continuously in rounds.
+The consensus protocol can be seen as a technique of running infinitly many rounds of some 
+Byzantine consensus algorithm\footnote{More accuratly it is ACS or asynchronous subset consensus, we describe ACS in \Cref{sec:acs}.},
+starting a new execution immediately after the previous one is completed.
+This is necessary because blockchain systems always need to reach consensus on new values, or CP blocks in our case.
 
-\begin{enumerate}
-    \item \emph{Validity}:
-        If any correct party outputs a set $C$,
-        then $|C| \ge n - t$ and $C$ contains the input of at least $n - 2t$ parties.
-    \item \emph{Agreement}:
-        If a correct party outputs $C$, then every party outputs $C$.
-    \item \emph{Totality}:
-        If $n - f$ party receive an input, then all correct parties produce an output.
-\end{enumerate}
-\end{definition}
-
-ACS has the nice property of censorship resiliance when compared to other consensus algorithms.
-For instance, Hyperledger and Tender mint uses Practical Byzantine Fault Tolerance (PBFT) as their consensus algorithm.
-In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectivly filter transactions.
-In contract, every party in ACS are involved in the proposal phase,
-and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
-
-The main drawback with ACS and other BFT protocols is the high message complexity.
-Typically, such protocols have the complexity of $O(n^2)$, where $n$ is the number of parties.
-Hence, it may work with a small number of nodes,
-but it is infeasible for blockchain systems where thousands of nodes are involved.
-
-% The literature on Byzantine consensus and atomic broadcast is rich, some noteable ones include TODO.
-% Thus in the ETC consensus protocol, we assume there exist an "off-the-shelf" Byzantine consensus algorithm which we can use
-% (in our implementation we use HoneyBadgerBFT and motivate our choice in TODO).
-
-% In our case, every node do not propose some arbitrary data, but a set of CP blocks.
-% Thus the result of the consensus is the set union of all the legitimate proposals.
-
-\subsubsection*{Consensus Protocol} 
-The consensus protocol runs continuously in rounds
-because a blockchain systems always need to reach consensus on new values, which are CP blocks in our case.
-This can be seen as running infinitly many rounds of some Byzantine consensus algorithm,
-starting a new execution immediately after the previous one is completed.
-
-As we mentioned earlier, the high message complexity prohibits us from running a Byzantine consensus algorithm on a large network.
+The high message complexity of Byzantine consensus algorithms prohibits us from running it on a large network.
 Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks and use them as the proposal.
 The facilitators are elected using a \emph{luck value}, which is computed using $H(\C_r || pk_i)$,
-where $\C_r$ is the consensus result in round $r$ and $pk_i$ is the public key of $i$.
+where $\C_r$ is the consensus result (a set of CP blocks) in round $r$ and $pk_i$ is the public key of $i$.
 Intuitively, the election is guaranteed to be random 
@@
 
-We proceed to describe the protocol step by step.
-Starting with the bootstrap phase and then moving on to the actual consensus phase.
+Before describing the protocol in detail, we take a brief detour to give background on the Byzantine consensus algorithm we use in the consensus protocol.
+% Starting with the bootstrap phase and then moving on to the actual consensus phase.
+
+\subsection{Background on Asynchronous Subset Consensus}
+
+The best way to explain asynchronous subset consensus (ACS) is to contrast it with the typical Byzantine consensus problem.
+We adapt the description from~\cite{podc}.
+%The Byzantine consensus problem is a classical problem is distributed systems literature.
+%It is first described by Leslie Lamport as the Byzantine general's problem in 1982~\cite{lamport1982byzantine}.
+%Byzantine consensus is described as follows (adapted from ).
+\begin{definition}
+\textbf{\emph{Byzantine Consensus}}
+
+There are $n$ nodes, of which at most $t$ might experience Byzantine fault.
+Node $i$ starts with an input value $v_i$.
+The nodes must decide for one of  those values, satisfying the the following.
+\begin{enumerate}
+    \item \emph{Agreement:}
+        All correct nodes decide for the same value.
+    \item \emph{Validity:}
+        The decision value must be the input value of a node.
+    \item \emph{Termination:}
+        All correct nodes terminate in finite time.
+\end{enumerate}
+\end{definition}
+A node under Byzantine fault means that it can have arbitrary behaviour.
+For example not sending message or colluding with other Byzantine nodes to undermine the entire system.
+Note that the decision is on a single value.
+This is in contrast to ACS which we describe next.
+
+ACS shares many similarities with Byzantine consensus.
+But it is an especially useful primitive for blockchain systems.
+It allows any party to propose a value and the result is the set union of all the proposed values by the majority.
+This is the primary difference with Byzantine consensus.
+Concretely, ACS needs to satisfy the following properties (adapted from~\cite{miller2016honey}).
+\begin{definition}
+\label{def:acs}
+\textbf{\emph{Asynchronous Subset Consensus}}
+
+There are $n$ nodes, of which at most $t$ might experience Byzantine fault.
+Node $i$ starts with an non-empty set of input values $C_i$.
+The nodes must decide an output $C$, satisfying the following.
+\begin{enumerate}
+    \item \emph{Validity:}
+        If any correct party outputs a set $C$,
+        then $|C| \ge n - t$ and $C$ contains the input of at least $n - 2t$ parties.
+    \item \emph{Agreement:}
+        If a correct party outputs $C$, then every party outputs $C$.
+    \item \emph{Totality:}
+        If $n - t$ party receive an input, then all correct parties produce an output.
+\end{enumerate}
+\end{definition}
+ACS has the nice property of censorship resiliance when compared to other consensus algorithms.
+For instance, Hyperledger and Tender mint uses Practical Byzantine Fault Tolerance (PBFT) as their consensus algorithm.
+In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectivly filter transactions.
+In contract, every party in ACS are involved in the proposal phase,
+and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
+Thus, some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
+For a detailed description of ACS we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.
+
+The main drawback with ACS and also Byzantine consensus algorithms is the high message complexity.
+Typically, such protocols have the complexity of $O(n^2)$, where $n$ is the number of parties.
+Hence, it may work with a small number of nodes,
+but it is infeasible for blockchain systems where thousands of nodes are involved.
+
+% The literature on Byzantine consensus and atomic broadcast is rich, some noteable ones include TODO.
+% Thus in the ETC consensus protocol, we assume there exist an "off-the-shelf" Byzantine consensus algorithm which we can use
+% (in our implementation we use HoneyBadgerBFT and motivate our choice in TODO).
+
+% In our case, every node do not propose some arbitrary data, but a set of CP blocks.
+% Thus the result of the consensus is the set union of all the legitimate proposals.
 

### 2017-06-18 thesis/chapters/ch5-related-work.tex
Better Byzantine consensus background

@@
     \item classical blockchain systems,
-    \item Byzantine fault tolerance consensus based consensus,
+    \item Byzantine consensus based consensus,
     \item hybrid systems and
@@
 
-\section{Byzantine Fault Tolerance Consensus Based Systems}
-
-The Byzantine fault tolerance problem is a classical problem is distributed systems literature.
-It is first described by Leslie Lamport as the Byzantine general's problem in 1982~\cite{lamport1982byzantine}.
+\section{Byzantine Consensus Based Systems}

### 2017-06-19 thesis/chapters/ch5-related-work.tex
Write more related work

@@
     \item classical blockchain systems,
-    \item Byzantine consensus based consensus,
+    \item classical blockchain with offchain transactions,
+    \item permissioned systems,
     \item hybrid systems and
-    \item blockchain without global consensus.
+    \item blockchains without global consensus.
 \end{enumerate}
@@
 Attackers can not out pace honest users in finding new blocks unless they have a majority of the hash power.
-The disadvantage however is that transactions are never in consensus with a probability of 1.
+The disadvantage however is that transactions are never in consensus with a probability of 1---no consensus finality.
 Also, the performance is limited due to the fact that blocks are of fixed size and are generated at fixed intervals.
 
-Our system significantly improves upon Bitcoin and other classical blockchain systems in the performance, scalability and consistency department.
+Our system significantly improves upon Bitcoin and other classical blockchain systems in performance, scalability and consensus finality.
 The results described in \Cref{ch:analysis} and \Cref{ch:implementation}, show that we have horizontal scalability,
@@
 Further, we do not have the aforementioned probabilistic behaviour,
-once some consensus result is decided, it cannot be orphaned.
+once some consensus result is decided, it cannot be orphaned, thus our consensus is final.
 The leadership election is also not ideal in classical blockchain systems.
@@
 
-\section{Byzantine Consensus Based Systems}
+\section{Offchain Transactions}
+Offchain transactions make use of the fact that, if two or more parties frequently make transactions,
+then it is not necessary to store every transaction on the blockchain, the only useful information is the net settlement.
+The best examples are the lightning network~\cite{lightningnetwork} and full duplex channels~\cite{decker2015fast}.
+These use the Bitcoin blockchain to store the net settlement and a payment channel to conduct offchain transactions.
+
+TODO explain payment channel
+
+\section{Permissioned Systems}
+
+This category of systems use Byzantine consensus algorithms (discussed in \Cref{sec:acs-background}).
+In essence, they contain a fixed set of nodes, sometimes called validators,
+that run a Byzantine consensus algorithm to decide on new blocks.
+This is known as a permissioned system where the validators must be pre-determined.
+Some examples include Hyperledger and Tendermint.
+The consensus algorithm used in these two systems is PBFT.
+
+A nice aspect of Byzantine consensus and in particular PBFT is that it can handle much more transactions than classical blockchain systems.
+But our system has the potential to perform beyond that of PBFT because we represent many transactions with a single CP block,
+enabling horizontal scalability.
+Furthermore, since PBFT relies on a leader, it is not censorship resiliant.
+Our system on the other hand has the benifits of ACS where CP blocks cannot be censored.
+Finally, our system is able to work in the permissionless setting by simply submitting new CP blocks to the facilitators.
+It can even be adapted to work in the permissioned by simply removing the luck value computation.
+
+What we wish to have which is in Hyperledger is a smart contract system (also known as chaincodes in Hyperledger).
+We hope to design and implement smart contracts by adding additional logic to the transaction protocol and the validation protocol.
+Such functionalities we believe is better to be built into the backbone rather than having it as an add on.
+
+\section{Hybrid Systems}
+Hybrid systems are very recent inventions.
+Just like our work, they are attempts on solving the problems of traditional blockchains.
+The main characteristic of these systems is that they use a classical blockchain technique,
+i.e. proof-of-work, to elect a committee and prevent Sybils.
+And then they use a Byzantine consensus algorithm to actually reach consensus on a set of transactions within the committee.
+Some examples are SCP~\cite{luu2015scp}, ByzCoin~\cite{kogias2016enhancing} and Solidus~\cite{abraham2016solidus}.
+
+Our approach share many similarities with the hybrid systems.
+First, we also elect a committee (facilitators) to drive consensus.
+But we do not have proof-of-work for Sybil defence because we believe 
+it is possible to do it efficiently, e.g. using NetFlow~\cite{pimotte}.
+Secondly, our use of CP blocks and ACS is also unique.
+This creates a much higher throughput and enables censorship resiliance as mentioned earlier.
+For instance, ByzCoin performs just below 1000 TPS with a thousand nodes whereas we peak at 8000 TPS.
+
+A major side effect of these hybrid systems is that they cannot guarantee correctness when there is a large number of malicious nodes.
+Our system has the same issue.
+For SCP, ByzCoin and OmniLedger, they all have some probability to elect more than $n$ Byzantine nodes into the committee.
+This problem is especially difficult solve because the committee is always much smaller than the population size which has more than $t$ Byzantine nodes,
+thus electing more than $t$ nodes into the committee in always a possibility.
+Classical blockchain do not have this problem because they do not use Byzantine consensus.
+The permissioned systems work around this problem by trusting validators.
+
+\section{Blockchains Without Global Consensus}
+
+Tangle~\cite{tangle}, Corda~\cite{corda} and the original TrustChain~\cite{trustchain} do not use global consensus at all.
+By avoiding global consensus, they are able to achieve extreme scalability.
+Just like our approach, these blockchains are also application neutral where transactions can contain arbitrary data.
+
+Our system can be considered as the same as these types of blockchains but with a lightweight consensus protocol.
+Consensus might not be applicable in all applications.
+But we believe it is important for detecting and preventing fraud.
+The example in~\Cref{fig:trustchain-bad} on page~\pageref{fig:trustchain-bad} demonstrates this.
+If $b$ makes a fork and $a$ and $c$ have no way to communicate (e.g. the adversary may control parts of the network) ,
+then $c$ is tricked to believe that her transaction with $b$ is valid.
+Only when $c$ sees a conflicting chain is she able to tell that the transaction is invalid.
+But $c$ does not know the true end-of-chain of $b$, thus she can never know whether her transaction is valid.
+This is not possible in our system because $b$ cannot convince $c$ unless he can compute exponential time algorithms
+(this is finding the second preimage for a hash function).
+
+However, consensus and validation comes at a cost that do not exist in Tangle, Corda or the original TrustChain.
+Our transaction rate is affected by the validation protocol in the worst case scenario as we saw in~\Cref{sec:evaluation}.

### 2017-06-21 thesis/chapters/ch2-model.tex
Write more related work
Change model to pure async

@@
 
-The primary goal guiding our design is scalability.
-As mentioned in the Introduction, having a scalable blockchain system while still keeping global consensus
+The goals guiding our system design are as follows.
+\begin{itemize}
+    \item Performance and scalability,
+    \item application neutrality and
+    \item security.
+\end{itemize}
+
+As mentioned in the Introduction, the primary goal is performance and scalability.
+Having a scalable blockchain system while still keeping global consensus
 allows the system to be ubiquitous and realise the full potential of blockchain.
@@
 The secondary goal is to design an application neutral system.
-In particular, it should act as a framework that provides the building blocks of blockchain based applications.
-Application developers using the framework should be able to create any application they wish.
+In particular, it should act as a backbone that provides the building blocks of blockchain based applications.
+It should be possible to buld any kind of application on top our system.
 Further, we do not impose on a consensus algorithm, as long as it satisfies the properties of atomic broadcast which we describe in~\Cref{sec:overview-cons}.
@@
 Online marketplaces such as Amazon use the rating of buyer and sellers.
-Thus it is not possible to design a Sybil defence mechanism with a a application neutral framework.
+Thus it is not possible to design a Sybil defence mechanism with a application neutral framework.
 On the other hand, our system also has no restrictions on the Sybil defence technique
-and application designers can pick the best mechanism for their application.
+and applications built on top of our system can use the best mechanism for their purpose.
 
@@
 Our system should be unaffected in the presence of powerful adversaries.
-Security is often difficult to verify, especially when it is not formalised, therefore we require our design to be provably secure.
-To summarise, our system design is designed with the following goals in mind.
-\begin{itemize}
-    \item Application neutrality,
-    \item scalability and
-    \item security.
-\end{itemize}
+In particular, adversaries are Byzantine meaning that they can have arbitrary behaviour.
+Thus anything is possible from simply ommitting messages to colluding with eachother in order to undermine the whole system.
 
@@
 \subsection{Extended TrustChain}
-Extended TrustChain naturally builds on top of TrustChain, thus we first describe the standard TrustChain.
+Extended TrustChain naturally builds on top of the standard TrustChain. 
+Thus we first describe the standard TrustChain.
 Our description has minor differences compared to the description in~\cite{trustchain}.
 This is to help with the description of the extended TrustChain.
+We remark the difference when it occurs.
 However, the two descriptions are functionally the same.
@@
 In TrustChain, every node has a ``personal'' chain. 
-Initially, the chain only contains a genesis block.
-When a node wishes to add a new transaction (TX), a new TX block is generated and is appended to the chain.
-A TX block must have a valid hash pointer pointing to the previous block
+Initially, the chain only contains a genesis block generated by the nodes themselves.
+When a node $A$ wishes to add a new transaction (TX) with $B$, a new TX block is generated and then appended to $A$'s chain.
+The TX block must have a valid hash pointer pointing to the previous block
 and a reference\footnote{This is different from the original TrustChain definition found in~\cite{trustchain}.
 In there, a TX block has two outgoing edges which are hash pointers to the two parties involved in the transaction.
-This work uses one outgoing edge and a reference.} to its \emph{pair}.
+This work uses one outgoing edge and a reference.} to its \emph{pair} on $B$'s chain.
 As a result, a single transaction generates two TX blocks, one on each party's chain.
@@
 If every node follows the rules of TrustChain and we only consider hash pointers,
-then the chain effectively forms a singly linked list.
+then every chain effectively forms a singly linked list.
 However, if a node violates the rules, then a \emph{fork} may happen.
@@
 \subsubsection*{Extended TrustChain}
-We are now ready to explain the Extended TrustChain, which we abbreviate to ETC.
-In ETC, we introduce a new type of block---checkpoint (CP) block.
+We are now ready to explain the Extended TrustChain.
+The primary difference is the introduction of a new type of block---checkpoint (CP) block.
 In contract to TX blocks, CP blocks do not store transactions or contain references.
@@
 A visual representation is shown in \Cref{fig:trustchain-bad-cp}.
+
+From this point onwards, we use TrustChain to mean the Extended TrustChain unless explicitly clarified.
 
@@
 The consensus protocol can be seen as a technique of running infinitly many rounds of some 
-Byzantine consensus algorithm\footnote{More accuratly it is ACS or asynchronous subset consensus, we describe ACS in \Cref{sec:acs}.},
+Byzantine consensus algorithm\footnote{More accuratly it is ACS or asynchronous subset consensus, we describe ACS in \Cref{sec:acs-background}.},
 starting a new execution immediately after the previous one is completed.
-This is necessary because blockchain systems always need to reach consensus on new values, or CP blocks in our case.
+This is necessary because blockchain systems always need to reach consensus on new values proposed by nodes in the system,
+or CP blocks in our case.
 
-The high message complexity of Byzantine consensus algorithms prohibits us from running it on a large network.
-Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks and use them as the proposal.
-The facilitators are elected using a \emph{luck value}, which is computed using $H(\C_r || pk_i)$,
-where $\C_r$ is the consensus result (a set of CP blocks) in round $r$ and $pk_i$ is the public key of $i$.
+The high communication complexity of Byzantine consensus algorithms prohibits us from running it on a large network.
+Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks from every other node and use them as the proposal.
+The facilitators are elected using a \emph{luck value}, which is computed using $H(\C_r || pk_u)$,
+where $\C_r$ is the consensus result 
+(which can be seen as the set union of all the CP blocks collected by the facilitators)
+in round $r$ and $pk_u$ is the public key of node $u$.
 Intuitively, the election is guaranteed to be random 
@@
 A visual explaination can be found in \Cref{app:consensus-example},
-it walks through the steps needed for a node to be selected as a facilitator.
+it walks through the steps needed for a node to be selected as a facilitator using an example.
 
@@
 When a node wish to validate one of its TX, it asks the counterparty for the \emph{fragment} of the TX.
-A fragment of a TX is a section of the chain beginning and ending with CP blocks that contains the TX.
-Upon the counterparty's response, the node checks that the CP blocks are in consensus, the hash pointers are valid and his TX is actually in the fragment.
+A fragment of a TX is a section of the chain beginning and ending with CP blocks that contains the TX and are in consensus.
+Upon the counterparty's response, the node checks that the CP blocks are indeed in consensus,
+the hash pointers are valid and his TX is actually in the fragment.
 The TX is valid if these conditions are satisfied.
-Intuitively, this works because it is hard (because hash collision is hard)
+Intuitively, this works because it is hard (because a cryptographically secure hash function is second-preimage resistant)
 to create a different chain that begins and ends with the same two CP blocks but with a different middle section.
@@
 In particular, we do not need to ``freeze'' the state of the chain for some communication to complete in order to create a block.
-For instance, a node may start the consensus protocol, and while it is running, the node may still perform transactions.
+For instance, a node may start the consensus protocol, and while it is running, create new transactions and validate old transactions.
 By the time the consensus protocol is done, the new CP block is added to whatever the state that the chain is in.
@@
 
-For notational clarify, we use the following convention (adapted from~\cite{miller2016honey}) throughout this work.
+This section and the ones following it give a technical treatment of what the content in System Overview.
+For notational clarity, we use the following convention (adapted from~\cite{miller2016honey}) throughout this work.
 \begin{itemize}
@@
 \end{itemize}
-Further, we use $a || b$ to dedote concatination of the binary representations of $a$ and $b$.
+Further, we use $a || b$ to denote concatination of the binary representations of $a$ and $b$.
 
-We assume a distributed network where nodes are fully connected,
-the channels are reliable\footnote{
+We assume purely asynchronous channels with eventual delivery.
+Thus in no circumstance do we make timing assumptions.
+The adversary has fully control of the delivery schedule and the message ordering.
+But they are not allowed to drop messages\footnote{
     Reliability can be achieved in unreliable networks by resending messages or using some error correction code.
-},
-but messages may be re-ordered and delayed by at most some time $\Delta$,
-this is sometimes known as a $\Delta$-synchronous network.
+}.
+
 We assume there exist a Public Key Infrastructure (PKI), and nodes are identified by their unique and permanent public key.
@@
 In our model we consider $N$ nodes, which is the population size.
-$n$ of them are facilitators, $t$ out of $n$ are malicious and the inequality
-$n \ge 3t + 1$ must hold.
+$n$ of them are facilitators, $t$ out of $n$ are malicious and the inequalities
+$n \ge 3t + 1$ and $N \ge n + t$ must hold.
+The first is from the FLP impossibility result TODO.
+The second is necessary due to our system design, which becomes clear in TODO.
 
+Our threat model is as follows. 
 We use a restricted version of the adaptive corruption model.
@@
 
-The primary data structure used in our system is Extended TrustChain.
+\section{Extended TrustChain}
+The primary data structure used in our system is the Extended TrustChain.
 Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
@@
 
-\subsection{TX Block}
+\subsection{Transaction Block}
 The TX block is a six-tuple, i.e $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$.
@@
 \item $\textsf{H}(b_{u, i - 1})$ is the hash pointer to the previous block.
-\item $seq_u$ is the sequence which should equal $i$.
-\item $txid$ is a cryptographically secure random number representing the transaction identifier.
-\item $pk_v$ is the public key of the counterparty.
+\item $seq_u$ is the sequence number which should equal $i$.
+\item $txid$ is the transaction identifier, it should be generated using a cryptographically secure random number generator by the initiator of the transaction.
+\item $pk_v$ is the public key of the counterparty $v$.
 \item $m$ is the transaction message.
@@
 TX blocks come in pairs.
-In particular, for every block $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$
-there exist one and only one \emph{pair} $t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), seq_v, txid, pk_u, m, sig_v \rangle$.
+In particular, for every block 
+$$t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$$
+there exist one and only one \emph{pair} 
+$$t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), seq_v, txid, pk_u, m, sig_v \rangle.$$
 Note that the $txid$ and $m$ are the same, and the public keys refer to each other.
@@
 
-\subsection{CP Block}
+\subsection{Checkpoint Block}
 
@@
 i.e. $c_{u, i} = \langle \textsf{H}(b_{u, i-1}), seq_u, \textsf{H}(\C_r), r, sig_u \rangle$,
-where $\C_r$ is the consensus result in round $r$, the other items are the same as the TX block definition.
+where $\C_r$ is the consensus result (which we describe in \Cref{sec:consensus-result}) in round $r$, the other items are the same as the TX block definition.
 Note that unlike in our prior work~\cite{implicitconsensus}, CP blocks and TX blocks do not have independent sequence numbers.
@@
 where $\textsf{H}(\bot)$ can be interpreted as applying the hash function on an empty string.
-The genesis block is unique due to every node due to $sig_u$.
+The genesis block is unique because every node has a unique public and private key pair.
 
 \subsection{Consensus Result}
+\label{sec:consensus-result}
 Our consensus protocol runs in rounds as discussed in \Cref{sec:system-overview}.
@@
 then $\langle c_{u,a}, c_{u, b} \rangle$ is the \emph{enclosure} of $t_{u, i}$.
-Some TX blocks may not have any subsequent CP blocks, then its enclosure is $\bot$.
+Note that $c_{u, a}$ is the more recent CP block.
+Also, some TX blocks may not have any subsequent CP blocks, then its enclosure is $\bot$.
 
 If the enclosure of some TX block is $\langle c_{u,a}, c_{u, b} \rangle$,
-then its \emph{fragment} is computed as $\{ b_{u, i} : a \le i \le b \}$.
+then its \emph{fragment} is defined as $\{ b_{u, i} : a \le i \le b \}$.
 For convenience, the function $\textsf{fragment}(\cdot)$ represents the fragment of some TX block if it exists, otherwise $\bot$.
@@
     then the agreed enclosure of $t_{i+2}$ is $\langle c_{i+1}, c_{i+3}\rangle$ and cannot be $\langle c_{i}, c_{i+3}\rangle$.
-Similarly, \emph{agreed fragment} is computed using the agreed enclosure.
-We define its function to be $\textsf{agreed\_fragment}(\cdot)$
+Similarly, \emph{agreed fragment} has the same definition as fragment but using agreed enclosure.
+We define its function to be $\textsf{agreed\_fragment}(\cdot)$ which we use later in the validation protocol (\Cref{sec:vd-protocol}).
 
-The length of the fragment is constrained by $L$,
-namely $\forall t |\textsf{fragment}(t)| \le L$.
-The purpose to prevent spam and encourage nodes to create more CP blocks.
-$L$ should be sufficiently high so that busy nodes are not hindered by it.
+% The length of the fragment is constrained by $L$,
+% namely $\forall t |\textsf{fragment}(t)| \le L$.
+% The purpose to prevent spam and encourage nodes to create more CP blocks.
+% $L$ should be sufficiently high so that busy nodes are not hindered by it.
 
@@
 
-Our consensus protocol runs on top of the model described above.
+Our consensus protocol runs on top of Extended TrustChain.
 It is directly related to the creation of CP blocks.
 The objectives of the protocol are to
-    allow honest nodes always make progress (by creating new CP blocks),
+    allow honest nodes always make progress (in the form of creating new CP blocks),
     compute correct consensus result in every round
     and have unbiased election of facilitators.
-Concretely, we define the necessary properties as follows,
-it is collectively known as the correctness properties of the consensus protocol.
+Concretely, we define the necessary properties as follows.
 \begin{definition}
 \label{def:consensus}
-\textbf{\emph{Correctness properties of the consensus protocol}}
+\textbf{\emph{Properties of the consensus protocol}}
 
@@
     \item \emph{Fairness}:
-        Every node should have an equal probability of becoming a facilitator.
+        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
     \item \emph{Termination}:
@@
 \end{enumerate}
+\end{definition}
+These properties may look like Byzantine consensus properties (which we describe next in \Cref{sec:acs-background}) but they have some subtle differences.
+Firstly, they are properties for every node in the network and not just the facilitators.
+Secondly, they must be satisfied for all rounds because the whole system falls apart if one of the property cannot be satisfied in one of the rounds.
 
-\end{definition}
-Note that the final three properties are on the facilitators rather than on every node,
-they are the properties ACS. 
-The fair lottery and consistent facilitator are properties on every node,
-they are prerequites for ACS.
-
-Before describing the protocol in detail, we take a brief detour to give background on the Byzantine consensus algorithm we use in the consensus protocol.
+Before describing the protocol in detail,
+we take a brief detour to give background on the asynchronous subset consensus.
+This is the primary building block of our protocol.
 % Starting with the bootstrap phase and then moving on to the actual consensus phase.
@@
 \subsection{Background on Asynchronous Subset Consensus}
+\label{sec:acs-background}
 
-The best way to explain asynchronous subset consensus (ACS) is to contrast it with the typical Byzantine consensus problem.
-We adapt the description from~\cite{podc}.
+The best way to explain asynchronous subset consensus (ACS) is to contrast it with the typical Byzantine consensus.
+We adapt the description from~\cite[Chapter 17]{podc}.
 %The Byzantine consensus problem is a classical problem is distributed systems literature.
@@
     \item \emph{Agreement:}
-        All correct nodes decide for the same value.
+        If a correct node outputs $v$, then every node outputs $v$.
     \item \emph{Validity:}
@@
 \begin{enumerate}
+    \item \emph{Agreement:}
+        If a correct node outputs $C$, then every node outputs $C$.
     \item \emph{Validity:}
-        If any correct party outputs a set $C$,
-        then $|C| \ge n - t$ and $C$ contains the input of at least $n - 2t$ parties.
-    \item \emph{Agreement:}
-        If a correct party outputs $C$, then every party outputs $C$.
+        If any correct node outputs a set $C$,
+        then $|C| \ge n - t$ and $C$ contains the input of at least $n - 2t$ nodes.
     \item \emph{Totality:}
-        If $n - t$ party receive an input, then all correct parties produce an output.
+        If $n - t$ nodes receive an input, then all correct nodes produce an output.
 \end{enumerate}
 \end{definition}
+
 ACS has the nice property of censorship resiliance when compared to other consensus algorithms.
-For instance, Hyperledger and Tender mint uses Practical Byzantine Fault Tolerance (PBFT) as their consensus algorithm.
+For instance, Hyperledger and Tendermint uses Practical Byzantine Fault Tolerance (PBFT)~\cite{castro1999practical} as their consensus algorithm.
 In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectivly filter transactions.
@@
 and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
-Thus, some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
+Thus, if some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
 For a detailed description of ACS we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.
@@
 The main drawback with ACS and also Byzantine consensus algorithms is the high message complexity.
-Typically, such protocols have the complexity of $O(n^2)$, where $n$ is the number of parties.
+Typically, such protocols have a message complexity of $O(n^2)$, where $n$ is the number of parties.
 Hence, it may work with a small number of nodes,
@@
 \label{sec:bootstrap}
+Now we have all the necessary information to describe our consensus protocol.
+We begin with the bootstrap phase and then move onto the actual consensus phase.
+
 Recall that facilitators are computed from the consensus result,
@@
 
-Imagine that there is some bootstrap oracle, that initiates the code on every node.
+To bootstrap, imagine that there is some bootstrap oracle, that initiates the code on every node.
 The code satisfied all the properties in \Cref{def:consensus}.
@@
 Note that $\F_r$ indicates the facilitators that were elected using result of round $r$ and are responsible for driving the ACS protocol in round $r + 1$.
-The goal is to reach agreement on a set of new facilitators $\F_{r + 1}$ that satisfied \Cref{def:consensus}.
+The goal is to reach agreement on a set of new facilitators $\F_{r + 1}$ that satisfies the four properties in \Cref{def:consensus}.
 
@@
 First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators.
-Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D \gg \Delta$.
+Second, if the node is a facilitator, it waits until it has received $N - t$ messages of type \texttt{cp\_msg}.
+% Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D \gg \Delta$.
 Invalid messages are removed, namely blocks with invalid signatures and duplicate blocks signed by the same key.
-After $D$ elapses, it begins the ACS protocol and some $\C_{r + 1}$ should be agreed upon by the end of it.
+With the sufficient number of \texttt{cp\_msg} messages,
+it begins the ACS algorithm and some $\C'_{r + 1}$ should be agreed upon by the end of it.
+Duplicates and blocks with invalid signatures are again removed from $\C'_{r+1}$ and we call the final result $\C_{r+1}$
+We have the remove invalid blocks a second time (after ACS) because the adversary may send different CP blocks to different facilitators,
+which results in invalid blocks in the ACS output.
 
@@
 While any ACS protocol that satisfies the standard definition will work,
-we use a simplification of HoneyBadgerBFT as our ACS protocol
+we use a simplification of HoneyBadgerBFT~\cite{miller2016honey} as our ACS protocol
 because it is the only (to the best of our knowledge) consensus algorithms designed for blockchain systems.
@@
 
-When $\F_{r}$ reaches agreement on $\C_{r+1}$,
+Continuing, when $\F_{r}$ reaches agreement on $\C_{r+1}$,
 they immediately broadcast two messages to all the nodes---
@@
 Recall that channels are not authenticated, 
-and there are no signatures in $\C_{r - 1}$.
-If a non-facilitator sees some $\C_{r - 1}$, it cannot immediately trust it because it may have been forged.
+and there are no signatures in $\C_{r + 1}$.
+If a non-facilitator sees some $\C_{r + 1}$, it cannot immediately trust it because it may have been forged.
 Thus, To guarantee authenticity, every facilitator sends an additional message that is the signature of $\C_{r + 1}$.
 
-Continuing, upon receiving $\C_{r + 1}$ and at least $n - f$ valid signatures by some $u$, $u$ performs two asks.
+Upon receiving $\C_{r + 1}$ and at least $n - f$ valid signatures by some node $u$, $u$ performs two asks.
 First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1}, r + 1)$ (\Cref{alg:new-cp}).
@@
 \end{algorithm}
+
+\subsection{Relationship with $\alpha$-Synchroniser}
 
@@
     \State $txid', pk'_u, m' \gets t_{v, j}$
-    \If{$m \ne m' \vee pk_u \ne pk'_u \vee |F_{v, j}| > L$}
+    \If{$m \ne m' \vee pk_u \ne pk'_u$}
         \State \Return \emph{invalid}
@@
 
-\section{Protocol Extensions}
-\label{sec:protocol-extensions}
+\section{Optimisations}
+\label{sec:optimisations}
 

### 2017-06-21 thesis/chapters/ch5-related-work.tex
Change model to pure async

@@
 Offchain transactions make use of the fact that, if two or more parties frequently make transactions,
-then it is not necessary to store every transaction on the blockchain, the only useful information is the net settlement.
-The best examples are the lightning network~\cite{lightningnetwork} and full duplex channels~\cite{decker2015fast}.
+then it is not necessary to store every transaction on the blockchain, only the net settlement is necessary.
+The best examples are Lightning Network~\cite{lightningnetwork} and full duplex channels~\cite{decker2015fast}.
 These use the Bitcoin blockchain to store the net settlement and a payment channel to conduct offchain transactions.
 
-TODO explain payment channel
+Offchain transaction systems are implemented using multi-signature addresses~\cite{bitcoinmultisig} and Bitcoin contracts.
+If two parties wish to make transactions, they open a timelocked payment channel with a multi-signature Bitcoin address (for two parties it would be a 2-of-2 signature address).
+Transactions happen off the Bitcoin blockchain and are signed by both parties.
+These transactions have a timestamp and only contain the net amount since the start of the payment channel.
+Before the payment channel expires, the latest of such transactions is sent to the multi-signature address.
+If multiple transactions are sent to the payment channel, the latest one is used.
+After payment channel is closed, the net transaction is propagated to the Bitcoin blockchain.
+
+The advantages of such systems is that they act as add-ons to Bitcoin wish already has a large number of users.
+Thus, if enough of the network wish for it (by setting a new block version),
+then a large number of users will instantly benefit from it.
+It also shares the advantages of Bitcoin such as security and incentives.
+
+On the other hand, offchain transactions also suffers from the problem of Bitcoin.
+Proof-of-work is still problematic as it consumes an unreasonable amount of power.
+TODO more?
 
@@
 Our system has the same issue.
-For SCP, ByzCoin and OmniLedger, they all have some probability to elect more than $n$ Byzantine nodes into the committee.
+For SCP, ByzCoin and Solidus, they all have some probability to elect more than $n$ Byzantine nodes into the committee.
 This problem is especially difficult solve because the committee is always much smaller than the population size which has more than $t$ Byzantine nodes,

### 2017-06-21 thesis/chapters/ch2-model.tex
Rewrite proof

@@
 
-$\forall r \in \mathbb{N}$, the following properties must be satisfied.
+$\forall r \in \mathbb{N}$, the following properties must hold.
 \begin{enumerate}
     \item \emph{Agreement}:
-        If one honest node outputs a list of facilitators $\F_r$,
-        every other node honest decides on $\F_r$
+        If one correct node outputs a list of facilitators $\F_r$,
+        then every node outputs $\F_r$
     \item \emph{Validity}:
-        If any honest node outputs $\F_r$, then $|\F_r| > n$ and $|\F_r|$ must contain at least $n - t$ honest nodes.
+        If any correct node outputs $\F_r$, then 
+            \begin{enumerate}
+                \item $|\C_r| \ge N - t$ must hold for the $\C_r$ which was used to create $\F_r$,
+                \item $\F_r$ must contain at least $n - t$ honest nodes and
+                \item $|\F_r| = n$.
+            \end{enumerate}
     \item \emph{Fairness}:
@@
     \item \emph{Termination}:
-        At the end of the round, every node outputs some $\F_r$.
+        Every correct node eventually outputs some $\F_r$.
 \end{enumerate}

### 2017-06-21 thesis/chapters/ch3-analysis.tex
Write implementation and experiment setup
Rewrite proof

@@
-\chapter{Fault Tolerance and Security Analysis}
+\chapter{Analysis of Correctness and Performance}
 \label{ch:analysis}
@@
 Of course, specification along does not establish any truths.
-In this chapter, we prove two aspects of our system.
-First is correctness,
-where we show that the consensus protocol and the validation protocol satisfies their desired properties
-(\Cref{def:consensus} and \Cref{def:validation} respectively).
-The second is performance,
-where we prove the lower bound of our throughput and show that it out performs classical blockchain systems.
+In this chapter, we analyse two aspects of our system.
+First we show it has the desired properties.
+That is, the properties in \Cref{def:consensus} and \Cref{def:validation} should hold.
+Then we analyse the performance, especially the throughput,
+and show that it out performs classical blockchain systems.
 
 \section{Correctness in the Presense of Faults}
-Our first objective in this section is to establish truths regarding the correctness of our protocol.
-We do this in two parts.
-First we use mathematical induction to show that properties in \Cref{def:consensus} holds for all round.
-Building on top that, if \Cref{def:consensus} is true, then we can show that many properties in \Cref{def:validation} is true.
+Our first objective is to show that \Cref{def:consensus} holds for our consensus protocol.
+Then, building on top of it, we show \Cref{def:validation} holds for the validation protocol.
+The resulting theorem shows that only using CP blocks in the consensus algorithm implies consensus on TX blocks,
+in other words, implicit consensus.
 
-\subsection{Correctness of Consensus}
+% Our first objective in this section is to establish truths regarding the correctness of our protocol.
+% We do this in two parts.
+% First we use mathematical induction to show that properties in \Cref{def:consensus} holds for all round.
+% Building on top that, if \Cref{def:consensus} is true, then we can show that many properties in \Cref{def:validation} is true.
 
-We begin our analysis by establishing the fact that the $\textsf{get\_facilitator}(\cdot)$ is fair.
+\subsection{Analysis of the Consensus Protocol}
+
+We begin our analysis by establishing truths on the four properties in \Cref{def:consensus},
+namely agreement, validity, fairness and termination.
+Using these results, we use mathematical induction to show that they hold for all rounds.
+
+\begin{lemma}
+\label{lemma:agreement}
+For an arbitrary round $r$
+if $\F_{r}$ is known by all correct nodes and one correct node outputs a list of facilitators $\F_{r+1}$,
+then all correct nodes output $\F_{r+1}$.
+\end{lemma}
+\begin{proof}
+The argument follows from the protocol description.
+Given that $\F_r$ is known,
+correct nodes will send CP blocks to all members in $\F_r$.
+The ACS algorithm starts independently whenever the facilitator has $N - t$ valid CP blocks
+(recall from \Cref{sec:consensus-phase} that invalid blocks are ones with an invalid signature or has a duplicate signature).
+It cannot make progress until $n-t$ honest facilitators start algorithm,
+but this eventually happens because there are $N - t$ correct nodes and all correct facilitator eventually receives $N -t$ valid CP blocks.
+At the end of ACS, some $\C_{r+1}$ is created, and is broadcasted along with the signature of the facilitators.
+Due to the agreement property of ACS (\Cref{def:acs}),
+every correct node should receive at least $n - t$ valid signatures on the agreed $\C_{r+1}$.
+Thus they use $\C_{r+1}$ to generate a new CP block and compute new facilitators.
+Since $\textsf{get\_facilitators}(\cdot)$ is a deterministic algorithm and the input $\C_{r+1}$ is in agreement, the output $\F_{r+1}$ is also in agreement.
+\end{proof}
+
+\begin{lemma}
+\label{lemma:validity}
+For an arbitrary round $r$,
+if $\F_r$ is known by all correct nodes and any correct node outputs $\F_{r+1}$,
+then (a) $|\C_{r+1}| \ge N - t$ must hold for the $\C_{r+1}$ which was used to create $\F_{r+1}$,
+(b) $\F_{r+1}$ must contain at least $n - t$ honest nodes and
+(c) $|\F_{r+1}| = n$.
+\end{lemma}
+\begin{proof}
+The validity follows from the validity property of ACS and the definition of our model,
+namely $N \ge n + t$ and $n \ge 3t + 1$.
+Given $\F_r$, since $N \ge n + t$ , there is at least $n$ nodes that would send their CP block to $\F_r$.
+From the validity property of ACS, we know the output must contain the input of at least $n - 2t$ nodes.
+But $n -t$ facilitators must have received $N - t$ valid CP blocks, so $|\C_{r+1}| \ge N-t$, this proves (a).
+There are $n-t$ honest nodes in $\F_{r+1}$ follows from the model, this proves (b).
+Finally, since $N-t \ge (n+t) -t = n$ and $\textsf{get\_facilitators}(\cdot)$ outputs $n$ items, $|\F_{r+1}| = n$ and this proves (c).
+\end{proof}
+
 \begin{lemma}
 \label{lemma:fairness}
-% (Fairness of facilitator election)
-Every node with a CP block in $\C_r$, should have an equal probability to be elected as a facilitator.
+For an arbitrary round $r$,
+if $\F_r$ is known by all correct nodes then
+every node with a CP block in $\C_{r+1}$, should have an equal probability to be elected as a facilitator in $\F_{r+1}$.
 \end{lemma}
 \begin{proof}
-This directly follows from the random oracle model.
-Recall that the luck value is computed using $\textsf{H}(\C_r, || pk_i)$.
-Since $pk_i$ is unique for every node that has a CP block in $\C_r$, the output of $\textsf{H}(\cdot)$ is uniformly random.
-This implies that the ordered sequence by luck value is uniformly random.
+We have already established that $|\C_{r+1}| \ge N - t \ge n$ from \Cref{lemma:validity}.
+Then the proof directly follows from the random oracle model.
+Recall that the luck value is computed using $\textsf{H}(\C_{r+1}, || pk_u)$.
+Since $pk_u$ is unique for every node that has a CP block in $\C_{r+1}$,
+the output of $\textsf{H}(\cdot)$ is uniformly random.
+This effectively generates a random permutation
+so every node has the same probability of being in the top $n$ for the ordered sequence,
+namely the output of $\textsf{get\_facilitators}(\cdot)$.
 \end{proof}
 
-Using \Cref{lemma:fairness}, we show that our consensus protocol satisfied \Cref{def:consensus}.
 \begin{lemma}
-\label{lemma:consensus}
-% (Correctness of consensus)
-$\forall r \in \mathbb{N}$, agreement, validity, fairness and termination holds.
+\label{lemma:termination}
+For an arbitrary round $r$,
+if $\F_r$ is known by all correct nodes then
+every correct node eventually outputs some $\F_{r + 1}$.
 \end{lemma}
+\begin{proof}
+This follows directly from the properties of the channel (eventual delivery)
+and the termination property of ACS.
+That is, $\F_r$ eventually receives all the CP blocks required to begin ACS.
+ACS eventually terminates.
+Finally the results are eventually dissemminated to all the nodes.
+\end{proof}
 
+From Lemmas \ref{lemma:agreement}, \ref{lemma:validity}, \ref{lemma:fairness} and \ref{lemma:termination},
+we have shown that the 4 properties of \Cref{def:consensus} holds when assuming the existance of some $\F_r$.
+Thus, to proof the whole of \Cref{def:consensus},
+we need to proof these 4 properties under the universal quantifier.
+We do this using mathematical induction.
+
+\begin{theorem}
+\label{theorem:consensus}
+For all rounds,
+the consensus protocol satisfies agreement, validity, fairness and termination (\Cref{def:consensus}).
+\end{theorem}
 \begin{proof}
-We proof by mathematical induction.
+We proof using mathematical induction.
 
-In the base case, agreement, validity fairness and termination follows directly from the bootstrap protocol, due to the bootstrap oracle.
-Note that the result is $\F_1$, which indicates the facilitators that are agreed in round 1 and are responsible for driving the ACS protocol in round 2.
+In the base case, agreement, validity fairness and termination follows directly from the bootstrap protocol,
+due to the bootstrap oracle.
+Note that the result is $\F_1$, which indicates the facilitators that are agreed in round 1, who are responsible for driving the ACS protocol in round 2.
 
-For the inductive step, we assume that the properties hold for round $r$.
-Then, to start round $r + 1$, the honest nodes begin sending CP blocks to $\F_r$. % TODO but they don't send at the same time!
-Since the honest region in $\F_r$ waits for $D$ and $D \gg \Delta$,
-the CP blocks of the honest nodes are guaranteed to be received by at least $n - t$ facilitators.
-The agreement property of ACS (from \Cref{def:acs}) ensures that the $\C_{r + 1}$ is in consensus. 
-Observe that $\F_{r + 1}$ is computed using the deterministic function $\textsf{get\_facilitators}(\cdot)$.
-Thus agreement of $\F_{r + 1}$ follows directly from the agreement of ACS.
-The validity property of ACS ensures that $\C_{r + 1}$ contains the input of at least $n - 2t$ parties, but this is a quorum containing at least one honest facilitator,
-thus $\C_{r + 1}$ contains the CP blocks of all nodes.
-Due to the assumption that the adversary cannot corrupt more than $t$ nodes,
-validity of $\F_{r + 1}$ also follows from validity of ACS.
-Since all honest nodes are in $\C_{r + 1}$, the fairness property follows directly from \Cref{lemma:fairness}.
-Finally, the termination property holds because $D$ eventually elapses and then ACS eventually terminates
-(the totality property of ACS).
-This completes the proof.
+For the inductive step,
+we assume that the 4 properties hold in round $r$ and prove that they also hold in round $r + 1$.
+Using Lemmas \ref{lemma:agreement}, \ref{lemma:validity}, \ref{lemma:fairness} and \ref{lemma:termination},
+it directly follows from Modus Poneus that these properties hold for $r + 1$.
+Due to the principals of mathematical induction, these properties hold for all $r$.
 \end{proof}
@@
 The consensus protocol (on CP blocks and facilitators) is the backbone for consensus on transactions.
-In this section we build on top of \Cref{lemma:consensus} to show that most properties in \Cref{def:validation} can be satisfied.
+In this section we build on top of \Cref{theorem:consensus} to show that most (except liveness) properties in \Cref{def:validation} can be satisfied.
 
@@
 There also exist a fragment $F'' = \{ \dots, t'', c''\}$ which $v$ received that does not contain or contains an invalid pair---$t''$.
-In both cases, the $\textsf{get\_validity}(\cdot)$ must have reached \Cref{line:valid-fragment}.
-Due to \Cref{lemma:consensus}, we have $c' = c''$, otherwise the result would be \emph{unknown}.
+In both cases, the $\textsf{get\_validity}(\cdot)$ function must have reached \Cref{line:valid-fragment}.
+Due to \Cref{theorem:consensus}, we have $c' = c''$, otherwise the result would be \emph{unknown}.
 Since $c' (= c'') = \langle \textsf{H}(t'), \dots \rangle$ we must have $\textsf{H}(t') = \textsf{H}(t'')$ and $t' \ne t''$ (because $t''$ is invalid).
-In other words, the sender of $F''$ must be able to create some $t''$ that has the same digest as $t'$.
-But we assumed that the adversary can only perform polynomial-time algorithm, but in order to find $t''$ it needs to query the random oracle exponentially many times.
-Thus we have a contradiction and this completes our proof.
+In other words, whoever sent $F''$ must be able to create some $t''$ that has the same digest as $t'$.
+But we assumed that the adversary can only perform polynomial-time algorithm,
+so in order to find $t''$ it needs to query the random oracle exponentially many times.
+Thus we have a contradiction and this completes the proof.
 \end{proof}
@@
 It shows that consensus on CP blocks would lead to consensus on TX blocks when the nodes are running the validation protocol.
+Just like in our prior work~\cite{implicitconsensus}, we call this behaviour implicit consensus.
 One of the main advantages over running a consensus algorithm on all the transactions is that 
-the rate of transaction is no longer dependent on the consensus algorithm---ACS (up to $L$ transactions).
+the rate of transaction is no longer dependent on the consensus algorithm---ACS.
 This enables horizontal scalability where adding new nodes would lead to higher global transaction rate.
@@
 
-However, not everything is perfect.
-Now we show a negative result, where the liveness property cannot be attained.
+\subsection{Impossibility of Liveness}
+While \Cref{theorem:validation-agreement} is a major result that allows significantly improved performance over traditional blockchain systems,
+it is not perfect.
+Now we show a negative result, where the liveness property of \Cref{def:validation} cannot be attained.
+Meaning that trasanctions with adversaries cannot always be validated.
+
 \begin{lemma}
@@
 then only one agreed fragment need to be exchanged as it contains all the transactions rather than $x$ agreed fragments.
+
+\section{Effect of A Highly Adverserial Environment}

### 2017-06-22 thesis/chapters/ch3-analysis.tex
Fix up performance analysis

@@
 \section{Performance}
-TODO intro
+This section aims to (analytically) answer our research question.
+That is, does the global throughput increase linearly with respect to the population size?
+We begin by looking at the communication and time complexity of ACS,
+and then the bandwidth requirement for a single transaction.
+These are prerequisites of the global throughput analysis.
 
-\subsection{Message Complexity of ACS}
+\subsection{Communication Complexity of ACS}
 \label{sec:acs-complexity}
-The message complexity of ACS is $O(n^2|v| + \lambda n^3 \log n)$~\cite{miller2016honey},
+The communication complexity of ACS is $O(n^2|v| + \lambda n^3 \log n)$~\cite{miller2016honey},
 where $|v|$ is the size of largest message and $\lambda$ is the security parameter.
@@
 Thus we consider the complexity as a function of $N$ rather than $n$ or $\lambda$.
-Since $|v|$ is at most all the CP blocks from every node, we have $|v| = N$.
-Therefore the message complexity of ACS in our system is $O(N)$.
-Since we use a constant $n$, $O(N)$ message complexity also holds for a single facilitator.
+Since $|v|$ is at most all the CP blocks from every node, we have $|v| = cN$,
+where $c$ is a constant representing the size of one CP block.
+Therefore the communication complexity of ACS in our system is $O(N)$.
+Since we use a constant $n$, $O(N)$ communication complexity also holds for a single facilitator.
 
@@
 
+\subsection{Time Complexity of ACS}
+\label{sec:acs-time-complexity}
+In order to make arguments on bandwidth or throughput, which are concepts that depend on time,
+we must make some assumptions on the time complexity of ACS.
+The time complexity we use here is not the same as what is typically used in distributed systems.
+In analysis of distributed systems, time complexity is often in number of rounds.
+For example, ACS runs in a constant number of rounds because its subprotocols---reliable broadcast and binary Byzantine consensus---also run in a constant number of rounds.
+However, in practice, making a unit of communication always has some overhead associated with it, for example serialising and writing it to some network socket.
+Hence, for the remainder of our performance analysis, add the following to our computational model.
+For every unit of communication, we assume they take some non-negligible but constant time to perform.
+From our result in \Cref{sec:acs-complexity} and the fact that ACS runs in a constant number of rounds,
+it follows that ACS has a time complexity of $O(N)$.
+
 \subsection{Bandwidth Requirement for Transactions}
-To make and validate a transaction, the bandwidth required is of $O(l)$,
+With our assumption on the time complexity of ACS,
+we are ready to analyse the bandwidth.
+We know that to create and then validate a transaction, the bandwidth required is of $O(l)$,
 where $l$ is the length of the agreed fragment.
@@
 If we assume that every node performs transactions at a constant rate of $r_{\text{tx}}$ per second.
-Then $l = (D_{\text{acs}} + D) \cdot r_{\text{tx}}$, where $D_{\text{acs}}$ is the duration an instance of ACS.
-But from \Cref{sec:acs-complexity}, we know that $D_{\text{acs}}$ is of $O(N)$, thus the bandwidth per transaction is $O(N)$.
+Then
+$$l = r_{\text{tx}} D_{\text{acs}},$$
+where $D_{\text{acs}}$ is the duration an instance of ACS.
+But from \Cref{sec:acs-time-complexity},
+we know that $D_{\text{acs}}$ is of $O(N)$, thus the bandwidth per transaction is $O(N)$.
 % TODO how does message complexity translate to time complexity?
 % TODO computational task minimal compared to messages, thus the sending of message is expensive?
-This is natural because consensus duration would be longer if there are more CP blocks, which means that the agreed fragments are longer.
+This is intuitive because consensus duration would be longer if there are more CP blocks, which means that the agreed fragments are longer (assuming nodes transact at a constant rate).
 This behaviour is also verified experimentally in \Cref{ch:implementation}.
@@
 \subsection{Global Throughput}
+Using our results so far, we are able to analyse the global throughput.
 Suppose every node has some fixed bandwidth capacity $C$ (unit of communication per second),
@@
 Then we have the inequality $C \ge r_{\text{tx}} l$, where $l$ is the length of the the agree fragment as before.
-Rearranging, we get $\frac{C}{l} \ge r_{\text{tx}}$.
-With this, we consider two cases, first is  when all the nodes are running at maximum capacity.
+Rearranging, we get
+$$\frac{C}{l} \ge r_{\text{tx}}.$$
+With this, we consider two cases, first is  when all the nodes are running at maximum capacity, i.e. $\frac{C}{l} = r_{\text{tx}}$.
 Recall that $l$ is of $O(N)$, so as the the population increases, 
@@
 the global throughput becomes $O(N)$ until $N$ is too large and we go back to the first case.
+A visualisation of the result is shown in \Cref{fig:throughput-visualisation}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.7\textwidth]{throughput-visualisation}
+  \caption{Visualisation of the worst case global throughput.}
+  \label{fig:throughput-visualisation}
+\end{figure}
 

### 2017-06-25 thesis/chapters/ch1-intro.tex
New intro and etc.

--- 2017-05-29 thesis/chapters/ch1-intro.tex
@@
-TODO some positive intro about blockchain
+\chapter{Introduction}
+\label{ch:intro}
 
-One of the key issues in many blockchain systems today is that they are not scalable.
-Bitcoin~\cite{bitcoin}, the largest permissionless\footnote{Explain permissionless} blockchain system
-in terms of market capitalisation~\cite{bitcoinmarketcap} has a maximum transaction rate of merely 7 transaction per second (TX/s).
-This is due to the consensus mechanism in Bitcoin, namely proof-of-work (PoW),
-miners can only create new blocks every 10 minutes and every block cannot be larger than 1 megabyte.
-Payment processors in use today such as Visa can handle transaction rates in the order of thousands~\cite{visa}.
-While Bitcoin may be a revolutionary phenomenon, it clearly cannot be ubiquitous in its current state.
+We live in a world where technologies have become vital for our welfare and success.
+The internet, for instance, gave us the ability of efficiently exchange information on a global scale.
+However, in contrast to its original design, the internet and in particular the world wide web is becoming increasingly centralised.
+The domain name system, certificate authorities, to name a few, carry enormous responsibilities and are a central point of failure.
+The  same can be said for many other services such are online marketplaces,
+cloud services, hospitality services and even our banking system.
+The 2008 financial crisis is an example of the banking system making poor choices which result in a
+decline in consumer wealth in the order of trillions~\cite{financialcrisis} and led to the European debt crisis.
 
-An different approach is to not reach global consensus at all.
-For instance in TrustChain~\cite{multichain} and Tangle~\cite{tangle},
-nodes in the network only store their personal ledger.
-Since consensus is left out, nodes can perform transactions as fast as their machine and network allows.
-The downside of this approach is that it cannot prevent fraud (it is possible to detect fraud).
-To examplify, a malicious node Mallory may claim she has 3 units of currency to Alice,
-but in reality Mallory already spent all of it on Bob.
-If there is no global consensus and Bob and Alice never communicate,
-then the 3 units that Alice is about to receive is nonexistent.
+Ironically, also in 2008, Satoshi Nakamoto published the Bitcoin whitepaper~\cite{bitcoin}.
+Which, for the first time, gave us a simple banking system in the form of an distributed ledger.
+It needs no central control but still incorruptable even if there are malicious parties that aim to undermine the system.
+We call such a ledger a blockchain.
 
-The scalability property of TrustChain and Tangle are exceptionally desirable.
-The global consensus mechanism of Bitcoin and many other blockchain systems are also worthwhile for detecting or preventing fraud.
-These two properties may seem mutually exclusive, but in this work, we demonstrate the opposite.
-Specifically, we answer the following research question in the affirmative.
-\emph{Is it possible to design a blockchain fabric that can reach global consensus on the state of the system 
-and also scalable?} We define scalability as a property where if more nodes join the system, then the transaction rate should increases.
-% this is not completely true, there's a cap
+The primary innovations are (1) its consensus model, which prevents double spending,
+and (2) its incentive mechanism that encourages anyone (with adaquate hardware) to participate in the network and keep it running.
+The double spending problem can be seen as an inconsistency issue.
+For example, $C$ has 5 units of currency in her account.
+At some instant in time, $A$ sees that $C$ transferred 5 units to $A$ and $B$ sees that $C$ transferred 5 units to $B$.
+Hence, the transaction is inconsistent with respect to the observer.
+Bitcoin and many other blockchain systems solve the inconsistency problem with a consensus algorithm.
+The goal of the algorithm is reach agreement on a set of values, e.g. transactions, which eliminates inconsistencies.
+In Bitcoin's case, the consensus algorithm is called \emph{proof-of-work}.
+Where miners (parties that runs the Bitcoin network) collect transactions and compete in solving (using brute force) a puzzle.
+The first miner to solve it generates a block containing all the collected transactions.
+The miner is also rewarded with new coins and transaction fees.
+It is important to note that every block, that is the solution of the puzzle,
+depends on the previous block.
+Hence the name blockchain.
+Further, honest miners always work on the largest chain.
+Due to the difficulty of the puzzle,
+it is unlikely for more than one blockchain to exist in the network for a long period of time.
+Thus every party sees a consistent blockchain which solves the double spending problem.
+Further, due to the reward mechanism,
+people are incentivised to act as miners and keep the system running.
 
+Bitcoin has had its ups and downs, but over it has grown into an enormous system.
+Its power consumption is as high as Republic of Ireland~\cite{o2014bitcoin}.
+Its market cap, at the time of writing, is over 40 Trillion USD~\cite{bitcoinmarketcap}.
+Many online marketplaces are using Bitcoin, for example Steam\footnote{\url{https://store.steampowered.com/}} and even Amazon\footnote{Not directly, but via \url{https://purse.io/}.}.
+Due to its success, people from many different disciplines began investigating in way to use blockchain technology.
+This includes finance~\cite{todo}, heath care~\cite{todo}, logistics~\cite{todo}, energy~\cite{todo} and so on.
+
+Sadly, as traditional blockchain systems began to gain popularity,
+their limitations also became apparent.
+Bitcoin has the infamous 7 transactions per second upper bound.
+This is due to the fact that blocks are fixed to 1 MB are only generated on average every 10 minutes.
+Since every Bitcoin transaction is at least 250 bytes, it computes to about 7 transactions per second.
+Due to this limitation, it is not uncommon to see a long backlog of about 20000 transactions on \url{https://blockchain.info}\footnote{\url{https://blockchain.info/unconfirmed-transactions/}}.
+A few months ago the backlog even reached 100000, which meant new transactions would take at least 11 hours to hit the blockchain~\cite{bitcoinbacklog}.
+The issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some calls it a civil war~\cite{bitcoincivilwar}.
+Parties that are for the increase in block size argue that a larger block would improve the transaction rate.
+Parties against it argue that it would make mining more centralised because blocks take longer to propogate through the network and it does not solve the fundamental problem.
+A recent empirical study Croman et el.~\cite{croman2016scaling} has shown that simply increasing the block size may help.
+But given the bandwidth and latency constraints, it not possible to have more than 758 transactions per second and it may give some miner an unfair lead over others.
+Thus fundamental changes are necessary run at the scale of centralised payment processors such as Visa~\cite{visa}, which is in the order of tens of thousands transactions per second.
+
+In this work, we take the advice of Croman et el.~\cite{croman2016scaling} and Marko~\cite{vukolic2015quest} and rethink the blockchain architecture.
 Our primary insight came from observing the differences between how transactional systems work in the real world and how they work in blockchain systems like Bitcoin. 
@@
 
-Using the aforementioned insight,
-we explore an alternative consensus model for blockchain systems where transactions themself do not reach consensus,
-but nevertheless verifiable at a later stage by any node in the network.
-Informally, out model works as follows.
-Every node stores its own blockchain and every block is one transaction, same as the TrustChain construction.
-We randomly selected nodes in every round, the selected ones are called facilitators.
-They reach consensus not on the individual transactions,
-but on the state of every chain represented by a single digest, we call this state the checkpoint.
-If a checkpoint of some node is in consensus, 
-then that node can proof to any other node that it holds a set of transactions that computes (form a chain) to the checkpoint.
-This immediately show that those transactions are tamper-proof.
+We describe and analyse the aforementioned idea in the remainder of this work.
+We begin with the problem description in~\Cref{ch:problem}.
+This is followed by a detailed formulation of the model in~\Cref{ch:model}. 
+Next, we analyse the correctness and performance of our design in~\Cref{ch:analysis}, this is where we present our key theorems.
+Implementation and experimental results are discussed in~\Cref{ch:implementation}.
+Finally, we compare our system with other state-of-the-art blockchain systems and conclude in Chapters \ref{ch:related} and \ref{ch:conclusion} respectively.
 
-We begin the detailed discussion by formulating the model \Cref{sec:model}. 
-Next, we analyse the correctness, security and performance of our design in \Cref{sec:analysis},
-this is where we present our theorems.
-Implementation and experimental results are discussed in \Cref{sec:implementation}.
-Finally, we compare our system with other state-of-the-art blockchain systems and conclude in Sections \ref{sec:related} and \ref{sec:conclusion} respectively.
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: "../thesis"
-%%% End:
+% However, not everyone is convinced on the practicality of blockchain systems.
+% Bram Cohen, the inventor of BitTorrent said
+% ``It's getting a lot more media attention than the actual impact it's having so far.''~\cite{cohen}
+% Forbes listed a few reasons on why we should be skeptical about blockchains~\cite{forbes},
+% some of them include its accessibility for the layperson and its scalability issues.

### 2017-06-25 thesis/chapters/ch2-problem.tex
New intro and etc.

\chapter{Problem Description}
\label{ch:problem}

- long history of BFT, since Lamport
- PBFT
- PBFT triggered a renaissance in BFT replication research, with protocols like Q/U

- classical blockchains are nice, but is it possible to also use 30 years of BFT research?
- non-consensus chains scales, but no consensus...
- alternative consensus model

% Scalability is indeed one of the key challenges we face in blockchain systems today.
% Bitcoin~\cite{bitcoin}, the largest permissionless\footnote{Explain permissionless} blockchain system
% in terms of market capitalisation~\cite{bitcoinmarketcap} has a maximum transaction rate of merely 7 transaction per second (TX/s).
% This is due to the consensus mechanism in Bitcoin, namely proof-of-work (PoW),
% miners can only create new blocks every 10 minutes and every block cannot be larger than 1 megabyte.
% Payment processors in use today such as Visa can handle transaction rates in the order of thousands~\cite{visa}.
% While Bitcoin may be a revolutionary phenomenon, it clearly cannot be ubiquitous in its current state.

% An different approach is to not reach global consensus at all.
% For instance in TrustChain~\cite{multichain} and Tangle~\cite{tangle},
% nodes in the network only store their personal ledger.
% Since consensus is left out, nodes can perform transactions as fast as their machine and network allows.
% The downside of this approach is that it cannot prevent fraud (it is possible to detect fraud).
% To examplify, a malicious node Mallory may claim she has 3 units of currency to Alice,
% but in reality Mallory already spent all of it on Bob.
% If there is no global consensus and Bob and Alice never communicate,
% then the 3 units that Alice is about to receive is nonexistent.

% The scalability property of TrustChain and Tangle are exceptionally desirable.
% The global consensus mechanism of Bitcoin and many other blockchain systems are also worthwhile for detecting or preventing fraud.
% These two properties may seem mutually exclusive, but in this work, we demonstrate the opposite.
% Specifically, we answer the following research question in the affirmative.
% \emph{Is it possible to design a blockchain fabric that can reach global consensus on the state of the system 
% and also scalable?} We define scalability as a property where if more nodes join the system, then the transaction rate should increases.
% % this is not completely true, there's a cap

% Using the aforementioned insight,
% we explore an alternative consensus model for blockchain systems where transactions themself do not reach consensus,
% but nevertheless verifiable at a later stage by any node in the network.
% Informally, out model works as follows.
% Every node stores its own blockchain and every block is one transaction, same as the TrustChain construction.
% We randomly selected nodes in every round, the selected ones are called facilitators.
% They reach consensus not on the individual transactions,
% but on the state of every chain represented by a single digest, we call this state the checkpoint.
% If a checkpoint of some node is in consensus, 
% then that node can proof to any other node that it holds a set of transactions that computes (form a chain) to the checkpoint.
% This immediately show that those transactions are tamper-proof.

### 2017-06-25 thesis/chapters/ch4-analysis.tex
New intro and etc.

--- 2017-06-22 thesis/chapters/ch3-analysis.tex
@@
 \section{Effect of A Highly Adverserial Environment}
+
+Our last study consideres the effect when the number of adversaries is more than $t$.
+This is useful because in practice it is difficult to guarantee that $t$ satisfied $n \ge 3t + 1$, especially when $N$ is large.
+Hence we are interested in the probability for this to happen under our facilitator election process.
+
+The problem can be formulated as follows.
+Suppose an urn contains $N$ balls, $t$ are black and $N-t$ are white.
+If $n$ balls are drawn uniformly at random without replacement,
+what is the probability that $\lfloor \frac{n-1}{3} \rfloor$ are black?
+The random variable $X$ in this case is the number of black balls, or the number of successful events.
+It follows the hypergeometric distribution since we pick balls \emph{without} replacement~\cite{skala2013hypergeometric}.
+Hence, we are interested in the following probability.
+$$
+1 - \sum_{k = 0}^{\lfloor \frac{n-1}{3} \rfloor} \Pr[X = k] = 
+1 - \sum_{k = 0}^{\lfloor \frac{n-1}{3} \rfloor} \frac{ \binom{t}{k} \binom{N-t}{n-k} }{ \binom{N}{n} }
+$$
+
+This is not in closed form,
+but we can visualise the effect in \Cref{fig:hypergeometric}.
+We set the population size $N$ to 2000
+and plot the probability of more than $\lfloor \frac{n-1}{3} \rfloor$ successful events for different numbers of draws.
+Evidently,
+if the number of black balls (traitors) is a third of the population (666 out of 2000)
+we have about 0.5 probability of electing more than $\lfloor \frac{n-1}{3} \rfloor$ black balls for sufficiently large $n$.
+Thus, we cannot expect the system to function correctly when 
+the expected value is close to the number of black balls that we can tolerate.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{hypergeometric}
+  \caption{Plot of the probability of selecting more than $\lfloor \frac{n-1}{3} \rfloor$ black balls for
+  different numbers of }
+  \label{fig:hypergeometric}
+\end{figure}
+
+On the other hand, due to the fact that hypergeometric distributions have light tails,
+that is ``faster-than-exponential fall-off''~\cite{skala2013hypergeometric},
+the probability for picking more than $\lfloor \frac{n-1}{3} \rfloor$ black balls when 
+the expected value is  much smaller than $\lfloor \frac{n-1}{3} \rfloor$ is small.
+For example, we can use tail inequality to bound the probability of picking more than $\lfloor \frac{n-1}{3} \rfloor$ black balls when only 5\% are black.
+The tail inequality is
+$$
+\Pr[X \ge E[X] + \tau n] \le e^{-2\tau^2n},
+$$
+where $E[X] = n/20$.
+We are interested in $\Pr[X \ge \lfloor \frac{n-1}{3} \rfloor + 1]$, so
+\begin{align*}
+\tau &= \frac{\lfloor \frac{n-1}{3} \rfloor + 1}{n} - \frac{1}{20} \\
+  &> \frac{\lfloor \frac{n-1}{3} \rfloor}{n} - \frac{1}{20} \\
+  &\ge \frac{1}{4} - \frac{1}{20} = \frac{1}{5} \quad \text{for $n \ge 4$}.
+\end{align*}
+Putting $\tau$ back into the tail inequality we get
+$$
+\Pr[X \ge \lfloor \frac{n-1}{3} \rfloor + 1] \le e^{-\frac{2}{25} n}.
+$$
+The bound is not tight, but it is clear that for sufficiently large $n$, the probability becomes vanishingly small.
+Hence, it is possible to use our system in practice when 5\% of the population are adversarial but $n \ge 3t + 1$ does not hold,
+as long as $n$ is large.

### 2017-06-25 thesis/chapters/ch6-related-work.tex
New intro and etc.

--- 2017-06-21 thesis/chapters/ch5-related-work.tex
@@
 Blockchain technology has seen a surge in recent years from both the industry and academia.
-There are many approaches and innovation.
-We classify by their consensus approach and divide them into the following categories: 
+We classify the various blockchain systems by their consensus approach and divide them into the following categories: 
 \begin{enumerate}
@@
 \end{enumerate}
+A few systems from each of these categories are compared with our design.
 
@@
 
-Offchain transaction systems are implemented using multi-signature addresses~\cite{bitcoinmultisig} and Bitcoin contracts.
-If two parties wish to make transactions, they open a timelocked payment channel with a multi-signature Bitcoin address (for two parties it would be a 2-of-2 signature address).
+Offchain transaction systems are implemented using multi-signature addresses~\cite{bitcoinmultisig} and hashed time-locked Bitcoin contracts~\cite{bitcointimelock}.
+If two parties wish to make transactions, they open a time-locked payment channel with a multi-signature Bitcoin address (for two parties it would be a 2-of-2 signature address).
 Transactions happen off the Bitcoin blockchain and are signed by both parties.
@@
 Proof-of-work is still problematic as it consumes an unreasonable amount of power.
-TODO more?
+Further, sidechain transactions are limited only to an exchange of cryptocurrency,
+it is less general than typical Bitcoin transactions which may be a simple smart contract.
+Time-locked contracts have a strong dependency on timing, thus disputes may arise when the payment channel is just about to close.
+Our system is purely asynchronous and make no assumption on timing,
+in fact we assume the adversary has control of the message delivery time and order.
 

### 2017-06-26 thesis/chapters/ch1-intro.tex
Write problem description

@@
 Its power consumption is as high as Republic of Ireland~\cite{o2014bitcoin}.
-Its market cap, at the time of writing, is over 40 Trillion USD~\cite{bitcoinmarketcap}.
+Its market cap, at the time of writing, is over 40 billion USD~\cite{bitcoinmarketcap}.
 Many online marketplaces are using Bitcoin, for example Steam\footnote{\url{https://store.steampowered.com/}} and even Amazon\footnote{Not directly, but via \url{https://purse.io/}.}.
@@
 Since every Bitcoin transaction is at least 250 bytes, it computes to about 7 transactions per second.
-Due to this limitation, it is not uncommon to see a long backlog of about 20000 transactions on \url{https://blockchain.info}\footnote{\url{https://blockchain.info/unconfirmed-transactions/}}.
+Due to this limitation, it is not uncommon to see a long backlog of about 20000 transactions\footnote{\url{https://blockchain.info/unconfirmed-transactions/}}.
 A few months ago the backlog even reached 100000, which meant new transactions would take at least 11 hours to hit the blockchain~\cite{bitcoinbacklog}.
@@
 Parties that are for the increase in block size argue that a larger block would improve the transaction rate.
-Parties against it argue that it would make mining more centralised because blocks take longer to propogate through the network and it does not solve the fundamental problem.
-A recent empirical study Croman et el.~\cite{croman2016scaling} has shown that simply increasing the block size may help.
-But given the bandwidth and latency constraints, it not possible to have more than 758 transactions per second and it may give some miner an unfair lead over others.
-Thus fundamental changes are necessary run at the scale of centralised payment processors such as Visa~\cite{visa}, which is in the order of tens of thousands transactions per second.
+Parties against it argue that it would make mining more centralised because blocks take longer to propogate through the network.
+It also requires a hard fork (not backward compatible), which risks consensus failure and devaluation.
+A recent empirical study Croman et el.~\cite{croman2016scaling} has shown that increasing the block size may help.
+But given the bandwidth and latency constraints,
+it not possible to have more than 758 transactions per second.
+Even worse, it may give some miner an unfair lead over others.
+Thus fundamental changes are necessary run at the scale of centralised payment processors such as Visa~\cite{visa},
+which is in the order of tens of thousands transactions per second.
+
+What we have described is in fact the current state of Bitcoin.
+Many proposals exist with the aim to make Bitcoin more scalable.
+The most prominant is Seggregated Witness or SegWit~\cite{segwit}.
+It moves the signature data to the end of the block and changes the block size to 4 MB.
+The effect is that the first 1 MB of the block is still backward compatible, thus there is no need for a hard fork.
+It also solves the transaction malleability problem~\cite{bitcoinmalleability} and paves the way for offchain transactions via Lightning Network~\cite{lightningnetwork}.
+Although SegWit is implemented, it is still uncertain whether the Bitcoin network will adopt it.
+At the time of writing, there is only 35\% of the network signalling for adoptation,
+this is only a 10\% increase since November last year\footnote{\url{https://blockchain.info/charts/bip-9-segwit}}.
+Thus it may take years if ever for SegWit to be adopted.
+Even if it is adopted, it cannot reach the transaction rate of centralised systems.
 
@@
 This consequently lead to limited scalability because every transaction must be validated by the authority even when most of the transaction are legitimate.
+
+Using the aforementioned insight,
+we explore an alternative consensus model for blockchain systems where transactions themself do not reach consensus,
+but nevertheless verifiable at a later stage by any node in the network.
+Informally, out model works as follows.
+Every node stores its own blockchain and every block is one transaction.
+We randomly selected nodes in every round, the selected ones are called facilitators.
+They reach consensus not on the individual transactions,
+but on the state of every chain represented by a single digest, we call this state the checkpoint.
+If a checkpoint of some node is in consensus, 
+then that node can proof to any other node that it holds a set of transactions that computes (form a chain) to the checkpoint.
+This immediately show that those transactions are tamper-proof.
 

### 2017-06-26 thesis/chapters/ch2-problem.tex
Write problem description

@@
 
-- long history of BFT, since Lamport
-- PBFT
-- PBFT triggered a renaissance in BFT replication research, with protocols like Q/U
+Blockchain technology has many advantages over centralised systems.
+Its decentralised and democratic nature removes central point of control or failure.
+However, as we described in the Introduction, it faces a genuine scalability problem.
+In particular, proof-of-work based systems are unlikely to reach the transaction rate of Visa even with parameter tuning.
+What we wish to see is horizontal scalability.
+That is, by adding new nodes into the network, the global transaction rate should increase proportionally.
+Hence, the research question we answer in this  work is the following.
+\begin{displayquote}
+\emph{Is it possible to design a blockchain fabric that is fault tolerant, scalable and can reach global consensus?}
+\end{displayquote}
 
-- classical blockchains are nice, but is it possible to also use 30 years of BFT research?
-- non-consensus chains scales, but no consensus...
-- alternative consensus model
+In order to make sense of our research question, we must first define \emph{blockchain fabric}.
+Typical blockchain systems are application specific.
+For example, the application in Bitcoin (any many others) is digital cash, the application in Namecoin is domain name registration~\cite{namecoin} and the application in Siacoin is cloud storage~\cite{siacoin}.
+Underneath these applications lies the blockchain fabric, which we call the consensus protocol, and it is abstract over the application.
+In the case of Bitcoin it is proof-of-work (PoW) because PoW is not concerned with the structure of the transactions,
+it works by simply treating transactions as a blob of data and hashes them with the previous block.
+Hence, we wish to design an application neutral system.
+It should act as a backbone that provides the building blocks of any blockchain based applications.
+Also, it should be possible to build any kind of application on top our system.
+We do not impose on a consensus algorithm, as long as it satisfies the properties of atomic broadcast which we describe in~\Cref{sec:overview-cons}.
+
+In the remainder of this chapter we expand on our research question and visit each of our requirement in detail and discuss our approach.
+
+\section{Global Consensus}
+
+The first objective is to reach consensus on a global \emph{state}.
+We emphasise the word state because in contrary to traditional blockchains,
+it does not neccessarily mean the set of all transactions that has ever happened.
+The state can be a verifiable representation of all the transactions.
+As an extreme example, a single digest of the complete history can represent the state.
+It is even verifiable as long as there are nodes that stores the history.
+
+Traditional blockchain systems use PoW as the consensus algorithm.
+However, consensus algorithm and especially Byzantine\footnote{Byzantine means participants can experience arbitrary failure, more on this in~\Cref{sec:acs-background}.}
+consensus algorithms have existece since 1982,
+where it is first introduced by Lamport et el.~\cite{lamport1982byzantine}.
+The Practical Byzantine Fault Tolerance (PBFT) algorithm introduced in 1999 by Castro and Liskov~\cite{castro1999practical} is widely regarded as the first high-performance consensus algorithm.
+Since then research in this area picked up pace and many more algorithms are introduced~\cite{abd2005fault, kotla2007zyzzyva}.
+With the introduction of Bitcoin's probabilistic consensus model, the field experienced a second renaissance.
+These include proof-of-stake protocols~\cite{bentov2016snow, micali2016algorand}, as well as protocols optimised for blockchain~\cite{liu2016xft, miller2016honey}.
+
+Our approach is to combine the vast literature in Byzantine fault tolerance research with blockchain technology,
+and introduce a new consensus model to record the global state.
+
+\section{Security and Fault Tolerance}
+
+The second objective is fault tolerance,
+where our system should be unaffected in the presence of powerful adversaries.
+In particular, adversaries are Byzantine meaning that they can have arbitrary behaviour.
+Thus anything is possible from simply ommitting messages to colluding with eachother in order to undermine the whole system.
+This asppect usually comes for free when using a Byzantine fault tolerant consensus algorithm.
+
+There is another aspect of fault tolerance that is transaction verifiability.
+Recall that it is not neccessary to reach consensus on all the transactions in order to record the state.
+But when using some digest (or a set of digets) to represent the state, it is also important for the transactions to be verifiable.
+Otherwise adversaries could simply store invalid transactions in the state and they would never be detected.
+
+\section{Performance and Scalability}
+
+The last but not least, the final objective is horizontal scalability.
+We wish to see the performance (global transaction rate) increase as more nodes join the network.
+BitTorrent~\cite{cohen2003incentives} is an example of such a system,
+where peers interact with eachother to exchange files without any global bottleneck.
+
+Traditional blockchain systems may run in a network of thousands of nodes, but the performance do not scale.
+On the other hand, as we will see in~\Cref{sec:acs-background}, Byzantine consensus algorithms performs well but only when in a small network.
+In this work we combine the ideas from both of these systems to create the first horizontally scalable blockchain fabric.
+
+\section{Non-goals}
+Due to the nature of our system, we do not explicitly address the Sybil attack~\cite{douceur2002sybil}.
+Sybil defence mechanism always require some form of reputation score from the application.
+For example, social network based Sybil defence mechanims use graph structure of real-world relationships~\cite{yu2006sybilguard}.
+Online marketplaces such as Amazon use the rating of buyer and sellers.
+Thus it is not possible to design a Sybil defence mechanism with a application neutral framework.
+On the other hand, our system also has no restrictions on the Sybil defence technique
+and applications built on top of our system can use the best mechanism for their purpose.
 
@@
 % % this is not completely true, there's a cap
-
-% Using the aforementioned insight,
-% we explore an alternative consensus model for blockchain systems where transactions themself do not reach consensus,
-% but nevertheless verifiable at a later stage by any node in the network.
-% Informally, out model works as follows.
-% Every node stores its own blockchain and every block is one transaction, same as the TrustChain construction.
-% We randomly selected nodes in every round, the selected ones are called facilitators.
-% They reach consensus not on the individual transactions,
-% but on the state of every chain represented by a single digest, we call this state the checkpoint.
-% If a checkpoint of some node is in consensus, 
-% then that node can proof to any other node that it holds a set of transactions that computes (form a chain) to the checkpoint.
-% This immediately show that those transactions are tamper-proof.

### 2017-06-26 thesis/chapters/ch3-model.tex
New intro and etc.
Write problem description

--- 2017-06-21 thesis/chapters/ch2-model.tex
@@
 
-The goals guiding our system design are as follows.
-\begin{itemize}
-    \item Performance and scalability,
-    \item application neutrality and
-    \item security.
-\end{itemize}
+% As was shown in~\Cref{ch:problem}, the requirements are the following.
+% \begin{itemize}
+%     \item Global consensus
+%     \item Security and fault tolerance
+%     \item Performance and scalability
+% \end{itemize}
 
-As mentioned in the Introduction, the primary goal is performance and scalability.
-Having a scalable blockchain system while still keeping global consensus
-allows the system to be ubiquitous and realise the full potential of blockchain.
+% As mentioned in the Introduction, the primary goal is performance and scalability.
+% Having a scalable blockchain system while still keeping global consensus
+% allows the system to be ubiquitous and realise the full potential of blockchain.
 
-The secondary goal is to design an application neutral system.
-In particular, it should act as a backbone that provides the building blocks of blockchain based applications.
-It should be possible to buld any kind of application on top our system.
-Further, we do not impose on a consensus algorithm, as long as it satisfies the properties of atomic broadcast which we describe in~\Cref{sec:overview-cons}.
+% The secondary goal is to design an application neutral system.
+% In particular, it should act as a backbone that provides the building blocks of blockchain based applications.
+% It should be possible to buld any kind of application on top our system.
+% Further, we do not impose on a consensus algorithm, as long as it satisfies the properties of atomic broadcast which we describe in~\Cref{sec:overview-cons}.
 
-Due to the nature of our system, we do not explicitly address the Sybil attack~\cite{douceur2002sybil}.
-Sybil defence mechanism always require some form of reputation score from the application.
-For example, social network based Sybil defence mechanims use graph structure of real-world relationships~\cite{yu2006sybilguard}.
-Online marketplaces such as Amazon use the rating of buyer and sellers.
-Thus it is not possible to design a Sybil defence mechanism with a application neutral framework.
-On the other hand, our system also has no restrictions on the Sybil defence technique
-and applications built on top of our system can use the best mechanism for their purpose.
+% The third and final goal is security.
+% Our system should be unaffected in the presence of powerful adversaries.
+% In particular, adversaries are Byzantine meaning that they can have arbitrary behaviour.
+% Thus anything is possible from simply ommitting messages to colluding with eachother in order to undermine the whole system.
 
-The third and final goal is security.
-Our system should be unaffected in the presence of powerful adversaries.
-In particular, adversaries are Byzantine meaning that they can have arbitrary behaviour.
-Thus anything is possible from simply ommitting messages to colluding with eachother in order to undermine the whole system.
-
-We begin the chapter with an intuitive overview of the architecture in \Cref{sec:system-overview}.
+The goal of this chapter is to detail our system specification.
+We begin our discussion with an intuitive overview of the architecture in \Cref{sec:system-overview}.
 Next, we give the formal description, starting with the model and assumptions in \Cref{sec:model-assumptions}.
@@
 namely consensus protocol (\Cref{sec:cons-protocol}), transaction protocol (\Cref{sec:tx-protocol}) and validation protocol (\Cref{sec:vd-protocol}).
-Finally, the possible extensions are described in \Cref{sec:protocol-extensions}.
+Finally, some optimisations are described in \Cref{sec:optimisations}.
 
@@
 \item Sans serif (e.g. $\textsf{fn}(\cdot)$) denotes a function.
-\item Typewrite (e.g. $\texttt{ack}$) denotes message type.
+\item Monospace (e.g. $\texttt{ack}$) denotes message type.
 \end{itemize}

### 2017-06-30 thesis/chapters/ch1-intro.tex
Better analysis

@@
 Ironically, also in 2008, Satoshi Nakamoto published the Bitcoin whitepaper~\cite{bitcoin}.
-Which, for the first time, gave us a simple banking system in the form of an distributed ledger.
+Which, for the first time, gave us a simple banking system in the form of a distributed ledger.
 It needs no central control but still incorruptable even if there are malicious parties that aim to undermine the system.
@@
 
-The primary innovations are (1) its consensus model, which prevents double spending,
+The primary innovations are (1) its consensus model which prevents double spending,
 and (2) its incentive mechanism that encourages anyone (with adaquate hardware) to participate in the network and keep it running.
@@
 At some instant in time, $A$ sees that $C$ transferred 5 units to $A$ and $B$ sees that $C$ transferred 5 units to $B$.
-Hence, the transaction is inconsistent with respect to the observer.
+Hence the transaction is inconsistent with respect to the observer.
 Bitcoin and many other blockchain systems solve the inconsistency problem with a consensus algorithm.
-The goal of the algorithm is reach agreement on a set of values, e.g. transactions, which eliminates inconsistencies.
+The goal of the algorithm is reach agreement on a set of valid values, e.g. transactions. 
+This effectively eliminates inconsistencies.
 In Bitcoin's case, the consensus algorithm is called \emph{proof-of-work}.
-Where miners (parties that runs the Bitcoin network) collect transactions and compete in solving (using brute force) a puzzle.
+Where miners (parties that runs the Bitcoin network) collect transactions and compete in solving (using brute force) a difficult puzzle.
 The first miner to solve it generates a block containing all the collected transactions.
@@
 Thus every party sees a consistent blockchain which solves the double spending problem.
-Further, due to the reward mechanism,
-people are incentivised to act as miners and keep the system running.
+% Further, due to the reward mechanism,
+% people are incentivised to act as miners and keep the system running.
 
@@
 Since every Bitcoin transaction is at least 250 bytes, it computes to about 7 transactions per second.
-Due to this limitation, it is not uncommon to see a long backlog of about 20000 transactions\footnote{\url{https://blockchain.info/unconfirmed-transactions/}}.
-A few months ago the backlog even reached 100000, which meant new transactions would take at least 11 hours to hit the blockchain~\cite{bitcoinbacklog}.
-The issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some calls it a civil war~\cite{bitcoincivilwar}.
+Due to this limitation, it is not uncommon to see a long backlog of about 20,000 transactions\footnote{\url{https://blockchain.info/unconfirmed-transactions/}}.
+A few months ago the backlog even reached 100,000, which meant new transactions would take at least 11 hours to hit the blockchain~\cite{bitcoinbacklog}.
+This issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some calls it a civil war~\cite{bitcoincivilwar}.
 Parties that are for the increase in block size argue that a larger block would improve the transaction rate.
@@
 It also requires a hard fork (not backward compatible), which risks consensus failure and devaluation.
-A recent empirical study Croman et el.~\cite{croman2016scaling} has shown that increasing the block size may help.
+A recent empirical study by Croman et el.~\cite{croman2016scaling} has shown that increasing the block size may help.
 But given the bandwidth and latency constraints,
@@
 
-We describe and analyse the aforementioned idea in the remainder of this work.
+We describe and analyse our idea in the remainder of this work.
 We begin with the problem description in~\Cref{ch:problem}.

### 2017-06-30 thesis/chapters/ch4-analysis.tex
Better analysis

@@
 
-\subsection{Bandwidth Requirement for Transactions}
+\subsection{Data Volume Requirement for Transactions}
 With our assumption on the time complexity of ACS,
-we are ready to analyse the bandwidth.
-We know that to create and then validate a transaction, the bandwidth required is of $O(l)$,
+we are ready to analyse the amount of data required to be transmitted over a link per transaction,
+which we call the data volume per transaction.
+% First, we define bandwidth as the ``amount of data a link or network path can deliver per unit of
+% time''~\cite{prasad2003bandwidth}.
+We know that to create and then validate a transaction,
+the data volume per transaction is of $O(l)$,
 where $l$ is the length of the agreed fragment.
@@
 But from \Cref{sec:acs-time-complexity},
-we know that $D_{\text{acs}}$ is of $O(N)$, thus the bandwidth per transaction is $O(N)$.
-% TODO how does message complexity translate to time complexity?
-% TODO computational task minimal compared to messages, thus the sending of message is expensive?
+we know that $D_{\text{acs}}$ is of $O(N)$, thus the data volume per transaction is $O(N)$.
 This is intuitive because consensus duration would be longer if there are more CP blocks, which means that the agreed fragments are longer (assuming nodes transact at a constant rate).
-This behaviour is also verified experimentally in \Cref{ch:implementation}.
+The behaviour is also verified experimentally in \Cref{ch:implementation}.
 
@@
 Using our results so far, we are able to analyse the global throughput.
-Suppose every node has some fixed bandwidth capacity $C$ (unit of communication per second),
-they make transactions at $r_{\text{tx}}$ per second.
-Then we have the inequality $C \ge r_{\text{tx}} l$, where $l$ is the length of the the agree fragment as before.
-Rearranging, we get
-$$\frac{C}{l} \ge r_{\text{tx}}.$$
-With this, we consider two cases, first is  when all the nodes are running at maximum capacity, i.e. $\frac{C}{l} = r_{\text{tx}}$.
-Recall that $l$ is of $O(N)$, so as the the population increases, 
-the transaction rate must decrease in order to maintain the inverse relationship,
-thus $r_\text{tx}$ is of $O(N^{-1})$. % TODO is this formal? use omega?
-Therefore, the global throughput would be $O(N^{-1})N = O(1)$.
-In the second case, nodes are not running at maximum capacity and $r_{\text{tx}}$ is maintained.
-Therefore, if every node runs at $r_{\text{tx}}$,
-the global throughput becomes $O(N)$ until $N$ is too large and we go back to the first case.
-A visualisation of the result is shown in \Cref{fig:throughput-visualisation}.
+First we clarify the bandwidth definition, which is ``the data rate at which a network link or a network path can transfer''~\cite{prasad2003bandwidth}.
+Now, suppose every node has some fixed bandwidth per link $C$ and $N$ links.
+They also make transactions at $r_{\text{tx}}$ per second.
+Then we have the inequality
+$$NC \ge r_{\text{tx}} l,$$
+where $l$ is the length of the the agree fragment as before.
+The inequality suggests that the rate for which transactions and validations are made cannot exceed the total bandwidth of all the links.
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.7\textwidth]{throughput-visualisation}
-  \caption{Visualisation of the worst case global throughput.}
-  \label{fig:throughput-visualisation}
-\end{figure}
+We note that the inequality does not hold if the node is only transacting with a small sample of some subset of the population.
+This is because it cannot use all the bandwidth available in all the links.
+For instance, if the node is only transacting with one other node, then it can only use the bandwidth of one link which is only $C$.
+However, if that is the case, we can intelligently cache the \texttt{vd\_resp} messages.
+Thus multiple transactions can be validated with a single \texttt{vd\_resp} message.
+For this work, we analyse the worst case where every node transacts with a random node from the population,
+and a new \texttt{vd\_resp} must be sent for every \texttt{vd\_req} message.
 
-The upshot of this analysis is that our system scales nicely at a global throughput of $O(N)$ until the population gets too large.
-Then we maintain a constant global throughput.
-This result falls a bit short of what we envisioned in the introduction.
-However, the population is not the number of nodes that use the system, but the number of nodes that are online during a single round.
-Furthermore, this result is for the worst case where every transaction needs an agreed fragment to be transmitted.
-In practice, nodes are able to cache agreed fragments.
-For instance, if $u$ and $v$ make $x$ transactions in a single round,
-then only one agreed fragment need to be exchanged as it contains all the transactions rather than $x$ agreed fragments.
+Consider the case where the system is making use of all the bandwidth, i.e. $NC = r_{\text{tx}} l$.
+Recall that $l$ is of $O(N)$, that means LHS and RHS both grow linearly with respect to $N$.
+Hence, there exist some constant $r_\text{tx}$ that makes use of all the available bandwidth regardless of $N$.
+Finally, if every node in the network is transacting at a constant rate, then the global throughput (in terms of transactions per second) is of $O(N)$.
+If the system is not making use of all the bandwidth.
+We also maintain a constant transaction rate by the same argument.
+Thus a global throughput is also $O(N)$.
+We verify both of these claims experimentally in~\Cref{sec:evaluation}.
+
+% The upshot of this analysis is that our system scales nicely at a global throughput of $O(N)$ until the population gets too large.
+% Then we maintain a constant global throughput.
+% This result falls a bit short of what we envisioned in the introduction.
+% However, the population is not the number of nodes that use the system, but the number of nodes that are online during a single round.
+% Furthermore, this result is for the worst case where every transaction needs an agreed fragment to be transmitted.
+% In practice, nodes are able to cache agreed fragments.
+% For instance, if $u$ and $v$ make $x$ transactions in a single round,
+% then only one agreed fragment need to be exchanged as it contains all the transactions rather than $x$ agreed fragments.
 
@@
 the expected value is  much smaller than $\lfloor \frac{n-1}{3} \rfloor$ is small.
-For example, we can use tail inequality to bound the probability of picking more than $\lfloor \frac{n-1}{3} \rfloor$ black balls when only 5\% are black.
+We can use tail inequality to bound the probability of picking more than $\lfloor \frac{n-1}{3} \rfloor$ black balls when only $n\alpha$ are black where $0 \le \alpha \le \lfloor \frac{n-1}{3} \rfloor / n$.
 The tail inequality is
@@
 $$
-where $E[X] = n/20$.
+where $E[X] = n\alpha$.
 We are interested in $\Pr[X \ge \lfloor \frac{n-1}{3} \rfloor + 1]$, so
 \begin{align*}
-\tau &= \frac{\lfloor \frac{n-1}{3} \rfloor + 1}{n} - \frac{1}{20} \\
-  &> \frac{\lfloor \frac{n-1}{3} \rfloor}{n} - \frac{1}{20} \\
-  &\ge \frac{1}{4} - \frac{1}{20} = \frac{1}{5} \quad \text{for $n \ge 4$}.
+\tau &= \frac{\lfloor \frac{n-1}{3} \rfloor + 1}{n} - \alpha
 \end{align*}
-Putting $\tau$ back into the tail inequality we get
+Putting $\tau$ back into the tail inequality we get the following bound.
 $$
-\Pr[X \ge \lfloor \frac{n-1}{3} \rfloor + 1] \le e^{-\frac{2}{25} n}.
+\Pr[X \ge \lfloor \frac{n-1}{3} \rfloor + 1] \le e^{-2 \big(\frac{\lfloor  \frac{n - 1}{3} \rfloor + 1}{n} - \alpha \big)^2 n}
 $$
-The bound is not tight, but it is clear that for sufficiently large $n$, the probability becomes vanishingly small.
-Hence, it is possible to use our system in practice when 5\% of the population are adversarial but $n \ge 3t + 1$ does not hold,
-as long as $n$ is large.
+
+The bound is not tight, but it is useful for for picking parameters.
+For some fixed $n$, since $0 \le \alpha \le \lfloor \frac{n-1}{3} \rfloor/n < (\lfloor  \frac{n - 1}{3} \rfloor + 1) / n$,
+the probability is maximum when the squared term is minimum at $\alpha = \lfloor \frac{n-1}{3} \rfloor/n$.
+The probability is minimum when the squared term is maximum at $\alpha = 0$.
+Hence, if $n$ is known, then we can pick a $\alpha$ such that the probability becomes small.
+On the other hand, if $\alpha$ is fixed, but small, we may increase $n$ to achieve the same.
+To put this into perspective, suppose $n = 1000$ and $\alpha = 1/20$, then the probability to draw more black balls than the threshold is only $2 \times 10^{-48}$.

### 2017-07-03 thesis/appendices/additional-graphs.tex
New graphs

\chapter{Additional Graphs}
\label{app:additional-graphs}

\begin{figure}[h]
  \centering
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/consensus-duration-vs-population}
    \caption{Transactions are with fixed neighbours}
  \end{subfigure}

  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/consensus-duration-vs-population}
    \caption{Transactions are with random neighbours}
  \end{subfigure}
  \caption{Communication cost for consensus}
  \label{fig:consensus-duration}
\end{figure}

### 2017-07-03 thesis/chapters/ch4-analysis.tex
New graphs

@@
 
-\subsection{Data Volume Requirement for Transactions}
+\subsection{Communication Cost for Transactions}
+\label{sec:communication-cost-for-tx}
 With our assumption on the time complexity of ACS,
 we are ready to analyse the amount of data required to be transmitted over a link per transaction,
-which we call the data volume per transaction.
+which we call the communication cost per transaction.
 % First, we define bandwidth as the ``amount of data a link or network path can deliver per unit of
@@
 We know that to create and then validate a transaction,
-the data volume per transaction is of $O(l)$,
+the communication cost per transaction is of $O(l)$,
 where $l$ is the length of the agreed fragment.
@@
 But from \Cref{sec:acs-time-complexity},
-we know that $D_{\text{acs}}$ is of $O(N)$, thus the data volume per transaction is $O(N)$.
+we know that $D_{\text{acs}}$ is of $O(N)$, thus the communication per transaction is $O(N)$.
 This is intuitive because consensus duration would be longer if there are more CP blocks, which means that the agreed fragments are longer (assuming nodes transact at a constant rate).
@@
 \subsection{Global Throughput}
+\label{sec:global-throughput}
 Using our results so far, we are able to analyse the global throughput.

### 2017-07-03 thesis/chapters/ch5-implementation.tex
New intro and etc.
New graphs

--- 2017-05-15 thesis/chapters/ch3-implementation.tex
@@
+\chapter{Implementation and Experimental Results}
+\label{ch:implementation}
+
+Henceforth, we evaluate out system experimentally and compare the results with the theoretical analysis.
+We begin this chapter by a description of the implementation in \Cref{sec:implementation}.
+Then, we move on to describing our experimental setup in \Cref{sec:experimental-setup}.
+Finally, \Cref{sec:evaluation} presents our experimental results and our evaluation.
+Our experiment primarily focuses on the consensus duration and the throughput.
+
 \section{Implementation}
-Python. Twisted.
+\label{sec:implementation}
 
-\section{Evaluation}
-\begin{itemize}
-  \item How fast is the consensus algorithm? Possibly plot graph of time versus
-    the number of nodes.
-  \item Does the promoter registration phase add a lot of extra overhead?
-  \item What's the rate of transaction such that they can be verified ``on
-    time'', i.e. without a growing backlog?
-  \item Our global validation rate is somewhat equivalent to the transaction
-    rate in other systems. Does the validation rate scale with respect to the
-    number of nodes? In theory it should. Plot validation rate vs number of
-    nodes, we expect it to be almost linear.
-\end{itemize}
+The prototype implementation can be found on GitHub\footnote{\url{https://github.com/kc1212/consensus-thesis-code}}.
+It is written in the event driven paradigm, using the Python programming language\footnote{\url{https://www.python.org/}} and
+the Twisted\footnote{\url{https://twistedmatrix.com/}} library for networking.
+
+The structure of the implementation is primarily made up of three modules---\texttt{acs}, \texttt{trustchain} and \texttt{node}.
+\texttt{acs}, as its name suggests, implements ACS.
+ACS uses erasure code in one of its subprotocols (reliable broadcast).
+Thus we use the liberasurecode\footnote{\url{https://github.com/openstack/liberasurecode}} library for its Reed-Solomon error correcting code functionality.
+An implementation detail is that liberasurecode cannot create more than 32 code blocks\footnote{
+  The 32 code blocks limitation is hardcoded in the source file, see
+  \url{https://github.com/openstack/liberasurecode/blob/0794b31c623e4cede76d66be730719d24debcca9/include/erasurecode/erasurecode.h\#L35}
+}, we discuss the effect of this in~\Cref{sec:experimental-setup}.
+The \texttt{acs} module provides a small interface to the caller to start and stop the consensus process and also retrieve results.
+The \texttt{trustchain} module implements the Extended TrustChain data structure.
+It also provides the essential algorithm necessary to interact with Extended TrustChain such as 
+$\textsf{new\_tx}(\cdot)$, $\textsf{new\_cp}(\cdot)$, $\textsf{agreed\_fragment}(\cdot)$ and so on.
+Since this is a prototype implementation, we only store the data structure in memory and not on disk.
+Finally, the \texttt{node} module ties everything together.
+It implements the the consensus protocol, the transaction protocol and the validation protocol.
+
+Every node keeps a persistent TCP connection with every other node.
+This creates a fully connected network for our experiment.
+It is certainly not ideal in real world scenarios where nodes may have limited resources (e.g. sockets).
+But as a prototype, it is sufficient to run a network of over a thousand nodes and experiment with it.
+
+\section{Experimental Setup}
+\label{sec:experimental-setup}
+
+The the goal of the experiment is to run the three protocols---consensus protocol,
+transaction protocol and validation protocol---simultaneously and analyse the throughput, consensus duration and other related metrics.
+We investigate these properties under the following four parameters.
+\begin{enumerate}
+  \item The transaction rate $r_{\text{tx}}$ per node. This is not comparable to the others because it is fixed at 2 TX/s.
+        Nevertheless, it is a good value because, as we show later,
+        it hits a bottleneck in extreme cases which helps us understand the limitations of our design.
+  \item The number of facilitators $n \in \{ 12, 16, \dots, 32\}$.
+        Unfortunately, the maximum $n$ is 32 because the limitation in liberasurecode mentioned in \Cref{sec:implementation}.
+  \item The population size $N \in \{200, \dots, 1200\}$.
+        $N$ stops at 1200 is due to our physical setup, which we describe next.
+  \item The two modes of transaction.
+        The first mode or the ideal mode is that nodes only transact with their immediate neighbour. 
+        This minimises the data volume per validated transaction because agreed fragment can be cached.
+        The second mode is in the other extreme, where every transaction is with a random node out of the $N$ nodes in the system,
+        thus the agreed fragment is unlikely to be cached.
+\end{enumerate}
+
+The experiment is run on the DAS-5 (The Distributed ASCI Supercomputer 5\footnote{\url{https://www.cs.vu.nl/das5/}}).
+From now on, we use "machines" to refer to DAS-5 nodes and nodes to refer to a running instance in our system.
+On DAS-5 we use up to 30 machine, for each machine we use 40 nodes.
+This gives us the aforementioned 1200 number.
+With this setup, we cannot run more nodes because the every machine only has 65535 ports available (minus the reserved ones).
+But 40 nodes each need 1200 TCP connections which is 48000 TCP connections per machine and that is inching close to the limit.
+While it is possible to have many more TCP connections per machine,
+but it requires additional network interface which is something we do not control on the DAS-5.
+
+To coordinate nodes on many different machine,
+we employ a discovery server to inform every node the IP addresses and port numbers of every other node.
+It is only run before the experiment and is not used during the experiment.
+
+\section{Communication Cost for Consensus}
+
+\begin{figure}[h]
+  \centering
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-fixed/consensus-communication-cost-vs-population}
+    \caption{Transactions are with fixed neighbours}
+    \label{fig:consensus-comms-fixed}
+  \end{subfigure}
+
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-random/consensus-communication-cost-vs-population}
+    \caption{Transactions are with random neighbours}
+    \label{fig:consensus-comms-random}
+  \end{subfigure}
+  \caption{Communication cost for consensus}
+  \label{fig:consensus-comms}
+\end{figure}
+
+\Cref{fig:consensus-comms} show the relationship between communication cost and population.
+These most important aspect is that these results show a linear increase.
+This reinforces our analytical result in~\Cref{sec:acs-complexity}.
+Further, note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
+the magnitude of the communication cost are similar.
+Both peak at about $1.4 \times 10^{10}$ bytes.
+This is expected because the consensus protocol is decoupled from the transaction protocol.
+
+We are also interested in how communication costs translate to time.
+Hence, for the same experiment we use duration in seconds on the vertical axis instead of communication cost and plot the~\Cref{fig:consensus-duration}
+These show a similar trend as the communication cost graphs.
+Hence our assumption (from~\Cref{sec:acs-time-complexity}) that for every unit of communication there is some fixed overhead is valid.
+
+\section{Communication Cost for Transaction and Validation}
+
+\begin{figure}[h]
+  \centering
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-fixed/tx-communication-cost-vs-population}
+    \caption{Transactions are with fixed neighbours}
+    \label{fig:tx-comms-fixed}
+  \end{subfigure}
+
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-random/tx-communication-cost-vs-population}
+    \caption{Transactions are with random neighbours}
+    \label{fig:tx-comms-random}
+  \end{subfigure}
+  \caption{Communication cost per verified transaction}
+  \label{fig:tx-comms}
+\end{figure}
+
+Recall that in~\Cref{sec:communication-cost-for-tx} we argued that the communication cost per verified transaction is of $O(N)$.
+To verify the argument, we plot the relationship between the communication cost and population size in~\Cref{fig:tx-comms}.
+Again, we observe a linear relationship, which verified the argument.
+
+However, note that there is a large difference in communication cost between the two modes of transaction.
+When transacting with only one neighbour, the communication cost is low because only one agreed fragment needs to be communicated for every round in order to validate all transactions of that round.
+This is because agreed fragments are cached.
+On the other hand,
+if every node is transacting with a random node, then it is likely the case that one agreed fragment needs to be communicated for every transaction.
+Hence the communication cost we see in~\Cref{fig:tx-comms-random} is much higher than in~\Cref{fig:tx-comms-fixed}.
+
+It is interesting to note that some fluctuations exist in the graphs.
+This is likely due to the way how the communication cost are recorded.
+We maintain the total amount of data sent and received in a few variables.
+On every message that is sent or received, we increment the relevent variable by the size of the message.
+However, it is infeasible to log the values every time a message is sent or received as there is simply too much data for our build server to hold.
+Therefore we only log every 5 seconds.
+This does not always align with the time when a round finishes, thus some of the data may not represent reality accurately.
+%  TODO this can be fixed
+
+\section{Global Throughput}
+The global throughput results are shown in~\Cref{fig:global-throughput}.
+Clearly, the throughput has a linear relationship with the population size.
+This result is a strong indication of the horizontal scalability which we aimed to achieve.
+It also verified our analytical result in~\Cref{sec:global-throughput}.
+
+More interestingly, is the magnitude in throughput, 
+and the difference in magnitude between the two modes of transaction.
+Firstly, recall that we fixed our $r_\text{tx}$ to 2, but how is it possible to have nearly 9000 transactions per second for 1200 nodes (which is about 7.5 transactions per second)?
+This is due to the way validated transactions are calculated.
+Transactions are between two parties, hence if every node makes two transactions per seoncd,
+every node also expects to receive two transactions per second.
+Hence, for every node, the TX blocks are created at 4 per second.
+Every TX block must be validated.
+So everyone sends 4 validation requests per second and also receives 4 per second.
+Thus during one second 8 validations are performed.
+The difference in magnitude is caused by the caching mechanism mentioned earlier.
+If an agreed fragment needs to be transmitted to validating every transaction then it puts a toll on the network infrastructure.
+The low transaction rate in~\Cref{fig:global-throughput-random} is caused by fact that the network infrastructure cannot keep up with our demand.
+We demonstrate this issue from a different angle in the next section.
+
+\begin{figure}[h]
+  \centering
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-fixed/throughput-vs-population}
+    \label{fig:global-throughput-fixed}
+  \end{subfigure}
+
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-random/throughput-vs-population}
+    \label{fig:global-throughput-random}
+  \end{subfigure}
+  \caption{Global throughput}
+  \label{fig:global-throughput}
+\end{figure}
+
+\section{Backlog}
+
+TODO
+
+\begin{figure}[h]
+  \centering
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-fixed/timeseries}
+  \end{subfigure}
+
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-random/timeseries}
+  \end{subfigure}
+\end{figure}

### 2017-07-04 thesis/chapters/ch2-problem.tex
Write some optimisation

@@
 
-\section{Non-goals}
-Due to the nature of our system, we do not explicitly address the Sybil attack~\cite{douceur2002sybil}.
-Sybil defence mechanism always require some form of reputation score from the application.
+\section{Limitations}
+Our work removes the transaction bound from the blockchain.
+However, the price to pay is that the Sybil attack~\cite{douceur2002sybil} prevention mechanism moves from the blockchain backbone to the application layer.
 For example, social network based Sybil defence mechanims use graph structure of real-world relationships~\cite{yu2006sybilguard}.
 Online marketplaces such as Amazon use the rating of buyer and sellers.
-Thus it is not possible to design a Sybil defence mechanism with a application neutral framework.
+Since our system is application neutral, we do not provide an explicit Sybil defence mechanism.
 On the other hand, our system also has no restrictions on the Sybil defence technique

### 2017-07-04 thesis/chapters/ch3-model.tex
Write some optimisation

@@
 % Thus anything is possible from simply ommitting messages to colluding with eachother in order to undermine the whole system.
+
+We present a design which removes the transaction rate restrictions seen in traditional blockchain systems.
+This is done by combining the recent HoneyBadgerBFT~\cite{miller2016honey}
+(a purely asynchronous consensus algorithm for blockchain systems)
+with a modification of our prior work on TrustChain~\cite{trustchain}.
+The design is inspired by the essence the (early) Internet architecture as well as how transactions are performed in the real world.
 
@@
 
-(Move to future work?)
+Up to this point,
+we have discussed our protocol in the context of the model and assumptions defined in \Cref{sec:model-assumptions}.
+In this section, 
+we explore a few optimisations which we can make if we relax our model.
+These relaxations do not have the same theoretical elegance as our original model.
+But they enable better performance and allow us to apply our design in the fully permissionless setting.
 
-Up to this point, we have discussed our protocol in the context of the model and assumptions defined in \Cref{sec:model-assumptions}.
-In this section, we remove a few assumptions and discuss how our architecture is adapted.
+\subsection{Use epidemic protocol instead of broadcast protocol}
+One of the final steps in our consensus protocol is to broadcast the consensus result and signatures to every other node (\Cref{sec:consensus-phase}).
+While this guarantees delivery (since we assumed reliable channel), it is wasteful.
+For example, if every facilitator is honest, a node would receive $n$ consensus results which are identical when only one is necessary.
 
-Gossip?
+An optimisation we can do is to use an epidemic protocol~\cite{eugster2004epidemic} (also known as gossiping)) instead of our broadcast approach.
+Typical epidemic protocols works as follows.
+Every node buffers every message it receives up to some buffer size $b$.
+Then it forwards the messages $t$ number of times.
+Every time the message is sent to $f$ random neighbours, $f$ is often called the fan-out.
+The upside of using epidemic protocol is a lower communication cost.
+This is especially true with a lazy push approach the node who just received a message would push the message ID to its $f$ random neighbours,
+and only push the full message if the neighbour explicitly requests for the message~\cite{leitao2007epidemic}.
+With this, nodes typically only need to receive one consensus result message instead of $n$.
 
-Churn?
+A down side of epidemic protocol is that it usually takes $O(\log N)$ time to infect the whole network,
+whereas broadcasting uses constant time.
+Another downside of some epidemic protocols (e.g. eager push) is that it is difficult to guarantee delivery.
+It is especially true when the parameters are not choosen correctly of if the network is not fully connected.
+If the delivery cannot be guaranteed, 
+then we cannot guarantee liveness in our consensus protocol because a future facilitator may miss the memo.
+Picking parameters are difficult in practice because the network configuration is unknown and the total number of nodes might also be unknown.
 
-Sybil?
+\subsection{Using timing assumption in the permissionless setting}
+Our model is purely asynchronous, where we make no timing assumptions anywhere in the protocol.
+In many applications, it is often fine to make timing assumptions.
+For example, TCP relies on timeout for its retransmission and the nLockTime property in Bitcoin transactions makes the transaction unspendable until some time in the future (either Unix time or block height)~\cite{bitcoindevguide}.
+One limitation of our system is that we use the parameter $N$ in our algorithms, which makes it unsuitable for the permissionless environment where users can join and leave at will.
+In this section we show how making a timing assumption would allow us to operate in the permissionless setting.
+
+At the start of our consensus phase (\Cref{sec:consensus-phase}), facilitators must wait for $N-f$ \text{cp\_msg} messages.
+This is the only place where we used $N$ as a parameter.
+To introduce timing, instead of waiting for $N-f$ messages, we wait for some time $D$,
+such that $D$ is sufficiently long for honest nodes to send their CP blocks to the facilitators.
+Again, choosing the parameter $D$ is difficult and depends on a number of factors such as the network condition, message size, and so on.
+Overestimating it would make agreed fragments much longer than usual.
+Underestimating it would lead to unfairness where users that are too late do not have a chance to be selected as a facilitator in the next round.
+Nevertheless, there is a significant gain for making the timing assumption and that is the ability to operate in the permissionless setting.
+
+TODO explain churn
 

### 2017-07-08 thesis/chapters/ch3-model.tex
Finish some TODOs

@@
 Our model is purely asynchronous, where we make no timing assumptions anywhere in the protocol.
-In many applications, it is often fine to make timing assumptions.
+In many applications however, it is often fine to make timing assumptions.
 For example, TCP relies on timeout for its retransmission and the nLockTime property in Bitcoin transactions makes the transaction unspendable until some time in the future (either Unix time or block height)~\cite{bitcoindevguide}.
@@
 
-At the start of our consensus phase (\Cref{sec:consensus-phase}), facilitators must wait for $N-f$ \text{cp\_msg} messages.
+At the start of our consensus phase (\Cref{sec:consensus-phase}), facilitators must wait for $N-f$ \texttt{cp\_msg} messages.
 This is the only place where we used $N$ as a parameter.
@@
 Again, choosing the parameter $D$ is difficult and depends on a number of factors such as the network condition, message size, and so on.
-Overestimating it would make agreed fragments much longer than usual.
+Overestimating it would make agreed fragments much longer than usual, which increases communication costs for validation.
 Underestimating it would lead to unfairness where users that are too late do not have a chance to be selected as a facilitator in the next round.
-Nevertheless, there is a significant gain for making the timing assumption and that is the ability to operate in the permissionless setting.
+Nevertheless, there is a significant gain for making the timing assumption and that is the ability to operate in the permissionless setting which we explain next.
 
-TODO explain churn
+Suppose a new node wish to join the network and the facilitators are known (this can be done with a public registry).
+It simply sends its latest CP block to the facilitators.
+Then, in the next round the node will have a chance to become a facilitator just like any existing node.
+To leave the network, nodes simply stop submitting CP blocks.
+There is a subtlety here which happens when the node is elected as a facilitator in the following round.
+In this case, the node must fulfill its oblication by completing the consensus protocol (but without proposing its own CP block) before leaving.
+Otherwise the $n \ge 3t + 1$ condition may be violated.
+
+% Another assumption we made in this work is the existance of a PKI.
+% This can also be removed with the introduction of timing by storing the public keys the consensus results,
+% much like in Bitcoin.
 

### 2017-07-11 thesis/chapters/ch2-problem.tex
More graphs; consensus protocol analysis

@@
 
-Traditional blockchain systems use PoW as the consensus algorithm.
-However, consensus algorithm and especially Byzantine\footnote{Byzantine means participants can experience arbitrary failure, more on this in~\Cref{sec:acs-background}.}
-consensus algorithms have existece since 1982,
-where it is first introduced by Lamport et el.~\cite{lamport1982byzantine}.
+Traditional blockchain systems use PoW as the consensus algorithm,
+but consensus algorithm and especially Byzantine\footnote{Byzantine means participants can experience arbitrary failure, more on this in~\Cref{sec:acs-background}.}
+consensus algorithms have existece since the 80s~\cite{pease1980reaching, lamport1982byzantine}.
 The Practical Byzantine Fault Tolerance (PBFT) algorithm introduced in 1999 by Castro and Liskov~\cite{castro1999practical} is widely regarded as the first high-performance consensus algorithm.
-Since then research in this area picked up pace and many more algorithms are introduced~\cite{abd2005fault, kotla2007zyzzyva}.
+The authors show that running PBFT in Network File System (NFS) was only 3\% slower than the standard unreplicated NFS.
+Since then, research in this area picked up pace and many more algorithms are introduced~\cite{abd2005fault, kotla2007zyzzyva}.
 With the introduction of Bitcoin's probabilistic consensus model, the field experienced a second renaissance.

### 2017-07-11 thesis/chapters/ch3-model.tex
More graphs; consensus protocol analysis

@@
 In our model we consider $N$ nodes, which is the population size.
-$n$ of them are facilitators, $t$ out of $n$ are malicious and the inequalities
-$n \ge 3t + 1$ and $N \ge n + t$ must hold.
-The first is from the FLP impossibility result TODO.
-The second is necessary due to our system design, which becomes clear in TODO.
+$n$ of them are facilitators, $t$ out of $n$ are malicious and the inequality
+$n \ge 3t + 1$ must hold.
+This is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$.
+Further, the inequality $N \ge n + t$ must also hold.
+This is due to our system design, which becomes clear in~\Cref{sec:consensus-phase}.
 
@@
 
+\subsection{Privacy preserving validation protocol}
+TODO
+
 % Another assumption we made in this work is the existance of a PKI.

### 2017-07-11 thesis/chapters/ch4-analysis.tex
More graphs; consensus protocol analysis

@@
 \section{Performance}
-This section aims to (analytically) answer our research question.
+This section aims to analytically answer our research question.
 That is, does the global throughput increase linearly with respect to the population size?
-We begin by looking at the communication and time complexity of ACS,
+We begin by looking at the communication and time complexity of the consensus protocol,
 and then the bandwidth requirement for a single transaction.
-These are prerequisites of the global throughput analysis.
+We build on top of those results to analyse do the global throughput analysis.
 
-\subsection{Communication Complexity of ACS}
+\subsection{Communication complexity of the consensus protocol}
 \label{sec:acs-complexity}
+The consensus protocol can bee seen as three parts.
+Thus the communication complexity is the sum of these three parts.
+The first part is when every node sends their CP block to all the facilitators, which is $O(Nn)$ since there are $N$ nodes and $n$ facilitators.
+Or simply $O(N)$ if we consider $n$ as a constant.
+
+The second part is ACS.
 The communication complexity of ACS is $O(n^2|v| + \lambda n^3 \log n)$~\cite{miller2016honey},
@@
 
+The third and final part is the dissemination, where the facilitators broadcast the consensus result along with their signatures.
+For the same reason as the first part, this is also $O(N)$.
+Thus the combined communication complexity when $n$ is a constant is $O(N)$.
+
 % We argue that with a $O(N)$ message complexity we also have $O(N)$ time complexity.
@@
 
-\subsection{Time Complexity of ACS}
+\subsection{Time complexity of the consensus protocol}
 \label{sec:acs-time-complexity}
 In order to make arguments on bandwidth or throughput, which are concepts that depend on time,
-we must make some assumptions on the time complexity of ACS.
+we must make some assumptions on the time complexity of consensus protocol.
 The time complexity we use here is not the same as what is typically used in distributed systems.
-In analysis of distributed systems, time complexity is often in number of rounds.
+In analysis of distributed systems, time complexity is often in terms of the number of rounds.
 For example, ACS runs in a constant number of rounds because its subprotocols---reliable broadcast and binary Byzantine consensus---also run in a constant number of rounds.
 However, in practice, making a unit of communication always has some overhead associated with it, for example serialising and writing it to some network socket.
-Hence, for the remainder of our performance analysis, add the following to our computational model.
+Hence, for the remainder of our performance analysis, we add the following to our computational model.
 For every unit of communication, we assume they take some non-negligible but constant time to perform.
-From our result in \Cref{sec:acs-complexity} and the fact that ACS runs in a constant number of rounds,
-it follows that ACS has a time complexity of $O(N)$.
+Hence, from~\Cref{sec:acs-complexity}, it directly follows that the consensus protocol has a time complexity of $O(N)$.
 
@@
 \label{sec:communication-cost-for-tx}
-With our assumption on the time complexity of ACS,
+With the analysis on the time complexity of consensus protocol,
 we are ready to analyse the amount of data required to be transmitted over a link per transaction,
@@
 This can be seen from the fact that the largest message by far is the \texttt{vd\_resp} message,
-which contains the agreed fragment,
-the other messages (\texttt{tx\_req}, \texttt{tx\_resp} and \texttt{vd\_req}) are constant factors.
+which contains the agreed fragment.
+The other messages (\texttt{tx\_req}, \texttt{tx\_resp} and \texttt{vd\_req}) are constant factors.
 If we assume that every node performs transactions at a constant rate of $r_{\text{tx}}$ per second.
 Then
-$$l = r_{\text{tx}} D_{\text{acs}},$$
-where $D_{\text{acs}}$ is the duration an instance of ACS.
+$$l = r_{\text{tx}} D_{\text{c}},$$
+where $D_{\text{c}}$ is the duration a round of the consensus protocol.
 But from \Cref{sec:acs-time-complexity},
-we know that $D_{\text{acs}}$ is of $O(N)$, thus the communication per transaction is $O(N)$.
-This is intuitive because consensus duration would be longer if there are more CP blocks, which means that the agreed fragments are longer (assuming nodes transact at a constant rate).
+we know that $D_{\text{c}}$ is of $O(N)$, thus the communication cost per transaction is $O(N)$.
+This is intuitive because round duration would be longer if there are more CP blocks (more $N$), which means that the agreed fragments are longer (assuming nodes transact at a constant rate).
 The behaviour is also verified experimentally in \Cref{ch:implementation}.

### 2017-07-11 thesis/chapters/ch5-implementation.tex
Write some optimisation
More graphs; consensus protocol analysis

@@
 
+Finally, the cryptography primitives we use are SHA256 for hash functions and Ed25519 for digital signatures.
+Both of which are provided by libnacl~\footnote{\url{https://pypi.python.org/pypi/libnacl}}.
+
 \section{Experimental Setup}
@@
 
-\section{Communication Cost for Consensus}
+\section{Communication cost for a round}
 
-\begin{figure}[h]
+\begin{figure}[tb]
   \centering
-  \begin{subfigure}{\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-fixed/consensus-communication-cost-vs-population}
+  \makebox[\linewidth][c]{%
+    \begin{subfigure}[b]{.7\textwidth}
+      \includegraphics[width=1.0\textwidth]{neighbour-fixed/round-communication-cost-vs-population}
       \caption{Transactions are with fixed neighbours}
       \label{fig:consensus-comms-fixed}
-  \end{subfigure}
-
-  \begin{subfigure}{\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-random/consensus-communication-cost-vs-population}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.7\textwidth}
+      \includegraphics[width=1.0\textwidth]{neighbour-random/round-communication-cost-vs-population}
       \caption{Transactions are with random neighbours}
       \label{fig:consensus-comms-random}
-  \end{subfigure}
-  \caption{Communication cost for consensus}
-  \label{fig:consensus-comms}
+    \end{subfigure}%
+  }
 \end{figure}

### 2017-07-12 thesis/chapters/ch3-model.tex
Refactor experimental results

@@
 namely consensus protocol (\Cref{sec:cons-protocol}), transaction protocol (\Cref{sec:tx-protocol}) and validation protocol (\Cref{sec:vd-protocol}).
-Finally, some optimisations are described in \Cref{sec:optimisations}.
+Finally, we discuss a few variations of our main design and their respective tradeoffs in \Cref{sec:tradeoffs}.
 
@@
 \subsection{Relationship with $\alpha$-Synchroniser}
+TODO
 
@@
 
-\section{Optimisations}
-\label{sec:optimisations}
+\section{Design variations and tradeoffs}
+\label{sec:tradeoffs}
 
@@
 In this section, 
-we explore a few optimisations which we can make if we relax our model.
-These relaxations do not have the same theoretical elegance as our original model.
-But they enable better performance and allow us to apply our design in the fully permissionless setting.
+we explore a few design variations which we can make if we relax our model.
+These variations do not have the same theoretical elegance as our original model.
+But they enable better performance, allow us to apply our design in the fully permissionless setting and improves privacy.
 
@@
 \subsection{Privacy preserving validation protocol}
+TODO
+
+\subsection{Global probabilistic fork detection}
 TODO

### 2017-07-12 thesis/chapters/ch4-analysis.tex
Refactor experimental results

@@
 \section{Performance}
+\label{sec:performance-analysis}
 This section aims to analytically answer our research question.
@@
 \subsection{Communication complexity of the consensus protocol}
-\label{sec:acs-complexity}
+\label{sec:cons-complexity}
 The consensus protocol can bee seen as three parts.
@@
 
-\subsection{Time complexity of the consensus protocol}
-\label{sec:acs-time-complexity}
+\subsection{Duration of the consensus protocol}
+\label{sec:cons-duration}
 In order to make arguments on bandwidth or throughput, which are concepts that depend on time,
-we must make some assumptions on the time complexity of consensus protocol.
-The time complexity we use here is not the same as what is typically used in distributed systems.
+we must make some assumptions regarding our computational model to make arguments on the duration of the consensus protocol.
+Note that it is not the same as the time complexity typically used in distributed systems.
 In analysis of distributed systems, time complexity is often in terms of the number of rounds.
@@
 For every unit of communication, we assume they take some non-negligible but constant time to perform.
-Hence, from~\Cref{sec:acs-complexity}, it directly follows that the consensus protocol has a time complexity of $O(N)$.
+Hence, from~\Cref{sec:cons-complexity} and the fact that the consensus protocol uses a constant number of rounds,
+it follows that the consensus protocol has a duration of $O(N)$.
 
@@
 \label{sec:communication-cost-for-tx}
-With the analysis on the time complexity of consensus protocol,
+With the analysis on the duration of the consensus protocol,
 we are ready to analyse the amount of data required to be transmitted over a link per transaction,
@@
 where $D_{\text{c}}$ is the duration a round of the consensus protocol.
-But from \Cref{sec:acs-time-complexity},
+But from \Cref{sec:cons-duration},
 we know that $D_{\text{c}}$ is of $O(N)$, thus the communication cost per transaction is $O(N)$.

### 2017-07-12 thesis/chapters/ch5-implementation.tex
Refactor experimental results

@@
 
+The remainder of this chapter follows the same structure as the performance analysis in~\Cref{sec:performance-analysis},
+where we check our theoretical evaluation with our experimental results.
+
 \begin{figure}[tb]
@@
     \begin{subfigure}[b]{.7\textwidth}
-      \includegraphics[width=1.0\textwidth]{neighbour-fixed/round-communication-cost-vs-population}
+      \includegraphics[width=\textwidth]{neighbour-fixed/round-communication-cost-vs-population}
       \caption{Transactions are with fixed neighbours}
-      \label{fig:consensus-comms-fixed}
+      \label{fig:round-comms-fixed}
     \end{subfigure}%
     \begin{subfigure}[b]{0.7\textwidth}
-      \includegraphics[width=1.0\textwidth]{neighbour-random/round-communication-cost-vs-population}
+      \includegraphics[width=\textwidth]{neighbour-random/round-communication-cost-vs-population}
       \caption{Transactions are with random neighbours}
-      \label{fig:consensus-comms-random}
+      \label{fig:round-comms-random}
     \end{subfigure}%
   }
+  \caption{}
+  \label{fig:round-comms}
 \end{figure}
 
-\Cref{fig:consensus-comms} show the relationship between communication cost and population.
-These most important aspect is that these results show a linear increase.
-This reinforces our analytical result in~\Cref{sec:acs-complexity}.
-Further, note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
+\begin{figure}[tb]
+  \centering
+  \makebox[\linewidth][c]{%
+    \begin{subfigure}[b]{.7\textwidth}
+      \includegraphics[width=\textwidth]{neighbour-fixed/round-duration-vs-population}
+      \caption{Transactions are with fixed neighbours}
+      \label{fig:round-duration-fixed}
+    \end{subfigure}%
+    \begin{subfigure}[b]{0.7\textwidth}
+      \includegraphics[width=\textwidth]{neighbour-random/round-duration-vs-population}
+      \caption{Transactions are with random neighbours}
+      \label{fig:round-duration-random}
+    \end{subfigure}%
+  }
+  \caption{}
+  \label{fig:round-duration}
+\end{figure}
+
+\Cref{fig:round-comms} show the relationship between communication cost and population.
+The most important aspect is that these results show a linear increase.
+This reinforces our analytical result in~\Cref{sec:cons-complexity}.
+Note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
 the magnitude of the communication cost are similar.
-Both peak at about $1.4 \times 10^{10}$ bytes.
-This is expected because the consensus protocol is decoupled from the transaction protocol.
+Both peak at about 100 MB.
+This is expected because the consensus protocol is decoupled from the transaction protocol and the validation protocol.
+Finally, the rate for which the communication cost increases is higher when the number of facilitators are higher.
+This is also expected because the ACS algorithm has a $n^2$ term in its communication complexity.
 
 We are also interested in how communication costs translate to time.
-Hence, for the same experiment we use duration in seconds on the vertical axis instead of communication cost and plot the~\Cref{fig:consensus-duration}
-These show a similar trend as the communication cost graphs.
-Hence our assumption (from~\Cref{sec:acs-time-complexity}) that for every unit of communication there is some fixed overhead is valid.
+Hence, for the same experiment we also record the duration in seconds, which is shown on the vertical axis instead of communication cost and plot as before (\Cref{fig:consensus-duration}).
+Interestingly, the duration is not entirely linear.
+We attribute this behaviour to the extra time needed to hash the CP blocks in the consensus result to compute the luck value.
+Since if $N$ increases, every node must also perform more hash operations.
+These result do not conform to analytical result in~\Cref{sec:cons-duration}.
+Nevertheless, the difference is minor and there are ways to optimise the luck value computation.
+For example, the luck value can be computed by the facilitators and are sent with the consensus result.
+Then the non-facilitator nodes simply checks that they are correct.
 
@@
   \centering
-  \begin{subfigure}{\textwidth}
+  \makebox[\linewidth][c]{%
+  \begin{subfigure}[b]{0.7\textwidth}
     \includegraphics[width=\textwidth]{neighbour-fixed/tx-communication-cost-vs-population}
@@
     \label{fig:tx-comms-fixed}
-  \end{subfigure}
-
-  \begin{subfigure}{\textwidth}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.7\textwidth}
     \includegraphics[width=\textwidth]{neighbour-random/tx-communication-cost-vs-population}
@@
   \end{subfigure}
+  }
   \caption{Communication cost per verified transaction}
@@
 To verify the argument, we plot the relationship between the communication cost and population size in~\Cref{fig:tx-comms}.
-Again, we observe a linear relationship, which verified the argument.
+We observe a nearly linear relationship, which is due to the nearly linear relationship of the communication duration mentioned before.
+Again, we believe the difference is minor and it is possible to remove the extra overhead.
 
-However, note that there is a large difference in communication cost between the two modes of transaction.
+More interestingly, there is a large difference in communication cost between the two modes of transaction.
 When transacting with only one neighbour, the communication cost is low because only one agreed fragment needs to be communicated for every round in order to validate all transactions of that round.
@@
 
-It is interesting to note that some fluctuations exist in the graphs.
-This is likely due to the way how the communication cost are recorded.
-We maintain the total amount of data sent and received in a few variables.
-On every message that is sent or received, we increment the relevent variable by the size of the message.
-However, it is infeasible to log the values every time a message is sent or received as there is simply too much data for our build server to hold.
-Therefore we only log every 5 seconds.
-This does not always align with the time when a round finishes, thus some of the data may not represent reality accurately.
+Some fluctuations exist in~\Cref{fig:tx-comms-fixed}, this is due to our caching mechanism.
+We send validation requests at the same rate as transactions.
+Upon receiving an (remote) agreed fragment, the caching mechanism inspects all the transactions in the agreed fragment and attempts to verify as many as it can,
+rather than just the transaction in the original validation request.
+However, it may be the case that the agreed fragments arrive later than the validation request interval.
+Then it is possible to have sent two or more validation requests for the transaction in the same local agreed fragment.
+In this case, the remote would respond with two or more of the same agreed fragments,
+which results in extra (wasted) communication cost, and this is the source of the fluctuation seen in~\Cref{fig:tx-comms-fixed}.
+The result in~\Cref{fig:tx-comms-random} reinforces our argument because it is a lot more stable because for every transaction it is almost always the case that an agreed fragment is needed to validate it.
+% It is interesting to note that some fluctuations exist in the graphs.
+% This is likely due to the way how the communication cost are recorded.
+% We maintain the total amount of data sent and received in a few variables.
+% On every message that is sent or received, we increment the relevent variable by the size of the message.
+% However, it is infeasible to log the values every time a message is sent or received as there is simply too much data for our build server to hold.
+% Therefore we only log every 5 seconds.
+% This does not always align with the time when a round finishes, thus some of the data may not represent reality accurately.
 %  TODO this can be fixed
@@
 \section{Global Throughput}
+\begin{figure}[h]
+  \centering
+  \makebox[\linewidth][c]{%
+  \begin{subfigure}[b]{0.7\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-fixed/throughput-vs-population}
+    \caption{}
+    \label{fig:global-throughput-fixed}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.7\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-random/throughput-vs-population}
+    \caption{}
+    \label{fig:global-throughput-random}
+  \end{subfigure}
+  }
+  \caption{Global throughput}
+  \label{fig:global-throughput}
+\end{figure}
+
 The global throughput results are shown in~\Cref{fig:global-throughput}.
-Clearly, the throughput has a linear relationship with the population size.
+Evidently, the throughput has a linear relationship with the population size.
 This result is a strong indication of the horizontal scalability which we aimed to achieve.
-It also verified our analytical result in~\Cref{sec:global-throughput}.
+It also matches with our analytical result.
 
-More interestingly, is the magnitude in throughput, 
-and the difference in magnitude between the two modes of transaction.
-Firstly, recall that we fixed our $r_\text{tx}$ to 2, but how is it possible to have nearly 9000 transactions per second for 1200 nodes (which is about 7.5 transactions per second)?
+Note that the throughput decreases as the number of facilitators increases.
+This is due to the additional communication cost for running ACS with a high number of facilitators.
+That is, if the network is congested then the nodes may not have enough bandwidth to send validation responses timely.
+
+For~\Cref{fig:global-throughput-fixed},
+the magnitude of our throughput may not be self-evident at first glance.
+Recall that we fixed our $r_\text{tx}$ to 2, but how is it possible to have around 4800 transactions per second for 1200 nodes (4 TX/s)?
 This is due to the way validated transactions are calculated.
@@
 Hence, for every node, the TX blocks are created at 4 per second.
-Every TX block must be validated.
-So everyone sends 4 validation requests per second and also receives 4 per second.
-Thus during one second 8 validations are performed.
-The difference in magnitude is caused by the caching mechanism mentioned earlier.
-If an agreed fragment needs to be transmitted to validating every transaction then it puts a toll on the network infrastructure.
+Validation requests are sent at the same rate, which explains the magnitude.
+
+The difference in magnitude between \Cref{fig:global-throughput-fixed} and \Cref{fig:global-throughput-random} is also interesting.
+It is caused by the caching mechanism mentioned earlier.
+If an agreed fragment needs to be transmitted to validate every transaction then it puts a toll on the network infrastructure.
 The low transaction rate in~\Cref{fig:global-throughput-random} is caused by fact that the network infrastructure cannot keep up with our demand.
-We demonstrate this issue from a different angle in the next section.
+We demonstrate this issue from a different perspective in~\Cref{fig:backlog}.
+The graph is plotted by counting the number of transactions and the number of validated transactions every 5 seconds for one node running in a network of 1200 nodes and 32 facilitators.
+In~\Cref{fig:backlog-fixed}, the number of validated transaction changes as a step function.
+This means that transactions are validated in bursts, and the validation protocol can ``keep up'' with the transactions.
+Note that the horizontal ``lines'' where no new validated transactions are made are roughly 70 seconds, this matches with our result in~\Cref{fig:round-duration-fixed}.
+On the other hand in~\Cref{fig:backlog-random}, the validation protocol clearly cannot ``keep up'' with the rate which the transactions are made.
+As a result, the global throughput is lower when transacting with random nodes than only with neighbours.
 
@@
   \centering
-  \begin{subfigure}{\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-fixed/throughput-vs-population}
-    \label{fig:global-throughput-fixed}
+  \makebox[\linewidth][c]{%
+  \begin{subfigure}[b]{0.7\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-fixed/timeseries}
+    \caption{}
+    \label{fig:backlog-fixed}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.7\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-random/timeseries}
+    \caption{}
+    \label{fig:backlog-random}
   \end{subfigure}
-
-  \begin{subfigure}{\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-random/throughput-vs-population}
-    \label{fig:global-throughput-random}
-  \end{subfigure}
-  \caption{Global throughput}
-  \label{fig:global-throughput}
+  }
+  \caption{}
+  \label{fig:backlog}
 \end{figure}
-
-\section{Backlog}
-
-TODO
-
-\begin{figure}[h]
-  \centering
-  \begin{subfigure}{\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-fixed/timeseries}
-  \end{subfigure}
-
-  \begin{subfigure}{\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-random/timeseries}
-  \end{subfigure}
-\end{figure}

### 2017-07-13 thesis/chapters/ch1-intro.tex
Finish some TODOs
Thanks Quinten

@@
 However, in contrast to its original design, the internet and in particular the world wide web is becoming increasingly centralised.
-The domain name system, certificate authorities, to name a few, carry enormous responsibilities and are a central point of failure.
+The domain name system (root name server), certificate authorities (root CA), to name a few, carry enormous responsibilities and are a central point of failure.
 The  same can be said for many other services such are online marketplaces,
@@
 Which, for the first time, gave us a simple banking system in the form of a distributed ledger.
-It needs no central control but still incorruptable even if there are malicious parties that aim to undermine the system.
+It needs no central control but still incorruptable with high probability even if there are malicious parties that aim to undermine the system.
 We call such a ledger a blockchain.
@@
 Hence the transaction is inconsistent with respect to the observer.
-Bitcoin and many other blockchain systems solve the inconsistency problem with a consensus algorithm.
+Bitcoin and many of its derivatives (e.g. Litecoin\footnote{\url{litecoin.org}} and Dogecoin\footnote{\url{dogecoin.org}}) solve the inconsistency problem with a consensus algorithm.
 The goal of the algorithm is reach agreement on a set of valid values, e.g. transactions. 
@@
 Hence the name blockchain.
-Further, honest miners always work on the largest chain.
-Due to the difficulty of the puzzle,
+Further, due to the difficulty of the puzzle and that the majority of the parties are honest,
 it is unlikely for more than one blockchain to exist in the network for a long period of time.
@@
 Due to its success, people from many different disciplines began investigating in way to use blockchain technology.
-This includes finance~\cite{todo}, heath care~\cite{todo}, logistics~\cite{todo}, energy~\cite{todo} and so on.
+This includes finance~\cite{finance}, heathcare~\cite{healthcare}, logistics~\cite{supplychain} and energy~\cite{energy}.
 
@@
 their limitations also became apparent.
-Bitcoin has the infamous 7 transactions per second upper bound.
+Bitcoin has the infamous 7 transactions per second upper bound~\cite{vukolic2015quest}.
 This is due to the fact that blocks are fixed to 1 MB are only generated on average every 10 minutes.
@@
 Finally, we compare our system with other state-of-the-art blockchain systems and conclude in Chapters \ref{ch:related} and \ref{ch:conclusion} respectively.
-
-% However, not everyone is convinced on the practicality of blockchain systems.
-% Bram Cohen, the inventor of BitTorrent said
-% ``It's getting a lot more media attention than the actual impact it's having so far.''~\cite{cohen}
-% Forbes listed a few reasons on why we should be skeptical about blockchains~\cite{forbes},
-% some of them include its accessibility for the layperson and its scalability issues.

### 2017-07-13 thesis/chapters/ch2-problem.tex
Thanks Quinten

@@
 
-Blockchain technology has many advantages over centralised systems.
-Its decentralised and democratic nature removes central point of control or failure.
+The decentralised and democratic nature of blockchain technology removes central point of control or failure.
 However, as we described in the Introduction, it faces a genuine scalability problem.
@@
 \begin{displayquote}
-\emph{Is it possible to design a blockchain fabric that is fault tolerant, scalable and can reach global consensus?}
+\emph{How do we design a blockchain fabric that is fault tolerant, scalable and can reach global consensus?}
 \end{displayquote}
@@
 Typical blockchain systems are application specific.
-For example, the application in Bitcoin (any many others) is digital cash, the application in Namecoin is domain name registration~\cite{namecoin} and the application in Siacoin is cloud storage~\cite{siacoin}.
-Underneath these applications lies the blockchain fabric, which we call the consensus protocol, and it is abstract over the application.
-In the case of Bitcoin it is proof-of-work (PoW) because PoW is not concerned with the structure of the transactions,
+For example, the application in Bitcoin (any many of its derivatives) is digital cash, the application in Namecoin is domain name registration~\cite{namecoin} and the application in Siacoin is cloud storage~\cite{siacoin}.
+Underneath these applications lies the blockchain fabric which has the goal of reaching consensus on transactions.
+In the case of Bitcoin it is proof-of-work (PoW).
+It is not concerned with the structure of the transactions,
 it works by simply treating transactions as a blob of data and hashes them with the previous block.
-Hence, we wish to design an application neutral system.
-It should act as a backbone that provides the building blocks of any blockchain based applications.
-Also, it should be possible to build any kind of application on top our system.
-We do not impose on a consensus algorithm, as long as it satisfies the properties of atomic broadcast which we describe in~\Cref{sec:overview-cons}.
+Hence, we focus on blockchain fabric rather than any specific application.
+Because it is the scalability bottleneck and is neccessary for any blockchain based application.
+% Also, it should be possible to build any kind of application on top our system.
+% We do not impose on a consensus algorithm, as long as it satisfies the properties of atomic broadcast which we describe in~\Cref{sec:overview-cons}.
 
-In the remainder of this chapter we expand on our research question and visit each of our requirement in detail and discuss our approach.
+In the remainder of this chapter we expand on our research question and visit each of our requirement in detail.
 
@@
 
-The first objective is to reach consensus on a global \emph{state}.
-We emphasise the word state because in contrary to traditional blockchains,
-it does not neccessarily mean the set of all transactions that has ever happened.
-The state can be a verifiable representation of all the transactions.
-As an extreme example, a single digest of the complete history can represent the state.
-It is even verifiable as long as there are nodes that stores the history.
+% why consensus is important
+% The first objective is to reach consensus on a global \emph{state}.
+% We emphasise the word state because in contrary to traditional blockchains,
+% it does not neccessarily mean the set of all transactions that has ever happened.
+% The state can be a verifiable representation of all the transactions.
+% As an extreme example, a single digest of the complete history can represent the state.
+% It is even verifiable as long as there are nodes that stores the history.
 
-Traditional blockchain systems use PoW as the consensus algorithm,
+Early blockchain systems use PoW as the consensus algorithm,
 but consensus algorithm and especially Byzantine\footnote{Byzantine means participants can experience arbitrary failure, more on this in~\Cref{sec:acs-background}.}
@@
 
-Our approach is to combine the vast literature in Byzantine fault tolerance research with blockchain technology,
-and introduce a new consensus model to record the global state.
+The first objective is to reach consensus on a global \emph{state} by standing on the shoulder of giants.
+Having consensus is essential in blockchain systems.
+It stops many types of malicious activities because agreed state must have the consent of the honest nodes in the network.
+We emphasise the word state because in contrary to traditional blockchains,
+it does not neccessarily mean the set of all transactions that has ever happened.
+The state can be a verifiable representation of all the transactions.
+As an extreme example, a single digest of the complete history can represent the state.
+It is even verifiable as long as there are nodes that stores the history.
+
+% Our approach is to combine the vast literature in Byzantine fault tolerance research with blockchain technology,
+% and introduce a new consensus model to record the global state.
 
@@
 There is another aspect of fault tolerance that is transaction verifiability.
-Recall that it is not neccessary to reach consensus on all the transactions in order to record the state.
-But when using some digest (or a set of digets) to represent the state, it is also important for the transactions to be verifiable.
-Otherwise adversaries could simply store invalid transactions in the state and they would never be detected.
+All transactions must be eventually verifiable to maintain system integrity.
+Further, the validation result must be consistent.
+For example, if two different honest nodes attempts verify the same transaction,
+it cannot be the case that one nodes thinks that the transaction is valid but the other thinks it is invalid.
+
+% Recall that it is not neccessary to reach consensus on all the transactions in order to record the state.
+% But when using some digest (or a set of digets) to represent the state, it is also important for the transactions to be verifiable.
+% Otherwise adversaries could simply store invalid transactions in the state and they would never be detected.
 
@@
 BitTorrent~\cite{cohen2003incentives} is an example of such a system,
-where peers interact with eachother to exchange files without any global bottleneck.
+where peers interact with each other to exchange files without any global bottleneck.
 
@@
 \section{Limitations}
-Our work removes the transaction bound from the blockchain.
-However, the price to pay is that the Sybil attack~\cite{douceur2002sybil} prevention mechanism moves from the blockchain backbone to the application layer.
+Our work aims to significantly lower the transaction bound.
+However, the price to pay is that the Sybil attack~\cite{douceur2002sybil} prevention mechanism moves from the blockchain fabric to the application layer.
 For example, social network based Sybil defence mechanims use graph structure of real-world relationships~\cite{yu2006sybilguard}.

### 2017-07-13 thesis/chapters/ch3-model.tex
Thanks Quinten

@@
 with a modification of our prior work on TrustChain~\cite{trustchain}.
-The design is inspired by the essence the (early) Internet architecture as well as how transactions are performed in the real world.
+The design is inspired by (the distributed part of) the Internet architecture as well as how transactions are performed in the real world.
 
@@
 By the time the consensus protocol is done, the new CP block is added to whatever the state that the chain is in.
-It is not necessary to keep the chain immutable while the consensus protocol is running.
+It is not necessary to lock the chain while the consensus protocol is running and then unlock it afterwards.
 
@@
 Thus in no circumstance do we make timing assumptions.
-The adversary has fully control of the delivery schedule and the message ordering.
-But they are not allowed to drop messages\footnote{
+The adversary has fully control of the delivery schedule and the message ordering of all messages.
+But they are not allowed to drop messages except for their own\footnote{
     Reliability can be achieved in unreliable networks by resending messages or using some error correction code.
@@
 but if the node recovers, then the adversary must forget the private key.
-Otherwise the adversary can eventually learn all the private keys and sabotage the system.
+This is realistic because otherwise the adversary can eventually learn all the private keys and sabotage the system.
 Finally, we assume computational security.
-That is, the adversary can run polynomial-time algorithms but not exponential-time algorithms.
+That is, the adversary can run polynomial-time algorithms but not exponential-time algorithms efficiently.
 
@@
 \item $seq_u$ is the sequence number which should equal $i$.
-\item $txid$ is the transaction identifier, it should be generated using a cryptographically secure random number generator by the initiator of the transaction.
+\item $txid$ is the transaction identifier, it should be generated using a cryptographically secure pseudo-random number generator by the initiator of the transaction.
 \item $pk_v$ is the public key of the counterparty $v$.
@@
 \label{def:validation}
-\textbf{\emph{Consensus on transactions}}
+\textbf{\emph{Properties of the validation protocol}}
 
@@
 Every time the message is sent to $f$ random neighbours, $f$ is often called the fan-out.
-The upside of using epidemic protocol is a lower communication cost.
+The upside of using epidemic protocol is that the communication cost is distributed more evenly between nodes.
 This is especially true with a lazy push approach the node who just received a message would push the message ID to its $f$ random neighbours,
@@
 Another downside of some epidemic protocols (e.g. eager push) is that it is difficult to guarantee delivery.
-It is especially true when the parameters are not choosen correctly of if the network is not fully connected.
+It is especially true when the parameters are not choosen correctly in a network that is only partially connected (but every node is nevertheless reachable).
 If the delivery cannot be guaranteed, 

### 2017-07-13 thesis/chapters/ch4-analysis.tex
Thanks Quinten

@@
 First we show it has the desired properties.
-That is, the properties in \Cref{def:consensus} and \Cref{def:validation} should hold.
+That is, the properties of the consensus protocol and the validation protocol (\Cref{def:consensus} and \Cref{def:validation} respectively) should hold.
 Then we analyse the performance, especially the throughput,
@@
 Using Lemmas \ref{lemma:agreement}, \ref{lemma:validity}, \ref{lemma:fairness} and \ref{lemma:termination},
-it directly follows from Modus Poneus that these properties hold for $r + 1$.
+it directly follows from modus ponens that these properties hold for $r + 1$.
 Due to the principals of mathematical induction, these properties hold for all $r$.

### 2017-07-13 thesis/chapters/ch5-implementation.tex
Thanks Quinten

@@
 
-The prototype implementation can be found on GitHub\footnote{\url{https://github.com/kc1212/consensus-thesis-code}}.
+The prototype implementation can be found on GitHub.
+\begin{displayquote}
+\url{https://github.com/kc1212/consensus-thesis-code}
+\end{displayquote}
 It is written in the event driven paradigm, using the Python programming language\footnote{\url{https://www.python.org/}} and
@@
 
-\begin{figure}[h]
+\begin{figure}[tb]
   \centering

### 2017-07-14 thesis/chapters/ch3-model.tex
Fix up implementation section

@@
 Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks from every other node and use them as the proposal.
-The facilitators are elected using a \emph{luck value}, which is computed using $H(\C_r || pk_u)$,
+The facilitators are elected using a \emph{luck value}, which is computed using $\textsf{H}(\C_r || pk_u)$,
 where $\C_r$ is the consensus result 
@@
 
+\subsection{Caching agreed fragments}
+
 \subsection{Global probabilistic fork detection}

### 2017-07-14 thesis/chapters/ch5-implementation.tex
Fix up implementation section

@@
 
-Henceforth, we evaluate out system experimentally and compare the results with the theoretical analysis.
+To build confidence of the horizontal scalability claim from our analytical results,
+we must verify the same property experimentally.
+Hence, we give an open source prototype implementation of our system.
+The nodes are capable of forming a network (tested up to 1500 nodes) and send transactions to each other,
+which allows us to confirm our analytical claims.
+
 We begin this chapter by a description of the implementation in \Cref{sec:implementation}.
-Then, we move on to describing our experimental setup in \Cref{sec:experimental-setup}.
-Finally, \Cref{sec:evaluation} presents our experimental results and our evaluation.
-Our experiment primarily focuses on the consensus duration and the throughput.
+Then, we move on to describing our experimental setup in \Cref{sec:experimental-setup} which includes a discussion on the system parameters and the physical infrastructure which we use.
+Finally, \Cref{sec:comms-cost-experiment} and onwards follows the same structure as the Performance analysis (\Cref{sec:performance-analysis}),
+where we analyse the performance experimentally rather than analytically.
 
@@
 \end{displayquote}
-It is written in the event driven paradigm, using the Python programming language\footnote{\url{https://www.python.org/}} and
-the Twisted\footnote{\url{https://twistedmatrix.com/}} library for networking.
+It is written in the event driven paradigm, using the Python programming language\footnote{\url{https://www.python.org/}}.
+We use the Twisted\footnote{\url{https://twistedmatrix.com/}} library for networking.
 
@@
         it hits a bottleneck in extreme cases which helps us understand the limitations of our design.
-  \item The number of facilitators $n \in \{ 12, 16, \dots, 32\}$.
+  \item The number of facilitators $n \in \{4, 8, \dots, 32\}$.
         Unfortunately, the maximum $n$ is 32 because the limitation in liberasurecode mentioned in \Cref{sec:implementation}.
-  \item The population size $N \in \{200, \dots, 1200\}$.
+  \item The population size $N \in \{200, 300, \dots, 1200\}$.
         $N$ stops at 1200 is due to our physical setup, which we describe next.
   \item The two modes of transaction.
-        The first mode or the ideal mode is that nodes only transact with their immediate neighbour. 
+        The first mode is the \emph{fixed-neighbour} mode where nodes only transact with their immediate neighbour which are predetermined. 
         This minimises the data volume per validated transaction because agreed fragment can be cached.
-        The second mode is in the other extreme, where every transaction is with a random node out of the $N$ nodes in the system,
-        thus the agreed fragment is unlikely to be cached.
+        The second mode is in the other extreme which we call the \emph{random-neighbour} mode,
+        where every transaction is with a random node out of the $N$ nodes in the system.
+        It is unlikely that the agreed fragments can be reused.
 \end{enumerate}
@@
 The experiment is run on the DAS-5 (The Distributed ASCI Supercomputer 5\footnote{\url{https://www.cs.vu.nl/das5/}}).
-From now on, we use "machines" to refer to DAS-5 nodes and nodes to refer to a running instance in our system.
+From now on, we use ``machines'' to refer to DAS-5 nodes and ``nodes'' to refer to a running instance in our system.
 On DAS-5 we use up to 30 machine, for each machine we use 40 nodes.
@@
 but it requires additional network interface which is something we do not control on the DAS-5.
+Nevertheless, running the system with 1200 nodes gives a good indication of its scalability properties as we show later.
 
@@
 
-\section{Communication cost for a round}
+Finally, since Bitcoin transactions are approximately 500 bytes~\cite{txsize},
+we use a uniformly random transaction size sampled between 400 and 600 bytes.
 
-The remainder of this chapter follows the same structure as the performance analysis in~\Cref{sec:performance-analysis},
-where we check our theoretical evaluation with our experimental results.
+\section{Communication cost for the consensus protocol}
+\label{sec:comms-cost-experiment}
+
+The remainder of this chapter follows the same structure as the performance analysis in~\Cref{sec:performance-analysis}.
+We check our analytical results experimentally and again demonstrate horizontal scalability.
 
@@
       \includegraphics[width=\textwidth]{neighbour-fixed/round-communication-cost-vs-population}
-      \caption{Transactions are with fixed neighbours}
+      \caption{Transactions are with fixed neighbours.}
       \label{fig:round-comms-fixed}
@@
       \includegraphics[width=\textwidth]{neighbour-random/round-communication-cost-vs-population}
-      \caption{Transactions are with random neighbours}
+      \caption{Transactions are with random neighbours.}
       \label{fig:round-comms-random}
@@
   }
-  \caption{}
+  \caption{Communication cost for the consensus protocol per round increases linearly with population size.
+  The error bars are larger for higher population size or higher number of facilitators is because rounds take longer thus they are repeated fewer times.}
   \label{fig:round-comms}
@@
       \includegraphics[width=\textwidth]{neighbour-fixed/round-duration-vs-population}
-      \caption{Transactions are with fixed neighbours}
+      \caption{Transactions are with fixed neighbours.}
       \label{fig:round-duration-fixed}
@@
       \includegraphics[width=\textwidth]{neighbour-random/round-duration-vs-population}
-      \caption{Transactions are with random neighbours}
+      \caption{Transactions are with random neighbours.}
       \label{fig:round-duration-random}
@@
   }
-  \caption{}
+  \caption{Round duration do not increase linearly with the population size. 
+  This is likely due to the additional hashing operation required for larger consensus result.}
   \label{fig:round-duration}
@@
 
-\Cref{fig:round-comms} show the relationship between communication cost and population.
+\Cref{fig:round-comms} show the relationship between communication cost of the consensus protocol per round and population size.
 The most important aspect is that these results show a linear increase.
@@
 We are also interested in how communication costs translate to time.
-Hence, for the same experiment we also record the duration in seconds, which is shown on the vertical axis instead of communication cost and plot as before (\Cref{fig:consensus-duration}).
+Hence, for the same experiment we also record the duration in seconds and the result is shown in~\Cref{fig:round-duration}.
 Interestingly, the duration is not entirely linear.
@@
 
-\section{Communication Cost for Transaction and Validation}
+\section{Communication cost for transaction and validation protocols}
 
@@
     \includegraphics[width=\textwidth]{neighbour-fixed/tx-communication-cost-vs-population}
-    \caption{Transactions are with fixed neighbours}
+    \caption{Transactions are with fixed neighbours.}
     \label{fig:tx-comms-fixed}
@@
     \includegraphics[width=\textwidth]{neighbour-random/tx-communication-cost-vs-population}
-    \caption{Transactions are with random neighbours}
+    \caption{Transactions are with random neighbours.}
     \label{fig:tx-comms-random}
@@
   }
-  \caption{Communication cost per verified transaction}
+  \caption{Communication cost per verified transaction has similar (near linear) trend as~\Cref{fig:round-duration}.
+  Fluctuation for the fixed-neighbour mode exists because the cache mechanism is unpredictable.}
   \label{fig:tx-comms}
@@
 
-Recall that in~\Cref{sec:communication-cost-for-tx} we argued that the communication cost per verified transaction is of $O(N)$.
-To verify the argument, we plot the relationship between the communication cost and population size in~\Cref{fig:tx-comms}.
-We observe a nearly linear relationship, which is due to the nearly linear relationship of the communication duration mentioned before.
+We argued that the communication cost per verified transaction is of $O(N)$ (\Cref{sec:communication-cost-for-tx}).
+To verify the argument, we plot the relationship between the communication cost of for every validated transaction and population size in~\Cref{fig:tx-comms}.
+We observe a near linear relationship, which is due to the near linear relationship of the communication duration mentioned before.
 Again, we believe the difference is minor and it is possible to remove the extra overhead.
@@
 However, it may be the case that the agreed fragments arrive later than the validation request interval.
-Then it is possible to have sent two or more validation requests for the transaction in the same local agreed fragment.
+Then it is possible to have sent two or more validation requests for some transactions in the same local agreed fragment.
 In this case, the remote would respond with two or more of the same agreed fragments,
 which results in extra (wasted) communication cost, and this is the source of the fluctuation seen in~\Cref{fig:tx-comms-fixed}.
-The result in~\Cref{fig:tx-comms-random} reinforces our argument because it is a lot more stable because for every transaction it is almost always the case that an agreed fragment is needed to validate it.
+The result in~\Cref{fig:tx-comms-random} reinforces our argument.
+It is a lot more stable because for every transaction it is almost always the case that an agreed fragment is needed to validate it.
 % It is interesting to note that some fluctuations exist in the graphs.
@@
     \includegraphics[width=\textwidth]{neighbour-fixed/throughput-vs-population}
-    \caption{}
+    \caption{Transactions are with fixed neighbours.}
     \label{fig:global-throughput-fixed}
@@
     \includegraphics[width=\textwidth]{neighbour-random/throughput-vs-population}
-    \caption{}
+    \caption{Transactions are with random neighbours.}
     \label{fig:global-throughput-random}
@@
 
-The global throughput results are shown in~\Cref{fig:global-throughput}.
+Finally, the global throughput results are shown in~\Cref{fig:global-throughput}.
 Evidently, the throughput has a linear relationship with the population size.
@@
 the magnitude of our throughput may not be self-evident at first glance.
-Recall that we fixed our $r_\text{tx}$ to 2, but how is it possible to have around 4800 transactions per second for 1200 nodes (4 TX/s)?
+Recall that we fixed our $r_\text{tx}$ to 2, but how is it possible to have around 4800 transactions per second for 1200 nodes (which is 4 TX/s)?
 This is due to the way validated transactions are calculated.
-Transactions are between two parties, hence if every node makes two transactions per seoncd,
+Transactions are between two parties, hence if every node makes two transactions per second,
 every node also expects to receive two transactions per second.
@@
 It is caused by the caching mechanism mentioned earlier.
-If an agreed fragment needs to be transmitted to validate every transaction then it puts a toll on the network infrastructure.
+If a new agreed fragment needs to be transmitted to validate every transaction then it puts a toll on the network infrastructure.
 The low transaction rate in~\Cref{fig:global-throughput-random} is caused by fact that the network infrastructure cannot keep up with our demand.
-We demonstrate this issue from a different perspective in~\Cref{fig:backlog}.
+
+We demonstrate the bottleneck issue from a different perspective in~\Cref{fig:backlog}.
 The graph is plotted by counting the number of transactions and the number of validated transactions every 5 seconds for one node running in a network of 1200 nodes and 32 facilitators.
@@
     \includegraphics[width=\textwidth]{neighbour-fixed/timeseries}
-    \caption{}
+    \caption{Transactions are with fixed neighbours.}
     \label{fig:backlog-fixed}
@@
     \includegraphics[width=\textwidth]{neighbour-random/timeseries}
-    \caption{}
+    \caption{Transactions are with random neighbours.}
     \label{fig:backlog-random}
@@
   }
-  \caption{}
+  \caption{A limitation of our system is when nodes are transacting with many random nodes,
+  then the network cannot keep up with the large number of agreed fragments that need to be communicated between nodes.
+  Hence the slope of validated transactions grows at a slower rate in~\Cref{fig:backlog-random}.}
   \label{fig:backlog}

### 2017-07-17 thesis/chapters/ch3-model.tex
Write more design variations

@@
-\chapter{System Architecture}
+\chapter{System architecture}
 \label{ch:model}
@@
 
-\section{System Overview}
+\section{System overview}
 \label{sec:system-overview}
@@
 
-\subsection{Consensus Protocol}\label{sec:overview-cons}
+\subsection{Consensus protocol}\label{sec:overview-cons}
 
@@
 
-\subsection{Transaction and Validation}
+\subsection{Transaction and validation}
 The TX protocol is a simple request and response protocol.
@@
 
-\subsection{Combined Protocol}
+\subsection{Combined protocol}
 \label{sec:combined-protocol}
@@
 
-\section{Model and Assumptions}
+\section{Model and assumptions}
 \label{sec:model-assumptions}
@@
 
-\subsection{Transaction Block}
+\subsection{Transaction block}
 The TX block is a six-tuple, i.e $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$.
@@
 
-\subsection{Checkpoint Block}
+\subsection{Checkpoint block}
 
@@
 
-\subsection{Consensus Result}
+\subsection{Consensus result}
 \label{sec:consensus-result}
@@
 
-\subsection{Chain Properties}
+\subsection{Chain properties}
 Here we define a few important properties which results from the interleaving nature of CP and TX blocks.
@@
 
-\section{Consensus Protocol}
+\section{Consensus protocol}
 \label{sec:cons-protocol}
@@
 
-\subsection{Background on Asynchronous Subset Consensus}
+\subsection{Background on asynchronous subset consensus}
 \label{sec:acs-background}
@@
 
-\subsection{Bootstrap Phase}
+\subsection{Bootstrap phase}
 \label{sec:bootstrap}
@@
 
-\subsection{Consensus Phase}
+\subsection{Consensus phase}
 \label{sec:consensus-phase}
@@
 
-\subsection{Relationship with $\alpha$-Synchroniser}
+\subsection{Relationship with $\alpha$-synchroniser}
 TODO
 
-\section{Transaction Protocol}
+\section{Transaction protocol}
 \label{sec:tx-protocol}
@@
 
-\section{Validation Protocol}
+\section{Validation protocol}
 \label{sec:vd-protocol}
@@
 
-\subsection{Validity Definition}
+\subsection{Validity definition}
 A transaction can be in one of three states in terms of validity---\emph{valid}, \emph{invalid} and \emph{unknown}.
@@
 
-    \If{$|F_{v, j}| > L$}
-        \State \Return \emph{unknown}
-    \EndIf \Comment fragment not too long
-    \State
+    % \If{$|F_{v, j}| > L$}
+    %     \State \Return \emph{unknown}
+    % \EndIf \Comment fragment not too long
+    % \State
 
@@
 
-\subsection{Validation Protocol}
+\subsection{Validation protocol}
 With the validity definition, we are ready to construct a protocol for determining the validity of transactions.
@@
 In this section, 
-we explore a few design variations which we can make if we relax our model.
-These variations do not have the same theoretical elegance as our original model.
-But they enable better performance, allow us to apply our design in the fully permissionless setting and improves privacy.
+we explore a few design variations which we can make, some of them require a relaxed version of our original model.
+They enable better performance, allow us to apply our design in the fully permissionless setting and improves privacy.
 
@@
 
-\subsection{Privacy preserving validation protocol}
-TODO
+\subsection{Privacy preserving validation protocol using compact blocks}
+\label{sec:compact}
+Our approach already has privacy preserving features in comparison to early blockchain systems.
+That is, the transactions for each node are only revealed during the validation protocol.
+Hence if two nodes never directly or indirectly interact with each other,
+their transactions are never revealed.
 
-\subsection{Caching agreed fragments}
+We can take our privacy-preserving property one step further by introducing another level of hash pointer indirection.
+The result is shown in~\Cref{fig:compact}.
+\begin{figure}
+    \includegraphics[width=0.7\textwidth]{compact}
+    \centering
+    \caption{The chain on the left represent direct chaining, where the digest in ``Prev'' is simply the digest of the previous block.
+    The chain on the right uses compact blocks, represented by the smaller squares.
+    Compact blocks also form a chain as before, but they each have a hash pointer to the full block, identified by ``Seq'' or ``Digest''.}
+    \label{fig:compact}
+\end{figure}
 
-\subsection{Global probabilistic fork detection}
-TODO
+Concretely, we introduce an additional block type,
+namely compact block.
+Such blocks only have three attributes,
+\begin{enumerate}
+\item Seq---the sequence number its corresponding block,
+\item Digest---the digest of its corresponding block, and
+\item Prev---the digest of the previous compact block.
+\end{enumerate}
+Each compact block has a corresponding full block (either a CP block or a TX block).
+The relationship is uniquely identified with Seq or Digest.
+Recall that our original validation protocol requires the nodes to send the full agreed fragment.
+With compact blocks, it is only necessary to send the compact version of the agreed fragment.
+The validation then proceeds in a similar fashion,
+provided that the pair of the to-be-validated TX block is known.
+
+The space saving of this approach of course depends on the size of the full blocks.
+If the full blocks are on average 500 bytes
+(which is the typical size of Bitcoin transactions ~\cite{txsize}),
+and the compact blocks are $32 + 32 + 8 = 72$ bytes
+(SHA256 digests are 32 bytes each and we use a 64 bit integer to represent the sequence number),
+then we have a 86\% saving in communication cost.
+
+%Using compact blocks, it is possible to reduce network communication costs in the validation protocol but also protect transaction privacy.
+
+\subsection{Optimising validation protocol using cached agreed fragments}
+\label{sec:caching}
+One more way to improve the efficiency of the validation protocol is to use a single agreed fragment to validate multiple transaction.
+Concretely, upon receiving an agreed fragment from node $A$,
+rather than validating a single transaction,
+we attempt to validate all transactions in the unknown state performed with $A$ in that fragment.
+
+For a node, the benefit of this technique is maximised when it only transacts with one other node.
+In this case, the communication of one fragment is sufficient to validate all transactions in that fragment.
+In the opposite extreme, if every transaction that the node makes is with another unique node,
+then the caching mechanism would have no effect.
+
+\subsection{Global fork detection}
+The validation algorithm guarantees that there are no forks within a single agreed fragment.
+This is sufficient for some applications such as proving the existance of some information.
+However, applications such as digital cash where every block depends on some previous block,
+then our scheme is not suitable.
+
+There are a variety of approaches to do global fork detection.
+First and the easiest solution is to simply ask for every agreed fragment from the one containing the genesis block up to the one that the to-be-proven TX block is in.
+If the hash pointers are correct and all the CP blocks are in consensus, then the verifier can be sure that there are no forks.
+We use this approach in our prior work on Implicit Consensus~\cite{implicitconsensus}.
+It may sound inefficient at first, but nodes can employ caching to minimise communication costs.
+In our work we call this ``spontaneous sharding''.
+
+The second approach is probabilistic but with only a constant communication overhead.
+For a node, observe that if every of its agreed fragment has a transaction with a honest node,
+then the complete chain is effectively validated in a distributed maner.
+The only way for an attacker to make a fork is to make sure that the agreed fragment containing the fork has no transactions with honest nodes.
+We claim such malicious behaviour is prevented probabilisticaly using a challenge-response protocol as follows.
+Suppose node $A$ wish to make a transaction with node $B$.
+$A$ first sends a challenge to $B$ asking it to proof that it holds a valid agreed fragment between some consensus round specified by $A$.
+If $B$ provides a correct proof, then they run the transaction protocol as usual.
+If $B$ provides an invalid proof or refuses to respond,
+then $A$ would refuse to make the transaction.
+The probability that an honest node catches out a malicious node is 
+$$p = \frac{f}{r},$$
+where $f$ is the number of bad agreed fragments and $r$ is the latest round number.
+If there are more nodes (say $n$) trying to make transactions with the malicious node,
+then the probability that the malicious node does gets caught at least once follows a binomial distribution
+$$\Pr[X > 0] = \sum_{k = 1}^{n}\binom{n}{k}p^k (1 - p)^{n - k}.$$
+
+% subtlety when fork happens on CP block
 

### 2017-07-17 thesis/chapters/ch4-analysis.tex
Fix up implementation section
Write more design variations

@@
-\chapter{Analysis of Correctness and Performance}
+\chapter{Analysis of correctness and performance}
 \label{ch:analysis}
@@
 
-\section{Correctness in the Presense of Faults}
+\section{Correctness in the presense of faults}
 Our first objective is to show that \Cref{def:consensus} holds for our consensus protocol.
@@
 
-\subsection{Analysis of the Consensus Protocol}
+\subsection{Analysis of the consensus protocol}
 
@@
 
-\subsection{Correctness of Validation}
+\subsection{Correctness of validation}
 The consensus protocol (on CP blocks and facilitators) is the backbone for consensus on transactions.
@@
 
-\subsection{Impossibility of Liveness}
+\subsection{Impossibility of liveness}
 While \Cref{theorem:validation-agreement} is a major result that allows significantly improved performance over traditional blockchain systems,
@@
 
-\section{Performance}
+\section{Performance analysis}
 \label{sec:performance-analysis}
@@
 
-\subsection{Communication Cost for Transactions}
+\subsection{Communication cost for transactions}
 \label{sec:communication-cost-for-tx}
@@
 
-\subsection{Global Throughput}
+\subsection{Global throughput}
 \label{sec:global-throughput}
@@
 
-\section{Effect of A Highly Adverserial Environment}
+\section{Effect of a highly adverserial environment}
 

### 2017-07-17 thesis/chapters/ch5-implementation.tex
Write more design variations

@@
-\chapter{Implementation and Experimental Results}
+\chapter{Implementation and experimental results}
 \label{ch:implementation}
@@
 \end{displayquote}
+It implements the three protocols and the Extended TrustChain discussed in~\Cref{ch:model}.
+We also implement two optimisations---privacy preserving validation protocol using compact blocks (\Cref{sec:compact})
+and optimised validation protocol using cached agreed fragments (\Cref{sec:caching}).
 It is written in the event driven paradigm, using the Python programming language\footnote{\url{https://www.python.org/}}.
@@
 
-\section{Experimental Setup}
+\section{Experimental setup}
 \label{sec:experimental-setup}
@@
 
-\section{Global Throughput}
+\section{Global throughput}
 \begin{figure}[h]

### 2017-07-17 thesis/chapters/ch6-related-work.tex
Write more design variations

@@
-\chapter{Related Work}
+\chapter{Related work}
 \label{ch:related}
@@
 
-\section{Classical Blockchain Systems}
+\section{Early blockchain systems}
 This category represent systems with a probabilistic probabilistic consensus algorithm.
@@
 
-\section{Offchain Transactions}
+\section{Offchain transactions}
 Offchain transactions make use of the fact that, if two or more parties frequently make transactions,
@@
 
-\section{Permissioned Systems}
+\section{Permissioned systems}
 
@@
 
-\section{Hybrid Systems}
+\section{Hybrid systems}
 Hybrid systems are very recent inventions.
@@
 
-\section{Blockchains Without Global Consensus}
+\section{Blockchains without global consensus}
 

### 2017-07-18 thesis/chapters/ch3-model.tex
Write synchroniser

@@
 
-\subsection{Relationship with $\alpha$-synchroniser}
-TODO
+Our protocol has some similarities with synchronisers~\cite[Chapter 10]{podc} because it is effectively a technique to introduce synchrony in an asynchronous environment.
+If we consider the facilitators is a collective authority, then it can be seen as a synchroniser that sends pulse messages (in the form of \texttt{cons\_msg} and \texttt{cons\_sig}) to indicate the start of a new clock pulse.
+Every node then sends a completion messages (in the form of \texttt{cp\_msg}) to the collective authority to indicate that they are ready to start a new pulse.
 
@@
 
-\subsection{Use epidemic protocol instead of broadcast protocol}
+\subsection{Using epidemic protocol to reduce communication cost}
 One of the final steps in our consensus protocol is to broadcast the consensus result and signatures to every other node (\Cref{sec:consensus-phase}).

### 2017-07-18 thesis/chapters/ch7-conclusion.tex
Write synchroniser

\chapter{Conclusion}
\label{ch:conclusion}

\section{Future work}

### 2017-07-26 presentation/presentation.tex
Update presentation

--- 2017-03-16 presentation.tex
@@
 
+\usepackage{amsfonts}
 \usepackage{graphicx}
+\graphicspath{{../figures/}}
+\usepackage[noend]{algpseudocode}
+\usepackage{algorithm}
 
-\title{Presentation Title}
+\newcommand{\A}{\mathcal{A}}
+\newcommand{\C}{\mathcal{C}}
+\renewcommand{\P}{\mathcal{P}}
+\newcommand{\F}{\mathcal{F}}
+\newcommand{\E}{\mathcal{E}}
+
+\algnewcommand\algorithmicupon{\textbf{Upon}}
+\algnewcommand\Upon{\item[\algorithmicupon]}
+
+\algnewcommand\algorithmicsend{\textbf{Send}}
+\algnewcommand\Send{\item[\algorithmicsend]}
+
+\title{Blockchain Consensus Protocol with Horizontal Scalability}
 
 % A subtitle is optional and this may be deleted
-\subtitle{Optional Subtitle}
+% \subtitle{Optional Subtitle}
 
@@
 % \date{Conference Name, 2013}
-\date{2017}
+\date{\today}
 % - Either use conference name or its abbreviation.
@@
 % the beginning of each subsection:
-\AtBeginSubsection[]
+\AtBeginSubsection[] % Do nothing for \subsection*
 {
-  \begin{frame}<beamer>{Outline}
-    \tableofcontents[currentsection,currentsubsection]
+  % do nothing
+}
+
+\AtBeginSection[] % Do nothing for \section*
+{
+  \begin{frame}<beamer>
+  \frametitle{Outline}
+  \tableofcontents[currentsection]
   \end{frame}
@@
 % and table of contents.
-\section{Background}
-
-\subsection{TrustChain}
-
-\begin{frame}{TrustChain}{}
-
-  \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{figures/trustchain-good}
-  \centering
-  \caption{Every node has a chain. \emph{TX block} is a six-tuple: $t_{i,j} =
-    (\texttt{h}(b_{i,j-1}), h_s, h_r, s_s, s_r, m)$, one transaction results in
-    two TX blocks---\emph{a pair}.}
-  \end{figure}
-
-\end{frame}
-
-\begin{frame}{TrustChain}{}
-
-  \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{figures/trustchain-bad}
-  \centering
-  \caption{Fork is two correctly signed TX blocks that belong to the same sender
-    and has the same hash pointer but involve different receivers. Only one TX
-    block may be in consensus.}
-  \end{figure}
-
-\end{frame}
-
-\begin{frame}{TrustChain}
+\section{Introduction}
+\subsection{Motivation}
+\begin{frame}{\subsecname}{}
   \begin{itemize}
-    \item Everyone has their own chain
-    \item Transactions are on arbitrary data $m$
-    % \item Transactions make the chains intertwined
-    \item Transactions are irrefutable due to hash pointers
-    \item No consensus (my thesis)
+    \item Blockchain systems offer an alternative to central authorities for the first time
+    \item Market cap and trade volume figures indicate they are here to stay
+    \item Early blockchain systems are not scalable (7 TX/s for Bitcoin)
+    \item Parameter tuning leads to centralisation
   \end{itemize}
@@
 
-\section{My Thesis}
-\subsection{TrustChain with Checkpoints}
-\begin{frame}{TrustChain with Checkpoints}
-
-  \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{figures/trustchain-bad-cp}
-  \centering
-  \caption{CP block is a five-tuple: $c_{i,j} = (\texttt{h}(b_{i,j-1}),
-    \texttt{h}(\mathcal{C}_r), r, p, s)$, $\mathcal{C}_r$ is the consensus
-    result at round $r$, $p =$ promoter indicator, $s =$ signature.}
-  \end{figure}
-
+\subsection{Research question}
+\begin{frame}{\subsecname}
+    How do we design a \emph{blockchain consensus protocol} that is \emph{fault tolerant},
+    \emph{scalable} and can reach \emph{global consensus?}
 \end{frame}
 
-\subsection{Protocol Overview}
-\begin{frame}{Protocol Overview}
+\subsection{Intuition}
+\begin{frame}{\subsecname}{}
+  \begin{itemize}
+    \item A restaurant owner does not report all of its transactions with an central authority
+    \item Ocassionally a customer may leave without paying and this event is reported to a central authority
+    \item Our blockchain system achieves scalability using the same idea
+  \end{itemize}
+\end{frame}
+
+\section{System architecture}
+\subsection{System model}
+\begin{frame}{\subsecname}
+  \begin{itemize}
+    \item Population size is $N$
+    \item $n$ nodes are facilitators, $t$ nodes are malicious (Byzantine)
+    \item $n \ge 3t + 1$
+    \item $N \ge n + t$
+    \item Purely asynchronous channels with eventual delivery
+    \item Public key infrastructure
+    \item Random oracle model
+    \item Application neutural
+  \end{itemize}
+\end{frame}
+
+\subsection{Architecture overview}
+\begin{frame}{\subsecname}
+  \begin{figure}[h]
+  \includegraphics[width=0.7\textwidth]{architecture}
+  \centering
+  \end{figure}
+\end{frame}
+
+\subsection{Extended TrustChain}
+\begin{frame}{\subsecname}
+  \begin{itemize}
+    \item Everyone has their own chain
+    \item Two types of blocks, transaction (TX) blocks and checkpoint (CP) blocks
+    \item A transaction involves two parties and results in two TX blocks
+    \item A CP block captures the chain state
+    \item TX and CP blocks are chained together using hash pointers
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{\subsecname}
+  \begin{figure}[h]
+  \includegraphics[width=0.7\textwidth]{trustchain-bad-cp}
+  \centering
+  \caption{TX block is a six-tuple: $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), i, txid, pk_v, m, sig_u \rangle$.
+    CP block is a five-tuple: $c_{u, i} = \langle \textsf{H}(b_{u, i-1}), i, \textsf{H}(\C_r), r, sig_u \rangle$.}
+  \end{figure}
+\end{frame}
+
+\subsection{Consensus protocol}
+\begin{frame}{\subsecname overview}
   \begin{enumerate}
-    \item $n$ lucky nodes are selected at random to act as promoters.
-    \item Promoters run a BFT (Byzantine Fault Tolerant) consensus algorithm to
-      agree on a set of CP blocks.
-    \item Disseminate the consensus result (the CP blocks).
-    \item Repeat for next round.
-    \item Any interested node can validate that their transaction.
+    \item In round $r$, $n$ out of $N$ lucky nodes are selected at random to act as facilitators
+    \item Facilitators run a BFT (Byzantine Fault Tolerant) consensus algorithm to
+      agree on a set of CP blocks
+    \item Disseminate the consensus result $\C_r$, i.e. the CP blocks
+    \item Repeat
   \end{enumerate}
@@
 
-\subsection{Promoter Registration}
-\begin{frame}{Promoter Registration}
-
-  \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{figures/trustchain-1}
-  \centering
-  \caption{We start in the state where $\mathcal{C}_{r - 1}$ has just been agreed but
-    not yet propagated.}
-  \end{figure}
-
+\begin{frame}{\subsecname properties}
+\label{def:consensus}
+$\forall r \in \mathbb{N}$, the following properties must hold.
+\begin{itemize}
+    \item \emph{Agreement}:
+        If one correct node outputs a list of facilitators $\F_r$,
+        then every node outputs $\F_r$
+    \item \emph{Validity}:
+        If any correct node outputs $\F_r$, then 
+            \begin{enumerate}
+                \item $|\C_r| \ge N - t$ must hold for the $\C_r$ which was used to create $\F_r$,
+                \item $\F_r$ must contain at least $n - t$ honest nodes and
+                \item $|\F_r| = n$.
+            \end{enumerate}
+    \item \emph{Fairness}:
+        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
+    \item \emph{Termination}:
+        Every correct node eventually outputs some $\F_r$.
+\end{itemize}
 \end{frame}
 
-\begin{frame}{Promoter Registration}
-
+\begin{frame}{\subsecname}
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{figures/trustchain-2}
+  \includegraphics[width=0.7\textwidth]{trustchain-1}
   \centering
-  \caption{Nodes receive consensus result and set promoters indicator $p$, then
-    send the new CP blocks to promoters of round $r$.}
+  \caption{Suppose we are in a state where $\mathcal{C}_{r - 1}$ has just been agreed by some facilitators but not yet propagated.}
   \end{figure}
-
 \end{frame}
 
-\begin{frame}{Promoter Registration}
-
+\begin{frame}{\subsecname}
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{figures/trustchain-3}
+  \includegraphics[width=0.7\textwidth]{trustchain-2}
   \centering
-  \caption{Transactions carry on as usual in round $r$. Note that our CP blocks
-    (round $r - 1$) has not reached consensus yet.}
+  \caption{Nodes receive consensus result $\mathcal{C}_{r - 1}$,
+    compute the new facilitators $\mathcal{F}_{r-1}$,
+    and send out the new CP blocks.}
   \end{figure}
-
 \end{frame}
 
-\begin{frame}{Promoter Registration}
-
+\begin{frame}{\subsecname}
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{figures/trustchain-4}
+  \includegraphics[width=0.7\textwidth]{trustchain-3}
   \centering
-  \caption{CP blocks at round $r-1$ should be in $\mathcal{C}_r$. If we're
-    lucky: $\texttt{h}(k_i || c_{i,j}) < T$, then we're responsible for
-    consensus of round $r + 1$.}
+  \caption{Transactions carry on as usual in round $r$,
+  while facilitators are trying to reach consensus on the new CP blocks concurrently.}
   \end{figure}
-
 \end{frame}
 
-\subsection{Consensus}
-\begin{frame}{Consensus}
-  \begin{enumerate}
-    \item Nodes send CP blocks to the promoters.
-    \item The promoters' identities are encoded in the consensus result.
-    \item Promoters run the some asynchronous BFT consensus algorithm to agree on
-      a set of CP blocks---$\mathcal{C}_r$.
-    \item $n \ge 3t + 1$ is the optimal for BFT consensus.
-    \item $\mathcal{C}_r$ and the signatures are disseminated.
-    \item Nodes create new CP blocks when they receive $t + 1$ good signatures and
-      $\mathcal{C}_r$.
-  \end{enumerate}
-
-  \begin{theorem}
-    Assuming the promoters satisfy $n \ge 3t + 1$. The promoter registration and
-    the consensus protocol satisfies agreement, total order and liveness.
-  \end{theorem}
+\begin{frame}{\subsecname}
+  \begin{figure}[h]
+  \includegraphics[width=0.7\textwidth]{trustchain-4}
+  \centering
+  \caption{CP blocks at round $r-1$ should be in $\mathcal{C}_r$.
+  The first $n$ nodes ordered by $\textsf{H}(\mathcal{C}_r || pk)$ become $\mathcal{F}_r$.}
+  \end{figure}
 \end{frame}
 
-\subsection{Validation}
-
-\begin{frame}{Validation}
-Assume node $u$ is aware of all the past consensus results $\mathcal{C}_r$.
-Suppose $u$ wish to validate $t_{i,j}$. It performs the following.
-
-\begin{enumerate}
-\item Determine the pair $t_{i', j'}$.
-\item Find the agreed enclosure for $t_{i,j}$ and $t_{i', j'}$ from
-  $\mathcal{C}_r$, otherwise return ``unknown''.
-%\item Find the smallest $c_{i', b}$ such that $b > j'$ is in consensus.
-\item Query $i$ and $i'$ for the agreed pieces and ensure hash pointers are
-  correct. Otherwise return ``unknown''.
-\item Check that $t_{i,j}$ and $t_{i', j'}$ are in the agreed pieces and are
-  created correctly using \texttt{newtx}. Otherwise return ``invalid''.
-\item Check the checkpoints $c_{i, k}$ and $c_{i', k'}$ that immediately follow
-  $t_{i,j}$ and $t_{i', j'}$ are in the agreed pieces and are created in the
-  same round, i.e. $\texttt{round}(c_{i, k}) = \texttt{round}(c_{i', k'})$.
-  Otherwise return ``invalid''.
-\item Return ``valid''.
-\end{enumerate}
+\subsection{Transaction protocol}
+\begin{frame}{\subsecname}
+\begin{itemize}
+\item Request (\texttt{tx\_req}) and response (\texttt{tx\_resp}) protocol
+\item Two TX blocks containing the same $txid$ are generated
+\item Non-blocking
+\end{itemize}
 \end{frame}
 
-\begin{frame}{Validation}
-  In essence, given a TX, ask the receiver to proof that it has a set of
-  transactions that produces some CP block, the CP block should be in the
-  consensus result and the pair of the TX should be in that set.
+\begin{frame}{\subsecname}
+Create $t_{u, h}$ and send $\langle \texttt{tx\_req}, t_{u, h} \rangle$ to start a transaction.
+\vspace{5 mm}
+\begin{algorithmic}
+    \Upon $\langle \texttt{tx\_req}, t_{v, j} \rangle$ from $v$
+    \State $\langle \_, \_, txid, pk_v, m, \_ \rangle \gets t_{v, j}$ \Comment unpack $t_{v, j}$
+    \State $\textsf{new\_tx}(pk_u, m, txid)$ \Comment create and store $t_{u, i}$
+    \State store $t_{v, j}$ as the pair of $t_{u, h}$
+    \State send $\langle \texttt{tx\_resp}, t_{u, h} \rangle$ to $v$
 
-  \begin{theorem}
-    If at least one party of every transaction is honest, then forking and other
-    forms of tampering is guaranteed to be detected if the malicious party is
-    alive.
-  \end{theorem}
-
-  \begin{block}{Question}
-    Given a starting digest and an ending digest, is it possible for the prover
-    to proof to the verifier that it has a set of blocks linked by hash pointers
-    that ``fits'' inside those digests in \emph{zero knowledge}.
-  \end{block}
+    \Upon $\langle \texttt{tx\_resp}, t_{v, j} \rangle$ from $v$  
+    \State $\langle \_, \_, txid, pk_v, m, \_ \rangle \gets t_{v, j}$ \Comment unpack $t_{v, j}$
+    \State store $t_{v, j}$ as the pair of the TX with identifier $txid$
+\end{algorithmic}
 \end{frame}
 
-\subsection{Implementation}
-\begin{frame}{Implementation}
+\subsection{Validation protocol}
+\begin{frame}{\subsecname overview}
 \begin{itemize}
-    \item Currently on going---using Python and Twisted
-    \item Completed BFT consensus algorithm
-    \item Completed local TrustChain with Checkpoints
-    \item Next step is networked TrustChain and validation protocol
+\item Request (\texttt{vd\_req}) and response (\texttt{vd\_resp}) protocol
+\item Transactions are in three states---\emph{valid}, \emph{invalid} and \emph{unknown}
+\item The goal is to identify which state a given transaction is in
 \end{itemize}
 \end{frame}
+
+\begin{frame}{\subsecname properties}
+\begin{itemize}
+    \item \emph{Correctness}:
+        The validation protocol outputs the correct result
+        according to the aforementioned validity definition.
+    \item \emph{Agreement}:
+        If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
+        then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
+    \item \emph{Liveness}:
+        Any valid transactions can be validated eventually.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Validity definition}
+\textbf{Function} $\textsf{get\_validity}()$ validates the transaction $t_{u, i}$
+\vspace{5 mm}
+\begin{algorithmic}
+    \State Check that $v$ sent the correct agreed fragment,
+    otherwise return \emph{unknown}.
+
+    \State
+    \State $\langle \_, \_, txid, pk_v, m, \_ \rangle \gets t_{u, i}$
+    \If{number of blocks of $txid$ in $F_{v, j} \ne 1$}
+        \State \Return \emph{invalid}
+    \EndIf
+    \State \Comment TX exists
+
+    \State $\langle \_, \_, txid', pk'_u, m', \_ \rangle \gets t_{v, j}$
+    \If{$m \ne m' \vee pk_u \ne pk'_u$}
+        \State \Return \emph{invalid}
+    \EndIf
+    \State \Comment no tampering
+    \State \Return \emph{valid}
+\end{algorithmic}
+\end{frame}
+
+\begin{frame}{\subsecname}
+Send $\langle \texttt{vd\_req}, txid \rangle$ to $v$ to begin.
+\vspace{5 mm}
+\begin{algorithmic}
+    \Upon $\langle \texttt{vd\_req}, txid \rangle$ from $v$
+        \State $t_{u, i} \gets \text{the transaction identified by } txid$
+        \State $F_{u, i} \gets \textsf{agreed\_fragment}(t_{u, i})$
+        \State send $\langle \texttt{vd\_resp}, txid, F_{u, i} \rangle$ to $v$
+
+    \Upon $\langle \texttt{vd\_resp}, txid, F_{v, j} \rangle$ from $v$
+        \State $t_{u, i} \gets \text{the transaction identified by } txid$
+        \State set the validity of $t_{u, i}$ to $\textsf{get\_validity}(t_{u, i}, F_{v, j})$
+\end{algorithmic}
+\end{frame}
+
+\section{Analysis of correctness and performance}
+
+\subsection{Correctness of the consensus protocol}
+\begin{frame}{\subsecname}
+\begin{theorem}
+For all rounds,
+the consensus protocol satisfies agreement, validity, fairness and termination.
+\end{theorem}
+\begin{proof}(sketch)
+Because consensus result are eventually delivered and the properties of ACS,
+we get agreement, validity and termination.
+Fairness is from the fact that we model $\textsf{H}(\cdot)$ as a random oracle (RO) and the input to the RO is different for every node,
+thus the list of nodes ordered by $\textsf{H}(\C_r || pk )$ is a random permutation of those nodes.
+\end{proof}
+\end{frame}
+
+\subsection{Correctness of the validation protocol}
+\begin{frame}{\subsecname}
+\begin{theorem}
+If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
+then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
+\end{theorem}
+\begin{proof}(sketch)
+Proof by contradiction.
+For this attack to work, the adversary must be able to create two different fragments but with the same checkpoint enclosure.
+We model $\textsf{H}(\cdot)$ as a RO, so the adversary need to query the RO a exponential\footnote{In terms of the security parameter.} number of times.
+But the adversary can only query the RO a polynomial number of times.
+\end{proof}
+\end{frame}
+
+\subsection{Linear global throughput argument}
+\begin{frame}{\subsecname}
+\begin{itemize}
+  \item Throughput has a notion of time but our model is asynchronous
+  \item Additional assumptions needed to make the argument---every unit of communication takes a non-negligible amount of time to process
+  \item Bandwidth relation---$NC \ge r_{\text{tx}} l$, where $l$ is $O(N)$
+  \item If $r_\text{tx}$ satisfies the inequality, then LHS and RHS grows at the same rate,
+  thus we have linear global throughput
+\end{itemize}
+\end{frame}
+
+\section{Experimental results}
+\begin{frame}{Implementation and experiment setup}
+  \begin{itemize}
+    \item Prototype implementation on Github\footnote{\url{https://github.com/kc1212/consensus-thesis-code}}
+    \item SHA256 for hash functions and Ed25519 for digital signature
+    \item Experiment on the DAS-5\footnote{\url{http://www.cs.vu.nl/das5/}}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Throughput vs population size (random neighbour)}
+  \begin{figure}[h]
+  \includegraphics[width=1.0\textwidth]{neighbour-random/throughput-vs-population}
+  \centering
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Throughput vs population size (fixed neighbour)}
+  \begin{figure}[h]
+  \includegraphics[width=1.0\textwidth]{neighbour-fixed/throughput-vs-population}
+  \centering
+  \end{figure}
+\end{frame}
+
+\section{Conclusion}
+\begin{frame}{\secname}
+    How do we design a \emph{blockchain consensus protocol} that is \emph{fault tolerant},
+    \emph{scalable} and can reach \emph{global consensus?}
+\begin{itemize}
+  \item[\checkmark] Fault tolerant up to $t$ nodes
+  \item[\checkmark] Horizontal scalability
+  \item[\checkmark] Global consensus
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Future work}
+\begin{itemize}
+\item Improve fault tolerance
+\item Improve fork detection
+\item Analyse the system in the permissionless environment
+\item Concrete application
+\end{itemize}
+\end{frame}
+
 \end{document}

### 2017-07-27 presentation/presentation.tex
Improve presentation

@@
 
+\setbeameroption{show notes}
+
 \title{Blockchain Consensus Protocol with Horizontal Scalability}
@@
 
-% \subject{Theoretical Computer Science}
-% This is only inserted into the PDF information catalog. Can be left
-% out.
+\titlegraphic{\includegraphics[width=4cm]{logo}}
 
-% If you have a file called "university-logo-filename.xxx", where xxx
-% is a graphic format that can be processed by latex or pdflatex,
-% resp., then you can add a logo as follows:
-
-% \pgfdeclareimage[height=0.5cm]{university-logo}{university-logo-filename}
-% \logo{\pgfuseimage{university-logo}}
-
-% Delete this, if you do not want the table of contents to pop up at
-% the beginning of each subsection:
 \AtBeginSubsection[] % Do nothing for \subsection*
@@
   \titlepage
+
 \end{frame}
@@
 \section{Introduction}
-\subsection{Motivation}
-\begin{frame}{\subsecname}{}
+\begin{frame}{Motivation}
   \begin{itemize}
     \item Blockchain systems offer an alternative to central authorities for the first time
-    \item Market cap and trade volume figures indicate they are here to stay
+    \item Market cap (40 billion for Bitcoin) and trade volume figures indicate they are here to stay
     \item Early blockchain systems are not scalable (7 TX/s for Bitcoin)
@@
 
-\subsection{Research question}
-\begin{frame}{\subsecname}
+\begin{frame}{Research question}
     How do we design a \emph{blockchain consensus protocol} that is \emph{fault tolerant},
     \emph{scalable} and can reach \emph{global consensus?}
+    \note{
+      \begin{itemize}
+      \item Blockchain consensus protocol---independent of the application, like PoW (not PoS)
+      \item Byzantine fault tolerance---if some number of nodes are malicious then the system should continue to function
+      \item Horizontal scalability---if every node make transactions at the same rate, then the global throughput should increase
+      \item Global consensus---all node should agree on a global state, useful for some applications, e.g. digital cash
+      \end{itemize}
+    }
 \end{frame}
 
-\subsection{Intuition}
-\begin{frame}{\subsecname}{}
+\begin{frame}{Inspiration}
   \begin{itemize}
@@
     \item Ocassionally a customer may leave without paying and this event is reported to a central authority
-    \item Our blockchain system achieves scalability using the same idea
+    \item Our blockchain system achieves scalability using a similar idea
   \end{itemize}
@@
     \item Random oracle model
-    \item Application neutural
+    \item Computational security
   \end{itemize}
+  \note{At this point, explain facilitators are nodes that have some extra task to do, temporarily.}
 \end{frame}
@@
   \end{figure}
+  \note{
+    \begin{itemize}
+    \item The primary data structure is the Extended TrustChain, extension of our prior work
+    \item The three protocols the tasks as their name suggests
+    \item They are independent and run concurrently
+    \item The only synchronisation happens via the Extended TrustChain
+    \item But in no part of those protocol do we lock the Extended TrustChain
+    \end{itemize}
+  }
 \end{frame}
@@
   \begin{itemize}
-    \item Everyone has their own chain
-    \item Two types of blocks, transaction (TX) blocks and checkpoint (CP) blocks
-    \item A transaction involves two parties and results in two TX blocks
+    \item Everyone has their own chain and genesis block
+    \item Two types of blocks
+      \begin{enumerate}
+      \item Transaction (TX) block
+      \item Checkpoint (CP) block
+      \end{enumerate}
+    \item A transaction involves two parties and results in two TX blocks (a pair)
     \item A CP block captures the chain state
@@
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{trustchain-bad-cp}
+  \includegraphics[width=0.7\textwidth]{trustchain-good-cp}
   \centering
@@
   \end{figure}
+  \note{
+    \begin{itemize}
+      \item In this example there are three nodes
+      \item Squares are TX blocks and circles are CP blocks
+      \item Explain the block content in caption
+      \item The dotted line represent pairs of TX blocks
+      \item Geared with the understanding of our data structure, we are ready to talk about the consensus protocol
+    \end{itemize}
+  }
+\end{frame}
+
+\begin{frame}{\subsecname:~fragment definition}
+\begin{itemize}
+\item A fragment $F_{u, i}$ is defined on a TX block $t_{u, i}$
+\item It is a section of the chain beginning and ending with CP blocks that contains the TX
+\end{itemize}
 \end{frame}
@@
 \subsection{Consensus protocol}
-\begin{frame}{\subsecname overview}
+\begin{frame}{\subsecname:~overview}
   \begin{enumerate}
-    \item In round $r$, $n$ out of $N$ lucky nodes are selected at random to act as facilitators
-    \item Facilitators run a BFT (Byzantine Fault Tolerant) consensus algorithm to
-      agree on a set of CP blocks
+    \item Runs in rounds
+    \item In round $r$, $n$ out of $N$ act as facilitators
+    \item The facilitators collect CP blocks from all nodes
+    \item Facilitators run a Byzantine consensus algorithm to agree on a set of CP blocks
     \item Disseminate the consensus result $\C_r$, i.e. the CP blocks
+    \item From $\C_r$, new facilitators are computed
     \item Repeat
@@
 
-\begin{frame}{\subsecname properties}
+\begin{frame}{\subsecname:~properties}
 \label{def:consensus}
@@
     \item \emph{Agreement}:
-        If one correct node outputs a list of facilitators $\F_r$,
+        If one correct node outputs a set of facilitators $\F_r$,
         then every node outputs $\F_r$
@@
   \centering
-  \caption{Suppose we are in a state where $\mathcal{C}_{r - 1}$ has just been agreed by some facilitators but not yet propagated.}
+  \caption{Suppose we are in a state where $\C_{r - 1}$ has just been agreed by some facilitators but not yet propagated.}
   \end{figure}
@@
   \centering
-  \caption{Nodes receive consensus result $\mathcal{C}_{r - 1}$,
-    compute the new facilitators $\mathcal{F}_{r-1}$,
-    and send out the new CP blocks.}
+  \caption{Nodes receive consensus result $\C_{r - 1}$,
+    first $n$ nodes ordered by $\textsf{H}(\C_r || pk)$ become $\F_{r-1}$,
+    send the new CP blocks to $\F_{r-1}$.}
   \end{figure}
@@
   \centering
-  \caption{CP blocks at round $r-1$ should be in $\mathcal{C}_r$.
-  The first $n$ nodes ordered by $\textsf{H}(\mathcal{C}_r || pk)$ become $\mathcal{F}_r$.}
+  \caption{$\F_{r-1}$ agree and disseminate $\C_r$,
+  CP blocks at round $r-1$ ($c_{a, 6}, c_{b, 4}, c_{c,8}$) should be in $\C_r$.}
   \end{figure}
@@
 \subsection{Validation protocol}
-\begin{frame}{\subsecname overview}
+\begin{frame}{\subsecname:~overview}
 \begin{itemize}
 \item Request (\texttt{vd\_req}) and response (\texttt{vd\_resp}) protocol
-\item Transactions are in three states---\emph{valid}, \emph{invalid} and \emph{unknown}
-\item The goal is to identify which state a given transaction is in
-\end{itemize}
-\end{frame}
-
-\begin{frame}{\subsecname properties}
-\begin{itemize}
-    \item \emph{Correctness}:
-        The validation protocol outputs the correct result
-        according to the aforementioned validity definition.
-    \item \emph{Agreement}:
-        If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
-        then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
-    \item \emph{Liveness}:
-        Any valid transactions can be validated eventually.
+\item Transactions are in three states---\emph{valid}, \emph{invalid} and \emph{unknown}---defined by $\textsf{get\_validity}(\cdot)$
+\item The goal is to identify which state a given transaction is in satisfying the validation protocol properties
 \end{itemize}
@@
 \begin{frame}{Validity definition}
-\textbf{Function} $\textsf{get\_validity}()$ validates the transaction $t_{u, i}$
+\textbf{Function} $\textsf{get\_validity}(t_{u, i}, F)$ validates the transaction $t_{u, i}$
 \vspace{5 mm}
 \begin{algorithmic}
-    \State Check that $v$ sent the correct agreed fragment,
+    \State Check that $v$ sent the correct $F$,
     otherwise return \emph{unknown}.
@@
     \State $\langle \_, \_, txid, pk_v, m, \_ \rangle \gets t_{u, i}$
-    \If{number of blocks of $txid$ in $F_{v, j} \ne 1$}
+    \If{number of blocks of $txid$ in $F \ne 1$}
         \State \Return \emph{invalid}
@@
 \end{algorithmic}
+\end{frame}
+
+\begin{frame}{\subsecname:~properties}
+\begin{itemize}
+    \item \emph{Agreement}:
+        If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
+        then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
+    \item \emph{Correctness}:
+        The validation protocol outputs the correct result
+        according to the aforementioned validity definition.
+    \item \emph{Liveness}:
+        Any valid (invalid) transaction is marked as validated (invalid) eventually.
+\end{itemize}
+\note{Liveness is not possible in our case because malicious nodes can simply go offline or refuse to respond,
+but the other two properties are possible.}
 \end{frame}
@@
 \begin{proof}(sketch)
-Because consensus result are eventually delivered and the properties of ACS,
+Because consensus result are eventually delivered and the properties of Byzantine consensus,
 we get agreement, validity and termination.
@@
 \begin{theorem}
-If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
+If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction according to the validity definition,
 then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
@@
     \item Experiment on the DAS-5\footnote{\url{http://www.cs.vu.nl/das5/}}
+      \begin{itemize}
+        \item Up to 30 machines
+        \item Each running 40 nodes
+      \end{itemize}
   \end{itemize}
@@
   \end{figure}
+\note{No need to request for fragment every time a TX needs to be validated. Upon receiving a fragment, validate as many TX as possible.}
 \end{frame}

### 2017-07-27 thesis/chapters/ch3-model.tex
Update presentation
Improve presentation

@@
 
-If the enclosure of some TX block is $\langle c_{u,a}, c_{u, b} \rangle$,
-then its \emph{fragment} is defined as $\{ b_{u, i} : a \le i \le b \}$.
+If the enclosure of some TX block $t_{u, i}$ is $\langle c_{u,a}, c_{u, b} \rangle$,
+then its \emph{fragment} $F_{u, i}$ is defined as $\{ b_{u, i} : a \le i \le b \}$.
 For convenience, the function $\textsf{fragment}(\cdot)$ represents the fragment of some TX block if it exists, otherwise $\bot$.
@@
 Second, it computes the new facilitators using $\textsf{get\_facilitator}(C, n)$ (\Cref{alg:facilitator})
-and updates its facilitator list to the result.
+and updates its facilitator set to the result.
 This concludes the consensus phase and brings us back to the beginning of the consensus phase.
@@
 \begin{algorithm}
-\caption{Function $\textsf{get\_facilitator}(C, n)$ takes a list of CP blocks $C$ and an integer $n$,
+\caption{Function $\textsf{get\_facilitator}(C, n)$ takes a set of CP blocks $C$ and an integer $n$,
 sort evey element in $C$ by its luck value (the $\lambda$-expression), and outputs the smallest $n$ elements.}
@@
         \Upon $\langle \texttt{tx\_req}, t_{v, j} \rangle$ from $v$
-        \State $txid, pk_v, m \gets t_{v, j}$ \Comment unpack the TX
+        \State $\langle \_, \_, txid, pk_v, m, \_ \rangle \gets t_{v, j}$
         \State $\textsf{new\_tx}(pk_u, m, txid)$
@@
         \Upon $\langle \texttt{tx\_resp}, t_{v, j} \rangle$ from $v$
-        \State $txid, pk_v, m \gets t_{v, j}$ \Comment unpack the TX
+        \State $\langle \_, \_, txid, pk_v, m, \_ \rangle \gets t_{v, j}$
         \State store $t_{v, j}$ as the pair of the TX with identifier $txid$
@@
         \State \Return \emph{unknown}
-    \EndIf \Comment $v$ has agreed fragment
-    \State
+    \EndIf
+    \State \Comment $v$ has agreed fragment
 
@@
         \EndIf
-    \EndIf \Comment correct TrustChain structure
-    \State
+    \EndIf
+    \State \Comment correct TrustChain structure
 
@@
         \State \Return \emph{unknown}
-    \EndIf \Comment correct consensus round
-    \State
+    \EndIf
+    \State \Comment correct consensus round
     \label{line:valid-fragment}
-    \State $txid, pk_v, m \gets t_{u, i}$
+    \State $\langle \_, \_, txid, pk_v, m, \_ \rangle \gets t_{u, i}$
     \If{number of blocks of $txid$ in $F_{v, j} \ne 1$}
         \State \Return \emph{invalid}
-    \EndIf \Comment TX exists
-    \State
+    \EndIf
+    \State \Comment TX exists
 
-    \State $txid', pk'_u, m' \gets t_{v, j}$
+    \State $\langle \_, \_, txid', pk'_u, m', \_ \rangle \gets t_{v, j}$
     \If{$m \ne m' \vee pk_u \ne pk'_u$}
         \State \Return \emph{invalid}
-    \EndIf \Comment no tampering
-    \State
-
+    \EndIf
+    \State \Comment no tampering
     \State \Return \emph{valid}
@@
 \begin{enumerate}
+    \item \emph{Agreement}:
+        If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
+        then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
     \item \emph{Correctness}:
@@
         according to the aforementioned validity definition.
-    \item \emph{Agreement}:
-        If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
-        then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
     % \item \emph{Unforgeability}:
@@
     \item \emph{Liveness}:
-        Any valid transactions can be validated eventually.
+        Any valid (invalid) transactions is marked as valid (invalid) eventually.
 \end{enumerate}

### 2017-07-27 presentation/presentation.tex
Use TUD template for presentation

@@
-% Copyright 2004 by Till Tantau <tantau@users.sourceforge.net>.
-%
-% In principle, this file can be redistributed and/or modified under
-% the terms of the GNU Public License, version 2.
-%
-% However, this file is supposed to be a template to be modified
-% for your own needs. For this reason, if you use this file as a
-% template and not specifically distribute it as part of a another
-% package/program, I grant the extra permission to freely copy and
-% modify this file as you see fit and even to delete this copyright
-% notice.
+% TU Delft beamer template
+% Author: Erwin Walraven (initial version was created by Maarten Abbink)
+% Delft Universiy of Technology
 
 \documentclass{beamer}
-
-% There are many different themes available for Beamer. A comprehensive
-% list with examples is given here:
-% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
-% You can uncomment the themes below if you would like to use a different
-% one:
-%\usetheme{AnnArbor}
-%\usetheme{Antibes}
-%\usetheme{Bergen}
-%\usetheme{Berkeley}
-%\usetheme{Berlin}
-%\usetheme{Boadilla}
-%\usetheme{boxes}
-%\usetheme{CambridgeUS}
-%\usetheme{Copenhagen}
-%\usetheme{Darmstadt}
-\usetheme{default}
-%\usetheme{Frankfurt}
-%\usetheme{Goettingen}
-%\usetheme{Hannover}
-%\usetheme{Ilmenau}
-%\usetheme{JuanLesPins}
-%\usetheme{Luebeck}
-%\usetheme{Madrid}
-%\usetheme{Malmoe}
-%\usetheme{Marburg}
-%\usetheme{Montpellier}
-%\usetheme{PaloAlto}
-%\usetheme{Pittsburgh}
-%\usetheme{Rochester}
-%\usetheme{Singapore}
-%\usetheme{Szeged}
-%\usetheme{Warsaw}
-
-\usepackage{amsfonts}
+\usepackage[english]{babel}
+\usepackage{calc}
+\usepackage[absolute,overlay]{textpos}
 \usepackage{graphicx}
 \graphicspath{{../figures/}}
+\usepackage{subfig}
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amsthm}
+\usepackage{mathtools}
+\usepackage{comment}
+\usepackage{MnSymbol,wasysym}
 \usepackage[noend]{algpseudocode}
 \usepackage{algorithm}
+
+\setbeamertemplate{navigation symbols}{} % remove navigation symbols
+\mode<presentation>{\usetheme{tud}}
 
@@
 
-\setbeameroption{show notes}
+% \setbeameroption{show notes}
 
-\title{Blockchain Consensus Protocol with Horizontal Scalability}
-
-% A subtitle is optional and this may be deleted
+\title{Blockchain Consensus Protocol with\\Horizontal Scalability}
 % \subtitle{Optional Subtitle}
@@
 \date{\today}
-% - Either use conference name or its abbreviation.
-% - Not really informative to the audience, more for people (including
-%   yourself) who are reading the slides online
-
-\titlegraphic{\includegraphics[width=4cm]{logo}}
 
@@
 \begin{frame}{Research question}
+  \begin{block}{}
     How do we design a \emph{blockchain consensus protocol} that is \emph{fault tolerant},
     \emph{scalable} and can reach \emph{global consensus?}
+  \end{block}
     \note{
@@
     \item A restaurant owner does not report all of its transactions with an central authority
-    \item Ocassionally a customer may leave without paying and this event is reported to a central authority
+    \item Occasionally a customer may leave without paying and this event is reported to a central authority
     \item Our blockchain system achieves scalability using a similar idea
@@
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{trustchain-good-cp}
+  \includegraphics[width=0.67\textwidth]{trustchain-good-cp}
   \centering
@@
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{trustchain-1}
+  \includegraphics[width=0.67\textwidth]{trustchain-1}
   \centering
@@
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{trustchain-2}
+  \includegraphics[width=0.67\textwidth]{trustchain-2}
   \centering
   \caption{Nodes receive consensus result $\C_{r - 1}$,
-    first $n$ nodes ordered by $\textsf{H}(\C_r || pk)$ become $\F_{r-1}$,
+    first $n$ nodes ordered by $\textsf{H}(\C_{r-1} || pk)$ become $\F_{r-1}$,
     send the new CP blocks to $\F_{r-1}$.}
@@
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{trustchain-3}
+  \includegraphics[width=0.67\textwidth]{trustchain-3}
   \centering
@@
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{trustchain-4}
+  \includegraphics[width=0.67\textwidth]{trustchain-4}
   \centering
@@
     \item \emph{Agreement}:
-        If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
+        If any correct node decides on the validity of a transaction (except when it is \emph{unknown}),
         then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
@@
 For all rounds,
-the consensus protocol satisfies agreement, validity, fairness and termination.
+the consensus protocol satisfies agreement, validity, fairness and termination properties.
 \end{theorem}
@@
 \begin{theorem}
-If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction according to the validity definition,
-then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
+  The validation protocol satisfies agreement and correctness properties.
 \end{theorem}
@@
 But the adversary can only query the RO a polynomial number of times.
+Hence we have agreement.
+Correctness follows directly the use of the $\textsf{get\_validity}(\cdot)$ function.
 \end{proof}
@@
   \begin{figure}[h]
-  \includegraphics[width=1.0\textwidth]{neighbour-random/throughput-vs-population}
+  \includegraphics[width=0.9\textwidth]{neighbour-random/throughput-vs-population}
   \centering
@@
   \begin{figure}[h]
-  \includegraphics[width=1.0\textwidth]{neighbour-fixed/throughput-vs-population}
+  \includegraphics[width=0.9\textwidth]{neighbour-fixed/throughput-vs-population}
   \centering
@@
 \begin{frame}{\secname}
+  \begin{block}{Research question}
     How do we design a \emph{blockchain consensus protocol} that is \emph{fault tolerant},
     \emph{scalable} and can reach \emph{global consensus?}
+  \end{block}
+Our system achieve the following
 \begin{itemize}
-  \item[\checkmark] Fault tolerant up to $t$ nodes
-  \item[\checkmark] Horizontal scalability
-  \item[\checkmark] Global consensus
+  \item Fault tolerant up to $t$ nodes
+  \item Horizontal scalability
+  \item Global consensus on CP blocks
 \end{itemize}

### 2017-07-29 presentation/presentation.tex
Edit ch1 and ch2

@@
   \centering
-  \caption{TX block is a six-tuple: $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), i, txid, pk_v, m, sig_u \rangle$.
-    CP block is a five-tuple: $c_{u, i} = \langle \textsf{H}(b_{u, i-1}), i, \textsf{H}(\C_r), r, sig_u \rangle$.}
+  \caption*{TX block is a six-tuple: $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), i, txid, pk_v, m, sig_u \rangle$\\
+    CP block is a five-tuple: $c_{u, i} = \langle \textsf{H}(b_{u, i-1}), i, \textsf{H}(\C_r), r, sig_u \rangle$}
   \end{figure}
@@
   \centering
-  \caption{Suppose we are in a state where $\C_{r - 1}$ has just been agreed by some facilitators but not yet propagated.}
+  \caption*{Suppose we are in a state where $\C_{r - 1}$ has just been agreed by some facilitators but not yet propagated.}
   \end{figure}
@@
   \centering
-  \caption{Nodes receive consensus result $\C_{r - 1}$,
+  \caption*{Nodes receive consensus result $\C_{r - 1}$,
     first $n$ nodes ordered by $\textsf{H}(\C_{r-1} || pk)$ become $\F_{r-1}$,
@@
   \centering
-  \caption{Transactions carry on as usual in round $r$,
+  \caption*{Transactions carry on as usual in round $r$,
   while facilitators are trying to reach consensus on the new CP blocks concurrently.}
@@
   \centering
-  \caption{$\F_{r-1}$ agree and disseminate $\C_r$,
+  \caption*{$\F_{r-1}$ agree and disseminate $\C_r$,
   CP blocks at round $r-1$ ($c_{a, 6}, c_{b, 4}, c_{c,8}$) should be in $\C_r$.}

### 2017-07-29 thesis/chapters/ch1-intro.tex
Write more design variations
Edit ch1 and ch2

@@
 cloud services, hospitality services and even our banking system.
-The 2008 financial crisis is an example of the banking system making poor choices which result in a
+The 2008 financial crisis is an example of the banking system making poor choices which resulted in a
 decline in consumer wealth in the order of trillions~\cite{financialcrisis} and led to the European debt crisis.
@@
 Ironically, also in 2008, Satoshi Nakamoto published the Bitcoin whitepaper~\cite{bitcoin}.
-Which, for the first time, gave us a simple banking system in the form of a distributed ledger.
+Which, for the first time, gave us a simple banking system in the form of a distributed ledger, also known as blockchain.
 It needs no central control but still incorruptable with high probability even if there are malicious parties that aim to undermine the system.
-We call such a ledger a blockchain.
 
@@
 For example, $C$ has 5 units of currency in her account.
-At some instant in time, $A$ sees that $C$ transferred 5 units to $A$ and $B$ sees that $C$ transferred 5 units to $B$.
-Hence the transaction is inconsistent with respect to the observer.
-Bitcoin and many of its derivatives (e.g. Litecoin\footnote{\url{litecoin.org}} and Dogecoin\footnote{\url{dogecoin.org}}) solve the inconsistency problem with a consensus algorithm.
-The goal of the algorithm is reach agreement on a set of valid values, e.g. transactions. 
+If $C$ can simultaneously claim that she transferred 5 units to $A$ but also 5 units to $B$, then the system is inconsistent.
+Bitcoin and many of its derivatives (e.g. Litecoin\footnote{\url{https://litecoin.org}} and Dogecoin\footnote{\url{http://dogecoin.org}}) solve the inconsistency problem with a consensus algorithm.
+The goal of the algorithm is to reach agreement on a set of valid transactions. 
 This effectively eliminates inconsistencies.
+
 In Bitcoin's case, the consensus algorithm is called \emph{proof-of-work}.
-Where miners (parties that runs the Bitcoin network) collect transactions and compete in solving (using brute force) a difficult puzzle.
-The first miner to solve it generates a block containing all the collected transactions.
+Where miners (parties that runs the Bitcoin network) collect transactions and compete in solving a puzzle.
+The puzzle is easy to verify but difficult to solve.
+Only the miner that solves it can generate a valid block containing all the collected transactions.
 The miner is also rewarded with new coins and transaction fees.
@@
 Hence the name blockchain.
-Further, due to the difficulty of the puzzle and that the majority of the parties are honest,
-it is unlikely for more than one blockchain to exist in the network for a long period of time.
+This property ensures malicious nodes cannot easily rewrite history if they do not have a majority of the network's computing power.
+Consequently, it is unlikely for more than one blockchain to exist in the network for a long period of time.
 Thus every party sees a consistent blockchain which solves the double spending problem.
@@
 Its market cap, at the time of writing, is over 40 billion USD~\cite{bitcoinmarketcap}.
+% TODO add graph
 Many online marketplaces are using Bitcoin, for example Steam\footnote{\url{https://store.steampowered.com/}} and even Amazon\footnote{Not directly, but via \url{https://purse.io/}.}.
@@
 Due to this limitation, it is not uncommon to see a long backlog of about 20,000 transactions\footnote{\url{https://blockchain.info/unconfirmed-transactions/}}.
-A few months ago the backlog even reached 100,000, which meant new transactions would take at least 11 hours to hit the blockchain~\cite{bitcoinbacklog}.
+A few months ago the backlog even reached 100,000, which meant new transactions would take at least 11 hours to be written to the blockchain~\cite{bitcoinbacklog}.
+% TODO update below
 This issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some calls it a civil war~\cite{bitcoincivilwar}.
@@
 It also requires a hard fork (not backward compatible), which risks consensus failure and devaluation.
+% TODO update the above
 A recent empirical study by Croman et el.~\cite{croman2016scaling} has shown that increasing the block size may help.
 But given the bandwidth and latency constraints,
-it not possible to have more than 758 transactions per second.
+it is not possible to have more than 758 transactions per second.
 Even worse, it may give some miner an unfair lead over others.
-Thus fundamental changes are necessary run at the scale of centralised payment processors such as Visa~\cite{visa},
-which is in the order of tens of thousands transactions per second.
+Thus fundamental changes are necessary run at the scale of centralised payment processors such as Visa,
+which is in the order of tens of thousands transactions per second~\cite{visa}.
 
+% TODO update
 What we have described is in fact the current state of Bitcoin.
@@
 Even if it is adopted, it cannot reach the transaction rate of centralised systems.
+% https://www.bloomberg.com/news/articles/2017-07-10/bitcoin-risks-splintering-as-civil-war-enters-critical-month
 
@@
 which can be seen as a collective authority.
-This consequently lead to limited scalability because every transaction must be validated by the authority even when most of the transaction are legitimate.
+This consequently lead to limited scalability because every transaction must be validated by the authority even when most of the transactions are legitimate.
 
@@
 but nevertheless verifiable at a later stage by any node in the network.
-Informally, out model works as follows.
+Informally, our system works as follows.
 Every node stores its own blockchain and every block is one transaction.
-We randomly selected nodes in every round, the selected ones are called facilitators.
+We randomly select nodes in every round, the selected ones are called facilitators.
 They reach consensus not on the individual transactions,
@@
 This immediately show that those transactions are tamper-proof.
+% TODO mention consensus and trade off?
 

### 2017-07-29 thesis/chapters/ch2-problem.tex
Write more design variations
Edit ch1 and ch2

@@
-\chapter{Problem Description}
+\chapter{Problem description}
 \label{ch:problem}
@@
 However, as we described in the Introduction, it faces a genuine scalability problem.
-In particular, proof-of-work based systems are unlikely to reach the transaction rate of Visa even with parameter tuning.
+In particular, proof-of-work based systems are unlikely to reach a transaction rate suitable for ubiquitous use, even with parameter tuning.
 What we wish to see is horizontal scalability.
@@
 \begin{displayquote}
-\emph{How do we design a blockchain fabric that is fault tolerant, scalable and can reach global consensus?}
+\emph{How do we design a blockchain consensus protocol that is fault tolerant, scalable and can reach global consensus?}
 \end{displayquote}
 
-In order to make sense of our research question, we must first define \emph{blockchain fabric}.
+In order to make sense of our research question, we must first define \emph{blockchain consensus protocol}.
 Typical blockchain systems are application specific.
 For example, the application in Bitcoin (any many of its derivatives) is digital cash, the application in Namecoin is domain name registration~\cite{namecoin} and the application in Siacoin is cloud storage~\cite{siacoin}.
-Underneath these applications lies the blockchain fabric which has the goal of reaching consensus on transactions.
+Underneath these applications lies the blockchain consensus protocol which has the goal of reaching consensus on transactions or state changes.
 In the case of Bitcoin it is proof-of-work (PoW).
@@
 it works by simply treating transactions as a blob of data and hashes them with the previous block.
-Hence, we focus on blockchain fabric rather than any specific application.
+In this work, we focus on the blockchain consensus protocol rather than any specific application.
 Because it is the scalability bottleneck and is neccessary for any blockchain based application.
@@
 In the remainder of this chapter we expand on our research question and visit each of our requirement in detail.
+At the end of this chapter we describe some issues that we do not consider as a part of our design,
+which is the result of not focusing on any specific application.
 
-\section{Global Consensus}
+\section{Global consensus}
 
@@
 The authors show that running PBFT in Network File System (NFS) was only 3\% slower than the standard unreplicated NFS.
-Since then, research in this area picked up pace and many more algorithms are introduced~\cite{abd2005fault, kotla2007zyzzyva}.
+Since then, research in this area picked up pace and many more algorithms are introduced, for example Q/U~\cite{abd2005fault} and Zyzzyva~\cite{kotla2007zyzzyva}.
 With the introduction of Bitcoin's probabilistic consensus model, the field experienced a second renaissance.
@@
 
-\section{Security and Fault Tolerance}
+\section{Security and fault tolerance}
 
@@
 
-\section{Performance and Scalability}
+\section{Performance and scalability}
 
@@
 
-Traditional blockchain systems may run in a network of thousands of nodes, but the performance do not scale.
+Early blockchain systems may run in a network of thousands of nodes, but the performance do not scale.
 On the other hand, as we will see in~\Cref{sec:acs-background}, Byzantine consensus algorithms performs well but only when in a small network.
-In this work we combine the ideas from both of these systems to create the first horizontally scalable blockchain fabric.
+In this work, our goal is to overcome the limitations of early blockchain systems and Byzantine consensus algorithms to create a horizontally scalable blockchain.
 
@@
 Our work aims to significantly lower the transaction bound.
-However, the price to pay is that the Sybil attack~\cite{douceur2002sybil} prevention mechanism moves from the blockchain fabric to the application layer.
+However, the price to pay is that the Sybil attack~\cite{douceur2002sybil} prevention mechanism moves from the blockchain consensus protocol to the application layer.
 For example, social network based Sybil defence mechanims use graph structure of real-world relationships~\cite{yu2006sybilguard}.
@@
 and applications built on top of our system can use the best mechanism for their purpose.
+
+For the same reason, we do not explicitly consider spam or denial of service attacks.
+Without a concrete application it is difficult to distinguish a super active user from an attacker.
 

### 2017-07-30 thesis/chapters/ch3-model.tex
Edit ch3, 4 and 5

@@
 
-We present a design which removes the transaction rate restrictions seen in traditional blockchain systems.
-This is done by combining the recent HoneyBadgerBFT~\cite{miller2016honey}
-(a purely asynchronous consensus algorithm for blockchain systems)
-with a modification of our prior work on TrustChain~\cite{trustchain}.
-The design is inspired by (the distributed part of) the Internet architecture as well as how transactions are performed in the real world.
+We design a novel system in order to achieve our ambitious goals outline in~\Cref{ch:problem}.
+The system consist of three protocols---the consensus protocol, the transaction protocol and the validation protocol.
+It is based on our prior work on TrustChain~\cite{trustchain}, where every node independently interact with each other via their own blockchain.
 
@@
 \label{sec:system-overview}
-The system consist of one data structure---Extended TrustChain,
-and three protocols---consensus protocol, transaction protocol and validation protocol.
-We first describe each component individually and then explain how they fit together in \Cref{sec:combined-protocol}.
+The architecture of our system is visualised in~\Cref{fig:architecture}.
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.7\textwidth]{architecture}
+\caption{Overview of the system architecture}
+\label{fig:architecture}
+\end{figure}
+It consist of one data structure---Extended TrustChain, and three protocols.
+The protocols have distinct roles which are evident from their names.
+They run concurrently and do not interact with each other.
+The only synchronisation happens on the Extended TrustChain layer, where all the protocols read and write from it.
+We first describe each component individually, starting with Extended TrustChain, and then explain how they fit together.
 
@@
 Initially, the chain only contains a genesis block generated by the nodes themselves.
-When a node $A$ wishes to add a new transaction (TX) with $B$, a new TX block is generated and then appended to $A$'s chain.
+When a node $A$ wishes to add a new transaction (TX) with $B$,
+a new TX block is generated for both $A$ and $B$ and is appended to their respective chains.
 The TX block must have a valid hash pointer pointing to the previous block
@@
 This work uses one outgoing edge and a reference.} to its \emph{pair} on $B$'s chain.
-As a result, a single transaction generates two TX blocks, one on each party's chain.
 An example of is shown in \Cref{fig:trustchain-bad}.
@@
 
-If every node follows the rules of TrustChain and we only consider hash pointers,
+If every node follows the rules of TrustChain and we only consider hash pointers and not references,
 then every chain effectively forms a singly linked list.
@@
 
-\subsection{Consensus protocol}\label{sec:overview-cons}
+\subsection{Consensus protocol}
+\label{sec:overview-cons}
 
@@
 because the output of a cryptographically secure hash function is unpredictable and $\C_r$ cannot be determined in advance.
+Finally, the elected facilitators run a Byzantine consensus algorithm to agree on a set of CP blocks.
+The next round begins when the new facilitators are computed from these CP blocks.
 
@@
 Thus we need a validation protocol to counteract malicious behaviour.
-When a node wish to validate one of its TX, it asks the counterparty for the \emph{fragment} of the TX.
-A fragment of a TX is a section of the chain beginning and ending with CP blocks that contains the TX and are in consensus.
+When a node wish to validate one of its TX, it asks the counterparty for the \emph{agreed fragment} of the TX.
+An agreed fragment of a TX is a section of the chain beginning and ending with CP blocks that contains the TX and are in consensus.
 Upon the counterparty's response, the node checks that the CP blocks are indeed in consensus,
@@
 Intuitively, this works because it is hard (because a cryptographically secure hash function is second-preimage resistant)
-to create a different chain that begins and ends with the same two CP blocks but with a different middle section.
+to create a different chain that begins and ends with the same two CP blocks.
 
@@
 
-This section and the ones following it give a technical treatment of what the content in System Overview.
-For notational clarity, we use the following convention (adapted from~\cite{miller2016honey}) throughout this work.
+This section and the ones following it give a technical treatment of what the content in System overview.
+For notational clarity, we use the following convention (adapted from~\cite{miller2016honey}) for most of this work.
 \begin{itemize}
@@
 We assume purely asynchronous channels with eventual delivery.
-Thus in no circumstance do we make timing assumptions.
+Thus in no stage of our protocol do we make timing assumptions.
 The adversary has fully control of the delivery schedule and the message ordering of all messages.
@@
 We assume there exist a Public Key Infrastructure (PKI), and nodes are identified by their unique and permanent public key.
-Finally, we use the random oracle model, i.e. calls to the random oracle are denoted by $\textsf{H}: \{0, 1\}^* \rightarrow \{0, 1\}^\lambda$,
-where $\{0, 1\}^*$ denotes the space of finite binary strings and $\lambda$ is the security parameter~\cite{bellare1993random}.
+Finally, we use the random oracle (RO) model, i.e. calls to the random oracle are denoted by $\textsf{H}: \{0, 1\}^* \rightarrow \{0, 1\}^\lambda$,
+where $\{0, 1\}^*$ denotes the space of finite binary strings and $\lambda$ is the security parameter.
+Under the RO model, the probability of successfully computing the inverse of the hash function is negligible with respect to $\lambda$~\cite{bellare1993random}.
 
+% TODO big N does not follow our convention
 In our model we consider $N$ nodes, which is the population size.
@@
 This is realistic because otherwise the adversary can eventually learn all the private keys and sabotage the system.
-Finally, we assume computational security.
-That is, the adversary can run polynomial-time algorithms but not exponential-time algorithms efficiently.
 
@@
 Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
-The chain consist of blocks $B_u = \{ b_{u, i} : k \in \{ 0, \dots, h - 1 \} \}$,
+The chain consist of blocks $B_u = \{ b_{u, i} : i \in \{ 0, \dots, h - 1 \} \}$,
 where $b_{u, i}$ is the $i$th block of $u$,
-and $h = |B_u|$.
+and $h$ is the height of the block (i.e. $h = |B_u|$).
 We often use $b_{u, h}$ to denote the latest block.
 There are two types of blocks, TX blocks and CP blocks.
-If $T_u$ is the set of TX blocks of $u$ and $C_u$ is the set of CP blocks of $u$,
+If $T_u$ is the set of all TX blocks in $B_u$ and $C_u$ is the set of all CP blocks is $B_u$,
 then it must be the case that $T_u \cup C_u = B_u$ and $T_u \cap C_u = \varnothing$.
@@
 
-\subsection{Transaction block}
+\begin{definition}
+\textbf{\emph{Transaction block}}
+
 The TX block is a six-tuple, i.e $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$.
@@
 Thus, given a TX block, these properties allow us to identify its pair.
+\end{definition}
 
@@
 
-\subsection{Checkpoint block}
+\begin{definition}
+\textbf{\emph{Checkpoint block}}
 
@@
 where $\C_r$ is the consensus result (which we describe in \Cref{sec:consensus-result}) in round $r$, the other items are the same as the TX block definition.
-Note that unlike in our prior work~\cite{implicitconsensus}, CP blocks and TX blocks do not have independent sequence numbers.
+% Note that unlike in our prior work~\cite{implicitconsensus}, CP blocks and TX blocks do not have independent sequence numbers.
 
@@
 The genesis block is unique because every node has a unique public and private key pair.
+\end{definition}
 
-\subsection{Consensus result}
+\begin{definition}
 \label{sec:consensus-result}
+\textbf{\emph{Consensus result}}
+
 Our consensus protocol runs in rounds as discussed in \Cref{sec:system-overview}.
@@
 where $C$ is a set of CP blocks agreed by the facilitators of round $r$.
+\end{definition}
 
-\subsection{Chain properties}
-Here we define a few important properties which results from the interleaving nature of CP and TX blocks.
+Here we define an important properties which results from the interleaving nature of CP and TX blocks.
+It is used in our validation protocol.
+\begin{definition}
+\textbf{\emph{Enclosure and agreed enclosure}}
 
@@
 $$b = \argmin_{k, k > i, \texttt{typeof}(b_{u,k}) = \gamma}(k - i),$$
-then $\langle c_{u,a}, c_{u, b} \rangle$ is the \emph{enclosure} of $t_{u, i}$.
+then $\langle c_{u,a}, c_{u, b} \rangle$ is the enclosure of $t_{u, i}$.
+Intuitively, that is the two closest CP blocks that surround $t_{u, i}$.
 Note that $c_{u, a}$ is the more recent CP block.
-Also, some TX blocks may not have any subsequent CP blocks, then its enclosure is $\bot$.
+Also, some TX blocks may not have any enclosure, then its enclosure is $\bot$.
+Agreed enclosure is the same as enclosure with an extra constraint where the CP blocks must be in some consensus result $\C_r$.
+
+\end{definition}
+
+Note that the ``closest CP blocks'' property must also apply to agreed enclosure. 
+Suppose a chain is in the form
+$\{c_{i}, c_{i+1}, t_{i+2}, c_{i+3}\}$\footnote{Usually the notation is of the form $c_{u, i}$, but the node identity is not important here so we simplify it to $c_{i}$}
+and $c_{i}, c_{i+1}, c_{i+3}$ are in $\C_r, \C_{r+1}, \C_{r+3}$ respectively,
+then the agreed enclosure of $t_{i+2}$ is $\langle c_{i+1}, c_{i+3}\rangle$ and cannot be $\langle c_{i}, c_{i+3}\rangle$.
+
+\begin{definition}
+\textbf{\emph{Fragment and agreed fragment}}
 
@@
 then its \emph{fragment} $F_{u, i}$ is defined as $\{ b_{u, i} : a \le i \le b \}$.
-For convenience, the function $\textsf{fragment}(\cdot)$ represents the fragment of some TX block if it exists, otherwise $\bot$.
-
-\emph{Agreed enclosure} is the same as enclosure with an extra constraint where the CP blocks must be in some consensus result $\C_r$ and also must be the smallest enclosure.
-That is, suppose a chain is in the form
-    $\{c_{i}, c_{i+1}, t_{i+2}, c_{i+3}\}$\footnote{Usually the notation is of the form $c_{u, i}$, but the node identity is not important here so we simplify it to $c_{i}$}
-    and $c_{i}, c_{i+1}, c_{i+3}$ are in $\C_r, \C_{r+1}, \C_{r+3}$ respectively,
-    then the agreed enclosure of $t_{i+2}$ is $\langle c_{i+1}, c_{i+3}\rangle$ and cannot be $\langle c_{i}, c_{i+3}\rangle$.
 Similarly, \emph{agreed fragment} has the same definition as fragment but using agreed enclosure.
-We define its function to be $\textsf{agreed\_fragment}(\cdot)$ which we use later in the validation protocol (\Cref{sec:vd-protocol}).
+For convenience, the function $\textsf{agreed\_fragment}(\cdot)$ represents the agreed fragment of some TX block if it exists, otherwise $\bot$.
+\end{definition}
 
@@
 
-Our consensus protocol runs on top of Extended TrustChain.
-It is directly related to the creation of CP blocks.
+Our scalable consensus protocol runs on top of Extended TrustChain.
+It uses an unmodified asynchronous subset consensus (ACS) algorithm as the key building block.
 The objectives of the protocol are to
@@
             \begin{enumerate}
-                \item $|\C_r| \ge N - t$ must hold for the $\C_r$ which was used to create $\F_r$,
+                \item $|\C_r| \ge N - t$ and each CP in $\C_r$ belong to a different node\footnote{
+                $\C_r$ is a tuple but we abuse the notation here by writing $|\C_r|$ to mean the number of CP blocks in the second element of $\C_r$.}.
                 \item $\F_r$ must contain at least $n - t$ honest nodes and
@@
 \end{definition}
-These properties may look like Byzantine consensus properties (which we describe next in \Cref{sec:acs-background}) but they have some subtle differences.
+These properties may look like Byzantine consensus properties (which we describe next in \Cref{sec:acs-background}) but there are subtle differences.
 Firstly, they are properties for every node in the network and not just the facilitators.
@@
 Before describing the protocol in detail,
-we take a brief detour to give background on the asynchronous subset consensus.
-This is the primary building block of our protocol.
-% Starting with the bootstrap phase and then moving on to the actual consensus phase.
+we take a brief detour to give background on the ACS algorithm as it is the primary building block.
+Then we move on to describe the two phases of our consensus protocol---bootstrap phase and consensus phase.
 
@@
 
-The best way to explain asynchronous subset consensus (ACS) is to contrast it with the typical Byzantine consensus.
+The best way to explain ACS is to contrast it with the typical Byzantine consensus.
 We adapt the description from~\cite[Chapter 17]{podc}.
@@
 \begin{definition}
-\textbf{\emph{Byzantine Consensus}}
+\textbf{\emph{Byzantine consensus}}
 
@@
 
-ACS shares many similarities with Byzantine consensus.
-But it is an especially useful primitive for blockchain systems.
+ACS is an especially useful primitive for blockchain systems.
 It allows any party to propose a value and the result is the set union of all the proposed values by the majority.
-This is the primary difference with Byzantine consensus.
 Concretely, ACS needs to satisfy the following properties (adapted from~\cite{miller2016honey}).
@@
 \label{def:acs}
-\textbf{\emph{Asynchronous Subset Consensus}}
+\textbf{\emph{Asynchronous subset consensus}}
 
@@
 ACS has the nice property of censorship resiliance when compared to other consensus algorithms.
-For instance, Hyperledger and Tendermint uses Practical Byzantine Fault Tolerance (PBFT)~\cite{castro1999practical} as their consensus algorithm.
+For instance, Hyperledger and Tendermint uses a variant of Practical Byzantine Fault Tolerance (PBFT)~\cite{castro1999practical} as their consensus algorithm.
 In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectivly filter transactions.
@@
 For a detailed description of ACS we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.
+To understand the remainder of this work, the knowledge of~\Cref{def:acs} is sufficient.
 
@@
 Now we have all the necessary information to describe our consensus protocol.
-We begin with the bootstrap phase and then move onto the actual consensus phase.
+As with all distributed systems, there must be a bootstrap phase which sets up the system.
 
-Recall that facilitators are computed from the consensus result,
+The reader may notice from the system overview (\Cref{sec:system-overview}) that the facilitators are computed from the consensus result,
 but the consensus result is agreed by the facilitators.
@@
 To bootstrap, imagine that there is some bootstrap oracle, that initiates the code on every node.
-The code satisfied all the properties in \Cref{def:consensus}.
+The code satisfies all the properties in \Cref{def:consensus}.
 Namely every node has the same set of valid facilitators $\F_1$ that are randomly choosen.
 This concludes the bootstrap phase.
+For any future rounds, the consensus phase is used.
 
@@
 The consensus phase begins when $\F_r$ is available to all the nodes.
-Note that $\F_r$ indicates the facilitators that were elected using result of round $r$ and are responsible for driving the ACS protocol in round $r + 1$.
+Note that $\F_r$ indicates the facilitators that were elected using results of round $r$ and are responsible for driving the ACS protocol in round $r + 1$.
 The goal is to reach agreement on a set of new facilitators $\F_{r + 1}$ that satisfies the four properties in \Cref{def:consensus}.
@@
 First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators.
-Second, if the node is a facilitator, it waits until it has received $N - t$ messages of type \texttt{cp\_msg}.
+Second, if the node is a facilitator, it waits until $N - t$ messages of type \texttt{cp\_msg} are received.
 % Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D \gg \Delta$.
-Invalid messages are removed, namely blocks with invalid signatures and duplicate blocks signed by the same key.
+Invalid messages are removed, namely blocks with invalid signatures and blocks signed by the same key.
 With the sufficient number of \texttt{cp\_msg} messages,
 it begins the ACS algorithm and some $\C'_{r + 1}$ should be agreed upon by the end of it.
-Duplicates and blocks with invalid signatures are again removed from $\C'_{r+1}$ and we call the final result $\C_{r+1}$
-We have the remove invalid blocks a second time (after ACS) because the adversary may send different CP blocks to different facilitators,
-which results in invalid blocks in the ACS output.
+Duplicates and blocks with invalid signatures are again removed from $\C'_{r+1}$ and we call the final result $\C_{r+1}$.
+We have to remove invalid blocks a second time (after ACS) because the adversary may send different CP blocks to different facilitators,
+which results in invalid blocks in the ACS output, but not in any of the inputs.
 
@@
 we use a simplification of HoneyBadgerBFT~\cite{miller2016honey} as our ACS protocol
-because it is the only (to the best of our knowledge) consensus algorithms designed for blockchain systems.
+because it is a consensus algorithms designed for blockchain systems.
 We do not use the full HoneyBadgerBFT due to the following.
@@
 Upon receiving $\C_{r + 1}$ and at least $n - f$ valid signatures by some node $u$, $u$ performs two asks.
-First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1}, r + 1)$ (\Cref{alg:new-cp}).
-Second, it computes the new facilitators using $\textsf{get\_facilitator}(C, n)$ (\Cref{alg:facilitator})
+First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1}, r + 1)$ using \Cref{alg:new-cp}.
+Second, it computes the new facilitators using $\textsf{get\_facilitator}(\C_{r + 1}, n)$ using \Cref{alg:facilitator},
 and updates its facilitator set to the result.
 This concludes the consensus phase and brings us back to the beginning of the consensus phase.
-
-\begin{algorithm}
-\caption{Function $\textsf{get\_facilitator}(C, n)$ takes a set of CP blocks $C$ and an integer $n$,
-sort evey element in $C$ by its luck value (the $\lambda$-expression), and outputs the smallest $n$ elements.}
-\label{alg:facilitator}
-\begin{algorithmic}
-\State $\textsf{take} (n, \textsf{sort\_by} (\lambda x.\textsf{H}(x || pk \text{ of } x), C))$
-\end{algorithmic}
-\end{algorithm}
 
@@
 
+\begin{algorithm}
+\caption{Function $\textsf{get\_facilitator}(\C_r, n)$ takes the consensus result $\C_r$ and an integer $n$,
+then sorts the CP blocks $C$ by the luck value (the $\lambda$-expression), and outputs the smallest $n$ elements.}
+\label{alg:facilitator}
+\begin{algorithmic}
+\State $\langle r, C \rangle \gets \C_r$
+\State $\textsf{take} (n, \textsf{sort\_by} (\lambda x.\textsf{H}(\C_r || pk \text{ of } x), C))$
+\end{algorithmic}
+\end{algorithm}
+
 Our protocol has some similarities with synchronisers~\cite[Chapter 10]{podc} because it is effectively a technique to introduce synchrony in an asynchronous environment.
-If we consider the facilitators is a collective authority, then it can be seen as a synchroniser that sends pulse messages (in the form of \texttt{cons\_msg} and \texttt{cons\_sig}) to indicate the start of a new clock pulse.
-Every node then sends a completion messages (in the form of \texttt{cp\_msg}) to the collective authority to indicate that they are ready to start a new pulse.
+If we consider the facilitators as a collective authority, then it can be seen as a synchroniser that sends pulse messages (in the form of \texttt{cons\_msg} and \texttt{cons\_sig}) to indicate the start of a new clock pulse.
+Every node then sends a completion messages (in the form of \texttt{cp\_msg}) to the new collective authority to indicate that they are ready for the next pulse.
 
@@
 This is likely to be the case for new transactions because $\textsf{agreed\_fragment}(\cdot)$ would be $\bot$.
-The next two conditions checks for tampering or missing blocks, if any of these misconducts are detected, then the TX is invalid.
+The next two conditions check for tampering or missing blocks, if any of these misconducts are detected, then the TX is invalid.
+
+We stress that the \emph{unknown} state means that the verifier does not have enough information to continue with the validation protocol.
+If enough information is available at a later time, then the verifier will output either \emph{valid} or \emph{invalid}.
 
@@
 
-Further, the caller of $\textsf{get\_validity}(t_u{u, i}, F_{v, i})$ is not necessarily 
+Further, the caller of $\textsf{get\_validity}(t_{u, i}, F_{v, i})$ is not necessarily 
 $u$\footnote{In practice it often is because after completing the TX protocol the parties are incentivised to check that the counterparty ``did the right thing''.}
@@
 \subsection{Validation protocol}
-With the validity definition, we are ready to construct a protocol for determining the validity of transactions.
-The protocol is a simple response and request protocol (\Cref{alg:vd-proto}).
+Our validation protocol is designed to classify transactions according to the aforementioned validity definition.
+The protocol is a simple response and request protocol shown in \Cref{alg:vd-proto}.
 If $u$ wishes to validate some TX with ID $txid$ and counterparty $v$, it sends $\langle \texttt{vd\_req}, txid \rangle$ to $v$.
@@
         then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
-    \item \emph{Correctness}:
+    \item \emph{Validity}:
         The validation protocol outputs the correct result
@@
 \begin{algorithm}
-\caption{Validation protocol}
+\caption{Validation protocol which runs in the context of $u$}
 \label{alg:vd-proto}
@@
 
-An optimisation we can do is to use an epidemic protocol~\cite{eugster2004epidemic} (also known as gossiping)) instead of our broadcast approach.
+An optimisation we can do is to use an epidemic protocol~\cite{eugster2004epidemic} (also known as gossiping) instead of our broadcast approach.
 Typical epidemic protocols works as follows.
@@
 The upside of using epidemic protocol is that the communication cost is distributed more evenly between nodes.
-This is especially true with a lazy push approach the node who just received a message would push the message ID to its $f$ random neighbours,
-and only push the full message if the neighbour explicitly requests for the message~\cite{leitao2007epidemic}.
+This is especially true with a lazy push approach where the node who just received a message would push the message ID (e.g. digest) to its $f$ random neighbours,
+and only push the full message if the neighbour explicitly requests the message~\cite{leitao2007epidemic}.
 With this, nodes typically only need to receive one consensus result message instead of $n$.
@@
 In many applications however, it is often fine to make timing assumptions.
-For example, TCP relies on timeout for its retransmission and the nLockTime property in Bitcoin transactions makes the transaction unspendable until some time in the future (either Unix time or block height)~\cite{bitcoindevguide}.
+For example, TCP relies on timeout for its retransmission and the \texttt{nLockTime} property in Bitcoin transactions makes the transaction unspendable until some time in the future (either Unix time or block height)~\cite{bitcoindevguide}.
 One limitation of our system is that we use the parameter $N$ in our algorithms, which makes it unsuitable for the permissionless environment where users can join and leave at will.
@@
 such that $D$ is sufficiently long for honest nodes to send their CP blocks to the facilitators.
-Again, choosing the parameter $D$ is difficult and depends on a number of factors such as the network condition, message size, and so on.
+Consequently, this removes the need for a PKI because the collected CP blocks may be from nodes that nobody has seen in the past.
+However, choosing the parameter $D$ is difficult and depends on a number of factors such as the network condition, message size, and so on.
 Overestimating it would make agreed fragments much longer than usual, which increases communication costs for validation.
-Underestimating it would lead to unfairness where users that are too late do not have a chance to be selected as a facilitator in the next round.
+Underestimating it would lead to unfairness where users with a poor internet connection will have a lower chance to be selected as a facilitator in the next round.
 Nevertheless, there is a significant gain for making the timing assumption and that is the ability to operate in the permissionless setting which we explain next.
@@
 There is a subtlety here which happens when the node is elected as a facilitator in the following round.
-In this case, the node must fulfill its oblication by completing the consensus protocol (but without proposing its own CP block) before leaving.
+In this case, the node must fulfill its obligation by completing the consensus protocol, but without proposing its own CP block, before leaving.
 Otherwise the $n \ge 3t + 1$ condition may be violated.
@@
     The chain on the right uses compact blocks, represented by the smaller squares.
-    Compact blocks also form a chain as before, but they each have a hash pointer to the full block, identified by ``Seq'' or ``Digest''.}
+    Compact blocks also form a chain as before, but they each have a hash pointer to the full block, identified by ``Seq'' and ``Digest''.}
     \label{fig:compact}
@@
 (SHA256 digests are 32 bytes each and we use a 64 bit integer to represent the sequence number),
-then we have a 86\% saving in communication cost.
+then the saving in communication cost is 86\%.
 
@@
 This is sufficient for some applications such as proving the existance of some information.
-However, applications such as digital cash where every block depends on some previous block,
-then our scheme is not suitable.
+However, for applications such as digital cash where every block depends on one or more previous blocks,
+our scheme is not suitable.
+For such applications we need to guarantee that there are no forks from the genesis block leading up to the TX block of interest.
 
@@
 We use this approach in our prior work on Implicit Consensus~\cite{implicitconsensus}.
-It may sound inefficient at first, but nodes can employ caching to minimise communication costs.
-In our work we call this ``spontaneous sharding''.
+It may sound inefficient at first, but nodes employ caching to minimise communication costs.
+In our work we call this effect ``spontaneous sharding''.
 
 The second approach is probabilistic but with only a constant communication overhead.
-For a node, observe that if every of its agreed fragment has a transaction with a honest node,
+For a node, observe that if all of its agreed fragment has a transaction with a honest node,
 then the complete chain is effectively validated in a distributed maner.
 The only way for an attacker to make a fork is to make sure that the agreed fragment containing the fork has no transactions with honest nodes.
-We claim such malicious behaviour is prevented probabilisticaly using a challenge-response protocol as follows.
+Such malicious behaviour is prevented probabilisticaly using a challenge-response protocol as follows.
 Suppose node $A$ wish to make a transaction with node $B$.
@@
 If there are more nodes (say $n$) trying to make transactions with the malicious node,
-then the probability that the malicious node does gets caught at least once follows a binomial distribution
+then the probability that the malicious node gets caught at least once follows a binomial distribution
 $$\Pr[X > 0] = \sum_{k = 1}^{n}\binom{n}{k}p^k (1 - p)^{n - k}.$$
+
+% TODO?
+% \subsection{Secure txid generation using a commitment protocol}
 

### 2017-07-30 thesis/chapters/ch4-analysis.tex
Edit ch3, 4 and 5

@@
 Up to this point we described our system specification in detail.
-Of course, specification along does not establish any truths.
-In this chapter, we analyse two aspects of our system.
+Of course, specification alone does not establish any truths.
+In this chapter, we evaluate our system analytically.
 First we show it has the desired properties.
@@
 and show that it out performs classical blockchain systems.
+Finally, we consider the case where the $n \ge 3t + 1$ assumption is violated,
+and show the effect of it in our system.
 
@@
 Then, building on top of it, we show \Cref{def:validation} holds for the validation protocol.
-The resulting theorem shows that only using CP blocks in the consensus algorithm implies consensus on TX blocks,
-in other words, implicit consensus.
+The resulting theorem shows that only using CP blocks in the consensus algorithm implies consensus on TX blocks.
 
@@
 
-\subsection{Analysis of the consensus protocol}
+\subsection{Correctness of the consensus protocol}
 
 We begin our analysis by establishing truths on the four properties in \Cref{def:consensus},
-namely agreement, validity, fairness and termination.
+namely agreement, validity, fairness and termination for an arbitrary round.
 Using these results, we use mathematical induction to show that they hold for all rounds.
@@
 But $n -t$ facilitators must have received $N - t$ valid CP blocks, so $|\C_{r+1}| \ge N-t$, this proves (a).
-There are $n-t$ honest nodes in $\F_{r+1}$ follows from the model, this proves (b).
+There are at least $n-t$ honest nodes in $\F_{r+1}$ which follows from our assumption, this proves (b).
 Finally, since $N-t \ge (n+t) -t = n$ and $\textsf{get\_facilitators}(\cdot)$ outputs $n$ items, $|\F_{r+1}| = n$ and this proves (c).
@@
 \begin{proof}
-We have already established that $|\C_{r+1}| \ge N - t \ge n$ from \Cref{lemma:validity}.
+We already established that $|\C_{r+1}| \ge N - t \ge n$ from \Cref{lemma:validity}.
 Then the proof directly follows from the random oracle model.
@@
 and the termination property of ACS.
-That is, $\F_r$ eventually receives all the CP blocks required to begin ACS.
-ACS eventually terminates.
-Finally the results are eventually dissemminated to all the nodes.
+That is, $\F_r$ eventually receives all the CP blocks required to begin ACS,
+and then ACS eventually terminates.
+Finally, the results are eventually dissemminated to all the nodes.
 \end{proof}
@@
 
-In the base case, agreement, validity fairness and termination follows directly from the bootstrap protocol,
-due to the bootstrap oracle.
+In the base case ($r = 1$), agreement, validity fairness and termination follows directly from the bootstrap protocol.
 Note that the result is $\F_1$, which indicates the facilitators that are agreed in round 1, who are responsible for driving the ACS protocol in round 2.
@@
 
-\subsection{Correctness of validation}
+\subsection{Correctness of the validation protocol}
 The consensus protocol (on CP blocks and facilitators) is the backbone for consensus on transactions.
-In this section we build on top of \Cref{theorem:consensus} to show that most (except liveness) properties in \Cref{def:validation} can be satisfied.
-
-\begin{lemma}
-% (Correctness of the validation protocol)
-The validation protocol outputs the correct result
-according to the validity definition.
-\end{lemma}
-\begin{proof}
-The algorithm (\Cref{alg:get-validity}) is the validity definition.
-\end{proof}
+In this section we build on top of \Cref{theorem:consensus} to show that the properties except liveness in \Cref{def:validation} can be satisfied.
 
@@
 % (Agreement of the validation protocol)
-If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
-then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
+The validation protocol satisfies agreement and validity properties.
 \end{theorem}
 \begin{proof}
-We proof by contradiction.
-Without loss of generality, for some transaction $t$ with an agreed fragment $F$,
+We proof the agreement property by contradiction.
+Without loss of generality, suppose for some transaction with TX block $t$,
 node $u$ decides \emph{valid} but node $v$ decides \emph{invalid}.
@@
 Since $c' (= c'') = \langle \textsf{H}(t'), \dots \rangle$ we must have $\textsf{H}(t') = \textsf{H}(t'')$ and $t' \ne t''$ (because $t''$ is invalid).
-In other words, whoever sent $F''$ must be able to create some $t''$ that has the same digest as $t'$.
-But we assumed that the adversary can only perform polynomial-time algorithm,
-so in order to find $t''$ it needs to query the random oracle exponentially many times.
-Thus we have a contradiction and this completes the proof.
+In other words, the sender of $F''$ must be able to create some $t''$ that has the same digest as $t'$.
+But this is only possible if the adversary can compute the inverse of $H(\cdot)$ with non-negligible probability.
+Thus we have a contradiction and this completes the agreement proof.
+Our protocol uses \Cref{alg:get-validity} which is also the validity definition,
+this completes the validity proof.
 \end{proof}
 
-\Cref{theorem:validation-agreement} is our first major result.
-It shows that consensus on CP blocks would lead to consensus on TX blocks when the nodes are running the validation protocol.
+\Cref{theorem:validation-agreement} shows that agreement on CP blocks would lead to agreement on TX blocks when the nodes are running the validation protocol.
 Just like in our prior work~\cite{implicitconsensus}, we call this behaviour implicit consensus.
-One of the main advantages over running a consensus algorithm on all the transactions is that 
+One of the main advantages of our scheme over running a consensus algorithm on all the transactions is that 
 the rate of transaction is no longer dependent on the consensus algorithm---ACS.
@@
 \subsection{Impossibility of liveness}
-While \Cref{theorem:validation-agreement} is a major result that allows significantly improved performance over traditional blockchain systems,
-it is not perfect.
+\Cref{theorem:validation-agreement} is a major result that allows significantly improved performance over traditional blockchain systems,
+but it does not have all the properties of a typical Byzantine consensus protocol.
 Now we show a negative result, where the liveness property of \Cref{def:validation} cannot be attained.
@@
 \begin{proof}
-We proof by providing a counterexample.
-Suppose nodes $u$ and $v$ correctly perfomed the TX protocol which resulted a transaction $t$.
-Then when $u$ wants to validate $t$, it does so by sending \texttt{vd\_req} message to $v$.
-$v$ can act maliciously and ignore all \texttt{vd\_req} message, thus $t$ can never be validated.
+Suppose nodes $u$ and $v$ correctly perfomed the TX protocol which resulted a transaction.
+Then when $u$ wants to validate it, it does so by sending \texttt{vd\_req} message to $v$.
+$v$ can act maliciously and ignore all \texttt{vd\_req} message, then the transaction can never be validated.
 \end{proof}
@@
 \label{sec:performance-analysis}
-This section aims to analytically answer our research question.
+This section aims to analytically answer the scalability aspect of our research question.
 That is, does the global throughput increase linearly with respect to the population size?
@@
 and then the bandwidth requirement for a single transaction.
-We build on top of those results to analyse do the global throughput analysis.
+We build on top of those results to analyse the global throughput.
 
@@
 \label{sec:cons-complexity}
-The consensus protocol can bee seen as three parts.
-Thus the communication complexity is the sum of these three parts.
+The consensus protocol can bee seen as three parts,
+so the communication complexity is the sum of these parts.
 The first part is when every node sends their CP block to all the facilitators, which is $O(Nn)$ since there are $N$ nodes and $n$ facilitators.
@@
 The communication complexity of ACS is $O(n^2|v| + \lambda n^3 \log n)$~\cite{miller2016honey},
-where $|v|$ is the size of largest message and $\lambda$ is the security parameter.
-Note that the security parameter is the same as the one for our random oracle described in \Cref{sec:model-assumptions}.
-In particular, it is from the use of $\textsf{H}(\cdot)$ in the reliable broadcast phase in ACS.
-In our system, we wish to understand the scalability properties.
+where $|v|$ is the size of largest message and $\lambda$ is the security parameter (described in \Cref{sec:model-assumptions}).
+In our system, we wish to understand the scalability property.
 Thus we consider the complexity as a function of $N$ rather than $n$ or $\lambda$.
-Since $|v|$ is at most all the CP blocks from every node, we have $|v| = cN$,
-where $c$ is a constant representing the size of one CP block.
+Since $|v|$ is at most all the CP blocks from every node, we have $|v| = kN$,
+where $k$ is a constant representing the size of one CP block.
 Therefore the communication complexity of ACS in our system is $O(N)$.
@@
 
+% TODO wrong?
 The third and final part is the dissemination, where the facilitators broadcast the consensus result along with their signatures.
@@
 \label{sec:cons-duration}
-In order to make arguments on bandwidth or throughput, which are concepts that depend on time,
-we must make some assumptions regarding our computational model to make arguments on the duration of the consensus protocol.
+In order to make arguments on bandwidth or throughput in our purely asynchronous model, which are concepts that depend on time,
+we must make some assumptions regarding our model to make arguments on the duration of the consensus protocol.
 Note that it is not the same as the time complexity typically used in distributed systems.
 In analysis of distributed systems, time complexity is often in terms of the number of rounds.
-For example, ACS runs in a constant number of rounds because its subprotocols---reliable broadcast and binary Byzantine consensus---also run in a constant number of rounds.
+For example, ACS runs in a constant number of rounds because its subprotocols---reliable broadcast and binary Byzantine consensus---also run in a constant number of rounds~\cite{miller2016honey}.
 However, in practice, making a unit of communication always has some overhead associated with it, for example serialising and writing it to some network socket.
@@
 % time''~\cite{prasad2003bandwidth}.
-We know that to create and then validate a transaction,
+To create and then validate a transaction,
 the communication cost per transaction is of $O(l)$,
@@
 This is intuitive because round duration would be longer if there are more CP blocks (more $N$), which means that the agreed fragments are longer (assuming nodes transact at a constant rate).
-The behaviour is also verified experimentally in \Cref{ch:implementation}.
+The behaviour is verified experimentally in \Cref{ch:implementation}.
 
-\subsection{Global throughput}
+\subsection{Linear global throughput}
 \label{sec:global-throughput}
@@
 First we clarify the bandwidth definition, which is ``the data rate at which a network link or a network path can transfer''~\cite{prasad2003bandwidth}.
-Now, suppose every node has some fixed bandwidth per link $C$ and $N$ links.
-They also make transactions at $r_{\text{tx}}$ per second.
+Now, suppose every node has some fixed bandwidth $C$ per link, $N$ links and are making transactions at $r_{\text{tx}}$ per second.
 Then we have the inequality
@@
 
-We note that the inequality does not hold if the node is only transacting with a small sample of some subset of the population.
+We note that the inequality does not hold if the node is only transacting with a subset of the population.
 This is because it cannot use all the bandwidth available in all the links.
-For instance, if the node is only transacting with one other node, then it can only use the bandwidth of one link which is only $C$.
-However, if that is the case, we can intelligently cache the \texttt{vd\_resp} messages.
-Thus multiple transactions can be validated with a single \texttt{vd\_resp} message.
+In the extreme case, if the node is only transacting with one other node, then it can only use the bandwidth of one link which is only $C$.
+However, if that is the case, we can intelligently cache the \texttt{vd\_resp} messages as described in~\Cref{sec:caching}.
 For this work, we analyse the worst case where every node transacts with a random node from the population,
@@
 Hence, there exist some constant $r_\text{tx}$ that makes use of all the available bandwidth regardless of $N$.
-Finally, if every node in the network is transacting at a constant rate, then the global throughput (in terms of transactions per second) is of $O(N)$.
+Finally, if every node in the network is transacting at a constant rate, then the global throughput (in terms of transactions per second) is linear w.r.t. the population size $N$.
 If the system is not making use of all the bandwidth.
 We also maintain a constant transaction rate by the same argument.
-Thus a global throughput is also $O(N)$.
-We verify both of these claims experimentally in~\Cref{sec:evaluation}.
+Thus a global throughput is also linear w.r.t. $N$.
+We verify both of these claims experimentally in~\Cref{ch:implementation}.
 
@@
 The problem can be formulated as follows.
-Suppose an urn contains $N$ balls, $t$ are black and $N-t$ are white.
+Suppose an urn contains $N$ balls, $t$ are black (malicious) and $N-t$ are white (honest).
 If $n$ balls are drawn uniformly at random without replacement,
-what is the probability that $\lfloor \frac{n-1}{3} \rfloor$ are black?
+what is the probability that more than $\lfloor \frac{n-1}{3} \rfloor$ are black?
 The random variable $X$ in this case is the number of black balls, or the number of successful events.
@@
   \caption{Plot of the probability of selecting more than $\lfloor \frac{n-1}{3} \rfloor$ black balls for
-  different numbers of }
+  different value of $t$ with $N$ fixed at 2000.}
   \label{fig:hypergeometric}
@@
 
-On the other hand, due to the fact that hypergeometric distributions have light tails,
-that is ``faster-than-exponential fall-off''~\cite{skala2013hypergeometric},
+On the other hand, due to the fact that hypergeometric distributions have light tails with ``faster-than-exponential fall-off''~\cite{skala2013hypergeometric},
 the probability for picking more than $\lfloor \frac{n-1}{3} \rfloor$ black balls when 

### 2017-07-30 thesis/chapters/ch5-implementation.tex
Edit ch3, 4 and 5

@@
 Hence, we give an open source prototype implementation of our system.
-The nodes are capable of forming a network (tested up to 1500 nodes) and send transactions to each other,
-which allows us to confirm our analytical claims.
+Our implementation runs on a network with over 1000 node.
+Each node send transactions to each other autonomously, 
+which allows us to evaluate our analytical claims.
 
-We begin this chapter by a description of the implementation in \Cref{sec:implementation}.
+We begin this chapter with a description of the implementation in \Cref{sec:implementation}.
 Then, we move on to describing our experimental setup in \Cref{sec:experimental-setup} which includes a discussion on the system parameters and the physical infrastructure which we use.
 Finally, \Cref{sec:comms-cost-experiment} and onwards follows the same structure as the Performance analysis (\Cref{sec:performance-analysis}),
-where we analyse the performance experimentally rather than analytically.
+where we analyse the performance experimentally rather than analytically and compare both results.
 
@@
 It also provides the essential algorithm necessary to interact with Extended TrustChain such as 
-$\textsf{new\_tx}(\cdot)$, $\textsf{new\_cp}(\cdot)$, $\textsf{agreed\_fragment}(\cdot)$ and so on.
-Since this is a prototype implementation, we only store the data structure in memory and not on disk.
+$\textsf{new\_tx}(\cdot)$, $\textsf{new\_cp}(\cdot)$ and $\textsf{agreed\_fragment}(\cdot)$.
 Finally, the \texttt{node} module ties everything together.
@@
 The the goal of the experiment is to run the three protocols---consensus protocol,
-transaction protocol and validation protocol---simultaneously and analyse the throughput, consensus duration and other related metrics.
+transaction protocol and validation protocol---simultaneously and analyse the communication cost, the consensus duration and the global throughput.
 We investigate these properties under the following four parameters.
@@
   \item The number of facilitators $n \in \{4, 8, \dots, 32\}$.
-        Unfortunately, the maximum $n$ is 32 because the limitation in liberasurecode mentioned in \Cref{sec:implementation}.
+        The maximum $n$ is 32 because the limitation in liberasurecode mentioned in \Cref{sec:implementation},
+        but our results give a good dindication of how our system may function for larger numbers of $n$.
   \item The population size $N \in \{200, 300, \dots, 1200\}$.
@@
 
-The experiment is run on the DAS-5 (The Distributed ASCI Supercomputer 5\footnote{\url{https://www.cs.vu.nl/das5/}}).
+The experiment is run on the DAS-5\footnote{\url{https://www.cs.vu.nl/das5/}}.
 From now on, we use ``machines'' to refer to DAS-5 nodes and ``nodes'' to refer to a running instance in our system.
-On DAS-5 we use up to 30 machine, for each machine we use 40 nodes.
+On DAS-5 we use up to 30 machine, for each machine we use up to 40 nodes.
 This gives us the aforementioned 1200 number.
@@
 The remainder of this chapter follows the same structure as the performance analysis in~\Cref{sec:performance-analysis}.
-We check our analytical results experimentally and again demonstrate horizontal scalability.
+We check our analytical results experimentally and demonstrate horizontal scalability.
 
@@
 
-\Cref{fig:round-comms} show the relationship between communication cost of the consensus protocol per round and population size.
+\Cref{fig:round-comms} show the relationship between the communication cost of the consensus protocol per round and the population size.
 The most important aspect is that these results show a linear increase.
@@
 We are also interested in how communication costs translate to time.
-Hence, for the same experiment we also record the duration in seconds and the result is shown in~\Cref{fig:round-duration}.
+Hence, for the same experiment we record the duration in seconds and the result is shown in~\Cref{fig:round-duration}.
 Interestingly, the duration is not entirely linear.
@@
 For example, the luck value can be computed by the facilitators and are sent with the consensus result.
-Then the non-facilitator nodes simply checks that they are correct.
+Then the non-facilitator nodes simply use it if there are enough signatures.
 
@@
 
-\section{Global throughput}
+\section{Linear global throughput}
 \begin{figure}[h]
@@
   }
-  \caption{Global throughput}
+  \caption{Global throughput increases as the population increases when every node transact at the same rate.
+  Using fixed neighbours results in a higher throughput because of the caching mechanism.}
   \label{fig:global-throughput}
@@
 
-Note that the throughput decreases as the number of facilitators increases.
+Note that the throughput decreases slightly as the number of facilitators increases.
 This is due to the additional communication cost for running ACS with a high number of facilitators.
@@
 
-The difference in magnitude between \Cref{fig:global-throughput-fixed} and \Cref{fig:global-throughput-random} is also interesting.
-It is caused by the caching mechanism mentioned earlier.
+The difference in magnitude between \Cref{fig:global-throughput-fixed} and \Cref{fig:global-throughput-random} is caused by the caching mechanism mentioned earlier.
 If a new agreed fragment needs to be transmitted to validate every transaction then it puts a toll on the network infrastructure.
@@
 This means that transactions are validated in bursts, and the validation protocol can ``keep up'' with the transactions.
-Note that the horizontal ``lines'' where no new validated transactions are made are roughly 70 seconds, this matches with our result in~\Cref{fig:round-duration-fixed}.
+Note that the horizontal ``lines'' where no new validated transactions are made are on average 76 seconds, this matches with the round duration result in~\Cref{fig:round-duration-fixed}.
 On the other hand in~\Cref{fig:backlog-random}, the validation protocol clearly cannot ``keep up'' with the rate which the transactions are made.

### 2017-07-31 thesis/chapters/ch3-model.tex
Write conclusion and fix references

@@
 The system consist of three protocols---the consensus protocol, the transaction protocol and the validation protocol.
-It is based on our prior work on TrustChain~\cite{trustchain}, where every node independently interact with each other via their own blockchain.
+It is based on our prior work on TrustChain~\cite{multichain},
+where every node independently interact with each other via their own blockchain\footnote{It is originally called MultiChain but the name has changed to TrustChain}.
 
@@
 Thus we first describe the standard TrustChain.
-Our description has minor differences compared to the description in~\cite{trustchain}.
+Our description has minor differences compared to the description in~\cite{multichain}.
 This is to help with the description of the extended TrustChain.
@@
 The TX block must have a valid hash pointer pointing to the previous block
-and a reference\footnote{This is different from the original TrustChain definition found in~\cite{trustchain}.
+and a reference\footnote{This is different from the original TrustChain definition found in~\cite{multichain}.
 In there, a TX block has two outgoing edges which are hash pointers to the two parties involved in the transaction.
@@
 \subsection{Using timing assumption in the permissionless setting}
+\label{sec:permissionless}
 Our model is purely asynchronous, where we make no timing assumptions anywhere in the protocol.

### 2017-07-31 thesis/chapters/ch7-conclusion.tex
Write conclusion and fix references

@@
 
+The research question we asked in~\Cref{ch:problem} is the following.
+\begin{displayquote}
+\emph{Is it possible to design a blockchain consensus protocol that is fault tolerant, scalable and can reach global consensus?}
+\end{displayquote}
+We answer it in the affirmative and our system achieves all the requirements.
+Fault tolerance is guaranteed if $n \ge 3t + 1$ from ACS.
+While $t$ may be small compared to the population size $N$,
+we show that the probability for the system to fail is low even when $n \ge 3t + 1$ does not hold as long as the proportion of malicious nodes is not close to a third of $N$.
+Horizontal scalability is demonstrated both analytically and experimentally.
+With 1200 nodes in the network we were able to achieve nearly 5000 TX/s---an order of magnitude improvement over Bitcoin.
+Finally, we achieve global consensus on transactions via consensus on checkpoint blocks.
+
+This work is the first step in building a new paradigm for blockchain consensus protocol.
+It has the potential to efficiently cultivate trust on the internet in the presence of faults without any central authority.
+We hope to improve our design by building a concrete application on top of it.
+
 \section{Future work}
+While our system has excellent scalability properties, it is not without limitations.
+Much of it is from the fact that it does not have a concrete application.
+We do not attempt to prevent the Sybil attack, spam or DoS because the accuracy of the defence depends on the understanding of the application.
+For instance, without any application it would be impossible to distinguish between a very active node from a spammer because the content of the transaction ($m$) carry no meaning.
+
+Our experiment is carried out in a somewhat idealised world (DAS-5).
+To understand our system further, we hope to experiment on a real platform with real users such as Tribler~\cite{pouwelse2008tribler}.
+For instance, the TrustChain in Tribler stores the number of bytes uploaded and downloaded between users, but it has no global consensus.
+It would be useful to integrate our consensus protocol into Tribler to prevent misreporting attacks.
+More importantly it gives us an opportunity to observe the characteristics of our system in the real world and how it evolves with new features or alterations.
+
+The fault tolerance property of our system is lower than that of typical blockchain systems,
+which usually aim for fewer than 50\%.
+For future work, we propose the following two ways to improve fault tolerance.
+First, use a reputation mechanism to select facilitators rather than simply selecting them randomly so that faulty nodes are less likely to be selected.
+This technique of course heavily depends on the application,
+for P2P file sharing systems the reputation score may be the number of bytes contributed to the network.
+Second, instead of using ACS we use a proof-of-stake (PoS) scheme which means that nodes with more stake in the system gets more vote.
+In an ideal PoS scheme, the system can tolerate any fault if the majority of the stake holders are honest.
+
+Finally, as mentioned in~\Cref{sec:permissionless}, we wish to evaluate our system (analytically and experimentally) in the permissionless settings.
+This would enable our protocol to be applied to a much wider range of applications.

### 2017-08-02 thesis/chapters/ch2-problem.tex
Write conclusion and fix references
Refactor validation protocol

@@
 
-The decentralised and democratic nature of blockchain technology removes central point of control or failure.
-However, as we described in the Introduction, it faces a genuine scalability problem.
-In particular, proof-of-work based systems are unlikely to reach a transaction rate suitable for ubiquitous use, even with parameter tuning.
+From the Introduction, we saw that early blockchain systems face a scalability problem.
+In particular, proof-of-work based systems are unlikely to reach a transaction rate suitable for ubiquitous use.
+Many attempts to fix the scalability problem exist in academia as well as the industry.
+In this chapter we describe the related work in the form of a mini taxonomy and identify areas for improvement and limitations.
+Then we state our research question which
+
+\section{Related work and mini taxonomy}
+
+\section{Research question}
+
+% TODO fix
 What we wish to see is horizontal scalability.
@@
 \begin{displayquote}
-\emph{How do we design a blockchain consensus protocol that is fault tolerant, scalable and can reach global consensus?}
+\emph{Is it possible to design a blockchain consensus protocol that is fault tolerant, scalable and can reach global consensus?}
 \end{displayquote}
@@
 
-\section{Global consensus}
+\subsection{Global consensus}
 
@@
 
-\section{Security and fault tolerance}
+\subsection{Security and fault tolerance}
 
@@
 
-\section{Performance and scalability}
+\subsection{Performance and scalability}
 
@@
 
-\section{Limitations}
+\subsection{Limitations}
 Our work aims to significantly lower the transaction bound.

### 2017-08-02 thesis/chapters/ch3-model.tex
Refactor validation protocol

@@
 A transaction can be in one of three states in terms of validity---\emph{valid}, \emph{invalid} and \emph{unknown}.
-Given a fragment $F_{v, j}$, the validity of the transaction $t_{u, i}$ is captured by the function $\textsf{get\_validity}(t, F)$ (\Cref{alg:get-validity}).
+Given a fragment $F_{v, j}$, the validity of the transaction $t_{u, i}$ is captured by the function $\textsf{get\_validity}(t, F_u, F_v)$ (\Cref{alg:get-validity}).
 The first four conditions (up to \Cref{line:valid-fragment}) essentially check whether the fragment is the one that the verifier needs.
@@
 
-Further, the caller of $\textsf{get\_validity}(t_{u, i}, F_{v, i})$ is not necessarily 
+Further, the caller of $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ is not necessarily 
 $u$\footnote{In practice it often is because after completing the TX protocol the parties are incentivised to check that the counterparty ``did the right thing''.}
-Any node $w$ may call $\textsf{get\_validity}(t_u{u, i}, F_{v, i})$ as long as the caller $w$ has an agreed fragment of $t_{u, i}$---$F_{u, i}$.
+Any node $w$ may call $\textsf{get\_validity}(t_u{u, i}, F_{u, i}, F_{v, j})$ as long as the caller $w$ has an agreed fragment of $t_{u, i}$---$F_{u, i}$.
 $F_{u, i}$ may be readily available if $w = u$ or it may be from some other \texttt{vd\_resp} message, which we describe next in the validation protocol.
@@
 \begin{algorithm}
-\caption{Function $\textsf{get\_validity}(t_{u, i}, F_{v, j})$ validates the transaction $t_{u, i}$.
+\caption{Function $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ validates the transaction represented by $t_{u, i}$.
+We assume $F_{u, i}$ is always correct and contains $t_{u, i}$.
 $F_{v, j}$ is the corresponding fragment received from $v$.}
-We assume there exist a valid $F_{u, i}$, namely the agreed fragment of $t_{u, i}$. 
-The caller is $w$, it may be $u$ but this is not necessary.
 \label{alg:get-validity}
@@
 
-    % \State $F_{u, i} \gets \textsf{agreed\_fragment}(t_{u, i})$
-    % \If{$F_{u, i} = \bot$}
-    %     \State \Return \emph{unknown}
-    % \EndIf \Comment $u$ has agreed fragment
-    % \State
-
-    \State $c_{v, a} \gets \textsf{first}(F_{v, j})$
-    \State $c_{v, b} \gets \textsf{last}(F_{v, j})$
-    \If{$c_{v, a}$ or $c_{v, b}$ are not in consensus}
+    \If{$F_{v, j}$ is not a fragment created in the same round as $F_{u, i}$}
         \State \Return \emph{unknown}
     \EndIf
-    \State \Comment $v$ has agreed fragment
+    \State
 
-    % \If{$|F_{v, j}| > L$}
-    %     \State \Return \emph{unknown}
-    % \EndIf \Comment fragment not too long
-    % \State
-
-    % txid are equal now, but bad hash pointer
-    \If{sequence number in $F_{v, j}$ is correct (sequential)}
-        \If{hash pointers in $F_{v, j}$ is wrong}
-            \State \Return \emph{unknown}
-        \EndIf
-    \EndIf
-    \State \Comment correct TrustChain structure
-
-    \State $c_{u, b} \gets \textsf{last}(F_{u, i})$
-    \If{$c_{u, b}$ is not created using the same $\C_r$ as $c_{v, b}$}
-        \State \Return \emph{unknown}
-    \EndIf
-    \State \Comment correct consensus round
     \label{line:valid-fragment}
@@
     \EndIf
-    \State \Comment TX exists
+    \State
 
-    \State $\langle \_, \_, txid', pk'_u, m', \_ \rangle \gets t_{v, j}$
-    \If{$m \ne m' \vee pk_u \ne pk'_u$}
+    \State $t_{v, j} \gets $ the TX block with $txid$ in $F_{v, j}$
+    \State $\langle \_, \_, txid', pk_u, m', \_ \rangle \gets t_{v, j}$
+
+    % message tampering
+    \If{$m \ne m' $} 
         \State \Return \emph{invalid}
     \EndIf
-    \State \Comment no tampering
+    \State
+
+    % signature tampering
+    \If{$t_{u, i} \text{ is not signed by } pk_u \vee t_{v, j} \text{ is not signed by } pk_v $}
+        \State \Return \emph{invalid}
+    \EndIf
     \State \Return \emph{valid}
@@
         \State $t_{u, i} \gets \text{the transaction identified by } txid$
-        \State set the validity of $t_{u, i}$ to $\textsf{get\_validity}(t_{u, i}, F_{v, j})$
+        \If{$F_{u, i}$ and $F_{v, j}$ are available and $F_{u, i}$ is the agreed fragment of $t_{u, i}$}
+            \State set the validity of $t_{u, i}$ to $\textsf{get\_validity}(t_{u, i}, F_{u, j}, F_{v, j})$
+        \EndIf
 \end{algorithmic}

### 2017-08-03 thesis/chapters/ch1-intro.tex
Edit ch3, 4 and 5
Rewrite problem description; add contribution

@@
 
+Our main contributions are the following.
+\begin{itemize}
+    \item We formally introduce a blockchain consensus protocol that uses individual hash chains on every node and achieves
+        horizontally scalable (through checkpoint blocks) for the first time.
+    \item We formally analyse our consensus protocol to ensure correctness as defined in our architecture.
+    \item We implement our consensus protocol with experiment with up to 1200 nodes,
+        our results show strong evidence of horizontal scalability.
+\end{itemize}
+
 We describe and analyse our idea in the remainder of this work.
-We begin with the problem description in~\Cref{ch:problem}.
-This is followed by a detailed formulation of the model in~\Cref{ch:model}. 
+We begin with the problem description in~\Cref{ch:problem} which describes the state-of-the-art work on making blockchain systems scalable and our research question.
+In~\Cref{ch:model} we give a formal description of our system model and architecture.
 Next, we analyse the correctness and performance of our design in~\Cref{ch:analysis}, this is where we present our key theorems.
 Implementation and experimental results are discussed in~\Cref{ch:implementation}.
-Finally, we compare our system with other state-of-the-art blockchain systems and conclude in Chapters \ref{ch:related} and \ref{ch:conclusion} respectively.
+Finally, we conclude in \ref{ch:conclusion} and discuss future work.

### 2017-08-03 thesis/chapters/ch2-problem.tex
Rewrite problem description; add contribution

@@
 From the Introduction, we saw that early blockchain systems face a scalability problem.
-In particular, proof-of-work based systems are unlikely to reach a transaction rate suitable for ubiquitous use.
+In particular, the use proof-of-work is unlikely to reach a transaction rate suitable for ubiquitous use.
 Many attempts to fix the scalability problem exist in academia as well as the industry.
 In this chapter we describe the related work in the form of a mini taxonomy and identify areas for improvement and limitations.
-Then we state our research question which
+Then we state our research question which aims to overcome some of the limitations.
 
-\section{Related work and mini taxonomy}
+\section{Related work and a mini taxonomy}
+\label{sec:taxonomy}
+
+Blockchain research has seen a surge in recent years from both the industry and academia,
+this resulted in a myriad of of blockchain systems with different approaches to scalability.
+To understand the current state-of-the-art,
+we classify various blockchain systems by their scalability approach.
+
+\subsection{Early blockchain systems}
+This category represents the baseline,
+which are systems with a probabilistic probabilistic consensus algorithm.
+That is, transactions never reach consensus with a probability of 1.
+The typical examples are proof-of-work based systems such as Bitcoin, Ethereum and other Altcoins.
+In Bitcoin, the level of consensus of a block containing many transactions is determined by how deep it is in the Bitcoin blockchain,
+also called the number of confirmation.
+The probability of a block being orphaned drops exponentially as the depth increases~\cite{bitcoin}.
+Nevertheless, the probability of the highest block being orphaned is non-negligible.
+
+The advantage of this approach is that it can be used in a large network.
+In fact, prior to Bitcoin there were no internet-scale consensus protocol.
+Bitcoin is also fault tolerant,
+because attackers can not out pace honest users in finding new blocks unless they have a majority of the hash power.
+As long as the majority is honest and reasonable measures are taken to prevent the selfish mining attack~\cite{eyal2014majority}.
+The disadvantage however is that transactions are never in consensus with a probability of 1---no consensus finality.
+Hence double spending is possible for transactions that do not have a large number of confirmations.
+Furthermore, as we mentioned in the Introduction,
+the performance is limited due to the fact that blocks are of fixed size and are generated at fixed intervals.
+Thus early blockchain systems do not achieve horizontal scalability.
+
+% Our system significantly improves upon Bitcoin and other classical blockchain systems in performance, scalability and consensus finality.
+% The results described in \Cref{ch:analysis} and \Cref{ch:implementation}, show that we have horizontal scalability,
+% where more nodes result in more global throughput.
+% Further, we do not have the aforementioned probabilistic behaviour,
+% once some consensus result is decided, it cannot be orphaned, thus our consensus is final.
+% The leadership election is also not ideal in classical blockchain systems.
+% Mining can be seen as a technique to elect a single leader.
+% The leader has full control of what goes into the block thus it may selectively censor transactions.
+% We use ACS, so as long as the CP block is in $n - 2t$ nodes, it is guaranteed to be in consensus.
+
+% However, the security aspect falls short of the ``honest majority'' security model that classical blockchains claim to 
+% have\footnote{Recently it was shown that doing selfish-ming would give the adversary an unfair advantage when she only controls 25\% of the mining power~\cite{eyal2014majority}}.
+% Our system risks going into erroneous state if the inequality $n \ge 3t + 1$ is not satisfied.
+% Classical blockchain systems also have an incentive mechanism, thus they do not depend on altruistic nodes and encourages participation.
+% Our system on the other hand does not have an incentive mechanism because we make no assumptions on the application.
+
+\subsection{Off-chain transactions and payment networks}
+Off-chain transactions make use of the following fact.
+If nodes make frequent transactions,
+then it is not necessary to store every transaction on the blockchain,
+only the net settlement is necessary.
+The best examples are Lightning Network~\cite{lightningnetwork} and Duplex Micropayment Channels~\cite{decker2015fast}.
+
+Off-chain transaction systems are implemented using multi-signature addresses~\cite{bitcoinmultisig} and hashed time-locked Bitcoin contracts~\cite{bitcointimelock}.
+In the simiplest case, 
+if two parties wish to make transactions,
+they open a payment channel in the form of of a multi-signature Bitcoin address (for two parties it would be a 2-of-2 signature address).
+Each of the party must also deposit some Bitcoins in the multi-signature address.
+Both parties keep the channel state which is not broadcasted to the network.
+The state is updated when the two parties make new transactions.
+The protocol disincentivises the parties from broadcasting old channel states.
+If this issue occurs, the counterparty can drain all the Bitcoins in the channel.
+After the payment channel is closed, the net transaction is broadcasted to the Bitcoin blockchain.
+The two-party scheme can be extended to a network of channels,
+which allows two parties to make off-chain transactions without an open channel as long as they are connected by nodes that do.
+
+The advantages of such systems is that they act as add-ons to Bitcoin wish already has a large number of users.
+Thus, if enough of the network wish for it (by setting a new block version),
+then a large number of users will instantly benefit from it.
+Some implementation already exist\footnote{For example \url{https://github.com/lightningnetwork/lnd}.},
+but at the time of writing Segwit is not activated yet which is a prerequisite for Lightning Network.
+
+On the other hand, off-chain transactions also suffers from the problem of Bitcoin.
+Proof-of-work still consumes an unreasonable amount of power.
+Further, payment channels complicates user experience.
+As we mentioned earlier, each node must deposit some Bitcoins into a multi-signature account.
+The users must pick a suitable amount;
+if the deposit is too low it would not allow large transactions,
+if the deposit is too high the user is locked out of much of its Bitcoins for use outside of the channel.
+In addition, the user must proactively check whether the counterparty has broadcasted an old channel state so that the user does not lose Bitcoins.
+Payment channel in theory solves the scalability problem of early blockchain systems,
+but to the best of our knowledge its exact scalability characteristics are not investigated.
+
+% Further, off-chain transactions are limited only to an exchange of cryptocurrency,
+% it is less general than typical Bitcoin transactions which may be a simple smart contract.
+% Time-locked contracts have a strong dependency on timing, thus disputes may arise when the payment channel is just about to close.
+% Our system is purely asynchronous and make no assumption on timing,
+% in fact we assume the adversary has control of the message delivery time and order.
+
+\subsection{Permissioned systems based on Byzantine consensus}
+
+This category of systems use traditional Byzantine consensus algorithms such as PBFT~\cite{castro1999practical}.
+In essence, they contain a fixed set of nodes (sometimes called validating peer) that run a Byzantine consensus algorithm to decide on new blocks.
+This is often used in permissioned system where the validators must be pre-determined,
+for example Hyperledger Fabric~\cite{cachin2016architecture}.
+
+A nice aspect of Byzantine consensus and in particular PBFT is that it can handle much more transactions than classical blockchain systems.
+PBFT can for example achieve 10,000 TX/s if the number of validating peer is under 16~\cite[Section 5.2]{miller2016honey}.
+Further, in contrast to proof-of-work, PBFT consensus is final.
+That is, transaction history cannot be re-written if it is already in the blockchain.
+
+The major drawback of Byzantine consensus based systems is that it does not scale in terms of the number of validating peers.
+Going back to PBFT, its transaction rate drops to under 5000 TX/s when the number of validating peer is 64~\cite[Section 5.2]{miller2016honey}.
+Moreover, the validating peers are pre-determined which makes the system unsuitable for the open internet.
+
+% But our system has the potential to perform beyond that of PBFT because we represent many transactions with a single CP block,
+% enabling horizontal scalability.
+% Furthermore, since PBFT relies on a leader, it is not censorship resiliant.
+% Our system on the other hand has the benifits of ACS where CP blocks cannot be censored.
+% Finally, our system is able to work in the permissionless setting by simply submitting new CP blocks to the facilitators.
+% It can even be adapted to work in the permissioned by simply removing the luck value computation.
+
+% What we wish to have which is in Hyperledger is a smart contract system (also known as chaincodes in Hyperledger).
+% We hope to design and implement smart contracts by adding additional logic to the transaction protocol and the validation protocol.
+% Such functionalities we believe is better to be built into the backbone rather than having it as an add on.
+
+\subsection{Combining proof-of-work with Byzantine consensus}
+Recent research has developed a class of hybrid systems which uses PoW for committe election,
+and Byzantine consensus algorithms to agree on transactions.
+This design is primarily for permissionless system because the PoW leader election aspect prevents Sybils.
+% And then they use a Byzantine consensus algorithm to actually reach consensus on a set of transactions within the committee.
+Some examples are SCP~\cite{luu2015scp}, ByzCoin~\cite{kogias2016enhancing} and Solidus~\cite{abraham2016solidus}.
+
+This technique overcomes the early blockchain scalability issue by delegating the transaction validation to the Byzantine consensus protocol (e.g. PBFT in ByzCoin~\cite{kogias2016enhancing}).
+A major tradeoff of such systems is that they cannot guarantee correctness when there is a large number of malicious nodes (but less than majority).
+For SCP, ByzCoin and Solidus, they all have some probability to elect more than $t$ Byzantine nodes into the committee,
+where $t$ is typically just under a third of the committee size (a lower bound of Byzantine consensus~\cite{pease1980reaching}).
+This problem is especially difficult to solve because the committee is always much smaller than the population size which has more than $t$ Byzantine nodes.
+Classical blockchain do not have this problem because they do not use Byzantine consensus.
+Further, due to the fact that these system must reach consensus on all transactions, none of them achieves horizontal scalability.
+
+% Our approach share many similarities with the hybrid systems.
+% First, we also elect a committee (facilitators) to drive consensus.
+% But we do not have proof-of-work for Sybil defence because we believe 
+% it is possible to do it efficiently, e.g. using NetFlow~\cite{pimotte}.
+% Secondly, our use of CP blocks and ACS is also unique.
+% This creates a much higher throughput and enables censorship resiliance as mentioned earlier.
+% For instance, ByzCoin performs just below 1000 TPS with a thousand nodes whereas we peak at 8000 TPS.
+
+\subsection{Sharding}
+
+Sharding is a technique to achieve horizontal scalability by grouping nodes into multiple committees or shards.
+Nodes within a single shard run a Byzantine consensus algorithm to agree on a set of transactions that belong to that specific shard.
+An intra-shard protocol is needed for transactions that involve nodes from more than one shard.
+The number of shard grow linears with respect to the total computational power in the network.
+This scheme achieves horizontal scalability because if every shard commits transactions at the same throughput,
+then adding more shard would naturally result in a linear increase of global throughput.
+Examples of blockchain systems that use sharding are Elastico~\cite{luu2016elastico} and OmniLedger~\cite{kokoris2017omniledger}.
+
+The biggest limitation of sharding is that it is only optimal is transactions stay in the same shard.
+In fact, Elastico cannot atomically process inter-shard transactions.
+OmniLedger has an inter-shard transaction protocol but choosing a good shard size is difficult.
+A large shard size would make the system less scalable, because the Byzantine consensus algorithm must be run by a large number of nodes.
+A small shard size would result in a large number of inter-shard transactions which also hinder scalability.
+The authors of OmniLedger noted that inter-shard transactions have significantly higher latency compared to the consensus protocol~\cite{kokoris2017omniledger}.
+Furthermore, since no shards can be compromised for the system to function correctly,
+the adversaries have more opportunites to compromise the system.
+
+\subsection{Blockchain without global consensus}
+
+Tangle~\cite{tangle}, Corda~\cite{corda} and the original TrustChain~\cite{multichain} do not use global consensus at all.
+In this scheme, every node has their own chain.
+Transactions between nodes are recorded on their respective chains.
+This effectively results in a directed acyclic graph.
+By avoiding global consensus, this approach achieves extremely desirable scalability properties similar to BitTorrent~\cite{cohen2003incentives}.
+
+On the other hand, the application and security is limited.
+A malicious node can easily lie to other nodes regarding its own chain by simply presenting one version of reality to a set of nodes and another version to a different set.
+Thus the network will have a conflicting view on the state of the malicious node.
+Applications such as digital cash can not be implemented on top of such systems because consistency is needed in solving the double spending problem.
+
+% Our system can be considered as the same as these types of blockchains but with a lightweight consensus protocol.
+% Consensus might not be applicable in all applications.
+% But we believe it is important for detecting and preventing fraud.
+% The example in~\Cref{fig:trustchain-bad} on page~\pageref{fig:trustchain-bad} demonstrates this.
+% If $b$ makes a fork and $a$ and $c$ have no way to communicate (e.g. the adversary may control parts of the network) ,
+% then $c$ is tricked to believe that her transaction with $b$ is valid.
+% Only when $c$ sees a conflicting chain is she able to tell that the transaction is invalid.
+% But $c$ does not know the true end-of-chain of $b$, thus she can never know whether her transaction is valid.
+% This is not possible in our system because $b$ cannot convince $c$ unless he can compute exponential time algorithms
+% (this is finding the second preimage for a hash function).
+
+% However, consensus and validation comes at a cost that do not exist in Tangle, Corda or the original TrustChain.
+% Our transaction rate is affected by the validation protocol in the worst case scenario as we saw in~\Cref{sec:evaluation}.
 
@@
 
-% TODO fix
-What we wish to see is horizontal scalability.
+We saw in~\Cref{sec:taxonomy} that the majority of the systems cannot achieve horizontal scalability.
 That is, by adding new nodes into the network, the global transaction rate should increase proportionally.
-Hence, the research question we answer in this  work is the following.
+To the best of our knowledge, only two systems---Elastico~\cite{luu2016elastico} and OmniLedger~\cite{kokoris2017omniledger}---has experimentally demonstrated horizontal scalability.
+However, these techniques rely on sharding, thus the performance highly depends on transaction characteristics and the shard size parameter because inter-shard transactions are slow.
+Furthermore, all the systems described in~\Cref{sec:taxonomy} are complete blockchain systems designed for a concrete application.
+We wish to design a blockchain consensus protocol that can be used in many applications.
+For example recording uploads and downloads between peers in Tribler~\cite{pouwelse2008tribler, multichain} and decentralised market\footnote{\url{https://github.com/Tribler/tribler/blob/devel/Tribler/community/tradechain/__init__.py}}.
+Hence our research question is the following.
 \begin{displayquote}
@@
 
-% why consensus is important
-% The first objective is to reach consensus on a global \emph{state}.
-% We emphasise the word state because in contrary to traditional blockchains,
-% it does not neccessarily mean the set of all transactions that has ever happened.
-% The state can be a verifiable representation of all the transactions.
-% As an extreme example, a single digest of the complete history can represent the state.
-% It is even verifiable as long as there are nodes that stores the history.
+% Early blockchain systems use PoW as the consensus algorithm,
+% but consensus algorithm and especially Byzantine\footnote{Byzantine means participants can experience arbitrary failure, more on this in~\Cref{sec:acs-background}.}
+% consensus algorithms have existece since the 80s~\cite{pease1980reaching, lamport1982byzantine}.
+% The Practical Byzantine Fault Tolerance (PBFT) algorithm introduced in 1999 by Castro and Liskov~\cite{castro1999practical} is widely regarded as the first high-performance consensus algorithm.
+% The authors show that running PBFT in Network File System (NFS) was only 3\% slower than the standard unreplicated NFS.
+% Since then, research in this area picked up pace and many more algorithms are introduced, for example Q/U~\cite{abd2005fault} and Zyzzyva~\cite{kotla2007zyzzyva}.
+% With the introduction of Bitcoin's probabilistic consensus model, the field experienced a second renaissance.
+% These include proof-of-stake protocols~\cite{bentov2016snow, micali2016algorand}, as well as protocols optimised for blockchain~\cite{liu2016xft, miller2016honey}.
 
-Early blockchain systems use PoW as the consensus algorithm,
-but consensus algorithm and especially Byzantine\footnote{Byzantine means participants can experience arbitrary failure, more on this in~\Cref{sec:acs-background}.}
-consensus algorithms have existece since the 80s~\cite{pease1980reaching, lamport1982byzantine}.
-The Practical Byzantine Fault Tolerance (PBFT) algorithm introduced in 1999 by Castro and Liskov~\cite{castro1999practical} is widely regarded as the first high-performance consensus algorithm.
-The authors show that running PBFT in Network File System (NFS) was only 3\% slower than the standard unreplicated NFS.
-Since then, research in this area picked up pace and many more algorithms are introduced, for example Q/U~\cite{abd2005fault} and Zyzzyva~\cite{kotla2007zyzzyva}.
-With the introduction of Bitcoin's probabilistic consensus model, the field experienced a second renaissance.
-These include proof-of-stake protocols~\cite{bentov2016snow, micali2016algorand}, as well as protocols optimised for blockchain~\cite{liu2016xft, miller2016honey}.
-
-The first objective is to reach consensus on a global \emph{state} by standing on the shoulder of giants.
+The first objective is to reach consensus on a global \emph{state} by building on top of decades of Byzantine consensus protocol research~\cite{pease1980reaching, lamport1982byzantine, castro1999practical, miller2016honey}.
 Having consensus is essential in blockchain systems.

### 2017-08-03 thesis/chapters/ch3-model.tex
Rewrite problem description; add contribution

@@
 $n \ge 3t + 1$ must hold.
-This is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$.
+This is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$~\cite{pease1980reaching}.
 Further, the inequality $N \ge n + t$ must also hold.

### 2017-08-04 thesis/appendices/consensus-example.tex
Write more design variations
Write varying facilitator; describe proto example

@@
-\chapter{Consensus Example}
+\chapter{Consensus example}
 \label{app:consensus-example}
@@
 \begin{figure}[htb]
-    \frame{\includegraphics[trim={0 0 0 11cm}, clip, width=1.0\textwidth]{trustchain-1}}
+    \includegraphics[trim={0 0 0 11cm}, clip, width=1.0\textwidth]{trustchain-1}
     \centering
-    \caption{Initial state}
+    \caption{We begin in a state where some facilitators $\F_{r-2}$ just agreed on the consensus result $\C_{r-1}$ but have not yet propogated it yet.}
     \label{fig:trustchain-1}
@@
 \begin{figure}[htb]
-    \frame{\includegraphics[trim={0 0 0 8cm}, clip, width=1.0\textwidth]{trustchain-2}}
+    \includegraphics[trim={0 0 0 8cm}, clip, width=1.0\textwidth]{trustchain-2}
     \centering
-    \caption{Initial state}
+    \caption{The facilitators propogates $\C_{r-1}$. Upon receiving it, the nodes perform two tasks;
+    (1) elect new facilitators $\F_{r-1}$ by selecting the first $n$ nodes ordered by $\textsf{H}(\C_{r-1} || pk)$,
+    and (2) create new CP blocks---$c_{a, 6}, c_{b, 4}$ and $c_{c, 8}$.
+    $\textsf{H}(\cdot)$ is a cryptographically secure hash function and $pk$ is the public key of the respective node.}
     \label{fig:trustchain-2}
@@
 \begin{figure}
-    \frame{\includegraphics[trim={0 0 0 2cm}, clip, width=1.0\textwidth]{trustchain-3}}
+    \includegraphics[trim={0 0 0 2cm}, clip, width=1.0\textwidth]{trustchain-3}
     \centering
-    \caption{Initial state}
+    \caption{Nodes now send their new CP blocks to the new facilitators $\F_{r-1}$.
+    While $\F_{r-1}$ is trying to reach consensus on a set of CP blocks,
+    nodes carry on making transactions as usual (concurrently).
+    We remark that the to-be-agreed consensus result $\C_r$ is created using CP blocks from the previous round---$c_{a, 6}, c_{b, 4}$ and $c_{c, 8}$.
+    }
     \label{fig:trustchain-3}
@@
 \begin{figure}
-    \frame{\includegraphics[width=1.0\textwidth]{trustchain-4}}
+    \includegraphics[width=1.0\textwidth]{trustchain-4}
     \centering
-    \caption{Initial state}
+    \caption{Finally, when $\F_{r-1}$ decides on $\C_r$ (which should contain $c_{a, 6}, c_{b, 4}$ and $c_{c, 8}$) and disseminates it.
+    Nodes compute new facilitators $\F_{r}$ and create new CP blocks.}
     \label{fig:trustchain-4}

### 2017-08-04 thesis/chapters/ch2-problem.tex
Write varying facilitator; describe proto example

@@
 they open a payment channel in the form of of a multi-signature Bitcoin address (for two parties it would be a 2-of-2 signature address).
-Each of the party must also deposit some Bitcoins in the multi-signature address.
+Each of the party must also deposit some Bitcoins in the multi-signature address,
+this is called the opening transaction.
 Both parties keep the channel state which is not broadcasted to the network.
@@
 If this issue occurs, the counterparty can drain all the Bitcoins in the channel.
-After the payment channel is closed, the net transaction is broadcasted to the Bitcoin blockchain.
+To close the channel, the parties simply boradcast the latest net settlement to the Bitcoin network,
+this is called the closing transaction.
+Effectively, only two transactions need to be broadcasted to the Bitcoin network---the opening and the closing transaction,
+even if the two parties made thousands or millions of transactions in between.
 The two-party scheme can be extended to a network of channels,

### 2017-08-04 thesis/chapters/ch5-implementation.tex
Write varying facilitator; describe proto example

@@
 \end{figure}
+
+\section{Communication cost with varying number of facilitators}
+Up to this point we focused on the effect of communication cost, throughput and so on with respect to the population size.
+In this section we consider a varying number of facilitators,
+which gives us an insight on our system performance may perform when the number of facilitators are higher than 32.
+
+\begin{figure}[tb]
+  \centering
+  \makebox[\linewidth][c]{%
+  \begin{subfigure}[b]{0.7\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-fixed/consensus-communication-cost-vs-facilitators}
+    \caption{Transactions are with fixed neighbours.}
+    \label{fig:facilitators-comms-fixed}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.7\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-random/consensus-communication-cost-vs-facilitators}
+    \caption{Transactions are with random neighbours.}
+    \label{fig:facilitators-comms-random}
+  \end{subfigure}
+  }
+  \caption{The communication cost of the consensus protocol increase polynomially with respect to the number of facilitators as expected from the ACS communication complexity.}
+  \label{fig:facilitators-comms}
+\end{figure}
+
+\Cref{fig:facilitators-comms} shows the communication cost of the consensus protocol for varying numbers of facilitators.
+There is strong evidence that the communication cost grows polynomially.
+We expect this because of there are polynomial terms in the ACS communication complexity---$O(n^2|v| + \lambda n^3 \log n)$.
+This also means that the rounds would take longer to complete and transactions would take longer to verify.
+On the other hand, we get better fault tolerance.
+Thus a linear increase in fault tolerance ($t$) costs a polynomial increase in communication cost.

### 2017-08-04 thesis/chapters/ch1-intro.tex
Remove unused files

@@
 This immediately show that those transactions are tamper-proof.
+We call our protocol ``Checkpoint Consensus''.
 % TODO mention consensus and trade off?
@@
 \begin{itemize}
-    \item We formally introduce a blockchain consensus protocol that uses individual hash chains on every node and achieves
-        horizontally scalable (through checkpoint blocks) for the first time.
-    \item We formally analyse our consensus protocol to ensure correctness as defined in our architecture.
-    \item We implement our consensus protocol with experiment with up to 1200 nodes,
+    \item We formally introduce a blockchain consensus protocol---Checkpoint Consensus.
+        It uses individual hash chains and checkpoints on every node to achieve
+        horizontally scalable for the first time.
+    \item We formally analyse Checkpoint Consensus to ensure correctness as defined in our architecture.
+    \item We implement Checkpiont Consensus with experiment with up to 1200 nodes,
         our results show strong evidence of horizontal scalability.

### 2017-08-04 thesis/chapters/ch3-model.tex
Remove unused files

@@
 
-We design a novel system in order to achieve our ambitious goals outline in~\Cref{ch:problem}.
-The system consist of three protocols---the consensus protocol, the transaction protocol and the validation protocol.
+Our system---Checkpoint Consensus consist of three protocols---the consensus protocol, the transaction protocol and the validation protocol.
 It is based on our prior work on TrustChain~\cite{multichain},
@@
 \label{sec:system-overview}
-The architecture of our system is visualised in~\Cref{fig:architecture}.
+The Checkpoint Consensus architecture is visualised in~\Cref{fig:architecture}.
 \begin{figure}[h]
@@
 \label{sec:combined-protocol}
-The final protocol is essentially the concurrent composition of the three aforementioned protocols,
+The final result is essentially the concurrent composition of the three aforementioned protocols,
 all making use the Extended TrustChain data structure.

### 2017-08-04 thesis/chapters/ch4-analysis.tex
Remove unused files

@@
 \Cref{theorem:validation-agreement} shows that agreement on CP blocks would lead to agreement on TX blocks when the nodes are running the validation protocol.
-Just like in our prior work~\cite{implicitconsensus}, we call this behaviour implicit consensus.
+Like in our prior work~\cite{implicitconsensus}, we call this behaviour implicit consensus.
 One of the main advantages of our scheme over running a consensus algorithm on all the transactions is that 
@@
 % Another way to think about it is if a transaction can be forged, then it cannot be agreed.
+
+A stronger version of the validity definition exists.
+That is, if two nodes are making transactions with each other,
+then the resulting transaction is always valid in addition to our current validity definition.
+Under our purely asynchronous model, we cannot guarantee this stronger version.
+Since the adversary can delay any message for any amount of time,
+it can make sure all \texttt{tx\_req} messages are delivered in a round later than the round which the message is sent.
+Effectively, the transaction pair would always be in different rounds and the validation protocol would not output \emph{valid}.
+We believe in a relaxed model, i.e. a weakly synchronous model, a stronger validity definition is possible.
 
@@
 On the other hand, if $\alpha$ is fixed, but small, we may increase $n$ to achieve the same.
-To put this into perspective, suppose $n = 1000$ and $\alpha = 1/20$, then the probability to draw more black balls than the threshold is only $2 \times 10^{-48}$.
+To put this into perspective,
+suppose $n = 1000$ and $\alpha = 1/5$, i.e. a fifth of the nodes are malicious.
+Then the probability to draw more black balls than the threshold is only $2.6 \times 10^{-16}$.

### 2017-08-04 thesis/chapters/ch6-conclusion.tex
Remove unused files

--- 2017-07-31 thesis/chapters/ch7-conclusion.tex
@@
 \label{ch:conclusion}
+
+In this work we introduced an application neutrual blockchain system which we call Checkpoint Consensus.
+We add checkpoint block to the existing TrustChain data structure for use in our consensus protocol.
+The round based consensus protocol uses ACS as a building block to reach consensus on checkpoint blocks.
+The consensus result lets nodes elect new facilitators and create new checkpoint blocks.
+Node make transactions via the transaction protocol,
+which is adapted from an existing work called True Halves.
+We introduce a validation protocol which ensure that if agreed fragments for some transaction exists,
+then nodes reach agreement on the validity of that transaction.
+The novelty of the validation protocol is that it uses point-to-point communication,
+i.e. nodes only validate the transactions of interest,
+this enables our horizontal scalability property.
 
@@
 \end{displayquote}
-We answer it in the affirmative and our system achieves all the requirements.
-Fault tolerance is guaranteed if $n \ge 3t + 1$ from ACS.
+We answer it in the affirmative.
+Fault tolerance is guaranteed if $n \ge 3t + 1$ by using ACS as a building block.
 While $t$ may be small compared to the population size $N$,
 we show that the probability for the system to fail is low even when $n \ge 3t + 1$ does not hold as long as the proportion of malicious nodes is not close to a third of $N$.
-Horizontal scalability is demonstrated both analytically and experimentally.
-With 1200 nodes in the network we were able to achieve nearly 5000 TX/s---an order of magnitude improvement over Bitcoin.
+For example, if there the population size is 1000 and 20\% of the nodes are malicious,
+the probability for a round to potentially to fail is bounded below $2.6 \times 10^{-16}$.
+This probability bound would decrease as the population size increases.
+Horizontal scalability property is demonstrated both analytically and experimentally.
+Unlike sharding protocols, the property holds regardless of transaction characteristics and needs no parameter selection.
 Finally, we achieve global consensus on transactions via consensus on checkpoint blocks.
@@
 This work is the first step in building a new paradigm for blockchain consensus protocol.
-It has the potential to efficiently cultivate trust on the internet in the presence of faults without any central authority.
+It has the potential to efficiently cultivate trust on the internet in the presence of faults without a central authority.
 We hope to improve our design by building a concrete application on top of it.
@@
 
-The fault tolerance property of our system is lower than that of typical blockchain systems,
-which usually aim for fewer than 50\%.
+The fault tolerance property is adaquate for permissioned systems where nodes have relatively more trust between each other.
+However, it is not adaquate for permissionless systems,
+which usually aim to tolerate fault that involve a minority of the network,
+which can be just shy of 50\%.
 For future work, we propose the following two ways to improve fault tolerance.

### 2017-08-04 thesis/chapters/ch1-intro.tex
Write abstract

@@
 This immediately show that those transactions are tamper-proof.
-We call our protocol ``Checkpoint Consensus''.
+We call our protocol CheckpointConsensys.
 % TODO mention consensus and trade off?
@@
 \begin{itemize}
-    \item We formally introduce a blockchain consensus protocol---Checkpoint Consensus.
+    \item We formally introduce a blockchain system---CheckpointConsensys.
         It uses individual hash chains and checkpoints on every node to achieve
         horizontally scalable for the first time.
-    \item We formally analyse Checkpoint Consensus to ensure correctness as defined in our architecture.
+    \item We formally analyse CheckpointConsensys to ensure correctness as defined in our architecture.
     \item We implement Checkpiont Consensus with experiment with up to 1200 nodes,

### 2017-08-04 thesis/chapters/ch3-model.tex
Write abstract

@@
 
-Our system---Checkpoint Consensus consist of three protocols---the consensus protocol, the transaction protocol and the validation protocol.
+Our system---CheckpointConsensys consist of three protocols---the consensus protocol, the transaction protocol and the validation protocol.
 It is based on our prior work on TrustChain~\cite{multichain},
@@
 \label{sec:system-overview}
-The Checkpoint Consensus architecture is visualised in~\Cref{fig:architecture}.
+The architecture is visualised in~\Cref{fig:architecture}.
 \begin{figure}[h]

### 2017-08-04 thesis/formalities/abstract.tex
New intro and etc.
Write synchroniser
Edit ch1 and ch2
Refactor validation protocol
Write abstract

@@
 \begin{tabular}{l}
-TODO AUTHOR\\
+Kelong Cong\\
 \\
@@
 \begin{tabular}{l}
-TODO TITLE\\
+Blockchain Consensus Protocol with Horizontal Scalability\\
 \\
@@
 % <MM> DD, YYYY (like \today)
-TODO GRADUATION DATE\\
+31st August 2017\\
 \\
@@
 \begin{tabular}{ll}
-TODO GRADUATION COMMITTEE & Delft University of Technology \\
 % The order of listing the names: Graduation prof, supervisor(s), others ordered by title + alphabetical
 %examples:
-%prof. dr. ir. H. J. Sips (chair) & Delft University of Technology \\
-%ir. dr. D. H. J. Epema           & Delft University of Technology \\
+Prof. dr. ir. D. H. J. Epema            & Delft University of Technology \\
+Dr. ir. J. A. Pouwelse                  & Delft University of Technology \\
+Dr. Z. Erkin                            & Delft University of Technology \\
 \end{tabular}
 
-\begin{abstract} %de abstract bevat alleen een korte samenvatting van de inhoud van het onderzoek
-TODO ABSTRACT
+\begin{abstract}
+% what's the problem?
+Blockchain systems have the potential to decentralise many traditionally centralised system.
+However, scalability remains a key challenge.
+Without a horizontally scalable solution, blockchain system remain unsuitable for ubiquitous use.
+% what's the solution?
+We design a novel blockchain system which we call CheckpointConsensys.
+Each node in our system maintain individual hash chains,
+which only stores transactions that involve the node.
+Consensus is reached on special blocks called checkpoint blocks rather than on all the transactions.
+Checkpoint blocks are effectively a hash pointer to the individual hash chains,
+thus a single checkpoint block may represent a large set of transaction blocks.
+The consensus protocol does not imply transaction validity.
+Hence we include a validation protocol which allows nodes to verify that the counterparty correctly recorded the transaction.
+% how good is the solution?
+We implement a CheckpointConsensys prototype and evaluate it experimentally.
+Our results show strong indication of horizontal scalability even in the worst case,
+where every transaction is with a randomly selected node.
+For 1200 nodes, we were able to perform almost 5000 transactions per second,
+orders of magnitude higher than the 7 transaction per second limit of Bitcoin.
 \end{abstract}

### 2017-08-05 thesis/chapters/ch1-intro.tex
Write preface and update Bitcoin situation

@@
 Its power consumption is as high as Republic of Ireland~\cite{o2014bitcoin}.
-Its market cap, at the time of writing, is over 40 billion USD~\cite{bitcoinmarketcap}.
-% TODO add graph
+Its market capitalisation, at the time of writing, is over 60 billion USD~\cite{bitcoinmarketcap} and growing (\Cref{fig:market-cap}).
+\begin{figure}[h]
+\centering
+\includegraphics[width=\textwidth]{market-cap}
+\caption{Market capitalisation and price of Bitcoin, graph is extracted from~\cite{bitcoinmarketcap}.}
+\label{fig:market-cap}
+\end{figure}
 Many online marketplaces are using Bitcoin, for example Steam\footnote{\url{https://store.steampowered.com/}} and even Amazon\footnote{Not directly, but via \url{https://purse.io/}.}.
@@
 A few months ago the backlog even reached 100,000, which meant new transactions would take at least 11 hours to be written to the blockchain~\cite{bitcoinbacklog}.
-% TODO update below
 This issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some calls it a civil war~\cite{bitcoincivilwar}.
@@
 It also requires a hard fork (not backward compatible), which risks consensus failure and devaluation.
-% TODO update the above
 A recent empirical study by Croman et el.~\cite{croman2016scaling} has shown that increasing the block size may help.
@@
 
-% TODO update
 What we have described is in fact the current state of Bitcoin.
@@
 The most prominant is Seggregated Witness or SegWit~\cite{segwit}.
-It moves the signature data to the end of the block and changes the block size to 4 MB.
+It moves the signature data to the end of the block and uses a variable block size up to 4 MB.
 The effect is that the first 1 MB of the block is still backward compatible, thus there is no need for a hard fork.
 It also solves the transaction malleability problem~\cite{bitcoinmalleability} and paves the way for offchain transactions via Lightning Network~\cite{lightningnetwork}.
-Although SegWit is implemented, it is still uncertain whether the Bitcoin network will adopt it.
-At the time of writing, there is only 35\% of the network signalling for adoptation,
-this is only a 10\% increase since November last year\footnote{\url{https://blockchain.info/charts/bip-9-segwit}}.
-Thus it may take years if ever for SegWit to be adopted.
-Even if it is adopted, it cannot reach the transaction rate of centralised systems.
-% https://www.bloomberg.com/news/articles/2017-07-10/bitcoin-risks-splintering-as-civil-war-enters-critical-month
+However, some fear SegWit is not enough and wish to see a hard fork that uses ``Adjustable Blocksize Cap'' that increase the block size up to 8 MB\footnote{\url{https://www.bitcoinabc.org/}}.
+The ongoing civil war resulted in a Bitcoin fork, called Bitcoin Cash\footnote{\url{https://www.bitcoincash.org/}}.
+At the time of writing, it is too early to speculate how these proposals and forks may affect the scalability properties of Bitcoin.
+But the conclusion that we can draw from the scalability debate is that the Bitcoin design is fundamentally not scalable,
+and it is highly non-trivial to ``patch'' the problem.
 
@@
 This immediately show that those transactions are tamper-proof.
-We call our protocol CheckpointConsensys.
+We call our protocol Checo, derived from \emph{che}ckpoint \emph{co}nsensus.
 % TODO mention consensus and trade off?
@@
 \begin{itemize}
-    \item We formally introduce a blockchain system---CheckpointConsensys.
+    \item We formally introduce a blockchain system---Checo.
         It uses individual hash chains and checkpoints on every node to achieve
         horizontally scalable for the first time.
-    \item We formally analyse CheckpointConsensys to ensure correctness as defined in our architecture.
+    \item We formally analyse Checo to ensure correctness as defined in our architecture.
     \item We implement Checkpiont Consensus with experiment with up to 1200 nodes,

### 2017-08-05 thesis/formalities/preface.tex
Write abstract
Write preface and update Bitcoin situation

@@
 \addcontentsline{toc}{chapter}{Preface}
-TODO MOTIVATION FOR RESEARCH TOPIC
+I mined my first Bitcoin in early 2013 with a GPU.
+Since then, I have been interested in decentralising the tradationally centralised systems.
+Following the Snowden revelations, the interest became a burning desire.
+Today, many blockchain systems are mature, but still not yet ready for ubiquitous use due to the scalability limitation.
+Many of us still do not use Bitcoin to purchase everyday goods, online or offline.
+This limitation as well as my personal ambitions led me to this thesis,
+with the goal of creating a truely horizontally scalable solution.
 
@@
 \noindent
-TODO ACKNOWLEDGEMENTS
+There are a number of people who have supported me in this endeavour who I would like to thank.
+My supervisor Dr. Johan Pouwelse, your unrivaled enthusiasm and guidance have kept me motivated and productive.
+Dr. Zhijie Ren, this work would not have been possible without your invalueable input and the numerous discussions which we have had.
+Members of the Tribler team---Martijn de Vos, Quinten Stokkink and Chengxin Ma,
+you have my gratitude for helping me with the implementation and proofreading my work.
+Members of the Cybersecurity group---Dr. Zekeriya Erkin and O\u{g}uzhan Ersoy,
+I thank both of your for your feedback on my presentations.
 
@@
 \noindent
-TODO AUTHOR
+Kelong Cong
 

### 2017-08-05 thesis/chapters/ch1-intro.tex
Grammar check ch1 and 2

@@
 Which, for the first time, gave us a simple banking system in the form of a distributed ledger, also known as blockchain.
-It needs no central control but still incorruptable with high probability even if there are malicious parties that aim to undermine the system.
+It needs no central control but still incorruptible with high probability even if there are malicious parties that aim to undermine the system.
 
-The primary innovations are (1) its consensus model which prevents double spending,
-and (2) its incentive mechanism that encourages anyone (with adaquate hardware) to participate in the network and keep it running.
+The primary innovations are (1) the consensus model which prevents double spending,
+and (2) the incentive mechanism that encourages anyone (with adequate hardware) to participate in the network and keep it running.
 The double spending problem can be seen as an inconsistency issue.
@@
 In Bitcoin's case, the consensus algorithm is called \emph{proof-of-work}.
-Where miners (parties that runs the Bitcoin network) collect transactions and compete in solving a puzzle.
+Where miners (parties that run the Bitcoin network) collect transactions and compete in solving a puzzle.
 The puzzle is easy to verify but difficult to solve.
@@
 \end{figure}
-Many online marketplaces are using Bitcoin, for example Steam\footnote{\url{https://store.steampowered.com/}} and even Amazon\footnote{Not directly, but via \url{https://purse.io/}.}.
-Due to its success, people from many different disciplines began investigating in way to use blockchain technology.
-This includes finance~\cite{finance}, heathcare~\cite{healthcare}, logistics~\cite{supplychain} and energy~\cite{energy}.
+Many online marketplaces are using Bitcoin.
+For example, Steam\footnote{\url{https://store.steampowered.com/}} and even Amazon\footnote{Not directly, but via \url{https://purse.io/}.}.
+Due to its success, people from many different disciplines began investigating in ways to use blockchain technology.
+This includes finance~\cite{finance}, health care~\cite{healthcare}, logistics~\cite{supplychain} and energy~\cite{energy}.
 
@@
 Bitcoin has the infamous 7 transactions per second upper bound~\cite{vukolic2015quest}.
-This is due to the fact that blocks are fixed to 1 MB are only generated on average every 10 minutes.
+This is due to the fact that blocks are fixed to 1 MB,
+and are only generated on average every 10 minutes.
 Since every Bitcoin transaction is at least 250 bytes, it computes to about 7 transactions per second.
@@
 Parties that are for the increase in block size argue that a larger block would improve the transaction rate.
-Parties against it argue that it would make mining more centralised because blocks take longer to propogate through the network.
-It also requires a hard fork (not backward compatible), which risks consensus failure and devaluation.
+Parties against it argue that it would make mining more centralised because blocks take longer to propagate through the network.
+It also requires a hard fork (not backwards compatible), which risks consensus failure and devaluation.
 A recent empirical study by Croman et el.~\cite{croman2016scaling} has shown that increasing the block size may help.
@@
 Even worse, it may give some miner an unfair lead over others.
-Thus fundamental changes are necessary run at the scale of centralised payment processors such as Visa,
+Thus fundamental changes are necessary to run at the scale of centralised payment processors such as Visa,
 which is in the order of tens of thousands transactions per second~\cite{visa}.
 
-What we have described is in fact the current state of Bitcoin.
+What we have described is, in fact, the current state of Bitcoin.
 Many proposals exist with the aim to make Bitcoin more scalable.
-The most prominant is Seggregated Witness or SegWit~\cite{segwit}.
+The most prominent one is Segregated Witness or SegWit~\cite{segwit}.
 It moves the signature data to the end of the block and uses a variable block size up to 4 MB.
-The effect is that the first 1 MB of the block is still backward compatible, thus there is no need for a hard fork.
-It also solves the transaction malleability problem~\cite{bitcoinmalleability} and paves the way for offchain transactions via Lightning Network~\cite{lightningnetwork}.
+The effect is that the first 1 MB of the block is still backwards compatible, thus there is no need for a hard fork.
+It also solves the transaction malleability problem~\cite{bitcoinmalleability} and paves the way for off-chain transactions via Lightning Network~\cite{lightningnetwork}.
 However, some fear SegWit is not enough and wish to see a hard fork that uses ``Adjustable Blocksize Cap'' that increase the block size up to 8 MB\footnote{\url{https://www.bitcoinabc.org/}}.
@@
 
-In this work, we take the advice of Croman et el.~\cite{croman2016scaling} and Marko~\cite{vukolic2015quest} and rethink the blockchain architecture.
-Our primary insight came from observing the differences between how transactional systems work in the real world and how they work in blockchain systems like Bitcoin. 
-Take a restaurant owner for example, most of the time the customer is honest and pays the bill.
+In this work, we take the advice of Croman et el.~\cite{croman2016scaling} and Vukoli\'{c}~\cite{vukolic2015quest} and rethink the blockchain architecture.
+Our primary insight came from observing the differences between how value transfer work in the real world and how they work in early blockchain systems.
+Take a restaurant owner, for example,
+most of the time the customer is honest and pays the bill.
 There is no need for the customer or the restaurant owner to report the transaction to any central authority 
@@
 which can be seen as a collective authority.
-This consequently lead to limited scalability because every transaction must be validated by the authority even when most of the transactions are legitimate.
+This consequently leads to limited scalability because every transaction must be validated by the authority even when most of the transactions are legitimate.
 
 Using the aforementioned insight,
-we explore an alternative consensus model for blockchain systems where transactions themself do not reach consensus,
+we explore an alternative consensus model for blockchain systems where transactions themselves do not reach consensus,
 but nevertheless verifiable at a later stage by any node in the network.
@@
 If a checkpoint of some node is in consensus, 
-then that node can proof to any other node that it holds a set of transactions that computes (form a chain) to the checkpoint.
-This immediately show that those transactions are tamper-proof.
+then that node can proof to any other node that it holds a set of transactions that compute (form a chain) to the checkpoint.
+This immediately shows that those transactions are tamper-proof.
 We call our protocol Checo, derived from \emph{che}ckpoint \emph{co}nsensus.
@@
     \item We formally analyse Checo to ensure correctness as defined in our architecture.
-    \item We implement Checkpiont Consensus with experiment with up to 1200 nodes,
+    \item We implement Checkpoint Consensus with experiment with up to 1200 nodes,
         our results show strong evidence of horizontal scalability.

### 2017-08-05 thesis/chapters/ch2-problem.tex
Grammar check ch1 and 2

@@
 Many attempts to fix the scalability problem exist in academia as well as the industry.
-In this chapter we describe the related work in the form of a mini taxonomy and identify areas for improvement and limitations.
+In this chapter, we describe the related work in the form of a mini taxonomy and identify areas for improvement and limitations.
 Then we state our research question which aims to overcome some of the limitations.
 
-\section{Related work and a mini taxonomy}
+\section{Related work and mini taxonomy}
 \label{sec:taxonomy}
@@
 Blockchain research has seen a surge in recent years from both the industry and academia,
-this resulted in a myriad of of blockchain systems with different approaches to scalability.
+this resulted in a myriad of blockchain systems with different approaches to scalability.
 To understand the current state-of-the-art,
@@
 This category represents the baseline,
-which are systems with a probabilistic probabilistic consensus algorithm.
+which are systems with a probabilistic consensus algorithm.
 That is, transactions never reach consensus with a probability of 1.
@@
 The advantage of this approach is that it can be used in a large network.
-In fact, prior to Bitcoin there were no internet-scale consensus protocol.
-Bitcoin is also fault tolerant,
-because attackers can not out pace honest users in finding new blocks unless they have a majority of the hash power.
+In fact, prior to Bitcoin, there was no internet-scale consensus protocol.
+Bitcoin is also fault-tolerant because attackers can not out pace honest users in finding new blocks unless they have a majority of the hash power.
 As long as the majority is honest and reasonable measures are taken to prevent the selfish mining attack~\cite{eyal2014majority}.
-The disadvantage however is that transactions are never in consensus with a probability of 1---no consensus finality.
+The disadvantage, however, is that transactions are never in consensus with a probability of 1---no consensus finality.
 Hence double spending is possible for transactions that do not have a large number of confirmations.
@@
 Off-chain transaction systems are implemented using multi-signature addresses~\cite{bitcoinmultisig} and hashed time-locked Bitcoin contracts~\cite{bitcointimelock}.
-In the simiplest case, 
+In the simplest case, 
 if two parties wish to make transactions,
-they open a payment channel in the form of of a multi-signature Bitcoin address (for two parties it would be a 2-of-2 signature address).
+they open a payment channel in the form of a multi-signature Bitcoin address (for two parties it would be a 2-of-2 signature address).
 Each of the party must also deposit some Bitcoins in the multi-signature address,
@@
 If this issue occurs, the counterparty can drain all the Bitcoins in the channel.
-To close the channel, the parties simply boradcast the latest net settlement to the Bitcoin network,
+To close the channel, the parties simply broadcast the latest net settlement to the Bitcoin network,
 this is called the closing transaction.
@@
 
-The advantages of such systems is that they act as add-ons to Bitcoin wish already has a large number of users.
+The advantage of such systems is that they act as add-ons to Bitcoin wish already has a large number of users.
 Thus, if enough of the network wish for it (by setting a new block version),
 then a large number of users will instantly benefit from it.
-Some implementation already exist\footnote{For example \url{https://github.com/lightningnetwork/lnd}.},
-but at the time of writing Segwit is not activated yet which is a prerequisite for Lightning Network.
+Some implementations already exist\footnote{For example \url{https://github.com/lightningnetwork/lnd}.},
+but at the time of writing SegWit is not activated yet which is a prerequisite for Lightning Network.
 
-On the other hand, off-chain transactions also suffers from the problem of Bitcoin.
+On the other hand, off-chain transactions also suffer from the problem of Bitcoin.
 Proof-of-work still consumes an unreasonable amount of power.
-Further, payment channels complicates user experience.
+Further, payment channels complicate user experience.
 As we mentioned earlier, each node must deposit some Bitcoins into a multi-signature account.
-The users must pick a suitable amount;
-if the deposit is too low it would not allow large transactions,
-if the deposit is too high the user is locked out of much of its Bitcoins for use outside of the channel.
+The users must pick a suitable amount.
+If the deposit is too low it would not allow large transactions.
+If the deposit is too high the user is locked out of much of its Bitcoins for use outside of the channel.
 In addition, the user must proactively check whether the counterparty has broadcasted an old channel state so that the user does not lose Bitcoins.
-Payment channel in theory solves the scalability problem of early blockchain systems,
-but to the best of our knowledge its exact scalability characteristics are not investigated.
+Payment channel, in theory, solves the scalability problem of early blockchain systems,
+but to the best of our knowledge, its exact scalability characteristics are not investigated.
 
@@
 
-This category of systems use traditional Byzantine consensus algorithms such as PBFT~\cite{castro1999practical}.
+This category of systems uses traditional Byzantine consensus algorithms such as PBFT~\cite{castro1999practical}.
 In essence, they contain a fixed set of nodes (sometimes called validating peer) that run a Byzantine consensus algorithm to decide on new blocks.
-This is often used in permissioned system where the validators must be pre-determined,
-for example Hyperledger Fabric~\cite{cachin2016architecture}.
+This is often used in permissioned system where the validators must be predetermined,
+for example, Hyperledger Fabric~\cite{cachin2016architecture}.
 
 A nice aspect of Byzantine consensus and in particular PBFT is that it can handle much more transactions than classical blockchain systems.
-PBFT can for example achieve 10,000 TX/s if the number of validating peer is under 16~\cite[Section 5.2]{miller2016honey}.
+PBFT can, for example, achieve 10,000 TX/s if the number of validating peer is under 16~\cite[Section 5.2]{miller2016honey}.
 Further, in contrast to proof-of-work, PBFT consensus is final.
@@
 Going back to PBFT, its transaction rate drops to under 5000 TX/s when the number of validating peer is 64~\cite[Section 5.2]{miller2016honey}.
-Moreover, the validating peers are pre-determined which makes the system unsuitable for the open internet.
+Moreover, the validating peers are predetermined which makes the system unsuitable for the open internet.
 
@@
 \subsection{Combining proof-of-work with Byzantine consensus}
-Recent research has developed a class of hybrid systems which uses PoW for committe election,
+Recent research has developed a class of hybrid systems which uses PoW for committee election,
 and Byzantine consensus algorithms to agree on transactions.
@@
 This technique overcomes the early blockchain scalability issue by delegating the transaction validation to the Byzantine consensus protocol (e.g. PBFT in ByzCoin~\cite{kogias2016enhancing}).
-A major tradeoff of such systems is that they cannot guarantee correctness when there is a large number of malicious nodes (but less than majority).
+A major trade-off of such systems is that they cannot guarantee correctness when there is a large number of malicious nodes (but less than a majority).
 For SCP, ByzCoin and Solidus, they all have some probability to elect more than $t$ Byzantine nodes into the committee,
@@
 This problem is especially difficult to solve because the committee is always much smaller than the population size which has more than $t$ Byzantine nodes.
-Classical blockchain do not have this problem because they do not use Byzantine consensus.
-Further, due to the fact that these system must reach consensus on all transactions, none of them achieves horizontal scalability.
+Classical blockchain does not have this problem because they do not use Byzantine consensus.
+Further, due to the fact that these systems must reach consensus on all transactions, none of them achieves horizontal scalability.
 
@@
 An intra-shard protocol is needed for transactions that involve nodes from more than one shard.
-The number of shard grow linears with respect to the total computational power in the network.
+The number of shards grows linearly with respect to the total computational power in the network.
 This scheme achieves horizontal scalability because if every shard commits transactions at the same throughput,
@@
 OmniLedger has an inter-shard transaction protocol but choosing a good shard size is difficult.
-A large shard size would make the system less scalable, because the Byzantine consensus algorithm must be run by a large number of nodes.
+A large shard size would make the system less scalable because the Byzantine consensus algorithm must be run by a large number of nodes.
 A small shard size would result in a large number of inter-shard transactions which also hinder scalability.
@@
 Furthermore, since no shards can be compromised for the system to function correctly,
-the adversaries have more opportunites to compromise the system.
+the adversaries have more opportunities to compromise the system.
 
@@
 
-On the other hand, the application and security is limited.
+On the other hand, the application and security are limited.
 A malicious node can easily lie to other nodes regarding its own chain by simply presenting one version of reality to a set of nodes and another version to a different set.
@@
 \begin{displayquote}
-\emph{Is it possible to design a blockchain consensus protocol that is fault tolerant, scalable and can reach global consensus?}
+\emph{Is it possible to design a blockchain consensus protocol that is fault-tolerant, scalable and can reach global consensus?}
 \end{displayquote}
@@
 Underneath these applications lies the blockchain consensus protocol which has the goal of reaching consensus on transactions or state changes.
-In the case of Bitcoin it is proof-of-work (PoW).
+In the case of Bitcoin, it is proof-of-work (PoW).
 It is not concerned with the structure of the transactions,
@@
 In this work, we focus on the blockchain consensus protocol rather than any specific application.
-Because it is the scalability bottleneck and is neccessary for any blockchain based application.
+Because it is the scalability bottleneck and is necessary for any blockchain based application.
 % Also, it should be possible to build any kind of application on top our system.
@@
 
-In the remainder of this chapter we expand on our research question and visit each of our requirement in detail.
-At the end of this chapter we describe some issues that we do not consider as a part of our design,
+In the remainder of this section, we expand on our research question and visit each of our requirement in detail.
+At the end of this section, we describe some issues that we do not consider as a part of our design,
 which is the result of not focusing on any specific application.
@@
 We emphasise the word state because in contrary to traditional blockchains,
-it does not neccessarily mean the set of all transactions that has ever happened.
+it does not necessarily mean the set of all transactions that have ever happened.
 The state can be a verifiable representation of all the transactions.
 As an extreme example, a single digest of the complete history can represent the state.
-It is even verifiable as long as there are nodes that stores the history.
+It is even verifiable as long as there are nodes that store the history.
 
@@
 
-\subsection{Security and fault tolerance}
+\subsection{Security and fault-tolerance}
 
-The second objective is fault tolerance,
-where our system should be unaffected in the presence of powerful adversaries.
+Our system should be unaffected in the presence of powerful adversaries.
 In particular, adversaries are Byzantine meaning that they can have arbitrary behaviour.
-Thus anything is possible from simply ommitting messages to colluding with eachother in order to undermine the whole system.
-This asppect usually comes for free when using a Byzantine fault tolerant consensus algorithm.
+Thus anything is possible from simply omitting messages to colluding with each other in order to undermine the whole system.
+This aspect usually comes for free when using a Byzantine fault-tolerant consensus algorithm.
 
@@
 For example, if two different honest nodes attempts verify the same transaction,
-it cannot be the case that one nodes thinks that the transaction is valid but the other thinks it is invalid.
+it cannot be the case that one node thinks that the transaction is valid but the other thinks it is invalid.
 
@@
 
-Early blockchain systems may run in a network of thousands of nodes, but the performance do not scale.
-On the other hand, as we will see in~\Cref{sec:acs-background}, Byzantine consensus algorithms performs well but only when in a small network.
+Early blockchain systems may run in a network of thousands of nodes, but the performance does not scale.
+On the other hand, as we will see in~\Cref{sec:acs-background}, Byzantine consensus algorithms perform well but only when in a small network.
 In this work, our goal is to overcome the limitations of early blockchain systems and Byzantine consensus algorithms to create a horizontally scalable blockchain.
@@
 However, the price to pay is that the Sybil attack~\cite{douceur2002sybil} prevention mechanism moves from the blockchain consensus protocol to the application layer.
-For example, social network based Sybil defence mechanims use graph structure of real-world relationships~\cite{yu2006sybilguard}.
+For example, social network based Sybil defence mechanisms use graph structure of real-world relationships~\cite{yu2006sybilguard}.
 Online marketplaces such as Amazon use the rating of buyer and sellers.
@@
 For the same reason, we do not explicitly consider spam or denial of service attacks.
-Without a concrete application it is difficult to distinguish a super active user from an attacker.
+Without a concrete application, it is difficult to distinguish a super active user from an attacker.
 

### 2017-08-05 thesis/formalities/preface.tex
Grammar check ch1 and 2

@@
 I mined my first Bitcoin in early 2013 with a GPU.
-Since then, I have been interested in decentralising the tradationally centralised systems.
+Since then, I have been interested in decentralising the traditionally centralised systems.
 Following the Snowden revelations, the interest became a burning desire.
@@
 Many of us still do not use Bitcoin to purchase everyday goods, online or offline.
-This limitation as well as my personal ambitions led me to this thesis,
-with the goal of creating a truely horizontally scalable solution.
+This limitation, as well as my personal ambitions, led me to this thesis,
+with the goal of creating a truly horizontally scalable solution.
 
@@
 There are a number of people who have supported me in this endeavour who I would like to thank.
-My supervisor Dr. Johan Pouwelse, your unrivaled enthusiasm and guidance have kept me motivated and productive.
-Dr. Zhijie Ren, this work would not have been possible without your invalueable input and the numerous discussions which we have had.
+My supervisor Dr. Johan Pouwelse, your unrivalled enthusiasm and guidance have kept me motivated and productive.
+Dr. Zhijie Ren, this work would not have been possible without your invaluable input and the numerous discussions which we have had.
 Members of the Tribler team---Martijn de Vos, Quinten Stokkink and Chengxin Ma,
 you have my gratitude for helping me with the implementation and proofreading my work.
-Members of the Cybersecurity group---Dr. Zekeriya Erkin and O\u{g}uzhan Ersoy,
-I thank both of your for your feedback on my presentations.
+Members of the Cyber Security group---Dr. Zekeriya Erkin and O\u{g}uzhan Ersoy,
+I thank both of you for your feedback on my presentations.
 

### 2017-08-06 thesis/chapters/ch2-problem.tex
Spelling/grammar check for ch3, 4, 5, 6

@@
 This technique overcomes the early blockchain scalability issue by delegating the transaction validation to the Byzantine consensus protocol (e.g. PBFT in ByzCoin~\cite{kogias2016enhancing}).
-A major trade-off of such systems is that they cannot guarantee correctness when there is a large number of malicious nodes (but less than a majority).
+A major tradeoff of such systems is that they cannot guarantee correctness when there is a large number of malicious nodes (but less than a majority).
 For SCP, ByzCoin and Solidus, they all have some probability to elect more than $t$ Byzantine nodes into the committee,

### 2017-08-06 thesis/chapters/ch3-model.tex
Write preface and update Bitcoin situation
Spelling/grammar check for ch3, 4, 5, 6

@@
 
-Our system---CheckpointConsensys consist of three protocols---the consensus protocol, the transaction protocol and the validation protocol.
+Our system consists of three protocols---the consensus protocol, the transaction protocol and the validation protocol.
 It is based on our prior work on TrustChain~\cite{multichain},
-where every node independently interact with each other via their own blockchain\footnote{It is originally called MultiChain but the name has changed to TrustChain}.
+where every node independently interacts with each other via their own blockchain\footnote{It is originally called MultiChain but the name has changed to TrustChain}.
 
@@
 \end{figure}
-It consist of one data structure---Extended TrustChain, and three protocols.
+It consists of one data structure---Extended TrustChain, and three protocols.
 The protocols have distinct roles which are evident from their names.
 They run concurrently and do not interact with each other.
-The only synchronisation happens on the Extended TrustChain layer, where all the protocols read and write from it.
+The only synchronization happens on the Extended TrustChain layer, where all the protocols read and write from it.
 We first describe each component individually, starting with Extended TrustChain, and then explain how they fit together.
@@
     The blocks at the ends of one dotted line are pairs of each other.
-    The red block after $t_{b, 5}$ indicate a fork.}
+    The red block after $t_{b, 5}$ indicates a fork.}
     \label{fig:trustchain-bad}
@@
 In \Cref{fig:trustchain-bad}, node $b$ (in the middle chain) created two TX blocks that both point to $t_{b, 5}$.
-If this is a ledger system it can be seen as a double spend, where the currency accumulated up until $t_{b, 5}$ are spent twice.
+If this is a ledger system it can be seen as a double spend, where the currency accumulated up until $t_{b, 5}$ is spent twice.
 
@@
 The primary difference is the introduction of a new type of block---checkpoint (CP) block.
-In contract to TX blocks, CP blocks do not store transactions or contain references.
+In contrast to TX blocks, CP blocks do not store transactions or contain references.
 Their purpose is to capture the state of the chain and the state of the whole system.
@@
 
-From this point onwards, we use TrustChain to mean the Extended TrustChain unless explicitly clarified.
+From this point on, we use TrustChain to mean the Extended TrustChain unless explicitly clarified.
 
@@
 % The consensus protocol uses an existing Byzantine consensus algorithm (explained in detail in \Cref{sec:xyz}) and runs it continuously in rounds.
-The consensus protocol can be seen as a technique of running infinitly many rounds of some 
-Byzantine consensus algorithm\footnote{More accuratly it is ACS or asynchronous subset consensus, we describe ACS in \Cref{sec:acs-background}.},
+The consensus protocol can be seen as a technique of running infinitely many rounds of some 
+Byzantine consensus algorithm\footnote{More accurately it is ACS or asynchronous subset consensus, we describe ACS in \Cref{sec:acs-background}.},
 starting a new execution immediately after the previous one is completed.
@@
 
-A visual explaination can be found in \Cref{app:consensus-example},
+A visual explanation can be found in \Cref{app:consensus-example},
 it walks through the steps needed for a node to be selected as a facilitator using an example.
@@
 
-The consensus and transaction protocol by themselves do not provide a mechanism to detect forks or other forms of tamperaing.
+The consensus and transaction protocol by themselves do not provide a mechanism to detect forks or other forms of tampering.
 Thus we need a validation protocol to counteract malicious behaviour.
-When a node wish to validate one of its TX, it asks the counterparty for the \emph{agreed fragment} of the TX.
-An agreed fragment of a TX is a section of the chain beginning and ending with CP blocks that contains the TX and are in consensus.
+When a node wishes to validate one of its TX, it asks the counterparty for the \emph{agreed fragment} of the TX.
+An agreed fragment of a TX block is a section of the chain beginning and ending with CP blocks that contain the TX block,
+the CP blocks must also be in consensus.
 Upon the counterparty's response, the node checks that the CP blocks are indeed in consensus,
@@
 
-Our subprotocol design gives us the highly desireable non-blocking property.
+Our sub-protocol design gives us the highly desirable non-blocking property.
 In particular, we do not need to ``freeze'' the state of the chain for some communication to complete in order to create a block.
@@
 \end{itemize}
-Further, we use $a || b$ to denote concatination of the binary representations of $a$ and $b$.
+Further, we use $a || b$ to denote concatenation of the binary representations of $a$ and $b$.
 
@@
 Thus in no stage of our protocol do we make timing assumptions.
-The adversary has fully control of the delivery schedule and the message ordering of all messages.
+The adversary has full control of the delivery schedule and the message ordering of all messages.
 But they are not allowed to drop messages except for their own\footnote{
@@
 
-We assume there exist a Public Key Infrastructure (PKI), and nodes are identified by their unique and permanent public key.
+We assume there exists a Public Key Infrastructure (PKI), and nodes are identified by their unique and permanent public key.
 Finally, we use the random oracle (RO) model, i.e. calls to the random oracle are denoted by $\textsf{H}: \{0, 1\}^* \rightarrow \{0, 1\}^\lambda$,
@@
 % TODO big N does not follow our convention
-In our model we consider $N$ nodes, which is the population size.
+In our model, we consider $N$ nodes, which is the population size.
 $n$ of them are facilitators, $t$ out of $n$ are malicious and the inequality
@@
 The first restriction is that corrupted node can only change across rounds.
-That is, if a round has started, the corrupted nodes cannot changed until the next round.
+That is, if a round has started, the corrupted nodes cannot change until the next round.
 The second restriction is that the adversary, presumably controlling all the corrupted nodes, is forgetful.
-Namely the adversary may learn the internal state such as the private key of a corrupted node,
+Namely, the adversary may learn the internal state such as the private key of a corrupted node,
 but if the node recovers, then the adversary must forget the private key.
@@
 The notation $b_{u, i}$ is generic over the block type.
-We assume there exist a function $\textsf{typeof}: B_u \rightarrow \{ \tau, \gamma \}$ that returns the type of the block,
+We assume there exists a function $\textsf{typeof}: B_u \rightarrow \{ \tau, \gamma \}$ that returns the type of the block,
 where $\tau$ represents the TX type and $\gamma$ represents the CP type.
@@
 \item $m$ is the transaction message.
-\item $sig_u$ is the signature created using $sk_u$ on the concatination of the binary representation of the five items above.
+\item $sig_u$ is the signature created using $sk_u$ on the concatenation of the binary representation of the five items above.
 \end{enumerate}
@@
 
-Here we define an important properties which results from the interleaving nature of CP and TX blocks.
+Here we define an important property which results from the interleaving nature of CP and TX blocks.
 It is used in our validation protocol.
@@
     compute correct consensus result in every round
-    and have unbiased election of facilitators.
+    and have an unbiased election of facilitators.
 Concretely, we define the necessary properties as follows.
@@
 Node $i$ starts with an input value $v_i$.
-The nodes must decide for one of  those values, satisfying the the following.
+The nodes must decide for one of  those values, satisfying the following.
 \begin{enumerate}
@@
 A node under Byzantine fault means that it can have arbitrary behaviour.
-For example not sending message or colluding with other Byzantine nodes to undermine the entire system.
+For example not sending a message or colluding with other Byzantine nodes to undermine the entire system.
 Note that the decision is on a single value.
@@
 There are $n$ nodes, of which at most $t$ might experience Byzantine fault.
-Node $i$ starts with an non-empty set of input values $C_i$.
+Node $i$ starts with a non-empty set of input values $C_i$.
 The nodes must decide an output $C$, satisfying the following.
@@
 
-ACS has the nice property of censorship resiliance when compared to other consensus algorithms.
+ACS has the nice property of censorship resilience when compared to other consensus algorithms.
 For instance, Hyperledger and Tendermint uses a variant of Practical Byzantine Fault Tolerance (PBFT)~\cite{castro1999practical} as their consensus algorithm.
-In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectivly filter transactions.
-In contract, every party in ACS are involved in the proposal phase,
+In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectively filter transactions.
+In contrast, every party in ACS are involved in the proposal phase,
 and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
 Thus, if some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
-For a detailed description of ACS we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.
+For a detailed description of ACS, we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.
 To understand the remainder of this work, the knowledge of~\Cref{def:acs} is sufficient.
@@
 The code satisfies all the properties in \Cref{def:consensus}.
-Namely every node has the same set of valid facilitators $\F_1$ that are randomly choosen.
+Namely, every node has the same set of valid facilitators $\F_1$ that are randomly chosen.
 This concludes the bootstrap phase.
@@
 % Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D \gg \Delta$.
-Invalid messages are removed, namely blocks with invalid signatures and blocks signed by the same key.
+Invalid messages are removed.
+That is blocks with invalid signatures and blocks signed by the same key.
 With the sufficient number of \texttt{cp\_msg} messages,
@@
 
-At the core of the consensus phase is the ACS protocol.
+The core of the consensus phase is the ACS protocol.
 While any ACS protocol that satisfies the standard definition will work,
 we use a simplification of HoneyBadgerBFT~\cite{miller2016honey} as our ACS protocol
-because it is a consensus algorithms designed for blockchain systems.
+because it is a consensus algorithm designed for blockchain systems.
 We do not use the full HoneyBadgerBFT due to the following.
@@
 Second, HoneyBadgerBFT uses threshold encryption to hide the content of the transactions.
-But we do not reach consensus on transactions, only CP blocks, so hiding CP blocks is meaningless for us as it contains no transactional information.
+But we do not reach consensus on transactions, only CP blocks, so hiding CP blocks is meaningless for us as it contains no transaction information.
 
@@
 
-Upon receiving $\C_{r + 1}$ and at least $n - f$ valid signatures by some node $u$, $u$ performs two asks.
+Upon receiving $\C_{r + 1}$ and at least $n - f$ valid signatures by some node $u$, $u$ performs two tasks.
 First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1}, r + 1)$ using \Cref{alg:new-cp}.
@@
 
-Our protocol has some similarities with synchronisers~\cite[Chapter 10]{podc} because it is effectively a technique to introduce synchrony in an asynchronous environment.
-If we consider the facilitators as a collective authority, then it can be seen as a synchroniser that sends pulse messages (in the form of \texttt{cons\_msg} and \texttt{cons\_sig}) to indicate the start of a new clock pulse.
+Our protocol has some similarities with synchronizers~\cite[Chapter 10]{podc} because it is effectively a technique to introduce synchrony in an asynchronous environment.
+If we consider the facilitators as a collective authority, then it can be seen as a synchronizer that sends pulse messages (in the form of \texttt{cons\_msg} and \texttt{cons\_sig}) to indicate the start of a new clock pulse.
 Every node then sends a completion messages (in the form of \texttt{cp\_msg}) to the new collective authority to indicate that they are ready for the next pulse.
@@
 It is also known as True Halves, first described by Veldhuisen~\cite[Chapter~3.2]{truehalves}.
-Node that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
+Nodes that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
 $txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$.
@@
 It is defined this way because the malicious sender may create new TX blocks in their own chain but never send \texttt{tx\_req} messages.
-In that case, it may seem that the counterparty, who is honest, purposefully ommitted TX blocks.
+In that case, it may seem that the counterparty, who is honest, purposefully omitted TX blocks.
 But in reality, it was the malicious sender who did not follow the protocol.
-Thus, in such cases the whole transaction, identified by its $txid$ is marked as invalid.
+Thus, in such cases, the whole transaction identified by its $txid$ is marked as invalid.
 
 Further, the caller of $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ is not necessarily 
-$u$\footnote{In practice it often is because after completing the TX protocol the parties are incentivised to check that the counterparty ``did the right thing''.}
+$u$\footnote{In practice, it often is because after completing the TX protocol the parties are incentivised to check that the counterparty ``did the right thing''.}
 Any node $w$ may call $\textsf{get\_validity}(t_u{u, i}, F_{u, i}, F_{v, j})$ as long as the caller $w$ has an agreed fragment of $t_{u, i}$---$F_{u, i}$.
@@
     \item \emph{Liveness}:
-        Any valid (invalid) transactions is marked as valid (invalid) eventually.
+        Any valid (invalid) transaction is marked as valid (invalid) eventually.
 \end{enumerate}
@@
 Our validity requirement only requires that there is only one $t_{v, j}$ in the correct consensus round.
-The counterparty may create any number of fake pairs in a later consensus rounds.
+The counterparty may create any number of fake pairs in later consensus rounds.
 But these fake pairs only pollutes the chain of $v$ and can never be validated because the round is incorrect.
@@
 
-A down side of epidemic protocol is that it usually takes $O(\log N)$ time to infect the whole network,
+A down side of an epidemic protocol is that it usually takes $O(\log N)$ time to infect the whole network,
 whereas broadcasting uses constant time.
 Another downside of some epidemic protocols (e.g. eager push) is that it is difficult to guarantee delivery.
-It is especially true when the parameters are not choosen correctly in a network that is only partially connected (but every node is nevertheless reachable).
+It is especially true when the parameters are not chosen correctly in a network that is only partially connected (but every node is nevertheless reachable).
 If the delivery cannot be guaranteed, 
@@
 Our model is purely asynchronous, where we make no timing assumptions anywhere in the protocol.
-In many applications however, it is often fine to make timing assumptions.
+However, in many applications it is often fine to make timing assumptions.
 For example, TCP relies on timeout for its retransmission and the \texttt{nLockTime} property in Bitcoin transactions makes the transaction unspendable until some time in the future (either Unix time or block height)~\cite{bitcoindevguide}.
@@
 It simply sends its latest CP block to the facilitators.
-Then, in the next round the node will have a chance to become a facilitator just like any existing node.
+Then, in the next round, the node will have a chance to become a facilitator just like any existing node.
 To leave the network, nodes simply stop submitting CP blocks.
 There is a subtlety here which happens when the node is elected as a facilitator in the following round.
-In this case, the node must fulfill its obligation by completing the consensus protocol, but without proposing its own CP block, before leaving.
-Otherwise the $n \ge 3t + 1$ condition may be violated.
+In this case, the node must fulfil its obligation by completing the consensus protocol, but without proposing its own CP block, before leaving.
+Otherwise, the $n \ge 3t + 1$ condition may be violated.
 
@@
 
-The space saving of this approach of course depends on the size of the full blocks.
+The space saving of this approach depends on the size of the full blocks.
 If the full blocks are on average 500 bytes
@@
 and the compact blocks are $32 + 32 + 8 = 72$ bytes
-(SHA256 digests are 32 bytes each and we use a 64 bit integer to represent the sequence number),
+(SHA256 digests are 32 bytes each and we use a 64-bit integer to represent the sequence number),
 then the saving in communication cost is 86\%.
@@
 \label{sec:caching}
-One more way to improve the efficiency of the validation protocol is to use a single agreed fragment to validate multiple transaction.
+One more way to improve the efficiency of the validation protocol is to use a single agreed fragment to validate multiple transactions.
 Concretely, upon receiving an agreed fragment from node $A$,
 rather than validating a single transaction,
-we attempt to validate all transactions in the unknown state performed with $A$ in that fragment.
+we attempt to validate all transactions with $A$ in the unknown state in that fragment.
 
@@
 The validation algorithm guarantees that there are no forks within a single agreed fragment.
-This is sufficient for some applications such as proving the existance of some information.
+This is sufficient for some applications such as proving the existence of some information.
 However, for applications such as digital cash where every block depends on one or more previous blocks,
 our scheme is not suitable.
-For such applications we need to guarantee that there are no forks from the genesis block leading up to the TX block of interest.
+For such applications, we need to guarantee that there are no forks from the genesis block leading up to the TX block of interest.
 
@@
 It may sound inefficient at first, but nodes employ caching to minimise communication costs.
-In our work we call this effect ``spontaneous sharding''.
+In our work, we call this effect ``spontaneous sharding''.
 
 The second approach is probabilistic but with only a constant communication overhead.
-For a node, observe that if all of its agreed fragment has a transaction with a honest node,
-then the complete chain is effectively validated in a distributed maner.
+For a node, observe that if all of its agreed fragment has a transaction with an honest node,
+then the complete chain is effectively validated in a distributed manner.
 The only way for an attacker to make a fork is to make sure that the agreed fragment containing the fork has no transactions with honest nodes.
-Such malicious behaviour is prevented probabilisticaly using a challenge-response protocol as follows.
+Such malicious behaviour is prevented probabilistically using a challenge-response protocol as follows.
 Suppose node $A$ wish to make a transaction with node $B$.
-$A$ first sends a challenge to $B$ asking it to proof that it holds a valid agreed fragment between some consensus round specified by $A$.
+$A$ first sends a challenge to $B$ asking it to prove that it holds a valid agreed fragment between some consensus round specified by $A$.
 If $B$ provides a correct proof, then they run the transaction protocol as usual.
@@
 
-% \subsubsection{Global clock for synchronisation}
+% \subsubsection{Global clock for synchronization}
 % Where there is no Byzantine corruption, we conjecture that our protocol runs in the asynchronous setting.

### 2017-08-06 thesis/chapters/ch4-analysis.tex
Spelling/grammar check for ch3, 4, 5, 6

@@
 
-Up to this point we described our system specification in detail.
+Up to this point, we described our system specification in detail.
 Of course, specification alone does not establish any truths.
 In this chapter, we evaluate our system analytically.
-First we show it has the desired properties.
+First, we show it has the desired properties.
 That is, the properties of the consensus protocol and the validation protocol (\Cref{def:consensus} and \Cref{def:validation} respectively) should hold.
 Then we analyse the performance, especially the throughput,
-and show that it out performs classical blockchain systems.
-Finally, we consider the case where the $n \ge 3t + 1$ assumption is violated,
-and show the effect of it in our system.
+and show that it out performs early blockchain systems.
+Finally, we consider the case where the $n \ge 3t + 1$ assumption is violated and show the effect of it in our system.
 
-\section{Correctness in the presense of faults}
+\section{Correctness in the presence of faults}
 Our first objective is to show that \Cref{def:consensus} holds for our consensus protocol.
@@
 but this eventually happens because there are $N - t$ correct nodes and all correct facilitator eventually receives $N -t$ valid CP blocks.
-At the end of ACS, some $\C_{r+1}$ is created, and is broadcasted along with the signature of the facilitators.
+At the end of ACS, $\C_{r+1}$ is created and then broadcasted along with the signature of the facilitators.
 Due to the agreement property of ACS (\Cref{def:acs}),
@@
 if $\F_r$ is known by all correct nodes then
-every node with a CP block in $\C_{r+1}$, should have an equal probability to be elected as a facilitator in $\F_{r+1}$.
+every node with a CP block in $\C_{r+1}$ should have an equal probability to be elected as a facilitator in $\F_{r+1}$.
 \end{lemma}
@@
 and then ACS eventually terminates.
-Finally, the results are eventually dissemminated to all the nodes.
+Finally, the results are eventually disseminated to all the nodes.
 \end{proof}
@@
 From Lemmas \ref{lemma:agreement}, \ref{lemma:validity}, \ref{lemma:fairness} and \ref{lemma:termination},
-we have shown that the 4 properties of \Cref{def:consensus} holds when assuming the existance of some $\F_r$.
+we have shown that the 4 properties of \Cref{def:consensus} holds when assuming the existence of some $\F_r$.
 Thus, to proof the whole of \Cref{def:consensus},
@@
 The consensus protocol (on CP blocks and facilitators) is the backbone for consensus on transactions.
-In this section we build on top of \Cref{theorem:consensus} to show that the properties except liveness in \Cref{def:validation} can be satisfied.
+In this section, we build on top of \Cref{theorem:consensus} to show that the properties except for liveness in \Cref{def:validation} can be satisfied.
 
@@
 node $u$ decides \emph{valid} but node $v$ decides \emph{invalid}.
-Then there exist a fragment $F' = \{ \dots, t', c'\}$ which $u$ received that contains a valid pair of $t$---$t'$.
-There also exist a fragment $F'' = \{ \dots, t'', c''\}$ which $v$ received that does not contain or contains an invalid pair---$t''$.
+Then there exists a fragment $F' = \{ \dots, t', c'\}$ which $u$ received that contains a valid pair of $t$---$t'$.
+There also exists a fragment $F'' = \{ \dots, t'', c''\}$ which $v$ received that does not contain or contains an invalid pair---$t''$.
 In both cases, the $\textsf{get\_validity}(\cdot)$ function must have reached \Cref{line:valid-fragment}.
@@
 One of the main advantages of our scheme over running a consensus algorithm on all the transactions is that 
-the rate of transaction is no longer dependent on the consensus algorithm---ACS.
+the transaction throughput is no longer dependent on the consensus algorithm---ACS.
 This enables horizontal scalability where adding new nodes would lead to higher global transaction rate.
@@
 Now we show a negative result, where the liveness property of \Cref{def:validation} cannot be attained.
-Meaning that trasanctions with adversaries cannot always be validated.
+Meaning that transactions with adversaries cannot always be validated.
 
 \begin{lemma}
-There exist a valid transactions that cannot be validated eventually.
+There exists a valid transaction that cannot be validated eventually.
 \end{lemma}
 \begin{proof}
-Suppose nodes $u$ and $v$ correctly perfomed the TX protocol which resulted a transaction.
+Suppose nodes $u$ and $v$ correctly performed the TX protocol to create a transaction.
 Then when $u$ wants to validate it, it does so by sending \texttt{vd\_req} message to $v$.
@@
 If the adversary is observed to ignore validation requests, then the honest nodes may prefer not to transact with her in the future.
-Thus, to stay relevant in the system, the adversary need to comply to the protocol.
+Thus, to stay relevant in the system, the adversary needs to comply with the protocol.
 
@@
 \label{sec:cons-complexity}
-The consensus protocol can bee seen as three parts,
+The consensus protocol can be seen as three parts,
 so the communication complexity is the sum of these parts.
@@
 The communication complexity of ACS is $O(n^2|v| + \lambda n^3 \log n)$~\cite{miller2016honey},
-where $|v|$ is the size of largest message and $\lambda$ is the security parameter (described in \Cref{sec:model-assumptions}).
+where $|v|$ is the size of the largest message and $\lambda$ is the security parameter (described in \Cref{sec:model-assumptions}).
 In our system, we wish to understand the scalability property.
@@
 For the same reason as the first part, this is also $O(N)$.
-Thus the combined communication complexity when $n$ is a constant is $O(N)$.
+Thus the combined communication complexity is $O(N)$ for a constant $n$.
 
@@
 Note that it is not the same as the time complexity typically used in distributed systems.
-In analysis of distributed systems, time complexity is often in terms of the number of rounds.
-For example, ACS runs in a constant number of rounds because its subprotocols---reliable broadcast and binary Byzantine consensus---also run in a constant number of rounds~\cite{miller2016honey}.
+In the analysis of distributed algorithms, time complexity is often in terms of the number of rounds.
+For example, ACS runs in a constant number of rounds because of its sub-protocols---reliable broadcast and binary Byzantine consensus---also run in a constant number of rounds~\cite{miller2016honey}.
 However, in practice, making a unit of communication always has some overhead associated with it, for example serialising and writing it to some network socket.
@@
 \label{sec:communication-cost-for-tx}
-With the analysis on the duration of the consensus protocol,
-we are ready to analyse the amount of data required to be transmitted over a link per transaction,
-which we call the communication cost per transaction.
-% First, we define bandwidth as the ``amount of data a link or network path can deliver per unit of
-% time''~\cite{prasad2003bandwidth}.
+Geared with the results above,
+we are ready to analyse the amount of data required to be transmitted over a link per transaction.
+Which we call the communication cost per transaction.
 To create and then validate a transaction,
@@
 Using our results so far, we are able to analyse the global throughput.
-First we clarify the bandwidth definition, which is ``the data rate at which a network link or a network path can transfer''~\cite{prasad2003bandwidth}.
+First, we clarify the bandwidth definition, which is ``the data rate at which a network link or a network path can transfer''~\cite{prasad2003bandwidth}.
 Now, suppose every node has some fixed bandwidth $C$ per link, $N$ links and are making transactions at $r_{\text{tx}}$ per second.
@@
 $$NC \ge r_{\text{tx}} l,$$
-where $l$ is the length of the the agree fragment as before.
+where $l$ is the length of the agreed fragment as before.
 The inequality suggests that the rate for which transactions and validations are made cannot exceed the total bandwidth of all the links.
@@
 Recall that $l$ is of $O(N)$, that means LHS and RHS both grow linearly with respect to $N$.
-Hence, there exist some constant $r_\text{tx}$ that makes use of all the available bandwidth regardless of $N$.
+Hence, there exists some constant $r_\text{tx}$ that makes use of all the available bandwidth regardless of $N$.
 Finally, if every node in the network is transacting at a constant rate, then the global throughput (in terms of transactions per second) is linear w.r.t. the population size $N$.
@@
 
-\section{Effect of a highly adverserial environment}
+\section{Effect of a highly adversarial environment}
 
-Our last study consideres the effect when the number of adversaries is more than $t$.
+Our last study considers the effect when the number of adversaries is more than $t$.
 This is useful because in practice it is difficult to guarantee that $t$ satisfied $n \ge 3t + 1$, especially when $N$ is large.
@@
 what is the probability that more than $\lfloor \frac{n-1}{3} \rfloor$ are black?
-The random variable $X$ in this case is the number of black balls, or the number of successful events.
+The random variable $X$, in this case, is the number of black balls, or the number of successful events.
 It follows the hypergeometric distribution since we pick balls \emph{without} replacement~\cite{skala2013hypergeometric}.
@@
 
-The bound is not tight, but it is useful for for picking parameters.
+The bound is not tight, but it is useful for picking parameters.
 For some fixed $n$, since $0 \le \alpha \le \lfloor \frac{n-1}{3} \rfloor/n < (\lfloor  \frac{n - 1}{3} \rfloor + 1) / n$,

### 2017-08-06 thesis/chapters/ch5-implementation.tex
Write preface and update Bitcoin situation
Spelling/grammar check for ch3, 4, 5, 6

@@
 
-To build confidence of the horizontal scalability claim from our analytical results,
+To build up the confidence of the horizontal scalability claim from our analytical results,
 we must verify the same property experimentally.
@@
 Our implementation runs on a network with over 1000 node.
-Each node send transactions to each other autonomously, 
+Each node sends transactions to each other autonomously, 
 which allows us to evaluate our analytical claims.
@@
 \texttt{acs}, as its name suggests, implements ACS.
-ACS uses erasure code in one of its subprotocols (reliable broadcast).
+ACS uses erasure code in one of its sub-protocols (reliable broadcast).
 Thus we use the liberasurecode\footnote{\url{https://github.com/openstack/liberasurecode}} library for its Reed-Solomon error correcting code functionality.
@@
 Finally, the \texttt{node} module ties everything together.
-It implements the the consensus protocol, the transaction protocol and the validation protocol.
+It implements the consensus protocol, the transaction protocol and the validation protocol.
 
@@
 
-The the goal of the experiment is to run the three protocols---consensus protocol,
+The goal of the experiment is to run the three protocols---consensus protocol,
 transaction protocol and validation protocol---simultaneously and analyse the communication cost, the consensus duration and the global throughput.
@@
         The maximum $n$ is 32 because the limitation in liberasurecode mentioned in \Cref{sec:implementation},
-        but our results give a good dindication of how our system may function for larger numbers of $n$.
+        but our results give a good indication of how our system may function for larger numbers of $n$.
   \item The population size $N \in \{200, 300, \dots, 1200\}$.
@@
   \item The two modes of transaction.
-        The first mode is the \emph{fixed-neighbour} mode where nodes only transact with their immediate neighbour which are predetermined. 
+        The first mode is the \emph{fixed-neighbour} mode where nodes only transact with their immediate neighbour which is predetermined. 
         This minimises the data volume per validated transaction because agreed fragment can be cached.
@@
 From now on, we use ``machines'' to refer to DAS-5 nodes and ``nodes'' to refer to a running instance in our system.
-On DAS-5 we use up to 30 machine, for each machine we use up to 40 nodes.
+On DAS-5 we use up to 30 machines, for each machine we use up to 40 nodes.
 This gives us the aforementioned 1200 number.
@@
 But 40 nodes each need 1200 TCP connections which is 48000 TCP connections per machine and that is inching close to the limit.
-While it is possible to have many more TCP connections per machine,
+While it is possible to have more TCP connections per machine,
 but it requires additional network interface which is something we do not control on the DAS-5.
@@
 
-To coordinate nodes on many different machine,
+To coordinate nodes on many different machines,
 we employ a discovery server to inform every node the IP addresses and port numbers of every other node.
@@
   }
-  \caption{Round duration do not increase linearly with the population size. 
+  \caption{Round duration does not increase linearly with the population size. 
   This is likely due to the additional hashing operation required for larger consensus result.}
@@
 
-\Cref{fig:round-comms} show the relationship between the communication cost of the consensus protocol per round and the population size.
+\Cref{fig:round-comms} shows the relationship between the communication cost of the consensus protocol per round and the population size.
 The most important aspect is that these results show a linear increase.
@@
 Note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
-the magnitude of the communication cost are similar.
+the magnitude of the communication cost is similar.
 Both peak at about 100 MB.
 This is expected because the consensus protocol is decoupled from the transaction protocol and the validation protocol.
-Finally, the rate for which the communication cost increases is higher when the number of facilitators are higher.
-This is also expected because the ACS algorithm has a $n^2$ term in its communication complexity.
+Finally, the rate for which the communication cost increases is higher when the number of facilitators is higher.
+This is also expected because the ACS algorithm has an $n^2$ term in its communication complexity.
 
 We are also interested in how communication costs translate to time.
-Hence, for the same experiment we record the duration in seconds and the result is shown in~\Cref{fig:round-duration}.
+Hence, for the same experiment, we record the duration in seconds and the result is shown in~\Cref{fig:round-duration}.
 Interestingly, the duration is not entirely linear.
@@
 Since if $N$ increases, every node must also perform more hash operations.
-These result do not conform to analytical result in~\Cref{sec:cons-duration}.
+These results do not conform to analytical result in~\Cref{sec:cons-duration}.
 Nevertheless, the difference is minor and there are ways to optimise the luck value computation.
@@
 We send validation requests at the same rate as transactions.
-Upon receiving an (remote) agreed fragment, the caching mechanism inspects all the transactions in the agreed fragment and attempts to verify as many as it can,
+Upon receiving a (remote) agreed fragment, the caching mechanism inspects all the transactions in the agreed fragment and attempts to verify as many as it can,
 rather than just the transaction in the original validation request.
@@
 It is a lot more stable because for every transaction it is almost always the case that an agreed fragment is needed to validate it.
-% It is interesting to note that some fluctuations exist in the graphs.
-% This is likely due to the way how the communication cost are recorded.
-% We maintain the total amount of data sent and received in a few variables.
-% On every message that is sent or received, we increment the relevent variable by the size of the message.
-% However, it is infeasible to log the values every time a message is sent or received as there is simply too much data for our build server to hold.
-% Therefore we only log every 5 seconds.
-% This does not always align with the time when a round finishes, thus some of the data may not represent reality accurately.
-%  TODO this can be fixed
 
@@
 If a new agreed fragment needs to be transmitted to validate every transaction then it puts a toll on the network infrastructure.
-The low transaction rate in~\Cref{fig:global-throughput-random} is caused by fact that the network infrastructure cannot keep up with our demand.
+The low transaction rate in~\Cref{fig:global-throughput-random} is caused by the fact that the network infrastructure cannot keep up with our demand.
 
@@
 Up to this point we focused on the effect of communication cost, throughput and so on with respect to the population size.
-In this section we consider a varying number of facilitators,
-which gives us an insight on our system performance may perform when the number of facilitators are higher than 32.
+In this section, we consider a varying number of facilitators,
+which gives us an insight into our system performance may perform when the number of facilitators is higher than 32.
 
@@
   }
-  \caption{The communication cost of the consensus protocol increase polynomially with respect to the number of facilitators as expected from the ACS communication complexity.}
+  \caption{The communication cost of the consensus protocol increases polynomially with respect to the number of facilitators as expected from the ACS communication complexity.}
   \label{fig:facilitators-comms}

### 2017-08-06 thesis/chapters/ch6-conclusion.tex
Write abstract
Write preface and update Bitcoin situation
Spelling/grammar check for ch3, 4, 5, 6

@@
 
-In this work we introduced an application neutrual blockchain system which we call Checkpoint Consensus.
+In this work we introduced an application neutral blockchain system which we call Checo.
 We add checkpoint block to the existing TrustChain data structure for use in our consensus protocol.
@@
 The consensus result lets nodes elect new facilitators and create new checkpoint blocks.
-Node make transactions via the transaction protocol,
+Nodes make transactions via the transaction protocol,
 which is adapted from an existing work called True Halves.
-We introduce a validation protocol which ensure that if agreed fragments for some transaction exists,
+We introduce a validation protocol which ensures that if agreed fragments for some transaction exists,
 then nodes reach agreement on the validity of that transaction.
@@
 We do not attempt to prevent the Sybil attack, spam or DoS because the accuracy of the defence depends on the understanding of the application.
-For instance, without any application it would be impossible to distinguish between a very active node from a spammer because the content of the transaction ($m$) carry no meaning.
+For instance, without any application, it would be impossible to distinguish between a very active node from a spammer because the content of the transaction ($m$) carry no meaning.
 
 Our experiment is carried out in a somewhat idealised world (DAS-5).
-To understand our system further, we hope to experiment on a real platform with real users such as Tribler~\cite{pouwelse2008tribler}.
+To understand our system further, we hope to experiment with a real platform with real users such as Tribler~\cite{pouwelse2008tribler}.
 For instance, the TrustChain in Tribler stores the number of bytes uploaded and downloaded between users, but it has no global consensus.
 It would be useful to integrate our consensus protocol into Tribler to prevent misreporting attacks.
-More importantly it gives us an opportunity to observe the characteristics of our system in the real world and how it evolves with new features or alterations.
+More importantly, it gives us an opportunity to observe the characteristics of our system in the real world and how it evolves with new features or alterations.
 
-The fault tolerance property is adaquate for permissioned systems where nodes have relatively more trust between each other.
-However, it is not adaquate for permissionless systems,
+The fault tolerance property is adequate for permissioned systems where nodes have relatively more trust between each other.
+However, it is not adequate for permissionless systems,
 which usually aim to tolerate fault that involve a minority of the network,
@@
 for P2P file sharing systems the reputation score may be the number of bytes contributed to the network.
-Second, instead of using ACS we use a proof-of-stake (PoS) scheme which means that nodes with more stake in the system gets more vote.
+Second, instead of using ACS we use a proof-of-stake (PoS) scheme which means that nodes with more stake in the system get more vote.
 In an ideal PoS scheme, the system can tolerate any fault if the majority of the stake holders are honest.

### 2017-08-07 thesis/chapters/ch1-intro.tex
Fixes

@@
 The  same can be said for many other services such are online marketplaces,
-cloud services, hospitality services and even our banking system.
+cloud services and even our banking system.
 The 2008 financial crisis is an example of the banking system making poor choices which resulted in a
@@
 
-The primary innovations are (1) the consensus model which prevents double spending,
-and (2) the incentive mechanism that encourages anyone (with adequate hardware) to participate in the network and keep it running.
+The primary innovation is the consensus mechanism which prevents double spending but also incentivises anyone (with adequate hardware) to participate and keep the network running.
 The double spending problem can be seen as an inconsistency issue.
@@
 The miner is also rewarded with new coins and transaction fees.
-It is important to note that every block, that is the solution of the puzzle,
-depends on the previous block.
+It is important to note that every block depends on the previous block.
 Hence the name blockchain.
@@
 
-Bitcoin has had its ups and downs, but over it has grown into an enormous system.
+Bitcoin has had its ups and downs, but overall it has grown into an enormous system.
 Its power consumption is as high as Republic of Ireland~\cite{o2014bitcoin}.
-Its market capitalisation, at the time of writing, is over 60 billion USD~\cite{bitcoinmarketcap} and growing (\Cref{fig:market-cap}).
+Its market capitalisation, at the time of writing, is over 60 billion USD and growing (see \Cref{fig:market-cap}).
 \begin{figure}[h]
@@
 Many online marketplaces are using Bitcoin.
-For example, Steam\footnote{\url{https://store.steampowered.com/}} and even Amazon\footnote{Not directly, but via \url{https://purse.io/}.}.
+For example, Steam\footnote{\url{https://store.steampowered.com/}} and Amazon\footnote{Not directly, but via \url{https://purse.io/}.}.
 Due to its success, people from many different disciplines began investigating in ways to use blockchain technology.
@@
 
-What we have described is, in fact, the current state of Bitcoin.
-Many proposals exist with the aim to make Bitcoin more scalable.
+The Bitcoin community understands the issue and many proposals exist with the aim to make Bitcoin more scalable.
 The most prominent one is Segregated Witness or SegWit~\cite{segwit}.
@@
 The effect is that the first 1 MB of the block is still backwards compatible, thus there is no need for a hard fork.
-It also solves the transaction malleability problem~\cite{bitcoinmalleability} and paves the way for off-chain transactions via Lightning Network~\cite{lightningnetwork}.
-However, some fear SegWit is not enough and wish to see a hard fork that uses ``Adjustable Blocksize Cap'' that increase the block size up to 8 MB\footnote{\url{https://www.bitcoinabc.org/}}.
-The ongoing civil war resulted in a Bitcoin fork, called Bitcoin Cash\footnote{\url{https://www.bitcoincash.org/}}.
+It also solves the transaction malleability problem~\cite{bitcoinmalleability} which paves the way for off-chain transactions via Lightning Network~\cite{lightningnetwork}.
+However, some fear SegWit is not enough and wish to see a hard fork that uses ``Adjustable Blocksize Cap'' that increase the maximum block size even further to 8 MB~\cite{bitcoincash}.
+The ongoing civil war resulted in a Bitcoin fork, called Bitcoin Cash~\cite{bitcoincasharticle}.
 At the time of writing, it is too early to speculate how these proposals and forks may affect the scalability properties of Bitcoin.
@@
 most of the time the customer is honest and pays the bill.
-There is no need for the customer or the restaurant owner to report the transaction to any central authority 
+There is no need for the customer or the owner to report the transaction to any central authority 
 because both parties are happy with the transaction.
 On the other hand, if the customer leaves without paying the bill,
-then the restaurant owner would report the incident to some central authority, e.g. the police.
+then the owner would report the incident to some central authority, e.g. the police.
 On the contrary, in blockchain systems, every transaction is effectively sent to the miners,
@@
 We randomly select nodes in every round, the selected ones are called facilitators.
-They reach consensus not on the individual transactions,
-but on the state of every chain represented by a single digest, we call this state the checkpoint.
-If a checkpoint of some node is in consensus, 
-then that node can proof to any other node that it holds a set of transactions that compute (form a chain) to the checkpoint.
+Facilitators reach consensus not on the individual transactions,
+but on the state of every chain represented by a single digest, called a checkpoint block.
+If a checkpoint block of some node is in consensus, 
+then that node can proof to any other node that it holds a set of transactions that compute (form a chain) to the checkpoint block.
 This immediately shows that those transactions are tamper-proof.
@@
     \item We formally analyse Checo to ensure correctness as defined in our architecture.
-    \item We implement Checkpoint Consensus with experiment with up to 1200 nodes,
+    \item We provide an implementation and then experiment with up to 1200 nodes,
         our results show strong evidence of horizontal scalability.

### 2017-08-07 thesis/chapters/ch2-problem.tex
Fixes

@@
 The typical examples are proof-of-work based systems such as Bitcoin, Ethereum and other Altcoins.
-In Bitcoin, the level of consensus of a block containing many transactions is determined by how deep it is in the Bitcoin blockchain,
+In Bitcoin, nodes compete in solving puzzles.
+The first node solves the puzzle has the privilege to create a new block.
+The level of consensus of a block containing many transactions is determined by how deep it is in the Bitcoin blockchain,
 also called the number of confirmation.
@@
 The protocol disincentivises the parties from broadcasting old channel states.
-If this issue occurs, the counterparty can drain all the Bitcoins in the channel.
+If this occurs, the counterparty can drain all the Bitcoins in the channel.
 To close the channel, the parties simply broadcast the latest net settlement to the Bitcoin network,
@@
 On the other hand, off-chain transactions also suffer from the problem of Bitcoin.
-Proof-of-work still consumes an unreasonable amount of power.
+Namely, proof-of-work consumes an unreasonable amount of power.
 Further, payment channels complicate user experience.

### 2017-08-07 thesis/chapters/ch3-model.tex
Fixes

@@
 
-The consensus and transaction protocol by themselves do not provide a mechanism to detect forks or other forms of tampering.
+The consensus and transaction protocol by themselves do not provide a mechanism to detect tampering.
 Thus we need a validation protocol to counteract malicious behaviour.
@@
 
-Up to this point, we do not provide a mechanism to detect forks or other forms of tampering or forging.
+Up to this point, we do not provide a mechanism to detect tampering.
 The validation protocol aims to solve this issue.
@@
 A transaction can be in one of three states in terms of validity---\emph{valid}, \emph{invalid} and \emph{unknown}.
-Given a fragment $F_{v, j}$, the validity of the transaction $t_{u, i}$ is captured by the function $\textsf{get\_validity}(t, F_u, F_v)$ (\Cref{alg:get-validity}).
-The first four conditions (up to \Cref{line:valid-fragment}) essentially check whether the fragment is the one that the verifier needs.
+Given a fragment $F_{v, j}$, the validity of the TX block $t_{u, i}$ and the agreed fragment of it is captured by the function $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ (\Cref{alg:get-validity}).
+Note that $t_{u, i}$ and $F_{u, i}$ are assumed to be valid,
+otherwise the node calling the function would have no point of reference.
+This is not difficult to achieve.
+Typically the caller is $u$, so it knows its own TX block and the corresponding agreed fragment.
+If the caller is not $u$, it can always query for the agreed fragment that contains the transaction of interest from $u$.
+
+\Cref{alg:get-validity} works as follows.
+Before~\Cref{line:valid-fragment},
+we essentially check whether the fragment is the one that the verifier needs.
 If it is not, then the verifier cannot make any decision and return \emph{unknown}.
-This is likely to be the case for new transactions because $\textsf{agreed\_fragment}(\cdot)$ would be $\bot$.
+This is likely to be the case for new transactions because the result of $\textsf{agreed\_fragment}(\cdot)$ would be $\bot$.
 The next two conditions check for tampering or missing blocks, if any of these misconducts are detected, then the TX is invalid.
@@
 $u$\footnote{In practice, it often is because after completing the TX protocol the parties are incentivised to check that the counterparty ``did the right thing''.}
-Any node $w$ may call $\textsf{get\_validity}(t_u{u, i}, F_{u, i}, F_{v, j})$ as long as the caller $w$ has an agreed fragment of $t_{u, i}$---$F_{u, i}$.
-$F_{u, i}$ may be readily available if $w = u$ or it may be from some other \texttt{vd\_resp} message, which we describe next in the validation protocol.
+Any node $w$ may call $\textsf{get\_validity}(t_u{u, i}, F_{u, i}, F_{v, j})$ as long as $w$ has an the necessary input parameters.
+$F_{u, i}$  and $t_{u, i}$ may be readily available if $w = u$.
+Otherwise, $w$ could query $u$ with the \texttt{vd\_req} message and then obtain the result from \texttt{vd\_resp}.
+This is the validation protocol which we describe next.
 
@@
     % message tampering
-    \If{$m \ne m' $} 
+    \If{$m \ne m' \vee txid \ne txid'$} 
         \State \Return \emph{invalid}
@@
 \subsection{Validation protocol}
-Our validation protocol is designed to classify transactions according to the aforementioned validity definition.
-The protocol is a simple response and request protocol shown in \Cref{alg:vd-proto}.
+Our validation protocol,
+shown in~\Cref{alg:vd-proto},
+is designed to classify transactions according to the aforementioned validity definition.
 If $u$ wishes to validate some TX with ID $txid$ and counterparty $v$, it sends $\langle \texttt{vd\_req}, txid \rangle$ to $v$.
@@
 There are a variety of approaches to do global fork detection.
-First and the easiest solution is to simply ask for every agreed fragment from the one containing the genesis block up to the one that the to-be-proven TX block is in.
-If the hash pointers are correct and all the CP blocks are in consensus, then the verifier can be sure that there are no forks.
+First and the easiest solution is to simply ask for the complete chain of the counterparty.
+The verifier can be sure that there are no forks if the following holds.
+\begin{enumerate}
+    \item The hash pointers are correct.
+    \item All the CP blocks are in consensus.
+    \item The TX of interest is in the chian.
+\end{enumerate}
 We use this approach in our prior work on Implicit Consensus~\cite{implicitconsensus}.
 It may sound inefficient at first, but nodes employ caching to minimise communication costs.
-In our work, we call this effect ``spontaneous sharding''.
 
-The second approach is probabilistic but with only a constant communication overhead.
-For a node, observe that if all of its agreed fragment has a transaction with an honest node,
+The second approach is probabilistic but with only a constant communication overhead over our current design.
+For a node, observe that if all of its agreed fragments has a transaction with an honest node,
 then the complete chain is effectively validated in a distributed manner.
@@
 If there are more nodes (say $n$) trying to make transactions with the malicious node,
-then the probability that the malicious node gets caught at least once follows a binomial distribution
+then the probability that the malicious node gets caught at least once (denoted by the random variable $X$) follows a binomial distribution, i.e.
 $$\Pr[X > 0] = \sum_{k = 1}^{n}\binom{n}{k}p^k (1 - p)^{n - k}.$$

### 2017-08-07 thesis/chapters/ch4-analysis.tex
Fixes

@@
 \subsection{Correctness of the validation protocol}
-The consensus protocol (on CP blocks and facilitators) is the backbone for consensus on transactions.
+The consensus protocol,
+which ensures consensus on CP blocks and facilitators,
+is the backbone for consensus on transactions.
 In this section, we build on top of \Cref{theorem:consensus} to show that the properties except for liveness in \Cref{def:validation} can be satisfied.
@@
 We proof the agreement property by contradiction.
+From~\Cref{alg:get-validity},
+we use the fact that $F_{u, i}$ and $t_{u, i}$ are valid.
 Without loss of generality, suppose for some transaction with TX block $t$,
@@
 Thus we have a contradiction and this completes the agreement proof.
+
 Our protocol uses \Cref{alg:get-validity} which is also the validity definition,
-this completes the validity proof.
+so the validity property is also satisfied.
 \end{proof}
@@
 In addition, a convenient consequence \Cref{theorem:validation-agreement} is unforgeability.
-That is, no polynomial time adversary is able to create two chains $F = \{ \dots, t, c\}$ $F' = \{ \dots, t', c\}$ with correct hash pointers and the same end of chain $c$.
+That is, no polynomial time adversary is able to create two chains, $F = \{ \dots, t, c\}$ and $F' = \{ \dots, t', c\}$, with correct hash pointers and the same end of chain $c$.
 % Another way to think about it is if a transaction can be forged, then it cannot be agreed.
@@
 A stronger version of the validity definition exists.
-That is, if two nodes are making transactions with each other,
-then the resulting transaction is always valid in addition to our current validity definition.
+That is, if two honest nodes make a transaction,
+then the transaction state is always valid in addition to our current validity definition.
 Under our purely asynchronous model, we cannot guarantee this stronger version.

### 2017-08-07 thesis/chapters/ch5-implementation.tex
Fixes

@@
   \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-fixed/consensus-communication-cost-vs-facilitators}
+    \includegraphics[width=\textwidth]{neighbour-fixed/round-communication-cost-vs-facilitators}
     \caption{Transactions are with fixed neighbours.}
-    \label{fig:facilitators-comms-fixed}
+    \label{fig:facilitators-round-comms-fixed}
   \end{subfigure}%
   \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-random/consensus-communication-cost-vs-facilitators}
+    \includegraphics[width=\textwidth]{neighbour-random/round-communication-cost-vs-facilitators}
     \caption{Transactions are with random neighbours.}
-    \label{fig:facilitators-comms-random}
+    \label{fig:facilitators-round-comms-random}
   \end{subfigure}
@@
   \caption{The communication cost of the consensus protocol increases polynomially with respect to the number of facilitators as expected from the ACS communication complexity.}
-  \label{fig:facilitators-comms}
+  \label{fig:facilitators-round-comms}
 \end{figure}
 
-\Cref{fig:facilitators-comms} shows the communication cost of the consensus protocol for varying numbers of facilitators.
+\Cref{fig:facilitators-round-comms} shows the communication cost of the consensus protocol for varying numbers of facilitators.
 There is strong evidence that the communication cost grows polynomially.

### 2017-08-07 thesis/chapters/ch6-conclusion.tex
Fixes

@@
 The round based consensus protocol uses ACS as a building block to reach consensus on checkpoint blocks.
-The consensus result lets nodes elect new facilitators and create new checkpoint blocks.
-Nodes make transactions via the transaction protocol,
+The consensus result, which is a set of checkpoint blocks, lets nodes elect new facilitators and create new checkpoint blocks.
+To make transactions, nodes use the transaction protocol,
 which is adapted from an existing work called True Halves.
-We introduce a validation protocol which ensures that if agreed fragments for some transaction exists,
+Finally, we introduce a validation protocol which ensures that if agreed fragments for some transaction exists,
 then nodes reach agreement on the validity of that transaction.
@@
 Our experiment is carried out in a somewhat idealised world (DAS-5).
+We hope to experiment our prototype in a more realistic environment.
+For example, DeterLab~\cite{mirkovic2012teaching} lets users simulate network delay,
+which is important for discovering issue that may only occur in hostile environments.
 To understand our system further, we hope to experiment with a real platform with real users such as Tribler~\cite{pouwelse2008tribler}.

### 2017-08-07 thesis/formalities/preface.tex
Fixes

@@
 \addcontentsline{toc}{chapter}{Preface}
-I mined my first Bitcoin in early 2013 with a GPU.
-Since then, I have been interested in decentralising the traditionally centralised systems.
-Following the Snowden revelations, the interest became a burning desire.
+Since early 2013, when I mined my first Bitcoin, 
+I have been interested in decentralising the traditionally centralised systems.
+Following the Snowden revelations, the interest became a desire that has stayed with me ever since.
 Today, many blockchain systems are mature, but still not yet ready for ubiquitous use due to the scalability limitation.

### 2017-08-07 thesis/chapters/ch1-intro.tex
Fixes

@@
 but nevertheless verifiable at a later stage by any node in the network.
+We call our protocol Checo, derived from \emph{che}ckpoint \emph{co}nsensus.
 Informally, our system works as follows.
@@
 This immediately shows that those transactions are tamper-proof.
-We call our protocol Checo, derived from \emph{che}ckpoint \emph{co}nsensus.
-% TODO mention consensus and trade off?
+The key difference from related work is the independence of the transaction validation phase.
+Prior systems such a Bitcoin and Ethereum entangled the consensus algorithm with the transaction validation, restricting performance.
+By splitting this functionality we achieved horizontal scalability in the general case.
 

### 2017-08-07 thesis/chapters/ch3-model.tex
Fixes

@@
 For example, if every facilitator is honest, a node would receive $n$ consensus results which are identical when only one is necessary.
+Furthermore, realistic networks do not offer a broadcast primitive,
+i.e. firewalls may block incoming messages from certain nodes.
 
-An optimisation we can do is to use an epidemic protocol~\cite{eugster2004epidemic} (also known as gossiping) instead of our broadcast approach.
+An improvement is use an epidemic protocol~\cite{eugster2004epidemic} (also known as gossiping) instead of our broadcast approach.
 Typical epidemic protocols works as follows.

### 2017-08-07 thesis/chapters/ch1-intro.tex
Change paper template; fix thesis template

@@
 
+\section{Contributions}
 In this work, we take the advice of Croman et el.~\cite{croman2016scaling} and Vukoli\'{c}~\cite{vukolic2015quest} and rethink the blockchain architecture.
@@
 
+\section{Document structure}
 We describe and analyse our idea in the remainder of this work.

### 2017-08-07 thesis/chapters/ch3-model.tex
Change paper template; fix thesis template

@@
 
-The TX block is a six-tuple, i.e $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$.
+The TX block is a six-tuple, i.e
+$$t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle.$$
 We describe each item in turn.
@@
 
-The CP block is a five-tuple, 
-i.e. $c_{u, i} = \langle \textsf{H}(b_{u, i-1}), seq_u, \textsf{H}(\C_r), r, sig_u \rangle$,
-where $\C_r$ is the consensus result (which we describe in \Cref{sec:consensus-result}) in round $r$, the other items are the same as the TX block definition.
+The CP block is a five-tuple, i.e. 
+$$c_{u, i} = \langle \textsf{H}(b_{u, i-1}), seq_u, \textsf{H}(\C_r), r, sig_u \rangle,$$
+where $\C_r$ is the consensus result (which we describe in \Cref{sec:consensus-result}) in round $r$,
+the other items are the same as the TX block definition.
 % Note that unlike in our prior work~\cite{implicitconsensus}, CP blocks and TX blocks do not have independent sequence numbers.
@@
 The genesis block in the chain must be a CP block in the form of
-$c_{u, 0} = \langle \textsf{H}(\bot), 0,  \textsf{H}(\bot), 0, sig_u \rangle$
+$$c_{u, 0} = \langle \textsf{H}(\bot), 0,  \textsf{H}(\bot), 0, sig_u \rangle,$$
 where $\textsf{H}(\bot)$ can be interpreted as applying the hash function on an empty string.

### 2017-08-07 thesis/chapters/ch5-implementation.tex
Fixes
Change paper template; fix thesis template

@@
 Thus we use the liberasurecode\footnote{\url{https://github.com/openstack/liberasurecode}} library for its Reed-Solomon error correcting code functionality.
-An implementation detail is that liberasurecode cannot create more than 32 code blocks\footnote{
-  The 32 code blocks limitation is hardcoded in the source file, see
-  \url{https://github.com/openstack/liberasurecode/blob/0794b31c623e4cede76d66be730719d24debcca9/include/erasurecode/erasurecode.h\#L35}
+An implementation detail is that liberasurecode cannot create more than 32 code blocks, which is hardcoded in the source file\footnote{\url{https://github.com/openstack/liberasurecode/blob/0794b31c623e4cede76d66be730719d24debcca9/include/erasurecode/erasurecode.h\#L35}
 }, we discuss the effect of this in~\Cref{sec:experimental-setup}.
@@
 The low transaction rate in~\Cref{fig:global-throughput-random} is caused by the fact that the network infrastructure cannot keep up with our demand.
+In practice, we do not expect such behaviour to occur as it is possible to cache agreed fragments.
 

### 2017-08-08 paper/s2-related.tex
Give paper a structure

--- 2017-08-07 thesis/chapters/ch2-problem.tex
@@
-\chapter{Problem description}
-\label{ch:problem}
+\section{Background and Related Work}
+% give background on ACS
+% give background on TrustChain?
 
-From the Introduction, we saw that early blockchain systems face a scalability problem.
-In particular, the use proof-of-work is unlikely to reach a transaction rate suitable for ubiquitous use.
-Many attempts to fix the scalability problem exist in academia as well as the industry.
-In this chapter, we describe the related work in the form of a mini taxonomy and identify areas for improvement and limitations.
-Then we state our research question which aims to overcome some of the limitations.
+\subsection{Background on asynchronous subset consensus}
+\label{sec:acs-background}
 
-\section{Related work and mini taxonomy}
-\label{sec:taxonomy}
+ACS is an especially useful primitive for blockchain systems.
+It allows any party to propose a value and the result is the set union of all the proposed values by the majority.
+Concretely, ACS needs to satisfy the following properties (adapted from~\cite{miller2016honey}).
+\begin{definition}
+\label{def:acs}
+\textbf{\emph{Asynchronous subset consensus}}
 
-Blockchain research has seen a surge in recent years from both the industry and academia,
-this resulted in a myriad of blockchain systems with different approaches to scalability.
-To understand the current state-of-the-art,
-we classify various blockchain systems by their scalability approach.
+There are $n$ nodes, of which at most $t$ might experience Byzantine fault.
+Node $i$ starts with a non-empty set of input values $C_i$.
+The nodes must decide an output $C$, satisfying the following.
+\begin{enumerate}
+    \item \emph{Agreement:}
+        If a correct node outputs $C$, then every node outputs $C$.
+    \item \emph{Validity:}
+        If any correct node outputs a set $C$,
+        then $|C| \ge n - t$ and $C$ contains the input of at least $n - 2t$ nodes.
+    \item \emph{Totality:}
+        If $n - t$ nodes receive an input, then all correct nodes produce an output.
+\end{enumerate}
+\end{definition}
 
-\subsection{Early blockchain systems}
-This category represents the baseline,
-which are systems with a probabilistic consensus algorithm.
-That is, transactions never reach consensus with a probability of 1.
-The typical examples are proof-of-work based systems such as Bitcoin, Ethereum and other Altcoins.
-In Bitcoin, nodes compete in solving puzzles.
-The first node solves the puzzle has the privilege to create a new block.
-The level of consensus of a block containing many transactions is determined by how deep it is in the Bitcoin blockchain,
-also called the number of confirmation.
-The probability of a block being orphaned drops exponentially as the depth increases~\cite{bitcoin}.
-Nevertheless, the probability of the highest block being orphaned is non-negligible.
+ACS has the nice property of censorship resilience when compared to other consensus algorithms.
+For instance, Hyperledger and Tendermint uses a variant of Practical Byzantine Fault Tolerance (PBFT)~\cite{castro1999practical} as their consensus algorithm.
+In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectively filter transactions.
+In contrast, every party in ACS are involved in the proposal phase,
+and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
+Thus, if some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
+For a detailed description of ACS, we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.
+To understand the remainder of this work, the knowledge of~\Cref{def:acs} is sufficient.
 
-The advantage of this approach is that it can be used in a large network.
-In fact, prior to Bitcoin, there was no internet-scale consensus protocol.
-Bitcoin is also fault-tolerant because attackers can not out pace honest users in finding new blocks unless they have a majority of the hash power.
-As long as the majority is honest and reasonable measures are taken to prevent the selfish mining attack~\cite{eyal2014majority}.
-The disadvantage, however, is that transactions are never in consensus with a probability of 1---no consensus finality.
-Hence double spending is possible for transactions that do not have a large number of confirmations.
-Furthermore, as we mentioned in the Introduction,
-the performance is limited due to the fact that blocks are of fixed size and are generated at fixed intervals.
-Thus early blockchain systems do not achieve horizontal scalability.
+The main drawback with ACS and also Byzantine consensus algorithms is the high message complexity.
+Typically, such protocols have a message complexity of $O(n^2)$, where $n$ is the number of parties.
+Hence, it may work with a small number of nodes,
+but it is infeasible for blockchain systems where thousands of nodes are involved.
 
-% Our system significantly improves upon Bitcoin and other classical blockchain systems in performance, scalability and consensus finality.
-% The results described in \Cref{ch:analysis} and \Cref{ch:implementation}, show that we have horizontal scalability,
-% where more nodes result in more global throughput.
-% Further, we do not have the aforementioned probabilistic behaviour,
-% once some consensus result is decided, it cannot be orphaned, thus our consensus is final.
-% The leadership election is also not ideal in classical blockchain systems.
-% Mining can be seen as a technique to elect a single leader.
-% The leader has full control of what goes into the block thus it may selectively censor transactions.
-% We use ACS, so as long as the CP block is in $n - 2t$ nodes, it is guaranteed to be in consensus.
+% The literature on Byzantine consensus and atomic broadcast is rich, some noteable ones include TODO.
+% Thus in the ETC consensus protocol, we assume there exist an "off-the-shelf" Byzantine consensus algorithm which we can use
+% (in our implementation we use HoneyBadgerBFT and motivate our choice in TODO).
 
-% However, the security aspect falls short of the ``honest majority'' security model that classical blockchains claim to 
-% have\footnote{Recently it was shown that doing selfish-ming would give the adversary an unfair advantage when she only controls 25\% of the mining power~\cite{eyal2014majority}}.
-% Our system risks going into erroneous state if the inequality $n \ge 3t + 1$ is not satisfied.
-% Classical blockchain systems also have an incentive mechanism, thus they do not depend on altruistic nodes and encourages participation.
-% Our system on the other hand does not have an incentive mechanism because we make no assumptions on the application.
+% In our case, every node do not propose some arbitrary data, but a set of CP blocks.
+% Thus the result of the consensus is the set union of all the legitimate proposals.
 
@@
 Applications such as digital cash can not be implemented on top of such systems because consistency is needed in solving the double spending problem.
-
-% Our system can be considered as the same as these types of blockchains but with a lightweight consensus protocol.
-% Consensus might not be applicable in all applications.
-% But we believe it is important for detecting and preventing fraud.
-% The example in~\Cref{fig:trustchain-bad} on page~\pageref{fig:trustchain-bad} demonstrates this.
-% If $b$ makes a fork and $a$ and $c$ have no way to communicate (e.g. the adversary may control parts of the network) ,
-% then $c$ is tricked to believe that her transaction with $b$ is valid.
-% Only when $c$ sees a conflicting chain is she able to tell that the transaction is invalid.
-% But $c$ does not know the true end-of-chain of $b$, thus she can never know whether her transaction is valid.
-% This is not possible in our system because $b$ cannot convince $c$ unless he can compute exponential time algorithms
-% (this is finding the second preimage for a hash function).
-
-% However, consensus and validation comes at a cost that do not exist in Tangle, Corda or the original TrustChain.
-% Our transaction rate is affected by the validation protocol in the worst case scenario as we saw in~\Cref{sec:evaluation}.
-
-\section{Research question}
-
-We saw in~\Cref{sec:taxonomy} that the majority of the systems cannot achieve horizontal scalability.
-That is, by adding new nodes into the network, the global transaction rate should increase proportionally.
-To the best of our knowledge, only two systems---Elastico~\cite{luu2016elastico} and OmniLedger~\cite{kokoris2017omniledger}---has experimentally demonstrated horizontal scalability.
-However, these techniques rely on sharding, thus the performance highly depends on transaction characteristics and the shard size parameter because inter-shard transactions are slow.
-Furthermore, all the systems described in~\Cref{sec:taxonomy} are complete blockchain systems designed for a concrete application.
-We wish to design a blockchain consensus protocol that can be used in many applications.
-For example recording uploads and downloads between peers in Tribler~\cite{pouwelse2008tribler, multichain} and decentralised market\footnote{\url{https://github.com/Tribler/tribler/blob/devel/Tribler/community/tradechain/__init__.py}}.
-Hence our research question is the following.
-\begin{displayquote}
-\emph{Is it possible to design a blockchain consensus protocol that is fault-tolerant, scalable and can reach global consensus?}
-\end{displayquote}
-
-In order to make sense of our research question, we must first define \emph{blockchain consensus protocol}.
-Typical blockchain systems are application specific.
-For example, the application in Bitcoin (any many of its derivatives) is digital cash, the application in Namecoin is domain name registration~\cite{namecoin} and the application in Siacoin is cloud storage~\cite{siacoin}.
-Underneath these applications lies the blockchain consensus protocol which has the goal of reaching consensus on transactions or state changes.
-In the case of Bitcoin, it is proof-of-work (PoW).
-It is not concerned with the structure of the transactions,
-it works by simply treating transactions as a blob of data and hashes them with the previous block.
-In this work, we focus on the blockchain consensus protocol rather than any specific application.
-Because it is the scalability bottleneck and is necessary for any blockchain based application.
-% Also, it should be possible to build any kind of application on top our system.
-% We do not impose on a consensus algorithm, as long as it satisfies the properties of atomic broadcast which we describe in~\Cref{sec:overview-cons}.
-
-In the remainder of this section, we expand on our research question and visit each of our requirement in detail.
-At the end of this section, we describe some issues that we do not consider as a part of our design,
-which is the result of not focusing on any specific application.
-
-\subsection{Global consensus}
-
-% Early blockchain systems use PoW as the consensus algorithm,
-% but consensus algorithm and especially Byzantine\footnote{Byzantine means participants can experience arbitrary failure, more on this in~\Cref{sec:acs-background}.}
-% consensus algorithms have existece since the 80s~\cite{pease1980reaching, lamport1982byzantine}.
-% The Practical Byzantine Fault Tolerance (PBFT) algorithm introduced in 1999 by Castro and Liskov~\cite{castro1999practical} is widely regarded as the first high-performance consensus algorithm.
-% The authors show that running PBFT in Network File System (NFS) was only 3\% slower than the standard unreplicated NFS.
-% Since then, research in this area picked up pace and many more algorithms are introduced, for example Q/U~\cite{abd2005fault} and Zyzzyva~\cite{kotla2007zyzzyva}.
-% With the introduction of Bitcoin's probabilistic consensus model, the field experienced a second renaissance.
-% These include proof-of-stake protocols~\cite{bentov2016snow, micali2016algorand}, as well as protocols optimised for blockchain~\cite{liu2016xft, miller2016honey}.
-
-The first objective is to reach consensus on a global \emph{state} by building on top of decades of Byzantine consensus protocol research~\cite{pease1980reaching, lamport1982byzantine, castro1999practical, miller2016honey}.
-Having consensus is essential in blockchain systems.
-It stops many types of malicious activities because agreed state must have the consent of the honest nodes in the network.
-We emphasise the word state because in contrary to traditional blockchains,
-it does not necessarily mean the set of all transactions that have ever happened.
-The state can be a verifiable representation of all the transactions.
-As an extreme example, a single digest of the complete history can represent the state.
-It is even verifiable as long as there are nodes that store the history.
-
-% Our approach is to combine the vast literature in Byzantine fault tolerance research with blockchain technology,
-% and introduce a new consensus model to record the global state.
-
-\subsection{Security and fault-tolerance}
-
-Our system should be unaffected in the presence of powerful adversaries.
-In particular, adversaries are Byzantine meaning that they can have arbitrary behaviour.
-Thus anything is possible from simply omitting messages to colluding with each other in order to undermine the whole system.
-This aspect usually comes for free when using a Byzantine fault-tolerant consensus algorithm.
-
-There is another aspect of fault tolerance that is transaction verifiability.
-All transactions must be eventually verifiable to maintain system integrity.
-Further, the validation result must be consistent.
-For example, if two different honest nodes attempts verify the same transaction,
-it cannot be the case that one node thinks that the transaction is valid but the other thinks it is invalid.
-
-% Recall that it is not neccessary to reach consensus on all the transactions in order to record the state.
-% But when using some digest (or a set of digets) to represent the state, it is also important for the transactions to be verifiable.
-% Otherwise adversaries could simply store invalid transactions in the state and they would never be detected.
-
-\subsection{Performance and scalability}
-
-The last but not least, the final objective is horizontal scalability.
-We wish to see the performance (global transaction rate) increase as more nodes join the network.
-BitTorrent~\cite{cohen2003incentives} is an example of such a system,
-where peers interact with each other to exchange files without any global bottleneck.
-
-Early blockchain systems may run in a network of thousands of nodes, but the performance does not scale.
-On the other hand, as we will see in~\Cref{sec:acs-background}, Byzantine consensus algorithms perform well but only when in a small network.
-In this work, our goal is to overcome the limitations of early blockchain systems and Byzantine consensus algorithms to create a horizontally scalable blockchain.
-
-\subsection{Limitations}
-Our work aims to significantly lower the transaction bound.
-However, the price to pay is that the Sybil attack~\cite{douceur2002sybil} prevention mechanism moves from the blockchain consensus protocol to the application layer.
-For example, social network based Sybil defence mechanisms use graph structure of real-world relationships~\cite{yu2006sybilguard}.
-Online marketplaces such as Amazon use the rating of buyer and sellers.
-Since our system is application neutral, we do not provide an explicit Sybil defence mechanism.
-On the other hand, our system also has no restrictions on the Sybil defence technique
-and applications built on top of our system can use the best mechanism for their purpose.
-
-For the same reason, we do not explicitly consider spam or denial of service attacks.
-Without a concrete application, it is difficult to distinguish a super active user from an attacker.
-
-% Scalability is indeed one of the key challenges we face in blockchain systems today.
-% Bitcoin~\cite{bitcoin}, the largest permissionless\footnote{Explain permissionless} blockchain system
-% in terms of market capitalisation~\cite{bitcoinmarketcap} has a maximum transaction rate of merely 7 transaction per second (TX/s).
-% This is due to the consensus mechanism in Bitcoin, namely proof-of-work (PoW),
-% miners can only create new blocks every 10 minutes and every block cannot be larger than 1 megabyte.
-% Payment processors in use today such as Visa can handle transaction rates in the order of thousands~\cite{visa}.
-% While Bitcoin may be a revolutionary phenomenon, it clearly cannot be ubiquitous in its current state.
-
-% An different approach is to not reach global consensus at all.
-% For instance in TrustChain~\cite{multichain} and Tangle~\cite{tangle},
-% nodes in the network only store their personal ledger.
-% Since consensus is left out, nodes can perform transactions as fast as their machine and network allows.
-% The downside of this approach is that it cannot prevent fraud (it is possible to detect fraud).
-% To examplify, a malicious node Mallory may claim she has 3 units of currency to Alice,
-% but in reality Mallory already spent all of it on Bob.
-% If there is no global consensus and Bob and Alice never communicate,
-% then the 3 units that Alice is about to receive is nonexistent.
-
-% The scalability property of TrustChain and Tangle are exceptionally desirable.
-% The global consensus mechanism of Bitcoin and many other blockchain systems are also worthwhile for detecting or preventing fraud.
-% These two properties may seem mutually exclusive, but in this work, we demonstrate the opposite.
-% Specifically, we answer the following research question in the affirmative.
-% \emph{Is it possible to design a blockchain fabric that can reach global consensus on the state of the system 
-% and also scalable?} We define scalability as a property where if more nodes join the system, then the transaction rate should increases.
-% % this is not completely true, there's a cap

### 2017-08-08 paper/s3-system.tex
Give paper a structure

--- 2017-08-07 thesis/chapters/ch3-model.tex
@@
-\chapter{System architecture}
-\label{ch:model}
-
-% As was shown in~\Cref{ch:problem}, the requirements are the following.
-% \begin{itemize}
-%     \item Global consensus
-%     \item Security and fault tolerance
-%     \item Performance and scalability
-% \end{itemize}
-
-% As mentioned in the Introduction, the primary goal is performance and scalability.
-% Having a scalable blockchain system while still keeping global consensus
-% allows the system to be ubiquitous and realise the full potential of blockchain.
-
-% The secondary goal is to design an application neutral system.
-% In particular, it should act as a backbone that provides the building blocks of blockchain based applications.
-% It should be possible to buld any kind of application on top our system.
-% Further, we do not impose on a consensus algorithm, as long as it satisfies the properties of atomic broadcast which we describe in~\Cref{sec:overview-cons}.
-
-% The third and final goal is security.
-% Our system should be unaffected in the presence of powerful adversaries.
-% In particular, adversaries are Byzantine meaning that they can have arbitrary behaviour.
-% Thus anything is possible from simply ommitting messages to colluding with eachother in order to undermine the whole system.
+\section{System Architecture}
 
@@
 
-The goal of this chapter is to detail our system specification.
-We begin our discussion with an intuitive overview of the architecture in \Cref{sec:system-overview}.
-Next, we give the formal description, starting with the model and assumptions in \Cref{sec:model-assumptions}.
-Then, the three protocols which make up the complete system,
-namely consensus protocol (\Cref{sec:cons-protocol}), transaction protocol (\Cref{sec:tx-protocol}) and validation protocol (\Cref{sec:vd-protocol}).
-Finally, we discuss a few variations of our main design and their respective tradeoffs in \Cref{sec:tradeoffs}.
-
-\section{System overview}
-\label{sec:system-overview}
-The architecture is visualised in~\Cref{fig:architecture}.
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.7\textwidth]{architecture}
-\caption{Overview of the system architecture}
-\label{fig:architecture}
-\end{figure}
-It consists of one data structure---Extended TrustChain, and three protocols.
-The protocols have distinct roles which are evident from their names.
-They run concurrently and do not interact with each other.
-The only synchronization happens on the Extended TrustChain layer, where all the protocols read and write from it.
-We first describe each component individually, starting with Extended TrustChain, and then explain how they fit together.
-
-\subsection{Extended TrustChain}
-Extended TrustChain naturally builds on top of the standard TrustChain. 
-Thus we first describe the standard TrustChain.
-Our description has minor differences compared to the description in~\cite{multichain}.
-This is to help with the description of the extended TrustChain.
-We remark the difference when it occurs.
-However, the two descriptions are functionally the same.
-
-\subsubsection*{Standard TrustChain}
-In TrustChain, every node has a ``personal'' chain. 
-Initially, the chain only contains a genesis block generated by the nodes themselves.
-When a node $A$ wishes to add a new transaction (TX) with $B$,
-a new TX block is generated for both $A$ and $B$ and is appended to their respective chains.
-The TX block must have a valid hash pointer pointing to the previous block
-and a reference\footnote{This is different from the original TrustChain definition found in~\cite{multichain}.
-In there, a TX block has two outgoing edges which are hash pointers to the two parties involved in the transaction.
-This work uses one outgoing edge and a reference.} to its \emph{pair} on $B$'s chain.
-An example of is shown in \Cref{fig:trustchain-bad}.
-
-\begin{figure}
-    \includegraphics[width=0.9\textwidth]{trustchain-bad}
-    \centering
-    \caption{Every block is denoted by $t_{i,j}$, where $i$ is the node ID and $j$ is the sequence number of the block.
-    Thus we have three nodes and three corresponding chains in this example.
-    The arrows represent hash pointers and the dotted lines represent references.
-    The blocks at the ends of one dotted line are pairs of each other.
-    The red block after $t_{b, 5}$ indicates a fork.}
-    \label{fig:trustchain-bad}
-\end{figure}
-
-If every node follows the rules of TrustChain and we only consider hash pointers and not references,
-then every chain effectively forms a singly linked list.
-However, if a node violates the rules, then a \emph{fork} may happen.
-That is, there may be more than one TX block with a hash pointer pointing back to the same block.
-In \Cref{fig:trustchain-bad}, node $b$ (in the middle chain) created two TX blocks that both point to $t_{b, 5}$.
-If this is a ledger system it can be seen as a double spend, where the currency accumulated up until $t_{b, 5}$ is spent twice.
-
-\subsubsection*{Extended TrustChain}
-We are now ready to explain the Extended TrustChain.
-The primary difference is the introduction of a new type of block---checkpoint (CP) block.
-In contrast to TX blocks, CP blocks do not store transactions or contain references.
-Their purpose is to capture the state of the chain and the state of the whole system.
-In particular, the state of the chain is captured with a hash pointer.
-The state of the whole system is captured in the content of the CP block,
-namely as a digest of the latest \emph{consensus result} which we explain in \Cref{sec:overview-cons}.
-A visual representation is shown in \Cref{fig:trustchain-bad-cp}.
-
-From this point on, we use TrustChain to mean the Extended TrustChain unless explicitly clarified.
-
-\begin{figure}
-    \includegraphics[width=0.9\textwidth]{trustchain-bad-cp}
-    \centering
-    \caption{The circles represent CP blocks,
-    they also have hash pointers (arrow) but do not have references (dotted line).
-    Note that the sequence number counter do not change, it is shared with TX blocks.}
-    \label{fig:trustchain-bad-cp}
-\end{figure}
-
-\subsection{Consensus protocol}
-\label{sec:overview-cons}
-
-% The consensus protocol uses an existing Byzantine consensus algorithm (explained in detail in \Cref{sec:xyz}) and runs it continuously in rounds.
-The consensus protocol can be seen as a technique of running infinitely many rounds of some 
-Byzantine consensus algorithm\footnote{More accurately it is ACS or asynchronous subset consensus, we describe ACS in \Cref{sec:acs-background}.},
-starting a new execution immediately after the previous one is completed.
-This is necessary because blockchain systems always need to reach consensus on new values proposed by nodes in the system,
-or CP blocks in our case.
-
-The high communication complexity of Byzantine consensus algorithms prohibits us from running it on a large network.
-Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks from every other node and use them as the proposal.
-The facilitators are elected using a \emph{luck value}, which is computed using $\textsf{H}(\C_r || pk_u)$,
-where $\C_r$ is the consensus result 
-(which can be seen as the set union of all the CP blocks collected by the facilitators)
-in round $r$ and $pk_u$ is the public key of node $u$.
-Intuitively, the election is guaranteed to be random 
-because the output of a cryptographically secure hash function is unpredictable and $\C_r$ cannot be determined in advance.
-Finally, the elected facilitators run a Byzantine consensus algorithm to agree on a set of CP blocks.
-The next round begins when the new facilitators are computed from these CP blocks.
-
-A visual explanation can be found in \Cref{app:consensus-example},
-it walks through the steps needed for a node to be selected as a facilitator using an example.
-
-\subsection{Transaction and validation}
-The TX protocol is a simple request and response protocol.
-The nodes exchange one round of messages and create new TX blocks on their respective chains.
-Thus, as we mentioned before, one transaction should result in two TX blocks.
-
-The consensus and transaction protocol by themselves do not provide a mechanism to detect tampering.
-Thus we need a validation protocol to counteract malicious behaviour.
-When a node wishes to validate one of its TX, it asks the counterparty for the \emph{agreed fragment} of the TX.
-An agreed fragment of a TX block is a section of the chain beginning and ending with CP blocks that contain the TX block,
-the CP blocks must also be in consensus.
-Upon the counterparty's response, the node checks that the CP blocks are indeed in consensus,
-the hash pointers are valid and his TX is actually in the fragment.
-The TX is valid if these conditions are satisfied.
-Intuitively, this works because it is hard (because a cryptographically secure hash function is second-preimage resistant)
-to create a different chain that begins and ends with the same two CP blocks.
-
-\subsection{Combined protocol}
-\label{sec:combined-protocol}
-The final result is essentially the concurrent composition of the three aforementioned protocols,
-all making use the Extended TrustChain data structure.
-
-Our sub-protocol design gives us the highly desirable non-blocking property.
-In particular, we do not need to ``freeze'' the state of the chain for some communication to complete in order to create a block.
-For instance, a node may start the consensus protocol, and while it is running, create new transactions and validate old transactions.
-By the time the consensus protocol is done, the new CP block is added to whatever the state that the chain is in.
-It is not necessary to lock the chain while the consensus protocol is running and then unlock it afterwards.
-
-\section{Model and assumptions}
+\subsection{Model and assumptions}
 \label{sec:model-assumptions}
@@
 
-\section{Extended TrustChain}
+\subsection{Extended TrustChain}
 The primary data structure used in our system is the Extended TrustChain.
@@
 
-% The length of the fragment is constrained by $L$,
-% namely $\forall t |\textsf{fragment}(t)| \le L$.
-% The purpose to prevent spam and encourage nodes to create more CP blocks.
-% $L$ should be sufficiently high so that busy nodes are not hindered by it.
-
-\section{Consensus protocol}
-\label{sec:cons-protocol}
-
-Our scalable consensus protocol runs on top of Extended TrustChain.
-It uses an unmodified asynchronous subset consensus (ACS) algorithm as the key building block.
-The objectives of the protocol are to
-    allow honest nodes always make progress (in the form of creating new CP blocks),
-    compute correct consensus result in every round
-    and have an unbiased election of facilitators.
-Concretely, we define the necessary properties as follows.
-\begin{definition}
-\label{def:consensus}
-\textbf{\emph{Properties of the consensus protocol}}
-
-$\forall r \in \mathbb{N}$, the following properties must hold.
-\begin{enumerate}
-    \item \emph{Agreement}:
-        If one correct node outputs a set of facilitators $\F_r$,
-        then every node outputs $\F_r$
-    \item \emph{Validity}:
-        If any correct node outputs $\F_r$, then 
-            \begin{enumerate}
-                \item $|\C_r| \ge N - t$ and each CP in $\C_r$ belong to a different node\footnote{
-                $\C_r$ is a tuple but we abuse the notation here by writing $|\C_r|$ to mean the number of CP blocks in the second element of $\C_r$.}.
-                \item $\F_r$ must contain at least $n - t$ honest nodes and
-                \item $|\F_r| = n$.
-            \end{enumerate}
-    \item \emph{Fairness}:
-        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
-    \item \emph{Termination}:
-        Every correct node eventually outputs some $\F_r$.
-\end{enumerate}
-\end{definition}
-These properties may look like Byzantine consensus properties (which we describe next in \Cref{sec:acs-background}) but there are subtle differences.
-Firstly, they are properties for every node in the network and not just the facilitators.
-Secondly, they must be satisfied for all rounds because the whole system falls apart if one of the property cannot be satisfied in one of the rounds.
-
-Before describing the protocol in detail,
-we take a brief detour to give background on the ACS algorithm as it is the primary building block.
-Then we move on to describe the two phases of our consensus protocol---bootstrap phase and consensus phase.
-
-\subsection{Background on asynchronous subset consensus}
-\label{sec:acs-background}
-
-The best way to explain ACS is to contrast it with the typical Byzantine consensus.
-We adapt the description from~\cite[Chapter 17]{podc}.
-%The Byzantine consensus problem is a classical problem is distributed systems literature.
-%It is first described by Leslie Lamport as the Byzantine general's problem in 1982~\cite{lamport1982byzantine}.
-%Byzantine consensus is described as follows (adapted from ).
-\begin{definition}
-\textbf{\emph{Byzantine consensus}}
-
-There are $n$ nodes, of which at most $t$ might experience Byzantine fault.
-Node $i$ starts with an input value $v_i$.
-The nodes must decide for one of  those values, satisfying the following.
-\begin{enumerate}
-    \item \emph{Agreement:}
-        If a correct node outputs $v$, then every node outputs $v$.
-    \item \emph{Validity:}
-        The decision value must be the input value of a node.
-    \item \emph{Termination:}
-        All correct nodes terminate in finite time.
-\end{enumerate}
-\end{definition}
-A node under Byzantine fault means that it can have arbitrary behaviour.
-For example not sending a message or colluding with other Byzantine nodes to undermine the entire system.
-Note that the decision is on a single value.
-This is in contrast to ACS which we describe next.
-
-ACS is an especially useful primitive for blockchain systems.
-It allows any party to propose a value and the result is the set union of all the proposed values by the majority.
-Concretely, ACS needs to satisfy the following properties (adapted from~\cite{miller2016honey}).
-\begin{definition}
-\label{def:acs}
-\textbf{\emph{Asynchronous subset consensus}}
-
-There are $n$ nodes, of which at most $t$ might experience Byzantine fault.
-Node $i$ starts with a non-empty set of input values $C_i$.
-The nodes must decide an output $C$, satisfying the following.
-\begin{enumerate}
-    \item \emph{Agreement:}
-        If a correct node outputs $C$, then every node outputs $C$.
-    \item \emph{Validity:}
-        If any correct node outputs a set $C$,
-        then $|C| \ge n - t$ and $C$ contains the input of at least $n - 2t$ nodes.
-    \item \emph{Totality:}
-        If $n - t$ nodes receive an input, then all correct nodes produce an output.
-\end{enumerate}
-\end{definition}
-
-ACS has the nice property of censorship resilience when compared to other consensus algorithms.
-For instance, Hyperledger and Tendermint uses a variant of Practical Byzantine Fault Tolerance (PBFT)~\cite{castro1999practical} as their consensus algorithm.
-In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectively filter transactions.
-In contrast, every party in ACS are involved in the proposal phase,
-and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
-Thus, if some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
-For a detailed description of ACS, we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.
-To understand the remainder of this work, the knowledge of~\Cref{def:acs} is sufficient.
-
-The main drawback with ACS and also Byzantine consensus algorithms is the high message complexity.
-Typically, such protocols have a message complexity of $O(n^2)$, where $n$ is the number of parties.
-Hence, it may work with a small number of nodes,
-but it is infeasible for blockchain systems where thousands of nodes are involved.
-
-% The literature on Byzantine consensus and atomic broadcast is rich, some noteable ones include TODO.
-% Thus in the ETC consensus protocol, we assume there exist an "off-the-shelf" Byzantine consensus algorithm which we can use
-% (in our implementation we use HoneyBadgerBFT and motivate our choice in TODO).
-
-% In our case, every node do not propose some arbitrary data, but a set of CP blocks.
-% Thus the result of the consensus is the set union of all the legitimate proposals.
-
-\subsection{Bootstrap phase}
+\subsection{Consensus Protocol}
+\subsubsection{Bootstrap phase}
 \label{sec:bootstrap}
@@
 
-\subsection{Consensus phase}
+\subsubsection{Consensus phase}
 \label{sec:consensus-phase}
@@
 
-\section{Transaction protocol}
+\subsection{Transaction Protocol}
 \label{sec:tx-protocol}
@@
 
-\section{Validation protocol}
+\subsection{Validation protocol}
 \label{sec:vd-protocol}
@@
 
-\subsection{Validity definition}
+\subsubsection{Validity definition}
 A transaction can be in one of three states in terms of validity---\emph{valid}, \emph{invalid} and \emph{unknown}.
@@
 
-\begin{algorithmic}[1]
+\begin{algorithmic}
 
@@
 
-\subsection{Validation protocol}
+\subsubsection{Validation protocol}
 Our validation protocol,
@@
 
-\section{Design variations and tradeoffs}
+\subsection{Design variations and tradeoffs}
 \label{sec:tradeoffs}
@@
 
-\subsection{Using epidemic protocol to reduce communication cost}
+\subsubsection{Using epidemic protocol to reduce communication cost}
 One of the final steps in our consensus protocol is to broadcast the consensus result and signatures to every other node (\Cref{sec:consensus-phase}).
@@
 
-\subsection{Using timing assumption in the permissionless setting}
+\subsubsection{Using timing assumption in the permissionless setting}
 \label{sec:permissionless}
@@
 
-\subsection{Privacy preserving validation protocol using compact blocks}
+\subsubsection{Privacy preserving validation protocol using compact blocks}
 \label{sec:compact}
@@
 \begin{figure}
-    \includegraphics[width=0.7\textwidth]{compact}
+    \includegraphics[width=0.7\linewidth]{compact}
     \centering
@@
 
-\subsection{Optimising validation protocol using cached agreed fragments}
+\subsubsection{Optimising validation protocol using cached agreed fragments}
 \label{sec:caching}
@@
 
-\subsection{Global fork detection}
+\subsubsection{Global fork detection}
 The validation algorithm guarantees that there are no forks within a single agreed fragment.
@@
 $$\Pr[X > 0] = \sum_{k = 1}^{n}\binom{n}{k}p^k (1 - p)^{n - k}.$$
-
-% TODO?
-% \subsection{Secure txid generation using a commitment protocol}
-
-% subtlety when fork happens on CP block
-
-% Another assumption we made in this work is the existance of a PKI.
-% This can also be removed with the introduction of timing by storing the public keys the consensus results,
-% much like in Bitcoin.
-
-% \section{Universally Composable Framework}
-% \label{sec:uc-intro}
-
-% In order to analyse the security of a system, a formal notion of security is required.
-% For instance, what does it mean if an encryption algorithm is secure?
-% One may say it is secure if the adversay cannot learn any information about the plaintext from the ciphertext.
-% But what if the adversay has some background information, for instance she may know it is English.
-% Do we then still say the encryption algorithm is secure?
-% Goldwasser and Micali introduced the notion of semantic security~\cite{goldwasser1982probabilistic}.
-% That is, imagine two worlds, a real world and an ideal world.
-% In the real world, the adversay is given the ciphertext,
-% and in the ideal world the adversay has nothing.
-% Then the encryption algorithm is semantically secure if the amount of information that the adversay can learn in the real world is just as much as the ideal world.
-% While our description is informal, the notion of security can be formally captured in this way.
-
-% Security sensitive distributed systems such as secure multi-party computation and blockchain systems also require a formal notion of security so that they can be analysed.
-% Fortunately, the idea from semantic security can also be applied in a distributed setting.
-% In an ideal world, we create an ideal functionality $\Fideal$ that performs all the computation on behalf of every node in the network.
-% The nodes act as dummies and only relay messages between $\Fideal$ and the environment.
-% Thus $\Fideal$ essentially becomes the specification of the distributed protocol.
-% If the execution of the real world protocol is indistinguishable from the ideal emulation in the presence of some adversay,
-% then the real world protocol is secure as per the ideal specification.
-% This is in essence the Universally Composable (UC) framework.
-
-% We use the UC framework not only because it suits our needs, 
-% it is also the only framework used in modelling blockchain systems to the best of our knowledge
-% ~\cite{todo}TODO.
-% In the remainder of this section we give an overview of the UC framework.
-% A detailed treatment  can be found in~\cite{canetti2001universally}.
-
-% \subsection{Model of Computation}
-% The model of computation considered in the UC framework is the interactive Turing machine (ITM).
-% Specifically, an ITM is an extension of the Turing machine with externally writable tapes which are the following.
-% \begin{enumerate}
-% \item input tape---TODO
-% \item incoming communication tape---TODO
-% \item subroutine output tape---TODO
-% \end{enumerate}
-% ITM can be seen as a specification or an algorithm, a machine running an ITM is an ITM instance (ITI).
-% To communicate, an ITI can write on the externally writable tapes of other ITIs.
-% The writing ITI then pauses exeuction, the receiving ITI begins execution.
-% Consequently, only one ITI is running at any point in time.
-
-% \subsection{Simulation-based Security}
-% Simulation-based security, also known as ideal/real paradigm is a model for defining security.
-% The model consist of a set of , the environment $\E$, the adversay $\A$,
-% the protocol ITM $\Preal$ and zero or more ideal functionalities $\mathcal{F}_0, \mathcal{F}_1, \dots$.
-% The 
-% Other than the machines running the protocol under consideration, the model contains two extra entities,
-% the environment $\E$ and the adversary $\A$.
-% $\E$ can be seen as users of the protocol, it can only provide input and receive output from $\A$ and the protocol.
-
-% Control function TODO
-
-% \subsection{Universal Composability}
-
-% \section{Formal Specification}
-% \label{sec:formal-model}
-
-% The formal model follows the same structure as \ref{fig:todo}TODO.
-
-% \begin{figure}[h]
-% \begin{framed}
-% \small  % end with \normalsize
-
-% \textbf{Protocol}\,$\Preal$
-
-% The ETC protocol. On initialisation do the following.
-% \begin{itemize}
-% \item Generate public and private key pair, $pk$ and $sk$.
-% \item Set personal chain $C := \{genesis()\}$.
-% \item Set checkpoint buffer $\C' := \varnothing$.
-% \item Set the facilitators $F$ to the bootstrap nodes provided by $\E$.
-% \item Set the latest round $r_{\text{latest}} = 0$.
-% \end{itemize}
-
-% Run the protocol as specified below after initialisation.
-% \begin{itemize}
-
-% \item Upon (\texttt{consensus}, $D$) from $\F^r_\text{BFT}$,
-%     if $r_{\text{latest}} >= r$ then do nothing and pause,
-%     otherwise do $try_add_cp(D, r)$
-%     send (\texttt{propose})
-
-% \item Upon (\texttt{consensus}, $D$, $r$) from $p \in \P$,
-
-% \item Upon (\texttt{checkpoint}, $c$) from $p \in \P$,
-
-% \item Upon (\texttt{tx\_init}, $m$, $p$) from $\E$, 
-%     $C := C \cup \{new\_tx(m)\}$,
-%     send (\texttt{tx\_req}, latext TX in $C$) to $p$.
-
-% \item Upon (\texttt{tx\_req}, $m$) from $p \in \P$,
-%     $C := C \cup \{new\_tx(m)\}$,
-%     let $m'$ be the latest TX in $C$ and set $m$ to be the other half of $m'$,
-%     send (\texttt{tx\_resp}, $m'$) to $p$.
-%     Output the complete TX to $\E$.
-
-% \item Upon (\texttt{tx\_resp}, $m'$) from $p \in \P$,
-%     find the corresponding pair of $m'$ and name it $m$,
-%     add $m'$ as the other half of $m$.
-%     Output the complete TX to $\E$.
-
-% \item Upon (\texttt{vd\_init}, $s$) from $\E$,
-%     if the sequence number $s$ does not exist in $C$ or the block with $s$ does not have the other half then do nothing.
-%     Otherwise send (\texttt{vd\_req}, $s'$) to $p'$ where $s'$ is the sequence number of the other half and $p'$ is the counterparty.
-
-% \item Upon (\texttt{vd\_req}, $s'$) from $p \in \P$,
-%     if the sequence number $s'$ does not exist then do nothing.
-%     Otherwise send (\texttt{vd\_resp}, $agreed\_fragment(s')$) to $p$.
-
-% \item Upon (\texttt{vd\_resp}, $x$) from $p \in \P$,
-%     run $validate(x)$ and output result to $\E$.
-
-% \end{itemize}
-
-% \normalsize
-% \end{framed}
-% \end{figure}
-
-% \begin{framed}
-% \textbf{Functionality}\,$\F^r_\text{BFT}$
-
-% The ideal Byzantine consensus protocol, parameterised by the round number $r$.
-
-% \begin{itemize}
-% \item Upon (\texttt{propose}, $r$, $C$),
-% \item Upon (\texttt{fetch}),
-% \end{itemize}
-% \end{framed}
-
-% \subsection{Discussion}
-
-% \subsubsection{Global clock for synchronization}
-% Where there is no Byzantine corruption, we conjecture that our protocol runs in the asynchronous setting.
-% However, security of many Byzantine consensus algorithms, especially the one we adopted---HoneyBadgerBFT,
-% fall apart when there is dynamic corruption.
-% In order to enforce that the corrupted machines do not change when running an instance of some Byzantine consensus algorithm,
-% we introduce synchrony.
-
-% We make use of a global clock... TODO
-
-% Dynamic corruption only between different clock ticks is enforced by the control function.
-
-% \subsubsection{Static corruption versus dynamic corruption}
-% A subtlety exists when modelling the type of corruption.
-% In dynamic corruption\footnote{In \cite{canetti2001universally}, dynamic corruption is termed Byzantine corruption.},
-% $\A$ take full control of a number of machines and also learns all of their states.
-% The states include private keys.
-% Thus, using the dynamic corruption model from \cite{canetti2001universally} we cannot guarantee security as
-% $\A$ can corrupt different machines over time and eventually learn all the private keys..
-
-% First is to use static corruption, where the corrupted machines are fixed.
-% While this circumvents our problem, it is a much weaker model.
-% Alternatively, we modify the aforementioned dynamic corruption model and weaken the adversary's abilities.
-% In particular, we introduce that forgetful adversary that only remembers the state of the currently corrupted nodes and forget the state of the recovered nodes.

### 2017-08-08 paper/s4-implementation.tex
Give paper a structure

--- 2017-08-06 thesis/chapters/ch5-implementation.tex
@@
-\chapter{Implementation and experimental results}
-\label{ch:implementation}
-
-To build up the confidence of the horizontal scalability claim from our analytical results,
-we must verify the same property experimentally.
-Hence, we give an open source prototype implementation of our system.
-Our implementation runs on a network with over 1000 node.
-Each node sends transactions to each other autonomously, 
-which allows us to evaluate our analytical claims.
-
-We begin this chapter with a description of the implementation in \Cref{sec:implementation}.
-Then, we move on to describing our experimental setup in \Cref{sec:experimental-setup} which includes a discussion on the system parameters and the physical infrastructure which we use.
-Finally, \Cref{sec:comms-cost-experiment} and onwards follows the same structure as the Performance analysis (\Cref{sec:performance-analysis}),
-where we analyse the performance experimentally rather than analytically and compare both results.
-
 \section{Implementation}
@@
 
-The structure of the implementation is primarily made up of three modules---\texttt{acs}, \texttt{trustchain} and \texttt{node}.
-\texttt{acs}, as its name suggests, implements ACS.
-ACS uses erasure code in one of its sub-protocols (reliable broadcast).
-Thus we use the liberasurecode\footnote{\url{https://github.com/openstack/liberasurecode}} library for its Reed-Solomon error correcting code functionality.
-An implementation detail is that liberasurecode cannot create more than 32 code blocks\footnote{
-  The 32 code blocks limitation is hardcoded in the source file, see
-  \url{https://github.com/openstack/liberasurecode/blob/0794b31c623e4cede76d66be730719d24debcca9/include/erasurecode/erasurecode.h\#L35}
-}, we discuss the effect of this in~\Cref{sec:experimental-setup}.
-The \texttt{acs} module provides a small interface to the caller to start and stop the consensus process and also retrieve results.
-The \texttt{trustchain} module implements the Extended TrustChain data structure.
-It also provides the essential algorithm necessary to interact with Extended TrustChain such as 
-$\textsf{new\_tx}(\cdot)$, $\textsf{new\_cp}(\cdot)$ and $\textsf{agreed\_fragment}(\cdot)$.
-Finally, the \texttt{node} module ties everything together.
-It implements the consensus protocol, the transaction protocol and the validation protocol.
-
 Every node keeps a persistent TCP connection with every other node.
@@
 
-\section{Experimental setup}
+\subsection{Experimental setup}
 \label{sec:experimental-setup}
@@
 transaction protocol and validation protocol---simultaneously and analyse the communication cost, the consensus duration and the global throughput.
-We investigate these properties under the following four parameters.
+We investigate these properties under the following parameters.
+2 TX/s
 \begin{enumerate}
-  \item The transaction rate $r_{\text{tx}}$ per node. This is not comparable to the others because it is fixed at 2 TX/s.
-        Nevertheless, it is a good value because, as we show later,
-        it hits a bottleneck in extreme cases which helps us understand the limitations of our design.
   \item The number of facilitators $n \in \{4, 8, \dots, 32\}$.
@@
         $N$ stops at 1200 is due to our physical setup, which we describe next.
-  \item The two modes of transaction.
-        The first mode is the \emph{fixed-neighbour} mode where nodes only transact with their immediate neighbour which is predetermined. 
-        This minimises the data volume per validated transaction because agreed fragment can be cached.
-        The second mode is in the other extreme which we call the \emph{random-neighbour} mode,
-        where every transaction is with a random node out of the $N$ nodes in the system.
-        It is unlikely that the agreed fragments can be reused.
 \end{enumerate}
@@
 we use a uniformly random transaction size sampled between 400 and 600 bytes.
-
-\section{Communication cost for the consensus protocol}
-\label{sec:comms-cost-experiment}
-
-The remainder of this chapter follows the same structure as the performance analysis in~\Cref{sec:performance-analysis}.
-We check our analytical results experimentally and demonstrate horizontal scalability.
-
-\begin{figure}[tb]
-  \centering
-  \makebox[\linewidth][c]{%
-    \begin{subfigure}[b]{.7\textwidth}
-      \includegraphics[width=\textwidth]{neighbour-fixed/round-communication-cost-vs-population}
-      \caption{Transactions are with fixed neighbours.}
-      \label{fig:round-comms-fixed}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.7\textwidth}
-      \includegraphics[width=\textwidth]{neighbour-random/round-communication-cost-vs-population}
-      \caption{Transactions are with random neighbours.}
-      \label{fig:round-comms-random}
-    \end{subfigure}%
-  }
-  \caption{Communication cost for the consensus protocol per round increases linearly with population size.
-  The error bars are larger for higher population size or higher number of facilitators is because rounds take longer thus they are repeated fewer times.}
-  \label{fig:round-comms}
-\end{figure}
-
-\begin{figure}[tb]
-  \centering
-  \makebox[\linewidth][c]{%
-    \begin{subfigure}[b]{.7\textwidth}
-      \includegraphics[width=\textwidth]{neighbour-fixed/round-duration-vs-population}
-      \caption{Transactions are with fixed neighbours.}
-      \label{fig:round-duration-fixed}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.7\textwidth}
-      \includegraphics[width=\textwidth]{neighbour-random/round-duration-vs-population}
-      \caption{Transactions are with random neighbours.}
-      \label{fig:round-duration-random}
-    \end{subfigure}%
-  }
-  \caption{Round duration does not increase linearly with the population size. 
-  This is likely due to the additional hashing operation required for larger consensus result.}
-  \label{fig:round-duration}
-\end{figure}
-
-\Cref{fig:round-comms} shows the relationship between the communication cost of the consensus protocol per round and the population size.
-The most important aspect is that these results show a linear increase.
-This reinforces our analytical result in~\Cref{sec:cons-complexity}.
-Note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
-the magnitude of the communication cost is similar.
-Both peak at about 100 MB.
-This is expected because the consensus protocol is decoupled from the transaction protocol and the validation protocol.
-Finally, the rate for which the communication cost increases is higher when the number of facilitators is higher.
-This is also expected because the ACS algorithm has an $n^2$ term in its communication complexity.
-
-We are also interested in how communication costs translate to time.
-Hence, for the same experiment, we record the duration in seconds and the result is shown in~\Cref{fig:round-duration}.
-Interestingly, the duration is not entirely linear.
-We attribute this behaviour to the extra time needed to hash the CP blocks in the consensus result to compute the luck value.
-Since if $N$ increases, every node must also perform more hash operations.
-These results do not conform to analytical result in~\Cref{sec:cons-duration}.
-Nevertheless, the difference is minor and there are ways to optimise the luck value computation.
-For example, the luck value can be computed by the facilitators and are sent with the consensus result.
-Then the non-facilitator nodes simply use it if there are enough signatures.
-
-\section{Communication cost for transaction and validation protocols}
-
-\begin{figure}[h]
-  \centering
-  \makebox[\linewidth][c]{%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-fixed/tx-communication-cost-vs-population}
-    \caption{Transactions are with fixed neighbours.}
-    \label{fig:tx-comms-fixed}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-random/tx-communication-cost-vs-population}
-    \caption{Transactions are with random neighbours.}
-    \label{fig:tx-comms-random}
-  \end{subfigure}
-  }
-  \caption{Communication cost per verified transaction has similar (near linear) trend as~\Cref{fig:round-duration}.
-  Fluctuation for the fixed-neighbour mode exists because the cache mechanism is unpredictable.}
-  \label{fig:tx-comms}
-\end{figure}
-
-We argued that the communication cost per verified transaction is of $O(N)$ (\Cref{sec:communication-cost-for-tx}).
-To verify the argument, we plot the relationship between the communication cost of for every validated transaction and population size in~\Cref{fig:tx-comms}.
-We observe a near linear relationship, which is due to the near linear relationship of the communication duration mentioned before.
-Again, we believe the difference is minor and it is possible to remove the extra overhead.
-
-More interestingly, there is a large difference in communication cost between the two modes of transaction.
-When transacting with only one neighbour, the communication cost is low because only one agreed fragment needs to be communicated for every round in order to validate all transactions of that round.
-This is because agreed fragments are cached.
-On the other hand,
-if every node is transacting with a random node, then it is likely the case that one agreed fragment needs to be communicated for every transaction.
-Hence the communication cost we see in~\Cref{fig:tx-comms-random} is much higher than in~\Cref{fig:tx-comms-fixed}.
-
-Some fluctuations exist in~\Cref{fig:tx-comms-fixed}, this is due to our caching mechanism.
-We send validation requests at the same rate as transactions.
-Upon receiving a (remote) agreed fragment, the caching mechanism inspects all the transactions in the agreed fragment and attempts to verify as many as it can,
-rather than just the transaction in the original validation request.
-However, it may be the case that the agreed fragments arrive later than the validation request interval.
-Then it is possible to have sent two or more validation requests for some transactions in the same local agreed fragment.
-In this case, the remote would respond with two or more of the same agreed fragments,
-which results in extra (wasted) communication cost, and this is the source of the fluctuation seen in~\Cref{fig:tx-comms-fixed}.
-The result in~\Cref{fig:tx-comms-random} reinforces our argument.
-It is a lot more stable because for every transaction it is almost always the case that an agreed fragment is needed to validate it.
-
-\section{Linear global throughput}
-\begin{figure}[h]
-  \centering
-  \makebox[\linewidth][c]{%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-fixed/throughput-vs-population}
-    \caption{Transactions are with fixed neighbours.}
-    \label{fig:global-throughput-fixed}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-random/throughput-vs-population}
-    \caption{Transactions are with random neighbours.}
-    \label{fig:global-throughput-random}
-  \end{subfigure}
-  }
-  \caption{Global throughput increases as the population increases when every node transact at the same rate.
-  Using fixed neighbours results in a higher throughput because of the caching mechanism.}
-  \label{fig:global-throughput}
-\end{figure}
-
-Finally, the global throughput results are shown in~\Cref{fig:global-throughput}.
-Evidently, the throughput has a linear relationship with the population size.
-This result is a strong indication of the horizontal scalability which we aimed to achieve.
-It also matches with our analytical result.
-
-Note that the throughput decreases slightly as the number of facilitators increases.
-This is due to the additional communication cost for running ACS with a high number of facilitators.
-That is, if the network is congested then the nodes may not have enough bandwidth to send validation responses timely.
-
-For~\Cref{fig:global-throughput-fixed},
-the magnitude of our throughput may not be self-evident at first glance.
-Recall that we fixed our $r_\text{tx}$ to 2, but how is it possible to have around 4800 transactions per second for 1200 nodes (which is 4 TX/s)?
-This is due to the way validated transactions are calculated.
-Transactions are between two parties, hence if every node makes two transactions per second,
-every node also expects to receive two transactions per second.
-Hence, for every node, the TX blocks are created at 4 per second.
-Validation requests are sent at the same rate, which explains the magnitude.
-
-The difference in magnitude between \Cref{fig:global-throughput-fixed} and \Cref{fig:global-throughput-random} is caused by the caching mechanism mentioned earlier.
-If a new agreed fragment needs to be transmitted to validate every transaction then it puts a toll on the network infrastructure.
-The low transaction rate in~\Cref{fig:global-throughput-random} is caused by the fact that the network infrastructure cannot keep up with our demand.
-
-We demonstrate the bottleneck issue from a different perspective in~\Cref{fig:backlog}.
-The graph is plotted by counting the number of transactions and the number of validated transactions every 5 seconds for one node running in a network of 1200 nodes and 32 facilitators.
-In~\Cref{fig:backlog-fixed}, the number of validated transaction changes as a step function.
-This means that transactions are validated in bursts, and the validation protocol can ``keep up'' with the transactions.
-Note that the horizontal ``lines'' where no new validated transactions are made are on average 76 seconds, this matches with the round duration result in~\Cref{fig:round-duration-fixed}.
-On the other hand in~\Cref{fig:backlog-random}, the validation protocol clearly cannot ``keep up'' with the rate which the transactions are made.
-As a result, the global throughput is lower when transacting with random nodes than only with neighbours.
-
-\begin{figure}[tb]
-  \centering
-  \makebox[\linewidth][c]{%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-fixed/timeseries}
-    \caption{Transactions are with fixed neighbours.}
-    \label{fig:backlog-fixed}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-random/timeseries}
-    \caption{Transactions are with random neighbours.}
-    \label{fig:backlog-random}
-  \end{subfigure}
-  }
-  \caption{A limitation of our system is when nodes are transacting with many random nodes,
-  then the network cannot keep up with the large number of agreed fragments that need to be communicated between nodes.
-  Hence the slope of validated transactions grows at a slower rate in~\Cref{fig:backlog-random}.}
-  \label{fig:backlog}
-\end{figure}
-
-\section{Communication cost with varying number of facilitators}
-Up to this point we focused on the effect of communication cost, throughput and so on with respect to the population size.
-In this section, we consider a varying number of facilitators,
-which gives us an insight into our system performance may perform when the number of facilitators is higher than 32.
-
-\begin{figure}[tb]
-  \centering
-  \makebox[\linewidth][c]{%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-fixed/consensus-communication-cost-vs-facilitators}
-    \caption{Transactions are with fixed neighbours.}
-    \label{fig:facilitators-comms-fixed}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-random/consensus-communication-cost-vs-facilitators}
-    \caption{Transactions are with random neighbours.}
-    \label{fig:facilitators-comms-random}
-  \end{subfigure}
-  }
-  \caption{The communication cost of the consensus protocol increases polynomially with respect to the number of facilitators as expected from the ACS communication complexity.}
-  \label{fig:facilitators-comms}
-\end{figure}
-
-\Cref{fig:facilitators-comms} shows the communication cost of the consensus protocol for varying numbers of facilitators.
-There is strong evidence that the communication cost grows polynomially.
-We expect this because of there are polynomial terms in the ACS communication complexity---$O(n^2|v| + \lambda n^3 \log n)$.
-This also means that the rounds would take longer to complete and transactions would take longer to verify.
-On the other hand, we get better fault tolerance.
-Thus a linear increase in fault tolerance ($t$) costs a polynomial increase in communication cost.

### 2017-08-08 paper/s5-evaluation.tex
Give paper a structure

--- 2017-08-06 thesis/chapters/ch5-implementation.tex
@@
-\chapter{Implementation and experimental results}
-\label{ch:implementation}
+\section{Experimental Evalutation}
 
-To build up the confidence of the horizontal scalability claim from our analytical results,
-we must verify the same property experimentally.
-Hence, we give an open source prototype implementation of our system.
-Our implementation runs on a network with over 1000 node.
-Each node sends transactions to each other autonomously, 
-which allows us to evaluate our analytical claims.
-
-We begin this chapter with a description of the implementation in \Cref{sec:implementation}.
-Then, we move on to describing our experimental setup in \Cref{sec:experimental-setup} which includes a discussion on the system parameters and the physical infrastructure which we use.
-Finally, \Cref{sec:comms-cost-experiment} and onwards follows the same structure as the Performance analysis (\Cref{sec:performance-analysis}),
-where we analyse the performance experimentally rather than analytically and compare both results.
-
-\section{Implementation}
-\label{sec:implementation}
-
-The prototype implementation can be found on GitHub.
-\begin{displayquote}
-\url{https://github.com/kc1212/consensus-thesis-code}
-\end{displayquote}
-It implements the three protocols and the Extended TrustChain discussed in~\Cref{ch:model}.
-We also implement two optimisations---privacy preserving validation protocol using compact blocks (\Cref{sec:compact})
-and optimised validation protocol using cached agreed fragments (\Cref{sec:caching}).
-It is written in the event driven paradigm, using the Python programming language\footnote{\url{https://www.python.org/}}.
-We use the Twisted\footnote{\url{https://twistedmatrix.com/}} library for networking.
-
-The structure of the implementation is primarily made up of three modules---\texttt{acs}, \texttt{trustchain} and \texttt{node}.
-\texttt{acs}, as its name suggests, implements ACS.
-ACS uses erasure code in one of its sub-protocols (reliable broadcast).
-Thus we use the liberasurecode\footnote{\url{https://github.com/openstack/liberasurecode}} library for its Reed-Solomon error correcting code functionality.
-An implementation detail is that liberasurecode cannot create more than 32 code blocks\footnote{
-  The 32 code blocks limitation is hardcoded in the source file, see
-  \url{https://github.com/openstack/liberasurecode/blob/0794b31c623e4cede76d66be730719d24debcca9/include/erasurecode/erasurecode.h\#L35}
-}, we discuss the effect of this in~\Cref{sec:experimental-setup}.
-The \texttt{acs} module provides a small interface to the caller to start and stop the consensus process and also retrieve results.
-The \texttt{trustchain} module implements the Extended TrustChain data structure.
-It also provides the essential algorithm necessary to interact with Extended TrustChain such as 
-$\textsf{new\_tx}(\cdot)$, $\textsf{new\_cp}(\cdot)$ and $\textsf{agreed\_fragment}(\cdot)$.
-Finally, the \texttt{node} module ties everything together.
-It implements the consensus protocol, the transaction protocol and the validation protocol.
-
-Every node keeps a persistent TCP connection with every other node.
-This creates a fully connected network for our experiment.
-It is certainly not ideal in real world scenarios where nodes may have limited resources (e.g. sockets).
-But as a prototype, it is sufficient to run a network of over a thousand nodes and experiment with it.
-
-Finally, the cryptography primitives we use are SHA256 for hash functions and Ed25519 for digital signatures.
-Both of which are provided by libnacl~\footnote{\url{https://pypi.python.org/pypi/libnacl}}.
-
-\section{Experimental setup}
-\label{sec:experimental-setup}
-
-The goal of the experiment is to run the three protocols---consensus protocol,
-transaction protocol and validation protocol---simultaneously and analyse the communication cost, the consensus duration and the global throughput.
-We investigate these properties under the following four parameters.
-\begin{enumerate}
-  \item The transaction rate $r_{\text{tx}}$ per node. This is not comparable to the others because it is fixed at 2 TX/s.
-        Nevertheless, it is a good value because, as we show later,
-        it hits a bottleneck in extreme cases which helps us understand the limitations of our design.
-  \item The number of facilitators $n \in \{4, 8, \dots, 32\}$.
-        The maximum $n$ is 32 because the limitation in liberasurecode mentioned in \Cref{sec:implementation},
-        but our results give a good indication of how our system may function for larger numbers of $n$.
-  \item The population size $N \in \{200, 300, \dots, 1200\}$.
-        $N$ stops at 1200 is due to our physical setup, which we describe next.
-  \item The two modes of transaction.
-        The first mode is the \emph{fixed-neighbour} mode where nodes only transact with their immediate neighbour which is predetermined. 
-        This minimises the data volume per validated transaction because agreed fragment can be cached.
-        The second mode is in the other extreme which we call the \emph{random-neighbour} mode,
-        where every transaction is with a random node out of the $N$ nodes in the system.
-        It is unlikely that the agreed fragments can be reused.
-\end{enumerate}
-
-The experiment is run on the DAS-5\footnote{\url{https://www.cs.vu.nl/das5/}}.
-From now on, we use ``machines'' to refer to DAS-5 nodes and ``nodes'' to refer to a running instance in our system.
-On DAS-5 we use up to 30 machines, for each machine we use up to 40 nodes.
-This gives us the aforementioned 1200 number.
-With this setup, we cannot run more nodes because the every machine only has 65535 ports available (minus the reserved ones).
-But 40 nodes each need 1200 TCP connections which is 48000 TCP connections per machine and that is inching close to the limit.
-While it is possible to have more TCP connections per machine,
-but it requires additional network interface which is something we do not control on the DAS-5.
-Nevertheless, running the system with 1200 nodes gives a good indication of its scalability properties as we show later.
-
-To coordinate nodes on many different machines,
-we employ a discovery server to inform every node the IP addresses and port numbers of every other node.
-It is only run before the experiment and is not used during the experiment.
-
-Finally, since Bitcoin transactions are approximately 500 bytes~\cite{txsize},
-we use a uniformly random transaction size sampled between 400 and 600 bytes.
-
-\section{Communication cost for the consensus protocol}
-\label{sec:comms-cost-experiment}
-
-The remainder of this chapter follows the same structure as the performance analysis in~\Cref{sec:performance-analysis}.
-We check our analytical results experimentally and demonstrate horizontal scalability.
+\subsection{Communication cost of the consensus protocol}
 
@@
   \centering
-  \makebox[\linewidth][c]{%
-    \begin{subfigure}[b]{.7\textwidth}
-      \includegraphics[width=\textwidth]{neighbour-fixed/round-communication-cost-vs-population}
-      \caption{Transactions are with fixed neighbours.}
-      \label{fig:round-comms-fixed}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.7\textwidth}
-      \includegraphics[width=\textwidth]{neighbour-random/round-communication-cost-vs-population}
-      \caption{Transactions are with random neighbours.}
-      \label{fig:round-comms-random}
-    \end{subfigure}%
-  }
+  \includegraphics[width=\linewidth]{neighbour-random/round-communication-cost-vs-population}
   \caption{Communication cost for the consensus protocol per round increases linearly with population size.
@@
   \centering
-  \makebox[\linewidth][c]{%
-    \begin{subfigure}[b]{.7\textwidth}
-      \includegraphics[width=\textwidth]{neighbour-fixed/round-duration-vs-population}
-      \caption{Transactions are with fixed neighbours.}
-      \label{fig:round-duration-fixed}
-    \end{subfigure}%
-    \begin{subfigure}[b]{0.7\textwidth}
-      \includegraphics[width=\textwidth]{neighbour-random/round-duration-vs-population}
-      \caption{Transactions are with random neighbours.}
-      \label{fig:round-duration-random}
-    \end{subfigure}%
-  }
+  \includegraphics[width=\linewidth]{neighbour-random/round-duration-vs-population}
   \caption{Round duration does not increase linearly with the population size. 
@@
 
-\section{Communication cost for transaction and validation protocols}
-
+\subsection{Communication cost of transaction and validation}
 \begin{figure}[h]
   \centering
-  \makebox[\linewidth][c]{%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-fixed/tx-communication-cost-vs-population}
-    \caption{Transactions are with fixed neighbours.}
-    \label{fig:tx-comms-fixed}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-random/tx-communication-cost-vs-population}
-    \caption{Transactions are with random neighbours.}
-    \label{fig:tx-comms-random}
-  \end{subfigure}
-  }
+  \includegraphics[width=\linewidth]{neighbour-random/tx-communication-cost-vs-population}
   \caption{Communication cost per verified transaction has similar (near linear) trend as~\Cref{fig:round-duration}.
@@
 
-We argued that the communication cost per verified transaction is of $O(N)$ (\Cref{sec:communication-cost-for-tx}).
-To verify the argument, we plot the relationship between the communication cost of for every validated transaction and population size in~\Cref{fig:tx-comms}.
-We observe a near linear relationship, which is due to the near linear relationship of the communication duration mentioned before.
-Again, we believe the difference is minor and it is possible to remove the extra overhead.
+\subsection{Linear Global Throughput}
 
-More interestingly, there is a large difference in communication cost between the two modes of transaction.
-When transacting with only one neighbour, the communication cost is low because only one agreed fragment needs to be communicated for every round in order to validate all transactions of that round.
-This is because agreed fragments are cached.
-On the other hand,
-if every node is transacting with a random node, then it is likely the case that one agreed fragment needs to be communicated for every transaction.
-Hence the communication cost we see in~\Cref{fig:tx-comms-random} is much higher than in~\Cref{fig:tx-comms-fixed}.
-
-Some fluctuations exist in~\Cref{fig:tx-comms-fixed}, this is due to our caching mechanism.
-We send validation requests at the same rate as transactions.
-Upon receiving a (remote) agreed fragment, the caching mechanism inspects all the transactions in the agreed fragment and attempts to verify as many as it can,
-rather than just the transaction in the original validation request.
-However, it may be the case that the agreed fragments arrive later than the validation request interval.
-Then it is possible to have sent two or more validation requests for some transactions in the same local agreed fragment.
-In this case, the remote would respond with two or more of the same agreed fragments,
-which results in extra (wasted) communication cost, and this is the source of the fluctuation seen in~\Cref{fig:tx-comms-fixed}.
-The result in~\Cref{fig:tx-comms-random} reinforces our argument.
-It is a lot more stable because for every transaction it is almost always the case that an agreed fragment is needed to validate it.
-
-\section{Linear global throughput}
 \begin{figure}[h]
   \centering
-  \makebox[\linewidth][c]{%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-fixed/throughput-vs-population}
-    \caption{Transactions are with fixed neighbours.}
-    \label{fig:global-throughput-fixed}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-random/throughput-vs-population}
-    \caption{Transactions are with random neighbours.}
-    \label{fig:global-throughput-random}
-  \end{subfigure}
-  }
+  \includegraphics[width=\linewidth]{neighbour-random/throughput-vs-population}
   \caption{Global throughput increases as the population increases when every node transact at the same rate.
@@
 This result is a strong indication of the horizontal scalability which we aimed to achieve.
-It also matches with our analytical result.
-
-Note that the throughput decreases slightly as the number of facilitators increases.
-This is due to the additional communication cost for running ACS with a high number of facilitators.
-That is, if the network is congested then the nodes may not have enough bandwidth to send validation responses timely.
-
-For~\Cref{fig:global-throughput-fixed},
-the magnitude of our throughput may not be self-evident at first glance.
-Recall that we fixed our $r_\text{tx}$ to 2, but how is it possible to have around 4800 transactions per second for 1200 nodes (which is 4 TX/s)?
-This is due to the way validated transactions are calculated.
-Transactions are between two parties, hence if every node makes two transactions per second,
-every node also expects to receive two transactions per second.
-Hence, for every node, the TX blocks are created at 4 per second.
-Validation requests are sent at the same rate, which explains the magnitude.
-
-The difference in magnitude between \Cref{fig:global-throughput-fixed} and \Cref{fig:global-throughput-random} is caused by the caching mechanism mentioned earlier.
-If a new agreed fragment needs to be transmitted to validate every transaction then it puts a toll on the network infrastructure.
-The low transaction rate in~\Cref{fig:global-throughput-random} is caused by the fact that the network infrastructure cannot keep up with our demand.
-
-We demonstrate the bottleneck issue from a different perspective in~\Cref{fig:backlog}.
-The graph is plotted by counting the number of transactions and the number of validated transactions every 5 seconds for one node running in a network of 1200 nodes and 32 facilitators.
-In~\Cref{fig:backlog-fixed}, the number of validated transaction changes as a step function.
-This means that transactions are validated in bursts, and the validation protocol can ``keep up'' with the transactions.
-Note that the horizontal ``lines'' where no new validated transactions are made are on average 76 seconds, this matches with the round duration result in~\Cref{fig:round-duration-fixed}.
-On the other hand in~\Cref{fig:backlog-random}, the validation protocol clearly cannot ``keep up'' with the rate which the transactions are made.
-As a result, the global throughput is lower when transacting with random nodes than only with neighbours.
-
-\begin{figure}[tb]
-  \centering
-  \makebox[\linewidth][c]{%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-fixed/timeseries}
-    \caption{Transactions are with fixed neighbours.}
-    \label{fig:backlog-fixed}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-random/timeseries}
-    \caption{Transactions are with random neighbours.}
-    \label{fig:backlog-random}
-  \end{subfigure}
-  }
-  \caption{A limitation of our system is when nodes are transacting with many random nodes,
-  then the network cannot keep up with the large number of agreed fragments that need to be communicated between nodes.
-  Hence the slope of validated transactions grows at a slower rate in~\Cref{fig:backlog-random}.}
-  \label{fig:backlog}
-\end{figure}
-
-\section{Communication cost with varying number of facilitators}
-Up to this point we focused on the effect of communication cost, throughput and so on with respect to the population size.
-In this section, we consider a varying number of facilitators,
-which gives us an insight into our system performance may perform when the number of facilitators is higher than 32.
-
-\begin{figure}[tb]
-  \centering
-  \makebox[\linewidth][c]{%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-fixed/consensus-communication-cost-vs-facilitators}
-    \caption{Transactions are with fixed neighbours.}
-    \label{fig:facilitators-comms-fixed}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-random/consensus-communication-cost-vs-facilitators}
-    \caption{Transactions are with random neighbours.}
-    \label{fig:facilitators-comms-random}
-  \end{subfigure}
-  }
-  \caption{The communication cost of the consensus protocol increases polynomially with respect to the number of facilitators as expected from the ACS communication complexity.}
-  \label{fig:facilitators-comms}
-\end{figure}
-
-\Cref{fig:facilitators-comms} shows the communication cost of the consensus protocol for varying numbers of facilitators.
-There is strong evidence that the communication cost grows polynomially.
-We expect this because of there are polynomial terms in the ACS communication complexity---$O(n^2|v| + \lambda n^3 \log n)$.
-This also means that the rounds would take longer to complete and transactions would take longer to verify.
-On the other hand, we get better fault tolerance.
-Thus a linear increase in fault tolerance ($t$) costs a polynomial increase in communication cost.

### 2017-08-08 paper/s6-conclusion.tex
Give paper a structure

--- 2017-08-07 thesis/chapters/ch6-conclusion.tex
@@
-\chapter{Conclusion}
-\label{ch:conclusion}
+\section{Conclusion}
 
@@
 We hope to improve our design by building a concrete application on top of it.
-
-\section{Future work}
-While our system has excellent scalability properties, it is not without limitations.
-Much of it is from the fact that it does not have a concrete application.
-We do not attempt to prevent the Sybil attack, spam or DoS because the accuracy of the defence depends on the understanding of the application.
-For instance, without any application, it would be impossible to distinguish between a very active node from a spammer because the content of the transaction ($m$) carry no meaning.
-
-Our experiment is carried out in a somewhat idealised world (DAS-5).
-We hope to experiment our prototype in a more realistic environment.
-For example, DeterLab~\cite{mirkovic2012teaching} lets users simulate network delay,
-which is important for discovering issue that may only occur in hostile environments.
-To understand our system further, we hope to experiment with a real platform with real users such as Tribler~\cite{pouwelse2008tribler}.
-For instance, the TrustChain in Tribler stores the number of bytes uploaded and downloaded between users, but it has no global consensus.
-It would be useful to integrate our consensus protocol into Tribler to prevent misreporting attacks.
-More importantly, it gives us an opportunity to observe the characteristics of our system in the real world and how it evolves with new features or alterations.
-
-The fault tolerance property is adequate for permissioned systems where nodes have relatively more trust between each other.
-However, it is not adequate for permissionless systems,
-which usually aim to tolerate fault that involve a minority of the network,
-which can be just shy of 50\%.
-For future work, we propose the following two ways to improve fault tolerance.
-First, use a reputation mechanism to select facilitators rather than simply selecting them randomly so that faulty nodes are less likely to be selected.
-This technique of course heavily depends on the application,
-for P2P file sharing systems the reputation score may be the number of bytes contributed to the network.
-Second, instead of using ACS we use a proof-of-stake (PoS) scheme which means that nodes with more stake in the system get more vote.
-In an ideal PoS scheme, the system can tolerate any fault if the majority of the stake holders are honest.
-
-Finally, as mentioned in~\Cref{sec:permissionless}, we wish to evaluate our system (analytically and experimentally) in the permissionless settings.
-This would enable our protocol to be applied to a much wider range of applications.

### 2017-08-09 paper/s1-intro.tex
Write/fix paper

\section{Introduction}

% problem description
% research question
% contribution
% organisation

### 2017-08-09 paper/s2-related.tex
Write/fix paper

@@
 \section{Background and Related Work}
-% give background on ACS
-% give background on TrustChain?
 
-\subsection{Background on asynchronous subset consensus}
-\label{sec:acs-background}
-
-ACS is an especially useful primitive for blockchain systems.
-It allows any party to propose a value and the result is the set union of all the proposed values by the majority.
-Concretely, ACS needs to satisfy the following properties (adapted from~\cite{miller2016honey}).
-\begin{definition}
-\label{def:acs}
-\textbf{\emph{Asynchronous subset consensus}}
-
-There are $n$ nodes, of which at most $t$ might experience Byzantine fault.
-Node $i$ starts with a non-empty set of input values $C_i$.
-The nodes must decide an output $C$, satisfying the following.
-\begin{enumerate}
-    \item \emph{Agreement:}
-        If a correct node outputs $C$, then every node outputs $C$.
-    \item \emph{Validity:}
-        If any correct node outputs a set $C$,
-        then $|C| \ge n - t$ and $C$ contains the input of at least $n - 2t$ nodes.
-    \item \emph{Totality:}
-        If $n - t$ nodes receive an input, then all correct nodes produce an output.
-\end{enumerate}
-\end{definition}
-
-ACS has the nice property of censorship resilience when compared to other consensus algorithms.
-For instance, Hyperledger and Tendermint uses a variant of Practical Byzantine Fault Tolerance (PBFT)~\cite{castro1999practical} as their consensus algorithm.
-In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectively filter transactions.
-In contrast, every party in ACS are involved in the proposal phase,
-and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
-Thus, if some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
-For a detailed description of ACS, we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.
-To understand the remainder of this work, the knowledge of~\Cref{def:acs} is sufficient.
-
-The main drawback with ACS and also Byzantine consensus algorithms is the high message complexity.
-Typically, such protocols have a message complexity of $O(n^2)$, where $n$ is the number of parties.
-Hence, it may work with a small number of nodes,
-but it is infeasible for blockchain systems where thousands of nodes are involved.
-
-% The literature on Byzantine consensus and atomic broadcast is rich, some noteable ones include TODO.
-% Thus in the ETC consensus protocol, we assume there exist an "off-the-shelf" Byzantine consensus algorithm which we can use
-% (in our implementation we use HoneyBadgerBFT and motivate our choice in TODO).
-
-% In our case, every node do not propose some arbitrary data, but a set of CP blocks.
-% Thus the result of the consensus is the set union of all the legitimate proposals.
+Making early blockchain systems scalable has received much attention from both the academia and the industry in recent years.
+This section gives an mini taxonomy of the state-of-the-art, categorised by the scalability approach.
+At the end of this section,
+we describe the important--asynchronous subset consensus (ACS),
+which is the key building block of our work.
 
@@
 
-Off-chain transaction systems are implemented using multi-signature addresses~\cite{bitcoinmultisig} and hashed time-locked Bitcoin contracts~\cite{bitcointimelock}.
-In the simplest case, 
-if two parties wish to make transactions,
-they open a payment channel in the form of a multi-signature Bitcoin address (for two parties it would be a 2-of-2 signature address).
-Each of the party must also deposit some Bitcoins in the multi-signature address,
-this is called the opening transaction.
-Both parties keep the channel state which is not broadcasted to the network.
-The state is updated when the two parties make new transactions.
-The protocol disincentivises the parties from broadcasting old channel states.
-If this occurs, the counterparty can drain all the Bitcoins in the channel.
-To close the channel, the parties simply broadcast the latest net settlement to the Bitcoin network,
-this is called the closing transaction.
-Effectively, only two transactions need to be broadcasted to the Bitcoin network---the opening and the closing transaction,
-even if the two parties made thousands or millions of transactions in between.
-The two-party scheme can be extended to a network of channels,
-which allows two parties to make off-chain transactions without an open channel as long as they are connected by nodes that do.
-
-The advantage of such systems is that they act as add-ons to Bitcoin wish already has a large number of users.
-Thus, if enough of the network wish for it (by setting a new block version),
-then a large number of users will instantly benefit from it.
-Some implementations already exist\footnote{For example \url{https://github.com/lightningnetwork/lnd}.},
-but at the time of writing SegWit is not activated yet which is a prerequisite for Lightning Network.
-
-On the other hand, off-chain transactions also suffer from the problem of Bitcoin.
+Off-chain transactions promises significant scalability improvements, but it suffer from the problem of Bitcoin.
 Namely, proof-of-work consumes an unreasonable amount of power.
@@
 
-\subsection{Blockchain without global consensus}
+\subsection{Asynchronous subset consensus}
+\label{sec:acs-background}
 
-Tangle~\cite{tangle}, Corda~\cite{corda} and the original TrustChain~\cite{multichain} do not use global consensus at all.
-In this scheme, every node has their own chain.
-Transactions between nodes are recorded on their respective chains.
-This effectively results in a directed acyclic graph.
-By avoiding global consensus, this approach achieves extremely desirable scalability properties similar to BitTorrent~\cite{cohen2003incentives}.
+ACS is an especially useful primitive for blockchain systems.
+It allows any party to propose a value and the result is the set union of all the proposed values by the majority.
+Concretely, ACS needs to satisfy the following properties (adapted from~\cite{miller2016honey}).
+\begin{definition}
+\label{def:acs}
+\textbf{\emph{Asynchronous subset consensus}}
 
-On the other hand, the application and security are limited.
-A malicious node can easily lie to other nodes regarding its own chain by simply presenting one version of reality to a set of nodes and another version to a different set.
-Thus the network will have a conflicting view on the state of the malicious node.
-Applications such as digital cash can not be implemented on top of such systems because consistency is needed in solving the double spending problem.
+There are $n$ nodes, of which at most $t$ might experience Byzantine fault.
+Node $i$ starts with a non-empty set of input values $C_i$.
+The nodes must decide an output $C$, satisfying the following.
+\begin{enumerate}
+    \item \emph{Agreement:}
+        If a correct node outputs $C$, then every node outputs $C$.
+    \item \emph{Validity:}
+        If any correct node outputs a set $C$,
+        then $|C| \ge n - t$ and $C$ contains the input of at least $n - 2t$ nodes.
+    \item \emph{Termination:}
+        If $n - t$ nodes receive an input, then all correct nodes produce an output.
+\end{enumerate}
+\end{definition}
+
+ACS has the nice property of censorship resilience when compared to other consensus algorithms.
+For instance, a leader is elected in PBFT~\cite{castro1999practical},
+if the leader is malicious but follows the protocol, then it can selectively filter transactions.
+In contrast, every party in ACS are involved in the proposal phase,
+and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
+Thus, if some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
+The total communication complexity of ACS is $O(n^2 |v| + \lambda n^3 \log(n))$,
+where $n$ is the number of nodes, $|v|$ is the message size and $\lambda$ is the security parameter.
+For a detailed description of ACS, we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.

### 2017-08-09 paper/s3-system.tex
Write/fix paper

@@
 
+\subsection{Overview and intuition}
+Unlike most blockchain systems, we do not use a global ledger.
+Instead, every node has their own hash chain.
+The nodes only store transactions (TX) that they are involved in on their hash chain.
+We introduce a special block called checkpoint (CP) block, which represents the state of a hash chain.
+The data structure is called Extended TrustChain, a visualisation can be seen on~\Cref{fig:trustchain-good-cp}.
+\begin{figure}[h]
+\centering
+\includegraphics[width=\linewidth]{trustchain-good-cp}
+\caption{Visualisation of the Extended TrustChain. $t_{u, i}$ represents a TX block on $u$'s chain with a sequence number of $i$.
+$c_{v, j}$ represents a CP block on $v$'s chain with a sequence number of $j$.}
+\label{fig:trustchain-good-cp}
+\end{figure}
+
+We randomly elect special nodes called facilitators in every round.
+The facilitators reach consensus on CP blocks and the consensus result is essentially a set of CP blocks of from (nearly) all the nodes.
+While this gives us consensus on a representation of the global state, it does not imply consensus on transactions.
+To this end, we introduce the validation protocol that allows any nodes to challenge any other node to produce a set of TX blocks that computes to the CP block in consensus.
+Hence we implicitly achieve consensus on transactions.
+
 \subsection{Model and assumptions}
 \label{sec:model-assumptions}
-
-This section and the ones following it give a technical treatment of what the content in System overview.
-For notational clarity, we use the following convention (adapted from~\cite{miller2016honey}) for most of this work.
-\begin{itemize}
-\item Lower case (e.g. $x$) denotes a scalar object or a tuple.
-\item Upper case (e.g. $X$) denotes a set or a constant.
-\item Sans serif (e.g. $\textsf{fn}(\cdot)$) denotes a function.
-\item Monospace (e.g. $\texttt{ack}$) denotes message type.
-\end{itemize}
-Further, we use $a || b$ to denote concatenation of the binary representations of $a$ and $b$.
 
@@
 }.
-
-We assume there exists a Public Key Infrastructure (PKI), and nodes are identified by their unique and permanent public key.
-Finally, we use the random oracle (RO) model, i.e. calls to the random oracle are denoted by $\textsf{H}: \{0, 1\}^* \rightarrow \{0, 1\}^\lambda$,
-where $\{0, 1\}^*$ denotes the space of finite binary strings and $\lambda$ is the security parameter.
-Under the RO model, the probability of successfully computing the inverse of the hash function is negligible with respect to $\lambda$~\cite{bellare1993random}.
+Further, we assume there exists a Public Key Infrastructure (PKI),
+and nodes are identified by their unique and permanent public key.
 
@@
 
-Our consensus protocol runs in rounds as discussed in \Cref{sec:system-overview}.
+Our consensus protocol runs in rounds.
 Every round is identified by a round number $r$, which is incremented on every new round.
@@
 Agreed enclosure is the same as enclosure with an extra constraint where the CP blocks must be in some consensus result $\C_r$.
-
 \end{definition}
-
-Note that the ``closest CP blocks'' property must also apply to agreed enclosure. 
-Suppose a chain is in the form
-$\{c_{i}, c_{i+1}, t_{i+2}, c_{i+3}\}$\footnote{Usually the notation is of the form $c_{u, i}$, but the node identity is not important here so we simplify it to $c_{i}$}
-and $c_{i}, c_{i+1}, c_{i+3}$ are in $\C_r, \C_{r+1}, \C_{r+3}$ respectively,
-then the agreed enclosure of $t_{i+2}$ is $\langle c_{i+1}, c_{i+3}\rangle$ and cannot be $\langle c_{i}, c_{i+3}\rangle$.
 
@@
 \subsection{Consensus Protocol}
+Our scalable consensus protocol runs on top of Extended TrustChain.
+It uses an unmodified asynchronous subset consensus (ACS) algorithm as the key building block.
+The objectives of the protocol are to
+    allow honest nodes always make progress (in the form of creating new CP blocks),
+    compute correct consensus result in every round
+    and have an unbiased election of facilitators.
+Concretely, we define the necessary properties as follows.
+\begin{definition}
+\label{def:consensus}
+\textbf{\emph{Properties of the consensus protocol}}
+
+$\forall r \in \mathbb{N}$, the following properties must hold.
+\begin{enumerate}
+    \item \emph{Agreement}:
+        If one correct node outputs a set of facilitators $\F_r$,
+        then every node outputs $\F_r$
+    \item \emph{Validity}:
+        If any correct node outputs $\F_r$, then 
+            \begin{enumerate}
+                \item $|\C_r| \ge N - t$ and each CP in $\C_r$ belong to a different node\footnote{
+                $\C_r$ is a tuple but we abuse the notation here by writing $|\C_r|$ to mean the number of CP blocks in the second element of $\C_r$.}.
+                \item $\F_r$ must contain at least $n - t$ honest nodes and
+                \item $|\F_r| = n$.
+            \end{enumerate}
+    \item \emph{Fairness}:
+        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
+    \item \emph{Termination}:
+        Every correct node eventually outputs some $\F_r$.
+\end{enumerate}
+\end{definition}
+These properties are similar to Byzantine consensus properties but there are subtle differences.
+Firstly, they are properties for every node in the network and not just the facilitators.
+Secondly, they must be satisfied for all rounds because the whole system falls apart if one of the property cannot be satisfied in one of the rounds.
+
 \subsubsection{Bootstrap phase}
 \label{sec:bootstrap}
-Now we have all the necessary information to describe our consensus protocol.
-As with all distributed systems, there must be a bootstrap phase which sets up the system.
-
-The reader may notice from the system overview (\Cref{sec:system-overview}) that the facilitators are computed from the consensus result,
-but the consensus result is agreed by the facilitators.
-Thus we have a dependency cycle.
-The goal of the bootstrap phase is to give us a starting point in the cycle.
-
-To bootstrap, imagine that there is some bootstrap oracle, that initiates the code on every node.
+As with many distributed systems, there must be a bootstrap phase which sets up the system.
+Imagine that there is some bootstrap oracle, that initiates the code on every node.
 The code satisfies all the properties in \Cref{def:consensus}.
@@
 Continuing, when $\F_{r}$ reaches agreement on $\C_{r+1}$,
-they immediately broadcast two messages to all the nodes---
+they immediately broadcast\footnote{An alternative to using broadcasting is to use an epidemic protocl~\cite{eugster2004epidemic}, 
+as long as it can guarantee delivery.} two messages to all the nodes---
 first the consensus message $\langle \texttt{cons\_msg}, \C_{r+1} \rangle$,
@@
     \EndIf
-    \State
 
@@
 
-\subsubsection{Using epidemic protocol to reduce communication cost}
-One of the final steps in our consensus protocol is to broadcast the consensus result and signatures to every other node (\Cref{sec:consensus-phase}).
-While this guarantees delivery (since we assumed reliable channel), it is wasteful.
-For example, if every facilitator is honest, a node would receive $n$ consensus results which are identical when only one is necessary.
-Furthermore, realistic networks do not offer a broadcast primitive,
-i.e. firewalls may block incoming messages from certain nodes.
-
-An improvement is use an epidemic protocol~\cite{eugster2004epidemic} (also known as gossiping) instead of our broadcast approach.
-Typical epidemic protocols works as follows.
-Every node buffers every message it receives up to some buffer size $b$.
-Then it forwards the messages $t$ number of times.
-Every time the message is sent to $f$ random neighbours, $f$ is often called the fan-out.
-The upside of using epidemic protocol is that the communication cost is distributed more evenly between nodes.
-This is especially true with a lazy push approach where the node who just received a message would push the message ID (e.g. digest) to its $f$ random neighbours,
-and only push the full message if the neighbour explicitly requests the message~\cite{leitao2007epidemic}.
-With this, nodes typically only need to receive one consensus result message instead of $n$.
-
-A down side of an epidemic protocol is that it usually takes $O(\log N)$ time to infect the whole network,
-whereas broadcasting uses constant time.
-Another downside of some epidemic protocols (e.g. eager push) is that it is difficult to guarantee delivery.
-It is especially true when the parameters are not chosen correctly in a network that is only partially connected (but every node is nevertheless reachable).
-If the delivery cannot be guaranteed, 
-then we cannot guarantee liveness in our consensus protocol because a future facilitator may miss the memo.
-Picking parameters are difficult in practice because the network configuration is unknown and the total number of nodes might also be unknown.
-
 \subsubsection{Using timing assumption in the permissionless setting}
@@
 
-The space saving of this approach depends on the size of the full blocks.
-If the full blocks are on average 500 bytes
-(which is the typical size of Bitcoin transactions ~\cite{txsize}),
-and the compact blocks are $32 + 32 + 8 = 72$ bytes
-(SHA256 digests are 32 bytes each and we use a 64-bit integer to represent the sequence number),
-then the saving in communication cost is 86\%.
-
-%Using compact blocks, it is possible to reduce network communication costs in the validation protocol but also protect transaction privacy.
-
 \subsubsection{Optimising validation protocol using cached agreed fragments}
@@
 then $A$ would refuse to make the transaction.
-The probability that an honest node catches out a malicious node is 
-$$p = \frac{f}{r},$$
-where $f$ is the number of bad agreed fragments and $r$ is the latest round number.
-If there are more nodes (say $n$) trying to make transactions with the malicious node,
-then the probability that the malicious node gets caught at least once (denoted by the random variable $X$) follows a binomial distribution, i.e.
-$$\Pr[X > 0] = \sum_{k = 1}^{n}\binom{n}{k}p^k (1 - p)^{n - k}.$$

### 2017-08-09 paper/s4-analysis.tex
Write/fix paper

\section{Correctness and fault tolerance analysis}
\label{sec:analysis}
Here we evaluate our system analytically to ensure the desired properties (\Cref{def:consensus} and \Cref{def:validation}) hold.
An informal argument is given in this section, we refer to~\cite{checo}??? for an indepth analysis.

\subsection{Correctness of the consensus protocol}
Recall that the consensus protocol properties (\Cref{def:consensus}) has
four conditions---agreement, validity, fairness and termination.
Agreement, validity and termination hold due to the fact that the following three facts
\begin{enumerate}
    \item The CP blocks sent to the facilitators are eventually delivered and then ACS eventually starts.
    \item Agreement, validity and termination holds for ACS.
    \item The consensus result and signatures are eventually disseminated to all the nodes.
\end{enumerate}

Fairnress holds if we model $\textsf{H}(\cdot)$ as a query to the random oracle.
Sorting nodes by the luck value $\textsf{H}(\C || pk)$ can be seen as a random permutation of a list of nodes.
Therefore every node has the same probability of becoming a facilitator.

\subsection{Correctness of the validation protocol}
Using the previous result,
we show that the agreement and validity properties (from~\Cref{def:validation}) hold for the validation protocol.

The validty property holds because we use $\textsf{get\_validity}(\cdot)$ in the validation protocol.
The agreement property holds due to the collision resistent property of cryptographically secure hash function.
Suppose two honest nodes decided on two different state, \emph{valid} and \emph{invalid} for the same transaction.
Due to the properties of the consensus protocol, the agreed enclosure for the transaction must be the same.
For that to happen, two agreed fragments must exist for the same transaction.
Recall that blocks form a hash chain.
So this cannot happen unless the hash function is not cryptographically secure.

Liveness, unfortunately does not hold in our model.
A malicious node can ask honestly when running the transaction protocol,
but then never respond to any validation requests.
Therefore there are transactions that can never be validated.
Nevertheless, the malicious node will be at an economical loss if it is not responsive because honest nodes are less likely to make contact with nodes that do not respond to validation messages.

A stronger version of the validity definition exists.
That is, if two honest nodes make a transaction,
then the transaction state is always valid in addition to our current validity definition.
Under our purely asynchronous model, we cannot guarantee this stronger version.
Since the adversary can delay any message for any amount of time,
it can make sure all \texttt{tx\_req} messages are delivered in a round later than the round which the message is sent.
Effectively, the transaction pair would always be in different rounds and the validation protocol would not output \emph{valid}.
We believe in a relaxed model, i.e. a weakly synchronous model, a stronger validity definition is possible.

\subsection{Fault tolerance}
Finally, we consider the effect when the number of adversaries is more than $t$.
This is useful because in practice it is difficult to guarantee that $t$ satisfies $n \ge 3t + 1$,
especially when $N$ is large.
Hence we are interested in the probability for this to happen under our facilitator election process.
The problem is modeled by a hypergeometric distribution,
where the random variable $X$ is the number of malicious nodes.
For our system to fail, we are interested in the following probability.
$$
\Pr[X \ge \lfloor \frac{n-1}{3} \rfloor + 1]
$$
Let $E[X] = n\alpha$ where $0 \le \alpha \le \lfloor \frac{n-1}{3} \rfloor/n < (\lfloor  \frac{n - 1}{3} \rfloor + 1) / n$.
Using the tail inequality found in~\cite{skala2013hypergeometric}
$$
\Pr[X \ge E[X] + \tau n] \le e^{-2\tau^2n},
$$
and setting 
$$
\tau = \frac{\lfloor \frac{n-1}{3} \rfloor + 1}{n} - \alpha,
$$
we arrive at the following bound
$$
\Pr[X \ge \lfloor \frac{n-1}{3} \rfloor + 1] \le e^{-2 \big(\frac{\lfloor  \frac{n - 1}{3} \rfloor + 1}{n} - \alpha \big)^2 n}.
$$

The bound is not tight, but it is useful for picking parameters.
If $n$ is known, then we can pick a $\alpha$ such that the probability becomes small.
On the other hand, if $\alpha$ is fixed, but small, we may increase $n$ to achieve the same.
Furthermore, hypergeometric distributions have light tails with ``faster-than-exponential fall-off''~\cite{skala2013hypergeometric},
the probability for picking more than $\lfloor \frac{n-1}{3} \rfloor$ malicious nodes drops of exponentially as the expected value moves away from $\lfloor \frac{n-1}{3} \rfloor$.
To put this into perspective,
suppose $n = 1000$ and $\alpha = 1/5$, i.e. a fifth of the nodes are malicious.
Then the probability to draw more black balls than the threshold is only $2.6 \times 10^{-16}$.

### 2017-08-09 paper/s5-evaluation.tex
Write/fix paper

@@
-\section{Experimental Evalutation}
+\section{Implementation and Evaluation}
 
-\subsection{Communication cost of the consensus protocol}
+The prototype implementation can be found on GitHub.
+\begin{displayquote}
+\url{https://github.com/kc1212/consensus-thesis-code}
+\end{displayquote}
+It implements the three protocols and the Extended TrustChain.
+We also implement two optimisations---privacy preserving validation protocol using compact blocks (\Cref{sec:compact})
+and optimised validation protocol using cached agreed fragments (\Cref{sec:caching}).
+It is written in the event driven paradigm, using the Python programming language\footnote{\url{https://www.python.org/}}.
+The cryptography primitives we use are SHA256 for hash functions and Ed25519 for digital signatures.
+Both of which are provided by libnacl~\footnote{\url{https://pypi.python.org/pypi/libnacl}}.
 
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=\linewidth]{neighbour-random/round-communication-cost-vs-population}
-  \caption{Communication cost for the consensus protocol per round increases linearly with population size.
-  The error bars are larger for higher population size or higher number of facilitators is because rounds take longer thus they are repeated fewer times.}
-  \label{fig:round-comms}
-\end{figure}
+\subsection{Experimental setup}
+\label{sec:experimental-setup}
 
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=\linewidth]{neighbour-random/round-duration-vs-population}
-  \caption{Round duration does not increase linearly with the population size. 
-  This is likely due to the additional hashing operation required for larger consensus result.}
-  \label{fig:round-duration}
-\end{figure}
+We run the experiment on the DAS-5\footnote{\url{https://www.cs.vu.nl/das5/}} with up to 1200 nodes.
+Every node makes a transaction with a random node twice every second.
+Since Bitcoin transactions are approximately 500 bytes~\cite{txsize},
+we use a uniformly random transaction size sampled between 400 and 600 bytes.
 
-\Cref{fig:round-comms} shows the relationship between the communication cost of the consensus protocol per round and the population size.
-The most important aspect is that these results show a linear increase.
-This reinforces our analytical result in~\Cref{sec:cons-complexity}.
-Note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
-the magnitude of the communication cost is similar.
-Both peak at about 100 MB.
-This is expected because the consensus protocol is decoupled from the transaction protocol and the validation protocol.
-Finally, the rate for which the communication cost increases is higher when the number of facilitators is higher.
-This is also expected because the ACS algorithm has an $n^2$ term in its communication complexity.
-
-We are also interested in how communication costs translate to time.
-Hence, for the same experiment, we record the duration in seconds and the result is shown in~\Cref{fig:round-duration}.
-Interestingly, the duration is not entirely linear.
-We attribute this behaviour to the extra time needed to hash the CP blocks in the consensus result to compute the luck value.
-Since if $N$ increases, every node must also perform more hash operations.
-These results do not conform to analytical result in~\Cref{sec:cons-duration}.
-Nevertheless, the difference is minor and there are ways to optimise the luck value computation.
-For example, the luck value can be computed by the facilitators and are sent with the consensus result.
-Then the non-facilitator nodes simply use it if there are enough signatures.
-
-\subsection{Communication cost of transaction and validation}
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{neighbour-random/tx-communication-cost-vs-population}
-  \caption{Communication cost per verified transaction has similar (near linear) trend as~\Cref{fig:round-duration}.
-  Fluctuation for the fixed-neighbour mode exists because the cache mechanism is unpredictable.}
-  \label{fig:tx-comms}
-\end{figure}
+To coordinate nodes on many different machines,
+we employ a discovery server to inform every node the IP addresses and port numbers of every other node.
+It is only run before the experiment and is not used during the experiment.
 
 \subsection{Linear Global Throughput}
+
+The global throughput results are shown in~\Cref{fig:global-throughput}.
+Evidently, the throughput has a linear relationship with the population size.
+This result is a strong indication of the horizontal scalability which we aimed to achieve.
 
@@
 \end{figure}
-
-Finally, the global throughput results are shown in~\Cref{fig:global-throughput}.
-Evidently, the throughput has a linear relationship with the population size.
-This result is a strong indication of the horizontal scalability which we aimed to achieve.

### 2017-08-09 paper/s6-conclusion.tex
Write/fix paper

@@
 this enables our horizontal scalability property.
-
-The research question we asked in~\Cref{ch:problem} is the following.
-\begin{displayquote}
-\emph{Is it possible to design a blockchain consensus protocol that is fault tolerant, scalable and can reach global consensus?}
-\end{displayquote}
-We answer it in the affirmative.
-Fault tolerance is guaranteed if $n \ge 3t + 1$ by using ACS as a building block.
-While $t$ may be small compared to the population size $N$,
-we show that the probability for the system to fail is low even when $n \ge 3t + 1$ does not hold as long as the proportion of malicious nodes is not close to a third of $N$.
-For example, if there the population size is 1000 and 20\% of the nodes are malicious,
-the probability for a round to potentially to fail is bounded below $2.6 \times 10^{-16}$.
-This probability bound would decrease as the population size increases.
-Horizontal scalability property is demonstrated both analytically and experimentally.
-Unlike sharding protocols, the property holds regardless of transaction characteristics and needs no parameter selection.
-Finally, we achieve global consensus on transactions via consensus on checkpoint blocks.
-
-This work is the first step in building a new paradigm for blockchain consensus protocol.
-It has the potential to efficiently cultivate trust on the internet in the presence of faults without a central authority.
-We hope to improve our design by building a concrete application on top of it.

### 2017-08-09 thesis/chapters/ch3-model.tex
Change code URL, etc.

@@
 \end{definition}
-These properties may look like Byzantine consensus properties (which we describe next in \Cref{sec:acs-background}) but there are subtle differences.
+These properties are similar to Byzantine consensus properties (which we describe next in \Cref{sec:acs-background}), but there are subtle differences.
 Firstly, they are properties for every node in the network and not just the facilitators.
@@
         then $|C| \ge n - t$ and $C$ contains the input of at least $n - 2t$ nodes.
-    \item \emph{Totality:}
+    \item \emph{Termination:}
         If $n - t$ nodes receive an input, then all correct nodes produce an output.
@@
 then $A$ would refuse to make the transaction.
+
 The probability that an honest node catches out a malicious node is 

### 2017-08-09 thesis/chapters/ch4-analysis.tex
Change code URL, etc.

@@
 Our last study considers the effect when the number of adversaries is more than $t$.
-This is useful because in practice it is difficult to guarantee that $t$ satisfied $n \ge 3t + 1$, especially when $N$ is large.
+This is useful because in practice it is difficult to guarantee that $t$ satisfies $n \ge 3t + 1$,
+especially when $N$ is large.
 Hence we are interested in the probability for this to happen under our facilitator election process.
@@
 We are interested in $\Pr[X \ge \lfloor \frac{n-1}{3} \rfloor + 1]$, so
-\begin{align*}
-\tau &= \frac{\lfloor \frac{n-1}{3} \rfloor + 1}{n} - \alpha
-\end{align*}
+$$
+\tau = \frac{\lfloor \frac{n-1}{3} \rfloor + 1}{n} - \alpha
+$$
 Putting $\tau$ back into the tail inequality we get the following bound.

### 2017-08-12 paper/s1-intro.tex
Write intro

@@
 % problem description
+Bitcoin is almost 10 years old and its market capitalisation is over \$60 billion USD and growing~\cite{bitcoinmarketcap}.
+We can be reasonably sure cryptocurrencies,
+even if their application is limited,
+are here to stay in the foreseeable future.
+Driven by the success of Bitcoin, we are seeing a renaissance of consensus research~\cite{miller2016honey, kogias2016enhancing, kokoris2017omniledger},
+where the primary focus is to improve the scalability of blockchain system.
+This is due to the consensus mechanism in early blockchain systems---proof-of-work (PoW).
+For example, Bitcoin can only do 7 transactions per second (TPS) at most~\cite{vukolic2015quest} prior to SegWit,
+which introduces a new block structure.
+While adjusting the block size and/or the block interval may increase TPS,
+it also leads to centralisation as larger blocks take longer to propagate in the network,
+giving miners that do not have a fast network a disadvantage~\cite{croman2016scaling}.
+In today's network, it is not possible to achieve more than 758 TPS
+if new blocks need to be propogated timely to 90\% of the network.
+
 % research question
+In this work, answer the following research question.
+\begin{displayquote}
+\emph{Is it possible to design a blockchain consensus protocol that is fault-tolerant and horizontally scalable?}
+\end{displayquote}
+A blockchain consensus protocol should be application neutural,
+for example PoW without the transaction logic.
+Our system should be Byzantine fault tolerant up to some threshold.
+The threshold may be made adjustable by trading off performance.
+Finally, we are interested in horizontal scalability as it enables ubiquitous use.
+That is, by adding new nodes into the network, the global transaction rate should increase proportionally.
+
 % contribution
+The key insight is to \emph{not} reach consensus using an existing consensus algorithm (a modified HoneyBadgerBFT~\cite{miller2016honey}) on transactions themselves but on special blocks called checkpoint blocks,
+such that transactions are nevertheless verifiable at a later stage by any node in the network.
+Our main contributions are the following.
+\begin{itemize}
+    \item We formally introduce a blockchain system---Checo\footnote{\emph{Che}ckpoint \emph{co}nsensus}.
+        It uses individual hash chains and checkpoints on every node to achieve
+        horizontally scalable for the first time.
+    \item We analyse Checo to ensure correctness as defined in our architecture.
+    \item We provide an implementation and then experiment with up to 1200 nodes,
+        our results show strong evidence of horizontal scalability.
+\end{itemize}
+
 % organisation
+In~\Cref{sec:related},
+we cover the current state-of-the-art and the necessary background.
+\Cref{sec:system} gives the formal system architecture.
+Next, we argue the correctness and fault tolerance properties of our system in~\Cref{sec:analysis}.
+We evaluate our system experimentally in~\Cref{sec:implementation}.
+Finally, we conclude our work in~\Cref{sec:conclusion}.

### 2017-08-13 paper/s1-intro.tex
Write paper

@@
 Bitcoin is almost 10 years old and its market capitalisation is over \$60 billion USD and growing~\cite{bitcoinmarketcap}.
-We can be reasonably sure cryptocurrencies,
-even if their application is limited,
+We can be reasonably sure that cryptocurrencies,
+even if their application is still somewhat limited,
 are here to stay in the foreseeable future.
@@
 where the primary focus is to improve the scalability of blockchain system.
-This is due to the consensus mechanism in early blockchain systems---proof-of-work (PoW).
-For example, Bitcoin can only do 7 transactions per second (TPS) at most~\cite{vukolic2015quest} prior to SegWit,
-which introduces a new block structure.
-While adjusting the block size and/or the block interval may increase TPS,
+This is due to the inefficiencies in the consensus mechanism in early blockchain systems---proof-of-work (PoW).
+For example, Bitcoin can only do 7 transactions per second (TPS) at most~\cite{vukolic2015quest}.
+While adjusting the block size (which Bitcoin has recently done via SegWit~\cite{segwit}) and/or the block interval may increase TPS,
 it also leads to centralisation as larger blocks take longer to propagate in the network,
@@
 In today's network, it is not possible to achieve more than 758 TPS
-if new blocks need to be propogated timely to 90\% of the network.
+if new blocks need to be propogated timely to 90\% of the network~\cite{croman2016scaling}.
+
+% related work
+Many approaches exist for improving the scalability of early blockchains.
+Off-chain transactions make use of the fact that if nodes make frequent transactions,
+then it is not necessary to store every transaction on the blockchain,
+only the net settlement is necessary.
+The best examples are Lightning Network~\cite{lightningnetwork} and Duplex Micropayment Channels~\cite{decker2015fast}.
+It promises significant scalability improvements,
+but complicates user experience and heavily relies on the semantics of transactions.
+Each node must deposit a suitable amount of Bitcoins into a multi-signature account.
+A low deposit would not allow large transactions.
+A high deposit locks the user from using much of her Bitcoins outside the channel.
+In addition, the user must proactively check whether the counterparty has broadcasted an old channel state so that the user does not lose Bitcoins.
+Payment channel, in theory, solves the scalability problem of early blockchain systems,
+but to the best of our knowledge, its exact scalability characteristics are not investigated.
+
+Another way to improve transaction rate is to use traditional Byzantine consensus algorithms such as PBFT~\cite{castro1999practical}.
+In essence, they contain a fixed set of nodes (sometimes called validating peer) that run a Byzantine consensus algorithm to decide on new blocks.
+This is often used in permissioned system where the validators must be predetermined, e.g. Hyperledger Fabric~\cite{cachin2016architecture}.
+Such systems can achieve much higher transaction rates, e.g. 10,000 TX/s if the number of validating peer is under 16 for PBFT~\cite[Section 5.2]{miller2016honey}.
+However, these systems do not scale, e.g. the transaction rate drops to under 5000 TX/s when the number of validating peer is 64~\cite[Section 5.2]{miller2016honey}.
+Moreover, the validating peers are predetermined which makes the system unsuitable for the open internet.
+
+Recent research has developed a class of hybrid systems which uses PoW for committee election,
+and Byzantine consensus algorithms to agree on transactions, e.g. SCP~\cite{luu2015scp}, ByzCoin~\cite{kogias2016enhancing} and Solidus~\cite{abraham2016solidus}.
+This design is primarily for permissionless system because the PoW leader election aspect prevents Sybils.
+This technique overcomes the early blockchain scalability issue by delegating the transaction validation to the Byzantine consensus protocol (e.g. PBFT in ByzCoin~\cite{kogias2016enhancing}).
+A major tradeoff of such systems is that they cannot guarantee correctness when there is a large number of malicious nodes (but less than a majority).
+For SCP, ByzCoin and Solidus, they all have some probability to elect more than $t$ Byzantine nodes into the committee,
+where $t$ is typically just under a third of the committee size (a lower bound of Byzantine consensus~\cite{pease1980reaching}).
+This problem is especially difficult to solve because the committee is always much smaller than the population size which has more than $t$ Byzantine nodes.
+Again, due to the fact that these systems must reach consensus on all transactions,
+none of them achieves horizontal scalability.
+
+Finally, a technique that does achieve horizontal scalability is sharding, e.g. Elastico~\cite{luu2016elastico} and OmniLedger~\cite{kokoris2017omniledger}.
+This is done by grouping nodes into multiple committees or shards of constant size.
+Nodes within a single shard run a Byzantine consensus algorithm to agree on a set of transactions that belong to that specific shard.
+The number of shards grows linearly with respect to the total computational power in the network, hence the transaction rate also grows linearly.
+The biggest limitation of sharding is that it is only optimal is transactions stay in the same shard.
+In fact, Elastico cannot atomically process inter-shard transactions.
+OmniLedger has an inter-shard transaction protocol but choosing a good shard size is difficult.
+A large shard size would make the system less scalable because the Byzantine consensus algorithm must be run by a large number of nodes.
+A small shard size would result in a large number of inter-shard transactions which also hinder scalability.
+The authors of OmniLedger noted that inter-shard transactions have significantly higher latency compared to the consensus protocol~\cite{kokoris2017omniledger}.
 
@@
 The threshold may be made adjustable by trading off performance.
-Finally, we are interested in horizontal scalability as it enables ubiquitous use.
+Finally, we are interested in horizontal scalability in the general case, as it enables ubiquitous use.
 That is, by adding new nodes into the network, the global transaction rate should increase proportionally.
@@
 \begin{itemize}
-    \item We formally introduce a blockchain system---Checo\footnote{\emph{Che}ckpoint \emph{co}nsensus}.
+    \item We formally introduce a blockchain system---\textsc{Checo}\footnote{\emph{Che}ckpoint \emph{co}nsensus}.
         It uses individual hash chains and checkpoints on every node to achieve
         horizontally scalable for the first time.
-    \item We analyse Checo to ensure correctness as defined in our architecture.
+    \item We analyse \textsc{Checo} to ensure correctness as defined in our architecture.
     \item We provide an implementation and then experiment with up to 1200 nodes,
@@
 % organisation
-In~\Cref{sec:related},
-we cover the current state-of-the-art and the necessary background.
+In~\Cref{sec:description},
+we give the problem description and our system model.
 \Cref{sec:system} gives the formal system architecture.

### 2017-08-13 paper/s2-problem.tex
Write paper

\section{Problem Description}
\label{sec:description}
We introduce the problem as a modified Byzantine consensus problem.
In our model, we consider $N$ nodes, $t$ of which are Byzantine.
Nodes in our system make transactions with each other.
Transactions can be in one of three states---\emph{valid}, \emph{invalid} and \emph{unknown}, defined by some function $\textsf{get\_validity}(\cdot)$.
We seek a protocol that defines these states given a transaction,
but also satisfies the following properties.
\begin{itemize}
    \item \emph{Agreement}:
        If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
        then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
    \item \emph{Validity}:
        The protocol outputs the correct result according to $\textsf{get\_validity}(\cdot).$
    \item \emph{Scalability}:
        If every node makes transactions at the same rate,
        then as $N$ increases,
        the global transaction rate should increase linearly w.r.t. $N$.
\end{itemize}

Note that the agreement and validity property are similar to what is often seen in Byzantine consensus problems but there are subtle differences and relaxations.
Firstly, the agreement property only holds if honest nodes do not output \emph{unknown}.
For example, it is ok if two honest nodes output \emph{valid} and \emph{unknown}, but it is not ok to output \emph{valid} and \emph{invalid}.
Secondly, the validity is a relaxation of what is typically seen in Byzantine consensus problems,
namely we do not make any guarantee on whether the transaction is \emph{valid} even if all nodes are honest,
only that it is correct according to the $\textsf{get\_validity}(\cdot)$ defined by the protocol.

Our problem is purposefully made to be application neutural,
i.e. there are no constraints on the semantics of transaction.
This is so that the protocol can act as a backbone to many applications.
Due to this fact, we also do not consider global fork prevention or detection,
as some application may not need such strong guarantees such as recording the number of bytes uploaded and downloaded in Tribler~\cite{pimotte, pouwelse2008tribler}.
On the other hand, we give an alternative construction that does have fork detection in~\Cref{sec:fork}.

\textbf{System model.}
We assume purely asynchronous channels with eventual delivery.
Thus in no stage of the protocol are we allowed to make timing assumptions.
The adversary has full control of the delivery schedule and the message ordering of all messages.
But they are not allowed to drop messages except for their own.
Reliability can be achieved in unreliable networks by resending messages or using some error correction code.

\textbf{Security assumptions.}
The malicious nodes are Byzantine and we use a static, round-adaptive corruption model.
That is, if a round has started, the corrupted nodes cannot change until the next round.
We assume there exists a Public Key Infrastructure (PKI), and nodes are identified by their unique and permanent public key.
Finally, we use the random oracle (RO) model, i.e. calls to the random oracle are denoted by $\textsf{H}: \{0, 1\}^* \rightarrow \{0, 1\}^\lambda$,
where $\{0, 1\}^*$ denotes the space of finite binary strings and $\lambda$ is the security parameter.
Under the RO model, the probability of successfully computing the inverse of the hash function is negligible with respect to $\lambda$~\cite{bellare1993random}.

### 2017-08-13 paper/s3-system.tex
Write intro
Write paper

@@
 \section{System Architecture}
+\label{sec:system}
 
-Our system consists of three protocols---the consensus protocol, the transaction protocol and the validation protocol.
-It is based on our prior work on TrustChain~\cite{multichain},
-where every node independently interacts with each other via their own blockchain\footnote{It is originally called MultiChain but the name has changed to TrustChain}.
-
-\subsection{Overview and intuition}
+To describe our system, we first give an informal overview and then move on to the formal description.
 Unlike most blockchain systems, we do not use a global ledger.
@@
 The nodes only store transactions (TX) that they are involved in on their hash chain.
+Transactions are stored in TX blocks and every block only contains one transaction.
 We introduce a special block called checkpoint (CP) block, which represents the state of a hash chain.
-The data structure is called Extended TrustChain, a visualisation can be seen on~\Cref{fig:trustchain-good-cp}.
+Such data structure is called Extended TrustChain,
+a visualisation can be seen on~\Cref{fig:trustchain-good-cp}.
 \begin{figure}[h]
@@
 
-We randomly elect special nodes called facilitators in every round.
-The facilitators reach consensus on CP blocks and the consensus result is essentially a set of CP blocks of from (nearly) all the nodes.
+We randomly elect $n$ special nodes called facilitators in every round.
+The facilitators reach consensus on CP blocks using an existing consensus algorithm---asynchronous subset consensus from HoneyBadgerBFT~\cite{miller2016honey}.
+The consensus result is essentially a set of CP blocks from (nearly) all the nodes.
 While this gives us consensus on a representation of the global state, it does not imply consensus on transactions.
-To this end, we introduce the validation protocol that allows any nodes to challenge any other node to produce a set of TX blocks that computes to the CP block in consensus.
+To this end, we introduce a validation protocol that allows any nodes to challenge any other node to produce a set of TX blocks that computes to the CP block in consensus.
 Hence we implicitly achieve consensus on transactions.
+Since the transaction and validation protocols only make point-to-point communication,
+we achieve horizontal scalability.
 
-\subsection{Model and assumptions}
-\label{sec:model-assumptions}
-
-We assume purely asynchronous channels with eventual delivery.
-Thus in no stage of our protocol do we make timing assumptions.
-The adversary has full control of the delivery schedule and the message ordering of all messages.
-But they are not allowed to drop messages except for their own\footnote{
-    Reliability can be achieved in unreliable networks by resending messages or using some error correction code.
-}.
-Further, we assume there exists a Public Key Infrastructure (PKI),
-and nodes are identified by their unique and permanent public key.
-
-% TODO big N does not follow our convention
-In our model, we consider $N$ nodes, which is the population size.
-$n$ of them are facilitators, $t$ out of $n$ are malicious and the inequality
-$n \ge 3t + 1$ must hold.
-This is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$~\cite{pease1980reaching}.
-Further, the inequality $N \ge n + t$ must also hold.
-This is due to our system design, which becomes clear in~\Cref{sec:consensus-phase}.
-
-Our threat model is as follows. 
-We use a restricted version of the adaptive corruption model.
-The first restriction is that corrupted node can only change across rounds.
-That is, if a round has started, the corrupted nodes cannot change until the next round.
-The second restriction is that the adversary, presumably controlling all the corrupted nodes, is forgetful.
-Namely, the adversary may learn the internal state such as the private key of a corrupted node,
-but if the node recovers, then the adversary must forget the private key.
-This is realistic because otherwise the adversary can eventually learn all the private keys and sabotage the system.
+The formal description is given next,
+it consist of one data structure---Extended TrustChain,
+and three protocols---consensus protocol, transaction protocol and validation protocol.
 
 \subsection{Extended TrustChain}
-The primary data structure used in our system is the Extended TrustChain.
 Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
@@
 \end{enumerate}
-The fact that we have no constraint on the content of $m$ is in alignment with our design goal---application neutrality.
+The fact that we have no constraint on the content of $m$ is in alignment with our problem description---application neutrality.
 
@@
 Every round is identified by a round number $r$, which is incremented on every new round.
-The consensus result is a tuple, i.e. $\C_r = \langle r, C \rangle$,
+The consensus result is a tuple, i.e.
+$$\C_r = \langle r, C \rangle,$$
 where $C$ is a set of CP blocks agreed by the facilitators of round $r$.
@@
 Our scalable consensus protocol runs on top of Extended TrustChain.
-It uses an unmodified asynchronous subset consensus (ACS) algorithm as the key building block.
+It uses an unmodified asynchronous subset consensus (ACS) algorithm as the key building block (we describe the reasoning later).
 The objectives of the protocol are to
@@
             \end{enumerate}
+    \item \emph{Termination}:
+        Every correct node eventually outputs some $\F_r$.
     \item \emph{Fairness}:
         Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
-    \item \emph{Termination}:
-        Every correct node eventually outputs some $\F_r$.
 \end{enumerate}
@@
 Firstly, they are properties for every node in the network and not just the facilitators.
-Secondly, they must be satisfied for all rounds because the whole system falls apart if one of the property cannot be satisfied in one of the rounds.
+Secondly, they must be satisfied for all rounds,
+because a failure in one round will lead to failures in sebsequent rounds.
 
@@
 For any future rounds, the consensus phase is used.
-
-In practice, the bootstrap oracle is most likely the software developer and some of the desired properties cannot be achieved.
-In particular, it is not possible to have the fairness property because it is unlikely that the developer knows the identity of every node in advance.
 
@@
 
-The core of the consensus phase is the ACS protocol.
-While any ACS protocol that satisfies the standard definition will work,
-we use a simplification of HoneyBadgerBFT~\cite{miller2016honey} as our ACS protocol
-because it is a consensus algorithm designed for blockchain systems.
+The core of the consensus phase is the ACS protocol,
+it is based on the construction described in HoneyBadgerBFT~\cite{miller2016honey}.
 We do not use the full HoneyBadgerBFT due to the following.
@@
 First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1}, r + 1)$ using \Cref{alg:new-cp}.
-Second, it computes the new facilitators using $\textsf{get\_facilitator}(\C_{r + 1}, n)$ using \Cref{alg:facilitator},
+Second, it computes the new facilitators using $\textsf{get\_facilitator}(\C_{r + 1}, n)$ (\Cref{alg:facilitator}),
 and updates its facilitator set to the result.
-This concludes the consensus phase and brings us back to the beginning of the consensus phase.
+This concludes the consensus phase and brings us back to the state at the beginning of the consensus phase,
+so a new round can be started.
 
@@
 
-Our protocol has some similarities with synchronizers~\cite[Chapter 10]{podc} because it is effectively a technique to introduce synchrony in an asynchronous environment.
-If we consider the facilitators as a collective authority, then it can be seen as a synchronizer that sends pulse messages (in the form of \texttt{cons\_msg} and \texttt{cons\_sig}) to indicate the start of a new clock pulse.
-Every node then sends a completion messages (in the form of \texttt{cp\_msg}) to the new collective authority to indicate that they are ready for the next pulse.
-
 \subsection{Transaction Protocol}
@@
 The TX protocol, shown in \Cref{alg:tx-proto}, is run by all nodes.
-It is also known as True Halves, first described by Veldhuisen~\cite[Chapter~3.2]{truehalves}.
+It is also known as True Halves, first described informally by Veldhuisen~\cite[Chapter~3.2]{truehalves}.
 Nodes that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
-$txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$.
-Then the initiator sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.
+$txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$,
+then sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.
 
@@
 At no time in \Cref{alg:new-tx} or \Cref{alg:tx-proto} do we need to hold the chain state and wait for some message to be delivered before committing a new block to the chain.
-This allows for high concurrency where we can call $\textsf{new\_tx}(\cdot)$ multiple times without waiting for the corresponding \texttt{tx\_resp} messages.
+This allows for a high level of concurrency where we can call $\textsf{new\_tx}(\cdot)$ multiple times without waiting for the corresponding \texttt{tx\_resp} messages.
 
@@
 A transaction can be in one of three states in terms of validity---\emph{valid}, \emph{invalid} and \emph{unknown}.
-Given a fragment $F_{v, j}$, the validity of the TX block $t_{u, i}$ and the agreed fragment of it is captured by the function $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ (\Cref{alg:get-validity}).
+Given a fragment $F_{v, j}$,
+the validity of the TX block $t_{u, i}$ with its corresponding fragment $F_{u, i}$ is captured by the function $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ (\Cref{alg:get-validity}).
 Note that $t_{u, i}$ and $F_{u, i}$ are assumed to be valid,
 otherwise the node calling the function would have no point of reference.
-This is not difficult to achieve.
-Typically the caller is $u$, so it knows its own TX block and the corresponding agreed fragment.
+This is not difficult to achieve,
+because typically the caller is $u$, so it knows its own TX block and the corresponding agreed fragment.
 If the caller is not $u$, it can always query for the agreed fragment that contains the transaction of interest from $u$.
 
-\Cref{alg:get-validity} works as follows.
-Before~\Cref{line:valid-fragment},
-we essentially check whether the fragment is the one that the verifier needs.
-If it is not, then the verifier cannot make any decision and return \emph{unknown}.
-This is likely to be the case for new transactions because the result of $\textsf{agreed\_fragment}(\cdot)$ would be $\bot$.
-The next two conditions check for tampering or missing blocks, if any of these misconducts are detected, then the TX is invalid.
+% \Cref{alg:get-validity} works as follows.
+% Before~\Cref{line:valid-fragment},
+% we essentially check whether the fragment is the one that the verifier needs.
+% If it is not, then the verifier cannot make any decision and return \emph{unknown}.
+% This is likely to be the case for new transactions because the result of $\textsf{agreed\_fragment}(\cdot)$ would be $\bot$.
+% The next two conditions check for tampering or missing blocks, if any of these misconducts are detected, then the TX is invalid.
 
@@
 Thus, in such cases, the whole transaction identified by its $txid$ is marked as invalid.
-
-Further, the caller of $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ is not necessarily 
-$u$\footnote{In practice, it often is because after completing the TX protocol the parties are incentivised to check that the counterparty ``did the right thing''.}
-Any node $w$ may call $\textsf{get\_validity}(t_u{u, i}, F_{u, i}, F_{v, j})$ as long as $w$ has an the necessary input parameters.
-$F_{u, i}$  and $t_{u, i}$ may be readily available if $w = u$.
-Otherwise, $w$ could query $u$ with the \texttt{vd\_req} message and then obtain the result from \texttt{vd\_resp}.
-This is the validation protocol which we describe next.
 
@@
 \label{sec:tradeoffs}
-
-Up to this point,
-we have discussed our protocol in the context of the model and assumptions defined in \Cref{sec:model-assumptions}.
 In this section, 
 we explore a few design variations which we can make, some of them require a relaxed version of our original model.
-They enable better performance, allow us to apply our design in the fully permissionless setting and improves privacy.
+They enable better performance and allow us to apply our design in the fully permissionless setting.
 
@@
 \label{sec:permissionless}
-Our model is purely asynchronous, where we make no timing assumptions anywhere in the protocol.
-However, in many applications it is often fine to make timing assumptions.
-For example, TCP relies on timeout for its retransmission and the \texttt{nLockTime} property in Bitcoin transactions makes the transaction unspendable until some time in the future (either Unix time or block height)~\cite{bitcoindevguide}.
 One limitation of our system is that we use the parameter $N$ in our algorithms, which makes it unsuitable for the permissionless environment where users can join and leave at will.
@@
 Consequently, this removes the need for a PKI because the collected CP blocks may be from nodes that nobody has seen in the past.
-However, choosing the parameter $D$ is difficult and depends on a number of factors such as the network condition, message size, and so on.
-Overestimating it would make agreed fragments much longer than usual, which increases communication costs for validation.
-Underestimating it would lead to unfairness where users with a poor internet connection will have a lower chance to be selected as a facilitator in the next round.
-Nevertheless, there is a significant gain for making the timing assumption and that is the ability to operate in the permissionless setting which we explain next.
 
+The new protocol handles churn as follows.
 Suppose a new node wish to join the network and the facilitators are known (this can be done with a public registry).
@@
 
-\subsubsection{Privacy preserving validation protocol using compact blocks}
-\label{sec:compact}
-Our approach already has privacy preserving features in comparison to early blockchain systems.
-That is, the transactions for each node are only revealed during the validation protocol.
-Hence if two nodes never directly or indirectly interact with each other,
-their transactions are never revealed.
-
-We can take our privacy-preserving property one step further by introducing another level of hash pointer indirection.
-The result is shown in~\Cref{fig:compact}.
-\begin{figure}
-    \includegraphics[width=0.7\linewidth]{compact}
-    \centering
-    \caption{The chain on the left represent direct chaining, where the digest in ``Prev'' is simply the digest of the previous block.
-    The chain on the right uses compact blocks, represented by the smaller squares.
-    Compact blocks also form a chain as before, but they each have a hash pointer to the full block, identified by ``Seq'' and ``Digest''.}
-    \label{fig:compact}
-\end{figure}
-
-Concretely, we introduce an additional block type,
-namely compact block.
-Such blocks only have three attributes,
-\begin{enumerate}
-\item Seq---the sequence number its corresponding block,
-\item Digest---the digest of its corresponding block, and
-\item Prev---the digest of the previous compact block.
-\end{enumerate}
-Each compact block has a corresponding full block (either a CP block or a TX block).
-The relationship is uniquely identified with Seq or Digest.
-Recall that our original validation protocol requires the nodes to send the full agreed fragment.
-With compact blocks, it is only necessary to send the compact version of the agreed fragment.
-The validation then proceeds in a similar fashion,
-provided that the pair of the to-be-validated TX block is known.
-
 \subsubsection{Optimising validation protocol using cached agreed fragments}
@@
 One more way to improve the efficiency of the validation protocol is to use a single agreed fragment to validate multiple transactions.
-Concretely, upon receiving an agreed fragment from node $A$,
+Concretely, for node $A$, upon receiving an agreed fragment from node $B$,
 rather than validating a single transaction,
-we attempt to validate all transactions with $A$ in the unknown state in that fragment.
+$A$ attempts to validate all transactions with performed with $B$ in the unknown state, in that fragment.
 
-For a node, the benefit of this technique is maximised when it only transacts with one other node.
+The benefit of this technique is maximised when a node only transacts with one other node.
 In this case, the communication of one fragment is sufficient to validate all transactions in that fragment.
@@
 
-\subsubsection{Global fork detection}
+\subsubsection{Total fork detection}
+\label{sec:fork}
 The validation algorithm guarantees that there are no forks within a single agreed fragment.
@@
 
-There are a variety of approaches to do global fork detection.
+We offer two approaches to do total fork detection.
 First and the easiest solution is to simply ask for the complete chain of the counterparty.
@@
 We use this approach in our prior work on Implicit Consensus~\cite{implicitconsensus}.
-It may sound inefficient at first, but nodes employ caching to minimise communication costs.
+Nodes employ caching to minimise communication costs, we call this effect spontaneous sharding.
 

### 2017-08-13 paper/s4-analysis.tex
Write paper

@@
 \label{sec:analysis}
-Here we evaluate our system analytically to ensure the desired properties (\Cref{def:consensus} and \Cref{def:validation}) hold.
-An informal argument is given in this section, we refer to~\cite{checo}??? for an indepth analysis.
+We evaluate our system analytically to ensure the desired properties (\Cref{def:consensus} and \Cref{def:validation}) hold.
+An informal argument is given in this section, we refer to~\cite[Chapter 4]{checo}??? for an indepth analysis.
 
@@
 Recall that the consensus protocol properties (\Cref{def:consensus}) has
-four conditions---agreement, validity, fairness and termination.
-Agreement, validity and termination hold due to the fact that the following three facts
-\begin{enumerate}
+four items---agreement, validity, termination and fairness.
+The first three hold due to the following facts.
+\begin{itemize}
     \item The CP blocks sent to the facilitators are eventually delivered and then ACS eventually starts.
-    \item Agreement, validity and termination holds for ACS.
-    \item The consensus result and signatures are eventually disseminated to all the nodes.
-\end{enumerate}
+    \item Agreement, validity and termination holds for ACS as they are the properties of ACS and are proven to hold in~\cite{miller2016honey}.
+    \item The consensus result and signatures are eventually disseminated to all the nodes,
+        so honest nodes must hold the same result as the honest facilitators.
+\end{itemize}
 
@@
 \subsection{Correctness of the validation protocol}
+\label{sec:correctness-of-validity}
 Using the previous result,
@@
 The validty property holds because we use $\textsf{get\_validity}(\cdot)$ in the validation protocol.
-The agreement property holds due to the collision resistent property of cryptographically secure hash function.
+The agreement property holds because we model $\textsf{H}(\cdot)$ as a query to a random oracle.
 Suppose two honest nodes decided on two different state, \emph{valid} and \emph{invalid} for the same transaction.
@@
 Recall that blocks form a hash chain.
-So this cannot happen unless the hash function is not cryptographically secure.
+So this cannot happen unless the adversary can compute the inverse of $\textsf{H}(\cdot)$ with high probability.
 
@@
 Therefore there are transactions that can never be validated.
-Nevertheless, the malicious node will be at an economical loss if it is not responsive because honest nodes are less likely to make contact with nodes that do not respond to validation messages.
+Nevertheless, the malicious node will be at an economical loss if it is not responsive because honest nodes are less likely to make contact with nodes that do not respond to validation requests.
 
@@
 
-The bound is not tight, but it is useful for picking parameters.
+The bound is not tight, but it is useful for picking parameters for a desired level of fault tolerance.
 If $n$ is known, then we can pick a $\alpha$ such that the probability becomes small.
 On the other hand, if $\alpha$ is fixed, but small, we may increase $n$ to achieve the same.
-Furthermore, hypergeometric distributions have light tails with ``faster-than-exponential fall-off''~\cite{skala2013hypergeometric},
-the probability for picking more than $\lfloor \frac{n-1}{3} \rfloor$ malicious nodes drops of exponentially as the expected value moves away from $\lfloor \frac{n-1}{3} \rfloor$.
+Furthermore, hypergeometric distributions have light tails with ``faster-than-exponential fall-off''~\cite{skala2013hypergeometric}.
+So the probability for picking more than $\lfloor \frac{n-1}{3} \rfloor$ malicious nodes drops of exponentially as the expected value moves away from $\lfloor \frac{n-1}{3} \rfloor$.
 To put this into perspective,
 suppose $n = 1000$ and $\alpha = 1/5$, i.e. a fifth of the nodes are malicious.
-Then the probability to draw more black balls than the threshold is only $2.6 \times 10^{-16}$.
+Then the probability to elect more malicious nodes than the threshold is only $2.6 \times 10^{-16}$.

### 2017-08-13 paper/s5-evaluation.tex
Write intro
Write paper

@@
 \section{Implementation and Evaluation}
+\label{sec:implementation}
+
+\begin{figure*}[t]
+  \centering
+  \makebox[\linewidth][c]{%
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\linewidth]{neighbour-fixed/throughput-vs-population}
+    \caption{Every node make transactions with a fixed node.}
+    \label{fig:global-throughput-fixed}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.5\linewidth}
+    \includegraphics[width=\linewidth]{neighbour-random/throughput-vs-population}
+    \caption{Every node make transactions with a random node.}
+    \label{fig:global-throughput-random}
+  \end{subfigure}
+  }
+  \caption{Global throughput increases as the population increases when every node transact at the same rate.
+  Making transactions with fixed nodes results in a higher throughput because of the caching mechanism.}
+  \label{fig:global-throughput}
+\end{figure*}
 
@@
 It implements the three protocols and the Extended TrustChain.
-We also implement two optimisations---privacy preserving validation protocol using compact blocks (\Cref{sec:compact})
-and optimised validation protocol using cached agreed fragments (\Cref{sec:caching}).
-It is written in the event driven paradigm, using the Python programming language\footnote{\url{https://www.python.org/}}.
+We also implement the caching optimisations discussed in \Cref{sec:caching}.
+It is written in the Python programming language\footnote{\url{https://www.python.org/}}.
 The cryptography primitives we use are SHA256 for hash functions and Ed25519 for digital signatures.
@@
 
-\subsection{Experimental setup}
-\label{sec:experimental-setup}
-
 We run the experiment on the DAS-5\footnote{\url{https://www.cs.vu.nl/das5/}} with up to 1200 nodes.
-Every node makes a transaction with a random node twice every second.
+Every node makes transactions at 2 per second.
 Since Bitcoin transactions are approximately 500 bytes~\cite{txsize},
@@
 
-\subsection{Linear Global Throughput}
+The global throughput results are shown in~\Cref{fig:global-throughput-fixed} and \Cref{fig:global-throughput-random}.
+We consider~\Cref{fig:global-throughput-fixed} as the ideal case,
+where nodes only make transactions with a fixed node.
+We consider \Cref{fig:global-throughput-random} as the worst case,
+where nodes make transactions with random nodes and the caching mechanism is unlikely to be used.
+The low transaction rate in~\Cref{fig:global-throughput-random} is caused by the fact that the network infrastructure cannot keep up with our demand.
+In practice, we do not expect such behaviour to occur as it is possible to cache agreed fragments.
 
-The global throughput results are shown in~\Cref{fig:global-throughput}.
-Evidently, the throughput has a linear relationship with the population size.
+For~\Cref{fig:global-throughput-fixed},
+the magnitude of our throughput may not be self-evident at first glance.
+Recall that we fixed the transaction rate to 2 TPS,
+but how is it possible to have around 4800 transactions per second for 1200 nodes (which is 4 TX/s)?
+This is due to the way validated transactions are calculated.
+Transactions are between two parties, hence if every node makes two transactions per second,
+every node also expects to receive two transactions per second.
+Hence, for every node, the TX blocks are created at 4 per second.
+Validation requests are sent at the same rate, which explains the magnitude.
+
+Overall, the throughput has a linear relationship with the population size.
 This result is a strong indication of the horizontal scalability which we aimed to achieve.
+For more experimental results, refer to~\cite[Chapter 5]{checo}???.
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{neighbour-random/throughput-vs-population}
-  \caption{Global throughput increases as the population increases when every node transact at the same rate.
-  Using fixed neighbours results in a higher throughput because of the caching mechanism.}
-  \label{fig:global-throughput}
-\end{figure}
+% The difference in magnitude between \Cref{fig:global-throughput-fixed} and \Cref{fig:global-throughput-random} is caused by the caching mechanism mentioned earlier.
+% If a new agreed fragment needs to be transmitted to validate every transaction then it puts a toll on the network infrastructure.
+% The low transaction rate in~\Cref{fig:global-throughput-random} is caused by the fact that the network infrastructure cannot keep up with our demand.
+% In practice, we do not expect such behaviour to occur as it is possible to cache agreed fragments.

### 2017-08-13 paper/s6-conclusion.tex
Write intro
Write paper

@@
 \section{Conclusion}
+\label{sec:conclusion}
 
-In this work we introduced an application neutral blockchain system which we call Checo.
+In this work we introduced an application neutral blockchain system which we call \textsc{Checo}.
 We add checkpoint block to the existing TrustChain data structure for use in our consensus protocol.
@@
 this enables our horizontal scalability property.
+
+We achieve the properties described in the Problem Description (\Cref{sec:description}).
+Namely, our protocol achieves agreement and validity as we argued in~\Cref{sec:correctness-of-validity}.
+Further, the horizontal scalability is demonstrated in~\Cref{sec:implementation},
+in the ideal case as well as the worst case.

### 2017-08-13 thesis/chapters/ch1-intro.tex
Give paper a structure
Style improvements in thesis

@@
 Parties against it argue that it would make mining more centralised because blocks take longer to propagate through the network.
-It also requires a hard fork (not backwards compatible), which risks consensus failure and devaluation.
-A recent empirical study by Croman et el.~\cite{croman2016scaling} has shown that increasing the block size may help.
-But given the bandwidth and latency constraints,
-it is not possible to have more than 758 transactions per second.
-Even worse, it may give some miner an unfair lead over others.
+It also requires a hard (non-backwards compatible) fork,
+which risks consensus failure and devaluation.
+A recent empirical study by Croman et el.~\cite{croman2016scaling} has shown that increasing the block size or the block interval may help.
+But given the bandwidth and latency constraints of today's network,
+it is not possible to have more than 758 transactions per second if messages need to be propagated timely to 90\% of the nodes.
+To achieve transaction rates higher than that, it will lead to unfairness for nodes that do not have a good network connection.
 Thus fundamental changes are necessary to run at the scale of centralised payment processors such as Visa,
@@
 but nevertheless verifiable at a later stage by any node in the network.
-We call our protocol Checo, derived from \emph{che}ckpoint \emph{co}nsensus.
+We call our protocol \textsf{Checo}, derived from \emph{che}ckpoint \emph{co}nsensus.
 Informally, our system works as follows.
@@
 By splitting this functionality we achieved horizontal scalability in the general case.
+That is, by adding new nodes into the network, the global transaction rate should increase proportionally.
 
@@
 \begin{itemize}
-    \item We formally introduce a blockchain system---Checo.
+    \item We formally introduce a blockchain system---\textsc{Checo}.
         It uses individual hash chains and checkpoints on every node to achieve
         horizontally scalable for the first time.
-    \item We formally analyse Checo to ensure correctness as defined in our architecture.
+    \item We formally analyse \textsc{Checo} to ensure correctness as defined in our architecture.
     \item We provide an implementation and then experiment with up to 1200 nodes,

### 2017-08-13 thesis/chapters/ch3-model.tex
Style improvements in thesis

@@
 Every round is identified by a round number $r$, which is incremented on every new round.
-The consensus result is a tuple, i.e. $\C_r = \langle r, C \rangle$,
+The consensus result is a tuple, i.e. 
+$$\C_r = \langle r, C \rangle,$$
 where $C$ is a set of CP blocks agreed by the facilitators of round $r$.
@@
 Upon receiving $\C_{r + 1}$ and at least $n - f$ valid signatures by some node $u$, $u$ performs two tasks.
-First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1}, r + 1)$ using \Cref{alg:new-cp}.
-Second, it computes the new facilitators using $\textsf{get\_facilitator}(\C_{r + 1}, n)$ using \Cref{alg:facilitator},
+First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1}, r + 1)$ (\Cref{alg:new-cp}).
+Second, it computes the new facilitators using $\textsf{get\_facilitator}(\C_{r + 1}, n)$ (\Cref{alg:facilitator}),
 and updates its facilitator set to the result.
-This concludes the consensus phase and brings us back to the beginning of the consensus phase.
+This concludes the consensus phase and brings us back to the state at the beginning of the consensus phase,
+so a new round can be started.
 
@@
 The TX protocol, shown in \Cref{alg:tx-proto}, is run by all nodes.
-It is also known as True Halves, first described by Veldhuisen~\cite[Chapter~3.2]{truehalves}.
+It is also known as True Halves, first described informally by Veldhuisen~\cite[Chapter~3.2]{truehalves}.
 Nodes that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
@@
 A transaction can be in one of three states in terms of validity---\emph{valid}, \emph{invalid} and \emph{unknown}.
-Given a fragment $F_{v, j}$, the validity of the TX block $t_{u, i}$ and the agreed fragment of it is captured by the function $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ (\Cref{alg:get-validity}).
+Given a fragment $F_{v, j}$, the validity of the TX block $t_{u, i}$ with its corresponding fragment $F_{u, i}$ is captured by the function $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ (\Cref{alg:get-validity}).
 Note that $t_{u, i}$ and $F_{u, i}$ are assumed to be valid,
 otherwise the node calling the function would have no point of reference.
-This is not difficult to achieve.
-Typically the caller is $u$, so it knows its own TX block and the corresponding agreed fragment.
+This is not difficult to achieve,
+because typically the caller is $u$, so it knows its own TX block and the corresponding agreed fragment.
 If the caller is not $u$, it can always query for the agreed fragment that contains the transaction of interest from $u$.
@@
 
-Further, the caller of $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ is not necessarily 
-$u$\footnote{In practice, it often is because after completing the TX protocol the parties are incentivised to check that the counterparty ``did the right thing''.}
-Any node $w$ may call $\textsf{get\_validity}(t_u{u, i}, F_{u, i}, F_{v, j})$ as long as $w$ has an the necessary input parameters.
-$F_{u, i}$  and $t_{u, i}$ may be readily available if $w = u$.
-Otherwise, $w$ could query $u$ with the \texttt{vd\_req} message and then obtain the result from \texttt{vd\_resp}.
-This is the validation protocol which we describe next.
+% Further, the caller of $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ is not necessarily 
+% $u$\footnote{In practice, it often is because after completing the TX protocol the parties are incentivised to check that the counterparty ``did the right thing''.}
+% Any node $w$ may call $\textsf{get\_validity}(t_u{u, i}, F_{u, i}, F_{v, j})$ as long as $w$ has an the necessary input parameters.
+% $F_{u, i}$  and $t_{u, i}$ may be readily available if $w = u$.
+% Otherwise, $w$ could query $u$ with the \texttt{vd\_req} message and then obtain the result from \texttt{vd\_resp}.
+% This is the validation protocol which we describe next.
 
@@
 One more way to improve the efficiency of the validation protocol is to use a single agreed fragment to validate multiple transactions.
-Concretely, upon receiving an agreed fragment from node $A$,
+Concretely, for node $A$, upon receiving an agreed fragment from node $B$,
 rather than validating a single transaction,
-we attempt to validate all transactions with $A$ in the unknown state in that fragment.
+$A$ attempts to validate all transactions performed with $B$ in the unknown state, in that fragment.
 
-For a node, the benefit of this technique is maximised when it only transacts with one other node.
+The benefit of this technique is maximised when a node only transacts with one other node.
 In this case, the communication of one fragment is sufficient to validate all transactions in that fragment.
@@
 
-\subsection{Global fork detection}
+\subsection{Total fork detection}
 The validation algorithm guarantees that there are no forks within a single agreed fragment.

### 2017-08-13 paper/s1-intro.tex
Spell check paper

@@
 In today's network, it is not possible to achieve more than 758 TPS
-if new blocks need to be propogated timely to 90\% of the network~\cite{croman2016scaling}.
+if new blocks need to be propagated timely to 90\% of the network~\cite{croman2016scaling}.
 
@@
 \end{displayquote}
-A blockchain consensus protocol should be application neutural,
+A blockchain consensus protocol should be application neutral,
 for example PoW without the transaction logic.

### 2017-08-13 paper/s2-problem.tex
Spell check paper

@@
 Firstly, the agreement property only holds if honest nodes do not output \emph{unknown}.
-For example, it is ok if two honest nodes output \emph{valid} and \emph{unknown}, but it is not ok to output \emph{valid} and \emph{invalid}.
+For example, it is OK if two honest nodes output \emph{valid} and \emph{unknown}, but it is not OK to output \emph{valid} and \emph{invalid}.
 Secondly, the validity is a relaxation of what is typically seen in Byzantine consensus problems,
@@
 
-Our problem is purposefully made to be application neutural,
+Our problem is purposefully made to be application neutral,
 i.e. there are no constraints on the semantics of transaction.

### 2017-08-13 paper/s3-system.tex
Spell check paper

@@
 Secondly, they must be satisfied for all rounds,
-because a failure in one round will lead to failures in sebsequent rounds.
+because a failure in one round will lead to failures in subsequent rounds.
 
@@
 Continuing, when $\F_{r}$ reaches agreement on $\C_{r+1}$,
-they immediately broadcast\footnote{An alternative to using broadcasting is to use an epidemic protocl~\cite{eugster2004epidemic}, 
+they immediately broadcast\footnote{An alternative to using broadcasting is to use an epidemic protocol~\cite{eugster2004epidemic}, 
 as long as it can guarantee delivery.} two messages to all the nodes---
@@
 Nodes that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
-$txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$,
-then sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.
+$txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$.
+Then the initiator sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.
 
@@
     \item All the CP blocks are in consensus.
-    \item The TX of interest is in the chian.
+    \item The TX of interest is in the chain.
 \end{enumerate}

### 2017-08-13 paper/s4-analysis.tex
Spell check paper

@@
 We evaluate our system analytically to ensure the desired properties (\Cref{def:consensus} and \Cref{def:validation}) hold.
-An informal argument is given in this section, we refer to~\cite[Chapter 4]{checo}??? for an indepth analysis.
+An informal argument is given in this section, we refer to~\cite[Chapter 4]{checo}??? for an in-depth analysis.
 
@@
 
-Fairnress holds if we model $\textsf{H}(\cdot)$ as a query to the random oracle.
+Fairness holds if we model $\textsf{H}(\cdot)$ as a query to the random oracle.
 Sorting nodes by the luck value $\textsf{H}(\C || pk)$ can be seen as a random permutation of a list of nodes.
@@
 
-The validty property holds because we use $\textsf{get\_validity}(\cdot)$ in the validation protocol.
+The validity property holds because we use $\textsf{get\_validity}(\cdot)$ in the validation protocol.
 The agreement property holds because we model $\textsf{H}(\cdot)$ as a query to a random oracle.
@@
 Hence we are interested in the probability for this to happen under our facilitator election process.
-The problem is modeled by a hypergeometric distribution,
+The problem is modelled by a hypergeometric distribution,
 where the random variable $X$ is the number of malicious nodes.

### 2017-08-13 thesis/formalities/abstract.tex
Write preface and update Bitcoin situation
Fixes
Improve abstract

@@
 % what's the solution?
-We design a novel blockchain system which we call CheckpointConsensys.
+We design a novel blockchain system called \textsc{Checo}.
 Each node in our system maintain individual hash chains,
-which only stores transactions that involve the node.
+which only stores transactions that the node is involved in.
 Consensus is reached on special blocks called checkpoint blocks rather than on all the transactions.
@@
 The consensus protocol does not imply transaction validity.
-Hence we include a validation protocol which allows nodes to verify that the counterparty correctly recorded the transaction.
+Hence we include a validation protocol which allows nodes to verify that transactions are correctly recorded,
+which indirectly achieves consensus on transactions.
 % how good is the solution?
-We implement a CheckpointConsensys prototype and evaluate it experimentally.
+We implement a prototype and evaluate it experimentally.
 Our results show strong indication of horizontal scalability even in the worst case,
-where every transaction is with a randomly selected node.
+where every transaction is made with a randomly selected node.
 For 1200 nodes, we were able to perform almost 5000 transactions per second,

### 2017-08-14 paper/s1-intro.tex
Improve abstract
Update code URL in README

@@
 none of them achieves horizontal scalability.
+That is, by adding new nodes into the network, the global transaction rate should increase proportionally.
 
@@
 % research question
+Thus far, there are no systems that achieve horizontal scalability in the general case.
 In this work, answer the following research question.
@@
 Finally, we are interested in horizontal scalability in the general case, as it enables ubiquitous use.
-That is, by adding new nodes into the network, the global transaction rate should increase proportionally.
 
@@
 \begin{itemize}
-    \item We formally introduce a blockchain system---\textsc{Checo}\footnote{\emph{Che}ckpoint \emph{co}nsensus}.
+    \item We formally introduce a blockchain system---\textsc{Checo}\footnote{Derived from ``\emph{Che}ckpoint \emph{co}nsensus''.}.
         It uses individual hash chains and checkpoints on every node to achieve
-        horizontally scalable for the first time.
+        horizontally scalable in the general case for the first time.
     \item We analyse \textsc{Checo} to ensure correctness as defined in our architecture.

### 2017-08-15 paper/s1-intro.tex
Edit paper

@@
 Driven by the success of Bitcoin, we are seeing a renaissance of consensus research~\cite{miller2016honey, kogias2016enhancing, kokoris2017omniledger},
-where the primary focus is to improve the scalability of blockchain system.
-This is due to the inefficiencies in the consensus mechanism in early blockchain systems---proof-of-work (PoW).
+where the primary focus is to improve the scalability of blockchain systems.
+This is due to the inefficiencies of the consensus mechanism---proof-of-work (PoW).
 For example, Bitcoin can only do 7 transactions per second (TPS) at most~\cite{vukolic2015quest}.
@@
 it also leads to centralisation as larger blocks take longer to propagate in the network,
-giving miners that do not have a fast network a disadvantage~\cite{croman2016scaling}.
-In today's network, it is not possible to achieve more than 758 TPS
-if new blocks need to be propagated timely to 90\% of the network~\cite{croman2016scaling}.
+putting miners that do not have a fast network at a disadvantage~\cite{croman2016scaling}.
+Furthermore, due to the bandwidth and latency in today's network,
+it is not possible to achieve more than 27 TPS from simply adjusting the block size~\cite{croman2016scaling}.
 
@@
 The best examples are Lightning Network~\cite{lightningnetwork} and Duplex Micropayment Channels~\cite{decker2015fast}.
-It promises significant scalability improvements,
-but complicates user experience and heavily relies on the semantics of transactions.
+It promises significant scalability improvements, but complicates user experience and leads to centralisation.
 Each node must deposit a suitable amount of Bitcoins into a multi-signature account.
@@
 In addition, the user must proactively check whether the counterparty has broadcasted an old channel state so that the user does not lose Bitcoins.
-Payment channel, in theory, solves the scalability problem of early blockchain systems,
-but to the best of our knowledge, its exact scalability characteristics are not investigated.
+Finally, creating channels with sufficient balance and also keeping it online to act as a router is expensive.
+A casual user is not capable of such tasks, thus it leads to centralisation.
 
-Another way to improve transaction rate is to use traditional Byzantine consensus algorithms such as PBFT~\cite{castro1999practical}.
+Another way to improve transaction rate is to use traditional Byzantine consensus algorithms such as PBFT~\cite{castro1999practical} in a permissioned ledger such as Hyperledger Fabric~\cite{cachin2016architecture}.
 In essence, they contain a fixed set of nodes (sometimes called validating peer) that run a Byzantine consensus algorithm to decide on new blocks.
-This is often used in permissioned system where the validators must be predetermined, e.g. Hyperledger Fabric~\cite{cachin2016architecture}.
 Such systems can achieve much higher transaction rates, e.g. 10,000 TX/s if the number of validating peer is under 16 for PBFT~\cite[Section 5.2]{miller2016honey}.
@@
 and Byzantine consensus algorithms to agree on transactions, e.g. SCP~\cite{luu2015scp}, ByzCoin~\cite{kogias2016enhancing} and Solidus~\cite{abraham2016solidus}.
-This design is primarily for permissionless system because the PoW leader election aspect prevents Sybils.
+This design is primarily for permissionless systems because the PoW leader election aspect prevents the Sybil attack~\cite{douceur2002sybil}.
 This technique overcomes the early blockchain scalability issue by delegating the transaction validation to the Byzantine consensus protocol (e.g. PBFT in ByzCoin~\cite{kogias2016enhancing}).
-A major tradeoff of such systems is that they cannot guarantee correctness when there is a large number of malicious nodes (but less than a majority).
-For SCP, ByzCoin and Solidus, they all have some probability to elect more than $t$ Byzantine nodes into the committee,
+A major tradeoff of such systems is that they cannot guarantee a high level of fault tolerance when there is a large number of malicious nodes (but less than a majority).
+SCP, ByzCoin and Solidus all have some probability to elect more than $t$ Byzantine nodes into the committee,
 where $t$ is typically just under a third of the committee size (a lower bound of Byzantine consensus~\cite{pease1980reaching}).
-This problem is especially difficult to solve because the committee is always much smaller than the population size which has more than $t$ Byzantine nodes.
+This problem is especially difficult to solve because the committee is always much smaller than the population size which usually has more than $t$ Byzantine nodes.
 Again, due to the fact that these systems must reach consensus on all transactions,
@@
 The number of shards grows linearly with respect to the total computational power in the network, hence the transaction rate also grows linearly.
-The biggest limitation of sharding is that it is only optimal is transactions stay in the same shard.
+The biggest limitation of sharding is that it is only optimal if transactions stay in the same shard.
 In fact, Elastico cannot atomically process inter-shard transactions.
@@
 A large shard size would make the system less scalable because the Byzantine consensus algorithm must be run by a large number of nodes.
-A small shard size would result in a large number of inter-shard transactions which also hinder scalability.
-The authors of OmniLedger noted that inter-shard transactions have significantly higher latency compared to the consensus protocol~\cite{kokoris2017omniledger}.
+A small shard size would result in a large number of inter-shard transactions which also hinders scalability.
+% The authors of OmniLedger noted that inter-shard transactions have significantly higher latency compared to the consensus protocol~\cite{kokoris2017omniledger}.
 
 % research question
-Thus far, there are no systems that achieve horizontal scalability in the general case.
-In this work, answer the following research question.
+Thus far, there are no systems that achieve horizontal scalability in the general case,
+which leads to the goal of this work.
+Concretely, we attempt to answer the following research question.
 \begin{displayquote}
@@
 A blockchain consensus protocol should be application neutral,
-for example PoW without the transaction logic.
-Our system should be Byzantine fault tolerant up to some threshold.
+for example, PoW without the transaction logic.
+The protocol should be Byzantine fault tolerant up to some threshold.
 The threshold may be made adjustable by trading off performance.
@@
 % contribution
-The key insight is to \emph{not} reach consensus using an existing consensus algorithm (a modified HoneyBadgerBFT~\cite{miller2016honey}) on transactions themselves but on special blocks called checkpoint blocks,
+The key insight is to \emph{not} reach consensus using an existing consensus algorithm (a modified HoneyBadgerBFT~\cite{miller2016honey}) on transactions themselves,
+but on special blocks called checkpoint blocks,
 such that transactions are nevertheless verifiable at a later stage by any node in the network.
@@
         It uses individual hash chains and checkpoints on every node to achieve
-        horizontally scalable in the general case for the first time.
-    \item We analyse \textsc{Checo} to ensure correctness as defined in our architecture.
+        horizontal scalability in the general case for the first time.
+    \item We analyse \textsc{Checo} to ensure correctness according to our definition.
     \item We provide an implementation and then experiment with up to 1200 nodes,

### 2017-08-15 paper/s2-problem.tex
Edit paper

@@
 Transactions can be in one of three states---\emph{valid}, \emph{invalid} and \emph{unknown}, defined by some function $\textsf{get\_validity}(\cdot)$.
-We seek a protocol that defines these states given a transaction,
+We seek a protocol that reasonably defines these states given a transaction,
 but also satisfies the following properties.
@@
 
-Note that the agreement and validity property are similar to what is often seen in Byzantine consensus problems but there are subtle differences and relaxations.
+Note that the agreement and validity property are similar to what is often seen in Byzantine consensus problems,
+but there are subtle differences and relaxations.
 Firstly, the agreement property only holds if honest nodes do not output \emph{unknown}.
-For example, it is OK if two honest nodes output \emph{valid} and \emph{unknown}, but it is not OK to output \emph{valid} and \emph{invalid}.
+For example, for a transaction, it is OK if two honest nodes output \emph{valid} and \emph{unknown}, but it is not OK to output \emph{valid} and \emph{invalid}.
 Secondly, the validity is a relaxation of what is typically seen in Byzantine consensus problems,
 namely we do not make any guarantee on whether the transaction is \emph{valid} even if all nodes are honest,
-only that it is correct according to the $\textsf{get\_validity}(\cdot)$ defined by the protocol.
+only that it is correct according to $\textsf{get\_validity}(\cdot)$ defined by the protocol.
 
-Our problem is purposefully made to be application neutral,
-i.e. there are no constraints on the semantics of transaction.
+The problem is purposefully made to be application neutral,
+i.e. there are no constraints on the semantics of transactions.
 This is so that the protocol can act as a backbone to many applications.
@@
 as some application may not need such strong guarantees such as recording the number of bytes uploaded and downloaded in Tribler~\cite{pimotte, pouwelse2008tribler}.
-On the other hand, we give an alternative construction that does have fork detection in~\Cref{sec:fork}.
+On the other hand, we give two alternative constructions that do have fork detection in~\Cref{sec:fork}.
 
@@
 The adversary has full control of the delivery schedule and the message ordering of all messages.
-But they are not allowed to drop messages except for their own.
-Reliability can be achieved in unreliable networks by resending messages or using some error correction code.
+But they are not allowed to drop messages except for their own\footnote{Reliability can be achieved in unreliable networks by resending messages or using some error correction code.}.
 
 \textbf{Security assumptions.}
-The malicious nodes are Byzantine and we use a static, round-adaptive corruption model.
+The malicious nodes are Byzantine,
+meaning that there are no restrictions on the type of failure.
+We use a static, round-adaptive corruption model.
 That is, if a round has started, the corrupted nodes cannot change until the next round.

### 2017-08-15 paper/s3-system.tex
Edit paper

@@
 Transactions are stored in TX blocks and every block only contains one transaction.
-We introduce a special block called checkpoint (CP) block, which represents the state of a hash chain.
+We introduce a special block called checkpoint (CP) block,
+which represents the state of a hash chain in the form of a hash pointer.
 Such data structure is called Extended TrustChain,
-a visualisation can be seen on~\Cref{fig:trustchain-good-cp}.
+a visualisation can be seen in~\Cref{fig:trustchain-good-cp}.
 \begin{figure}[h]
@@
 \caption{Visualisation of the Extended TrustChain. $t_{u, i}$ represents a TX block on $u$'s chain with a sequence number of $i$.
-$c_{v, j}$ represents a CP block on $v$'s chain with a sequence number of $j$.}
+$c_{v, j}$ represents a CP block on $v$'s chain with a sequence number of $j$.
+A transaction is represented by a \emph{pair} of TX blocks,
+one on each of the party's chain.
+}
 \label{fig:trustchain-good-cp}
@@
 The consensus result is essentially a set of CP blocks from (nearly) all the nodes.
-While this gives us consensus on a representation of the global state, it does not imply consensus on transactions.
-To this end, we introduce a validation protocol that allows any nodes to challenge any other node to produce a set of TX blocks that computes to the CP block in consensus.
-Hence we implicitly achieve consensus on transactions.
+While this gives us a consensus on a representation of the global state, it does not imply consensus on transactions.
+To this end,
+we introduce a validation protocol that allows any nodes to challenge any other node to produce a set of TX blocks that computes to the CP block in consensus,
+achieving implicit consensus on transactions.
 Since the transaction and validation protocols only make point-to-point communication,
@@
 The formal description is given next,
-it consist of one data structure---Extended TrustChain,
+it consists of one data structure---Extended TrustChain,
 and three protocols---consensus protocol, transaction protocol and validation protocol.
+The complete protocol can be seen as the concurrent composition of the sub-protocols.
 
@@
 Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
-The chain consist of blocks $B_u = \{ b_{u, i} : i \in \{ 0, \dots, h - 1 \} \}$,
+The chain consists of blocks $B_u = \{ b_{u, i} : i \in \{ 0, \dots, h - 1 \} \}$,
 where $b_{u, i}$ is the $i$th block of $u$,
@@
 \item $pk_v$ is the public key of the counterparty $v$.
-\item $m$ is the transaction message.
+\item $m$ is the transaction message,
+    which can be seen as an arbitrary string.
 \item $sig_u$ is the signature created using $sk_u$ on the concatenation of the binary representation of the five items above.
 \end{enumerate}
-The fact that we have no constraint on the content of $m$ is in alignment with our problem description---application neutrality.
+% The fact that we have no constraint on the content of $m$ is in alignment with our problem description---application neutrality.
 
@@
 there exist one and only one \emph{pair} 
-$$t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), seq_v, txid, pk_u, m, sig_v \rangle.$$
+$$t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), seq_v, txid, pk_u, m, sig_v \rangle,$$
+if the nodes follow the transaction protocol (described in~\Cref{sec:tx-protocol}).
 Note that the $txid$ and $m$ are the same, and the public keys refer to each other.
@@
 Here we define an important property which results from the interleaving nature of CP and TX blocks.
-It is used in our validation protocol.
+It is used in our validation protocol (discussed in~\Cref{sec:vd-protocol}).
 \begin{definition}
@@
 where 
-$$a = \argmin_{k, k < i, \texttt{typeof}(b_{u,k}) = \gamma}(i - k)$$
-$$b = \argmin_{k, k > i, \texttt{typeof}(b_{u,k}) = \gamma}(k - i),$$
+$$a = \argmin_{k, k < i, \textsf{typeof}(b_{u,k}) = \gamma}(i - k)$$
+$$b = \argmin_{k, k > i, \textsf{typeof}(b_{u,k}) = \gamma}(k - i),$$
 then $\langle c_{u,a}, c_{u, b} \rangle$ is the enclosure of $t_{u, i}$.
 Intuitively, that is the two closest CP blocks that surround $t_{u, i}$.
-Note that $c_{u, a}$ is the more recent CP block.
-Also, some TX blocks may not have any enclosure, then its enclosure is $\bot$.
+Some TX blocks may not have any enclosure, then its enclosure is $\bot$.
 Agreed enclosure is the same as enclosure with an extra constraint where the CP blocks must be in some consensus result $\C_r$.
@@
 Our scalable consensus protocol runs on top of Extended TrustChain.
-It uses an unmodified asynchronous subset consensus (ACS) algorithm as the key building block (we describe the reasoning later).
+It uses an unmodified asynchronous subset consensus (ACS) algorithm as the key building block.
 The objectives of the protocol are to
@@
 
-\subsubsection{Bootstrap phase}
+\subsubsection{Bootstrap Phase}
 \label{sec:bootstrap}
@@
 
-\subsubsection{Consensus phase}
+\subsubsection{Consensus Phase}
 \label{sec:consensus-phase}
 The consensus phase begins when $\F_r$ is available to all the nodes.
-Note that $\F_r$ indicates the facilitators that were elected using results of round $r$ and are responsible for driving the ACS protocol in round $r + 1$.
+Note that $\F_r$ indicates the facilitators that were elected using results of round $r$ and are responsible for driving the ACS algorithm in round $r + 1$.
 The goal is to reach agreement on a set of new facilitators $\F_{r + 1}$ that satisfies the four properties in \Cref{def:consensus}.
@@
 Upon receiving $\C_{r + 1}$ and at least $n - f$ valid signatures by some node $u$, $u$ performs two tasks.
-First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1}, r + 1)$ using \Cref{alg:new-cp}.
+First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1})$ (\Cref{alg:new-cp}).
 Second, it computes the new facilitators using $\textsf{get\_facilitator}(\C_{r + 1}, n)$ (\Cref{alg:facilitator}),
@@
 \begin{algorithm}
-\caption{Function $\textsf{new\_cp}(\C_r, r)$ runs in the context of the caller $u$.
+\caption{Function $\textsf{new\_cp}(\C_r)$ runs in the context of the caller $u$.
 It creates a new CP block and appends it to $u$'s chain.}
@@
 \begin{algorithmic}
+\State $\langle r, \_ \rangle \gets \C_r$
 \State $h \gets |B_u|$
@@
 
-\subsection{Validation protocol}
+\subsection{Validation Protocol}
 \label{sec:vd-protocol}
@@
 The validation protocol aims to solve this issue.
-The protocol is also a request-response protocol, just like the transaction protocol.
+The protocol is also a request-response protocol.
 But before explaining the protocol itself, we first define what it means for a transaction to be valid.
 
-\subsubsection{Validity definition}
+\subsubsection{Validity Definition}
 A transaction can be in one of three states in terms of validity---\emph{valid}, \emph{invalid} and \emph{unknown}.
@@
 otherwise the node calling the function would have no point of reference.
-This is not difficult to achieve,
-because typically the caller is $u$, so it knows its own TX block and the corresponding agreed fragment.
+This is not difficult to achieve because typically the caller is $u$,
+so it knows its own TX block and the corresponding agreed fragment.
 If the caller is not $u$, it can always query for the agreed fragment that contains the transaction of interest from $u$.
@@
 
-Note that the validity is on a transaction (two TX blocks with the same $txid$), rather than on one TX block owned by a single party.
+Note that the validity is on a transaction, i.e. two TX blocks that form a pair.
 It is defined this way because the malicious sender may create new TX blocks in their own chain but never send \texttt{tx\_req} messages.
@@
 
-\subsubsection{Validation protocol}
+\subsubsection{Validation Protocol}
 Our validation protocol,
@@
 We make two remarks.
-First, just like the TX protocol, we do not block at any part of the protocol.
+First, just like the TX protocol, we do not block any part of the protocol.
 Second, suppose some $F_{v, j}$ validates $t_{u, i}$, then that does not imply that $t_{u, i}$ only has one pair $t_{v, j}$.
@@
 The counterparty may create any number of fake pairs in later consensus rounds.
-But these fake pairs only pollutes the chain of $v$ and can never be validated because the round is incorrect.
+But these fake pairs only pollutes the chain of $v$ and can never be validated.
 
-\subsection{Design variations and tradeoffs}
+\subsection{Design Variations and Tradeoffs}
 \label{sec:tradeoffs}
 In this section, 
-we explore a few design variations which we can make, some of them require a relaxed version of our original model.
+we explore a few design variations, some of them require a relaxed version of our original model.
 They enable better performance and allow us to apply our design in the fully permissionless setting.
 
-\subsubsection{Using timing assumption in the permissionless setting}
+\subsubsection{Using Timing Assumption in the Permissionless Setting}
 \label{sec:permissionless}
-One limitation of our system is that we use the parameter $N$ in our algorithms, which makes it unsuitable for the permissionless environment where users can join and leave at will.
-In this section we show how making a timing assumption would allow us to operate in the permissionless setting.
-
 At the start of our consensus phase (\Cref{sec:consensus-phase}), facilitators must wait for $N-f$ \texttt{cp\_msg} messages.
-This is the only place where we used $N$ as a parameter.
+The use of $N$ makes our system unsuitable for the permissionless setting.
 To introduce timing, instead of waiting for $N-f$ messages, we wait for some time $D$,
@@
 
-\subsubsection{Optimising validation protocol using cached agreed fragments}
+\subsubsection{Optimising Validation Protocol Using Cached Agreed Fragments}
 \label{sec:caching}
@@
 rather than validating a single transaction,
-$A$ attempts to validate all transactions with performed with $B$ in the unknown state, in that fragment.
+$A$ attempts to validate all transactions performed with $B$ in the unknown state, in that fragment.
 
@@
 
-\subsubsection{Total fork detection}
+\subsubsection{Total Fork Detection}
 \label{sec:fork}
@@
 $A$ first sends a challenge to $B$ asking it to prove that it holds a valid agreed fragment between some consensus round specified by $A$.
-If $B$ provides a correct proof, then they run the transaction protocol as usual.
-If $B$ provides an invalid proof or refuses to respond,
-then $A$ would refuse to make the transaction.
+If $B$ provides a correct and timely response, then they run the transaction protocol as usual.
+Otherwise, $A$ would refuse to make the transaction.

### 2017-08-15 paper/s4-analysis.tex
Edit paper

@@
-\section{Correctness and fault tolerance analysis}
+\section{Correctness and Fault Tolerance Analysis}
 \label{sec:analysis}
@@
 
-\subsection{Correctness of the consensus protocol}
+\subsection{Correctness of the Consensus Protocol}
 Recall that the consensus protocol properties (\Cref{def:consensus}) has
@@
     \item The CP blocks sent to the facilitators are eventually delivered and then ACS eventually starts.
-    \item Agreement, validity and termination holds for ACS as they are the properties of ACS and are proven to hold in~\cite{miller2016honey}.
+    \item Agreement, validity and termination hold for ACS as they are the properties of ACS and are proven to hold in~\cite{miller2016honey}.
     \item The consensus result and signatures are eventually disseminated to all the nodes,
@@
 
-Fairness holds if we model $\textsf{H}(\cdot)$ as a query to the random oracle.
+Fairness holds because we model $\textsf{H}(\cdot)$ as a query to the random oracle.
 Sorting nodes by the luck value $\textsf{H}(\C || pk)$ can be seen as a random permutation of a list of nodes.
@@
 
-\subsection{Correctness of the validation protocol}
+\subsection{Correctness of the Validation Protocol}
 \label{sec:correctness-of-validity}
@@
 The agreement property holds because we model $\textsf{H}(\cdot)$ as a query to a random oracle.
-Suppose two honest nodes decided on two different state, \emph{valid} and \emph{invalid} for the same transaction.
-Due to the properties of the consensus protocol, the agreed enclosure for the transaction must be the same.
-For that to happen, two agreed fragments must exist for the same transaction.
+Suppose two honest nodes decided on two different states, \emph{valid} and \emph{invalid} for the same transaction.
+For that to happen, two agreed fragments must exist for the same transaction,
+but these fragments must also have the same agreed enclosure.
 Recall that blocks form a hash chain.
-So this cannot happen unless the adversary can compute the inverse of $\textsf{H}(\cdot)$ with high probability.
+So this is not possible unless the adversary can compute the inverse of $\textsf{H}(\cdot)$ with high probability.
 
-Liveness, unfortunately does not hold in our model.
-A malicious node can ask honestly when running the transaction protocol,
+Liveness, unfortunately, does not hold in our model.
+A malicious node can act honestly when running the transaction protocol,
 but then never respond to any validation requests.
 Therefore there are transactions that can never be validated.
-Nevertheless, the malicious node will be at an economical loss if it is not responsive because honest nodes are less likely to make contact with nodes that do not respond to validation requests.
+Nevertheless, the malicious node will be at an economic loss if it is not responsive because honest nodes are less likely to make contact with nodes that do not respond to validation requests.
 
@@
 
-\subsection{Fault tolerance}
+\subsection{Fault Tolerance}
 Finally, we consider the effect when the number of adversaries is more than $t$.
@@
 $$
-Let $E[X] = n\alpha$ where $0 \le \alpha \le \lfloor \frac{n-1}{3} \rfloor/n < (\lfloor  \frac{n - 1}{3} \rfloor + 1) / n$.
+Let $E[X] = n\alpha$ where $0 \le \alpha \le \lfloor \frac{n-1}{3} \rfloor/n$.
 Using the tail inequality found in~\cite{skala2013hypergeometric}
@@
 
-The bound is not tight, but it is useful for picking parameters for a desired level of fault tolerance.
+The bound is not tight, but it is useful for picking parameters for the desired level of fault tolerance.
 If $n$ is known, then we can pick a $\alpha$ such that the probability becomes small.
-On the other hand, if $\alpha$ is fixed, but small, we may increase $n$ to achieve the same.
+On the other hand, if $\alpha$ is fixed, we may increase $n$ to achieve the same.
 Furthermore, hypergeometric distributions have light tails with ``faster-than-exponential fall-off''~\cite{skala2013hypergeometric}.
-So the probability for picking more than $\lfloor \frac{n-1}{3} \rfloor$ malicious nodes drops of exponentially as the expected value moves away from $\lfloor \frac{n-1}{3} \rfloor$.
+So the probability of picking more than $\lfloor \frac{n-1}{3} \rfloor$ malicious nodes drops of exponentially as the expected value moves away from $\lfloor \frac{n-1}{3} \rfloor$.
 To put this into perspective,

### 2017-08-15 paper/s5-evaluation.tex
Edit paper

@@
 \begin{displayquote}
-\url{https://github.com/kc1212/consensus-thesis-code}
+\url{https://github.com/kc1212/checo}
 \end{displayquote}
 It implements the three protocols and the Extended TrustChain.
-We also implement the caching optimisations discussed in \Cref{sec:caching}.
+We also implement the caching optimisation discussed in \Cref{sec:caching}.
 It is written in the Python programming language\footnote{\url{https://www.python.org/}}.
@@
 
-To coordinate nodes on many different machines,
-we employ a discovery server to inform every node the IP addresses and port numbers of every other node.
-It is only run before the experiment and is not used during the experiment.
-
-The global throughput results are shown in~\Cref{fig:global-throughput-fixed} and \Cref{fig:global-throughput-random}.
+The global throughput results are shown in~\Cref{fig:global-throughput}.
 We consider~\Cref{fig:global-throughput-fixed} as the ideal case,
@@
 where nodes make transactions with random nodes and the caching mechanism is unlikely to be used.
-The low transaction rate in~\Cref{fig:global-throughput-random} is caused by the fact that the network infrastructure cannot keep up with our demand.
+Observe that the transaction rate is much lower in~\Cref{fig:global-throughput-random}.
+This is because the communication of an agreed fragment is necessary to verify every transaction (no caching),
+putting a strain on our network infrastructure.
 In practice, we do not expect such behaviour to occur as it is possible to cache agreed fragments.
@@
 This result is a strong indication of the horizontal scalability which we aimed to achieve.
-For more experimental results, refer to~\cite[Chapter 5]{checo}???.
+For more experimental results refer to~\cite[Chapter 5]{checo}???.
 

### 2017-08-15 paper/s6-conclusion.tex
Edit paper

@@
 
-In this work we introduced an application neutral blockchain system which we call \textsc{Checo}.
-We add checkpoint block to the existing TrustChain data structure for use in our consensus protocol.
+We introduced an application neutral blockchain system which we call \textsc{Checo}.
+The checkpoint block is added to the existing TrustChain data structure for use in our consensus protocol.
 The round based consensus protocol uses ACS as a building block to reach consensus on checkpoint blocks.
-The consensus result, which is a set of checkpoint blocks, lets nodes elect new facilitators and create new checkpoint blocks.
+The consensus result lets nodes elect new facilitators and create new checkpoint blocks.
 To make transactions, nodes use the transaction protocol,
 which is adapted from an existing work called True Halves.
-Finally, we introduce a validation protocol which ensures that if agreed fragments for some transaction exists,
+Finally, we introduce a validation protocol which ensures that if an agreed fragment for some transaction exists,
 then nodes reach agreement on the validity of that transaction.
@@
 in the ideal case as well as the worst case.
+In the future, we hope to apply our system to a concrete application and evaluate its performance.

### 2017-08-15 thesis/chapters/ch1-intro.tex
Update code URL in README
Add payment channel centralisation, etc.

@@
 But given the bandwidth and latency constraints of today's network,
-it is not possible to have more than 758 transactions per second if messages need to be propagated timely to 90\% of the nodes.
+it is not possible to have more than 27 transactions per second.
 To achieve transaction rates higher than that, it will lead to unfairness for nodes that do not have a good network connection.
@@
         It uses individual hash chains and checkpoints on every node to achieve
-        horizontally scalable for the first time.
+        horizontally scalable in the general case for the first time.
     \item We formally analyse \textsc{Checo} to ensure correctness as defined in our architecture.

### 2017-08-15 thesis/chapters/ch2-problem.tex
Give paper a structure
Add payment channel centralisation, etc.

@@
 Hence double spending is possible for transactions that do not have a large number of confirmations.
-Furthermore, as we mentioned in the Introduction,
+Furthermore, as we mentioned in the~\Cref{ch:intro},
 the performance is limited due to the fact that blocks are of fixed size and are generated at fixed intervals.
@@
 
-On the other hand, off-chain transactions also suffer from the problem of Bitcoin.
-Namely, proof-of-work consumes an unreasonable amount of power.
-Further, payment channels complicate user experience.
+On the other hand, payment channel complicates user experience and leads to centralisation.
 As we mentioned earlier, each node must deposit some Bitcoins into a multi-signature account.
@@
 In addition, the user must proactively check whether the counterparty has broadcasted an old channel state so that the user does not lose Bitcoins.
-Payment channel, in theory, solves the scalability problem of early blockchain systems,
+Furthermore, creating channels with sufficient balance and also keeping it online to act as a router is expensive.
+A casual user is not capable of such tasks, thus it leads to centralisation.
+Payment channels, in theory, solve the scalability problem of early blockchain systems,
 but to the best of our knowledge, its exact scalability characteristics are not investigated.
@@
 We saw in~\Cref{sec:taxonomy} that the majority of the systems cannot achieve horizontal scalability.
-That is, by adding new nodes into the network, the global transaction rate should increase proportionally.
 To the best of our knowledge, only two systems---Elastico~\cite{luu2016elastico} and OmniLedger~\cite{kokoris2017omniledger}---has experimentally demonstrated horizontal scalability.

### 2017-08-15 thesis/chapters/ch3-model.tex
Improve abstract
Add payment channel centralisation, etc.

@@
 where 
-$$a = \argmin_{k, k < i, \texttt{typeof}(b_{u,k}) = \gamma}(i - k)$$
-$$b = \argmin_{k, k > i, \texttt{typeof}(b_{u,k}) = \gamma}(k - i),$$
+$$a = \argmin_{k, k < i, \textsf{typeof}(b_{u,k}) = \gamma}(i - k)$$
+$$b = \argmin_{k, k > i, \textsf{typeof}(b_{u,k}) = \gamma}(k - i),$$
 then $\langle c_{u,a}, c_{u, b} \rangle$ is the enclosure of $t_{u, i}$.
@@
 Before describing the protocol in detail,
-we take a brief detour to give background on the ACS algorithm as it is the primary building block.
+we take a brief detour to give background on the ACS protocol as it is the primary building block.
 Then we move on to describe the two phases of our consensus protocol---bootstrap phase and consensus phase.
@@
 Upon receiving $\C_{r + 1}$ and at least $n - f$ valid signatures by some node $u$, $u$ performs two tasks.
-First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1}, r + 1)$ (\Cref{alg:new-cp}).
+First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1})$ (\Cref{alg:new-cp}).
 Second, it computes the new facilitators using $\textsf{get\_facilitator}(\C_{r + 1}, n)$ (\Cref{alg:facilitator}),
@@
 $A$ first sends a challenge to $B$ asking it to prove that it holds a valid agreed fragment between some consensus round specified by $A$.
-If $B$ provides a correct proof, then they run the transaction protocol as usual.
-If $B$ provides an invalid proof or refuses to respond,
-then $A$ would refuse to make the transaction.
+If $B$ provides a correct and timely response, then they run the transaction protocol as usual.
+Otherwise, $A$ would refuse to make the transaction.
 

### 2017-08-15 thesis/formalities/abstract.tex
Add payment channel centralisation, etc.

@@
 % what's the problem?
-Blockchain systems have the potential to decentralise many traditionally centralised system.
+Blockchain systems have the potential to decentralise many traditionally centralised systems.
 However, scalability remains a key challenge.
-Without a horizontally scalable solution, blockchain system remain unsuitable for ubiquitous use.
+Without a horizontally scalable solution, blockchain systems remain unsuitable for ubiquitous use.
 % what's the solution?
 We design a novel blockchain system called \textsc{Checo}.
-Each node in our system maintain individual hash chains,
+Each node in our system maintains individual hash chains,
 which only stores transactions that the node is involved in.
-Consensus is reached on special blocks called checkpoint blocks rather than on all the transactions.
+A consensus is reached on special blocks called checkpoint blocks rather than on all the transactions.
 Checkpoint blocks are effectively a hash pointer to the individual hash chains,
@@
 We implement a prototype and evaluate it experimentally.
-Our results show strong indication of horizontal scalability even in the worst case,
+Our results show a strong indication of horizontal scalability even in the worst case,
 where every transaction is made with a randomly selected node.
 For 1200 nodes, we were able to perform almost 5000 transactions per second,
-orders of magnitude higher than the 7 transaction per second limit of Bitcoin.
+orders of magnitude higher than the 7 transactions per second limit of Bitcoin.
 \end{abstract}

### 2017-08-17 thesis/chapters/ch1-intro.tex
Rework ch1 and ch2; fix title

@@
 
-Ironically, also in 2008, Satoshi Nakamoto published the Bitcoin whitepaper~\cite{bitcoin}.
+Ironically, during the midst of the financial crisis, Satoshi Nakamoto published the Bitcoin whitepaper~\cite{bitcoin}.
 Which, for the first time, gave us a simple banking system in the form of a distributed ledger, also known as blockchain.
@@
 their limitations also became apparent.
-Bitcoin has the infamous 7 transactions per second upper bound~\cite{vukolic2015quest}.
-This is due to the fact that blocks are fixed to 1 MB,
-and are only generated on average every 10 minutes.
+Bitcoin has the infamous 7 transactions per second upper bound~\cite{vukolic2015quest}, 
+which is due to the fact that blocks are fixed to 1 MB and are only generated on average every 10 minutes.
 Since every Bitcoin transaction is at least 250 bytes, it computes to about 7 transactions per second.
@@
 A few months ago the backlog even reached 100,000, which meant new transactions would take at least 11 hours to be written to the blockchain~\cite{bitcoinbacklog}.
-This issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some calls it a civil war~\cite{bitcoincivilwar}.
+This issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some call a civil war~\cite{bitcoincivilwar}.
 Parties that are for the increase in block size argue that a larger block would improve the transaction rate.
@@
 which risks consensus failure and devaluation.
-A recent empirical study by Croman et el.~\cite{croman2016scaling} has shown that increasing the block size or the block interval may help.
+A recent empirical study by Croman et al.~\cite{croman2016scaling} has shown that increasing the block size or the block interval may help.
 But given the bandwidth and latency constraints of today's network,
@@
 \section{Contributions}
-In this work, we take the advice of Croman et el.~\cite{croman2016scaling} and Vukoli\'{c}~\cite{vukolic2015quest} and rethink the blockchain architecture.
+In this work, we take the advice of Croman et al.~\cite{croman2016scaling} and Vukoli\'{c}~\cite{vukolic2015quest} and rethink the blockchain architecture.
 Our primary insight came from observing the differences between how value transfer work in the real world and how they work in early blockchain systems.
@@
 but nevertheless verifiable at a later stage by any node in the network.
-We call our protocol \textsf{Checo}, derived from \emph{che}ckpoint \emph{co}nsensus.
+We call our protocol \textsc{Checo}, derived from \emph{che}ckpoint \emph{co}nsensus.
 Informally, our system works as follows.
@@
 If a checkpoint block of some node is in consensus, 
-then that node can proof to any other node that it holds a set of transactions that compute (form a chain) to the checkpoint block.
+then that node can prove to any other node that it holds a set of transactions that compute (form a chain) to the checkpoint block.
 This immediately shows that those transactions are tamper-proof.
@@
 By splitting this functionality we achieved horizontal scalability in the general case.
-That is, by adding new nodes into the network, the global transaction rate should increase proportionally.
+That is, by adding new nodes into the network, the global transaction rate should increase proportionally,
+regardless of the choice of system parameters.
 
@@
 \section{Document structure}
-We describe and analyse our idea in the remainder of this work.
-We begin with the problem description in~\Cref{ch:problem} which describes the state-of-the-art work on making blockchain systems scalable and our research question.
+We begin with the problem description in~\Cref{ch:problem},
+it first describes the state-of-the-art in scalable blockchain systems and their shortcomings,
+which leads to our research question.
 In~\Cref{ch:model} we give a formal description of our system model and architecture.
@@
 Implementation and experimental results are discussed in~\Cref{ch:implementation}.
-Finally, we conclude in \ref{ch:conclusion} and discuss future work.
+Finally, we conclude in~\Cref{ch:conclusion} and discuss future work.

### 2017-08-17 thesis/chapters/ch2-problem.tex
Rework ch1 and ch2; fix title

@@
 That is, transactions never reach consensus with a probability of 1.
+Given sufficient computational power, adversaries are able to undo any transaction.
 The typical examples are proof-of-work based systems such as Bitcoin, Ethereum and other Altcoins.
@@
 In essence, they contain a fixed set of nodes (sometimes called validating peer) that run a Byzantine consensus algorithm to decide on new blocks.
-This is often used in permissioned system where the validators must be predetermined,
-for example, Hyperledger Fabric~\cite{cachin2016architecture}.
+This technique is often used in permissioned system where the validators must be predetermined.
+For example, Hyperledger Fabric~\cite{cachin2016architecture}.
 
@@
 We wish to design a blockchain consensus protocol that can be used in many applications.
-For example recording uploads and downloads between peers in Tribler~\cite{pouwelse2008tribler, multichain} and decentralised market\footnote{\url{https://github.com/Tribler/tribler/blob/devel/Tribler/community/tradechain/__init__.py}}.
+For example, the accounting of detailed internet traffic in Tribler~\cite{pouwelse2008tribler, multichain} and decentralised market\footnote{\url{https://github.com/Tribler/tribler/blob/devel/Tribler/community/tradechain/__init__.py}}.
 Hence our research question is the following.
 \begin{displayquote}
-\emph{Is it possible to design a blockchain consensus protocol that is fault-tolerant, scalable and can reach global consensus?}
+\emph{How can we design a blockchain consensus protocol that is fault-tolerant,
+horizontally scalable,
+and able to reach global consensus?}
 \end{displayquote}
@@
 
-In the remainder of this section, we expand on our research question and visit each of our requirement in detail.
+In the remainder of this section, we expand on our research question and visit each of our requirements in detail.
 At the end of this section, we describe some issues that we do not consider as a part of our design,
@@
 
-\subsection{Global consensus}
-
-% Early blockchain systems use PoW as the consensus algorithm,
-% but consensus algorithm and especially Byzantine\footnote{Byzantine means participants can experience arbitrary failure, more on this in~\Cref{sec:acs-background}.}
-% consensus algorithms have existece since the 80s~\cite{pease1980reaching, lamport1982byzantine}.
-% The Practical Byzantine Fault Tolerance (PBFT) algorithm introduced in 1999 by Castro and Liskov~\cite{castro1999practical} is widely regarded as the first high-performance consensus algorithm.
-% The authors show that running PBFT in Network File System (NFS) was only 3\% slower than the standard unreplicated NFS.
-% Since then, research in this area picked up pace and many more algorithms are introduced, for example Q/U~\cite{abd2005fault} and Zyzzyva~\cite{kotla2007zyzzyva}.
-% With the introduction of Bitcoin's probabilistic consensus model, the field experienced a second renaissance.
-% These include proof-of-stake protocols~\cite{bentov2016snow, micali2016algorand}, as well as protocols optimised for blockchain~\cite{liu2016xft, miller2016honey}.
-
-The first objective is to reach consensus on a global \emph{state} by building on top of decades of Byzantine consensus protocol research~\cite{pease1980reaching, lamport1982byzantine, castro1999practical, miller2016honey}.
+\subsection*{Global consensus}
+% WHY WE NEED GLOBAL CONSENSUS?
 Having consensus is essential in blockchain systems.
-It stops many types of malicious activities because agreed state must have the consent of the honest nodes in the network.
-We emphasise the word state because in contrary to traditional blockchains,
+It stops many types of malicious activities because the agreed state, or a representation of it, must have the consent of the honest nodes in the network.
+We emphasise on the notion of a representation of the state because in contrary to traditional blockchains,
 it does not necessarily mean the set of all transactions that have ever happened.
-The state can be a verifiable representation of all the transactions.
+The state can be a small, but verifiable representation of all the transactions.
 As an extreme example, a single digest of the complete history can represent the state.
@@
 
-% Our approach is to combine the vast literature in Byzantine fault tolerance research with blockchain technology,
-% and introduce a new consensus model to record the global state.
-
-\subsection{Security and fault-tolerance}
-
-Our system should be unaffected in the presence of powerful adversaries.
+\subsection*{Fault-tolerance}
+% WHY WE NEED FAULT-TOLERANCE
+Similar to many blockchain systems,
+our system should be unaffected in the presence of powerful adversaries.
 In particular, adversaries are Byzantine meaning that they can have arbitrary behaviour.
 Thus anything is possible from simply omitting messages to colluding with each other in order to undermine the whole system.
-This aspect usually comes for free when using a Byzantine fault-tolerant consensus algorithm.
 
@@
 Further, the validation result must be consistent.
-For example, if two different honest nodes attempts verify the same transaction,
+For example, if two different honest nodes attempt to verify the same transaction,
 it cannot be the case that one node thinks that the transaction is valid but the other thinks it is invalid.
 
-% Recall that it is not neccessary to reach consensus on all the transactions in order to record the state.
-% But when using some digest (or a set of digets) to represent the state, it is also important for the transactions to be verifiable.
-% Otherwise adversaries could simply store invalid transactions in the state and they would never be detected.
+\subsection*{Horizontal scalability}
 
-\subsection{Performance and scalability}
-
-The last but not least, the final objective is horizontal scalability.
-We wish to see the performance (global transaction rate) increase as more nodes join the network.
+To enable ubiquitous use, we demand horizontal scalability.
+That is, the global transaction rate should increase as more nodes join the network.
 BitTorrent~\cite{cohen2003incentives} is an example of such a system,
 where peers interact with each other to exchange files without any global bottleneck.
+% Early blockchain systems may run in a network of thousands of nodes, but the performance does not scale.
+% On the other hand, as we will see in~\Cref{sec:acs-background}, Byzantine consensus algorithms perform well but only when in a small network.
+% In this work, our goal is to overcome the limitations of early blockchain systems and Byzantine consensus algorithms to create a horizontally scalable blockchain.
 
-Early blockchain systems may run in a network of thousands of nodes, but the performance does not scale.
-On the other hand, as we will see in~\Cref{sec:acs-background}, Byzantine consensus algorithms perform well but only when in a small network.
-In this work, our goal is to overcome the limitations of early blockchain systems and Byzantine consensus algorithms to create a horizontally scalable blockchain.
-
-\subsection{Limitations}
-Our work aims to significantly lower the transaction bound.
-However, the price to pay is that the Sybil attack~\cite{douceur2002sybil} prevention mechanism moves from the blockchain consensus protocol to the application layer.
-For example, social network based Sybil defence mechanisms use graph structure of real-world relationships~\cite{yu2006sybilguard}.
+\subsection*{Limitations}
+Since our consensus protocol is application neutral, it does not attempt to prevent the Sybil attack~\cite{douceur2002sybil}.
+A good defence mechanism requires application specific information.
+For example, a social network based Sybil defence mechanisms use graph structure of real-world relationships~\cite{yu2006sybilguard}.
 Online marketplaces such as Amazon use the rating of buyer and sellers.
-Since our system is application neutral, we do not provide an explicit Sybil defence mechanism.
-On the other hand, our system also has no restrictions on the Sybil defence technique
+% Since our system is application neutral, we do not provide an explicit Sybil defence mechanism.
+On the other hand, our system also has no restrictions on the Sybil defence mechanism
 and applications built on top of our system can use the best mechanism for their purpose.

### 2017-08-17 thesis/formalities/abstract.tex
Rework ch1 and ch2; fix title

@@
 \begin{tabular}{l}
-Blockchain Consensus Protocol with Horizontal Scalability\\
+A Blockchain Consensus Protocol With Horizontal Scalability\\
 \\
@@
 where every transaction is made with a randomly selected node.
-For 1200 nodes, we were able to perform almost 5000 transactions per second,
-orders of magnitude higher than the 7 transactions per second limit of Bitcoin.
 \end{abstract}

### 2017-08-17 thesis/formalities/preface.tex
Rework ch1 and ch2; fix title

@@
 \addcontentsline{toc}{chapter}{Preface}
-Since early 2013, when I mined my first Bitcoin, 
+Since early 2013, the time when I mined my first Bitcoin, 
 I have been interested in decentralising the traditionally centralised systems.

### 2017-08-18 thesis/chapters/ch1-intro.tex
Edit ch3

@@
 If $C$ can simultaneously claim that she transferred 5 units to $A$ but also 5 units to $B$, then the system is inconsistent.
-Bitcoin and many of its derivatives (e.g. Litecoin\footnote{\url{https://litecoin.org}} and Dogecoin\footnote{\url{http://dogecoin.org}}) solve the inconsistency problem with a consensus algorithm.
+Bitcoin and many of its derivatives (e.g., Litecoin\footnote{\url{https://litecoin.org}} and Dogecoin\footnote{\url{http://dogecoin.org}}) solve the inconsistency problem with a consensus algorithm.
 The goal of the algorithm is to reach agreement on a set of valid transactions. 
@@
 On the other hand, if the customer leaves without paying the bill,
-then the owner would report the incident to some central authority, e.g. the police.
+then the owner would report the incident to some central authority, e.g., the police.
 On the contrary, in blockchain systems, every transaction is effectively sent to the miners,

### 2017-08-18 thesis/chapters/ch2-problem.tex
Edit ch3

@@
 
-This technique overcomes the early blockchain scalability issue by delegating the transaction validation to the Byzantine consensus protocol (e.g. PBFT in ByzCoin~\cite{kogias2016enhancing}).
+This technique overcomes the early blockchain scalability issue by delegating the transaction validation to the Byzantine consensus protocol (e.g., PBFT in ByzCoin~\cite{kogias2016enhancing}).
 A major tradeoff of such systems is that they cannot guarantee correctness when there is a large number of malicious nodes (but less than a majority).
@@
 Further, due to the fact that these systems must reach consensus on all transactions, none of them achieves horizontal scalability.
-
-% Our approach share many similarities with the hybrid systems.
-% First, we also elect a committee (facilitators) to drive consensus.
-% But we do not have proof-of-work for Sybil defence because we believe 
-% it is possible to do it efficiently, e.g. using NetFlow~\cite{pimotte}.
-% Secondly, our use of CP blocks and ACS is also unique.
-% This creates a much higher throughput and enables censorship resiliance as mentioned earlier.
-% For instance, ByzCoin performs just below 1000 TPS with a thousand nodes whereas we peak at 8000 TPS.
 

### 2017-08-18 thesis/chapters/ch3-model.tex
Edit ch3

@@
 \label{ch:model}
-
-% As was shown in~\Cref{ch:problem}, the requirements are the following.
-% \begin{itemize}
-%     \item Global consensus
-%     \item Security and fault tolerance
-%     \item Performance and scalability
-% \end{itemize}
-
-% As mentioned in the Introduction, the primary goal is performance and scalability.
-% Having a scalable blockchain system while still keeping global consensus
-% allows the system to be ubiquitous and realise the full potential of blockchain.
-
-% The secondary goal is to design an application neutral system.
-% In particular, it should act as a backbone that provides the building blocks of blockchain based applications.
-% It should be possible to buld any kind of application on top our system.
-% Further, we do not impose on a consensus algorithm, as long as it satisfies the properties of atomic broadcast which we describe in~\Cref{sec:overview-cons}.
-
-% The third and final goal is security.
-% Our system should be unaffected in the presence of powerful adversaries.
-% In particular, adversaries are Byzantine meaning that they can have arbitrary behaviour.
-% Thus anything is possible from simply ommitting messages to colluding with eachother in order to undermine the whole system.
 
@@
 
-\subsection{Extended TrustChain}
+\subsection*{TrustChain and Extended TrustChain}
 Extended TrustChain naturally builds on top of the standard TrustChain. 
 Thus we first describe the standard TrustChain.
-Our description has minor differences compared to the description in~\cite{multichain}.
-This is to help with the description of the extended TrustChain.
-We remark the difference when it occurs.
-However, the two descriptions are functionally the same.
+Our description has minor differences compared to the description in~\cite{multichain},
+which is to help with the description of the Extended TrustChain.
+We remark the differences when they occur.
 
-\subsubsection*{Standard TrustChain}
+% \subsubsection*{Standard TrustChain}
 In TrustChain, every node has a ``personal'' chain. 
@@
 a new TX block is generated for both $A$ and $B$ and is appended to their respective chains.
-The TX block must have a valid hash pointer pointing to the previous block
-and a reference\footnote{This is different from the original TrustChain definition found in~\cite{multichain}.
-In there, a TX block has two outgoing edges which are hash pointers to the two parties involved in the transaction.
+The TX block must have a valid hash pointer pointing to the previous block and a reference\footnote{This definition is different from the original TrustChain definition found in~\cite{multichain}.
+    In the original definition, a TX block has two outgoing edges which are hash pointers to the two parties involved in the transaction.
     This work uses one outgoing edge and a reference.} to its \emph{pair} on $B$'s chain.
@@
     \centering
-    \caption{Every block is denoted by $t_{i,j}$, where $i$ is the node ID and $j$ is the sequence number of the block.
-    Thus we have three nodes and three corresponding chains in this example.
+    \caption{
+    A TrustChain example with three nodes.
+    Every block is denoted by $t_{i,j}$, where $i$ is the node ID and $j$ is the sequence number of the block.
     The arrows represent hash pointers and the dotted lines represent references.
@@
 then every chain effectively forms a singly linked list.
-However, if a node violates the rules, then a \emph{fork} may happen.
+However, if a node violates the rules,
+either by accident or for malicious intent,
+then a \emph{fork} may happen.
 That is, there may be more than one TX block with a hash pointer pointing back to the same block.
@@
 
-\subsubsection*{Extended TrustChain}
-We are now ready to explain the Extended TrustChain.
+% \subsubsection*{Extended TrustChain}
 The primary difference is the introduction of a new type of block---checkpoint (CP) block.
-In contrast to TX blocks, CP blocks do not store transactions or contain references.
-Their purpose is to capture the state of the chain and the state of the whole system.
-In particular, the state of the chain is captured with a hash pointer.
-The state of the whole system is captured in the content of the CP block,
-namely as a digest of the latest \emph{consensus result} which we explain in \Cref{sec:overview-cons}.
+The most important function of the CP block is to represent the state of a node, in the form of a hash pointer.
+A collection of CP blocks from all nodes represents the state of the whole system.
+Therefore, nodes need to reach consensus on CP blocks using some consensus protocol, which we explain next.
 A visual representation is shown in \Cref{fig:trustchain-bad-cp}.
@@
     \centering
-    \caption{The circles represent CP blocks,
+    \caption{
+    An example of Extended TrustChain.
+    The circles represent CP blocks,
     they also have hash pointers (arrow) but do not have references (dotted line).
@@
 
-\subsection{Consensus protocol}
-\label{sec:overview-cons}
+\subsection*{Consensus protocol}
 
-% The consensus protocol uses an existing Byzantine consensus algorithm (explained in detail in \Cref{sec:xyz}) and runs it continuously in rounds.
-The consensus protocol can be seen as a technique of running infinitely many rounds of some 
-Byzantine consensus algorithm\footnote{More accurately it is ACS or asynchronous subset consensus, we describe ACS in \Cref{sec:acs-background}.},
+The consensus protocol can be seen as a technique of running infinitely many times of some 
+Byzantine consensus algorithm\footnote{More accurately it is asynchronous common subset (ACS) which we describe in \Cref{sec:acs-background}.},
 starting a new execution immediately after the previous one is completed.
-This is necessary because blockchain systems always need to reach consensus on new values proposed by nodes in the system,
+This approach is necessary because blockchain systems always need to reach consensus on new values proposed by nodes in the system,
 or CP blocks in our case.
@@
 The high communication complexity of Byzantine consensus algorithms prohibits us from running it on a large network.
-Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks from every other node and use them as the proposal.
-The facilitators are elected using a \emph{luck value}, which is computed using $\textsf{H}(\C_r || pk_u)$,
-where $\C_r$ is the consensus result 
-(which can be seen as the set union of all the CP blocks collected by the facilitators)
-in round $r$ and $pk_u$ is the public key of node $u$.
-Intuitively, the election is guaranteed to be random 
-because the output of a cryptographically secure hash function is unpredictable and $\C_r$ cannot be determined in advance.
-Finally, the elected facilitators run a Byzantine consensus algorithm to agree on a set of CP blocks.
-The next round begins when the new facilitators are computed from these CP blocks.
+Thus, at the beginning of every Byzantine consensus algorithm execution,
+we randomly elect a set of nodes---called facilitators---to collect CP blocks from every other node and use them as the input to the Byzantine consensus algorithm, and then run it.
+After running the algorithm,
+the facilitators output a set of CP blocks which we call the consensus result.
+The result is then propagated to the network and the next algorithm execution begins.
 
-A visual explanation can be found in \Cref{app:consensus-example},
-it walks through the steps needed for a node to be selected as a facilitator using an example.
-
-\subsection{Transaction and validation}
+\subsection*{Transaction and validation}
 The TX protocol is a simple request and response protocol.
@@
 
-The consensus and transaction protocol by themselves do not provide a mechanism to detect tampering.
-Thus we need a validation protocol to counteract malicious behaviour.
-When a node wishes to validate one of its TX, it asks the counterparty for the \emph{agreed fragment} of the TX.
-An agreed fragment of a TX block is a section of the chain beginning and ending with CP blocks that contain the TX block,
+The consensus and transaction protocol by themselves do not provide a mechanism to detect malicious behaviour such as tampering the TrustChain.
+Thus, we need a validation protocol to counteract such behaviour.
+When a node wishes to validate one of its transactions, it asks the counterparty for the \emph{agreed fragment} of the transaction.
+Which is a fragment of the counterparty's chain beginning and ending with CP blocks but contains the TX block belonging to that transaction,
 the CP blocks must also be in consensus.
-Upon the counterparty's response, the node checks that the CP blocks are indeed in consensus,
+Upon the counterparty's response, the node checks whether the CP blocks are in some consensus result,
 the hash pointers are valid and his TX is actually in the fragment.
 The TX is valid if these conditions are satisfied.
-Intuitively, this works because it is hard (because a cryptographically secure hash function is second-preimage resistant)
+Intuitively, this works because it is hard (because a cryptographically secure hash function is collision resistant)
 to create a different chain that begins and ends with the same two CP blocks.
 
-\subsection{Combined protocol}
-\label{sec:combined-protocol}
-The final result is essentially the concurrent composition of the three aforementioned protocols,
+\subsection*{Complete protocol}
+The complete protocol is essentially the concurrent composition of the three aforementioned protocols,
 all making use the Extended TrustChain data structure.
@@
 For instance, a node may start the consensus protocol, and while it is running, create new transactions and validate old transactions.
-By the time the consensus protocol is done, the new CP block is added to whatever the state that the chain is in.
+By the time the consensus protocol is done, the new CP block is added to whatever state the chain is in.
 It is not necessary to lock the chain while the consensus protocol is running and then unlock it afterwards.
@@
 \begin{itemize}
-\item Lower case (e.g. $x$) denotes a scalar object or a tuple.
-\item Upper case (e.g. $X$) denotes a set or a constant.
-\item Sans serif (e.g. $\textsf{fn}(\cdot)$) denotes a function.
-\item Monospace (e.g. $\texttt{ack}$) denotes message type.
+\item Lower case (e.g., $x$) denotes a scalar object or a tuple.
+\item Upper case (e.g., $X$) denotes a set or a constant.
+\item Sans serif (e.g., $\textsf{fn}(\cdot)$) denotes a function.
+\item Monospace (e.g., $\texttt{ack}$) denotes message type.
 \end{itemize}
-Further, we use $a || b$ to denote concatenation of the binary representations of $a$ and $b$.
+Further, we use $a || b$ to denote the concatenation of the binary representations of $a$ and $b$.
 
@@
 The adversary has full control of the delivery schedule and the message ordering of all messages.
-But they are not allowed to drop messages except for their own\footnote{
-    Reliability can be achieved in unreliable networks by resending messages or using some error correction code.
-}.
+% But they are not allowed to drop messages except for their own\footnote{
+%     Reliability can be achieved in unreliable networks by resending messages or using some error correction code.
+% }.
 
@@
 $n \ge 3t + 1$ must hold.
-This is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$~\cite{pease1980reaching}.
-Further, the inequality $N \ge n + t$ must also hold.
-This is due to our system design, which becomes clear in~\Cref{sec:consensus-phase}.
+This inequality is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$~\cite{pease1980reaching}.
+Further, the inequality $N \ge n + t$ must also hold so that there are enough honest nodes in the network,
+the exact reason becomes clear in~\Cref{sec:consensus-phase}.
 
@@
 but if the node recovers, then the adversary must forget the private key.
-This is realistic because otherwise the adversary can eventually learn all the private keys and sabotage the system.
+This restriction is realistic because otherwise the adversary can eventually learn all the private keys and sabotage the system.
 
@@
 If $T_u$ is the set of all TX blocks in $B_u$ and $C_u$ is the set of all CP blocks is $B_u$,
-then it must be the case that $T_u \cup C_u = B_u$ and $T_u \cap C_u = \varnothing$.
+then $T_u \cup C_u = B_u$ and $T_u \cap C_u = \varnothing$.
 The notation $b_{u, i}$ is generic over the block type.
-We assume there exists a function $\textsf{typeof}: B_u \rightarrow \{ \tau, \gamma \}$ that returns the type of the block,
-where $\tau$ represents the TX type and $\gamma$ represents the CP type.
+% We assume there exists a function $\textsf{typeof}: B_u \rightarrow \{ \tau, \gamma \}$ that returns the type of the block,
+% where $\tau$ represents the TX type and $\gamma$ represents the CP type.
 
@@
 
-Our consensus protocol runs in rounds as discussed in \Cref{sec:system-overview}.
-Every round is identified by a round number $r$, which is incremented on every new round.
+Our consensus protocol runs in rounds,
+where the first round is defined to be 1 and it is incremented after every execution of the consensus protocol.
 The consensus result is a tuple, i.e. 
@@
 
-If there exist a tuple $\langle c_{u,a}, c_{u, b} \rangle$ for a TX block $t_{u, i}$,
-where 
-$$a = \argmin_{k, k < i, \textsf{typeof}(b_{u,k}) = \gamma}(i - k)$$
-$$b = \argmin_{k, k > i, \textsf{typeof}(b_{u,k}) = \gamma}(k - i),$$
+If there exist a tuple $\langle c_{u,a}, c_{u, b} \rangle$ for a TX block $t_{u, i}$, where
+\begin{itemize}
+\item $c_{u, a}$ is the closest CP block to $t_{u, i}$ with a lower sequence number and
+\item $c_{u, b}$ is the closest CP block to $t_{u, i}$ with a higher sequence number,
+\end{itemize}
 then $\langle c_{u,a}, c_{u, b} \rangle$ is the enclosure of $t_{u, i}$.
-Intuitively, that is the two closest CP blocks that surround $t_{u, i}$.
-Note that $c_{u, a}$ is the more recent CP block.
-Also, some TX blocks may not have any enclosure, then its enclosure is $\bot$.
+Some TX blocks may not have any enclosure, then their enclosure is $\bot$.
 Agreed enclosure is the same as enclosure with an extra constraint where the CP blocks must be in some consensus result $\C_r$.
@@
 
-Note that the ``closest CP blocks'' property must also apply to agreed enclosure. 
-Suppose a chain is in the form
-$\{c_{i}, c_{i+1}, t_{i+2}, c_{i+3}\}$\footnote{Usually the notation is of the form $c_{u, i}$, but the node identity is not important here so we simplify it to $c_{i}$}
-and $c_{i}, c_{i+1}, c_{i+3}$ are in $\C_r, \C_{r+1}, \C_{r+3}$ respectively,
-then the agreed enclosure of $t_{i+2}$ is $\langle c_{i+1}, c_{i+3}\rangle$ and cannot be $\langle c_{i}, c_{i+3}\rangle$.
+% Note that the ``closest CP blocks'' property must also apply to agreed enclosure. 
+% Suppose a chain is in the form
+% $\{c_{i}, c_{i+1}, t_{i+2}, c_{i+3}\}$\footnote{Usually the notation is of the form $c_{u, i}$, but the node identity is not important here so we simplify it to $c_{i}$}
+% and $c_{i}, c_{i+1}, c_{i+3}$ are in $\C_r, \C_{r+1}, \C_{r+3}$ respectively,
+% then the agreed enclosure of $t_{i+2}$ is $\langle c_{i+1}, c_{i+3}\rangle$ and cannot be $\langle c_{i}, c_{i+3}\rangle$.
 
@@
 
-Our scalable consensus protocol runs on top of Extended TrustChain.
-It uses an unmodified asynchronous subset consensus (ACS) algorithm as the key building block.
+Our consensus protocol runs on top of Extended TrustChain.
+It uses an unmodified asynchronous common subset (ACS) algorithm as the key building block.
 The objectives of the protocol are to
@@
     and have an unbiased election of facilitators.
-Concretely, we define the necessary properties as follows.
+Concretely, we define the following.
 \begin{definition}
 \label{def:consensus}
-\textbf{\emph{Properties of the consensus protocol}}
+\textbf{\emph{TrustChain consensus}}
 
-$\forall r \in \mathbb{N}$, the following properties must hold.
+A TrustChain consensus protocol is correct if the following holds for every round $r$:
 \begin{enumerate}
@@
                 $\C_r$ is a tuple but we abuse the notation here by writing $|\C_r|$ to mean the number of CP blocks in the second element of $\C_r$.}.
-                \item $\F_r$ must contain at least $n - t$ honest nodes and
                 \item $|\F_r| = n$.
@@
 Firstly, they are properties for every node in the network and not just the facilitators.
-Secondly, they must be satisfied for all rounds,
-because a failure in one round will lead to failures in sebsequent rounds.
+Secondly, they must be satisfied for all rounds because a failure in one round will lead to failures in subsequent rounds.
 
 Before describing the protocol in detail,
-we take a brief detour to give background on the ACS protocol as it is the primary building block.
+we take a brief detour to give background on ACS as it is the primary building block.
 Then we move on to describe the two phases of our consensus protocol---bootstrap phase and consensus phase.
@@
     \item \emph{Agreement:}
-        If a correct node outputs $v$, then every node outputs $v$.
+        If a correct node outputs $v$, then every correct node outputs $v$.
     \item \emph{Validity:}
-        The decision value must be the input value of a node.
+        The decision value must be the input value of a correct node.
     \item \emph{Termination:}
@@
 \end{definition}
-A node under Byzantine fault means that it can have arbitrary behaviour.
-For example not sending a message or colluding with other Byzantine nodes to undermine the entire system.
-Note that the decision is on a single value.
-This is in contrast to ACS which we describe next.
+A node under Byzantine fault means that it can have arbitrary behaviour,
+for example, not sending a message or colluding with other Byzantine nodes to undermine the entire system.
+Note that the decision is on a single value,
+which is in contrast to ACS that we describe next.
 
@@
 ACS has the nice property of censorship resilience when compared to other consensus algorithms.
-For instance, Hyperledger and Tendermint uses a variant of Practical Byzantine Fault Tolerance (PBFT)~\cite{castro1999practical} as their consensus algorithm.
+For instance, Hyperledger and Tendermint use a variant of Practical Byzantine Fault Tolerance (PBFT)~\cite{castro1999practical} as their consensus algorithm.
 In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectively filter transactions.
@@
 
-To bootstrap, imagine that there is some bootstrap oracle, that initiates the code on every node.
-The code satisfies all the properties in \Cref{def:consensus}.
-Namely, every node has the same set of valid facilitators $\F_1$ that are randomly chosen.
+To bootstrap, imagine that there is some bootstrap oracle that initiates the correct program on every node,
+meaning that it satisfied the properties in \Cref{def:consensus}.
 This concludes the bootstrap phase.
@@
 The consensus phase begins when $\F_r$ is available to all the nodes.
-Note that $\F_r$ indicates the facilitators that were elected using results of round $r$ and are responsible for driving the ACS protocol in round $r + 1$.
+Note that $\F_r$ indicates the facilitators that were elected using results of round $r$ and are responsible for driving the ACS algorithm in round $r + 1$.
 The goal is to reach agreement on a set of new facilitators $\F_{r + 1}$ that satisfies the four properties in \Cref{def:consensus}.
@@
 There are two scenarios in the consensus phase.
-First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators.
+First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators,
+where $c_{u, h}$ is the most recent CP block.
 Second, if the node is a facilitator, it waits until $N - t$ messages of type \texttt{cp\_msg} are received.
@@
 
-The core of the consensus phase is the ACS protocol.
-While any ACS protocol that satisfies the standard definition will work,
-we use a simplification of HoneyBadgerBFT~\cite{miller2016honey} as our ACS protocol
+The core of the consensus phase is the ACS algorithm.
+While any ACS algorithm that satisfies the standard definition will work,
+we use a simplification of HoneyBadgerBFT~\cite{miller2016honey} as our ACS algorithm 
 because it is a consensus algorithm designed for blockchain systems.
@@
 
-Continuing, when $\F_{r}$ reaches agreement on $\C_{r+1}$,
+Continuing, when $\F_{r}$ finish the ACS execution and reach agreement on $\C_{r+1}$,
 they immediately broadcast two messages to all the nodes---
@@
 Every node then sends a completion messages (in the form of \texttt{cp\_msg}) to the new collective authority to indicate that they are ready for the next pulse.
+
+A visual explanation can be found in \Cref{app:consensus-example},
+it walks through the steps needed for a node to be selected as a facilitator using an example.
 
@@
 otherwise the node calling the function would have no point of reference.
-This is not difficult to achieve,
-because typically the caller is $u$, so it knows its own TX block and the corresponding agreed fragment.
+This assumption is reasonable in practice because typically the caller is $u$,
+so it knows its own TX block and the corresponding agreed fragment.
 If the caller is not $u$, it can always query for the agreed fragment that contains the transaction of interest from $u$.
@@
 we essentially check whether the fragment is the one that the verifier needs.
-If it is not, then the verifier cannot make any decision and return \emph{unknown}.
-This is likely to be the case for new transactions because the result of $\textsf{agreed\_fragment}(\cdot)$ would be $\bot$.
+If it is not, then the verifier cannot make any decision and return \emph{unknown},
+which is likely to be the case for new transactions because the result of $\textsf{agreed\_fragment}(\cdot)$ would be $\bot$.
 The next two conditions check for tampering or missing blocks, if any of these misconducts are detected, then the TX is invalid.
@@
 Every time the message is sent to $f$ random neighbours, $f$ is often called the fan-out.
-The upside of using epidemic protocol is that the communication cost is distributed more evenly between nodes.
-This is especially true with a lazy push approach where the node who just received a message would push the message ID (e.g. digest) to its $f$ random neighbours,
+The upside of using epidemic protocol is that the communication cost is distributed more evenly between nodes,
+which is especially true with a lazy push approach where the node who just received a message would push the message ID (e.g., digest) to its $f$ random neighbours,
 and only push the full message if the neighbour explicitly requests the message~\cite{leitao2007epidemic}.
@@
 whereas broadcasting uses constant time.
-Another downside of some epidemic protocols (e.g. eager push) is that it is difficult to guarantee delivery.
+Another downside of some epidemic protocols (e.g., eager push) is that it is difficult to guarantee delivery.
 It is especially true when the parameters are not chosen correctly in a network that is only partially connected (but every node is nevertheless reachable).
@@
 
-At the start of our consensus phase (\Cref{sec:consensus-phase}), facilitators must wait for $N-f$ \texttt{cp\_msg} messages.
-This is the only place where we used $N$ as a parameter.
+At the start of our consensus phase (\Cref{sec:consensus-phase}), facilitators must wait for $N-f$ \texttt{cp\_msg} messages,
+which is the only place where we used $N$ as a parameter.
 To introduce timing, instead of waiting for $N-f$ messages, we wait for some time $D$,
@@
 \subsection{Total fork detection}
-The validation algorithm guarantees that there are no forks within a single agreed fragment.
-This is sufficient for some applications such as proving the existence of some information.
+The validation algorithm guarantees that there are no forks within a single agreed fragment,
+which is sufficient for some applications such as proving the existence of some information.
 However, for applications such as digital cash where every block depends on one or more previous blocks,

### 2017-08-19 thesis/chapters/ch3-model.tex
Edit ch4 and 5

@@
 
-Our consensus protocol runs on top of Extended TrustChain.
+Our consensus protocol $\Pi_c$ runs on top of Extended TrustChain.
 It uses an unmodified asynchronous common subset (ACS) algorithm as the key building block.
@@
 \label{def:consensus}
-\textbf{\emph{TrustChain consensus}}
+\textbf{\emph{TrustChain consensus protocol}}
 
@@
 
-\subsection{Background on asynchronous subset consensus}
+\subsection{Background on asynchronous common subset}
 \label{sec:acs-background}
@@
 \label{def:acs}
-\textbf{\emph{Asynchronous subset consensus}}
+\textbf{\emph{Asynchronous common subset}}
 
@@
 
-The TX protocol, shown in \Cref{alg:tx-proto}, is run by all nodes.
+The TX protocol $\Pi_t$, shown in \Cref{alg:tx-proto}, is run by all nodes.
 It is also known as True Halves, first described informally by Veldhuisen~\cite[Chapter~3.2]{truehalves}.
@@
 \begin{algorithm}
-    \caption{The TX protocol which runs in the context of node $u$.}
+    \caption{The TX protocol $\Pi_t$ runs in the context of node $u$.}
     \label{alg:tx-proto}
@@
 \subsection{Validation protocol}
-Our validation protocol,
+Our validation protocol $\Pi_v$,
 shown in~\Cref{alg:vd-proto},
@@
 \label{def:validation}
-\textbf{\emph{Properties of the validation protocol}}
+\textbf{\emph{TrustChain validation protocol}}
 
+A TrustChain validation protocol is correct if the following properties hold.
 \begin{enumerate}
@@
 \begin{algorithm}
-\caption{Validation protocol which runs in the context of $u$}
+\caption{Validation protocol $\Pi_v$ which runs in the context of $u$}
 \label{alg:vd-proto}

### 2017-08-19 thesis/chapters/ch4-analysis.tex
Style improvements in thesis
Edit ch4 and 5

@@
 
-Up to this point, we described our system specification in detail.
-Of course, specification alone does not establish any truths.
-In this chapter, we evaluate our system analytically.
-First, we show it has the desired properties.
-That is, the properties of the consensus protocol and the validation protocol (\Cref{def:consensus} and \Cref{def:validation} respectively) should hold.
-Then we analyse the performance, especially the throughput,
-and show that it out performs early blockchain systems.
-Finally, we consider the case where the $n \ge 3t + 1$ assumption is violated and show the effect of it in our system.
+In this chapter,
+we evaluate our system analytically to check whether it correctly implements the TrustChain consensus protocol (\Cref{def:consensus})
+and the TrustChain validation protocol (\Cref{def:validation}) in~\Cref{sec:correctness}.
+Further, we analyse the performance and especially the throughput to understand the scalability characteristics in~\Cref{sec:performance-analysis}.
+Finally, we consider the case where the $n \ge 3t + 1$ assumption is violated and show the effect of it in our system in~\Cref{sec:highly-adversarial}.
 
 \section{Correctness in the presence of faults}
-Our first objective is to show that \Cref{def:consensus} holds for our consensus protocol.
-Then, building on top of it, we show \Cref{def:validation} holds for the validation protocol.
+\label{sec:correctness}
+Our first objective is to show that our consensus protocol $\Pi_c$ described in~\Cref{sec:cons-protocol} is in fact a TrustChain consensus protocol defined in~\Cref{def:consensus}.
+Then, building on top of it, we show that our validation protocol $\Pi_v$ described in~\Cref{sec:vd-protocol} conforms to the TrustChain validation protocol defined in~\Cref{def:validation}.
 The resulting theorem shows that only using CP blocks in the consensus algorithm implies consensus on TX blocks.
@@
 
-We begin our analysis by establishing truths on the four properties in \Cref{def:consensus},
+We begin our analysis by establishing the four properties in \Cref{def:consensus},
 namely agreement, validity, fairness and termination for an arbitrary round.
@@
 \label{lemma:agreement}
-For an arbitrary round $r$
-if $\F_{r}$ is known by all correct nodes and one correct node outputs a list of facilitators $\F_{r+1}$,
+If for an arbitrary round $r$,
+$\F_{r}$ is known by all correct nodes and one correct node outputs a list of facilitators $\F_{r+1}$,
 then all correct nodes output $\F_{r+1}$.
@@
 \label{lemma:validity}
-For an arbitrary round $r$,
-if $\F_r$ is known by all correct nodes and any correct node outputs $\F_{r+1}$,
-then (a) $|\C_{r+1}| \ge N - t$ must hold for the $\C_{r+1}$ which was used to create $\F_{r+1}$,
-(b) $\F_{r+1}$ must contain at least $n - t$ honest nodes and
-(c) $|\F_{r+1}| = n$.
+If for an arbitrary round $r$,
+$\F_r$ is known by all correct nodes and any correct node outputs $\F_{r+1}$,
+then (a) $|\C_{r+1}| \ge N - t$ must hold for the $\C_{r+1}$ which was used to create $\F_{r+1}$ and (b) $|\F_{r+1}| = n$.
 \end{lemma}
@@
 From the validity property of ACS, we know the output must contain the input of at least $n - 2t$ nodes.
-But $n -t$ facilitators must have received $N - t$ valid CP blocks, so $|\C_{r+1}| \ge N-t$, this proves (a).
-There are at least $n-t$ honest nodes in $\F_{r+1}$ which follows from our assumption, this proves (b).
-Finally, since $N-t \ge (n+t) -t = n$ and $\textsf{get\_facilitators}(\cdot)$ outputs $n$ items, $|\F_{r+1}| = n$ and this proves (c).
+But $n -t$ facilitators must have received $N - t$ valid CP blocks, so $|\C_{r+1}| \ge N-t$ and this proves (a).
+Since $N-t \ge (n+t) -t = n$ and $\textsf{get\_facilitators}(\cdot)$ outputs $n$ items, so $|\F_{r+1}| = n$ and this proves (b).
 \end{proof}
@@
 \label{lemma:fairness}
-For an arbitrary round $r$,
-if $\F_r$ is known by all correct nodes then
+If for an arbitrary round $r$,
+$\F_r$ is known by all correct nodes then
 every node with a CP block in $\C_{r+1}$ should have an equal probability to be elected as a facilitator in $\F_{r+1}$.
@@
 \label{lemma:termination}
-For an arbitrary round $r$,
-if $\F_r$ is known by all correct nodes then
+If for an arbitrary round $r$,
+$\F_r$ is known by all correct nodes then
 every correct node eventually outputs some $\F_{r + 1}$.
@@
 From Lemmas \ref{lemma:agreement}, \ref{lemma:validity}, \ref{lemma:fairness} and \ref{lemma:termination},
-we have shown that the 4 properties of \Cref{def:consensus} holds when assuming the existence of some $\F_r$.
-Thus, to proof the whole of \Cref{def:consensus},
-we need to proof these 4 properties under the universal quantifier.
+we have shown that the four properties of \Cref{def:consensus} holds when assuming the existence of some $\F_r$.
+Now we need to proof these properties under the universal quantifier on $r$.
 We do this using mathematical induction.
@@
 \label{theorem:consensus}
-For all rounds,
-the consensus protocol satisfies agreement, validity, fairness and termination (\Cref{def:consensus}).
+For all rounds, $\Pi_c$ implements a TrustChain consensus protocol (\Cref{def:consensus}).
 \end{theorem}
 \begin{proof}
-We proof using mathematical induction.
+We prove using mathematical induction.
 
@@
 For the inductive step,
-we assume that the 4 properties hold in round $r$ and prove that they also hold in round $r + 1$.
+we assume that the four properties hold in round $r$ and prove that they also hold in round $r + 1$.
 Using Lemmas \ref{lemma:agreement}, \ref{lemma:validity}, \ref{lemma:fairness} and \ref{lemma:termination},
@@
 \subsection{Correctness of the validation protocol}
-The consensus protocol,
-which ensures consensus on CP blocks and facilitators,
-is the backbone for consensus on transactions.
-In this section, we build on top of \Cref{theorem:consensus} to show that the properties except for liveness in \Cref{def:validation} can be satisfied.
+ $\Pi_c$ ensures consensus on CP blocks and facilitators,
+which makes it the building block for reaching consensus on transactions.
+In this section, we build on top of \Cref{theorem:consensus} to show that our validation protocol described in \Cref{sec:vd-protocol} has the agreement and validity properties of a TrustChain validation protocol defined in \Cref{def:validation}.
 
@@
 % (Agreement of the validation protocol)
-The validation protocol satisfies agreement and validity properties.
+$\Pi_v$ satisfies the agreement and validity properties of a TrustChain validation protocol (\Cref{def:validation}).
 \end{theorem}
@@
 \Cref{theorem:validation-agreement} is a major result that allows significantly improved performance over traditional blockchain systems,
-but it does not have all the properties of a typical Byzantine consensus protocol.
-Now we show a negative result, where the liveness property of \Cref{def:validation} cannot be attained.
-Meaning that transactions with adversaries cannot always be validated.
+but it does not have all the properties of a typical Byzantine consensus problem.
+Now we show a negative result, where the liveness property of \Cref{def:validation} cannot be attained in $\Pi_v$,
+meaning that transactions made with adversaries cannot always be validated.
 
@@
 \label{sec:performance-analysis}
-This section aims to analytically answer the scalability aspect of our research question.
-That is, does the global throughput increase linearly with respect to the population size?
-We begin by looking at the communication and time complexity of the consensus protocol,
+This section aims to analytically answer whether our system is horizontally scalable.
+We begin by looking at the communication and time complexity of the consensus protocol $\Pi_c$,
 and then the bandwidth requirement for a single transaction.
@@
 \label{sec:cons-complexity}
-The consensus protocol can be seen as three parts,
+$\Pi_c$ can be seen as three parts,
 so the communication complexity is the sum of these parts.
@@
 \section{Effect of a highly adversarial environment}
+\label{sec:highly-adversarial}
 
@@
   \includegraphics[width=1.0\textwidth]{hypergeometric}
-  \caption{Plot of the probability of selecting more than $\lfloor \frac{n-1}{3} \rfloor$ black balls for
+  \caption{Probability of selecting more than $\lfloor \frac{n-1}{3} \rfloor$ black balls for
   different value of $t$ with $N$ fixed at 2000.}
@@
 
-The bound is not tight, but it is useful for picking parameters.
+The bound is not tight, but it is useful for picking parameters for a desired level of fault tolerance.
 For some fixed $n$, since $0 \le \alpha \le \lfloor \frac{n-1}{3} \rfloor/n < (\lfloor  \frac{n - 1}{3} \rfloor + 1) / n$,

### 2017-08-19 thesis/chapters/ch5-implementation.tex
Change code URL, etc.
Edit ch4 and 5

@@
 
-To build up the confidence of the horizontal scalability claim from our analytical results,
-we must verify the same property experimentally.
-Hence, we give an open source prototype implementation of our system.
-Our implementation runs on a network with over 1000 node.
-Each node sends transactions to each other autonomously, 
-which allows us to evaluate our analytical claims.
+We evaluate our system experimentally to ultimately understand its scalability characteristics as well as limitations.
+Concretely,
+we give an open source prototype implementation that runs on a network with over 1000 node.
+Nodes in our implementation make transactions autonomously with each other and then we collect the results for analysis.
 
-We begin this chapter with a description of the implementation in \Cref{sec:implementation}.
-Then, we move on to describing our experimental setup in \Cref{sec:experimental-setup} which includes a discussion on the system parameters and the physical infrastructure which we use.
-Finally, \Cref{sec:comms-cost-experiment} and onwards follows the same structure as the Performance analysis (\Cref{sec:performance-analysis}),
-where we analyse the performance experimentally rather than analytically and compare both results.
+This chapter begins with a description of the implementation in \Cref{sec:implementation}.
+Then, we describe our experimental setup in \Cref{sec:experimental-setup},
+which includes a discussion on the system parameters and the physical infrastructure which we use.
+Finally, \Cref{sec:comms-cost-experiment} and onwards follows the same structure as \Cref{sec:performance-analysis},
+where we analyse the performance experimentally rather than analytically and compare the results.
 
@@
 \begin{displayquote}
-\url{https://github.com/kc1212/consensus-thesis-code}
+\url{https://github.com/kc1212/checo}
 \end{displayquote}
@@
 transaction protocol and validation protocol---simultaneously and analyse the communication cost, the consensus duration and the global throughput.
-We investigate these properties under the following four parameters.
+We investigate these properties under the following parameters.
 \begin{enumerate}
-  \item The transaction rate $r_{\text{tx}}$ per node. This is not comparable to the others because it is fixed at 2 TX/s.
-        Nevertheless, it is a good value because, as we show later,
-        it hits a bottleneck in extreme cases which helps us understand the limitations of our design.
   \item The number of facilitators $n \in \{4, 8, \dots, 32\}$.
@@
 
-Finally, since Bitcoin transactions are approximately 500 bytes~\cite{txsize},
+Finally, the transaction rate $r_{\text{tx}}$ is fixed at 2 TX/s for every node.
+This value, as we show later,
+creates a bottleneck in extreme cases which helps us understand the limitations of our design.
+Since Bitcoin transactions are approximately 500 bytes~\cite{txsize},
 we use a uniformly random transaction size sampled between 400 and 600 bytes.
@@
   }
-  \caption{Round duration does not increase linearly with the population size. 
-  This is likely due to the additional hashing operation required for larger consensus result.}
+  \caption{Round duration does not increase linearly with the population size,
+  which is likely due to the additional hashing operation required for larger consensus result.}
   \label{fig:round-duration}
@@
 the magnitude of the communication cost is similar.
-Both peak at about 100 MB.
-This is expected because the consensus protocol is decoupled from the transaction protocol and the validation protocol.
-Finally, the rate for which the communication cost increases is higher when the number of facilitators is higher.
-This is also expected because the ACS algorithm has an $n^2$ term in its communication complexity.
+Both peak at about 100 MB,
+which is expected because the consensus protocol is decoupled from the transaction protocol and the validation protocol.
+Finally, the rate for which the communication cost increases is higher when the number of facilitators is higher,
+which is also expected because the ACS algorithm has a polynomial term with respect to $n$ in its communication complexity.
 
@@
 More interestingly, there is a large difference in communication cost between the two modes of transaction.
-When transacting with only one neighbour, the communication cost is low because only one agreed fragment needs to be communicated for every round in order to validate all transactions of that round.
-This is because agreed fragments are cached.
+When transacting with only one neighbour, the communication cost is low because only one agreed fragment needs to be communicated for every round in order to validate all transactions of that round,
+which is because agreed fragments are cached.
 On the other hand,
@@
 
-Note that the throughput decreases slightly as the number of facilitators increases.
-This is due to the additional communication cost for running ACS with a high number of facilitators.
+Note that the throughput decreases slightly as the number of facilitators increases,
+which is due to the additional communication cost for running ACS with a high number of facilitators.
 That is, if the network is congested then the nodes may not have enough bandwidth to send validation responses timely.
@@
 the magnitude of our throughput may not be self-evident at first glance.
-Recall that we fixed our $r_\text{tx}$ to 2, but how is it possible to have around 4800 transactions per second for 1200 nodes (which is 4 TX/s)?
-This is due to the way validated transactions are calculated.
+Recall that we fixed our $r_\text{tx}$ to 2, but how is it possible to have around 4800 transactions per second for 1200 nodes (4 TX/s)?
+This result is due to the way validated transactions are calculated.
 Transactions are between two parties, hence if every node makes two transactions per second,

### 2017-08-19 thesis/chapters/ch6-conclusion.tex
Style improvements in thesis
Edit ch4 and 5

@@
 
-In this work we introduced an application neutral blockchain system which we call Checo.
+In this work we introduced an application neutral blockchain system which we call \textsc{Checo}.
 We add checkpoint block to the existing TrustChain data structure for use in our consensus protocol.
@@
 
-The research question we asked in~\Cref{ch:problem} is the following.
+This work answers the research question stated in~\Cref{ch:problem}.
 \begin{displayquote}
@@
 \end{displayquote}
-We answer it in the affirmative.
-Fault tolerance is guaranteed if $n \ge 3t + 1$ by using ACS as a building block.
+Concretely, fault tolerance is guaranteed if $n \ge 3t + 1$ by using ACS as a building block.
 While $t$ may be small compared to the population size $N$,

### 2017-08-20 thesis/appendices/consensus-example.tex
Write abstract
Edit ch4 and 5
Edit everything

@@
-\chapter{Consensus example}
+\chapter{Consensus protocol example}
 \label{app:consensus-example}
+
+Visualisations of the different stages in the consensus protocol is shown in Figures \ref{fig:trustchain-1} to \ref{fig:trustchain-4}.
+The captions detail every step.
 
@@
     \centering
-    \caption{We begin in a state where some facilitators $\F_{r-2}$ just agreed on the consensus result $\C_{r-1}$ but have not yet propogated it yet.}
+    \caption{We consider three nodes and begin in a state where some facilitators $\F_{r-2}$ just agreed on the consensus result $\C_{r-1}$ but have not yet propogated it yet.}
     \label{fig:trustchain-1}
@@
     \centering
-    \caption{The facilitators propogates $\C_{r-1}$. Upon receiving it, the nodes perform two tasks;
+    \caption{The facilitators propogates $\C_{r-1}$. Upon receiving it, the nodes perform two tasks:
     (1) elect new facilitators $\F_{r-1}$ by selecting the first $n$ nodes ordered by $\textsf{H}(\C_{r-1} || pk)$,
     and (2) create new CP blocks---$c_{a, 6}, c_{b, 4}$ and $c_{c, 8}$.
-    $\textsf{H}(\cdot)$ is a cryptographically secure hash function and $pk$ is the public key of the respective node.}
+    For example, $c_{a, 6} = \langle \textsf{H}(t_{a, 5}), 6, \textsf{H}(\C_{r-1}), r-1, sig_a\rangle$.}
     \label{fig:trustchain-2}
@@
     \caption{Finally, when $\F_{r-1}$ decides on $\C_r$ (which should contain $c_{a, 6}, c_{b, 4}$ and $c_{c, 8}$) and disseminates it.
-    Nodes compute new facilitators $\F_{r}$ and create new CP blocks.}
+    Nodes compute new facilitators $\F_{r}$ and create new CP blocks.
+    For example, $c_{c, 11} = \langle \textsf{H}(t_{c, 10}), 11, \textsf{H}(\C_r), r, sig_c \rangle$.}
     \label{fig:trustchain-4}

### 2017-08-20 thesis/chapters/ch1-intro.tex
Edit everything

@@
 We live in a world where technologies have become vital for our welfare and success.
-The internet, for instance, gave us the ability of efficiently exchange information on a global scale.
+The internet, for instance, gave us the ability to efficiently exchange information on a global scale.
 However, in contrast to its original design, the internet and in particular the world wide web is becoming increasingly centralised.
-The domain name system (root name server), certificate authorities (root CA), to name a few, carry enormous responsibilities and are a central point of failure.
+The root name server in the domain name system,
+the root certificate authority in our public key infrastructure,
+to name a few, carry enormous responsibilities and are a central point of failure.
 The  same can be said for many other services such are online marketplaces,
@@
 If $C$ can simultaneously claim that she transferred 5 units to $A$ but also 5 units to $B$, then the system is inconsistent.
-Bitcoin and many of its derivatives (e.g., Litecoin\footnote{\url{https://litecoin.org}} and Dogecoin\footnote{\url{http://dogecoin.org}}) solve the inconsistency problem with a consensus algorithm.
+Bitcoin and many of its derivatives (e.g., Litecoin\footnote{\url{https://litecoin.org}} and Dogecoin\footnote{\url{http://dogecoin.com}}) solve the inconsistency problem with a consensus algorithm.
 The goal of the algorithm is to reach agreement on a set of valid transactions. 
@@
 Consequently, it is unlikely for more than one blockchain to exist in the network for a long period of time.
-Thus every party sees a consistent blockchain which solves the double spending problem.
+Thus, every party sees a consistent blockchain which solves the double spending problem.
 % Further, due to the reward mechanism,
@@
 To achieve transaction rates higher than that, it will lead to unfairness for nodes that do not have a good network connection.
-Thus fundamental changes are necessary to run at the scale of centralised payment processors such as Visa,
+Thus, fundamental changes are necessary to run at the scale of centralised payment processors such as Visa,
 which is in the order of tens of thousands transactions per second~\cite{visa}.
@@
 then the owner would report the incident to some central authority, e.g., the police.
-On the contrary, in blockchain systems, every transaction is effectively sent to the miners,
+On the contrary, in blockchain systems, every transaction, regardless of whether it is valid or invalid, are sent to the miners,
 which can be seen as a collective authority.
@@
 but nevertheless verifiable at a later stage by any node in the network.
-We call our protocol \textsc{Checo}, derived from \emph{che}ckpoint \emph{co}nsensus.
-Informally, our system works as follows.
-Every node stores its own blockchain and every block is one transaction.
+We call our system \textsc{Checo}, derived from \emph{che}ckpoint \emph{co}nsensus.
+From a high level, our system works as follows.
+Every node stores its own hash chain and every block contains only one transaction.
 We randomly select nodes in every round, the selected ones are called facilitators.
@@
 then that node can prove to any other node that it holds a set of transactions that compute (form a chain) to the checkpoint block.
-This immediately shows that those transactions are tamper-proof.
-The key difference from related work is the independence of the transaction validation phase.
+As a result, those transactions are tamper-proof.
+The key difference from the early blockchain systems is the independence of the transaction validation phase.
 Prior systems such a Bitcoin and Ethereum entangled the consensus algorithm with the transaction validation, restricting performance.
@@
     \item We formally introduce a blockchain system---\textsc{Checo}.
-        It uses individual hash chains and checkpoints on every node to achieve
+        It uses individual hash chains and checkpoint blocks on every node to achieve
         horizontally scalable in the general case for the first time.
-    \item We formally analyse \textsc{Checo} to ensure correctness as defined in our architecture.
+    \item We formally analyse \textsc{Checo} to ensure correctness as defined in our system architecture.
     \item We provide an implementation and then experiment with up to 1200 nodes,

### 2017-08-20 thesis/chapters/ch2-problem.tex
Edit everything

@@
 
-From the Introduction, we saw that early blockchain systems face a scalability problem.
+We saw that early blockchain systems face a scalability problem.
 In particular, the use proof-of-work is unlikely to reach a transaction rate suitable for ubiquitous use.
@@
 In this chapter, we describe the related work in the form of a mini taxonomy and identify areas for improvement and limitations.
-Then we state our research question which aims to overcome some of the limitations.
+Then, we state our research question which aims to overcome some of the limitations.
 
@@
 To understand the current state-of-the-art,
-we classify various blockchain systems by their scalability approach.
+we classify various block\-chain systems by their scalability approach,
+beginning with the baseline.
 
@@
 The level of consensus of a block containing many transactions is determined by how deep it is in the Bitcoin blockchain,
-also called the number of confirmation.
+which is also called the number of confirmation.
 The probability of a block being orphaned drops exponentially as the depth increases~\cite{bitcoin}.
@@
 In fact, prior to Bitcoin, there was no internet-scale consensus protocol.
-Bitcoin is also fault-tolerant because attackers can not out pace honest users in finding new blocks unless they have a majority of the hash power.
-As long as the majority is honest and reasonable measures are taken to prevent the selfish mining attack~\cite{eyal2014majority}.
+Bitcoin is also fault-tolerant because attackers can not out pace honest users in finding new blocks unless they have a majority of the computational power,
+as long as the majority is honest and reasonable measures are taken to prevent the selfish mining attack~\cite{eyal2014majority}.
 The disadvantage, however, is that transactions are never in consensus with a probability of 1---no consensus finality.
@@
 the performance is limited due to the fact that blocks are of fixed size and are generated at fixed intervals.
-Thus early blockchain systems do not achieve horizontal scalability.
+Thus early blockchain systems do not scale.
 
@@
 In the simplest case, 
-if two parties wish to make transactions,
+if two parties wish to make frequent transactions,
 they open a payment channel in the form of a multi-signature Bitcoin address (for two parties it would be a 2-of-2 signature address).
@@
 
-The advantage of such systems is that they act as add-ons to Bitcoin wish already has a large number of users.
+The advantage of such systems is that they act as add-ons to Bitcoin, which already has a large number of users.
 Thus, if enough of the network wish for it (by setting a new block version),
@@
 On the other hand, payment channel complicates user experience and leads to centralisation.
-As we mentioned earlier, each node must deposit some Bitcoins into a multi-signature account.
-The users must pick a suitable amount.
+As we mentioned earlier, each node must deposit some Bitcoins into a multi-signature account,
+so the users are responsible for picking a suitable amount.
 If the deposit is too low it would not allow large transactions.
@@
 In essence, they contain a fixed set of nodes (sometimes called validating peer) that run a Byzantine consensus algorithm to decide on new blocks.
-This technique is often used in permissioned system where the validators must be predetermined.
-For example, Hyperledger Fabric~\cite{cachin2016architecture}.
+This technique is often used in permissioned system where the validators must be predetermined,
+for example, Hyperledger Fabric~\cite{cachin2016architecture}.
 
@@
 \subsection{Combining proof-of-work with Byzantine consensus}
-Recent research has developed a class of hybrid systems which uses PoW for committee election,
+Recent research has developed a class of hybrid systems which uses proof-of-work (PoW) for committee election,
 and Byzantine consensus algorithms to agree on transactions.
-This design is primarily for permissionless system because the PoW leader election aspect prevents Sybils.
 % And then they use a Byzantine consensus algorithm to actually reach consensus on a set of transactions within the committee.
@@
 
-This technique overcomes the early blockchain scalability issue by delegating the transaction validation to the Byzantine consensus protocol (e.g., PBFT in ByzCoin~\cite{kogias2016enhancing}).
+This technique overcomes the early blockchain scalability issue by delegating the transaction validation to the Byzantine consensus protocol (e.g., PBFT in ByzCoin~\cite{kogias2016enhancing})
+while keeping the permissionless nature by using PoW.
 A major tradeoff of such systems is that they cannot guarantee correctness when there is a large number of malicious nodes (but less than a majority).
@@
 
-Sharding is a technique to achieve horizontal scalability by grouping nodes into multiple committees or shards.
+Sharding is a technique to achieve horizontal scalability by grouping nodes into multiple committees, also known as shards.
 Nodes within a single shard run a Byzantine consensus algorithm to agree on a set of transactions that belong to that specific shard.
@@
 
-The biggest limitation of sharding is that it is only optimal is transactions stay in the same shard.
+The biggest limitation of sharding is that it is only optimal if transactions stay in the same shard.
 In fact, Elastico cannot atomically process inter-shard transactions.
@@
 Tangle~\cite{tangle}, Corda~\cite{corda} and the original TrustChain~\cite{multichain} do not use global consensus at all.
-In this scheme, every node has their own chain.
+In these schemes, every node has their own chain.
 Transactions between nodes are recorded on their respective chains.
@@
 
-On the other hand, the application and security are limited.
+On the other hand, the application and security properties are limited.
 A malicious node can easily lie to other nodes regarding its own chain by simply presenting one version of reality to a set of nodes and another version to a different set.
@@
 \section{Research question}
+\label{sec:research-question}
 
 We saw in~\Cref{sec:taxonomy} that the majority of the systems cannot achieve horizontal scalability.
-To the best of our knowledge, only two systems---Elastico~\cite{luu2016elastico} and OmniLedger~\cite{kokoris2017omniledger}---has experimentally demonstrated horizontal scalability.
+To the best of our knowledge, only two systems---Elastico~\cite{luu2016elastico} and OmniLedger~\cite{kokoris2017omniledger}---have experimentally demonstrated horizontal scalability.
 However, these techniques rely on sharding, thus the performance highly depends on transaction characteristics and the shard size parameter because inter-shard transactions are slow.
@@
 We wish to design a blockchain consensus protocol that can be used in many applications.
-For example, the accounting of detailed internet traffic in Tribler~\cite{pouwelse2008tribler, multichain} and decentralised market\footnote{\url{https://github.com/Tribler/tribler/blob/devel/Tribler/community/tradechain/__init__.py}}.
+For example, the accounting of detailed internet traffic in Tribler~\cite{pouwelse2008tribler, multichain} and decentralised market in Tribler\footnote{\url{https://github.com/Tribler/tribler/blob/devel/Tribler/community/tradechain/__init__.py}}.
 Hence our research question is the following.
@@
 
-In order to make sense of our research question, we must first define \emph{blockchain consensus protocol}.
+In order to make sense of our research question, we must first explain \emph{blockchain consensus protocol}.
 Typical blockchain systems are application specific.
@@
 Underneath these applications lies the blockchain consensus protocol which has the goal of reaching consensus on transactions or state changes.
-In the case of Bitcoin, it is proof-of-work (PoW).
-It is not concerned with the structure of the transactions,
-it works by simply treating transactions as a blob of data and hashes them with the previous block.
-In this work, we focus on the blockchain consensus protocol rather than any specific application.
-Because it is the scalability bottleneck and is necessary for any blockchain based application.
+In the case of Bitcoin, it is PoW,
+because it is not concerned with the structure of the transactions.
+In this work, we focus on the blockchain consensus protocol rather than any specific application because it is the scalability bottleneck and is necessary for any blockchain based application.
 % Also, it should be possible to build any kind of application on top our system.
@@
 
-In the remainder of this section, we expand on our research question and visit each of our requirements in detail.
+In the remainder of this section, we expand on our research question and visit each aspect in detail.
 At the end of this section, we describe some issues that we do not consider as a part of our design,
@@
 The state can be a small, but verifiable representation of all the transactions.
-As an extreme example, a single digest of the complete history can represent the state.
-It is even verifiable as long as there are nodes that store the history.
+As an extreme example, a single digest computed from the complete history can represent the state.
+It is verifiable as long as there are nodes that store the history.
 
@@
 our system should be unaffected in the presence of powerful adversaries.
-In particular, adversaries are Byzantine meaning that they can have arbitrary behaviour.
+In particular, adversaries are Byzantine,
+meaning that they can have arbitrary behaviour.
 Thus anything is possible from simply omitting messages to colluding with each other in order to undermine the whole system.
@@
 % Since our system is application neutral, we do not provide an explicit Sybil defence mechanism.
-On the other hand, our system also has no restrictions on the Sybil defence mechanism
+On the other hand, our system has no restrictions on the Sybil defence mechanism
 and applications built on top of our system can use the best mechanism for their purpose.

### 2017-08-20 thesis/chapters/ch3-model.tex
Edit everything

@@
 It is based on our prior work on TrustChain~\cite{multichain},
-where every node independently interacts with each other via their own blockchain\footnote{It is originally called MultiChain but the name has changed to TrustChain}.
+where every node independently interacts with each other via their own hash chain\footnote{TrustChain is originally called MultiChain but the name has changed to TrustChain}.
 
@@
 We begin our discussion with an intuitive overview of the architecture in \Cref{sec:system-overview}.
-Next, we give the formal description, starting with the model and assumptions in \Cref{sec:model-assumptions}.
-Then, the three protocols which make up the complete system,
-namely consensus protocol (\Cref{sec:cons-protocol}), transaction protocol (\Cref{sec:tx-protocol}) and validation protocol (\Cref{sec:vd-protocol}).
-Finally, we discuss a few variations of our main design and their respective tradeoffs in \Cref{sec:tradeoffs}.
+Next, we give the formal description, starting with the model and assumptions in \Cref{sec:model-assumptions}, 
+and then the three protocols which make up the complete system.
+Finally, we discuss a few variations of our design and their respective tradeoffs in \Cref{sec:tradeoffs}.
 
@@
 Thus we first describe the standard TrustChain.
-Our description has minor differences compared to the description in~\cite{multichain},
-which is to help with the description of the Extended TrustChain.
-We remark the differences when they occur.
+Our description has a minor difference when compared to the description in~\cite{multichain},
+we remark the difference when it occurs.
 
 % \subsubsection*{Standard TrustChain}
-In TrustChain, every node has a ``personal'' chain. 
+In TrustChain, every node has a ``personal'' hash chain. 
 Initially, the chain only contains a genesis block generated by the nodes themselves.
-When a node $A$ wishes to add a new transaction (TX) with $B$,
+When node $A$ wishes to add a new transaction (TX) with $B$,
 a new TX block is generated for both $A$ and $B$ and is appended to their respective chains.
@@
     The arrows represent hash pointers and the dotted lines represent references.
-    The blocks at the ends of one dotted line are pairs of each other.
-    The red block after $t_{b, 5}$ indicates a fork.}
+    The blocks at the ends of the dotted lines are pairs of each other.
+    The red block behind $t_{b, 6}$ indicates a fork.}
     \label{fig:trustchain-bad}
@@
 % \subsubsection*{Extended TrustChain}
-The primary difference is the introduction of a new type of block---checkpoint (CP) block.
-The most important function of the CP block is to represent the state of a node, in the form of a hash pointer.
+Extended TrustChain introduces a new block type---checkpoint (CP) block.
+The primary function of the CP block is to represent the state of a node, in the form of a hash pointer.
 A collection of CP blocks from all nodes represents the state of the whole system.
@@
 
-The consensus protocol can be seen as a technique of running infinitely many times of some 
-Byzantine consensus algorithm\footnote{More accurately it is asynchronous common subset (ACS) which we describe in \Cref{sec:acs-background}.},
+The consensus protocol can be seen as a technique of running infinitely many times of a Byzantine consensus algorithm\footnote{More accurately it is an asynchronous common subset (ACS) algorithm, which we describe in \Cref{sec:acs-background}.},
 starting a new execution immediately after the previous one is completed.
-This approach is necessary because blockchain systems always need to reach consensus on new values proposed by nodes in the system,
+This approach is necessary because blockchain systems always need to reach consensus on new values proposed by the nodes in the system,
 or CP blocks in our case.
@@
 Thus, at the beginning of every Byzantine consensus algorithm execution,
-we randomly elect a set of nodes---called facilitators---to collect CP blocks from every other node and use them as the input to the Byzantine consensus algorithm, and then run it.
-After running the algorithm,
+we randomly elect a set of nodes---called facilitators---to collect CP blocks from every other node and use those blocks as the input to the Byzantine consensus algorithm, and then run it.
+After the algorithm completes,
 the facilitators output a set of CP blocks which we call the consensus result.
@@
 
-\subsection*{Transaction and validation}
+\subsection*{Transaction and validation protocols}
 The TX protocol is a simple request and response protocol.
@@
 the CP blocks must also be in consensus.
-Upon the counterparty's response, the node checks whether the CP blocks are in some consensus result,
-the hash pointers are valid and his TX is actually in the fragment.
+Upon the counterparty's response,
+the node checks whether the CP blocks are in some consensus result and some other conditions.
 The TX is valid if these conditions are satisfied.
@@
 
-\subsection*{Complete protocol}
-The complete protocol is essentially the concurrent composition of the three aforementioned protocols,
+\subsection*{The complete system}
+The complete system is essentially the concurrent composition of the three aforementioned protocols,
 all making use the Extended TrustChain data structure.
@@
 
-This section and the ones following it give a technical treatment of what the content in System overview.
+This section and the ones following it give a technical explanation of \textsc{Checo}.
 For notational clarity, we use the following convention (adapted from~\cite{miller2016honey}) for most of this work.
@@
 
-We assume there exists a Public Key Infrastructure (PKI), and nodes are identified by their unique and permanent public key.
+Further, we assume there exists a Public Key Infrastructure (PKI), and nodes are identified by their unique and permanent public key.
 Finally, we use the random oracle (RO) model, i.e. calls to the random oracle are denoted by $\textsf{H}: \{0, 1\}^* \rightarrow \{0, 1\}^\lambda$,
 where $\{0, 1\}^*$ denotes the space of finite binary strings and $\lambda$ is the security parameter.
-Under the RO model, the probability of successfully computing the inverse of the hash function is negligible with respect to $\lambda$~\cite{bellare1993random}.
+Under the RO model, the probability of successfully computing the inverse of $\textsf{H}(\cdot)$ is negligible with respect to $\lambda$~\cite{bellare1993random}.
 
@@
 In our model, we consider $N$ nodes, which is the population size.
-$n$ of them are facilitators, $t$ out of $n$ are malicious and the inequality
+$n$ of them are facilitators, $t$ nodes are malicious and the inequality
 $n \ge 3t + 1$ must hold.
 This inequality is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$~\cite{pease1980reaching}.
-Further, the inequality $N \ge n + t$ must also hold so that there are enough honest nodes in the network,
+Further, the inequality $N \ge n + t$ must also hold so that there are enough honest nodes in the network to run the consensus protocol,
 the exact reason becomes clear in~\Cref{sec:consensus-phase}.
@@
 and $h$ is the height of the block (i.e. $h = |B_u|$).
-We often use $b_{u, h}$ to denote the latest block.
+We use $b_{u, h}$ to denote the latest block.
 There are two types of blocks, TX blocks and CP blocks.
@@
 \item $\textsf{H}(b_{u, i - 1})$ is the hash pointer to the previous block.
-\item $seq_u$ is the sequence number which should equal $i$.
+\item $seq_u$ is the sequence number, which should equal $i$.
 \item $txid$ is the transaction identifier, it should be generated using a cryptographically secure pseudo-random number generator by the initiator of the transaction.
@@
 TX blocks come in pairs.
-In particular, for every block 
+In particular, for every TX block 
 $$t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$$
 there exist one and only one \emph{pair} 
-$$t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), seq_v, txid, pk_u, m, sig_v \rangle.$$
+$$t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), seq_v, txid, pk_u, m, sig_v \rangle,$$
+if they are created by honest nodes running the TX protocol.
 Note that the $txid$ and $m$ are the same, and the public keys refer to each other.
@@
 \begin{definition}
-\textbf{\emph{Checkpoint block}}
+\textbf{\emph{Checkpoint block and genesis block}}
 
@@
 $$c_{u, i} = \langle \textsf{H}(b_{u, i-1}), seq_u, \textsf{H}(\C_r), r, sig_u \rangle,$$
-where $\C_r$ is the consensus result (which we describe in \Cref{sec:consensus-result}) in round $r$,
+where $\C_r$ is the consensus result (which we describe next in \Cref{sec:consensus-result}) in round $r$,
 the other items are the same as the TX block definition.
@@
 Similarly, \emph{agreed fragment} has the same definition as fragment but using agreed enclosure.
-For convenience, the function $\textsf{agreed\_fragment}(\cdot)$ represents the agreed fragment of some TX block if it exists, otherwise $\bot$.
+For convenience, the function $\textsf{agreed\_fragment}(t_{u, i})$ outputs the agreed fragment of $t_{u, i}$ if it exists, otherwise $\bot$.
 \end{definition}
@@
 
-Our consensus protocol $\Pi_c$ runs on top of Extended TrustChain.
+Our consensus protocol $\Pi_c$ runs on top of the Extended TrustChain.
 It uses an unmodified asynchronous common subset (ACS) algorithm as the key building block.
@@
     and have an unbiased election of facilitators.
-Concretely, we define the following.
+We formally define the desired properties below.
 \begin{definition}
@@
 
-A TrustChain consensus protocol is correct if the following holds for every round $r$:
+A TrustChain consensus protocol is correct if the following holds for every round $r$.
 \begin{enumerate}
@@
             \begin{enumerate}
-                \item $|\C_r| \ge N - t$ and each CP in $\C_r$ belong to a different node\footnote{
-                $\C_r$ is a tuple but we abuse the notation here by writing $|\C_r|$ to mean the number of CP blocks in the second element of $\C_r$.}.
+                \item $|\C_r| \ge N - t$\footnote{
+                $\C_r$ is a tuple but we abuse the notation here by writing $|\C_r|$ to mean the number of CP blocks in the second element of $\C_r$.}, and
                 \item $|\F_r| = n$.
@@
 \end{definition}
-These properties are similar to Byzantine consensus properties (which we describe next in \Cref{sec:acs-background}), but there are subtle differences.
+These properties are similar to the typical Byzantine consensus properties, but there are subtle differences.
 Firstly, they are properties for every node in the network and not just the facilitators.
@@
 ACS has the nice property of censorship resilience when compared to other consensus algorithms.
-For instance, Hyperledger and Tendermint use a variant of Practical Byzantine Fault Tolerance (PBFT)~\cite{castro1999practical} as their consensus algorithm.
-In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectively filter transactions.
+In PBFT~\cite{castro1999practical}, a leader is elected, if the leader is malicious but follows the protocol, then it can selectively filter transactions.
 In contrast, every party in ACS are involved in the proposal phase,
@@
 
-The reader may notice from the system overview (\Cref{sec:system-overview}) that the facilitators are computed from the consensus result,
-but the consensus result is agreed by the facilitators.
-Thus we have a dependency cycle.
-The goal of the bootstrap phase is to give us a starting point in the cycle.
+% The reader may notice from the system overview (\Cref{sec:system-overview}) that the facilitators are computed from the consensus result,
+% but the consensus result is agreed by the facilitators.
+% Thus we have a dependency cycle.
+% The goal of the bootstrap phase is to give us a starting point in the cycle.
 
@@
 
-In practice, the bootstrap oracle is most likely the software developer and some of the desired properties cannot be achieved.
-In particular, it is not possible to have the fairness property because it is unlikely that the developer knows the identity of every node in advance.
+In practice, the bootstrap oracle is most likely the software developer that sets up the system and assigns the facilitators of round 1 in a fair way.
 
@@
 \label{sec:consensus-phase}
-The consensus phase begins when $\F_r$ is available to all the nodes.
+For any node $u$,
+the consensus phase begins when $\F_r$ is available and the latest block is $c_{u, h}$.
 Note that $\F_r$ indicates the facilitators that were elected using results of round $r$ and are responsible for driving the ACS algorithm in round $r + 1$.
@@
 There are two scenarios in the consensus phase.
-First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators,
-where $c_{u, h}$ is the most recent CP block.
-Second, if the node is a facilitator, it waits until $N - t$ messages of type \texttt{cp\_msg} are received.
+First, if $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators.
+Second, if $u$ is a facilitator, it waits until $N - t$ messages of type \texttt{cp\_msg} are received.
 % Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D \gg \Delta$.
@@
 Duplicates and blocks with invalid signatures are again removed from $\C'_{r+1}$ and we call the final result $\C_{r+1}$.
-We have to remove invalid blocks a second time (after ACS) because the adversary may send different CP blocks to different facilitators,
+We have to remove invalid blocks a second time because the adversary may send different CP blocks to different facilitators,
 which results in invalid blocks in the ACS output, but not in any of the inputs.
@@
 The core of the consensus phase is the ACS algorithm.
-While any ACS algorithm that satisfies the standard definition will work,
+While any algorithm that solves the ACS problem will work,
 we use a simplification of HoneyBadgerBFT~\cite{miller2016honey} as our ACS algorithm 
@@
 Second, HoneyBadgerBFT uses threshold encryption to hide the content of the transactions.
-But we do not reach consensus on transactions, only CP blocks, so hiding CP blocks is meaningless for us as it contains no transaction information.
+But we do not reach consensus on transactions, only CP blocks, so hiding CP blocks is meaningless for us.
 
@@
 The reason for sending \texttt{cons\_sig} is the following.
-Recall that channels are not authenticated, 
+The channels are not authenticated, 
 and there are no signatures in $\C_{r + 1}$.
 If a non-facilitator sees some $\C_{r + 1}$, it cannot immediately trust it because it may have been forged.
-Thus, To guarantee authenticity, every facilitator sends an additional message that is the signature of $\C_{r + 1}$.
+Thus, to guarantee authenticity, every facilitator sends an additional message that is the signature of $\C_{r + 1}$.
 
-Upon receiving $\C_{r + 1}$ and at least $n - f$ valid signatures by some node $u$, $u$ performs two tasks.
+Upon receiving $\C_{r + 1}$ and at least $n - t$ valid signatures, $u$ performs two tasks.
 First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1})$ (\Cref{alg:new-cp}).
@@
 \State $\langle r, C \rangle \gets \C_r$
-\State $\textsf{take} (n, \textsf{sort\_by} (\lambda x.\textsf{H}(\C_r || pk \text{ of } x), C))$
+\State \Return $\textsf{take} (n, \textsf{sort\_by} (\lambda x.\textsf{H}(\C_r || pk \text{ of } x), C))$
 \end{algorithmic}
@@
 $txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$.
-Then the initiator sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.
+Then the initiator then sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.
 
@@
     \caption{Function $\textsf{new\_tx}(pk_v, m, txid)$ generates a new TX block and appends it to the caller $u$'s chain.
-    It is executed in the private context of $u$, i.e. it has access to the $sk_u$ and $B_u$.
-    The necessary arguments are the public key of the counterparty $pk_v$, the transaction message $m$ and the transaction identifier $txid$.}
+    It is executed in the private context of $u$, i.e. it has access to the $sk_u$ and $B_u$.}
     \label{alg:new-tx}
@@
 At no time in \Cref{alg:new-tx} or \Cref{alg:tx-proto} do we need to hold the chain state and wait for some message to be delivered before committing a new block to the chain.
-This allows for a high level of concurrency where we can call $\textsf{new\_tx}(\cdot)$ multiple times without waiting for the corresponding \texttt{tx\_resp} messages.
+This allows for a high level of concurrency where we can call $\textsf{new\_tx}(\cdot)$ and send \texttt{tx\_req} messages multiple times without waiting for the corresponding \texttt{tx\_resp} messages.
 
@@
 
-Up to this point, we do not provide a mechanism to detect tampering.
+The consensus protocol and the validation protocol do not provide a mechanism to detect tampering.
 The validation protocol aims to solve this issue.
 The protocol is also a request-response protocol, just like the transaction protocol.
+It assumes that nodes store all the consensus results that are aware of.
 But before explaining the protocol itself, we first define what it means for a transaction to be valid.
@@
 \subsection{Validity definition}
-A transaction can be in one of three states in terms of validity---\emph{valid}, \emph{invalid} and \emph{unknown}.
+A transaction can be in one of three states---\emph{valid}, \emph{invalid} and \emph{unknown}.
 Given a fragment $F_{v, j}$, the validity of the TX block $t_{u, i}$ with its corresponding fragment $F_{u, i}$ is captured by the function $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ (\Cref{alg:get-validity}).
@@
 we essentially check whether the fragment is the one that the verifier needs.
-If it is not, then the verifier cannot make any decision and return \emph{unknown},
-which is likely to be the case for new transactions because the result of $\textsf{agreed\_fragment}(\cdot)$ would be $\bot$.
-The next two conditions check for tampering or missing blocks, if any of these misconducts are detected, then the TX is invalid.
+If it is not, then the verifier cannot make any decision and return \emph{unknown}.
+Such case is possible because the result of $\textsf{agreed\_fragment}(\cdot)$ would be $\bot$ for new transactions.
+The next three conditions check for tampering or missing blocks, if any of these misconducts are detected, then the TX is invalid.
 
@@
 But in reality, it was the malicious sender who did not follow the protocol.
-Thus, in such cases, the whole transaction identified by its $txid$ is marked as invalid.
+Thus, in such cases, the whole transaction is marked as invalid.
 
@@
         If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
-        then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
+        then all other correct nodes are able to reach the same conclusion or output \emph{unknown}.
     \item \emph{Validity}:
@@
 In this section, 
-we explore a few design variations which we can make, some of them require a relaxed version of our original model.
+we explore a few design variations, some of them require a relaxed version of our original model.
 They enable better performance, allow us to apply our design in the fully permissionless setting and improves privacy.
@@
 One of the final steps in our consensus protocol is to broadcast the consensus result and signatures to every other node (\Cref{sec:consensus-phase}).
-While this guarantees delivery (since we assumed reliable channel), it is wasteful.
+While this guarantees delivery, it is wasteful.
 For example, if every facilitator is honest, a node would receive $n$ consensus results which are identical when only one is necessary.
@@
 
-An improvement is use an epidemic protocol~\cite{eugster2004epidemic} (also known as gossiping) instead of our broadcast approach.
-Typical epidemic protocols works as follows.
+An improvement is use an epidemic protocol~\cite{eugster2004epidemic} instead, it is also known as gossip protocol.
+Typical epidemic protocols work as follows.
 Every node buffers every message it receives up to some buffer size $b$.
@@
 Every time the message is sent to $f$ random neighbours, $f$ is often called the fan-out.
-The upside of using epidemic protocol is that the communication cost is distributed more evenly between nodes,
+The upside of using an epidemic protocol is that the communication cost is distributed more evenly between nodes,
 which is especially true with a lazy push approach where the node who just received a message would push the message ID (e.g., digest) to its $f$ random neighbours,
@@
 A down side of an epidemic protocol is that it usually takes $O(\log N)$ time to infect the whole network,
-whereas broadcasting uses constant time.
+whereas broadcasting is constant time.
 Another downside of some epidemic protocols (e.g., eager push) is that it is difficult to guarantee delivery.
@@
 For example, TCP relies on timeout for its retransmission and the \texttt{nLockTime} property in Bitcoin transactions makes the transaction unspendable until some time in the future (either Unix time or block height)~\cite{bitcoindevguide}.
-One limitation of our system is that we use the parameter $N$ in our algorithms, which makes it unsuitable for the permissionless environment where users can join and leave at will.
+A limitation of our system is that we use the parameter $N$ in our algorithms, which makes it unsuitable for the permissionless environment where users can join and leave at will.
 In this section we show how making a timing assumption would allow us to operate in the permissionless setting.
 
-At the start of our consensus phase (\Cref{sec:consensus-phase}), facilitators must wait for $N-f$ \texttt{cp\_msg} messages,
+At the start of our consensus phase (\Cref{sec:consensus-phase}), facilitators must wait for $N-t$ \texttt{cp\_msg} messages,
 which is the only place where we used $N$ as a parameter.
-To introduce timing, instead of waiting for $N-f$ messages, we wait for some time $D$,
+To introduce timing, instead of waiting for $N-t$ messages, we wait for some time $D$,
 such that $D$ is sufficiently long for honest nodes to send their CP blocks to the facilitators.
@@
 To leave the network, nodes simply stop submitting CP blocks.
-There is a subtlety here which happens when the node is elected as a facilitator in the following round.
-In this case, the node must fulfil its obligation by completing the consensus protocol, but without proposing its own CP block, before leaving.
+There is a subtlety here which happens when the node is elected as a facilitator but still wish to leave.
+In this case, the node must fulfil its obligation by completing the consensus protocol,
+but without proposing its own CP block before leaving.
 Otherwise, the $n \ge 3t + 1$ condition may be violated.
@@
 \label{sec:compact}
-Our approach already has privacy preserving features in comparison to early blockchain systems.
+Our approach already has privacy preserving features in comparison to early block\-chain systems.
 That is, the transactions for each node are only revealed during the validation protocol.
@@
 their transactions are never revealed.
-
 We can take our privacy-preserving property one step further by introducing another level of hash pointer indirection.
 The result is shown in~\Cref{fig:compact}.
+
 \begin{figure}
@@
     \centering
-    \caption{The chain on the left represent direct chaining, where the digest in ``Prev'' is simply the digest of the previous block.
+    \caption{The chain on the left represent direct chaining of full blocks,
+    where the digest in ``Prev'' is simply the digest of the previous full block.
     The chain on the right uses compact blocks, represented by the smaller squares.
@@
 \begin{enumerate}
-\item Seq---the sequence number its corresponding block,
-\item Digest---the digest of its corresponding block, and
+\item Seq---the sequence number of the corresponding full block,
+\item Digest---the digest of the corresponding full block, and
 \item Prev---the digest of the previous compact block.
@@
 The relationship is uniquely identified with Seq or Digest.
-Recall that our original validation protocol requires the nodes to send the full agreed fragment.
+Recall that our original validation protocol requires the nodes to send the agreed fragment containing the full blocks.
 With compact blocks, it is only necessary to send the compact version of the agreed fragment.
@@
 rather than validating a single transaction,
-$A$ attempts to validate all transactions performed with $B$ in the unknown state, in that fragment.
+$A$ attempts to validate all transactions performed with $B$, which are in the unknown state but also in that fragment.
 
@@
 We offer two approaches to do total fork detection.
-First and the easiest solution is to simply ask for the complete chain of the counterparty.
-The verifier can be sure that there are no forks if the following holds.
+First and the easiest solution is to simply ask for the complete hash chain of the counterparty.
+The verifier can be sure that there are no forks if the following conditions hold.
 \begin{enumerate}

### 2017-08-20 thesis/chapters/ch4-analysis.tex
Edit everything

@@
 we evaluate our system analytically to check whether it correctly implements the TrustChain consensus protocol (\Cref{def:consensus})
-and the TrustChain validation protocol (\Cref{def:validation}) in~\Cref{sec:correctness}.
-Further, we analyse the performance and especially the throughput to understand the scalability characteristics in~\Cref{sec:performance-analysis}.
-Finally, we consider the case where the $n \ge 3t + 1$ assumption is violated and show the effect of it in our system in~\Cref{sec:highly-adversarial}.
+and the TrustChain validation protocol (\Cref{def:validation}).
+Further, we analyse the performance and especially the throughput to understand the scalability characteristics.
+Finally, we consider the case where the $n \ge 3t + 1$ assumption is violated and show the effect of it in our system.
 
-\section{Correctness in the presence of faults}
+\section{Correctness analysis}
 \label{sec:correctness}
@@
 $\F_r$ is known by all correct nodes and any correct node outputs $\F_{r+1}$,
-then (a) $|\C_{r+1}| \ge N - t$ must hold for the $\C_{r+1}$ which was used to create $\F_{r+1}$ and (b) $|\F_{r+1}| = n$.
+then (a) $|\C_{r+1}| \ge N - t$ and (b) $|\F_{r+1}| = n$.
 \end{lemma}
@@
 But $n -t$ facilitators must have received $N - t$ valid CP blocks, so $|\C_{r+1}| \ge N-t$ and this proves (a).
-Since $N-t \ge (n+t) -t = n$ and $\textsf{get\_facilitators}(\cdot)$ outputs $n$ items, so $|\F_{r+1}| = n$ and this proves (b).
+Since $N-t \ge (n+t) -t = n$ and $\textsf{get\_facilitators}(\C_{r+1})$ outputs $n$ items, so $|\F_{r+1}| = n$ and this proves (b).
 \end{proof}
@@
 Then the proof directly follows from the random oracle model.
-Recall that the luck value is computed using $\textsf{H}(\C_{r+1}, || pk_u)$.
+Recall that the luck value is computed using $\textsf{H}(\C_{r+1} || pk_u)$.
 Since $pk_u$ is unique for every node that has a CP block in $\C_{r+1}$,
 the output of $\textsf{H}(\cdot)$ is uniformly random.
-This effectively generates a random permutation
+This effectively generates a random permutation,
 so every node has the same probability of being in the top $n$ for the ordered sequence,
-namely the output of $\textsf{get\_facilitators}(\cdot)$.
+namely the output of $\textsf{get\_facilitators}(\C_{r+1})$.
 \end{proof}
@@
 \label{theorem:consensus}
-For all rounds, $\Pi_c$ implements a TrustChain consensus protocol (\Cref{def:consensus}).
+$\Pi_c$ implements a TrustChain consensus protocol (\Cref{def:consensus}).
 \end{theorem}
@@
 In the base case ($r = 1$), agreement, validity fairness and termination follows directly from the bootstrap protocol.
-Note that the result is $\F_1$, which indicates the facilitators that are agreed in round 1, who are responsible for driving the ACS protocol in round 2.
+Note that the result is $\F_1$,
+which represents the set of facilitators agreed in round 1,
+are responsible for driving the ACS protocol in round 2.
 
@@
 which makes it the building block for reaching consensus on transactions.
-In this section, we build on top of \Cref{theorem:consensus} to show that our validation protocol described in \Cref{sec:vd-protocol} has the agreement and validity properties of a TrustChain validation protocol defined in \Cref{def:validation}.
+In this section, we build on top of \Cref{theorem:consensus} to show that our validation protocol $\Pi_v$ described in \Cref{sec:vd-protocol} has the agreement and validity properties of a TrustChain validation protocol defined in \Cref{def:validation}.
 
@@
 We proof the agreement property by contradiction.
-From~\Cref{alg:get-validity},
-we use the fact that $F_{u, i}$ and $t_{u, i}$ are valid.
+Recall that $F_{u, i}$ and $t_{u, i}$ used in \Cref{alg:get-validity} are valid.
 Without loss of generality, suppose for some transaction with TX block $t$,
@@
 
-Our protocol uses \Cref{alg:get-validity} which is also the validity definition,
+$\Pi_v$ directly uses \Cref{alg:get-validity} which is also the validity definition,
 so the validity property is also satisfied.
@@
 This enables horizontal scalability where adding new nodes would lead to higher global transaction rate.
-In addition, a convenient consequence \Cref{theorem:validation-agreement} is unforgeability.
-That is, no polynomial time adversary is able to create two chains, $F = \{ \dots, t, c\}$ and $F' = \{ \dots, t', c\}$, with correct hash pointers and the same end of chain $c$.
+In addition, a convenient consequence of \Cref{theorem:validation-agreement} is unforgeability.
+That is, no adversary is able to create two chains,
+$F = \{ \dots, t, c\}$ and $F' = \{ \dots, t', c\}$,
+with correct hash pointers and the same end of chain $c$ with non-negligible probability.
 % Another way to think about it is if a transaction can be forged, then it cannot be agreed.
@@
 it can make sure all \texttt{tx\_req} messages are delivered in a round later than the round which the message is sent.
-Effectively, the transaction pair would always be in different rounds and the validation protocol would not output \emph{valid}.
+Effectively, the pair would always be in different rounds and the validation protocol would not output \emph{valid}.
 We believe in a relaxed model, i.e. a weakly synchronous model, a stronger validity definition is possible.
@@
 \begin{lemma}
-There exists a valid transaction that cannot be validated eventually.
+There exists a valid transaction that cannot be eventually validated.
 \end{lemma}
@@
 Then when $u$ wants to validate it, it does so by sending \texttt{vd\_req} message to $v$.
-$v$ can act maliciously and ignore all \texttt{vd\_req} message, then the transaction can never be validated.
+$v$ can act maliciously and ignore all \texttt{vd\_req} message from $u$, then the transaction can never be validated.
 \end{proof}
@@
 \label{sec:cons-duration}
-In order to make arguments on bandwidth or throughput in our purely asynchronous model, which are concepts that depend on time,
-we must make some assumptions regarding our model to make arguments on the duration of the consensus protocol.
-Note that it is not the same as the time complexity typically used in distributed systems.
+In order to make arguments on the duration of $\Pi_c$, the bandwidth or throughput in our purely asynchronous model,
+which are concepts that depend on time,
+we must make additional assumptions.
+Note that the duration we are interested in is not the same as the time complexity typically used in distributed systems.
 In the analysis of distributed algorithms, time complexity is often in terms of the number of rounds.
@@
 For every unit of communication, we assume they take some non-negligible but constant time to perform.
-Hence, from~\Cref{sec:cons-complexity} and the fact that the consensus protocol uses a constant number of rounds,
-it follows that the consensus protocol has a duration of $O(N)$.
+Hence, from~\Cref{sec:cons-complexity} and the fact that $\Pi_c$ runs in a constant number of rounds,
+it follows that $\Pi_c$ has a duration of $O(N)$.
 
@@
 Geared with the results above,
-we are ready to analyse the amount of data required to be transmitted over a link per transaction.
-Which we call the communication cost per transaction.
+we are ready to analyse the amount of data required to be transmitted over a link per transaction,
+which we call the communication cost per transaction.
 To create and then validate a transaction,
@@
 $$l = r_{\text{tx}} D_{\text{c}},$$
-where $D_{\text{c}}$ is the duration a round of the consensus protocol.
+where $D_{\text{c}}$ is the duration of a round of the consensus protocol.
 But from \Cref{sec:cons-duration},
 we know that $D_{\text{c}}$ is of $O(N)$, thus the communication cost per transaction is $O(N)$.
-This is intuitive because round duration would be longer if there are more CP blocks (more $N$), which means that the agreed fragments are longer (assuming nodes transact at a constant rate).
-The behaviour is verified experimentally in \Cref{ch:implementation}.
+This result is intuitive because round duration would be longer if there are more CP blocks (more $N$), which means that the agreed fragments are longer (assuming nodes transact at a constant rate).
+This behaviour is verified experimentally in \Cref{ch:implementation}.
 
@@
 First, we clarify the bandwidth definition, which is ``the data rate at which a network link or a network path can transfer''~\cite{prasad2003bandwidth}.
-Now, suppose every node has some fixed bandwidth $C$ per link, $N$ links and are making transactions at $r_{\text{tx}}$ per second.
+Now, suppose every node has $N$ links which connect to every other node,
+every link has a fix bandwidth $C$ and every node makes $r_{\text{tx}}$ transactions per second.
 Then we have the inequality
@@
 However, if that is the case, we can intelligently cache the \texttt{vd\_resp} messages as described in~\Cref{sec:caching}.
-For this work, we analyse the worst case where every node transacts with a random node from the population,
+Hence, we analyse the worst case where every node transacts with a random node from the population,
 and a new \texttt{vd\_resp} must be sent for every \texttt{vd\_req} message.
@@
 Recall that $l$ is of $O(N)$, that means LHS and RHS both grow linearly with respect to $N$.
-Hence, there exists some constant $r_\text{tx}$ that makes use of all the available bandwidth regardless of $N$.
-Finally, if every node in the network is transacting at a constant rate, then the global throughput (in terms of transactions per second) is linear w.r.t. the population size $N$.
+Hence, there exists some $r_\text{tx}$ that makes use of all the available bandwidth regardless of $N$.
+Finally, if every node in the network is transacting at $r_{\text{tx}}$ per second,
+then the global throughput (in terms of transactions per second) is linear w.r.t. the population size $N$.
 If the system is not making use of all the bandwidth.
-We also maintain a constant transaction rate by the same argument.
-Thus a global throughput is also linear w.r.t. $N$.
+We also maintain a constant transaction rate by the same argument,
+so the global throughput is also linear w.r.t. $N$.
 We verify both of these claims experimentally in~\Cref{ch:implementation}.
@@
 The bound is not tight, but it is useful for picking parameters for a desired level of fault tolerance.
-For some fixed $n$, since $0 \le \alpha \le \lfloor \frac{n-1}{3} \rfloor/n < (\lfloor  \frac{n - 1}{3} \rfloor + 1) / n$,
+If $n$ is fixed, since $0 \le \alpha \le \lfloor \frac{n-1}{3} \rfloor/n < (\lfloor  \frac{n - 1}{3} \rfloor + 1) / n$,
 the probability is maximum when the squared term is minimum at $\alpha = \lfloor \frac{n-1}{3} \rfloor/n$.
@@
 To put this into perspective,
-suppose $n = 1000$ and $\alpha = 1/5$, i.e. a fifth of the nodes are malicious.
-Then the probability to draw more black balls than the threshold is only $2.6 \times 10^{-16}$.
+suppose $n = 100$ and $\alpha = 1/10$, i.e. 10\% of the nodes are malicious,
+then the probability to draw more black balls than the threshold is $9.9 \times 10^{-6}$.

### 2017-08-20 thesis/chapters/ch5-implementation.tex
Edit everything

@@
 
-We evaluate our system experimentally to ultimately understand its scalability characteristics as well as limitations.
+We evaluate our system experimentally to ultimately understand its scalability characteristics as well as its limitations.
 Concretely,
@@
 Finally, \Cref{sec:comms-cost-experiment} and onwards follows the same structure as \Cref{sec:performance-analysis},
-where we analyse the performance experimentally rather than analytically and compare the results.
+where we analyse the performance experimentally.
 
@@
 
-The goal of the experiment is to run the three protocols---consensus protocol,
-transaction protocol and validation protocol---simultaneously and analyse the communication cost, the consensus duration and the global throughput.
+The goal of the experiment is to run \textsc{Checo} and analyse the communication cost, the consensus duration and the global throughput.
 We investigate these properties under the following parameters.
@@
         The maximum $n$ is 32 because the limitation in liberasurecode mentioned in \Cref{sec:implementation},
-        but our results give a good indication of how our system may function for larger numbers of $n$.
+        but our results in~\Cref{sec:varying-facilitator-experiment} give a good indication of how our system may function for a larger $n$.
   \item The population size $N \in \{200, 300, \dots, 1200\}$.
@@
   \item The two modes of transaction.
-        The first mode is the \emph{fixed-neighbour} mode where nodes only transact with their immediate neighbour which is predetermined. 
+        The first mode is the \emph{fixed-neighbour} mode where nodes only transact with one predetermined node. 
         This minimises the data volume per validated transaction because agreed fragment can be cached.
@@
   }
-  \caption{Communication cost for the consensus protocol per round increases linearly with population size.
+  \caption{Experimental result of the communication cost per round for the consensus protocol.
   The error bars are larger for higher population size or higher number of facilitators is because rounds take longer thus they are repeated fewer times.}
@@
   }
-  \caption{Round duration does not increase linearly with the population size,
-  which is likely due to the additional hashing operation required for larger consensus result.}
+  \caption{Experimental result of the round duration per round for the consensus protocol.}
   \label{fig:round-duration}
@@
   }
-  \caption{Communication cost per verified transaction has similar (near linear) trend as~\Cref{fig:round-duration}.
+  \caption{Experimental result of the communication cost per verified transaction.
+  It has a similar, near linear, trend as~\Cref{fig:round-duration}.
   Fluctuation for the fixed-neighbour mode exists because the cache mechanism is unpredictable.}
@@
 More interestingly, there is a large difference in communication cost between the two modes of transaction.
-When transacting with only one neighbour, the communication cost is low because only one agreed fragment needs to be communicated for every round in order to validate all transactions of that round,
-which is because agreed fragments are cached.
+When transacting with only one neighbour, the communication cost is low because only one agreed fragment needs to be communicated for every round in order to validate all transactions of that round.
 On the other hand,
@@
 The result in~\Cref{fig:tx-comms-random} reinforces our argument.
-It is a lot more stable because for every transaction it is almost always the case that an agreed fragment is needed to validate it.
+It is a lot more stable because for every transaction it is almost always the case that a new agreed fragment is needed to validate it.
 
 \section{Linear global throughput}
-\begin{figure}[h]
+\begin{figure}[tb]
   \centering
@@
 
-Note that the throughput decreases slightly as the number of facilitators increases,
-which is due to the additional communication cost for running ACS with a high number of facilitators.
+Note that the throughput decreases slightly as the number of facilitators increases.
+This effect is due to the additional communication cost for running ACS with a high number of facilitators.
 That is, if the network is congested then the nodes may not have enough bandwidth to send validation responses timely.
@@
   }
-  \caption{A limitation of our system is when nodes are transacting with many random nodes,
-  then the network cannot keep up with the large number of agreed fragments that need to be communicated between nodes.
-  Hence the slope of validated transactions grows at a slower rate in~\Cref{fig:backlog-random}.}
+  \caption{Demonstration of the network bottleneck.
+  Note that the slope of validated transactions grows at a slower rate in~\Cref{fig:backlog-random} than in~\Cref{fig:backlog-fixed}.}
   \label{fig:backlog}
@@
 \section{Communication cost with varying number of facilitators}
+\label{sec:varying-facilitator-experiment}
 Up to this point we focused on the effect of communication cost, throughput and so on with respect to the population size.
@@
 There is strong evidence that the communication cost grows polynomially.
-We expect this because of there are polynomial terms in the ACS communication complexity---$O(n^2|v| + \lambda n^3 \log n)$.
+We expect this behaviour because of there are polynomial terms in the ACS communication complexity---$O(n^2|v| + \lambda n^3 \log n)$.
 This also means that the rounds would take longer to complete and transactions would take longer to verify.

### 2017-08-20 thesis/chapters/ch6-conclusion.tex
Edit everything

@@
 In this work we introduced an application neutral blockchain system which we call \textsc{Checo}.
-We add checkpoint block to the existing TrustChain data structure for use in our consensus protocol.
+We add checkpoint block to the existing TrustChain data structure to capture the system state.
 The round based consensus protocol uses ACS as a building block to reach consensus on checkpoint blocks.
@@
 
-This work answers the research question stated in~\Cref{ch:problem}.
+This work answers the research question stated in~\Cref{sec:research-question}.
 \begin{displayquote}
-\emph{Is it possible to design a blockchain consensus protocol that is fault tolerant, scalable and can reach global consensus?}
+\emph{How can we design a blockchain consensus protocol that is fault-tolerant,
+horizontally scalable,
+and able to reach global consensus?}
 \end{displayquote}
@@
 we show that the probability for the system to fail is low even when $n \ge 3t + 1$ does not hold as long as the proportion of malicious nodes is not close to a third of $N$.
-For example, if there the population size is 1000 and 20\% of the nodes are malicious,
-the probability for a round to potentially to fail is bounded below $2.6 \times 10^{-16}$.
-This probability bound would decrease as the population size increases.
-Horizontal scalability property is demonstrated both analytically and experimentally.
+The horizontal scalability property is demonstrated both analytically and experimentally.
 Unlike sharding protocols, the property holds regardless of transaction characteristics and needs no parameter selection.
@@
 To understand our system further, we hope to experiment with a real platform with real users such as Tribler~\cite{pouwelse2008tribler}.
-For instance, the TrustChain in Tribler stores the number of bytes uploaded and downloaded between users, but it has no global consensus.
+For instance, the standard TrustChain in Tribler is used for accounting internet traffic between users,
+but it has no global consensus.
 It would be useful to integrate our consensus protocol into Tribler to prevent misreporting attacks.
@@
 for P2P file sharing systems the reputation score may be the number of bytes contributed to the network.
-Second, instead of using ACS we use a proof-of-stake (PoS) scheme which means that nodes with more stake in the system get more vote.
+Second, instead of using ACS we use a proof-of-stake (PoS) scheme (e.g., Ouroboros~\cite{kiayias2017ouroboros}) which means that nodes with more stake in the system get more vote.
 In an ideal PoS scheme, the system can tolerate any fault if the majority of the stake holders are honest.

### 2017-08-20 thesis/formalities/abstract.tex
Edit everything

@@
 We design a novel blockchain system called \textsc{Checo}.
-Each node in our system maintains individual hash chains,
+Each node in our system maintains a personal hash chain,
 which only stores transactions that the node is involved in.
 A consensus is reached on special blocks called checkpoint blocks rather than on all the transactions.
-Checkpoint blocks are effectively a hash pointer to the individual hash chains,
-thus a single checkpoint block may represent a large set of transaction blocks.
+Checkpoint blocks are effectively a hash pointer to the personal hash chains,
+thus a single checkpoint block may represent a large set of transactions.
 The consensus protocol does not imply transaction validity.
-Hence we include a validation protocol which allows nodes to verify that transactions are correctly recorded,
-which indirectly achieves consensus on transactions.
+Hence we introduce a validation protocol which allows nodes to verify that the transactions are correctly recorded,
+meaning that consensus on transactions is achieved indirectly via checkpoint blocks.
 % how good is the solution?
-We implement a prototype and evaluate it experimentally.
+We evaluate \textsc{Checo} analytically and experimentally.
 Our results show a strong indication of horizontal scalability even in the worst case,

### 2017-08-20 thesis/formalities/preface.tex
Edit everything

--- 2017-08-07 thesis/formalities/preface.tex
@@
 \addcontentsline{toc}{chapter}{Preface}
-Since early 2013, when I mined my first Bitcoin, 
-I have been interested in decentralising the traditionally centralised systems.
+Since early 2013, the time when I mined my first Bitcoin, 
+I became cautious of the security and privacy issues surrounding centralised systems,
+and became interested in building a decentralised society.
 Following the Snowden revelations, the interest became a desire that has stayed with me ever since.
@@
 Dr. Zhijie Ren, this work would not have been possible without your invaluable input and the numerous discussions which we have had.
+Prof. Dick Epema, you taught me a lot on report writing and the use (and misuse) of mathematical formalism.
 Members of the Tribler team---Martijn de Vos, Quinten Stokkink and Chengxin Ma,
-you have my gratitude for helping me with the implementation and proofreading my work.
+you have my gratitude for helping me with the implementation and proofreading my work,
+and of course the free coffee.
 Members of the Cyber Security group---Dr. Zekeriya Erkin and O\u{g}uzhan Ersoy,

### 2017-08-21 thesis/chapters/ch1-intro.tex
Thanks Johan

@@
 
-In Bitcoin's case, the consensus algorithm is called \emph{proof-of-work}.
+In Bitcoin's case, the consensus algorithm is called \emph{proof-of-work} (PoW).
 Where miners (parties that run the Bitcoin network) collect transactions and compete in solving a puzzle.
@@
 Bitcoin has had its ups and downs, but overall it has grown into an enormous system.
-Its power consumption is as high as Republic of Ireland~\cite{o2014bitcoin}.
+Its power consumption is as high as the Republic of Ireland~\cite{o2014bitcoin}.
 Its market capitalisation, at the time of writing, is over 60 billion USD and growing (see \Cref{fig:market-cap}).
@@
 This issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some call a civil war~\cite{bitcoincivilwar}.
-Parties that are for the increase in block size argue that a larger block would improve the transaction rate.
+Parties which are for an increase in block size argue that a larger block would improve the transaction rate.
 Parties against it argue that it would make mining more centralised because blocks take longer to propagate through the network.
@@
 The effect is that the first 1 MB of the block is still backwards compatible, thus there is no need for a hard fork.
-It also solves the transaction malleability problem~\cite{bitcoinmalleability} which paves the way for off-chain transactions via Lightning Network~\cite{lightningnetwork}.
-However, some fear SegWit is not enough and wish to see a hard fork that uses ``Adjustable Blocksize Cap'' that increase the maximum block size even further to 8 MB~\cite{bitcoincash}.
+It also solves the transaction malleability problem~\cite{bitcoinmalleability} which paves the way for off-chain transactions via the Lightning Network~\cite{lightningnetwork}.
+However, some fear SegWit is not sufficient and wish to see a hard fork that uses an ``Adjustable Blocksize Cap'' that increases the maximum block size even further to 8 MB~\cite{bitcoincash}.
 The ongoing civil war resulted in a Bitcoin fork, called Bitcoin Cash~\cite{bitcoincasharticle}.
@@
 and it is highly non-trivial to ``patch'' the problem.
+
+Moreover, Bitcoin is far from distributed.
+It is more accurate to say Bitcoin is made in China~\cite{madeinchina}.
+The combination of cheap electronics and electricity has made China one of the most profitable place to mine Bitcoins.
+At the time of writing, China holds about 80\% of the hashing power~\cite{bitcoinpool}.
+Effectively putting China in control of the 60 billion-dollar ecosystem.
+If the government decides to flex its totalitarian muscles,
+then it would spell the end of the ecosystem.
 
@@
 but nevertheless verifiable at a later stage by any node in the network.
-We call our system \textsc{Checo}, derived from \emph{che}ckpoint \emph{co}nsensus.
-From a high level, our system works as follows.
-Every node stores its own hash chain and every block contains only one transaction.
-We randomly select nodes in every round, the selected ones are called facilitators.
-Facilitators reach consensus not on the individual transactions,
-but on the state of every chain represented by a single digest, called a checkpoint block.
-If a checkpoint block of some node is in consensus, 
-then that node can prove to any other node that it holds a set of transactions that compute (form a chain) to the checkpoint block.
-As a result, those transactions are tamper-proof.
-The key difference from the early blockchain systems is the independence of the transaction validation phase.
-Prior systems such a Bitcoin and Ethereum entangled the consensus algorithm with the transaction validation, restricting performance.
-By splitting this functionality we achieved horizontal scalability in the general case.
-That is, by adding new nodes into the network, the global transaction rate should increase proportionally,
-regardless of the choice of system parameters.
-
+A verified transaction is guaranteed to be tamper-proof and irrefutable.
+% We call our system \textsc{Checo}, derived from \emph{che}ckpoint \emph{co}nsensus.
+% From a high level, our system works as follows.
+% Every node stores its own hash chain and every block contains only one transaction.
+% We randomly select nodes in every round, the selected ones are called facilitators.
+% Facilitators reach consensus not on the individual transactions,
+% but on the state of every chain represented by a single digest, called a checkpoint block.
+% If a checkpoint block of some node is in consensus, 
+% then that node can prove to any other node that it holds a set of transactions that compute (form a chain) to the checkpoint block.
+% As a result, those transactions are tamper-proof.
+% The key difference from the early blockchain systems is the independence of the transaction validation phase.
+% Prior systems such a Bitcoin and Ethereum entangled the consensus algorithm with the transaction validation, restricting performance.
+% By splitting this functionality we achieved horizontal scalability in the general case.
+% That is, by adding new nodes into the network, the global transaction rate should increase proportionally,
+% regardless of the choice of system parameters.
 Our main contributions are the following.
 \begin{itemize}
-    \item We formally introduce a blockchain system---\textsc{Checo}.
+    \item We formally introduce a blockchain system---\textsc{Checo}\footnote{Derived from \emph{che}ckpoint \emph{co}nsensus.}.
         It uses individual hash chains and checkpoint blocks on every node to achieve

### 2017-08-21 thesis/chapters/ch2-problem.tex
Thanks Johan

@@
 We saw that early blockchain systems face a scalability problem.
-In particular, the use proof-of-work is unlikely to reach a transaction rate suitable for ubiquitous use.
+In particular, the use PoW is unlikely to reach a transaction rate suitable for ubiquitous use.
 Many attempts to fix the scalability problem exist in academia as well as the industry.
-In this chapter, we describe the related work in the form of a mini taxonomy and identify areas for improvement and limitations.
-Then, we state our research question which aims to overcome some of the limitations.
+In this chapter, we describe the related work and identify their limitations.
+Then, we state our research question which aims to overcome some of these.
 
-\section{Related work and mini taxonomy}
+\section{Related work}
 \label{sec:taxonomy}
@@
 To understand the current state-of-the-art,
-we classify various block\-chain systems by their scalability approach,
-beginning with the baseline.
+we classify various block\-chain systems by their scalability approach.
 
@@
 Given sufficient computational power, adversaries are able to undo any transaction.
-The typical examples are proof-of-work based systems such as Bitcoin, Ethereum and other Altcoins.
+The typical examples are PoW based systems such as Bitcoin, Ethereum and other Altcoins.
 In Bitcoin, nodes compete in solving puzzles.
@@
 The advantage of this approach is that it can be used in a large network.
-In fact, prior to Bitcoin, there was no internet-scale consensus protocol.
-Bitcoin is also fault-tolerant because attackers can not out pace honest users in finding new blocks unless they have a majority of the computational power,
+In fact, prior to Bitcoin, there was no Internet-scale consensus protocol.
+Bitcoin is also fault-tolerant because attackers can not outpace honest users in finding new blocks unless they have a majority of the computational power,
 as long as the majority is honest and reasonable measures are taken to prevent the selfish mining attack~\cite{eyal2014majority}.
@@
 Hence double spending is possible for transactions that do not have a large number of confirmations.
-Furthermore, as we mentioned in the~\Cref{ch:intro},
+Furthermore, as we mentioned in~\Cref{ch:intro},
 the performance is limited due to the fact that blocks are of fixed size and are generated at fixed intervals.
@@
 even if the two parties made thousands or millions of transactions in between.
+
 The two-party scheme can be extended to a network of channels,
 which allows two parties to make off-chain transactions without an open channel as long as they are connected by nodes that do.
+For example, if a channel exists between node $A$ and $B$, and a channel exists between node $B$ and $C$,
+then $A$ can pay $C$ via the intermediary $B$.
+The result is that there is no need to create a payment channel with every node in the network.
+If some key nodes, acting as payment hubs, are well-connected,
+then nodes can make off-chain transactions to a large number of other nodes by opening only one channel with a payment hub.
 
@@
 In essence, they contain a fixed set of nodes (sometimes called validating peer) that run a Byzantine consensus algorithm to decide on new blocks.
-This technique is often used in permissioned system where the validators must be predetermined,
+This technique is often used in permissioned systems where the validators must be predetermined,
 for example, Hyperledger Fabric~\cite{cachin2016architecture}.
@@
 PBFT can, for example, achieve 10,000 TX/s if the number of validating peer is under 16~\cite[Section 5.2]{miller2016honey}.
-Further, in contrast to proof-of-work, PBFT consensus is final.
+Further, in contrast to PoW, PBFT consensus is final.
 That is, transaction history cannot be re-written if it is already in the blockchain.
@@
 
-\subsection{Combining proof-of-work with Byzantine consensus}
-Recent research has developed a class of hybrid systems which uses proof-of-work (PoW) for committee election,
+\subsection{Combining PoW with Byzantine consensus}
+Recent research has developed a class of hybrid systems which uses PoW for committee election,
 and Byzantine consensus algorithms to agree on transactions.
@@
 \subsection{Sharding}
+\label{sec:sharding}
 
@@
 then adding more shard would naturally result in a linear increase of global throughput.
-Examples of blockchain systems that use sharding are Elastico~\cite{luu2016elastico} and OmniLedger~\cite{kokoris2017omniledger}.
+Examples of blockchain systems that use sharding are Elastico~\cite{luu2016elastico} and Omni\-Ledger~\cite{kokoris2017omniledger}.
 
@@
 The authors of OmniLedger noted that inter-shard transactions have significantly higher latency compared to the consensus protocol~\cite{kokoris2017omniledger}.
-Furthermore, since no shards can be compromised for the system to function correctly,
-the adversaries have more opportunities to compromise the system.
+Furthermore, using smaller shards is a security risk.
+Every shard is fault-tolerant up to at most a third of the shard size.
+If the total number of malicious nodes remain fixed, then as the shard size decreases,
+it is more probably for a shard to fail.
 
@@
 
-Tangle~\cite{tangle}, Corda~\cite{corda} and the original TrustChain~\cite{multichain} do not use global consensus at all.
+Tangle~\cite{tangle}, Corda~\cite{corda} and the original TrustChain~\cite{multichain} avoid the need of global consensus.
 In these schemes, every node has their own chain.
@@
 We saw in~\Cref{sec:taxonomy} that the majority of the systems cannot achieve horizontal scalability.
-To the best of our knowledge, only two systems---Elastico~\cite{luu2016elastico} and OmniLedger~\cite{kokoris2017omniledger}---have experimentally demonstrated horizontal scalability.
-However, these techniques rely on sharding, thus the performance highly depends on transaction characteristics and the shard size parameter because inter-shard transactions are slow.
+To the best of our knowledge, only the two systems described in~\Cref{sec:sharding} have experimentally demonstrated horizontal scalability.
+However, these techniques rely on sharding, thus the performance highly depends on the transaction characteristics and the shard size parameter.
 Furthermore, all the systems described in~\Cref{sec:taxonomy} are complete blockchain systems designed for a concrete application.
@@
 A good defence mechanism requires application specific information.
-For example, a social network based Sybil defence mechanisms use graph structure of real-world relationships~\cite{yu2006sybilguard}.
+For example, a social network based Sybil defence mechanisms use random walks in real-world social graphs~\cite{yu2006sybilguard}.
 Online marketplaces such as Amazon use the rating of buyer and sellers.
+Cryptocurrencies use PoW to vote on the latest block.
 % Since our system is application neutral, we do not provide an explicit Sybil defence mechanism.
-On the other hand, our system has no restrictions on the Sybil defence mechanism
-and applications built on top of our system can use the best mechanism for their purpose.
-
-For the same reason, we do not explicitly consider spam or denial of service attacks.
-Without a concrete application, it is difficult to distinguish a super active user from an attacker.
+For the same reason, we do not explicitly consider spam or denial of service attacks 
+because without a concrete application,
+it is difficult to distinguish a super active user from an attacker.
+For this work,
+we assume that it is the role of application developer to find a reputation mechanism or a trust model suitable for defending against such attacks.
 

### 2017-08-21 thesis/chapters/ch3-model.tex
Include genesis block in figures
Thanks Johan

@@
 
-Our system consists of three protocols---the consensus protocol, the transaction protocol and the validation protocol.
-It is based on our prior work on TrustChain~\cite{multichain},
+In this chapter, we break away from the early blockchain paradigm and introduce a novel system that is application neutral,
+horizontally scalable and will never have a block size debate.
+In particular, our system consists of three protocols---the consensus protocol, the transaction protocol and the validation protocol.
+It is based on our prior research on TrustChain~\cite{multichain} at Delft,
 where every node independently interacts with each other via their own hash chain\footnote{TrustChain is originally called MultiChain but the name has changed to TrustChain}.
 
-The goal of this chapter is to detail our system specification.
 We begin our discussion with an intuitive overview of the architecture in \Cref{sec:system-overview}.
@@
 \label{sec:system-overview}
-The architecture is visualised in~\Cref{fig:architecture}.
+
 \begin{figure}[h]
@@
 \includegraphics[width=0.7\textwidth]{architecture}
-\caption{Overview of the system architecture}
+\caption{Overview of \textsc{Checo} architecture.}
 \label{fig:architecture}
 \end{figure}
+
+The architecture of \textsc{Checo} is visualised in~\Cref{fig:architecture}.
 It consists of one data structure---Extended TrustChain, and three protocols.
@@
 \subsection*{TrustChain and Extended TrustChain}
-Extended TrustChain naturally builds on top of the standard TrustChain. 
-Thus we first describe the standard TrustChain.
-Our description has a minor difference when compared to the description in~\cite{multichain},
+We first describe the standard TrustChain and then the Extended TrustChain.
+Our description of the standard TrustChain has a minor difference when compared to the description in~\cite{multichain},
 we remark the difference when it occurs.
@@
     The blocks at the ends of the dotted lines are pairs of each other.
-    The red block behind $t_{b, 6}$ indicates a fork.}
+    The red block behind $t_{b, 6}$ indicates a fork.
+    Blocks of index 0 (e.g. $t_{b, 0}$) are genesis blocks.}
     \label{fig:trustchain-bad}
@@
     \centering
-    \caption{
-    An example of Extended TrustChain.
+    \caption{An example of Extended TrustChain.
     The circles represent CP blocks,
     they also have hash pointers (arrow) but do not have references (dotted line).
-    Note that the sequence number counter do not change, it is shared with TX blocks.}
+    Note that the sequence number counter do not change, it is shared with TX blocks.
+    CP blocks of index 0 (e.g. $c_{b, 0}$) are genesis blocks.}
     \label{fig:trustchain-bad-cp}
@@
 
-The high communication complexity of Byzantine consensus algorithms prohibits us from running it on a large network.
+The communication complexity of Byzantine consensus algorithms typically grows polynomially w.r.t the number of nodes,
+which prohibits us from running it on a large network.
+For example, one of the most recent work tested with up to 104 nodes and the algorithm took almost 6 minutes to complete~\cite{miller2016honey}.
 Thus, at the beginning of every Byzantine consensus algorithm execution,

### 2017-08-21 thesis/formalities/abstract.tex
Thanks Johan

@@
 However, scalability remains a key challenge.
-Without a horizontally scalable solution, blockchain systems remain unsuitable for ubiquitous use.
+Without a horizontally scalable solution,
+where performance increases by adding more nodes to the system,
+blockchain systems remain unsuitable for ubiquitous use.
 % what's the solution?
@@
 Checkpoint blocks are effectively a hash pointer to the personal hash chains,
-thus a single checkpoint block may represent a large set of transactions.
-The consensus protocol does not imply transaction validity.
-Hence we introduce a validation protocol which allows nodes to verify that the transactions are correctly recorded,
-meaning that consensus on transactions is achieved indirectly via checkpoint blocks.
+thus a single checkpoint block may represent an arbitrarily large set of transactions.
+We introduce a validation protocol so that any node can check the validity of any transaction.
+Since it is a point-to-point protocol, we achieve horizontal scalability.
 % how good is the solution?
-We evaluate \textsc{Checo} analytically and experimentally.
-Our results show a strong indication of horizontal scalability even in the worst case,
-where every transaction is made with a randomly selected node.
+We analytically evaluate our system and proof a number of highly desirable correctness properties.
+Further, we give a free and open-source implementation of \textsc{Checo} and evaluate it experimentally.
+Our results show a strong indication of horizontal scalability.
 \end{abstract}

### 2017-08-21 thesis/formalities/preface.tex
Thanks Johan

@@
 Since early 2013, the time when I mined my first Bitcoin, 
-I became cautious of the security and privacy issues surrounding centralised systems,
+I became cautious of the security and privacy issues surrounding centralised systems
 and became interested in building a decentralised society.

### 2017-08-22 thesis/chapters/ch1-intro.tex
Edit grammar etc.

@@
 Due to this limitation, it is not uncommon to see a long backlog of about 20,000 transactions\footnote{\url{https://blockchain.info/unconfirmed-transactions/}}.
-A few months ago the backlog even reached 100,000, which meant new transactions would take at least 11 hours to be written to the blockchain~\cite{bitcoinbacklog}.
+The backlog even reached 100,000 in May 2017, which meant new transactions would take at least 11 hours to be written to the blockchain~\cite{bitcoinbacklog}.
 This issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some call a civil war~\cite{bitcoincivilwar}.

### 2017-08-22 thesis/chapters/ch2-problem.tex
Edit grammar etc.

@@
 we classify various block\-chain systems by their scalability approach.
+A summary is given in~\Cref{tab:scalability}.
+
+\begin{table}[htb]
+    \centering
+    \caption{Summary of the scalability properties of many blockchain systems.
+    Scalability gets better from left to right.}
+    \label{tab:scalability}
+    \begin{tabular}{|l|l|l|l|}
+    \hline
+    \textbf{\thead{Not\\scalable}} & \textbf{\thead{Somewhat\\scalable}} & \textbf{\thead{Limited horizontal\\scalability}} & \textbf{\thead{Full horizontal\\scalability}} \\ \hline
+    Bitcoin~\cite{bitcoin}           & Hyperledger~\cite{cachin2016architecture} & Elastico~\cite{luu2016elastico}           & \textsc{Checo} (this work)        \\
+    Ethereum~\cite{wood2014ethereum} & ByzCoin~\cite{kogias2016enhancing}       & OmniLedger~\cite{kokoris2017omniledger}   &                                   \\
+    etc.                             & Solidius~\cite{abraham2016solidus}       &                                           &                                   \\
+    \hline
+    \end{tabular}
+\end{table}
 
@@
 Each of the party must also deposit some Bitcoins in the multi-signature address,
-this is called the opening transaction.
+which is called the opening transaction.
 Both parties keep the channel state which is not broadcasted to the network.
@@
 To close the channel, the parties simply broadcast the latest net settlement to the Bitcoin network,
-this is called the closing transaction.
+which is called the closing transaction.
 Effectively, only two transactions need to be broadcasted to the Bitcoin network---the opening and the closing transaction,
@@
 % And then they use a Byzantine consensus algorithm to actually reach consensus on a set of transactions within the committee.
-Some examples are SCP~\cite{luu2015scp}, ByzCoin~\cite{kogias2016enhancing} and Solidus~\cite{abraham2016solidus}.
+Some examples are ByzCoin~\cite{kogias2016enhancing} and Solidus~\cite{abraham2016solidus}.
 
@@
 A major tradeoff of such systems is that they cannot guarantee correctness when there is a large number of malicious nodes (but less than a majority).
-For SCP, ByzCoin and Solidus, they all have some probability to elect more than $t$ Byzantine nodes into the committee,
+For ByzCoin and Solidus, they all have some probability to elect more than $t$ Byzantine nodes into the committee,
 where $t$ is typically just under a third of the committee size (a lower bound of Byzantine consensus~\cite{pease1980reaching}).
@@
 Nodes within a single shard run a Byzantine consensus algorithm to agree on a set of transactions that belong to that specific shard.
-An intra-shard protocol is needed for transactions that involve nodes from more than one shard.
+An inter-shard protocol is needed for transactions that involve nodes from more than one shard.
 The number of shards grows linearly with respect to the total computational power in the network.
@@
 To the best of our knowledge, only the two systems described in~\Cref{sec:sharding} have experimentally demonstrated horizontal scalability.
-However, these techniques rely on sharding, thus the performance highly depends on the transaction characteristics and the shard size parameter.
+However, these techniques rely on sharding; thus the performance highly depends on the transaction characteristics and the shard size parameter.
 Furthermore, all the systems described in~\Cref{sec:taxonomy} are complete blockchain systems designed for a concrete application.
@@
 
-In order to make sense of our research question, we must first explain \emph{blockchain consensus protocol}.
+To make sense of our research question, we must first explain \emph{blockchain consensus protocol}.
 Typical blockchain systems are application specific.
-For example, the application in Bitcoin (any many of its derivatives) is digital cash, the application in Namecoin is domain name registration~\cite{namecoin} and the application in Siacoin is cloud storage~\cite{siacoin}.
+For example, the application in Bitcoin (any many of its derivatives) is digital cash,
+the application in Namecoin is domain name registration~\cite{namecoin},
+and the application in Siacoin is cloud storage~\cite{siacoin}.
 Underneath these applications lies the blockchain consensus protocol which has the goal of reaching consensus on transactions or state changes.
 In the case of Bitcoin, it is PoW,
-because it is not concerned with the structure of the transactions.
+because it is not concerned with the structure or the semantics of the transactions.
 In this work, we focus on the blockchain consensus protocol rather than any specific application because it is the scalability bottleneck and is necessary for any blockchain based application.
@@
 At the end of this section, we describe some issues that we do not consider as a part of our design,
-which is the result of not focusing on any specific application.
+which is the result of not focusing on any particular application.
 
@@
 meaning that they can have arbitrary behaviour.
-Thus anything is possible from simply omitting messages to colluding with each other in order to undermine the whole system.
+Thus anything is possible from simply omitting messages to colluding with each other to undermine the whole system.
 
@@
 For example, if two different honest nodes attempt to verify the same transaction,
-it cannot be the case that one node thinks that the transaction is valid but the other thinks it is invalid.
+it cannot be the case that one node thinks that the transaction is valid, but the other thinks it is invalid.
+Furthermore, just like transactions in cryptocurrencies, they should be tamper-proof and irrefutable.
 
@@
 Online marketplaces such as Amazon use the rating of buyer and sellers.
-Cryptocurrencies use PoW to vote on the latest block.
+Cryptocurrencies use PoW to decide on the latest block.
 % Since our system is application neutral, we do not provide an explicit Sybil defence mechanism.
-For the same reason, we do not explicitly consider spam or denial of service attacks 
-because without a concrete application,
-it is difficult to distinguish a super active user from an attacker.
+For the same reason, we do not explicitly consider spam or denial of service attacks because,
+without a concrete application,
+it is hard to distinguish a super active user from an attacker.
 For this work,
-we assume that it is the role of application developer to find a reputation mechanism or a trust model suitable for defending against such attacks.
+we assume that it is the role of the application developer to find a reputation mechanism or a trust model suitable for defending against such attacks.
 

### 2017-08-22 thesis/chapters/ch3-model.tex
Edit grammar etc.

@@
 \includegraphics[width=0.7\textwidth]{architecture}
-\caption{Overview of \textsc{Checo} architecture.}
+\caption{The four components of \textsc{Checo}.}
 \label{fig:architecture}
@@
     they also have hash pointers (arrow) but do not have references (dotted line).
-    Note that the sequence number counter do not change, it is shared with TX blocks.
+    Note that the sequence number counter does not change, it is shared with TX blocks.
     CP blocks of index 0 (e.g. $c_{b, 0}$) are genesis blocks.}
@@
 which prohibits us from running it on a large network.
-For example, one of the most recent work tested with up to 104 nodes and the algorithm took almost 6 minutes to complete~\cite{miller2016honey}.
+For example, one of the most recent works tested with up to 104 nodes and the algorithm took almost 6 minutes to complete~\cite{miller2016honey}.
 Thus, at the beginning of every Byzantine consensus algorithm execution,
@@
 the facilitators output a set of CP blocks which we call the consensus result.
-The result is then propagated to the network and the next algorithm execution begins.
+The result is then propagated to the network,
+and the next algorithm execution begins.
 
@@
 \subsection*{The complete system}
-The complete system is essentially the concurrent composition of the three aforementioned protocols,
+The complete system is essentially the concurrent composition of the three protocols above;
 all making use the Extended TrustChain data structure.
@@
 This inequality is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$~\cite{pease1980reaching}.
-Further, the inequality $N \ge n + t$ must also hold so that there are enough honest nodes in the network to run the consensus protocol,
-the exact reason becomes clear in~\Cref{sec:consensus-phase}.
+Further, the inequality $N \ge n + t$ must also hold so that there are enough honest nodes in the network to run the consensus protocol;
+the exact reason becomes apparent in~\Cref{sec:consensus-phase}.
 
@@
 
-An improvement is use an epidemic protocol~\cite{eugster2004epidemic} instead, it is also known as gossip protocol.
+An improvement is to use an epidemic protocol~\cite{eugster2004epidemic} instead, also known as gossip protocol.
 Typical epidemic protocols work as follows.
@@
 Every time the message is sent to $f$ random neighbours, $f$ is often called the fan-out.
-The upside of using an epidemic protocol is that the communication cost is distributed more evenly between nodes,
-which is especially true with a lazy push approach where the node who just received a message would push the message ID (e.g., digest) to its $f$ random neighbours,
+The upside of using an epidemic protocol is that the communication cost is distributed more evenly between nodes.
+It is especially true with a lazy push approach where the node who just received a message would push the message ID (e.g., digest) to its $f$ random neighbours,
 and only push the full message if the neighbour explicitly requests the message~\cite{leitao2007epidemic}.
@@
 Another downside of some epidemic protocols (e.g., eager push) is that it is difficult to guarantee delivery.
-It is especially true when the parameters are not chosen correctly in a network that is only partially connected (but every node is nevertheless reachable).
+It is particularly the case when the parameters are not chosen correctly in a network that is only partially connected (but every node is nevertheless reachable).
 If the delivery cannot be guaranteed, 
@@
 Our model is purely asynchronous, where we make no timing assumptions anywhere in the protocol.
-However, in many applications it is often fine to make timing assumptions.
+However, in many applications, it is often fine to make timing assumptions.
 For example, TCP relies on timeout for its retransmission and the \texttt{nLockTime} property in Bitcoin transactions makes the transaction unspendable until some time in the future (either Unix time or block height)~\cite{bitcoindevguide}.
@@
 Consequently, this removes the need for a PKI because the collected CP blocks may be from nodes that nobody has seen in the past.
-However, choosing the parameter $D$ is difficult and depends on a number of factors such as the network condition, message size, and so on.
+However, choosing the parameter $D$ is difficult and depends on many factors such as the network condition, message size, and so on.
 Overestimating it would make agreed fragments much longer than usual, which increases communication costs for validation.
 Underestimating it would lead to unfairness where users with a poor internet connection will have a lower chance to be selected as a facilitator in the next round.
-Nevertheless, there is a significant gain for making the timing assumption and that is the ability to operate in the permissionless setting which we explain next.
+Nevertheless, there is a significant gain for making the timing assumption,
+and that is the ability to operate in the permissionless setting which we explain next.
 
@@
 \begin{figure}
-    \includegraphics[width=0.7\textwidth]{compact}
+    \includegraphics[width=0.6\textwidth]{compact}
     \centering
-    \caption{The chain on the left represent direct chaining of full blocks,
+    \caption{The chain on the left represents direct chaining of full blocks,
     where the digest in ``Prev'' is simply the digest of the previous full block.
@@
 With compact blocks, it is only necessary to send the compact version of the agreed fragment.
-The validation then proceeds in a similar fashion,
+The validation then proceeds similarly,
 provided that the pair of the to-be-validated TX block is known.
@@
 We offer two approaches to do total fork detection.
-First and the easiest solution is to simply ask for the complete hash chain of the counterparty.
+First and the easiest solution is to ask for the complete hash chain of the counterparty.
 The verifier can be sure that there are no forks if the following conditions hold.
@@
 We use this approach in our prior work on Implicit Consensus~\cite{implicitconsensus}.
-Nodes employ caching to minimise communication costs, we call this effect spontaneous sharding.
+Nodes employ caching to minimise communication costs, and we call this effect spontaneous sharding.
 
@@
 then the complete chain is effectively validated in a distributed manner.
-The only way for an attacker to make a fork is to make sure that the agreed fragment containing the fork has no transactions with honest nodes.
+The only way for an attacker to make a fork is to ensure that the agreed fragment containing the fork has no transactions with honest nodes.
 Such malicious behaviour is prevented probabilistically using a challenge-response protocol as follows.

### 2017-08-22 thesis/chapters/ch4-analysis.tex
Edit grammar etc.

@@
 and the TrustChain validation protocol (\Cref{def:validation}).
-Further, we analyse the performance and especially the throughput to understand the scalability characteristics.
+Further, we analyse the performance and in particular the throughput to understand the scalability characteristics.
 Finally, we consider the case where the $n \ge 3t + 1$ assumption is violated and show the effect of it in our system.
@@
 \label{sec:correctness}
-Our first objective is to show that our consensus protocol $\Pi_c$ described in~\Cref{sec:cons-protocol} is in fact a TrustChain consensus protocol defined in~\Cref{def:consensus}.
+Our first objective is to show that our consensus protocol $\Pi_c$ described in~\Cref{sec:cons-protocol} is,
+in fact, a TrustChain consensus protocol defined in~\Cref{def:consensus}.
 Then, building on top of it, we show that our validation protocol $\Pi_v$ described in~\Cref{sec:vd-protocol} conforms to the TrustChain validation protocol defined in~\Cref{def:validation}.
@@
 Given that $\F_r$ is known,
-correct nodes will send CP blocks to all members in $\F_r$.
-The ACS algorithm starts independently whenever the facilitator has $N - t$ valid CP blocks
+correct nodes will send CP blocks to all members of $\F_r$.
+For each facilitator, the ACS algorithm starts independently whenever the facilitator has $N - t$ valid CP blocks
 (recall from \Cref{sec:consensus-phase} that invalid blocks are ones with an invalid signature or has a duplicate signature).
-It cannot make progress until $n-t$ honest facilitators start algorithm,
-but this eventually happens because there are $N - t$ correct nodes and all correct facilitator eventually receives $N -t$ valid CP blocks.
+No honest facilitator can make progress until $n-t$ honest facilitators start algorithm,
+but this eventually happens because there are $N - t$ correct nodes and all correct facilitator eventually receive $N -t$ valid CP blocks.
 At the end of ACS, $\C_{r+1}$ is created and then broadcasted along with the signature of the facilitators.
@@
 the output of $\textsf{H}(\cdot)$ is uniformly random.
-This effectively generates a random permutation,
-so every node has the same probability of being in the top $n$ for the ordered sequence,
-namely the output of $\textsf{get\_facilitators}(\C_{r+1})$.
+As a result, the output of $\textsf{get\_facilitators}(\C_{r+1})$ is a random permutation of all the nodes in $\C_{r+1}$ truncated to $n$.
+Thus, every node has the same probability of becoming a facilitator.
 \end{proof}
@@
 
-$\Pi_v$ directly uses \Cref{alg:get-validity} which is also the validity definition,
+$\Pi_v$ directly uses \Cref{alg:get-validity}, which is also the validity definition,
 so the validity property is also satisfied.
@@
 In addition, a convenient consequence of \Cref{theorem:validation-agreement} is unforgeability.
-That is, no adversary is able to create two chains,
+That is, no adversary can create two chains,
 $F = \{ \dots, t, c\}$ and $F' = \{ \dots, t', c\}$,
@@
 Then when $u$ wants to validate it, it does so by sending \texttt{vd\_req} message to $v$.
-$v$ can act maliciously and ignore all \texttt{vd\_req} message from $u$, then the transaction can never be validated.
+$v$ can act maliciously and ignore all \texttt{vd\_req} message from $u$, and then the transaction can never be validated.
 \end{proof}
@@
 \label{sec:performance-analysis}
-This section aims to analytically answer whether our system is horizontally scalable.
+This section aims to answer whether our system is horizontally scalable analytically.
 We begin by looking at the communication and time complexity of the consensus protocol $\Pi_c$,
@@
 \label{sec:cons-duration}
-In order to make arguments on the duration of $\Pi_c$, the bandwidth or throughput in our purely asynchronous model,
+To make arguments on the duration of $\Pi_c$, the bandwidth or throughput in our purely asynchronous model,
 which are concepts that depend on time,
@@
 For every unit of communication, we assume they take some non-negligible but constant time to perform.
-Hence, from~\Cref{sec:cons-complexity} and the fact that $\Pi_c$ runs in a constant number of rounds,
+Hence, from our results in~\Cref{sec:cons-complexity} and the fact that $\Pi_c$ runs in a constant number of rounds,
 it follows that $\Pi_c$ has a duration of $O(N)$.
@@
 where $l$ is the length of the agreed fragment.
-This can be seen from the fact that the largest message by far is the \texttt{vd\_resp} message,
+This result can be seen from the fact that the largest message by far is the \texttt{vd\_resp} message,
 which contains the agreed fragment.
 The other messages (\texttt{tx\_req}, \texttt{tx\_resp} and \texttt{vd\_req}) are constant factors.
-If we assume that every node performs transactions at a constant rate of $r_{\text{tx}}$ per second.
-Then
+If we assume that every node performs transactions at a constant rate of $r_{\text{tx}}$ per second, then
 $$l = r_{\text{tx}} D_{\text{c}},$$
@@
 But from \Cref{sec:cons-duration},
-we know that $D_{\text{c}}$ is of $O(N)$, thus the communication cost per transaction is $O(N)$.
+we know that $D_{\text{c}}$ is of $O(N)$; thus the communication cost per transaction is $O(N)$.
 This result is intuitive because round duration would be longer if there are more CP blocks (more $N$), which means that the agreed fragments are longer (assuming nodes transact at a constant rate).
@@
 \label{sec:global-throughput}
-Using our results so far, we are able to analyse the global throughput.
+Using our results so far, we can analyse the global throughput.
 First, we clarify the bandwidth definition, which is ``the data rate at which a network link or a network path can transfer''~\cite{prasad2003bandwidth}.
@@
 
-We note that the inequality does not hold if the node is only transacting with a subset of the population.
-This is because it cannot use all the bandwidth available in all the links.
+We note that the inequality does not hold if the node is only transacting with a subset of the population,
+which is because it cannot use all the bandwidth available in all the links.
 In the extreme case, if the node is only transacting with one other node, then it can only use the bandwidth of one link which is only $C$.
@@
 Consider the case where the system is making use of all the bandwidth, i.e. $NC = r_{\text{tx}} l$.
-Recall that $l$ is of $O(N)$, that means LHS and RHS both grow linearly with respect to $N$.
+Recall that $l$ is of $O(N)$, which means LHS and RHS both grow linearly with respect to $N$.
 Hence, there exists some $r_\text{tx}$ that makes use of all the available bandwidth regardless of $N$.
@@
 then the global throughput (in terms of transactions per second) is linear w.r.t. the population size $N$.
-If the system is not making use of all the bandwidth.
-We also maintain a constant transaction rate by the same argument,
-so the global throughput is also linear w.r.t. $N$.
-We verify both of these claims experimentally in~\Cref{ch:implementation}.
+We also maintain a constant transaction rate by the same argument if the system is not making use of all the bandwidth.
+Both claims are verified experimentally in~\Cref{ch:implementation}.
 

### 2017-08-22 thesis/chapters/ch5-implementation.tex
Thanks Johan
Edit grammar etc.

@@
 
-We evaluate our system experimentally to ultimately understand its scalability characteristics as well as its limitations.
+We evaluate our system experimentally to understand its scalability characteristics as well as its limitations.
 Concretely,
-we give an open source prototype implementation that runs on a network with over 1000 node.
-Nodes in our implementation make transactions autonomously with each other and then we collect the results for analysis.
+we give a free and open-source prototype implementation that runs on a network with over 1000 node.
+Nodes in our implementation make transactions autonomously with each other, and then we collect the results for analysis.
 
@@
 Thus we use the liberasurecode\footnote{\url{https://github.com/openstack/liberasurecode}} library for its Reed-Solomon error correcting code functionality.
-An implementation detail is that liberasurecode cannot create more than 32 code blocks, which is hardcoded in the source file\footnote{\url{https://github.com/openstack/liberasurecode/blob/0794b31c623e4cede76d66be730719d24debcca9/include/erasurecode/erasurecode.h\#L35}
+An implementation detail is that liberasurecode cannot create more than 32 code blocks, which is hard-coded in the source file\footnote{\url{https://github.com/openstack/liberasurecode/blob/0794b31c623e4cede76d66be730719d24debcca9/include/erasurecode/erasurecode.h\#L35}
 }, we discuss the effect of this in~\Cref{sec:experimental-setup}.
@@
 
-Every node keeps a persistent TCP connection with every other node.
-This creates a fully connected network for our experiment.
+Every node keeps a persistent TCP connection with every other node,
+which creates a fully connected network for our experiment.
 It is certainly not ideal in real world scenarios where nodes may have limited resources (e.g. sockets).
@@
   \item The population size $N \in \{200, 300, \dots, 1200\}$.
-        $N$ stops at 1200 is due to our physical setup, which we describe next.
+        The reason why $N$ only goes up to 1200 is due to our physical setup, which we describe below.
   \item The two modes of transaction.
         The first mode is the \emph{fixed-neighbour} mode where nodes only transact with one predetermined node. 
-        This minimises the data volume per validated transaction because agreed fragment can be cached.
+        It minimises the data volume per validated transaction because agreed fragment can be cached.
         The second mode is in the other extreme which we call the \emph{random-neighbour} mode,
-        where every transaction is with a random node out of the $N$ nodes in the system.
+        where every transaction is performed with a random node out of the $N$ nodes in the system.
         It is unlikely that the agreed fragments can be reused.
@@
 From now on, we use ``machines'' to refer to DAS-5 nodes and ``nodes'' to refer to a running instance in our system.
-On DAS-5 we use up to 30 machines, for each machine we use up to 40 nodes.
-This gives us the aforementioned 1200 number.
+On DAS-5 we use up to 30 machines, for each machine we use up to 40 nodes,
+which gives us the aforementioned 1200 number.
 With this setup, we cannot run more nodes because the every machine only has 65535 ports available (minus the reserved ones).
@@
 \Cref{fig:round-comms} shows the relationship between the communication cost of the consensus protocol per round and the population size.
-The most important aspect is that these results show a linear increase.
-This reinforces our analytical result in~\Cref{sec:cons-complexity}.
+The most important aspect is that these results show a linear increase,
+which reinforces our analytical result in~\Cref{sec:cons-complexity}.
 Note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
@@
 Finally, the rate for which the communication cost increases is higher when the number of facilitators is higher,
-which is also expected because the ACS algorithm has a polynomial term with respect to $n$ in its communication complexity.
+which is also expected because the ACS algorithm has a polynomial term w.r.t. $n$ in its communication complexity.
 
@@
 These results do not conform to analytical result in~\Cref{sec:cons-duration}.
-Nevertheless, the difference is minor and there are ways to optimise the luck value computation.
+Nevertheless, the difference is minor, and there are ways to optimise the luck value computation.
 For example, the luck value can be computed by the facilitators and are sent with the consensus result.
@@
   \caption{Experimental result of the communication cost per verified transaction.
-  It has a similar, near linear, trend as~\Cref{fig:round-duration}.
+  It has a similar, nearly linear, trend as~\Cref{fig:round-duration}.
   Fluctuation for the fixed-neighbour mode exists because the cache mechanism is unpredictable.}
@@
 To verify the argument, we plot the relationship between the communication cost of for every validated transaction and population size in~\Cref{fig:tx-comms}.
-We observe a near linear relationship, which is due to the near linear relationship of the communication duration mentioned before.
-Again, we believe the difference is minor and it is possible to remove the extra overhead.
+We observe a nearly linear relationship, which is due to the nearly linear relationship of the communication duration mentioned before.
+Again, we believe the difference is minor, and it is possible to remove the extra overhead.
 
 More interestingly, there is a large difference in communication cost between the two modes of transaction.
-When transacting with only one neighbour, the communication cost is low because only one agreed fragment needs to be communicated for every round in order to validate all transactions of that round.
+When transacting with only one neighbour, the communication cost is low because only one agreed fragment needs to be communicated for every round to validate all transactions of that round.
 On the other hand,
@@
 
-Some fluctuations exist in~\Cref{fig:tx-comms-fixed}, this is due to our caching mechanism.
+Some fluctuations exist in~\Cref{fig:tx-comms-fixed}, which is due to our caching mechanism.
 We send validation requests at the same rate as transactions.
@@
 The difference in magnitude between \Cref{fig:global-throughput-fixed} and \Cref{fig:global-throughput-random} is caused by the caching mechanism mentioned earlier.
-If a new agreed fragment needs to be transmitted to validate every transaction then it puts a toll on the network infrastructure.
+If a new agreed fragment needs to be transmitted to validate every transaction,
+then it puts a toll on the network infrastructure.
 The low transaction rate in~\Cref{fig:global-throughput-random} is caused by the fact that the network infrastructure cannot keep up with our demand.
@@
 \label{sec:varying-facilitator-experiment}
-Up to this point we focused on the effect of communication cost, throughput and so on with respect to the population size.
+Up to this point, we focused on the effect of communication cost, throughput and so on with respect to the population size.
 In this section, we consider a varying number of facilitators,
@@
 \Cref{fig:facilitators-round-comms} shows the communication cost of the consensus protocol for varying numbers of facilitators.
-There is strong evidence that the communication cost grows polynomially.
-We expect this behaviour because of there are polynomial terms in the ACS communication complexity---$O(n^2|v| + \lambda n^3 \log n)$.
-This also means that the rounds would take longer to complete and transactions would take longer to verify.
+There is substantial evidence that the communication cost grows polynomially.
+We expect this behaviour because of there are polynomial terms in the ACS communication complexity---$O(n^2|v| + \lambda n^3 \log n)$,
+which means that the rounds would take longer to complete and transactions would take longer to verify.
 On the other hand, we get better fault tolerance.

### 2017-08-22 thesis/chapters/ch6-conclusion.tex
Edit grammar etc.

@@
 
-In this work we introduced an application neutral blockchain system which we call \textsc{Checo}.
+In this work, we introduced an application neutral blockchain system which we call \textsc{Checo}.
 We add checkpoint block to the existing TrustChain data structure to capture the system state.
@@
 The consensus result, which is a set of checkpoint blocks, lets nodes elect new facilitators and create new checkpoint blocks.
-To make transactions, nodes use the transaction protocol,
-which is adapted from an existing work called True Halves.
+The existing work on True Halves is formalised to act as our transaction protocol.
 Finally, we introduce a validation protocol which ensures that if agreed fragments for some transaction exists,
@@
 
-Our experiment is carried out in a somewhat idealised world (DAS-5).
-We hope to experiment our prototype in a more realistic environment.
+Our experiment is conducted in a somewhat idealised world (DAS-5).
+We hope to study our prototype in a more realistic environment.
 For example, DeterLab~\cite{mirkovic2012teaching} lets users simulate network delay,
-which is important for discovering issue that may only occur in hostile environments.
+which is essential for discovering issues that may only occur in hostile environments.
 To understand our system further, we hope to experiment with a real platform with real users such as Tribler~\cite{pouwelse2008tribler}.
@@
 However, it is not adequate for permissionless systems,
-which usually aim to tolerate fault that involve a minority of the network,
+which usually aim to tolerate fault that involves a minority of the network,
 which can be just shy of 50\%.
@@
 Second, instead of using ACS we use a proof-of-stake (PoS) scheme (e.g., Ouroboros~\cite{kiayias2017ouroboros}) which means that nodes with more stake in the system get more vote.
-In an ideal PoS scheme, the system can tolerate any fault if the majority of the stake holders are honest.
+In an ideal PoS scheme, the system can tolerate any fault if the majority of the stakeholders are honest.
 
 Finally, as mentioned in~\Cref{sec:permissionless}, we wish to evaluate our system (analytically and experimentally) in the permissionless settings.
-This would enable our protocol to be applied to a much wider range of applications.
+It would enable our protocol to be applied to a much wider range of applications.

### 2017-08-22 thesis/formalities/abstract.tex
Edit grammar etc.

@@
 A consensus is reached on special blocks called checkpoint blocks rather than on all the transactions.
-Checkpoint blocks are effectively a hash pointer to the personal hash chains,
+Checkpoint blocks are effectively a hash pointer to the personal hash chains;
 thus a single checkpoint block may represent an arbitrarily large set of transactions.

### 2017-08-22 thesis/formalities/preface.tex
Edit grammar etc.

@@
 and became interested in building a decentralised society.
-Following the Snowden revelations, the interest became a desire that has stayed with me ever since.
+Following the Snowden revelations, the interest grew into a desire that has stayed with me ever since.
 Today, many blockchain systems are mature, but still not yet ready for ubiquitous use due to the scalability limitation.
 Many of us still do not use Bitcoin to purchase everyday goods, online or offline.
-This limitation, as well as my personal ambitions, led me to this thesis,
+This limitation, as well as my ambitions, led me to this thesis,
 with the goal of creating a truly horizontally scalable solution.
@@
 \noindent
-There are a number of people who have supported me in this endeavour who I would like to thank.
-My supervisor Dr. Johan Pouwelse, your unrivalled enthusiasm and guidance have kept me motivated and productive.
-Dr. Zhijie Ren, this work would not have been possible without your invaluable input and the numerous discussions which we have had.
-Prof. Dick Epema, you taught me a lot on report writing and the use (and misuse) of mathematical formalism.
+Many people have supported me in this endeavour which I would like to thank.
+My supervisor Dr Johan Pouwelse, your unrivalled enthusiasm and guidance have kept me motivated and productive.
+Dr Zhijie Ren, this work would not have been possible without your invaluable input and the numerous discussions which we have had.
+Prof Dick Epema, you taught me a lot about report writing and the use (and misuse) of mathematical formalism.
 Members of the Tribler team---Martijn de Vos, Quinten Stokkink and Chengxin Ma,
@@
 and of course the free coffee.
-Members of the Cyber Security group---Dr. Zekeriya Erkin and O\u{g}uzhan Ersoy,
-I thank both of you for your feedback on my presentations.
+Members of the Cyber Security group---Dr Zekeriya Erkin and O\u{g}uzhan Ersoy,
+I thank you both for your feedback on my presentations.
 

### 2017-08-23 thesis/chapters/ch1-intro.tex
Rearrange figures; edit

@@
 The combination of cheap electronics and electricity has made China one of the most profitable place to mine Bitcoins.
-At the time of writing, China holds about 80\% of the hashing power~\cite{bitcoinpool}.
-Effectively putting China in control of the 60 billion-dollar ecosystem.
+At the time of writing, China holds about 80\% of the hashing power~\cite{bitcoinpool},
+effectively putting it in control of the 60 billion-dollar ecosystem.
 If the government decides to flex its totalitarian muscles,

### 2017-08-23 thesis/chapters/ch2-problem.tex
Rearrange figures; edit

@@
 In fact, prior to Bitcoin, there was no Internet-scale consensus protocol.
-Bitcoin is also fault-tolerant because attackers can not outpace honest users in finding new blocks unless they have a majority of the computational power,
-as long as the majority is honest and reasonable measures are taken to prevent the selfish mining attack~\cite{eyal2014majority}.
+Bitcoin is also fault-tolerant because attackers can not outpace honest users in finding new blocks unless they have a majority of the computational power
+(assuming reasonable measures are taken to prevent the selfish mining attack~\cite{eyal2014majority}).
 The disadvantage, however, is that transactions are never in consensus with a probability of 1---no consensus finality.
@@
 where $t$ is typically just under a third of the committee size (a lower bound of Byzantine consensus~\cite{pease1980reaching}).
-This problem is especially difficult to solve because the committee is always much smaller than the population size which has more than $t$ Byzantine nodes.
+This problem is difficult to solve because the committee is always much smaller than the population size which typically has more than $t$ Byzantine nodes.
 Classical blockchain does not have this problem because they do not use Byzantine consensus.
@@
 A small shard size would result in a large number of inter-shard transactions which also hinder scalability.
-The authors of OmniLedger noted that inter-shard transactions have significantly higher latency compared to the consensus protocol~\cite{kokoris2017omniledger}.
 Furthermore, using smaller shards is a security risk.
@@
 Tangle~\cite{tangle}, Corda~\cite{corda} and the original TrustChain~\cite{multichain} avoid the need of global consensus.
-In these schemes, every node has their own chain.
+In these schemes, every node has their own hash chain.
 Transactions between nodes are recorded on their respective chains.
@@
 In the remainder of this section, we expand on our research question and visit each aspect in detail.
-At the end of this section, we describe some issues that we do not consider as a part of our design,
+At the end of this section, we describe some issues that out of the scope of this work,
 which is the result of not focusing on any particular application.
@@
 Having consensus is essential in blockchain systems.
-It stops many types of malicious activities because the agreed state, or a representation of it, must have the consent of the honest nodes in the network.
-We emphasise on the notion of a representation of the state because in contrary to traditional blockchains,
-it does not necessarily mean the set of all transactions that have ever happened.
-The state can be a small, but verifiable representation of all the transactions.
-As an extreme example, a single digest computed from the complete history can represent the state.
-It is verifiable as long as there are nodes that store the history.
+It stops many types of malicious activities because the agreed state,
+or a representation of it, must have the consent of the honest nodes in the network.
+For example, if two honest nodes attempt to verify the same transaction,
+it cannot be the case that one node thinks that the transaction is valid, but the other thinks it is invalid,
+as long as the number of malicious nodes is below the threshold that the system can tolerate.
 
@@
 Thus anything is possible from simply omitting messages to colluding with each other to undermine the whole system.
-
-There is another aspect of fault tolerance that is transaction verifiability.
-All transactions must be eventually verifiable to maintain system integrity.
-Further, the validation result must be consistent.
-For example, if two different honest nodes attempt to verify the same transaction,
-it cannot be the case that one node thinks that the transaction is valid, but the other thinks it is invalid.
-Furthermore, just like transactions in cryptocurrencies, they should be tamper-proof and irrefutable.
+As a result, just like cryptocurrencies, our transactions should be tamper-proof and irrefutable.
 
@@
 A good defence mechanism requires application specific information.
-For example, a social network based Sybil defence mechanisms use random walks in real-world social graphs~\cite{yu2006sybilguard}.
+For example, a social network based Sybil defence mechanism uses random walks in real-world social graphs~\cite{yu2006sybilguard}.
 Online marketplaces such as Amazon use the rating of buyer and sellers.
-Cryptocurrencies use PoW to decide on the latest block.
+Cryptocurrencies use PoW to decide on new blocks.
 % Since our system is application neutral, we do not provide an explicit Sybil defence mechanism.

### 2017-08-23 thesis/chapters/ch3-model.tex
Rearrange figures; edit

@@
 It is based on our prior research on TrustChain~\cite{multichain} at Delft,
-where every node independently interacts with each other via their own hash chain\footnote{TrustChain is originally called MultiChain but the name has changed to TrustChain}.
+where every node independently interacts with each other via their own hash chain\footnote{TrustChain is originally called MultiChain.}.
 
@@
 \subsection*{TrustChain and Extended TrustChain}
-We first describe the standard TrustChain and then the Extended TrustChain.
+We explain the standard TrustChain first and then the Extended TrustChain.
 Our description of the standard TrustChain has a minor difference when compared to the description in~\cite{multichain},
@@
 % \subsubsection*{Standard TrustChain}
-In TrustChain, every node has a ``personal'' hash chain. 
+In TrustChain, every node has a personal hash chain. 
 Initially, the chain only contains a genesis block generated by the nodes themselves.
-When node $A$ wishes to add a new transaction (TX) with $B$,
+When node $A$ wishes to make a transaction (TX) with $B$,
 a new TX block is generated for both $A$ and $B$ and is appended to their respective chains.
@@
 \begin{figure}
-    \includegraphics[width=0.9\textwidth]{trustchain-bad}
+    \includegraphics[width=1\textwidth]{trustchain-bad}
     \centering
@@
 A collection of CP blocks from all nodes represents the state of the whole system.
-Therefore, nodes need to reach consensus on CP blocks using some consensus protocol, which we explain next.
+Therefore, nodes need to reach consensus on CP blocks using some consensus protocol in order to reason about the state,
+which we explain next.
 A visual representation is shown in \Cref{fig:trustchain-bad-cp}.
@@
 \begin{figure}
-    \includegraphics[width=0.9\textwidth]{trustchain-bad-cp}
+    \includegraphics[width=1\textwidth]{trustchain-bad-cp}
     \centering
@@
 starting a new execution immediately after the previous one is completed.
+Nodes create new CP blocks at the end of every execution.
 This approach is necessary because blockchain systems always need to reach consensus on new values proposed by the nodes in the system,
@@
 After the algorithm completes,
-the facilitators output a set of CP blocks which we call the consensus result.
-The result is then propagated to the network,
-and the next algorithm execution begins.
+the facilitators output a set of CP blocks which we call the consensus result,
+which is then propagated to the network.
+Using the result, nodes are allowed to create new CP blocks,
+and then the next algorithm execution begins.
 
@@
 Which is a fragment of the counterparty's chain beginning and ending with CP blocks but contains the TX block belonging to that transaction,
-the CP blocks must also be in consensus.
+and the CP blocks must be in consensus.
 Upon the counterparty's response,
 the node checks whether the CP blocks are in some consensus result and some other conditions.
-The TX is valid if these conditions are satisfied.
+The transaction is valid if these conditions are satisfied.
 Intuitively, this works because it is hard (because a cryptographically secure hash function is collision resistant)
@@
 Thus in no stage of our protocol do we make timing assumptions.
-The adversary has full control of the delivery schedule and the message ordering of all messages.
+The adversary has full control of the delivery schedule and the ordering of all messages.
 % But they are not allowed to drop messages except for their own\footnote{
@@
 $n \ge 3t + 1$ must hold.
-This inequality is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$~\cite{pease1980reaching}.
+This inequality is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement for $t \ge n/3$~\cite{pease1980reaching}.
 Further, the inequality $N \ge n + t$ must also hold so that there are enough honest nodes in the network to run the consensus protocol;
@@
 and $h$ is the height of the block (i.e. $h = |B_u|$).
-We use $b_{u, h}$ to denote the latest block.
+We use $b_{u, h-1}$ to denote the latest block.
 There are two types of blocks, TX blocks and CP blocks.
@@
 
-The TX block is a six-tuple, i.e
+The TX block is a six-tuple, i.e.
 $$t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle.$$
@@
 $$t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$$
-there exist one and only one \emph{pair} 
+there exist one and only one pair 
 $$t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), seq_v, txid, pk_u, m, sig_v \rangle,$$
@@
 where the first round is defined to be 1 and it is incremented after every execution of the consensus protocol.
-The consensus result is a tuple, i.e. 
+The consensus result, output of the consensus protocol, is a tuple, i.e. 
 $$\C_r = \langle r, C \rangle,$$
@@
 
-Here we define an important property which results from the interleaving nature of CP and TX blocks.
+Now we define an important property which results from the interleaving nature of CP and TX blocks.
 It is used in our validation protocol.
@@
 If the enclosure of some TX block $t_{u, i}$ is $\langle c_{u,a}, c_{u, b} \rangle$,
-then its \emph{fragment} $F_{u, i}$ is defined as $\{ b_{u, i} : a \le i \le b \}$.
-Similarly, \emph{agreed fragment} has the same definition as fragment but using agreed enclosure.
+then its fragment $F_{u, i}$ is defined as $\{ b_{u, i} : a \le i \le b \}$.
+Similarly, agreed fragment has the same definition as fragment but using agreed enclosure.
 For convenience, the function $\textsf{agreed\_fragment}(t_{u, i})$ outputs the agreed fragment of $t_{u, i}$ if it exists, otherwise $\bot$.
@@
 The objectives of the protocol are to
-    allow honest nodes always make progress (in the form of creating new CP blocks),
-    compute correct consensus result in every round
+    allow honest nodes to make progress (in the form of creating new CP blocks),
+    compute correct the consensus result in every round
     and have an unbiased election of facilitators.
@@
     \item \emph{Agreement:}
-        If a correct node outputs $C$, then every node outputs $C$.
+        If a correct node outputs $C$, then every correct node outputs $C$.
     \item \emph{Validity:}
@@
 ACS has the nice property of censorship resilience when compared to other consensus algorithms.
-In PBFT~\cite{castro1999practical}, a leader is elected, if the leader is malicious but follows the protocol, then it can selectively filter transactions.
+In PBFT~\cite{castro1999practical}, a leader is elected, if the leader is malicious but follows the protocol, then it can selectively filter values.
 In contrast, every party in ACS are involved in the proposal phase,
-and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
-Thus, if some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
+and it guarantees that if $n - 2t$ parties propose the same value, then it must be in the agreed output.
+In other words, if some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
 For a detailed description of ACS, we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.
@@
 meaning that it satisfied the properties in \Cref{def:consensus}.
+In practice, the bootstrap oracle is most likely the software developer that sets up the system and assigns the facilitators of round 1 in a fair way.
 This concludes the bootstrap phase.
 For any future rounds, the consensus phase is used.
-
-In practice, the bootstrap oracle is most likely the software developer that sets up the system and assigns the facilitators of round 1 in a fair way.
 
@@
 For any node $u$,
-the consensus phase begins when $\F_r$ is available and the latest block is $c_{u, h}$.
+the consensus phase begins when $\F_r$ is available and the latest block is $c_{u, h-1}$.
 Note that $\F_r$ indicates the facilitators that were elected using results of round $r$ and are responsible for driving the ACS algorithm in round $r + 1$.
@@
 There are two scenarios in the consensus phase.
-First, if $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators.
+First, if $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h-1} \rangle$ to all the facilitators.
 Second, if $u$ is a facilitator, it waits until $N - t$ messages of type \texttt{cp\_msg} are received.
@@
 Continuing, when $\F_{r}$ finish the ACS execution and reach agreement on $\C_{r+1}$,
-they immediately broadcast two messages to all the nodes---
-first the consensus message $\langle \texttt{cons\_msg}, \C_{r+1} \rangle$,
+they immediately broadcast two messages to all the nodes---first the consensus message $\langle \texttt{cons\_msg}, \C_{r+1} \rangle$,
 and second the signature message $\langle \texttt{cons\_sig}, r, sig \rangle$.
@@
 so a new round can be started.
+
+Our protocol has some similarities with synchronizers~\cite[Chapter 10]{podc} because it is effectively a technique to introduce synchrony in an asynchronous environment.
+If we consider the facilitators as a collective authority, then it can be seen as a synchronizer that sends pulse messages (in the form of \texttt{cons\_msg} and \texttt{cons\_sig}) to indicate the start of a new clock pulse.
+Every node then sends a completion messages (in the form of \texttt{cp\_msg}) to the new collective authority to indicate that they are ready for the next pulse.
+
+A visual explanation can be found in \Cref{app:consensus-example},
+it walks through the steps needed for a node to be selected as a facilitator using an example.
 
@@
 
-Our protocol has some similarities with synchronizers~\cite[Chapter 10]{podc} because it is effectively a technique to introduce synchrony in an asynchronous environment.
-If we consider the facilitators as a collective authority, then it can be seen as a synchronizer that sends pulse messages (in the form of \texttt{cons\_msg} and \texttt{cons\_sig}) to indicate the start of a new clock pulse.
-Every node then sends a completion messages (in the form of \texttt{cp\_msg}) to the new collective authority to indicate that they are ready for the next pulse.
-
-A visual explanation can be found in \Cref{app:consensus-example},
-it walks through the steps needed for a node to be selected as a facilitator using an example.
-
 \section{Transaction protocol}
@@
 $txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$.
-Then the initiator then sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.
+Then the initiator sends $\langle \texttt{tx\_req}, t_{u, h}\rangle$ to $v$.
 
@@
 
-The consensus protocol and the validation protocol do not provide a mechanism to detect tampering.
+The consensus protocol and the transaction protocol do not provide a mechanism to detect tampering.
 The validation protocol aims to solve this issue.
-The protocol is also a request-response protocol, just like the transaction protocol.
-It assumes that nodes store all the consensus results that are aware of.
+The protocol is also a request-response protocol.
+It assumes that nodes store all the consensus results as well as the corresponding signatures that they are aware of.
+Further, we assume nodes that recovered from a crash or a corruption are able to recover its missed \texttt{cons\_msg} and \texttt{cons\_sig} messages,
+which can be achieved by querying other nodes.
 But before explaining the protocol itself, we first define what it means for a transaction to be valid.
@@
 \caption{Function $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ validates the transaction represented by $t_{u, i}$.
-We assume $F_{u, i}$ is always correct and contains $t_{u, i}$.
+We assume $F_{u, i}$ is correct and contains $t_{u, i}$.
 $F_{v, j}$ is the corresponding fragment received from $v$.}
@@
     \item \emph{Agreement}:
-        If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
+        If any correct node decides on the validity of a transaction, except when it is \emph{unknown},
         then all other correct nodes are able to reach the same conclusion or output \emph{unknown}.
@@
         The validation protocol outputs the correct result
-        according to the aforementioned validity definition.
+        according to the validity definition above.
     % \item \emph{Unforgeability}:

### 2017-08-23 thesis/chapters/ch4-analysis.tex
Rearrange figures; edit

@@
 But $n -t$ facilitators must have received $N - t$ valid CP blocks, so $|\C_{r+1}| \ge N-t$ and this proves (a).
-Since $N-t \ge (n+t) -t = n$ and $\textsf{get\_facilitators}(\C_{r+1})$ outputs $n$ items, so $|\F_{r+1}| = n$ and this proves (b).
+Since $N-t \ge (n+t) -t = n$ and $\textsf{get\_facilitators}(\C_{r+1}, n)$ outputs $n$ items, so $|\F_{r+1}| = n$ and this proves (b).
 \end{proof}
@@
 the output of $\textsf{H}(\cdot)$ is uniformly random.
-As a result, the output of $\textsf{get\_facilitators}(\C_{r+1})$ is a random permutation of all the nodes in $\C_{r+1}$ truncated to $n$.
+As a result, the output of $\textsf{get\_facilitators}(\C_{r+1}, n)$ is a random permutation of all the nodes in $\C_{r+1}$ truncated to $n$.
 Thus, every node has the same probability of becoming a facilitator.

### 2017-08-23 thesis/chapters/ch5-implementation.tex
Rearrange figures; edit

@@
 
+\afterpage{
 \begin{figure}[tb]
@@
 \end{figure}
-
+\clearpage
+}
 \Cref{fig:round-comms} shows the relationship between the communication cost of the consensus protocol per round and the population size.
@@
 \section{Linear global throughput}
+
+\afterpage{
 \begin{figure}[tb]
@@
 \end{figure}
+
+\begin{figure}[tb]
+  \centering
+  \makebox[\linewidth][c]{%
+  \begin{subfigure}[b]{0.7\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-fixed/timeseries}
+    \caption{Transactions are with fixed neighbours.}
+    \label{fig:backlog-fixed}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.7\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-random/timeseries}
+    \caption{Transactions are with random neighbours.}
+    \label{fig:backlog-random}
+  \end{subfigure}
+  }
+  \caption{Demonstration of the network bottleneck.
+  Note that the slope of validated transactions grows at a slower rate in~\Cref{fig:backlog-random} than in~\Cref{fig:backlog-fixed}.}
+  \label{fig:backlog}
+\end{figure}
+\clearpage
+}
 
@@
 
-\begin{figure}[tb]
-  \centering
-  \makebox[\linewidth][c]{%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-fixed/timeseries}
-    \caption{Transactions are with fixed neighbours.}
-    \label{fig:backlog-fixed}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.7\textwidth}
-    \includegraphics[width=\textwidth]{neighbour-random/timeseries}
-    \caption{Transactions are with random neighbours.}
-    \label{fig:backlog-random}
-  \end{subfigure}
-  }
-  \caption{Demonstration of the network bottleneck.
-  Note that the slope of validated transactions grows at a slower rate in~\Cref{fig:backlog-random} than in~\Cref{fig:backlog-fixed}.}
-  \label{fig:backlog}
-\end{figure}
-
 \section{Communication cost with varying number of facilitators}
@@
 
-\begin{figure}[tb]
+\Cref{fig:facilitators-round-comms} shows the communication cost of the consensus protocol for varying numbers of facilitators.
+There is substantial evidence that the communication cost grows polynomially.
+We expect this behaviour because of there are polynomial terms in the ACS communication complexity---$O(n^2|v| + \lambda n^3 \log n)$,
+which means that the rounds would take longer to complete and transactions would take longer to verify.
+On the other hand, we get better fault tolerance.
+Thus a linear increase in fault tolerance ($t$) costs a polynomial increase in communication cost.
+
+The same experiment viewed from the perspective of time is shown in~\Cref{fig:facilitators-round-duration}.
+Similarly, the duration increases polynomially for a linear increase in fault tolerance.
+
+\afterpage{
+\begin{figure}[bt]
   \centering
@@
 
-\Cref{fig:facilitators-round-comms} shows the communication cost of the consensus protocol for varying numbers of facilitators.
-There is substantial evidence that the communication cost grows polynomially.
-We expect this behaviour because of there are polynomial terms in the ACS communication complexity---$O(n^2|v| + \lambda n^3 \log n)$,
-which means that the rounds would take longer to complete and transactions would take longer to verify.
-On the other hand, we get better fault tolerance.
-Thus a linear increase in fault tolerance ($t$) costs a polynomial increase in communication cost.
+\begin{figure}[bt]
+  \centering
+  \makebox[\linewidth][c]{%
+  \begin{subfigure}[b]{0.7\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-fixed/round-duration-vs-facilitators}
+    \caption{Transactions are with fixed neighbours.}
+    \label{fig:facilitators-round-duration-fixed}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.7\textwidth}
+    \includegraphics[width=\textwidth]{neighbour-random/round-duration-vs-facilitators}
+    \caption{Transactions are with random neighbours.}
+    \label{fig:facilitators-round-duration-random}
+  \end{subfigure}
+  }
+  \caption{The consensus duration increases polynomially with respect to the number of facilitators.}
+  \label{fig:facilitators-round-duration}
+\end{figure}
+\clearpage
+}

### 2017-08-26 presentation/presentation.tex
Work on presentation

--- 2017-07-27 presentation/presentation.tex
@@
 \usepackage{graphicx}
-\graphicspath{{../figures/}}
+\graphicspath{{./figures/}{../figures/}}
 \usepackage{subfig}
@@
 \usepackage{algorithm}
+\usepackage{makecell}
+\usepackage[style=ieee]{biblatex}
+\addbibresource{references.bib}
 
@@
 \begin{frame}{Outline}
-  \tableofcontents
+  \tableofcontents[]
   % You might wish to add the option [pausesections]
@@
 \section{Introduction}
-\begin{frame}{Motivation}
+\subsection{The dangers of centralisation}
+\begin{frame}{The dangers of centralisation}
   \begin{itemize}
-    \item Blockchain systems offer an alternative to central authorities for the first time
-    \item Market cap (40 billion for Bitcoin) and trade volume figures indicate they are here to stay
-    \item Early blockchain systems are not scalable (7 TX/s for Bitcoin)
-    \item Parameter tuning leads to centralisation
+    \item Technological advancements give us convenience
+    \item But it puts central authorities in control
+    \item Many are motivated by profit
+    \item Not always in the interest of the ``users''\footnote{Typically users of some free service X are, in fact, used by X.}
   \end{itemize}
@@
 
-\begin{frame}{Research question}
+\begin{frame}{The dangers of centralisation:~Examples}
+  \begin{itemize}
+    \item Baidu's promoted search result on experimental medical care caused death of a student~\cite{weizexi}
+    \item Facebook can predict your opinions and desires better than your spouse~\cite{youyou2015computer}
+    \item With intimate knowledge of the individuals,
+      Facebook creates ``psychographic'' profiles in political campaigns~\cite{psychographic}
+  \end{itemize}
+  \note{
+    \begin{itemize}
+      \item The student had cancer---(synovial sarcoma)
+      \item The treatment was phased out at clinical stage in the US because the success rate is too low
+      \item ``The study was conducted on 86,220 volunteers who have a Facebook account and who
+      completed a hundred-item personality questionnaire. The Facebook algorithm predicted the volunteers'
+      answers based on monitoring their Facebook Likes – which webpages, images and clips they tagged with the
+      Like button. The more Likes, the more accurate the predictions. The algorithm's predictions were compared
+      with those of work colleagues, friends, family members and spouses. Amazingly, the algorithm needed a set
+      of only ten Likes in order to outperform the predictions of work colleagues. It needed seventy Likes to
+      outperform friends, 150 Likes to outperform family members and 300 Likes to outperform spouses. In other
+      words, if you happen to have clicked 300 Likes on your Facebook account, the Facebook algorithm can
+      predict your opinions and desires better than your husband or wife!'' -Harari, \emph{Homo Deus}
+      \item Summary: we live in an age where companies and governments are able to control our thoughts,
+      where is the freedom and liberty in that?
+    \end{itemize}
+  }
+\end{frame}
+
+\begin{frame}{Blockchain:~a new hope?}
+  \begin{itemize}
+    \item Blockchains are distributed ledgers
+    \item They enable large scale consensus
+    \item An alternative to central authorities for the first time
+    \item Some applications include:
+    \begin{itemize}
+      \item Digital cash (e.g., Bitcoin)
+      \item Domain name system (e.g., Namecoin)
+      \item File sharing (e.g., Filecoin)
+      \item General purpose (e.g., Ethereum)
+    \end{itemize}
+  \end{itemize}
+  \note{
+    \begin{itemize}
+      \item Explain blockchain systems---consensus
+    \end{itemize}
+  }
+\end{frame}
+
+\begin{frame}{Blockchain:~not there yet}
+  \begin{itemize}
+    \item Early blockchain systems do not scale
+    \item Bitcoin is limited to 7 transactions per second
+    \item 100,000 transaction backlog in May 2017
+    \item We require horizontal scalability for ubiquitous use
+    \item More nodes = better performance and throughput
+  \end{itemize}
+\end{frame}
+
+\subsection{Related work}
+\begin{frame}{Related work}
+  \begin{table}[htb]
+    \centering
+    \caption{Summary of the scalability properties of many blockchain systems.
+    Scalability gets better from left to right.}
+    \label{tab:scalability}
+    \begin{tabular}{|l|l|l|l|}
+    \hline
+    \textbf{\thead{Not\\scalable}} & \textbf{\thead{Somewhat\\scalable}} & \textbf{\thead{Limited horizontal\\scalability}} & \textbf{\thead{True horizontal\\scalability}} \\ \hline
+    Bitcoin  & Hyperledger & Elastico          & \textsc{Checo} (this work) \\
+    Ethereum & ByzCoin     & OmniLedger        &                \\
+    etc.     & Solidius    & Lightning Network &                \\
+    \hline
+    \end{tabular}
+\end{table}
+\end{frame}
+
+\begin{frame}{Related work}
+  State-of-the-art: Sharding
+  \begin{itemize}
+  \item Split state into multiple shards
+  \item Shards run consensus algorithm in parallel
+  \item Difficult to perform atomic inter-shard transactions
+    \begin{itemize}
+      \item Elastico: not possible
+      \item OmniLedger: via Atomic Commit protocol
+    \end{itemize}
+  \item Additional complexity
+  \item Diminishing return
+  \end{itemize}
+\end{frame}
+
+\subsection{Research question}
+\begin{frame}{\subsecname}
   \begin{block}{}
-    How do we design a \emph{blockchain consensus protocol} that is \emph{fault tolerant},
-    \emph{scalable} and can reach \emph{global consensus?}
+    \Large{
+    How can we design a \emph{blockchain consensus protocol} that is \emph{fault tolerant},
+    \emph{scalable} and able to reach \emph{global consensus?}
+    }
   \end{block}
@@
 
-\begin{frame}{Inspiration}
+\section{System architecture}
+\begin{frame}{Intuition}
   \begin{itemize}
-    \item A restaurant owner does not report all of its transactions with an central authority
-    \item Occasionally a customer may leave without paying and this event is reported to a central authority
-    \item Our blockchain system achieves scalability using a similar idea
+    \item Traditionally: to reach consensus and check the validity of all transactions is expensive
+    \item Our idea: we decouple consensus and validation
+    \item Use a single digest to represent an arbitrarily large number of transactions
+    \item Reach consensus on the small digest
+    \item Nodes then independently check the validity of the transactions of interest
   \end{itemize}
@@
 
-\section{System architecture}
 \subsection{System model}
@@
   \begin{itemize}
-    \item Population size is $N$
-    \item $n$ nodes are facilitators, $t$ nodes are malicious (Byzantine)
+    \item $N$ is the population size
+    \item Purely asynchronous channels with eventual delivery
+    \item $n$ nodes are facilitators
+    \item $t$ nodes are malicious, i.e. Byzantine
     \item $n \ge 3t + 1$
     \item $N \ge n + t$
-    \item Purely asynchronous channels with eventual delivery
-    \item Public key infrastructure
-    \item Random oracle model
-    \item Computational security
   \end{itemize}
-  \note{At this point, explain facilitators are nodes that have some extra task to do, temporarily.}
+  \note{At this point, explain facilitators are nodes that have some extra task to do, temporarily.
+  Explain that $N$ is typically a lot larger than $n$ instead of explaining $N \ge n + t$.}
 \end{frame}
@@
 \begin{frame}{\subsecname}
+  \begin{center}
+  \Large{The four components of \textsc{Checo}}
   \begin{figure}[h]
-  \includegraphics[width=0.7\textwidth]{architecture}
+  \includegraphics[width=0.7\textwidth]{architecture-wo-title}
   \centering
   \end{figure}
+  \end{center}
   \note{
@@
 \begin{frame}{\subsecname}
-  \begin{itemize}
-    \item Everyone has their own chain and genesis block
-    \item Two types of blocks
-      \begin{enumerate}
-      \item Transaction (TX) block
-      \item Checkpoint (CP) block
-      \end{enumerate}
-    \item A transaction involves two parties and results in two TX blocks (a pair)
-    \item A CP block captures the chain state
-    \item TX and CP blocks are chained together using hash pointers
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{\subsecname}
   \begin{figure}[h]
-  \includegraphics[width=0.67\textwidth]{trustchain-good-cp}
+  \includegraphics[width=0.75\textwidth]{trustchain-good-cp}
   \centering
-  \caption{TX block is a six-tuple: $t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), i, txid, pk_v, m, sig_u \rangle$.
-    CP block is a five-tuple: $c_{u, i} = \langle \textsf{H}(b_{u, i-1}), i, \textsf{H}(\C_r), r, sig_u \rangle$.}
   \end{figure}
@@
 
-\begin{frame}{\subsecname:~fragment definition}
-\begin{itemize}
-\item A fragment $F_{u, i}$ is defined on a TX block $t_{u, i}$
-\item It is a section of the chain beginning and ending with CP blocks that contains the TX
-\end{itemize}
+\begin{frame}{\subsecname:~TX block}
+  \begin{enumerate}
+    \item Hash pointer to the previous block
+    \item Sequence number
+    \item Transaction ID
+    \item Public key of the counterparty
+    \item Transaction message $m$
+    \item Signature the five items above
+  \end{enumerate}
+  \vfill
+  A transaction is represented by a \emph{pair} of TX blocks
+\end{frame}
+
+\begin{frame}{\subsecname:~CP block}
+  \begin{enumerate}
+    \item Hash pointer to the previous block
+    \item Sequence number
+    \item Digest of consensus result, i.e.~a set of CP blocks
+    \item Round number $r$
+    \item Signature on the four items above
+  \end{enumerate}
+\end{frame}
+
+\begin{frame}{\subsecname:~Fragment of a TX block}
+  \begin{figure}[h]
+  \includegraphics[width=0.75\textwidth]{trustchain-good-cp-frag}
+  \centering
+  \end{figure}
 \end{frame}
@@
 \subsection{Consensus protocol}
-\begin{frame}{\subsecname:~overview}
-  \begin{enumerate}
-    \item Runs in rounds
-    \item In round $r$, $n$ out of $N$ act as facilitators
-    \item The facilitators collect CP blocks from all nodes
-    \item Facilitators run a Byzantine consensus algorithm to agree on a set of CP blocks
-    \item Disseminate the consensus result $\C_r$, i.e. the CP blocks
-    \item From $\C_r$, new facilitators are computed
-    \item Repeat
-  \end{enumerate}
-\end{frame}
 
-\begin{frame}{\subsecname:~properties}
-\label{def:consensus}
-$\forall r \in \mathbb{N}$, the following properties must hold.
+\begin{frame}{\subsecname---Background on ACS}
   \begin{itemize}
-    \item \emph{Agreement}:
-        If one correct node outputs a set of facilitators $\F_r$,
-        then every node outputs $\F_r$
-    \item \emph{Validity}:
-        If any correct node outputs $\F_r$, then 
-            \begin{enumerate}
-                \item $|\C_r| \ge N - t$ must hold for the $\C_r$ which was used to create $\F_r$,
-                \item $\F_r$ must contain at least $n - t$ honest nodes and
-                \item $|\F_r| = n$.
-            \end{enumerate}
-    \item \emph{Fairness}:
-        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
-    \item \emph{Termination}:
-        Every correct node eventually outputs some $\F_r$.
+    \item Asynchronous common subset
+    \item A simplification of HoneyBadgerBFT~\cite{miller2016honey}
+    \item $n$ nodes
+    \item $t$ nodes may be malicious
+    \item Input: every node proposes a set of values, e.g., $\{A, B\}, \{B, C\}, \dots$
+    \item Output: set union of the majority, e.g., $\{A, B, C, \dots \}$
   \end{itemize}
@@
   \begin{figure}[h]
-  \includegraphics[width=0.67\textwidth]{trustchain-1}
+  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-1}
   \centering
-  \caption{Suppose we are in a state where $\C_{r - 1}$ has just been agreed by some facilitators but not yet propagated.}
+  \end{figure}
+  \note{Suppose we are in a state where $\C_{r - 1}$ has just been agreed by some facilitators but not yet propagated.}
+\end{frame}
+
+\begin{frame}{\subsecname}
+  \begin{figure}[h]
+  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-2}
+  \centering
+  \end{figure}
+  \note{Nodes receive consensus result $\C_{r - 1}$,
+    first $n$ nodes ordered by $\textsf{H}(\C_{r-1} || pk)$ become $\F_{r-1}$,
+    send the new CP blocks to $\F_{r-1}$.}
+\end{frame}
+
+\begin{frame}{\subsecname}
+  \begin{figure}[h]
+  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-3}
+  \centering
   \end{figure}
@@
   \begin{figure}[h]
-  \includegraphics[width=0.67\textwidth]{trustchain-2}
+  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-4}
   \centering
-  \caption{Nodes receive consensus result $\C_{r - 1}$,
-    first $n$ nodes ordered by $\textsf{H}(\C_{r-1} || pk)$ become $\F_{r-1}$,
-    send the new CP blocks to $\F_{r-1}$.}
   \end{figure}
@@
   \begin{figure}[h]
-  \includegraphics[width=0.67\textwidth]{trustchain-3}
+  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-5}
   \centering
-  \caption{Transactions carry on as usual in round $r$,
+  \end{figure}
+  \note{Transactions carry on as usual in round $r$,
   while facilitators are trying to reach consensus on the new CP blocks concurrently.}
-  \end{figure}
 \end{frame}
@@
   \begin{figure}[h]
-  \includegraphics[width=0.67\textwidth]{trustchain-4}
+  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-6}
   \centering
-  \caption{$\F_{r-1}$ agree and disseminate $\C_r$,
+  \end{figure}
+  \note{$\F_{r-1}$ agree and disseminate $\C_r$,
   CP blocks at round $r-1$ ($c_{a, 6}, c_{b, 4}, c_{c,8}$) should be in $\C_r$.}
-  \end{figure}
+\end{frame}
+
+\begin{frame}{\subsecname:~properties}
+\label{def:consensus}
+The consensus protocol has the following properties in every round $r$.
+\begin{enumerate}
+    \item \emph{Agreement}:
+        Every correct outputs the same set of facilitators.
+    \item \emph{Validity}:
+      The consensus results is valid such that a new set of facilitators can be computed from it.
+    \item \emph{Fairness}:
+        Every node with a CP block in the consensus result should have an equal probability of becoming a facilitator.
+    \item \emph{Termination}:
+        Every correct node eventually outputs a set of facilitators.
+\end{enumerate}
 \end{frame}
@@
 \begin{frame}{\subsecname}
+  \begin{figure}[h]
+  \includegraphics[width=1.0\textwidth]{tx-proto}
+  \centering
+  \end{figure}
 \begin{itemize}
-\item Request (\texttt{tx\_req}) and response (\texttt{tx\_resp}) protocol
+\item Request and response protocol
 \item Two TX blocks containing the same $txid$ are generated
@@
 
+\subsection{Validation protocol}
 \begin{frame}{\subsecname}
-Create $t_{u, h}$ and send $\langle \texttt{tx\_req}, t_{u, h} \rangle$ to start a transaction.
-\vspace{5 mm}
-\begin{algorithmic}
-    \Upon $\langle \texttt{tx\_req}, t_{v, j} \rangle$ from $v$
-    \State $\langle \_, \_, txid, pk_v, m, \_ \rangle \gets t_{v, j}$ \Comment unpack $t_{v, j}$
-    \State $\textsf{new\_tx}(pk_u, m, txid)$ \Comment create and store $t_{u, i}$
-    \State store $t_{v, j}$ as the pair of $t_{u, h}$
-    \State send $\langle \texttt{tx\_resp}, t_{u, h} \rangle$ to $v$
-
-    \Upon $\langle \texttt{tx\_resp}, t_{v, j} \rangle$ from $v$  
-    \State $\langle \_, \_, txid, pk_v, m, \_ \rangle \gets t_{v, j}$ \Comment unpack $t_{v, j}$
-    \State store $t_{v, j}$ as the pair of the TX with identifier $txid$
-\end{algorithmic}
-\end{frame}
-
-\subsection{Validation protocol}
-\begin{frame}{\subsecname:~overview}
+  \begin{figure}[h]
+  \includegraphics[width=1.0\textwidth]{vd-proto}
+  \centering
+  \end{figure}
 \begin{itemize}
-\item Request (\texttt{vd\_req}) and response (\texttt{vd\_resp}) protocol
-\item Transactions are in three states---\emph{valid}, \emph{invalid} and \emph{unknown}---defined by $\textsf{get\_validity}(\cdot)$
-\item The goal is to identify which state a given transaction is in satisfying the validation protocol properties
+\item Alice needs the two fragments that belong to the transaction
+\item Validation function checks whether the fragments are OK and contain the transaction
+\item Non-blocking
 \end{itemize}
-\end{frame}
-
-\begin{frame}{Validity definition}
-\textbf{Function} $\textsf{get\_validity}(t_{u, i}, F)$ validates the transaction $t_{u, i}$
-\vspace{5 mm}
-\begin{algorithmic}
-    \State Check that $v$ sent the correct $F$,
-    otherwise return \emph{unknown}.
-
-    \State
-    \State $\langle \_, \_, txid, pk_v, m, \_ \rangle \gets t_{u, i}$
-    \If{number of blocks of $txid$ in $F \ne 1$}
-        \State \Return \emph{invalid}
-    \EndIf
-    \State \Comment TX exists
-
-    \State $\langle \_, \_, txid', pk'_u, m', \_ \rangle \gets t_{v, j}$
-    \If{$m \ne m' \vee pk_u \ne pk'_u$}
-        \State \Return \emph{invalid}
-    \EndIf
-    \State \Comment no tampering
-    \State \Return \emph{valid}
-\end{algorithmic}
 \end{frame}
@@
 \begin{itemize}
-    \item \emph{Agreement}:
-        If any correct node decides on the validity of a transaction (except when it is \emph{unknown}),
-        then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
-    \item \emph{Correctness}:
-        The validation protocol outputs the correct result
-        according to the aforementioned validity definition.
-    \item \emph{Liveness}:
-        Any valid (invalid) transaction is marked as validated (invalid) eventually.
-\end{itemize}
-\note{Liveness is not possible in our case because malicious nodes can simply go offline or refuse to respond,
-but the other two properties are possible.}
-\end{frame}
-
-\begin{frame}{\subsecname}
-Send $\langle \texttt{vd\_req}, txid \rangle$ to $v$ to begin.
-\vspace{5 mm}
-\begin{algorithmic}
-    \Upon $\langle \texttt{vd\_req}, txid \rangle$ from $v$
-        \State $t_{u, i} \gets \text{the transaction identified by } txid$
-        \State $F_{u, i} \gets \textsf{agreed\_fragment}(t_{u, i})$
-        \State send $\langle \texttt{vd\_resp}, txid, F_{u, i} \rangle$ to $v$
-
-    \Upon $\langle \texttt{vd\_resp}, txid, F_{v, j} \rangle$ from $v$
-        \State $t_{u, i} \gets \text{the transaction identified by } txid$
-        \State set the validity of $t_{u, i}$ to $\textsf{get\_validity}(t_{u, i}, F_{v, j})$
-\end{algorithmic}
-\end{frame}
-
-\section{Analysis of correctness and performance}
-
-\subsection{Correctness of the consensus protocol}
-\begin{frame}{\subsecname}
-\begin{theorem}
-For all rounds,
-the consensus protocol satisfies agreement, validity, fairness and termination properties.
-\end{theorem}
-\begin{proof}(sketch)
-Because consensus result are eventually delivered and the properties of Byzantine consensus,
-we get agreement, validity and termination.
-Fairness is from the fact that we model $\textsf{H}(\cdot)$ as a random oracle (RO) and the input to the RO is different for every node,
-thus the list of nodes ordered by $\textsf{H}(\C_r || pk )$ is a random permutation of those nodes.
-\end{proof}
-\end{frame}
-
-\subsection{Correctness of the validation protocol}
-\begin{frame}{\subsecname}
-\begin{theorem}
-  The validation protocol satisfies agreement and correctness properties.
-\end{theorem}
-\begin{proof}(sketch)
-Proof by contradiction.
-For this attack to work, the adversary must be able to create two different fragments but with the same checkpoint enclosure.
-We model $\textsf{H}(\cdot)$ as a RO, so the adversary need to query the RO a exponential\footnote{In terms of the security parameter.} number of times.
-But the adversary can only query the RO a polynomial number of times.
-Hence we have agreement.
-Correctness follows directly the use of the $\textsf{get\_validity}(\cdot)$ function.
-\end{proof}
-\end{frame}
-
-\subsection{Linear global throughput argument}
-\begin{frame}{\subsecname}
-\begin{itemize}
-  \item Throughput has a notion of time but our model is asynchronous
-  \item Additional assumptions needed to make the argument---every unit of communication takes a non-negligible amount of time to process
-  \item Bandwidth relation---$NC \ge r_{\text{tx}} l$, where $l$ is $O(N)$
-  \item If $r_\text{tx}$ satisfies the inequality, then LHS and RHS grows at the same rate,
-  thus we have linear global throughput
+  \item CP blocks of the fragments are ``anchored'' due to the consensus protocol
+  \item It is difficult to modify the fragment once ``anchored''
+  \item Implicit consensus on transactions via CP blocks
 \end{itemize}
@@
   \begin{itemize}
-    \item Prototype implementation on Github\footnote{\url{https://github.com/kc1212/consensus-thesis-code}}
+    \item Free and open source implementation on Github:
+      \url{https://github.com/kc1212/checo}
     \item SHA256 for hash functions and Ed25519 for digital signature
     \item Experiment on the DAS-5\footnote{\url{http://www.cs.vu.nl/das5/}}
-      \begin{itemize}
-        \item Up to 30 machines
-        \item Each running 40 nodes
-      \end{itemize}
   \end{itemize}
@@
   \begin{block}{Research question}
-    How do we design a \emph{blockchain consensus protocol} that is \emph{fault tolerant},
-    \emph{scalable} and can reach \emph{global consensus?}
+    How can we design a \emph{blockchain consensus protocol} that is \emph{fault tolerant},
+    \emph{scalable} and able to reach \emph{global consensus?}
   \end{block}
@@
 
+\begin{frame}{Bibliography}
+\printbibliography
+\end{frame}
+
+\section*{Extras}
 \end{document}
+
+% Sybils
+% facilitator set?
+% incentivise?
+% attack (predict) of facilitator
+% appoint facilitator
+% liveness / non-responsive / slow
+% drawbacks
+%

### 2017-08-26 presentation/presentation.tex
Add large throughput graph

@@
   \begin{itemize}
-    \item Baidu's promoted search result on experimental medical care caused death of a student~\cite{weizexi}
     \item Facebook can predict your opinions and desires better than your spouse~\cite{youyou2015computer}
@@
       Facebook creates ``psychographic'' profiles in political campaigns~\cite{psychographic}
+    \item Baidu's promoted search result on experimental medical care caused death of a student~\cite{weizexi}
   \end{itemize}
@@
   \begin{itemize}
-    \item Early blockchain systems do not scale
-    \item Bitcoin is limited to 7 transactions per second
+    \item Consensus algorithm of early blockchain systems do not scale
+    \item Bitcoin (PoW) is limited to 7 transactions per second
     \item 100,000 transaction backlog in May 2017
@@
     \Large{
-    How can we design a \emph{blockchain consensus protocol} that is \emph{fault tolerant},
-    \emph{scalable} and able to reach \emph{global consensus?}
+    How can we design a \emph{blockchain consensus protocol} that is \emph{fault-tolerant},
+    \emph{horizontally scalable}, and able to reach \emph{global consensus?}
     }
@@
 The consensus protocol has the following properties in every round $r$.
-\begin{enumerate}
+\begin{itemize}
     \item \emph{Agreement}:
@@
         Every correct node eventually outputs a set of facilitators.
-\end{enumerate}
+\end{itemize}
 \end{frame}
@@
 
+\begin{frame}{Throughput vs population size (fixed neighbour)}
+  \begin{figure}[h]
+  \includegraphics[width=0.9\textwidth]{throughput-vs-population-large}
+  \centering
+  \end{figure}
+\end{frame}
+
 \section{Conclusion}
 \begin{frame}{\secname}
-  \begin{block}{Research question}
+  Out work answers the research question.
+  \begin{block}{}
     How can we design a \emph{blockchain consensus protocol} that is \emph{fault tolerant},
-    \emph{scalable} and able to reach \emph{global consensus?}
+    \emph{horizontally-scalable}, and able to reach \emph{global consensus?}
   \end{block}
-Our system achieve the following
-\begin{itemize}
-  \item Fault tolerant up to $t$ nodes
-  \item Horizontal scalability
-  \item Global consensus on CP blocks
-\end{itemize}
 \end{frame}
@@
 \begin{itemize}
+\item Implement and experiment with a concrete application
+\item Analyse the system in the permissionless environment
 \item Improve fault tolerance
-\item Improve fork detection
-\item Analyse the system in the permissionless environment
-\item Concrete application
 \end{itemize}

### 2017-08-27 paper/s1-intro.tex
Fix paper

@@
 
-% related work
+\textbf{Related work.\quad}
 Many approaches exist for improving the scalability of early blockchains.
@@
 
-% research question
+\textbf{Research question.\quad}
 Thus far, there are no systems that achieve horizontal scalability in the general case,
@@
 \begin{displayquote}
-\emph{Is it possible to design a blockchain consensus protocol that is fault-tolerant and horizontally scalable?}
+    \emph{How can we design a blockchain consensus protocol that is fault-tolerant,
+    horizontally scalable,
+    and able to reach global consensus?}
     \end{displayquote}
+We remark the different aspects of our research question.
 A blockchain consensus protocol should be application neutral,
@@
 
-% contribution
+\textbf{Contribution.\quad}
 The key insight is to \emph{not} reach consensus using an existing consensus algorithm (a modified HoneyBadgerBFT~\cite{miller2016honey}) on transactions themselves,
@@
 
-% organisation
+\textbf{Roadmap.\quad}
 In~\Cref{sec:description},

### 2017-08-27 paper/s2-problem.tex
Fix paper

@@
 We introduce the problem as a modified Byzantine consensus problem.
+The modification is primarily derived from the need of horizontal scalability,
+which is not a part of a typical Byzantine consensus problem.
 In our model, we consider $N$ nodes, $t$ of which are Byzantine.
 Nodes in our system make transactions with each other.
-Transactions can be in one of three states---\emph{valid}, \emph{invalid} and \emph{unknown}, defined by some function $\textsf{get\_validity}(\cdot)$.
-We seek a protocol that reasonably defines these states given a transaction,
-but also satisfies the following properties.
+Transactions can be in one of three states---\emph{valid}, \emph{invalid} and \emph{unknown}.
+We seek a protocol that satisfies the following properties.
 \begin{itemize}
@@
         then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
-    \item \emph{Validity}:
-        The protocol outputs the correct result according to $\textsf{get\_validity}(\cdot).$
     \item \emph{Scalability}:
@@
 
-Note that the agreement and validity property are similar to what is often seen in Byzantine consensus problems,
-but there are subtle differences and relaxations.
-Firstly, the agreement property only holds if honest nodes do not output \emph{unknown}.
+Note that the agreement property is similar, but a relaxed version of what is often seen in Byzantine consensus problems.
+Namely, the property only holds if honest nodes do not output \emph{unknown}.
 For example, for a transaction, it is OK if two honest nodes output \emph{valid} and \emph{unknown}, but it is not OK to output \emph{valid} and \emph{invalid}.
-Secondly, the validity is a relaxation of what is typically seen in Byzantine consensus problems,
-namely we do not make any guarantee on whether the transaction is \emph{valid} even if all nodes are honest,
-only that it is correct according to $\textsf{get\_validity}(\cdot)$ defined by the protocol.
+Our problem does not have a termination property, also known as liveness.
+Instead, nodes are incentivised to complete the protocol execution, we describe this phenomenon in~\Cref{sec:correctness-of-validity}.
 
@@
 Due to this fact, we also do not consider global fork prevention or detection,
-as some application may not need such strong guarantees such as recording the number of bytes uploaded and downloaded in Tribler~\cite{pimotte, pouwelse2008tribler}.
+as some application may not need such strong guarantees such as the accounting of detailed internet traffic in Tribler~\cite{pimotte, pouwelse2008tribler}.
 On the other hand, we give two alternative constructions that do have fork detection in~\Cref{sec:fork}.
 
-\textbf{System model.}
+\textbf{System model.\quad}
 We assume purely asynchronous channels with eventual delivery.
@@
 
-\textbf{Security assumptions.}
+\textbf{Security assumptions.\quad}
 The malicious nodes are Byzantine,

### 2017-08-27 paper/s3-system.tex
Fix paper

@@
 
-To describe our system, we first give an informal overview and then move on to the formal description.
-Unlike most blockchain systems, we do not use a global ledger.
-Instead, every node has their own hash chain.
+To describe \textsc{Checo}, we first give an informal overview and then move on to the formal description.
+
+Early blockchain systems that use a global ledger are difficult to scale because every node must reach consensus on all the transactions that ever existed.
+Instead, we introduce an alternative architecture where every node has their own genesis block and hash chain.
 The nodes only store transactions (TX) that they are involved in on their hash chain.
 Transactions are stored in TX blocks and every block only contains one transaction.
+A transaction between two nodes should therefore result in two TX blocks on the hash chains of those nodes.
 We introduce a special block called checkpoint (CP) block,
 which represents the state of a hash chain in the form of a hash pointer.
-Such data structure is called Extended TrustChain,
-a visualisation can be seen in~\Cref{fig:trustchain-good-cp}.
+Then, a collection of CP blocks from all nodes would represent the state of the whole system.
+A visualisation can be seen in~\Cref{fig:trustchain-good-cp}.
 \begin{figure}[h]
@@
 \includegraphics[width=\linewidth]{trustchain-good-cp}
-\caption{Visualisation of the Extended TrustChain. $t_{u, i}$ represents a TX block on $u$'s chain with a sequence number of $i$.
+\caption{Visualisation of the data structure used in \textsc{Checo}.
+$t_{u, i}$ represents a TX block on $u$'s chain with a sequence number of $i$.
 $c_{v, j}$ represents a CP block on $v$'s chain with a sequence number of $j$.
-A transaction is represented by a \emph{pair} of TX blocks,
-one on each of the party's chain.
-}
+The blocks at the ends of the dotted lines are pairs of each other.
+Blocks of sequence number 0 (e.g. $t_{b, 0}$) are genesis blocks.}
 \label{fig:trustchain-good-cp}
@@
 
-We randomly elect $n$ special nodes called facilitators in every round.
-The facilitators reach consensus on CP blocks using an existing consensus algorithm---asynchronous subset consensus from HoneyBadgerBFT~\cite{miller2016honey}.
-The consensus result is essentially a set of CP blocks from (nearly) all the nodes.
-While this gives us a consensus on a representation of the global state, it does not imply consensus on transactions.
-To this end,
-we introduce a validation protocol that allows any nodes to challenge any other node to produce a set of TX blocks that computes to the CP block in consensus,
-achieving implicit consensus on transactions.
+\textsc{Checo} consist of three protocols---consensus protocol, transaction protocol and validation protocol---all interacting with the data distributed hash chain structure described above.
+The primary protocol is the consensus protocol,
+which can be seen as a technique of running infinitely many times of an existing Byzantine consensus algorithm (in this work we use asynchronous common subset from HoneyBadgerBFT~\cite{miller2016honey}),
+starting a new execution immediately after the previous one is completed.
+Nodes create new CP blocks at the end of every execution.
+This approach is necessary because blockchain systems always need to reach consensus on new values proposed by the nodes in the system,
+or CP blocks in our case.
+
+The communication complexity of Byzantine consensus algorithms typically grows polynomially w.r.t the number of nodes,
+which prohibits us from running it on a large network.
+Thus, at the beginning of every Byzantine consensus algorithm execution,
+we randomly elect a set of nodes---called facilitators---to collect CP blocks from every other node and use those blocks as the input to the Byzantine consensus algorithm, and then run it.
+After the algorithm completes,
+the facilitators output a set of CP blocks which we call the consensus result,
+which is then propagated to the network.
+Using the result, nodes are allowed to create new CP blocks,
+and then the next algorithm execution begins.
+
+The transaction protocol is a simple request and response protocol.
+The nodes exchange one round of messages and create new TX blocks on their respective chains.
+Thus, as we mentioned before, one transaction should result in two TX blocks.
+
+The consensus and transaction protocol by themselves do not provide a mechanism to detect malicious behaviour such as tampering.
+Thus, we need a validation protocol to counteract such behaviour.
+When a node wishes to validate one of its transactions, it asks the counterparty for the \emph{agreed fragment} of the transaction.
+Which is a fragment of the counterparty's chain beginning and ending with CP blocks but contains the TX block belonging to that transaction,
+and the CP blocks must be in consensus.
+Upon the counterparty's response,
+the node checks whether the CP blocks are in some consensus result and among other conditions.
+The transaction is valid if these conditions are satisfied.
+% Intuitively, this works because it is hard (because a cryptographically secure hash function is collision resistant)
+% to create a different chain that begins and ends with the same two CP blocks.
 Since the transaction and validation protocols only make point-to-point communication,
@@
 
-The formal description is given next,
-it consists of one data structure---Extended TrustChain,
-and three protocols---consensus protocol, transaction protocol and validation protocol.
-The complete protocol can be seen as the concurrent composition of the sub-protocols.
-
-\subsection{Extended TrustChain}
+\subsection{\textsc{Checo} data structure}
 Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
-The chain consists of blocks $B_u = \{ b_{u, i} : i \in \{ 0, \dots, h - 1 \} \}$,
+The chain consist of blocks $B_u = \{ b_{u, i} : i \in \{ 0, \dots, h - 1 \} \}$,
 where $b_{u, i}$ is the $i$th block of $u$,
 and $h$ is the height of the block (i.e. $h = |B_u|$).
-We often use $b_{u, h}$ to denote the latest block.
+We use $b_{u, h-1}$ to denote the latest block.
 There are two types of blocks, TX blocks and CP blocks.
 If $T_u$ is the set of all TX blocks in $B_u$ and $C_u$ is the set of all CP blocks is $B_u$,
-then it must be the case that $T_u \cup C_u = B_u$ and $T_u \cap C_u = \varnothing$.
+then $T_u \cup C_u = B_u$ and $T_u \cap C_u = \varnothing$.
 The notation $b_{u, i}$ is generic over the block type.
-We assume there exists a function $\textsf{typeof}: B_u \rightarrow \{ \tau, \gamma \}$ that returns the type of the block,
-where $\tau$ represents the TX type and $\gamma$ represents the CP type.
 
@@
 \begin{definition}
-\textbf{\emph{Checkpoint block}}
-
+\textbf{\emph{Checkpoint block and genesis block}}
 The CP block is a five-tuple, i.e. 
 $$c_{u, i} = \langle \textsf{H}(b_{u, i-1}), seq_u, \textsf{H}(\C_r), r, sig_u \rangle,$$
-where $\C_r$ is the consensus result (which we describe in \Cref{sec:consensus-result}) in round $r$,
+where $\C_r$ is the consensus result (which we describe next in \Cref{sec:consensus-result}) in round $r$,
 the other items are the same as the TX block definition.
-% Note that unlike in our prior work~\cite{implicitconsensus}, CP blocks and TX blocks do not have independent sequence numbers.
 
@@
 \textbf{\emph{Consensus result}}
-
-Our consensus protocol runs in rounds.
-Every round is identified by a round number $r$, which is incremented on every new round.
-The consensus result is a tuple, i.e.
+Our consensus protocol runs in rounds,
+where the first round is defined to be 1 and it is incremented after every execution of the consensus protocol.
+The consensus result, output of the consensus protocol, is a tuple, i.e. 
 $$\C_r = \langle r, C \rangle,$$
@@
 It is used in our validation protocol (discussed in~\Cref{sec:vd-protocol}).
+
 \begin{definition}
 \textbf{\emph{Enclosure and agreed enclosure}}
-
-If there exist a tuple $\langle c_{u,a}, c_{u, b} \rangle$ for a TX block $t_{u, i}$,
-where 
-$$a = \argmin_{k, k < i, \textsf{typeof}(b_{u,k}) = \gamma}(i - k)$$
-$$b = \argmin_{k, k > i, \textsf{typeof}(b_{u,k}) = \gamma}(k - i),$$
+If there exist a tuple $\langle c_{u,a}, c_{u, b} \rangle$ for a TX block $t_{u, i}$, where
+\begin{itemize}
+\item $c_{u, a}$ is the closest CP block to $t_{u, i}$ with a lower sequence number and
+\item $c_{u, b}$ is the closest CP block to $t_{u, i}$ with a higher sequence number,
+\end{itemize}
 then $\langle c_{u,a}, c_{u, b} \rangle$ is the enclosure of $t_{u, i}$.
-Intuitively, that is the two closest CP blocks that surround $t_{u, i}$.
-Some TX blocks may not have any enclosure, then its enclosure is $\bot$.
+Some TX blocks may not have any enclosure, then their enclosure is $\bot$.
 Agreed enclosure is the same as enclosure with an extra constraint where the CP blocks must be in some consensus result $\C_r$.
@@
 \subsection{Consensus Protocol}
-Our scalable consensus protocol runs on top of Extended TrustChain.
-It uses an unmodified asynchronous subset consensus (ACS) algorithm as the key building block.
+Our scalable consensus protocol $\Pi_c$ uses an unmodified asynchronous subset consensus (ACS) algorithm as the key building block.
 The objectives of the protocol are to
@@
 and have an unbiased election of facilitators.
-Concretely, we define the necessary properties as follows.
+We formally define the desired properties below.
 \begin{definition}
 \label{def:consensus}
-\textbf{\emph{Properties of the consensus protocol}}
+\textbf{\emph{\textsc{Checo} consensus protocol}}
 
-$\forall r \in \mathbb{N}$, the following properties must hold.
+A \textsc{Checo} consensus protocol is correct if the following holds for every round $r$.
 \begin{enumerate}
@@
             \begin{enumerate}
-                \item $|\C_r| \ge N - t$ and each CP in $\C_r$ belong to a different node\footnote{
-                $\C_r$ is a tuple but we abuse the notation here by writing $|\C_r|$ to mean the number of CP blocks in the second element of $\C_r$.}.
-                \item $\F_r$ must contain at least $n - t$ honest nodes and
+                \item $|\C_r| \ge N - t$\footnote{
+                $\C_r$ is a tuple but we abuse the notation here by writing $|\C_r|$ to mean the number of CP blocks in the second element of $\C_r$.}, and
                 \item $|\F_r| = n$.
             \end{enumerate}
+    \item \emph{Fairness}:
+        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
     \item \emph{Termination}:
         Every correct node eventually outputs some $\F_r$.
-    \item \emph{Fairness}:
-        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
 \end{enumerate}
 \end{definition}
-These properties are similar to Byzantine consensus properties but there are subtle differences.
-Firstly, they are properties for every node in the network and not just the facilitators.
-Secondly, they must be satisfied for all rounds,
-because a failure in one round will lead to failures in subsequent rounds.
 
@@
 \label{sec:bootstrap}
-As with many distributed systems, there must be a bootstrap phase which sets up the system.
-Imagine that there is some bootstrap oracle, that initiates the code on every node.
-The code satisfies all the properties in \Cref{def:consensus}.
-Namely, every node has the same set of valid facilitators $\F_1$ that are randomly chosen.
+To bootstrap, imagine that there is some bootstrap oracle that initiates the correct program on every node,
+meaning that it satisfied the properties in \Cref{def:consensus}.
+In practice, the bootstrap oracle is most likely the software developer that sets up the system and assigns the facilitators of round 1 in a fair way.
 This concludes the bootstrap phase.
@@
 
-The TX protocol, shown in \Cref{alg:tx-proto}, is run by all nodes.
-It is also known as True Halves, first described informally by Veldhuisen~\cite[Chapter~3.2]{truehalves}.
+The TX protocol $\Pi_t$ is shown in \Cref{alg:tx-proto}.
 Nodes that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
@@
 Up to this point, we do not provide a mechanism to detect tampering.
-The validation protocol aims to solve this issue.
+The validation protocol $\Pi_v$ aims to solve this issue.
 The protocol is also a request-response protocol.
@@
 \label{def:validation}
-\textbf{\emph{Properties of the validation protocol}}
+\textbf{\emph{\textsc{Checo} validation protocol}}
 
+A \textsc{Checo} validation protocol is correct if the following properties hold.
 \begin{enumerate}
     \item \emph{Agreement}:
-        If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
-        then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
+        If any correct node decides on the validity of a transaction, except when it is \emph{unknown},
+        then all other correct nodes are able to reach the same conclusion or output \emph{unknown}.
     \item \emph{Validity}:
         The validation protocol outputs the correct result
-        according to the aforementioned validity definition.
+        according to the validity definition above.
     % \item \emph{Unforgeability}:
@@
 
-\subsection{Design Variations and Tradeoffs}
+\section{Design Variations and Tradeoffs}
 \label{sec:tradeoffs}
@@
 
-\subsubsection{Using Timing Assumption in the Permissionless Setting}
+\subsection{Using Timing Assumption in the Permissionless Setting}
 \label{sec:permissionless}
@@
 
-\subsubsection{Optimising Validation Protocol Using Cached Agreed Fragments}
+\subsection{Optimising Validation Protocol Using Cached Agreed Fragments}
 \label{sec:caching}
@@
 
-\subsubsection{Total Fork Detection}
+\subsection{Total Fork Detection}
 \label{sec:fork}

### 2017-08-27 paper/s5-evaluation.tex
Fix paper

@@
 
-\begin{figure*}[t]
+\begin{figure*}[ht]
   \centering
@@
 
-The prototype implementation can be found on GitHub.
-\begin{displayquote}
-\url{https://github.com/kc1212/checo}
-\end{displayquote}
+A free and open source implementation can be found on GitHub: \url{https://github.com/kc1212/checo}.
 It implements the three protocols and the Extended TrustChain.
 We also implement the caching optimisation discussed in \Cref{sec:caching}.
-It is written in the Python programming language\footnote{\url{https://www.python.org/}}.
+It is written in the Python programming language.
 The cryptography primitives we use are SHA256 for hash functions and Ed25519 for digital signatures.
-Both of which are provided by libnacl~\footnote{\url{https://pypi.python.org/pypi/libnacl}}.
 
@@
 Validation requests are sent at the same rate, which explains the magnitude.
-
 Overall, the throughput has a linear relationship with the population size.
 This result is a strong indication of the horizontal scalability which we aimed to achieve.
-For more experimental results refer to~\cite[Chapter 5]{checo}???.
+
+The communication complexity of the consensus protocol grows polynomially as the number of facilitators grows linearly.
+Hence, the consensus protocol will take longer to complete and larger fragments must be sent for transaction verification.
+The result is that transactions verification will take longer.
+On the other hand, it does not significantly impact the throughput.
+We refer the reader to~\cite[Chapter 5]{checo} for additional analysis of the effect of the number of facilitators as well as other experimental results.
 

### 2017-08-27 presentation/presentation.tex
Fix paper

@@
 
-\title{Blockchain Consensus Protocol with\\Horizontal Scalability}
+\title{A Blockchain Consensus Protocol With\\Horizontal Scalability}
 % \subtitle{Optional Subtitle}
@@
       Facebook creates ``psychographic'' profiles in political campaigns~\cite{psychographic}
-    \item Baidu's promoted search result on experimental medical care caused death of a student~\cite{weizexi}
+    \item Baidu's promoted search result on medical care caused death of a student~\cite{weizexi}
   \end{itemize}
@@
     \item We require horizontal scalability for ubiquitous use
-    \item More nodes = better performance and throughput
+    \item More nodes = better transaction throughput
   \end{itemize}
@@
 
-\begin{frame}{Related work}
-  State-of-the-art: Sharding
-  \begin{itemize}
-  \item Split state into multiple shards
-  \item Shards run consensus algorithm in parallel
-  \item Difficult to perform atomic inter-shard transactions
-    \begin{itemize}
-      \item Elastico: not possible
-      \item OmniLedger: via Atomic Commit protocol
-    \end{itemize}
-  \item Additional complexity
-  \item Diminishing return
-  \end{itemize}
-\end{frame}
+% \begin{frame}{Related work}
+%   State-of-the-art: Sharding
+%   \begin{itemize}
+%   \item Split state into multiple shards
+%   \item Shards run consensus algorithm in parallel
+%   \item Difficult to perform atomic inter-shard transactions
+%     \begin{itemize}
+%       \item Elastico: not possible
+%       \item OmniLedger: via Atomic Commit protocol
+%     \end{itemize}
+%   \item Additional complexity
+%   \item Diminishing return
+%   \end{itemize}
+% \end{frame}
 

### 2017-08-27 paper/s1-intro.tex
Grammar fixes etc.

@@
 % problem description
-Bitcoin is almost 10 years old and its market capitalisation is over \$60 billion USD and growing~\cite{bitcoinmarketcap}.
+Bitcoin is almost ten years old, and its market capitalisation is over \$60 billion USD and growing~\cite{bitcoinmarketcap}.
 We can be reasonably sure that cryptocurrencies,
@@
 are here to stay in the foreseeable future.
-Driven by the success of Bitcoin, we are seeing a renaissance of consensus research~\cite{miller2016honey, kogias2016enhancing, kokoris2017omniledger},
-where the primary focus is to improve the scalability of blockchain systems.
-This is due to the inefficiencies of the consensus mechanism---proof-of-work (PoW).
+Driven by the success of Bitcoin, we see a renaissance of consensus research~\cite{miller2016honey, kogias2016enhancing, kokoris2017omniledger},
+where the primary focus is to improve the scalability of blockchain systems,
+which is due to the inefficiencies of the consensus mechanism---proof-of-work (PoW).
 For example, Bitcoin can only do 7 transactions per second (TPS) at most~\cite{vukolic2015quest}.
 While adjusting the block size (which Bitcoin has recently done via SegWit~\cite{segwit}) and/or the block interval may increase TPS,
-it also leads to centralisation as larger blocks take longer to propagate in the network,
+it also leads to centralisation as larger blocks take longer to propagate through the network,
 putting miners that do not have a fast network at a disadvantage~\cite{croman2016scaling}.
-Furthermore, due to the bandwidth and latency in today's network,
-it is not possible to achieve more than 27 TPS from simply adjusting the block size~\cite{croman2016scaling}.
+Furthermore, due to the bandwidth and latency of today's network,
+it is not possible to achieve more than 27 TPS from simply adjusting the block size or block interval~\cite{croman2016scaling}.
 
 \textbf{Related work.\quad}
-Many approaches exist for improving the scalability of early blockchains.
+Many approaches exist for improving the scalability of early blockchain systems.
 Off-chain transactions make use of the fact that if nodes make frequent transactions,
 then it is not necessary to store every transaction on the blockchain,
-only the net settlement is necessary.
+only the net settlement is needed.
 The best examples are Lightning Network~\cite{lightningnetwork} and Duplex Micropayment Channels~\cite{decker2015fast}.
@@
 In addition, the user must proactively check whether the counterparty has broadcasted an old channel state so that the user does not lose Bitcoins.
-Finally, creating channels with sufficient balance and also keeping it online to act as a router is expensive.
-A casual user is not capable of such tasks, thus it leads to centralisation.
+Moreover, creating channels with sufficient balance and also keeping it online to act as a router is expensive.
+A casual user is not capable of such tasks, leading to centralisation.
 
 Another way to improve transaction rate is to use traditional Byzantine consensus algorithms such as PBFT~\cite{castro1999practical} in a permissioned ledger such as Hyperledger Fabric~\cite{cachin2016architecture}.
-In essence, they contain a fixed set of nodes (sometimes called validating peer) that run a Byzantine consensus algorithm to decide on new blocks.
-Such systems can achieve much higher transaction rates, e.g. 10,000 TX/s if the number of validating peer is under 16 for PBFT~\cite[Section 5.2]{miller2016honey}.
-However, these systems do not scale, e.g. the transaction rate drops to under 5000 TX/s when the number of validating peer is 64~\cite[Section 5.2]{miller2016honey}.
+In essence, such systems contain a fixed set of nodes (sometimes called validating peer) that run a Byzantine consensus algorithm to decide on new blocks.
+Such systems can achieve much higher transaction rates, e.g., 10,000 TPS if the number of validating peer is under 16 for PBFT~\cite[Section 5.2]{miller2016honey}.
+However, these systems do not scale, e.g., the transaction rate drops to under 5000 TPS when the number of validating peer is 64~\cite[Section 5.2]{miller2016honey}.
 Moreover, the validating peers are predetermined which makes the system unsuitable for the open internet.
@@
 Recent research has developed a class of hybrid systems which uses PoW for committee election,
-and Byzantine consensus algorithms to agree on transactions, e.g. SCP~\cite{luu2015scp}, ByzCoin~\cite{kogias2016enhancing} and Solidus~\cite{abraham2016solidus}.
+and Byzantine consensus algorithms to agree on transactions, e.g., ByzCoin~\cite{kogias2016enhancing} and Solidus~\cite{abraham2016solidus}.
 This design is primarily for permissionless systems because the PoW leader election aspect prevents the Sybil attack~\cite{douceur2002sybil}.
-This technique overcomes the early blockchain scalability issue by delegating the transaction validation to the Byzantine consensus protocol (e.g. PBFT in ByzCoin~\cite{kogias2016enhancing}).
-A major tradeoff of such systems is that they cannot guarantee a high level of fault tolerance when there is a large number of malicious nodes (but less than a majority).
-SCP, ByzCoin and Solidus all have some probability to elect more than $t$ Byzantine nodes into the committee,
+It overcomes the early blockchain scalability issue by delegating the transaction validation to a Byzantine consensus protocol.
+A tradeoff of such systems is that they cannot guarantee a high level of fault tolerance when there is a large number of malicious nodes (but less than a majority).
+ByzCoin and Solidus all have some probability of electing more than $t$ Byzantine nodes into the committee,
 where $t$ is typically just under a third of the committee size (a lower bound of Byzantine consensus~\cite{pease1980reaching}).
-This problem is especially difficult to solve because the committee is always much smaller than the population size which usually has more than $t$ Byzantine nodes.
-Again, due to the fact that these systems must reach consensus on all transactions,
+Again, because these systems must reach consensus on all transactions,
 none of them achieves horizontal scalability.
-That is, by adding new nodes into the network, the global transaction rate should increase proportionally.
 
-Finally, a technique that does achieve horizontal scalability is sharding, e.g. Elastico~\cite{luu2016elastico} and OmniLedger~\cite{kokoris2017omniledger}.
-This is done by grouping nodes into multiple committees or shards of constant size.
-Nodes within a single shard run a Byzantine consensus algorithm to agree on a set of transactions that belong to that specific shard.
-The number of shards grows linearly with respect to the total computational power in the network, hence the transaction rate also grows linearly.
-The biggest limitation of sharding is that it is only optimal if transactions stay in the same shard.
+Finally, a technique that does achieve horizontal scalability is sharding, e.g., Elastico~\cite{luu2016elastico} and OmniLedger~\cite{kokoris2017omniledger}.
+It involves grouping nodes into multiple committees of constant size, also known as shards,
+and nodes within a single shard run a Byzantine consensus algorithm to agree on a set of transactions that belong to that specific shard.
+The number of shards grows linearly with respect to the total computational power of the network;
+hence the transaction rate also grows linearly.
+The limitation of sharding is that it is only optimal if transactions stay in the same shard.
 In fact, Elastico cannot atomically process inter-shard transactions.
 OmniLedger has an inter-shard transaction protocol but choosing a good shard size is difficult.
-A large shard size would make the system less scalable because the Byzantine consensus algorithm must be run by a large number of nodes.
-A small shard size would result in a large number of inter-shard transactions which also hinders scalability.
+A large shard size would degrade the transaction throughput because the Byzantine consensus algorithm must be run by a large number of nodes.
+A small shard size would result in a large number of inter-shard transactions which hinders scalability.
 % The authors of OmniLedger noted that inter-shard transactions have significantly higher latency compared to the consensus protocol~\cite{kokoris2017omniledger}.
@@
 which leads to the goal of this work.
-Concretely, we attempt to answer the following research question.
+Hence the research question which we wish to answer is as follows.
 \begin{displayquote}
-    \emph{How can we design a blockchain consensus protocol that is fault-tolerant,
-    horizontally scalable,
-    and able to reach global consensus?}
+    \emph{How can we design a horizontally scalable blockchain consensus protocol?}
     \end{displayquote}
-We remark the different aspects of our research question.
-A blockchain consensus protocol should be application neutral,
-for example, PoW without the transaction logic.
-The protocol should be Byzantine fault tolerant up to some threshold.
-The threshold may be made adjustable by trading off performance.
-Finally, we are interested in horizontal scalability in the general case, as it enables ubiquitous use.
+Concretely, a blockchain consensus protocol should be application neutral.
+For example, PoW is application neutral because transaction semantics does not affect it,
+i.e. it can be applied in different applications such as cryptocurrency (Bitcoin) and domain name system (Namecoin~\cite{namecoin}).
+Further, we are interested in horizontal scalability in the general case as it enables ubiquitous use.
+That is, adding more nodes to the network should result in higher transaction throughput.
 
 \textbf{Contribution.\quad}
-The key insight is to \emph{not} reach consensus using an existing consensus algorithm (a modified HoneyBadgerBFT~\cite{miller2016honey}) on transactions themselves,
+The key insight is not to reach consensus using an existing consensus algorithm (a modified HoneyBadgerBFT~\cite{miller2016honey}) on transactions themselves,
 but on special blocks called checkpoint blocks,
@@
 \Cref{sec:system} gives the formal system architecture.
-Next, we argue the correctness and fault tolerance properties of our system in~\Cref{sec:analysis}.
-We evaluate our system experimentally in~\Cref{sec:implementation}.
+In~\Cref{sec:tradeoffs}, we discuss a few design variations and their tradeoffs.
+We argue the correctness and fault tolerance properties of our system in~\Cref{sec:analysis}.
+Then we evaluate our system experimentally in~\Cref{sec:implementation}.
 Finally, we conclude our work in~\Cref{sec:conclusion}.

### 2017-08-27 paper/s2-problem.tex
Grammar fixes etc.

@@
     \item \emph{Agreement}:
-        If any correct node decides on the validity (except when it is \emph{unknown}) of a transaction,
-        then all other correct nodes are able to reach the same conclusion or \emph{unknown}.
+        If any correct node decides on the validity of a transaction, except when it is \emph{unknown},
+        then all other correct nodes are able to reach the same conclusion or decide \emph{unknown}.
     \item \emph{Scalability}:
@@
 
-Note that the agreement property is similar, but a relaxed version of what is often seen in Byzantine consensus problems.
+Note that the agreement property is similar, but a relaxed version of what is often seen in a Byzantine consensus problem.
 Namely, the property only holds if honest nodes do not output \emph{unknown}.
@@
 Our problem does not have a termination property, also known as liveness.
-Instead, nodes are incentivised to complete the protocol execution, we describe this phenomenon in~\Cref{sec:correctness-of-validity}.
+Instead, nodes are incentivised to complete the protocol execution;
+we describe this phenomenon in~\Cref{sec:correctness-of-validity}.
 
@@
 i.e. there are no constraints on the semantics of transactions.
-This is so that the protocol can act as a backbone to many applications.
+This formulation is so that the protocol can act as a building block to many applications.
 Due to this fact, we also do not consider global fork prevention or detection,
-as some application may not need such strong guarantees such as the accounting of detailed internet traffic in Tribler~\cite{pimotte, pouwelse2008tribler}.
+as some application may not need such strong guarantees such as the accounting of internet traffic in Tribler~\cite{pimotte, pouwelse2008tribler}.
 On the other hand, we give two alternative constructions that do have fork detection in~\Cref{sec:fork}.
@@
 The adversary has full control of the delivery schedule and the message ordering of all messages.
-But they are not allowed to drop messages except for their own\footnote{Reliability can be achieved in unreliable networks by resending messages or using some error correction code.}.
 

### 2017-08-27 paper/s3-system.tex
Grammar fixes etc.

@@
 The nodes only store transactions (TX) that they are involved in on their hash chain.
-Transactions are stored in TX blocks and every block only contains one transaction.
-A transaction between two nodes should therefore result in two TX blocks on the hash chains of those nodes.
+Transactions are stored in TX blocks, and every block only contains one transaction.
+A transaction between two nodes should, therefore, result in two TX blocks on their respective hash chains.
 We introduce a special block called checkpoint (CP) block,
@@
 \caption{Visualisation of the data structure used in \textsc{Checo}.
-$t_{u, i}$ represents a TX block on $u$'s chain with a sequence number of $i$.
-$c_{v, j}$ represents a CP block on $v$'s chain with a sequence number of $j$.
+$t_{u, i}$ represents a TX block on $u$'s chain with a sequence number $i$.
+$c_{v, j}$ represents a CP block on $v$'s chain with a sequence number $j$.
 The blocks at the ends of the dotted lines are pairs of each other.
-Blocks of sequence number 0 (e.g. $t_{b, 0}$) are genesis blocks.}
+Blocks of sequence number 0 (e.g., $t_{c, 0}$) are genesis blocks.}
 \label{fig:trustchain-good-cp}
@@
 
-\textsc{Checo} consist of three protocols---consensus protocol, transaction protocol and validation protocol---all interacting with the data distributed hash chain structure described above.
+\textsc{Checo} consists of three protocols---consensus protocol, transaction protocol and validation protocol---all interacting with the distributed hash chain data structure described above.
 The primary protocol is the consensus protocol,
@@
 Thus, at the beginning of every Byzantine consensus algorithm execution,
-we randomly elect a set of nodes---called facilitators---to collect CP blocks from every other node and use those blocks as the input to the Byzantine consensus algorithm, and then run it.
+we randomly elect a set of nodes---called facilitators---to collect CP blocks from every other node and use those blocks as the input to the Byzantine consensus algorithm.
 After the algorithm completes,
@@
 When a node wishes to validate one of its transactions, it asks the counterparty for the \emph{agreed fragment} of the transaction.
-Which is a fragment of the counterparty's chain beginning and ending with CP blocks but contains the TX block belonging to that transaction,
-and the CP blocks must be in consensus.
+Which is a section of the counterparty's chain beginning and ending with CP blocks but contains the TX block belonging to that transaction,
+where the CP blocks must be in consensus.
 Upon the counterparty's response,
-the node checks whether the CP blocks are in some consensus result and among other conditions.
+the node checks whether the CP blocks are, in fact, in some consensus result and among other conditions.
 The transaction is valid if these conditions are satisfied.
@@
 
+For brevity, we left out issues such as selecting facilitators in a fair way and the exact validity condition.
+The following sections give the formal description and will answer these issues.
+
 \subsection{\textsc{Checo} data structure}
-Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
+Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a hash chain $B_u$.
 The chain consist of blocks $B_u = \{ b_{u, i} : i \in \{ 0, \dots, h - 1 \} \}$,
@@
 \textbf{\emph{Checkpoint block and genesis block}}
+
 The CP block is a five-tuple, i.e. 
@@
 \textbf{\emph{Consensus result}}
+
 Our consensus protocol runs in rounds,
@@
 
-Here we define an important property which results from the interleaving nature of CP and TX blocks.
+Here we define a property which results from the interleaving nature of CP and TX blocks.
 It is used in our validation protocol (discussed in~\Cref{sec:vd-protocol}).
@@
 \textbf{\emph{Enclosure and agreed enclosure}}
+
 If there exist a tuple $\langle c_{u,a}, c_{u, b} \rangle$ for a TX block $t_{u, i}$, where
@@
 If the enclosure of some TX block $t_{u, i}$ is $\langle c_{u,a}, c_{u, b} \rangle$,
-then its \emph{fragment} $F_{u, i}$ is defined as $\{ b_{u, i} : a \le i \le b \}$.
-Similarly, \emph{agreed fragment} has the same definition as fragment but using agreed enclosure.
-For convenience, the function $\textsf{agreed\_fragment}(\cdot)$ represents the agreed fragment of some TX block if it exists, otherwise $\bot$.
+then its fragment $F_{u, i}$ is defined as $\{ b_{u, i} : a \le i \le b \}$.
+Similarly, agreed fragment has the same definition as fragment but using agreed enclosure.
+For convenience, the function $\textsf{agreed\_fragment}(t_{u, i})$ outputs the agreed fragment of $t_{u, i}$ if it exists, otherwise $\bot$.
 \end{definition}
@@
 \subsection{Consensus Protocol}
-Our scalable consensus protocol $\Pi_c$ uses an unmodified asynchronous subset consensus (ACS) algorithm as the key building block.
+Our scalable consensus protocol $\Pi_c$ uses an unmodified asynchronous common subset (ACS) algorithm as the key building block.
 The objectives of the protocol are to
@@
 \label{sec:consensus-phase}
-The consensus phase begins when $\F_r$ is available to all the nodes.
+
+For any node $u$,
+the consensus phase begins when $\F_r$ is available and the latest block is $c_{u, h-1}$.
 Note that $\F_r$ indicates the facilitators that were elected using results of round $r$ and are responsible for driving the ACS algorithm in round $r + 1$.
@@
 There are two scenarios in the consensus phase.
-First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators.
-Second, if the node is a facilitator, it waits until $N - t$ messages of type \texttt{cp\_msg} are received.
+First, if $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h-1} \rangle$ to all the facilitators.
+Second, if $u$ is a facilitator, it waits until $N - t$ messages of type \texttt{cp\_msg} are received.
 % Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D \gg \Delta$.
@@
 Duplicates and blocks with invalid signatures are again removed from $\C'_{r+1}$ and we call the final result $\C_{r+1}$.
-We have to remove invalid blocks a second time (after ACS) because the adversary may send different CP blocks to different facilitators,
+We have to remove invalid blocks a second time because the adversary may send different CP blocks to different facilitators,
 which results in invalid blocks in the ACS output, but not in any of the inputs.
 
-The core of the consensus phase is the ACS protocol,
-it is based on the construction described in HoneyBadgerBFT~\cite{miller2016honey}.
+The core of the consensus phase is the ACS algorithm,
+which is derived from HoneyBadgerBFT~\cite{miller2016honey}.
 We do not use the full HoneyBadgerBFT due to the following.
@@
 Second, HoneyBadgerBFT uses threshold encryption to hide the content of the transactions.
-But we do not reach consensus on transactions, only CP blocks, so hiding CP blocks is meaningless for us as it contains no transaction information.
+But we do not reach consensus on transactions, only CP blocks, so hiding CP blocks is meaningless for us.
 
-Continuing, when $\F_{r}$ reaches agreement on $\C_{r+1}$,
-they immediately broadcast\footnote{An alternative to using broadcasting is to use an epidemic protocol~\cite{eugster2004epidemic}, 
-as long as it can guarantee delivery.} two messages to all the nodes---
-first the consensus message $\langle \texttt{cons\_msg}, \C_{r+1} \rangle$,
+Continuing, when $\F_{r}$ finish the ACS execution and reach agreement on $\C_{r+1}$,
+they immediately broadcast two messages to all the nodes---first the consensus message $\langle \texttt{cons\_msg}, \C_{r+1} \rangle$,
 and second the signature message $\langle \texttt{cons\_sig}, r, sig \rangle$.
 The reason for sending \texttt{cons\_sig} is the following.
-Recall that channels are not authenticated, 
+The channels are not authenticated, 
 and there are no signatures in $\C_{r + 1}$.
 If a non-facilitator sees some $\C_{r + 1}$, it cannot immediately trust it because it may have been forged.
-Thus, To guarantee authenticity, every facilitator sends an additional message that is the signature of $\C_{r + 1}$.
+Thus, to guarantee authenticity, every facilitator sends an additional message that is the signature of $\C_{r + 1}$.
 
-Upon receiving $\C_{r + 1}$ and at least $n - f$ valid signatures by some node $u$, $u$ performs two tasks.
+Upon receiving $\C_{r + 1}$ and at least $n - t$ valid signatures, $u$ performs two tasks.
 First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1})$ (\Cref{alg:new-cp}).
@@
 so a new round can be started.
+
+Our protocol has some similarities with synchronizers~\cite[Chapter 10]{podc} because it is effectively a technique to introduce synchrony in an asynchronous environment.
+If we consider the facilitators as a collective authority, then it can be seen as a synchronizer that sends pulse messages (in the form of \texttt{cons\_msg} and \texttt{cons\_sig}) to indicate the start of a new clock pulse.
+Every node then sends a completion messages (in the form of \texttt{cp\_msg}) to the new collective authority to indicate that they are ready for the next pulse.
 
@@
 
-The TX protocol $\Pi_t$ is shown in \Cref{alg:tx-proto}.
+The TX protocol $\Pi_t$, shown in \Cref{alg:tx-proto}, is run by all nodes.
 Nodes that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
@@
 \begin{algorithm}
-    \caption{The TX protocol which runs in the context of node $u$.}
+    \caption{$\Pi_t$ runs in the context of node $u$.}
     \label{alg:tx-proto}
@@
 
-A key feature of the TX protocol is that it is non-blocking.
+A key feature of $\Pi_t$ is that it is non-blocking.
 At no time in \Cref{alg:new-tx} or \Cref{alg:tx-proto} do we need to hold the chain state and wait for some message to be delivered before committing a new block to the chain.
-This allows for a high level of concurrency where we can call $\textsf{new\_tx}(\cdot)$ multiple times without waiting for the corresponding \texttt{tx\_resp} messages.
+This allows for a high level of concurrency where we can call $\textsf{new\_tx}(\cdot)$ and send \texttt{tx\_req} messages multiple times without waiting for the corresponding \texttt{tx\_resp} messages.
 
@@
 
-We stress that the \emph{unknown} state means that the verifier does not have enough information to continue with the validation protocol.
+We stress that the \emph{unknown} state means that the verifier does not have enough information to make progress in $\Pi_v$.
 If enough information is available at a later time, then the verifier will output either \emph{valid} or \emph{invalid}.
@@
 \subsubsection{Validation Protocol}
-Our validation protocol,
+Our validation protocol $\Pi_v$,
 shown in~\Cref{alg:vd-proto},
@@
 If $u$ wishes to validate some TX with ID $txid$ and counterparty $v$, it sends $\langle \texttt{vd\_req}, txid \rangle$ to $v$.
-The desired properties of the validation protocol are as follows.
+The desired properties are as follows.
 \begin{definition}
@@
         If any correct node decides on the validity of a transaction, except when it is \emph{unknown},
-        then all other correct nodes are able to reach the same conclusion or output \emph{unknown}.
+        then all other correct nodes are able to reach the same conclusion or decide \emph{unknown}.
     \item \emph{Validity}:
@@
 \begin{algorithm}
-\caption{Validation protocol which runs in the context of $u$}
+\caption{$\Pi_v$ which runs in the context of $u$}
 \label{alg:vd-proto}
@@
 We make two remarks.
-First, just like the TX protocol, we do not block any part of the protocol.
+First, just like $\Pi_t$, we do not block any part of the protocol.
 Second, suppose some $F_{v, j}$ validates $t_{u, i}$, then that does not imply that $t_{u, i}$ only has one pair $t_{v, j}$.
@@
 \label{sec:caching}
-One more way to improve the efficiency of the validation protocol is to use a single agreed fragment to validate multiple transactions.
+One more way to improve the efficiency of $\Pi_v$ is to use a single agreed fragment to validate multiple transactions.
 Concretely, for node $A$, upon receiving an agreed fragment from node $B$,
 rather than validating a single transaction,
-$A$ attempts to validate all transactions performed with $B$ in the unknown state, in that fragment.
+$A$ attempts to validate all transactions performed with $B$, which are in the unknown state but also in that fragment.
 
@@
 \label{sec:fork}
-The validation algorithm guarantees that there are no forks within a single agreed fragment.
-This is sufficient for some applications such as proving the existence of some information.
+The validation algorithm guarantees that there are no forks within a single agreed fragment,
+which is sufficient for some applications such as proving the existence of some information.
 However, for applications such as digital cash where every block depends on one or more previous blocks,
@@
 We offer two approaches to do total fork detection.
-First and the easiest solution is to simply ask for the complete chain of the counterparty.
-The verifier can be sure that there are no forks if the following holds.
+First and the easiest solution is to ask for the complete hash chain of the counterparty.
+The verifier can be sure that there are no forks if the following conditions hold.
 \begin{enumerate}

### 2017-08-27 paper/s4-analysis.tex
Fix paper
Grammar fixes etc.

@@
 We evaluate our system analytically to ensure the desired properties (\Cref{def:consensus} and \Cref{def:validation}) hold.
-An informal argument is given in this section, we refer to~\cite[Chapter 4]{checo}??? for an in-depth analysis.
+An informal argument is given in this section.
+We refer to~\cite[Chapter 4]{checo} for an in-depth analysis.
 
 \subsection{Correctness of the Consensus Protocol}
-Recall that the consensus protocol properties (\Cref{def:consensus}) has
-four items---agreement, validity, termination and fairness.
-The first three hold due to the following facts.
+$\Pi_c$ correctly implements the \textsc{Checo} consensus protocol (\Cref{def:consensus}) due to the following.
+The agreement, validity and termination properties hold because:
 \begin{itemize}
-    \item The CP blocks sent to the facilitators are eventually delivered and then ACS eventually starts.
+    \item The CP blocks sent to the facilitators are eventually delivered, and then ACS eventually starts.
     \item Agreement, validity and termination hold for ACS as they are the properties of ACS and are proven to hold in~\cite{miller2016honey}.
@@
 \end{itemize}
-
 Fairness holds because we model $\textsf{H}(\cdot)$ as a query to the random oracle.
 Sorting nodes by the luck value $\textsf{H}(\C || pk)$ can be seen as a random permutation of a list of nodes.
-Therefore every node has the same probability of becoming a facilitator.
+Therefore every node with a CP block in the consensus result has the same probability of becoming a facilitator.
 
@@
 Using the previous result,
-we show that the agreement and validity properties (from~\Cref{def:validation}) hold for the validation protocol.
+we show that $\Pi_v$ implements the agreement and validity properties of a \textsc{Checo} validation protocol (\Cref{def:validation}).
 
@@
 The agreement property holds because we model $\textsf{H}(\cdot)$ as a query to a random oracle.
-Suppose two honest nodes decided on two different states, \emph{valid} and \emph{invalid} for the same transaction.
+That is, suppose two honest nodes decided on two different states, \emph{valid} and \emph{invalid} for the same transaction.
 For that to happen, two agreed fragments must exist for the same transaction,
@@
 but then never respond to any validation requests.
-Therefore there are transactions that can never be validated.
+Therefore some transactions can never be validated.
 Nevertheless, the malicious node will be at an economic loss if it is not responsive because honest nodes are less likely to make contact with nodes that do not respond to validation requests.
+If the probabilistic fork detection proposal (\Cref{sec:fork}) is used,
+the uncooperative nodes will have more incentive to participate in the protocol.
 
-A stronger version of the validity definition exists.
-That is, if two honest nodes make a transaction,
-then the transaction state is always valid in addition to our current validity definition.
-Under our purely asynchronous model, we cannot guarantee this stronger version.
-Since the adversary can delay any message for any amount of time,
-it can make sure all \texttt{tx\_req} messages are delivered in a round later than the round which the message is sent.
-Effectively, the transaction pair would always be in different rounds and the validation protocol would not output \emph{valid}.
-We believe in a relaxed model, i.e. a weakly synchronous model, a stronger validity definition is possible.
+% A stronger version of the validity definition exists.
+% That is, if two honest nodes make a transaction,
+% then the transaction state is always valid in addition to our current validity definition.
+% Under our purely asynchronous model, we cannot guarantee this stronger version.
+% Since the adversary can delay any message for any amount of time,
+% it can make sure all \texttt{tx\_req} messages are delivered in a round later than the round which the message is sent.
+% Effectively, the transaction pair would always be in different rounds and the validation protocol would not output \emph{valid}.
+% We believe in a relaxed model, i.e. a weakly synchronous model, a stronger validity definition is possible.
 
 \subsection{Fault Tolerance}
-Finally, we consider the effect when the number of adversaries is more than $t$.
-This is useful because in practice it is difficult to guarantee that $t$ satisfies $n \ge 3t + 1$,
+Finally, we consider the effect when the number of adversaries is more than $t$,
+which is useful because in practice it is difficult to guarantee that $t$ satisfies $n \ge 3t + 1$,
 especially when $N$ is large.
@@
 Furthermore, hypergeometric distributions have light tails with ``faster-than-exponential fall-off''~\cite{skala2013hypergeometric}.
-So the probability of picking more than $\lfloor \frac{n-1}{3} \rfloor$ malicious nodes drops of exponentially as the expected value moves away from $\lfloor \frac{n-1}{3} \rfloor$.
+So the probability of picking more than $\lfloor \frac{n-1}{3} \rfloor$ malicious nodes drops off exponentially as the expected value moves away from $\lfloor \frac{n-1}{3} \rfloor$.
 To put this into perspective,
-suppose $n = 1000$ and $\alpha = 1/5$, i.e. a fifth of the nodes are malicious.
-Then the probability to elect more malicious nodes than the threshold is only $2.6 \times 10^{-16}$.
+suppose $n = 100$ and $\alpha = 1/10$, i.e. 10\% of the nodes are malicious,
+then the probability to elect more malicious nodes than the threshold is $9.9 \times 10^{-6}$.
+% To put this into perspective,
+% suppose $n = 1000$ and $\alpha = 1/5$, i.e. a fifth of the nodes are malicious.
+% Then the probability to elect more malicious nodes than the threshold is only $2.6 \times 10^{-16}$.

### 2017-08-27 paper/s5-evaluation.tex
Grammar fixes etc.

@@
 where nodes only make transactions with a fixed node.
-We consider \Cref{fig:global-throughput-random} as the worst case,
+\Cref{fig:global-throughput-random} is the worst case,
 where nodes make transactions with random nodes and the caching mechanism is unlikely to be used.
-Observe that the transaction rate is much lower in~\Cref{fig:global-throughput-random}.
-This is because the communication of an agreed fragment is necessary to verify every transaction (no caching),
+Observe that the transaction rate is much lower in~\Cref{fig:global-throughput-random},
+which is because the communication of an agreed fragment is necessary to verify every transaction (no caching),
 putting a strain on our network infrastructure.
-In practice, we do not expect such behaviour to occur as it is possible to cache agreed fragments.
+% In practice, we do not expect such behaviour to occur as it is possible to cache agreed fragments.
 
@@
 Recall that we fixed the transaction rate to 2 TPS,
-but how is it possible to have around 4800 transactions per second for 1200 nodes (which is 4 TX/s)?
-This is due to the way validated transactions are calculated.
+but how is it possible to have around 4800 transactions per second for 1200 nodes (which is 4 TPS)?
+It is due to the way validated transactions are calculated.
 Transactions are between two parties, hence if every node makes two transactions per second,
@@
 
-The communication complexity of the consensus protocol grows polynomially as the number of facilitators grows linearly.
-Hence, the consensus protocol will take longer to complete and larger fragments must be sent for transaction verification.
-The result is that transactions verification will take longer.
-On the other hand, it does not significantly impact the throughput.
+The downside of our design is that the communication complexity of the consensus protocol grows polynomially as the number of facilitators grows linearly.
+Hence, the consensus protocol will take longer to complete, and larger fragments must be sent for transaction verification.
+On the other hand, it does not significantly impact the throughput;
+only the transaction verification delay is affected.
 We refer the reader to~\cite[Chapter 5]{checo} for additional analysis of the effect of the number of facilitators as well as other experimental results.

### 2017-08-27 paper/s6-conclusion.tex
Grammar fixes etc.

@@
 
-We introduced an application neutral blockchain system which we call \textsc{Checo}.
-The checkpoint block is added to the existing TrustChain data structure for use in our consensus protocol.
-The round based consensus protocol uses ACS as a building block to reach consensus on checkpoint blocks.
+In this work, we described \textsc{Checo}, an application neutral blockchain system with horizontal scalability.
+Our novel data structure allows nodes to efficiently store transactions and record state using CP blocks.
+The round based consensus protocol uses ACS as a building block to reach consensus on CP blocks.
 The consensus result lets nodes elect new facilitators and create new checkpoint blocks.
-To make transactions, nodes use the transaction protocol,
-which is adapted from an existing work called True Halves.
+To make transactions, nodes use the simple and non-blocking transaction protocol.
 Finally, we introduce a validation protocol which ensures that if an agreed fragment for some transaction exists,
 then nodes reach agreement on the validity of that transaction.
-The novelty of the validation protocol is that it uses point-to-point communication,
-i.e. nodes only validate the transactions of interest,
-this enables our horizontal scalability property.
+The novelty of \textsc{Checo} is that it decouples consensus and transaction validation,
+which enables the desirable horizontal scalability property.
 
-We achieve the properties described in the Problem Description (\Cref{sec:description}).
-Namely, our protocol achieves agreement and validity as we argued in~\Cref{sec:correctness-of-validity}.
+We achieve the properties described in~\Cref{sec:description}.
+Namely, our protocol achieves agreement on transactions as we argued in~\Cref{sec:correctness-of-validity}.
 Further, the horizontal scalability is demonstrated in~\Cref{sec:implementation},

### 2017-08-29 presentation/presentation.tex
Simplify presentation

@@
     \item But it puts central authorities in control
-    \item Many are motivated by profit
+    \item Most are motivated exclusively by profit
     \item Not always in the interest of the ``users''\footnote{Typically users of some free service X are, in fact, used by X.}
@@
   \begin{itemize}
-    \item Facebook can predict your opinions and desires better than your spouse~\cite{youyou2015computer}
+    \item Facebook can predict your opinions and desires better than your spouse from your ``liked'' posts~\cite{youyou2015computer}
     \item With intimate knowledge of the individuals,
@@
   \begin{itemize}
-    \item Blockchains are distributed ledgers
-    \item They enable large scale consensus
+    \item Blockchains are distributed (replicated) ledgers
+    \item They enable internet-scale consensus
     \item An alternative to central authorities for the first time
-    \item Some applications include:
+    \item Some initial applications include:
     \begin{itemize}
@@
       \item Domain name system (e.g., Namecoin)
-      \item File sharing (e.g., Filecoin)
+      \item Bandwidth rental (e.g., Filecoin)
       \item General purpose (e.g., Ethereum)
@@
     \item We require horizontal scalability for ubiquitous use
-    \item More nodes = better transaction throughput
+    \item More users = more transactions per second globally
   \end{itemize}
@@
     \centering
-    \caption{Summary of the scalability properties of many blockchain systems.
-    Scalability gets better from left to right.}
+    \caption{Summary of the scalability properties of many blockchain systems.}
     \label{tab:scalability}
@@
 \end{table}
+\begin{center}
+\Large{
+$\xrightarrow{\qquad\text{scalability}\qquad}$
+}
+\end{center}
 \end{frame}
@@
   \end{block}
-    \note{
   \begin{itemize}
-      \item Blockchain consensus protocol---independent of the application, like PoW (not PoS)
-      \item Byzantine fault tolerance---if some number of nodes are malicious then the system should continue to function
-      \item Horizontal scalability---if every node make transactions at the same rate, then the global throughput should increase
-      \item Global consensus---all node should agree on a global state, useful for some applications, e.g. digital cash
+    \item Blockchain consensus protocol---application neutral, e.g., PoW
+    \item Fault-tolerant---tolerate up to some number of malicious nodes
+    \item Horizontal scalability---more nodes in the network leads to higher transaction throughput
+    \item Global consensus---all node should agree on a global state
   \end{itemize}
-    }
 \end{frame}
@@
 \section{System architecture}
-\begin{frame}{Intuition}
+\begin{frame}{Intuition and idea explored in this thesis}
   \begin{itemize}
-    \item Traditionally: to reach consensus and check the validity of all transactions is expensive
+    \item It is expensive to reach consensus on all transactions
     \item Our idea: we decouple consensus and validation
-    \item Use a single digest to represent an arbitrarily large number of transactions
+    \item A single digest represents an arbitrarily large number of transactions
     \item Reach consensus on the small digest
@@
   \end{itemize}
+  \begin{figure}[h]
+  \includegraphics[width=0.7\textwidth]{idea}
+  \centering
+  \end{figure}
 \end{frame}
@@
 \subsection{System model}
-\begin{frame}{\subsecname}
-  \begin{itemize}
-    \item $N$ is the population size
-    \item Purely asynchronous channels with eventual delivery
-    \item $n$ nodes are facilitators
-    \item $t$ nodes are malicious, i.e. Byzantine
-    \item $n \ge 3t + 1$
-    \item $N \ge n + t$
-  \end{itemize}
-  \note{At this point, explain facilitators are nodes that have some extra task to do, temporarily.
-  Explain that $N$ is typically a lot larger than $n$ instead of explaining $N \ge n + t$.}
-\end{frame}
-
 \subsection{Architecture overview}
@@
       \item In this example there are three nodes
+      \item Each node maintains their personal hash chain and genesis block
       \item Squares are TX blocks and circles are CP blocks
@@
 
-\begin{frame}{\subsecname:~TX block}
-  \begin{enumerate}
-    \item Hash pointer to the previous block
-    \item Sequence number
-    \item Transaction ID
-    \item Public key of the counterparty
-    \item Transaction message $m$
-    \item Signature the five items above
-  \end{enumerate}
-  \vfill
-  A transaction is represented by a \emph{pair} of TX blocks
+\begin{frame}{\subsecname:~Transaction (TX) block}
+\begin{itemize}
+  \item Goal: record transactions
+  \item A transaction is represented by a pair of TX blocks,
+  i.e. a contract signed by both parties
+\end{itemize}
+\begin{figure}[h]
+  \includegraphics[trim={7cm 8.2cm 2cm 2.55cm},clip,width=0.7\textwidth]{trustchain-good-cp}
+  \centering
+\end{figure}
 \end{frame}
 
-\begin{frame}{\subsecname:~CP block}
-  \begin{enumerate}
-    \item Hash pointer to the previous block
-    \item Sequence number
-    \item Digest of consensus result, i.e.~a set of CP blocks
-    \item Round number $r$
-    \item Signature on the four items above
-  \end{enumerate}
+\begin{frame}{\subsecname:~Checkpoint (CP) block}
+  \begin{itemize}
+    \item Goal: represent the state of the chain using a single digest
+    \item A collection of CP blocks from all the nodes represent the state of the system
+    \item Nodes agree on the system state using our consensus protocol
+  \end{itemize}
+\begin{figure}[h]
+  \includegraphics[trim={7cm 2.505cm 2cm 8.55cm},clip,width=0.7\textwidth]{trustchain-good-cp}
+  \centering
+\end{figure}
 \end{frame}
@@
 
-\begin{frame}{\subsecname---Background on ACS}
+\begin{frame}{\subsecname}
   \begin{itemize}
-    \item Asynchronous common subset
-    \item A simplification of HoneyBadgerBFT~\cite{miller2016honey}
-    \item $n$ nodes
-    \item $t$ nodes may be malicious
-    \item Input: every node proposes a set of values, e.g., $\{A, B\}, \{B, C\}, \dots$
-    \item Output: set union of the majority, e.g., $\{A, B, C, \dots \}$
+    \item Goal: reach consensus on a collection of CP blocks amongst all the nodes
+    \item Uses an existing fault-tolerant consensus algorithm (HoneyBadgerBFT~\cite{miller2016honey}) as the building block
+    \item But it cannot be used in a large network due to high communication complexity
+    \item We overcome this limitation by selecting a small number of \emph{facilitators} to run HoneyBadgerBFT
   \end{itemize}
@@
 \begin{frame}{\subsecname}
-  \begin{figure}[h]
-  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-1}
-  \centering
-  \end{figure}
-  \note{Suppose we are in a state where $\C_{r - 1}$ has just been agreed by some facilitators but not yet propagated.}
-\end{frame}
-
-\begin{frame}{\subsecname}
-  \begin{figure}[h]
-  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-2}
-  \centering
-  \end{figure}
-  \note{Nodes receive consensus result $\C_{r - 1}$,
-    first $n$ nodes ordered by $\textsf{H}(\C_{r-1} || pk)$ become $\F_{r-1}$,
-    send the new CP blocks to $\F_{r-1}$.}
-\end{frame}
-
-\begin{frame}{\subsecname}
-  \begin{figure}[h]
-  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-3}
+  \begin{figure}
+    \includegraphics[width=0.65\textwidth]{consensus-overview}
     \centering
@@
 
-\begin{frame}{\subsecname}
-  \begin{figure}[h]
-  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-4}
-  \centering
-  \end{figure}
-\end{frame}
-
-\begin{frame}{\subsecname}
-  \begin{figure}[h]
-  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-5}
-  \centering
-  \end{figure}
-  \note{Transactions carry on as usual in round $r$,
-  while facilitators are trying to reach consensus on the new CP blocks concurrently.}
-\end{frame}
-
-\begin{frame}{\subsecname}
-  \begin{figure}[h]
-  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-6}
-  \centering
-  \end{figure}
-  \note{$\F_{r-1}$ agree and disseminate $\C_r$,
-  CP blocks at round $r-1$ ($c_{a, 6}, c_{b, 4}, c_{c,8}$) should be in $\C_r$.}
-\end{frame}
-
 \begin{frame}{\subsecname:~properties}
 \label{def:consensus}
-The consensus protocol has the following properties in every round $r$.
 \begin{itemize}
@@
 \begin{itemize}
-\item Request and response protocol
-\item Two TX blocks containing the same $txid$ are generated
-\item Non-blocking
+\item Two TX blocks are generated on the chains of Alice and Bob
+\item No guarantee that nodes follow this protocol
 \end{itemize}
@@
 \begin{itemize}
-\item Alice needs the two fragments that belong to the transaction
+\item To check that the transaction protocol is correctly followed
+\item Alice needs the fragments of the TX on Bob's hash chain 
 \item Validation function checks whether the fragments are OK and contain the transaction
-\item Non-blocking
 \end{itemize}
@@
 \begin{itemize}
+  \item Every node may run the validation protocol on a transaction
+  \item We reach consensus on transactions via CP blocks
   \item CP blocks of the fragments are ``anchored'' due to the consensus protocol
   \item It is difficult to modify the fragment once ``anchored''
-  \item Implicit consensus on transactions via CP blocks
+  \item Since transaction protocol and validation protocol only use point-to-point communication,
+  we achieve horizontal scalability.
 \end{itemize}
@@
 \begin{frame}{\secname}
-  Out work answers the research question.
+  Our work answers the research question.
   \begin{block}{}
-    How can we design a \emph{blockchain consensus protocol} that is \emph{fault tolerant},
+    How can we design a \emph{blockchain consensus protocol} that is \emph{fault-tolerant},
     \emph{horizontally-scalable}, and able to reach \emph{global consensus?}
   \end{block}
+  \begin{itemize}
+    \item Fault-tolerance is achieved using HoneyBadgerBFT
+    \item Horizontal-scalability is achieved by untangling consensus and validation
+    \item Global-consensus on transactions is achieved via consensus on CP blocks
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Bibliography}
+\printbibliography
+\end{frame}
+
+\begin{frame}{Thank you}
+  \begin{center}
+    \Large{
+  Any questions?
+    }
+  \end{center}
+\end{frame}
+
+\section*{Extras}
+\begin{frame}{TX block}
+  \begin{enumerate}
+    \item Hash pointer to the previous block
+    \item Sequence number
+    \item Transaction ID
+    \item Public key of the counterparty
+    \item Transaction message $m$
+    \item Signature the five items above
+  \end{enumerate}
+  \vfill
+  A transaction is represented by a \emph{pair} of TX blocks
+\end{frame}
+
+\begin{frame}{CP block}
+  \begin{enumerate}
+    \item Hash pointer to the previous block
+    \item Sequence number
+    \item Digest of consensus result, i.e.~a set of CP blocks
+    \item Round number $r$
+    \item Signature on the four items above
+  \end{enumerate}
+\end{frame}
+
+\begin{frame}{Background on ACS}
+  \begin{itemize}
+    \item Asynchronous common subset
+    \item A simplification of HoneyBadgerBFT~\cite{miller2016honey}
+    \item $n$ nodes
+    \item $t$ nodes may be malicious
+    \item Input: every node proposes a set of values, e.g., $\{A, B\}, \{B, C\}, \dots$
+    \item Output: set union of the majority, e.g., $\{A, B, C, \dots \}$
+  \end{itemize}
+\end{frame}
+
+\subsection*{Consensus protocol}
+\begin{frame}{\subsecname:~part 1}
+  \begin{figure}[h]
+  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-1}
+  \centering
+  \end{figure}
+  \note{Suppose we are in a state where $\C_{r - 1}$ has just been agreed by some facilitators but not yet propagated.}
+\end{frame}
+
+\begin{frame}{\subsecname:~part 2}
+  \begin{figure}[h]
+  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-2}
+  \centering
+  \end{figure}
+  \note{Nodes receive consensus result $\C_{r - 1}$,
+    first $n$ nodes ordered by $\textsf{H}(\C_{r-1} || pk)$ become $\F_{r-1}$,
+    send the new CP blocks to $\F_{r-1}$.}
+\end{frame}
+
+\begin{frame}{\subsecname:~part 3}
+  \begin{figure}[h]
+  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-3}
+  \centering
+  \end{figure}
+\end{frame}
+
+\begin{frame}{\subsecname:~part 4}
+  \begin{figure}[h]
+  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-4}
+  \centering
+  \end{figure}
+\end{frame}
+
+\begin{frame}{\subsecname:~part 5}
+  \begin{figure}[h]
+  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-5}
+  \centering
+  \end{figure}
+  \note{Transactions carry on as usual in round $r$,
+  while facilitators are trying to reach consensus on the new CP blocks concurrently.}
+\end{frame}
+
+\begin{frame}{\subsecname:~part 6}
+  \begin{figure}[h]
+  \includegraphics[trim={2cm 1.5cm 2cm 1.5cm},clip,width=0.75\textwidth]{trustchain-6}
+  \centering
+  \end{figure}
+  \note{$\F_{r-1}$ agree and disseminate $\C_r$,
+  CP blocks at round $r-1$ ($c_{a, 6}, c_{b, 4}, c_{c,8}$) should be in $\C_r$.}
 \end{frame}
@@
 
-\begin{frame}{Bibliography}
-\printbibliography
-\end{frame}
-
-\section*{Extras}
 \end{document}

### 2017-08-29 presentation/presentation.tex
Edit presentation

@@
   \begin{itemize}
-    \item Facebook can predict your opinions and desires better than your spouse from your ``liked'' posts~\cite{youyou2015computer}
-    \item With intimate knowledge of the individuals,
-      Facebook creates ``psychographic'' profiles in political campaigns~\cite{psychographic}
+    \item Facebook can predict your opinions and desires from your ``likes''~\cite{youyou2015computer},
+      and then create a ``psychographic'' profile for use in political campaigns~\cite{psychographic}
     \item Baidu's promoted search result on medical care caused death of a student~\cite{weizexi}
@@
     \begin{itemize}
-      \item The student had cancer---(synovial sarcoma)
       \item The treatment was phased out at clinical stage in the US because the success rate is too low
@@
       predict your opinions and desires better than your husband or wife!'' -Harari, \emph{Homo Deus}
+      \item The student had cancer---(synovial sarcoma)
       \item Summary: we live in an age where companies and governments are able to control our thoughts,
@@
   \begin{itemize}
-    \item Blockchains are distributed (replicated) ledgers
-    \item They enable internet-scale consensus
-    \item An alternative to central authorities for the first time
+    \item Blockchains are distributed (replicated) ledgers with no central control\footnote{In practice, some central control exist.}
+    \item They enable internet-scale consensus for the first time
     \item Some initial applications include:
@@
       \item Domain name system (e.g., Namecoin)
-      \item Bandwidth rental (e.g., Filecoin)
+      \item Storage rental (e.g., Filecoin)
       \item General purpose (e.g., Ethereum)
@@
   \begin{itemize}
-    \item Consensus algorithm of early blockchain systems do not scale
-    \item Bitcoin (PoW) is limited to 7 transactions per second
+    \item All blockchain systems have a consensus algorithm
+    \item Early consensus algorithms (PoW) do not scale
+    \item Bitcoin is limited to 7 transactions per second
     \item 100,000 transaction backlog in May 2017
@@
     Bitcoin  & Hyperledger & Elastico          & \textsc{Checo} (this work) \\
-    Ethereum & ByzCoin     & OmniLedger        &                \\
-    etc.     & Solidius    & Lightning Network &                \\
+    Litecoin & ByzCoin     & OmniLedger        &                \\
     \hline
@@
   \begin{itemize}
-    \item Goal: reach consensus on a collection of CP blocks amongst all the nodes
+    \item Goal 1: reach consensus on a collection of CP blocks amongst all the nodes
+    \item Goal 2: create CP blocks at the end of the protocol
     \item Uses an existing fault-tolerant consensus algorithm (HoneyBadgerBFT~\cite{miller2016honey}) as the building block
     \item But it cannot be used in a large network due to high communication complexity
-    \item We overcome this limitation by selecting a small number of \emph{facilitators} to run HoneyBadgerBFT
+    \item We overcome this limitation by selecting a small number of \emph{facilitators} from the network to run HoneyBadgerBFT
   \end{itemize}
@@
   \begin{figure}
-    \includegraphics[width=0.65\textwidth]{consensus-overview}
+    \includegraphics[width=1.0\textwidth]{consensus-overview}
     \centering
@@
 \item To check that the transaction protocol is correctly followed
-\item Alice needs the fragments of the TX on Bob's hash chain 
-\item Validation function checks whether the fragments are OK and contain the transaction
+\item Alice needs the fragment of the TX on Bob's hash chain 
+\item Validation function checks whether the fragment is OK and contain the transaction
+\item Can be generalised---any node may run the validation protocol on any transaction (does not need to be their own)
 \end{itemize}
@@
 \begin{itemize}
-  \item Every node may run the validation protocol on a transaction
-  \item We reach consensus on transactions via CP blocks
+  \item Result: we reach consensus on transactions via CP blocks
   \item CP blocks of the fragments are ``anchored'' due to the consensus protocol
   \item It is difficult to modify the fragment once ``anchored''
-  \item Since transaction protocol and validation protocol only use point-to-point communication,
+  \item Since the transaction protocol and the validation protocol only use point-to-point communication,
   we achieve horizontal scalability.
@@
     \item Experiment on the DAS-5\footnote{\url{http://www.cs.vu.nl/das5/}}
+    \item Up to 1500 nodes
   \end{itemize}
@@
 
-\begin{frame}{Throughput vs population size (random neighbour)}
+\begin{frame}{Validated transaction throughput (random neighbour)}
   \begin{figure}[h]
@@
 
-\begin{frame}{Throughput vs population size (fixed neighbour)}
+\begin{frame}{Validated transaction throughput (fixed neighbour)}
   \begin{figure}[h]
@@
 
-\begin{frame}{Throughput vs population size (fixed neighbour)}
+\begin{frame}{Stress test (fixed neighbour)}
   \begin{figure}[h]

### 2017-08-30 paper/s1-intro.tex
Edit paper before defence

@@
 % problem description
-Bitcoin is almost ten years old, and its market capitalisation is over \$60 billion USD and growing~\cite{bitcoinmarketcap}.
-We can be reasonably sure that cryptocurrencies,
+The first blockchain system---Bitcoin---is almost ten years old. 
+Its market capitalisation is over \$60 billion USD and growing~\cite{bitcoinmarketcap}.
+We can be reasonably sure that such systems,
 even if their application is still somewhat limited,
 are here to stay in the foreseeable future.
-Driven by the success of Bitcoin, we see a renaissance of consensus research~\cite{miller2016honey, kogias2016enhancing, kokoris2017omniledger},
+Driven by the success of Bitcoin, we see a renaissance of consensus research~\cite{luu2016elastico, kogias2016enhancing, kokoris2017omniledger},
 where the primary focus is to improve the scalability of blockchain systems,
@@
 A casual user is not capable of such tasks, leading to centralisation.
+% TODO other side-chain schemes? polkadot? perum? dfinity?
 
@@
 In fact, Elastico cannot atomically process inter-shard transactions.
-OmniLedger has an inter-shard transaction protocol but choosing a good shard size is difficult.
-A large shard size would degrade the transaction throughput because the Byzantine consensus algorithm must be run by a large number of nodes.
-A small shard size would result in a large number of inter-shard transactions which hinders scalability.
+OmniLedger has an inter-shard transaction protocol but choosing a shard size that matches the transaction characteristics of the network is difficult.
+An inadequate shard size would result in a large number of inter-shard transactions which would hinder scalability.
 % The authors of OmniLedger noted that inter-shard transactions have significantly higher latency compared to the consensus protocol~\cite{kokoris2017omniledger}.

### 2017-08-30 paper/s2-problem.tex
Edit paper before defence

@@
         then all other correct nodes are able to reach the same conclusion or decide \emph{unknown}.
+    \item \emph{Validity}:
+        If a transaction is valid, then it must have been created by two honest nodes.
     \item \emph{Scalability}:
@@
 Our problem does not have a termination property, also known as liveness.
-Instead, nodes are incentivised to complete the protocol execution;
+Instead, nodes are incentivised to complete the protocol execution otherwise they risk economical loss;
 we describe this phenomenon in~\Cref{sec:correctness-of-validity}.
@@
 On the other hand, we give two alternative constructions that do have fork detection in~\Cref{sec:fork}.
+% TODO our transactions are rather different from a typical cryptocurrency transaction,
+% should we still call it transactions or something else?
 

### 2017-08-30 paper/s3-system.tex
Edit paper before defence

@@
 Instead, we introduce an alternative architecture where every node has their own genesis block and hash chain.
-The nodes only store transactions (TX) that they are involved in on their hash chain.
+The nodes only store transactions (TX) that they are involved in on their hash chains.
 Transactions are stored in TX blocks, and every block only contains one transaction.
@@
 \begin{definition}
-\textbf{\emph{Transaction block}}
+\textbf{Transaction block}
 
@@
 $$t_{u, i} = \langle \textsf{H}(b_{u, i - 1}), seq_u, txid, pk_v, m, sig_u \rangle$$
-there exist one and only one \emph{pair} 
+there exists one and only one \emph{pair} 
 $$t_{v, j} = \langle \textsf{H}(b_{v, j - 1}), seq_v, txid, pk_u, m, sig_v \rangle,$$
@@
 \begin{definition}
-\textbf{\emph{Checkpoint block and genesis block}}
+\textbf{Checkpoint block and genesis block}
 
@@
 The genesis block in the chain must be a CP block in the form of
-$$c_{u, 0} = \langle \textsf{H}(\bot), 0,  \textsf{H}(\bot), 0, sig_u \rangle,$$
-where $\textsf{H}(\bot)$ can be interpreted as applying the hash function on an empty string.
+$$c_{u, 0} = \langle \textsf{H}(\epsilon), 0,  \textsf{H}(\epsilon), 0, sig_u \rangle,$$
+where $\epsilon$ is the empty string.
 The genesis block is unique because every node has a unique public and private key pair.
@@
 \label{sec:consensus-result}
-\textbf{\emph{Consensus result}}
+\textbf{Consensus result}
 
@@
 
-Here we define a property which results from the interleaving nature of CP and TX blocks.
+Next we define a property which results from the interleaving nature of CP and TX blocks.
 It is used in our validation protocol (discussed in~\Cref{sec:vd-protocol}).
@@
 \begin{definition}
-\textbf{\emph{Enclosure and agreed enclosure}}
+\textbf{Enclosure and agreed enclosure}
 
-If there exist a tuple $\langle c_{u,a}, c_{u, b} \rangle$ for a TX block $t_{u, i}$, where
+If there exists a tuple $\langle c_{u,a}, c_{u, b} \rangle$ for a TX block $t_{u, i}$, where
 \begin{itemize}
@@
 \begin{definition}
-\textbf{\emph{Fragment and agreed fragment}}
+\textbf{Fragment and agreed fragment}
 
@@
 \label{def:consensus}
-\textbf{\emph{\textsc{Checo} consensus protocol}}
+\textbf{\textsc{Checo} consensus protocol}
 
@@
 Invalid messages are removed.
-That is blocks with invalid signatures and blocks signed by the same key.
+Those are blocks with invalid signatures and blocks signed by the same key.
 With the sufficient number of \texttt{cp\_msg} messages,
@@
 \label{def:validation}
-\textbf{\emph{\textsc{Checo} validation protocol}}
+\textbf{\textsc{Checo} validation protocol}
 
@@
 
-\subsection{Using Timing Assumption in the Permissionless Setting}
+\subsection{Open membership using timing assumption}
 \label{sec:permissionless}
 At the start of our consensus phase (\Cref{sec:consensus-phase}), facilitators must wait for $N-f$ \texttt{cp\_msg} messages.
-The use of $N$ makes our system unsuitable for the permissionless setting.
-To introduce timing, instead of waiting for $N-f$ messages, we wait for some time $D$,
+The use of $N$ makes our system unsuitable for the open membership setting,
+where nodes may join and leave at will (churn).
+We over come this problem by introducing a timing assumption.
+Concretely, instead of waiting for $N-f$ messages, we wait for some time $D$,
 such that $D$ is sufficiently long for honest nodes to send their CP blocks to the facilitators.
@@
 which is sufficient for some applications such as proving the existence of some information.
-However, for applications such as digital cash where every block depends on one or more previous blocks,
+However, for applications such as cryptocurrency where every block depends on one or more previous blocks,
 our scheme is not suitable.

### 2017-08-30 presentation/presentation.tex
Edit presentation

@@
 \mode<presentation>{\usetheme{tud}}
+
+% https://tex.stackexchange.com/questions/124256/how-do-i-get-numbered-entries-in-a-beamer-bibliography
+\setbeamertemplate{bibliography item}{\insertbiblabel}
 
@@
     \begin{itemize}
-      \item Digital cash (e.g., Bitcoin)
+      \item Digital cash (e.g., Bitcoin, Litecoin)
       \item Domain name system (e.g., Namecoin)
@@
     \item We require horizontal scalability for ubiquitous use
-    \item More users = more transactions per second globally
+    \item More users $\rightarrow$ more transactions per second globally
   \end{itemize}
@@
     \item Blockchain consensus protocol---application neutral, e.g., PoW
-    \item Fault-tolerant---tolerate up to some number of malicious nodes
+    \item Fault-tolerant---tolerate a number of malicious nodes
     \item Horizontal scalability---more nodes in the network leads to higher transaction throughput
@@
   \begin{itemize}
-    \item It is expensive to reach consensus on all transactions
+    \item It is expensive to verify and reach consensus on all transactions
     \item Our idea: we decouple consensus and validation
@@
   \begin{center}
-  \Large{The four components of \textsc{Checo}}
+  {\Large The four components of \textsc{Checo}}
   \begin{figure}[h]
@@
     \item A collection of CP blocks from all the nodes represent the state of the system
-    \item Nodes agree on the system state using our consensus protocol
+    \item Nodes become aware of the system state by running our consensus protocol
   \end{itemize}
@@
     \item Goal 1: reach consensus on a collection of CP blocks amongst all the nodes
-    \item Goal 2: create CP blocks at the end of the protocol
+    \item Goal 2: create new CP blocks at the end of the protocol
     \item Uses an existing fault-tolerant consensus algorithm (HoneyBadgerBFT~\cite{miller2016honey}) as the building block
@@
     \item \emph{Validity}:
-      The consensus results is valid such that a new set of facilitators can be computed from it.
+      The consensus result is valid such that a new set of facilitators can be computed from it.
     \item \emph{Fairness}:
@@
 \begin{frame}{\subsecname:~properties}
+\emph{Consensus on CP blocks $\rightarrow$ consensus on transactions}
+\bigskip
 \begin{itemize}
-  \item Result: we reach consensus on transactions via CP blocks
   \item CP blocks of the fragments are ``anchored'' due to the consensus protocol
@@
 
-\begin{frame}{Validated transaction throughput (random neighbour)}
+\begin{frame}{Validated transaction throughput (random node)}
   \begin{figure}[h]
@@
     \item Fault-tolerance is achieved using HoneyBadgerBFT
-    \item Horizontal-scalability is achieved by untangling consensus and validation
+    \item Horizontal-scalability is achieved by separating consensus and validation, demonstrated experimentally
     \item Global-consensus on transactions is achieved via consensus on CP blocks
@@
   \begin{center}
-    \Large{
-  Any questions?
-    }
+  {\Large Any questions?}
   \end{center}
@@
 \section*{Extras}
-\begin{frame}{TX block}
+\begin{frame}[noframenumbering]{TX block}
   \begin{enumerate}
@@
 
-\begin{frame}{CP block}
+\begin{frame}[noframenumbering]{CP block}
   \begin{enumerate}
@@
 
-\begin{frame}{Background on ACS}
+\begin{frame}[noframenumbering]{Background on ACS}
   \begin{itemize}
@@
 \subsection*{Consensus protocol}
-\begin{frame}{\subsecname:~part 1}
+\begin{frame}[noframenumbering]{\subsecname:~part 1}
   \begin{figure}[h]
@@
 
-\begin{frame}{\subsecname:~part 2}
+\begin{frame}[noframenumbering]{\subsecname:~part 2}
   \begin{figure}[h]
@@
 
-\begin{frame}{\subsecname:~part 3}
+\begin{frame}[noframenumbering]{\subsecname:~part 3}
   \begin{figure}[h]
@@
 
-\begin{frame}{\subsecname:~part 4}
+\begin{frame}[noframenumbering]{\subsecname:~part 4}
   \begin{figure}[h]
@@
 
-\begin{frame}{\subsecname:~part 5}
+\begin{frame}[noframenumbering]{\subsecname:~part 5}
   \begin{figure}[h]
@@
 
-\begin{frame}{\subsecname:~part 6}
+\begin{frame}[noframenumbering]{\subsecname:~part 6}
   \begin{figure}[h]
@@
 
-\begin{frame}{Future work}
+\begin{frame}[noframenumbering]{Effect of the number of facilitators (fixed neighbours)}
+  \begin{figure}
+    \includegraphics[width=0.9\textwidth]{neighbour-fixed/round-duration-vs-facilitators}
+    \centering
+  \end{figure}
+\end{frame}
+
+\begin{frame}[noframenumbering]{Effect of the number of facilitators (random neighbours)}
+  \begin{figure}
+    \includegraphics[width=0.9\textwidth]{neighbour-random/round-duration-vs-facilitators}
+    \centering
+  \end{figure}
+\end{frame}
+
+\begin{frame}[noframenumbering]{Future work}
 \begin{itemize}

### 2017-10-31 paper/s1-intro.tex
Address unfair facilitator election

@@
 It promises significant scalability improvements, but complicates user experience and leads to centralisation.
-Each node must deposit a suitable amount of Bitcoins into a multi-signature account.
+That is, each node must deposit a suitable amount of Bitcoins into a multi-signature account.
 A low deposit would not allow large transactions.
-A high deposit locks the user from using much of her Bitcoins outside the channel.
+A high deposit locks the user from using much of their Bitcoins outside the channel.
 In addition, the user must proactively check whether the counterparty has broadcasted an old channel state so that the user does not lose Bitcoins.

### 2017-10-31 paper/s2-problem.tex
Address unfair facilitator election

@@
 Namely, the property only holds if honest nodes do not output \emph{unknown}.
-For example, for a transaction, it is OK if two honest nodes output \emph{valid} and \emph{unknown}, but it is not OK to output \emph{valid} and \emph{invalid}.
-Our problem does not have a termination property, also known as liveness.
+For example, for a transaction, it is fine if two honest nodes output \emph{valid} and \emph{unknown}, but it is not fine to output \emph{valid} and \emph{invalid}.
+Our problem does not have a termination property.
 Instead, nodes are incentivised to complete the protocol execution otherwise they risk economical loss;
@@
 as some application may not need such strong guarantees such as the accounting of internet traffic in Tribler~\cite{pimotte, pouwelse2008tribler}.
-On the other hand, we give two alternative constructions that do have fork detection in~\Cref{sec:fork}.
+On the other hand, we give two alternative constructions that do perform fork detection in~\Cref{sec:fork}.
 % TODO our transactions are rather different from a typical cryptocurrency transaction,

### 2017-10-31 paper/s3-system.tex
Address unfair facilitator election

@@
 The blocks at the ends of the dotted lines are pairs of each other.
-Blocks of sequence number 0 (e.g., $t_{c, 0}$) are genesis blocks.}
+Blocks of sequence number 0 (e.g., $c_{c, 0}$) are genesis blocks.}
 \label{fig:trustchain-good-cp}
@@
 The primary protocol is the consensus protocol,
-which can be seen as a technique of running infinitely many times of an existing Byzantine consensus algorithm (in this work we use asynchronous common subset from HoneyBadgerBFT~\cite{miller2016honey}),
+which can be seen as a technique of running infinitely many times of an existing Byzantine consensus algorithm (in this work we use the asynchronous common subset protocol described in HoneyBadgerBFT~\cite{miller2016honey}),
 starting a new execution immediately after the previous one is completed.
@@
 
-For brevity, we left out issues such as selecting facilitators in a fair way and the exact validity condition.
-The following sections give the formal description and will answer these issues.
+The following sections give the formal description.
 
@@
 
-\begin{definition}
-\textbf{Transaction block}
-
+\begin{definition}[Transaction block]
 The TX block is a six-tuple, i.e
@@
 
-\begin{definition}
-\textbf{Checkpoint block and genesis block}
-
+\begin{definition}[Checkpoint block and genesis block]
 The CP block is a five-tuple, i.e. 
@@
 
-\begin{definition}
+\begin{definition}[Consensus result]
 \label{sec:consensus-result}
-\textbf{Consensus result}
-
 Our consensus protocol runs in rounds,
@@
 
-\begin{definition}
-\textbf{Enclosure and agreed enclosure}
-
+\begin{definition}[Enclosure and agreed enclosure]
 If there exists a tuple $\langle c_{u,a}, c_{u, b} \rangle$ for a TX block $t_{u, i}$, where
@@
 
-\begin{definition}
-\textbf{Fragment and agreed fragment}
-
+\begin{definition}[Fragment and agreed fragment]
 If the enclosure of some TX block $t_{u, i}$ is $\langle c_{u,a}, c_{u, b} \rangle$,
@@
 \subsection{Consensus Protocol}
-Our scalable consensus protocol $\Pi_c$ uses an unmodified asynchronous common subset (ACS) algorithm as the key building block.
+Our scalable consensus protocol $\Pi_c$ uses an asynchronous common subset (ACS) protocol as the key building block.
 The objectives of the protocol are to
@@
 We formally define the desired properties below.
-\begin{definition}
-\label{def:consensus}
-\textbf{\textsc{Checo} consensus protocol}
-
+\begin{definition}[\textsc{Checo} consensus protocol]\label{def:consensus}
 A \textsc{Checo} consensus protocol is correct if the following holds for every round $r$.
@@
             \end{enumerate}
-    \item \emph{Fairness}:
-        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
     \item \emph{Termination}:
@@
 meaning that it satisfied the properties in \Cref{def:consensus}.
-In practice, the bootstrap oracle is most likely the software developer that sets up the system and assigns the facilitators of round 1 in a fair way.
+In practice, the bootstrap oracle is most likely the software developer that sets up the system and assigns the facilitators of round 1.
 This concludes the bootstrap phase.
@@
 Second, HoneyBadgerBFT uses threshold encryption to hide the content of the transactions.
-But we do not reach consensus on transactions, only CP blocks, so hiding CP blocks is meaningless for us.
+But we do not reach consensus on transactions, only CP blocks;
+the content of the CP blocks are not sensitive so there is no need to hide it.
 
@@
 Upon receiving $\C_{r + 1}$ and at least $n - t$ valid signatures, $u$ performs two tasks.
-First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1})$ (\Cref{alg:new-cp}).
-Second, it computes the new facilitators using $\textsf{get\_facilitator}(\C_{r + 1}, n)$ (\Cref{alg:facilitator}),
+First, it creates a new CP block using $\textsf{new\_cp}(\C_{r + 1})$, described in \Cref{alg:new-cp}.
+Second, it computes the new facilitators using $\textsf{get\_facilitator}(\C_{r + 1}, n)$, described in \Cref{alg:facilitator},
 and updates its facilitator set to the result.
@@
 The TX protocol $\Pi_t$, shown in \Cref{alg:tx-proto}, is run by all nodes.
-Nodes that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$ (\Cref{alg:new-tx}) with the intended counterparty $v$ identified by $pk_v$ and message $m$.
+Nodes that wish to initiate a transaction calls $\textsf{new\_tx}(pk_v, m, txid)$, described in \Cref{alg:new-tx}, with the intended counterparty $v$ identified by $pk_v$ and message $m$.
 $txid$ should be a uniformly distributed random value, i.e. $txid \in_R \{0, 1\}^{256}$.
@@
 At no time in \Cref{alg:new-tx} or \Cref{alg:tx-proto} do we need to hold the chain state and wait for some message to be delivered before committing a new block to the chain.
-This allows for a high level of concurrency where we can call $\textsf{new\_tx}(\cdot)$ and send \texttt{tx\_req} messages multiple times without waiting for the corresponding \texttt{tx\_resp} messages.
+This allows for a high level of concurrency where we can call many $\textsf{new\_tx}(\cdot)$ and send multiple \texttt{tx\_req} messages simultaneously without waiting for the corresponding \texttt{tx\_resp} messages.
 
@@
 Given a fragment $F_{v, j}$,
-the validity of the TX block $t_{u, i}$ with its corresponding fragment $F_{u, i}$ is captured by the function $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ (\Cref{alg:get-validity}).
+the validity of the TX block $t_{u, i}$ with its corresponding fragment $F_{u, i}$ is captured by the function $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ in \Cref{alg:get-validity}.
 Note that $t_{u, i}$ and $F_{u, i}$ are assumed to be valid,
@@
 The desired properties are as follows.
-\begin{definition}
-\label{def:validation}
-\textbf{\textsc{Checo} validation protocol}
-
+\begin{definition}[\textsc{Checo} validation protocol]\label{def:validation}
 A \textsc{Checo} validation protocol is correct if the following properties hold.
@@
 Otherwise, $A$ would refuse to make the transaction.
+
+\subsection{Fair Facilitator Election}
+Our consensus protocol does not guarantee fair facilitator election when dedicated attackers are present.
+If a malicious facilitator is elected, it can delay, eavesdrop and collect all CP block messages before sending its own.
+Effectively, it can predict the consensus result before the facilitators start the consensus protocol, and generate a CP block such that the malicious node has an unfair advantage of being elected as a facilitator in the next round.
+
+To address the issue above,
+the facilitators run an extra protocol after the consensus protocol to produce some unbiased randomness.
+Concretely, they invoke RandHound~\cite{syta2017scalable} and then propagated the randomness and the signatures in the same way as the consensus result.
+Upon receiving the randomness,
+every node uses it in the hash function of \Cref{alg:facilitator} (i.e. $\textsf{H}(\text{randomness} || \C_r || pk \text{ of } x)$) to compute the new set of facilitators.

### 2017-10-31 paper/s4-analysis.tex
Edit paper before defence
Address unfair facilitator election

@@
 \end{itemize}
-Fairness holds because we model $\textsf{H}(\cdot)$ as a query to the random oracle.
-Sorting nodes by the luck value $\textsf{H}(\C || pk)$ can be seen as a random permutation of a list of nodes.
-Therefore every node with a CP block in the consensus result has the same probability of becoming a facilitator.
+% Fairness holds because we model $\textsf{H}(\cdot)$ as a query to the random oracle.
+% Sorting nodes by the luck value $\textsf{H}(\C || pk)$ can be seen as a random permutation of a list of nodes.
+% Therefore every node with a CP block in the consensus result has the same probability of becoming a facilitator.
 
@@
 
-\subsection{Fault Tolerance}
-Finally, we consider the effect when the number of adversaries is more than $t$,
-which is useful because in practice it is difficult to guarantee that $t$ satisfies $n \ge 3t + 1$,
-especially when $N$ is large.
-Hence we are interested in the probability for this to happen under our facilitator election process.
-The problem is modelled by a hypergeometric distribution,
-where the random variable $X$ is the number of malicious nodes.
-For our system to fail, we are interested in the following probability.
-$$
-\Pr[X \ge \lfloor \frac{n-1}{3} \rfloor + 1]
-$$
-Let $E[X] = n\alpha$ where $0 \le \alpha \le \lfloor \frac{n-1}{3} \rfloor/n$.
-Using the tail inequality found in~\cite{skala2013hypergeometric}
-$$
-\Pr[X \ge E[X] + \tau n] \le e^{-2\tau^2n},
-$$
-and setting 
-$$
-\tau = \frac{\lfloor \frac{n-1}{3} \rfloor + 1}{n} - \alpha,
-$$
-we arrive at the following bound
-$$
-\Pr[X \ge \lfloor \frac{n-1}{3} \rfloor + 1] \le e^{-2 \big(\frac{\lfloor  \frac{n - 1}{3} \rfloor + 1}{n} - \alpha \big)^2 n}.
-$$
+% \subsection{Fault Tolerance}
+% Finally, we consider the effect when the number of adversaries is more than $t$,
+% which is useful because in practice it is difficult to guarantee that $t$ satisfies $n \ge 3t + 1$,
+% especially when $N$ is large.
+% Hence we are interested in the probability for this to happen under our facilitator election process.
+% The problem is modelled by a hypergeometric distribution,
+% where the random variable $X$ is the number of malicious nodes.
+% For our system to fail, we are interested in the following probability.
+% $$
+% \Pr[X \ge \lfloor \frac{n-1}{3} \rfloor + 1]
+% $$
+% Let $E[X] = n\alpha$ where $0 \le \alpha \le \lfloor \frac{n-1}{3} \rfloor/n$.
+% Using the tail inequality found in~\cite{skala2013hypergeometric}
+% $$
+% \Pr[X \ge E[X] + \tau n] \le e^{-2\tau^2n},
+% $$
+% and setting 
+% $$
+% \tau = \frac{\lfloor \frac{n-1}{3} \rfloor + 1}{n} - \alpha,
+% $$
+% we arrive at the following bound
+% $$
+% \Pr[X \ge \lfloor \frac{n-1}{3} \rfloor + 1] \le e^{-2 \big(\frac{\lfloor  \frac{n - 1}{3} \rfloor + 1}{n} - \alpha \big)^2 n}.
+% $$
+% 
+% The bound is not tight, but it is useful for picking parameters for the desired level of fault tolerance.
+% If $n$ is known, then we can pick a $\alpha$ such that the probability becomes small.
+% On the other hand, if $\alpha$ is fixed, we may increase $n$ to achieve the same.
+% Furthermore, hypergeometric distributions have light tails with ``faster-than-exponential fall-off''~\cite{skala2013hypergeometric}.
+% So the probability of picking more than $\lfloor \frac{n-1}{3} \rfloor$ malicious nodes drops off exponentially as the expected value moves away from $\lfloor \frac{n-1}{3} \rfloor$.
 
-The bound is not tight, but it is useful for picking parameters for the desired level of fault tolerance.
-If $n$ is known, then we can pick a $\alpha$ such that the probability becomes small.
-On the other hand, if $\alpha$ is fixed, we may increase $n$ to achieve the same.
-Furthermore, hypergeometric distributions have light tails with ``faster-than-exponential fall-off''~\cite{skala2013hypergeometric}.
-So the probability of picking more than $\lfloor \frac{n-1}{3} \rfloor$ malicious nodes drops off exponentially as the expected value moves away from $\lfloor \frac{n-1}{3} \rfloor$.
-To put this into perspective,
-suppose $n = 100$ and $\alpha = 1/10$, i.e. 10\% of the nodes are malicious,
-then the probability to elect more malicious nodes than the threshold is $9.9 \times 10^{-6}$.
+% TODO what's a good example?
+
+% To put this into perspective,
+% suppose $n = 100$ and $\alpha = 1/10$, i.e. 10\% of the nodes are malicious,
+% then the probability to elect more malicious nodes than the threshold is $9.9 \times 10^{-6}$.
+
 % To put this into perspective,
@@
 % Then the probability to elect more malicious nodes than the threshold is only $2.6 \times 10^{-16}$.
+
+% TODO not unbiased! 
+% workaround: Include random (dummy) block in proposal, for every facilitator

### 2017-10-31 paper/s6-conclusion.tex
Edit paper before defence
Address unfair facilitator election

@@
 The novelty of \textsc{Checo} is that it decouples consensus and transaction validation,
-which enables the desirable horizontal scalability property.
+which enables the desirable horizontal scalability property, without employing sharding.
 
@@
 Namely, our protocol achieves agreement on transactions as we argued in~\Cref{sec:correctness-of-validity}.
+Validity is achieved because honest nodes run the $\textsf{get\_validity}(\cdot)$ function, which, in fact, is the validity definition.
 Further, the horizontal scalability is demonstrated in~\Cref{sec:implementation},
 in the ideal case as well as the worst case.
+
 In the future, we hope to apply our system to a concrete application and evaluate its performance.
+Furthermore, we plan to explore a few useful design alternatives such as open membership, total fork detection and fair facilitator election.

### 2017-12-05 paper/s1-intro.tex
Minor fixes

@@
 The first blockchain system---Bitcoin---is almost ten years old. 
-Its market capitalisation is over \$60 billion USD and growing~\cite{bitcoinmarketcap}.
+Its market capitalisation is nearly \$200 billion USD at the time of writing~\cite{bitcoinmarketcap}.
 We can be reasonably sure that such systems,
@@
 In essence, such systems contain a fixed set of nodes (sometimes called validating peer) that run a Byzantine consensus algorithm to decide on new blocks.
-Such systems can achieve much higher transaction rates, e.g., 10,000 TPS if the number of validating peer is under 16 for PBFT~\cite[Section 5.2]{miller2016honey}.
+They can achieve much higher transaction rates, e.g., 10,000 TPS if the number of validating peer is under 16 for PBFT~\cite[Section 5.2]{miller2016honey}.
 However, these systems do not scale, e.g., the transaction rate drops to under 5000 TPS when the number of validating peer is 64~\cite[Section 5.2]{miller2016honey}.
@@
 \textbf{Contribution.\quad}
-The key insight is not to reach consensus using an existing consensus algorithm (a modified HoneyBadgerBFT~\cite{miller2016honey}) on transactions themselves,
+The key insight is not to reach consensus using an existing consensus algorithm on transactions themselves,
 but on special blocks called checkpoint blocks,
@@
 \begin{itemize}
-    \item We formally introduce a blockchain system---\textsc{Checo}\footnote{Derived from ``\emph{Che}ckpoint \emph{co}nsensus''.}.
+    \item We formally introduce a blockchain system---\textsc{Checo}\footnote{Derived from ``\textsc{Che}ckpoint \textsc{co}nsensus''.}.
         It uses individual hash chains and checkpoints on every node to achieve

### 2017-12-05 paper/s2-problem.tex
Minor fixes

@@
 Namely, the property only holds if honest nodes do not output \emph{unknown}.
-For example, for a transaction, it is fine if two honest nodes output \emph{valid} and \emph{unknown}, but it is not fine to output \emph{valid} and \emph{invalid}.
+For example, for a transaction, it is fine if two honest nodes output \emph{valid} and \emph{unknown}, but they should never output \emph{valid} and \emph{invalid}.
 Our problem does not have a termination property.
@@
 This formulation is so that the protocol can act as a building block to many applications.
-Due to this fact, we also do not consider global fork prevention or detection,
+Thus, we do not consider global fork prevention or detection,
 as some application may not need such strong guarantees such as the accounting of internet traffic in Tribler~\cite{pimotte, pouwelse2008tribler}.

### 2018-04-04 paper/s1-intro.tex
Polish

@@
 Another way to improve transaction rate is to use traditional Byzantine consensus algorithms such as PBFT~\cite{castro1999practical} in a permissioned ledger such as Hyperledger Fabric~\cite{cachin2016architecture}.
-In essence, such systems contain a fixed set of nodes (sometimes called validating peer) that run a Byzantine consensus algorithm to decide on new blocks.
-They can achieve much higher transaction rates, e.g., 10,000 TPS if the number of validating peer is under 16 for PBFT~\cite[Section 5.2]{miller2016honey}.
+In essence, such systems contain a fixed set of nodes, called validating peers, that run a Byzantine consensus algorithm to decide on new blocks.
+They can achieve much higher transaction rates, e.g., 10,000 TPS if the number of validating peers is under 16 for PBFT~\cite[Section 5.2]{miller2016honey}.
 However, these systems do not scale, e.g., the transaction rate drops to under 5000 TPS when the number of validating peer is 64~\cite[Section 5.2]{miller2016honey}.
@@
 which leads to the goal of this work.
-Hence the research question which we wish to answer is as follows.
+Hence, the research question which we wish to answer is as follows.
 \begin{displayquote}

### 2018-04-04 paper/s3-system.tex
Polish

@@
 meaning that it satisfied the properties in \Cref{def:consensus}.
-In practice, the bootstrap oracle is most likely the software developer that sets up the system and assigns the facilitators of round 1.
+In practice, the bootstrap oracle is most likely a group of software developers
+(representing different organisations)
+that agreed to work together to set up the system and assign the facilitators of round 1.
+The number of facilitators is $n$, we discuss the trade-offs for different values of $n$ in~\Cref{sec:implementation}.
 This concludes the bootstrap phase.
@@
 First, if $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h-1} \rangle$ to all the facilitators.
-Second, if $u$ is a facilitator, it waits until $N - t$ messages of type \texttt{cp\_msg} are received.
+Second, if $u$ is a facilitator, it waits for $N - t$ messages of type \texttt{cp\_msg}.
 % Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D \gg \Delta$.
-Invalid messages are removed.
-Those are blocks with invalid signatures and blocks signed by the same key.
-With the sufficient number of \texttt{cp\_msg} messages,
+Invalid messages are removed,
+which are blocks with invalid signatures and blocks signed by the same key.
+With a sufficient number of \texttt{cp\_msg} messages,
 it begins the ACS algorithm and some $\C'_{r + 1}$ should be agreed upon by the end of it.
@@
 The core of the consensus phase is the ACS algorithm,
-which is derived from HoneyBadgerBFT~\cite{miller2016honey}.
+which is described in HoneyBadgerBFT~\cite{miller2016honey}.
 We do not use the full HoneyBadgerBFT due to the following.
@@
 
-Continuing, when $\F_{r}$ finish the ACS execution and reach agreement on $\C_{r+1}$,
+When $\F_{r}$ finish the ACS execution and reach agreement on $\C_{r+1}$,
 they immediately broadcast two messages to all the nodes---first the consensus message $\langle \texttt{cons\_msg}, \C_{r+1} \rangle$,
@@
 The protocol is also a request-response protocol.
-But before explaining the protocol itself, we first define what it means for a transaction to be valid.
+Before explaining the protocol itself, we first define what it means for a transaction to be valid.
 
@@
 
-\subsection{Fair Facilitator Election}
-Our consensus protocol does not guarantee fair facilitator election when dedicated attackers are present.
+\subsection{Unbiased Facilitator Election}
+Our consensus protocol does not guarantee unbiased facilitator election when dedicated attackers are present.
 If a malicious facilitator is elected, it can delay, eavesdrop and collect all CP block messages before sending its own.
-Effectively, it can predict the consensus result before the facilitators start the consensus protocol, and generate a CP block such that the malicious node has an unfair advantage of being elected as a facilitator in the next round.
+Effectively, it can generate a CP block such that it has an unfair advantage of being elected as a facilitator in the next round.
 

### 2018-04-04 paper/s5-evaluation.tex
Edit paper before defence
Polish

@@
-\section{Implementation and Evaluation}
-\label{sec:implementation}
-
 \begin{figure*}[ht]
@@
 
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=\linewidth]{neighbour-random/round-duration-vs-facilitators}
+    \caption{The consensus duration increases polynomially with respect to the number of facilitators.}
+    \label{fig:round-duration-vs-facilitators}
+\end{figure}
+
+\section{Implementation and Evaluation}
+\label{sec:implementation}
+
 A free and open source implementation can be found on GitHub: \url{https://github.com/kc1212/checo}.
@@
 We also implement the caching optimisation discussed in \Cref{sec:caching}.
-It is written in the Python programming language.
 The cryptography primitives we use are SHA256 for hash functions and Ed25519 for digital signatures.
@@
 
-The downside of our design is that the communication complexity of the consensus protocol grows polynomially as the number of facilitators grows linearly.
+The downside of our design is that the communication complexity of the consensus protocol grows polynomially with respect to the number of facilitators.
 Hence, the consensus protocol will take longer to complete, and larger fragments must be sent for transaction verification.
@@
 only the transaction verification delay is affected.
+The experimental results in \Cref{fig:round-duration-vs-facilitators} demonstrates this issue, it uses the same experimental-setup as before.
 We refer the reader to~\cite[Chapter 5]{checo} for additional analysis of the effect of the number of facilitators as well as other experimental results.

### 2018-04-23 conference-presentation/presentation.tex
Prepare conference presentation

--- 2017-08-30 presentation/presentation.tex
@@
 
-\setbeamertemplate{navigation symbols}{} % remove navigation symbols
-\mode<presentation>{\usetheme{tud}}
-
 % https://tex.stackexchange.com/questions/124256/how-do-i-get-numbered-entries-in-a-beamer-bibliography
@@
 \newcommand{\A}{\mathcal{A}}
-\newcommand{\C}{\mathcal{C}}
+\renewcommand{\C}{\mathcal{C}}
 \renewcommand{\P}{\mathcal{P}}
@@
 
-% \setbeameroption{show notes}
+\setbeameroption{show notes}
 
@@
 
-\author{K.~Cong}
+% \author{Kelong Cong \inst{1} \and Zhijie Ren \inst{2} \and Johan~Pouwelse \inst{2}}
 % - Give the names in the same order as the appear in the paper.
@@
 
-\institute[Delft University of Technology] % (optional, but mostly needed)
-{
-  Faculty of Electrical Engineering, Mathematics and Computer Science\\
-  Delft University of Technology}
-% - Use the \inst command only if there are several affiliations.
-% - Keep it simple, no one is interested in your street address.
+% \institute{\inst{1} EPFL \and \inst{2} TU Delft}
+
+\author[Short Name (U ABC)]{%
+  \texorpdfstring{%
+    \begin{columns}
+      \column{.3333\linewidth}
+      \centering
+      Kelong~Cong \\ \small kelong.cong@epfl.ch \\ EPFL
+      \column{.3333\linewidth}
+      \centering
+      Zhijie~Ren \\ \small z.ren@tudelft.nl \\ TU Delft
+      \column{.3333\linewidth}
+      \centering
+      Johan~Pouwelse \\ \small peer2peer@gmail.com \\ TU Delft
+    \end{columns}
+ }
+ {Author 1, Author 2, Author 3}
+}
 
 % \date{Conference Name, 2013}
-\date{\today}
+\date{IFIP Networking, 2018}
 
@@
 
-\begin{frame}{The dangers of centralisation:~Examples}
-  \begin{itemize}
-    \item Facebook can predict your opinions and desires from your ``likes''~\cite{youyou2015computer},
-      and then create a ``psychographic'' profile for use in political campaigns~\cite{psychographic}
-    \item Baidu's promoted search result on medical care caused death of a student~\cite{weizexi}
-  \end{itemize}
-  \note{
-    \begin{itemize}
-      \item The treatment was phased out at clinical stage in the US because the success rate is too low
-      \item ``The study was conducted on 86,220 volunteers who have a Facebook account and who
-      completed a hundred-item personality questionnaire. The Facebook algorithm predicted the volunteers'
-      answers based on monitoring their Facebook Likes – which webpages, images and clips they tagged with the
-      Like button. The more Likes, the more accurate the predictions. The algorithm's predictions were compared
-      with those of work colleagues, friends, family members and spouses. Amazingly, the algorithm needed a set
-      of only ten Likes in order to outperform the predictions of work colleagues. It needed seventy Likes to
-      outperform friends, 150 Likes to outperform family members and 300 Likes to outperform spouses. In other
-      words, if you happen to have clicked 300 Likes on your Facebook account, the Facebook algorithm can
-      predict your opinions and desires better than your husband or wife!'' -Harari, \emph{Homo Deus}
-      \item The student had cancer---(synovial sarcoma)
-      \item Summary: we live in an age where companies and governments are able to control our thoughts,
-      where is the freedom and liberty in that?
-    \end{itemize}
-  }
-\end{frame}
+% \begin{frame}{The dangers of centralisation:~Examples}
+%   \begin{itemize}
+%     \item Facebook can predict your opinions and desires from your ``likes''~\cite{youyou2015computer},
+%       and then create a ``psychographic'' profile for use in political campaigns~\cite{psychographic}
+%     \item Baidu's promoted search result on medical care caused death of a student~\cite{weizexi}
+%   \end{itemize}
+%   \note{
+%     \begin{itemize}
+%       \item The treatment was phased out at clinical stage in the US because the success rate is too low
+%       \item ``The study was conducted on 86,220 volunteers who have a Facebook account and who
+%       completed a hundred-item personality questionnaire. The Facebook algorithm predicted the volunteers'
+%       answers based on monitoring their Facebook Likes – which webpages, images and clips they tagged with the
+%       Like button. The more Likes, the more accurate the predictions. The algorithm's predictions were compared
+%       with those of work colleagues, friends, family members and spouses. Amazingly, the algorithm needed a set
+%       of only ten Likes in order to outperform the predictions of work colleagues. It needed seventy Likes to
+%       outperform friends, 150 Likes to outperform family members and 300 Likes to outperform spouses. In other
+%       words, if you happen to have clicked 300 Likes on your Facebook account, the Facebook algorithm can
+%       predict your opinions and desires better than your husband or wife!'' -Harari, \emph{Homo Deus}
+%       \item The student had cancer---(synovial sarcoma)
+%       \item Summary: we live in an age where companies and governments are able to control our thoughts,
+%       where is the freedom and liberty in that?
+%     \end{itemize}
+%   }
+% \end{frame}
 
@@
   \begin{itemize}
-    \item Blockchains are distributed (replicated) ledgers with no central control\footnote{In practice, some central control exist.}
+    \item Blockchains are distributed (replicated) ledgers with no central control
     \item They enable internet-scale consensus for the first time
@@
 \begin{frame}{Related work}
-  \begin{table}[htb]
-    \centering
-    \caption{Summary of the scalability properties of many blockchain systems.}
-    \label{tab:scalability}
-    \begin{tabular}{|l|l|l|l|}
-    \hline
-    \textbf{\thead{Not\\scalable}} & \textbf{\thead{Somewhat\\scalable}} & \textbf{\thead{Limited horizontal\\scalability}} & \textbf{\thead{True horizontal\\scalability}} \\ \hline
-    Bitcoin  & Hyperledger & Elastico          & \textsc{Checo} (this work) \\
-    Litecoin & ByzCoin     & OmniLedger        &                \\
-    \hline
-    \end{tabular}
-\end{table}
-\begin{center}
-\Large{
-$\xrightarrow{\qquad\text{scalability}\qquad}$
-}
-\end{center}
+    \begin{itemize}
+        \item Off-chain solution
+            \begin{itemize}
+                \item Lightning Network
+                \item Perun
+            \end{itemize}
+        \item On-chain solution
+            \begin{itemize}
+                \item Parameter tuning
+                \item BFT consensus (e.g. Tendermint, ByzCoin)
+                \item Sharding (e.g. Elastico, OmniLedger)
+            \end{itemize}
+    \end{itemize}
 \end{frame}
 
-% \begin{frame}{Related work}
-%   State-of-the-art: Sharding
-%   \begin{itemize}
-%   \item Split state into multiple shards
-%   \item Shards run consensus algorithm in parallel
-%   \item Difficult to perform atomic inter-shard transactions
-%     \begin{itemize}
-%       \item Elastico: not possible
-%       \item OmniLedger: via Atomic Commit protocol
-%     \end{itemize}
-%   \item Additional complexity
-%   \item Diminishing return
-%   \end{itemize}
-% \end{frame}
+\begin{frame}{Related work}
+  State-of-the-art---Sharding:
+  \begin{itemize}
+  \item Split state into multiple shards
+  \item Shards run consensus algorithm in parallel
+  \end{itemize}
+  Challenges:
+  \begin{itemize}
+      \item Choosing and evolving the shard size
+      \item Perform atomic inter-shard transactions
+      \item Parameter choice highly depends on the application
+  \end{itemize}
+\end{frame}
 

### 2018-04-24 conference-presentation/presentation.tex
Remove fairness
Clean up presentation

@@
 \usepackage{graphicx}
-\graphicspath{{./figures/}{../figures/}}
+\graphicspath{{../presentation/figures/}{../figures/}}
 \usepackage{subfig}
@@
 \usepackage[style=ieee]{biblatex}
-\addbibresource{references.bib}
+% \addbibresource{references.bib}
 
 % https://tex.stackexchange.com/questions/124256/how-do-i-get-numbered-entries-in-a-beamer-bibliography
-\setbeamertemplate{bibliography item}{\insertbiblabel}
+% \setbeamertemplate{bibliography item}{\insertbiblabel}
 
@@
 
-\setbeameroption{show notes}
+% \setbeameroption{show notes}
 
@@
     \item But it puts central authorities in control
-    \item Most are motivated exclusively by profit
-    \item Not always in the interest of the ``users''\footnote{Typically users of some free service X are, in fact, used by X.}
+    \item Many are motivated by profit or the goals of the local government
+    \item The interest of the authorities do not align with the users
   \end{itemize}
 \end{frame}
-
-% \begin{frame}{The dangers of centralisation:~Examples}
-%   \begin{itemize}
-%     \item Facebook can predict your opinions and desires from your ``likes''~\cite{youyou2015computer},
-%       and then create a ``psychographic'' profile for use in political campaigns~\cite{psychographic}
-%     \item Baidu's promoted search result on medical care caused death of a student~\cite{weizexi}
-%   \end{itemize}
-%   \note{
-%     \begin{itemize}
-%       \item The treatment was phased out at clinical stage in the US because the success rate is too low
-%       \item ``The study was conducted on 86,220 volunteers who have a Facebook account and who
-%       completed a hundred-item personality questionnaire. The Facebook algorithm predicted the volunteers'
-%       answers based on monitoring their Facebook Likes – which webpages, images and clips they tagged with the
-%       Like button. The more Likes, the more accurate the predictions. The algorithm's predictions were compared
-%       with those of work colleagues, friends, family members and spouses. Amazingly, the algorithm needed a set
-%       of only ten Likes in order to outperform the predictions of work colleagues. It needed seventy Likes to
-%       outperform friends, 150 Likes to outperform family members and 300 Likes to outperform spouses. In other
-%       words, if you happen to have clicked 300 Likes on your Facebook account, the Facebook algorithm can
-%       predict your opinions and desires better than your husband or wife!'' -Harari, \emph{Homo Deus}
-%       \item The student had cancer---(synovial sarcoma)
-%       \item Summary: we live in an age where companies and governments are able to control our thoughts,
-%       where is the freedom and liberty in that?
-%     \end{itemize}
-%   }
-% \end{frame}
 
@@
             \begin{itemize}
-                \item Lightning Network
-                \item Perun
+                \item Lightning Network\footnote{\url{https://lightning.network/}}
+                \item Perun\footnote{\url{https://perun.network}}
             \end{itemize}
@@
                 \item Parameter tuning
-                \item BFT consensus (e.g. Tendermint, ByzCoin)
-                \item Sharding (e.g. Elastico, OmniLedger)
+                \item BFT consensus (e.g. Tendermint\footnote{\url{https://www.tendermint.com/}}, ByzCoin\footnote{KJGKGF, USENIXSecurity16})
+                \item Sharding (e.g. Elastico\footnote{LNZBGS, CCS16}, OmniLedger\footnote{KJGGSF, S\&P18})
             \end{itemize}
@@
     \item Goal 2: create new CP blocks at the end of the protocol
-    \item Uses an existing fault-tolerant consensus algorithm (HoneyBadgerBFT~\cite{miller2016honey}) as the building block
+    \item Uses an existing fault-tolerant consensus algorithm (HoneyBadgerBFT\footnote{MXCSS, CCS16}) as the building block
     \item But it cannot be used in a large network due to high communication complexity
@@
       The consensus result is valid such that a new set of facilitators can be computed from it.
-    \item \emph{Fairness}:
-        Every node with a CP block in the consensus result should have an equal probability of becoming a facilitator.
+    % \item \emph{Fairness}:
+    %     Every node with a CP block in the consensus result should have an equal probability of becoming a facilitator.
     \item \emph{Termination}:
@@
     \item Experiment on the DAS-5\footnote{\url{http://www.cs.vu.nl/das5/}}
-    \item Up to 1500 nodes
+    \item Up to 1200 nodes
   \end{itemize}
@@
 
-\begin{frame}{Stress test (fixed neighbour)}
-  \begin{figure}[h]
-  \includegraphics[width=0.9\textwidth]{throughput-vs-population-large}
-  \centering
-  \end{figure}
-\end{frame}
-
 \section{Conclusion}
@@
 
-\begin{frame}{Bibliography}
-\printbibliography
-\end{frame}
+% \begin{frame}{Bibliography}
+% \printbibliography
+% \end{frame}
 
@@
     \item Asynchronous common subset
-    \item A simplification of HoneyBadgerBFT~\cite{miller2016honey}
+    \item A simplification of HoneyBadgerBFT
     \item $n$ nodes
@@
 
+\begin{frame}{Stress test (fixed neighbour)}
+  \begin{figure}[h]
+  \includegraphics[width=0.9\textwidth]{throughput-vs-population-large}
+  \centering
+  \end{figure}
+\end{frame}
+
 \end{document}

### 2018-05-15 conference-presentation/presentation.tex
Add conference location

@@
 
+\beamertemplatenavigationsymbolsempty
+\setbeamertemplate{footline}[frame number]
+
 \newcommand{\A}{\mathcal{A}}
@@
 
-\title{A Blockchain Consensus Protocol With\\Horizontal Scalability}
+\title{A Blockchain Consensus Protocol with\\Horizontal Scalability}
 % \subtitle{Optional Subtitle}
@@
  }
- {Author 1, Author 2, Author 3}
+ {Kelong~Cong, Zhijie~Ren, Johan~Pouwelse}
 }
@@
 % \date{Conference Name, 2013}
-\date{IFIP Networking, 2018}
+\date{IFIP Networking, Zurich 2018}
 
@@
     \item All blockchain systems have a consensus algorithm
-    \item Early consensus algorithms (PoW) do not scale
+    \item Early consensus algorithms do not scale
     \item Bitcoin is limited to 7 transactions per second
@@
                 \item Perun\footnote{\url{https://perun.network}}
+                \item Easy to deploy but application specific
             \end{itemize}
@@
   \begin{itemize}
-    \item Blockchain consensus protocol---application neutral, e.g., PoW
+    \item Blockchain consensus protocol---application neutral
     \item Fault-tolerant---tolerate a number of malicious nodes
@@
   \begin{center}
-  {\Large The four components of \textsc{Checo}}
+  {\Large The four components of \textsc{Checo}\footnote{\textsc{Che}ckpoint \textsc{co}nsensus}} 
   \begin{figure}[h]
@@
   \begin{figure}[h]
-  \includegraphics[width=0.75\textwidth]{trustchain-good-cp}
+  \includegraphics[width=0.84\textwidth]{trustchain-good-cp}
   \centering
@@
   \item Goal: record transactions
-  \item A transaction is represented by a pair of TX blocks,
-  i.e. a contract signed by both parties
+  \item A transaction is an agreement on a piece of data, i.e. digitally signed by both parties
+  \item It is represented by a pair of TX blocks
 \end{itemize}
@@
   \begin{figure}[h]
-  \includegraphics[width=0.75\textwidth]{trustchain-good-cp-frag}
+  \includegraphics[width=0.84\textwidth]{trustchain-good-cp-frag}
   \centering
@@
 
-\begin{frame}{\subsecname:~properties}
-\label{def:consensus}
-\begin{itemize}
-    \item \emph{Agreement}:
-        Every correct outputs the same set of facilitators.
-    \item \emph{Validity}:
-      The consensus result is valid such that a new set of facilitators can be computed from it.
-    % \item \emph{Fairness}:
-    %     Every node with a CP block in the consensus result should have an equal probability of becoming a facilitator.
-    \item \emph{Termination}:
-        Every correct node eventually outputs a set of facilitators.
-\end{itemize}
-\end{frame}
-
 \subsection{Transaction protocol}
@@
 \item No guarantee that nodes follow this protocol
+\item Contrary to early blockchain systems, we do not broadcast transactions
 \end{itemize}
@@
     \item Up to 1200 nodes
+    \item A third of the facilitators are malicious
   \end{itemize}
@@
   \begin{figure}[h]
-  \includegraphics[width=0.9\textwidth]{neighbour-random/throughput-vs-population}
+  \includegraphics[width=0.95\textwidth]{neighbour-random/throughput-vs-population}
   \centering
@@
   \begin{figure}[h]
-  \includegraphics[width=0.9\textwidth]{neighbour-fixed/throughput-vs-population}
+  \includegraphics[width=0.95\textwidth]{neighbour-fixed/throughput-vs-population}
   \centering
@@
 \begin{frame}{\secname}
-  Our work answers the research question.
-  \begin{block}{}
-    How can we design a \emph{blockchain consensus protocol} that is \emph{fault-tolerant},
-    \emph{horizontally-scalable}, and able to reach \emph{global consensus?}
-  \end{block}
+    Our contribution:
   \begin{itemize}
-    \item Fault-tolerance is achieved using HoneyBadgerBFT
-    \item Horizontal-scalability is achieved by separating consensus and validation, demonstrated experimentally
-    \item Global-consensus on transactions is achieved via consensus on CP blocks
+      \item We design a blockchain consensus protocol---\textsc{Checo} with horizontal scalability
+      \item We formally analyse CHECO to ensure correctness
+      \item Our experimental evaluation show strong evidence of our claim.
+  \end{itemize}
+      Future perspective:
+  \begin{itemize}
+      \item Improve fault-tolerance using ByzCoinX from OmniLedger
+      \item Design and evaluate UTXO-style transactions (new paper to appear in Crypto Valley Conference)
+      \item Assume a permissionless model
   \end{itemize}
@@
 % \end{frame}
-
-\begin{frame}{Thank you}
-  \begin{center}
-  {\Large Any questions?}
-  \end{center}
-\end{frame}
 
@@
 
-\begin{frame}[noframenumbering]{Future work}
-\begin{itemize}
-\item Implement and experiment with a concrete application
-\item Analyse the system in the permissionless environment
-\item Improve fault tolerance
-\end{itemize}
-\end{frame}
-
-\begin{frame}{Stress test (fixed neighbour)}
+\begin{frame}[noframenumbering]{Stress test (fixed neighbour)}
   \begin{figure}[h]

### 2018-05-16 conference-presentation/presentation.tex
Final tweaks

@@
 \section{System architecture}
-\begin{frame}{Intuition and idea explored in this thesis}
+\begin{frame}{Intuition}
   \begin{itemize}
@@
 
-\subsection{System model}
 \subsection{Architecture overview}
@@
 \item Two TX blocks are generated on the chains of Alice and Bob
+\item Contrary to early blockchain systems, we do not broadcast transactions
 \item No guarantee that nodes follow this protocol
-\item Contrary to early blockchain systems, we do not broadcast transactions
 \end{itemize}
@@
   \item CP blocks of the fragments are ``anchored'' due to the consensus protocol
-  \item It is difficult to modify the fragment once ``anchored''
+  \item It is not possible to modify the fragment once ``anchored''
   \item Since the transaction protocol and the validation protocol only use point-to-point communication,
@@
       \item We design a blockchain consensus protocol---\textsc{Checo} with horizontal scalability
-      \item We formally analyse CHECO to ensure correctness
+      \item We formally analyse \textsc{Checo} to ensure correctness
       \item Our experimental evaluation show strong evidence of our claim.
@@
       \item Improve fault-tolerance using ByzCoinX from OmniLedger
-      \item Design and evaluate UTXO-style transactions (new paper to appear in Crypto Valley Conference)
+      \item Design and evaluate UTXO-style transactions (new paper to appear in the Crypto Valley Conference)
       \item Assume a permissionless model
